### 2018-03-21 paper/trento-partons.tex
add paper to repo

\documentclass[aps,prc,reprint,amsmath,nofootinbib]{revtex4-1}

\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{multirow}

\usepackage{color}
\definecolor{theblue}{RGB}{0,50,230}

\usepackage{hyperref}
\hypersetup{
  colorlinks=true,
  linkcolor=theblue,
  citecolor=theblue,
  urlcolor=theblue
}

\usepackage{tikz}

\usepackage{graphicx}
\graphicspath{{../plots/}}

\newcommand{\trento}{T\raisebox{-0.5ex}{R}ENTo}
\newcommand{\avg}[1]{\langle #1 \rangle}
\newcommand{\nch}{N_\text{ch}}
\newcommand{\npart}{N_\text{part}}
\newcommand{\sqrts}{\sqrt{s_{NN}}}
\newcommand{\T}{\tilde{T}}
\newcommand{\vnk}[2]{v_#1\{#2\}}
\newcommand{\paddedhline}{\noalign{\smallskip}\hline\noalign{\smallskip}}
\newcommand{\order}[1]{$\mathcal O(10^{#1})$}
\newcommand{\note}{\textcolor{theblue}{[?]}}

% Convenient figure macro.  Usage:
%
%   \fig[placement specifier = t]{filename}{caption}
%
% This creates a figure environment, includes the given filename as graphics,
% puts the given caption below the graphics, and labels it 'fig:filename'.
% The optional placement specifier defaults to 't' and is passed directly to
% the figure environment.
% Use \fig* to make a figure* environment, i.e. a wide figure.
\usepackage{xparse}
\NewDocumentCommand\fig{sO{t}mm}{
  \begin{figure\IfBooleanT{#1}{*}}[#2]
    \includegraphics{#3}
    \caption{\label{fig:#3}#4}
  \end{figure\IfBooleanT{#1}{*}}
}

\begin{document}

\title{
  Estimating initial state and quark-gluon plasma medium properties\\
  using a hydrodynamic model with nucleon substructure\\calibrated to p-Pb and Pb-Pb collisions at $\sqrts=5.02$~TeV
}

\author{J.\ Scott Moreland}
\author{Jonah E.\ Bernhard}
\author{Steffen A.\ Bass}
\affiliation{Department of Physics, Duke University, Durham, NC 27708-0305}

\date{\today}

\begin{abstract}
  We posit a unified hydrodynamic description of ultrarelativistic $p$-Pb and Pb-Pb nuclear collisions at $\sqrts=5.02$~TeV and evaluate the likelihood of our assertion using Bayesian inference.
  Specifically, we model the dynamics of both collision systems using QGP initial conditions with parametric nucleon substructure, a pre-equilibrium free-streaming stage, event-by-event viscous hydrodynamics, and a microscopic hadronic afterburner.
  Free parameters of the model which describe the initial state and QGP medium are then simultaneously calibrated to fit charged particle yields, mean $p_T$ and flow cumulants of both systems.
  We argue that the global agreement of the calibrated model with available experimental data strongly supports the existence of hydrodynamic flow in small collision systems at ultrarelativistic energies, and that this flow necessarily develops at length scales smaller than a proton.
  Finally, we present posterior distributions on our model's initial state and QGP medium parameters, including new constraints on the shape and structure of the proton.
\end{abstract}

\maketitle

\section{Introduction}
  Ultrarelativistic collisions of light and light-heavy \mbox{(e.g.\ $p$-$p$ and $p$-Pb)} ion pairs produce long-range multiparticle correlations which are striking similar to those observed in heavy-ion collisions \note, where the observed collectivity is commonly attributed to hydrodynamic flow \note.
  This observation suggests that hydrodynamic behavior could be manifest in small droplets of quark-gluon-plasma (QGP), and that flow might develop at length scales smaller than a proton.

 Lorem ipsum dolor sit amet, consectetur adipiscing elit. In tincidunt elit a libero luctus, ut pharetra sem accumsan. Nullam eros dui, egestas ut arcu et, commodo suscipit purus. Quisque dapibus tincidunt nulla at vehicula. Nulla ut libero sapien. Suspendisse ac consectetur metus. Interdum et malesuada fames ac ante ipsum primis in faucibus. Nunc facilisis volutpat suscipit. Pellentesque nec est orci. Donec id pharetra enim.

Aliquam porttitor massa risus, hendrerit dignissim purus placerat vel. Suspendisse dui ante, porttitor pretium lorem et, placerat posuere mi. Donec lacinia nisl leo, ac luctus libero tincidunt non. Donec pretium ultricies felis vitae tempus. Duis pulvinar nisi justo, sit amet finibus orci volutpat non. Proin fermentum a magna id rhoncus. Vestibulum posuere, ipsum vel dignissim laoreet, arcu nibh venenatis neque, et aliquet dui dui a ipsum. Praesent sagittis condimentum neque, sed aliquam odio volutpat nec. Aenean consequat ante eget mauris lacinia, sed ultrices est dapibus. Morbi fermentum malesuada turpis, ac placerat est bibendum tincidunt. Nullam leo ipsum, sodales sit amet viverra non, iaculis vel nisl. Vestibulum purus ipsum, egestas ut bibendum sed, congue quis purus. Curabitur iaculis tortor nisl, nec venenatis tortor placerat a. Fusce feugiat nisl vitae porta porta. Praesent non arcu leo. Sed gravida eget neque in auctor.

Sed ac ullamcorper ipsum. Sed quis eleifend mauris. Phasellus quis massa neque. Proin nec lectus quis ante aliquam tincidunt ut sit amet eros. Nullam et odio eu elit aliquam molestie quis ut arcu. Aliquam gravida enim non tortor bibendum, at semper tellus egestas. In libero lectus, convallis in suscipit nec, bibendum eu felis. Vestibulum ante ipsum primis in faucibus orci luctus et ultrices posuere cubilia Curae; Vivamus quis lectus neque. Sed ac placerat justo. Nullam ultricies vulputate ligula, vel porta mauris scelerisque et. Sed hendrerit blandit semper. Sed vel bibendum elit, eu posuere elit. Phasellus nunc massa, tincidunt ac tempus quis, blandit quis erat.

\section{Nuclear collision model}

 Lorem ipsum dolor sit amet, consectetur adipiscing elit. In tincidunt elit a libero luctus, ut pharetra sem accumsan. Nullam eros dui, egestas ut arcu et, commodo suscipit purus. Quisque dapibus tincidunt nulla at vehicula. Nulla ut libero sapien. Suspendisse ac consectetur metus. Interdum et malesuada fames ac ante ipsum primis in faucibus. Nunc facilisis volutpat suscipit. Pellentesque nec est orci. Donec id pharetra enim.

\subsection{Initial conditions}

Aliquam porttitor massa risus, hendrerit dignissim purus placerat vel. Suspendisse dui ante, porttitor pretium lorem et, placerat posuere mi. Donec lacinia nisl leo, ac luctus libero tincidunt non. Donec pretium ultricies felis vitae tempus. Duis pulvinar nisi justo, sit amet finibus orci volutpat non. Proin fermentum a magna id rhoncus. Vestibulum posuere, ipsum vel dignissim laoreet, arcu nibh venenatis neque, et aliquet dui dui a ipsum. Praesent sagittis condimentum neque, sed aliquam odio volutpat nec. Aenean consequat ante eget mauris lacinia, sed ultrices est dapibus. Morbi fermentum malesuada turpis, ac placerat est bibendum tincidunt. Nullam leo ipsum, sodales sit amet viverra non, iaculis vel nisl. Vestibulum purus ipsum, egestas ut bibendum sed, congue quis purus. Curabitur iaculis tortor nisl, nec venenatis tortor placerat a. Fusce feugiat nisl vitae porta porta. Praesent non arcu leo. Sed gravida eget neque in auctor.

\subsection{Pre-equilibrium dynamics}

Sed ac ullamcorper ipsum. Sed quis eleifend mauris. Phasellus quis massa neque. Proin nec lectus quis ante aliquam tincidunt ut sit amet eros. Nullam et odio eu elit aliquam molestie quis ut arcu. Aliquam gravida enim non tortor bibendum, at semper tellus egestas. In libero lectus, convallis in suscipit nec, bibendum eu felis. Vestibulum ante ipsum primis in faucibus orci luctus et ultrices posuere cubilia Curae; Vivamus quis lectus neque. Sed ac placerat justo. Nullam ultricies vulputate ligula, vel porta mauris scelerisque et. Sed hendrerit blandit semper. Sed vel bibendum elit, eu posuere elit. Phasellus nunc massa, tincidunt ac tempus quis, blandit quis erat.

\subsection{Hydrodynamics and Boltzmann transport}

 Lorem ipsum dolor sit amet, consectetur adipiscing elit. In tincidunt elit a libero luctus, ut pharetra sem accumsan. Nullam eros dui, egestas ut arcu et, commodo suscipit purus. Quisque dapibus tincidunt nulla at vehicula. Nulla ut libero sapien. Suspendisse ac consectetur metus. Interdum et malesuada fames ac ante ipsum primis in faucibus. Nunc facilisis volutpat suscipit. Pellentesque nec est orci. Donec id pharetra enim.

To first approximation---as with nearly every object in physics---it is convenient to approximate the nucleons in heavy-ion initial condition models as a spherically symmetric particles of finite spatial extent. 
Commonly, a Gaussian is used 
\begin{figure}
  \begin{tikzpicture}
    % spherical proton
    \draw[dashed, xshift=-2cm] (0,0) circle (.8cm);
    % three partons
    \draw[dashed] (0,0) circle (.8cm);
    \foreach \theta in {0, 120, 240}{
      \draw ({\theta}:.39) circle (.4cm);
    }
    % ten partons
    \draw[dashed, xshift=2cm] (0,0) circle (.8cm);
    \foreach \theta/\radius in {
      36/0.4, 72/0.5, 108/0.5, 140/0.3, 172/0.1,
      190/0.5, 240/0.3, 272/0.5, 308/0.4, 360/0.5
    }{
      \draw[xshift=2cm] ({\theta}:\radius) circle (.2cm);
    }
  \end{tikzpicture}
  \caption{\label{proton_shapes} Schematic of plausible proton shapes.
  The sketch on the left shows a spherically symmetric proton (dashed line), while the middle and right illustrations depict a fluctuating proton with three and ten partons respectively (solid lines).
  }
\end{figure}

\section{Parameter estimation}

\subsection{Computer experiment design}

\subsection{Gaussian process emulators}

\subsection{Bayesian calibration}

Aliquam porttitor massa risus, hendrerit dignissim purus placerat vel. Suspendisse dui ante, porttitor pretium lorem et, placerat posuere mi. Donec lacinia nisl leo, ac luctus libero tincidunt non. Donec pretium ultricies felis vitae tempus. Duis pulvinar nisi justo, sit amet finibus orci volutpat non. Proin fermentum a magna id rhoncus. Vestibulum posuere, ipsum vel dignissim laoreet, arcu nibh venenatis neque, et aliquet dui dui a ipsum. Praesent sagittis condimentum neque, sed aliquam odio volutpat nec. Aenean consequat ante eget mauris lacinia, sed ultrices est dapibus. Morbi fermentum malesuada turpis, ac placerat est bibendum tincidunt. Nullam leo ipsum, sodales sit amet viverra non, iaculis vel nisl. Vestibulum purus ipsum, egestas ut bibendum sed, congue quis purus. Curabitur iaculis tortor nisl, nec venenatis tortor placerat a. Fusce feugiat nisl vitae porta porta. Praesent non arcu leo. Sed gravida eget neque in auctor.

Sed ac ullamcorper ipsum. Sed quis eleifend mauris. Phasellus quis massa neque. Proin nec lectus quis ante aliquam tincidunt ut sit amet eros. Nullam et odio eu elit aliquam molestie quis ut arcu. Aliquam gravida enim non tortor bibendum, at semper tellus egestas. In libero lectus, convallis in suscipit nec, bibendum eu felis. Vestibulum ante ipsum primis in faucibus orci luctus et ultrices posuere cubilia Curae; Vivamus quis lectus neque. Sed ac placerat justo. Nullam ultricies vulputate ligula, vel porta mauris scelerisque et. Sed hendrerit blandit semper. Sed vel bibendum elit, eu posuere elit. Phasellus nunc massa, tincidunt ac tempus quis, blandit quis erat.

\section{Results}

\subsection{Initial condition properties}

\subsection{QGP medium properties}

Aliquam porttitor massa risus, hendrerit dignissim purus placerat vel. Suspendisse dui ante, porttitor pretium lorem et, placerat posuere mi. Donec lacinia nisl leo, ac luctus libero tincidunt non. Donec pretium ultricies felis vitae tempus. Duis pulvinar nisi justo, sit amet finibus orci volutpat non. Proin fermentum a magna id rhoncus. Vestibulum posuere, ipsum vel dignissim laoreet, arcu nibh venenatis neque, et aliquet dui dui a ipsum. Praesent sagittis condimentum neque, sed aliquam odio volutpat nec. Aenean consequat ante eget mauris lacinia, sed ultrices est dapibus. Morbi fermentum malesuada turpis, ac placerat est bibendum tincidunt. Nullam leo ipsum, sodales sit amet viverra non, iaculis vel nisl. Vestibulum purus ipsum, egestas ut bibendum sed, congue quis purus. Curabitur iaculis tortor nisl, nec venenatis tortor placerat a. Fusce feugiat nisl vitae porta porta. Praesent non arcu leo. Sed gravida eget neque in auctor.

Sed ac ullamcorper ipsum. Sed quis eleifend mauris. Phasellus quis massa neque. Proin nec lectus quis ante aliquam tincidunt ut sit amet eros. Nullam et odio eu elit aliquam molestie quis ut arcu. Aliquam gravida enim non tortor bibendum, at semper tellus egestas. In libero lectus, convallis in suscipit nec, bibendum eu felis. Vestibulum ante ipsum primis in faucibus orci luctus et ultrices posuere cubilia Curae; Vivamus quis lectus neque. Sed ac placerat justo. Nullam ultricies vulputate ligula, vel porta mauris scelerisque et. Sed hendrerit blandit semper. Sed vel bibendum elit, eu posuere elit. Phasellus nunc massa, tincidunt ac tempus quis, blandit quis erat.

\section{Summary and conclusions}

Aliquam porttitor massa risus, hendrerit dignissim purus placerat vel. Suspendisse dui ante, porttitor pretium lorem et, placerat posuere mi. Donec lacinia nisl leo, ac luctus libero tincidunt non. Donec pretium ultricies felis vitae tempus. Duis pulvinar nisi justo, sit amet finibus orci volutpat non. Proin fermentum a magna id rhoncus. Vestibulum posuere, ipsum vel dignissim laoreet, arcu nibh venenatis neque, et aliquet dui dui a ipsum. Praesent sagittis condimentum neque, sed aliquam odio volutpat nec. Aenean consequat ante eget mauris lacinia, sed ultrices est dapibus. Morbi fermentum malesuada turpis, ac placerat est bibendum tincidunt. Nullam leo ipsum, sodales sit amet viverra non, iaculis vel nisl. Vestibulum purus ipsum, egestas ut bibendum sed, congue quis purus. Curabitur iaculis tortor nisl, nec venenatis tortor placerat a. Fusce feugiat nisl vitae porta porta. Praesent non arcu leo. Sed gravida eget neque in auctor.

Sed ac ullamcorper ipsum. Sed quis eleifend mauris. Phasellus quis massa neque. Proin nec lectus quis ante aliquam tincidunt ut sit amet eros. Nullam et odio eu elit aliquam molestie quis ut arcu. Aliquam gravida enim non tortor bibendum, at semper tellus egestas. In libero lectus, convallis in suscipit nec, bibendum eu felis. Vestibulum ante ipsum primis in faucibus orci luctus et ultrices posuere cubilia Curae; Vivamus quis lectus neque. Sed ac placerat justo. Nullam ultricies vulputate ligula, vel porta mauris scelerisque et. Sed hendrerit blandit semper. Sed vel bibendum elit, eu posuere elit. Phasellus nunc massa, tincidunt ac tempus quis, blandit quis erat.

\begin{acknowledgments}
  The authors thank ...
\end{acknowledgments}

\bibliography{trento-partons}

\end{document}

### 2018-03-21 src/__init__.py

@@ -24,7 +24,7 @@ cachedir.mkdir(parents=True, exist_ok=True)
 #: such as ``'PbPb2760'``, ``'AuAu200'``, ``'pPb5020'``.
 #: Even if the project uses only a single system,
 #: this should still be a list of one system string.
-systems = ['pPb5020', 'pPb5020']
+systems = ['pPb5020', 'PbPb5020']
 
 
 def parse_system(system):

### 2018-03-21 src/plots.py

@@ -159,14 +159,14 @@ def plot(f):
     return wrapper
 
 
-def figsize(relwidth=1, aspect=.618, refwidth=6, nrows=1, ncols=1):
+def figsize(relwidth=1, aspect=.618, refwidth=6):
     """
     Return figure dimensions from a relative width (to a reference width) and
     aspect ratio (default: 1/golden ratio).
 
     """
     width = relwidth * refwidth
-    return width, width*aspect/ncols*nrows
+    return width, width*aspect
 
 
 def set_tight(fig=None, **kwargs):
@@ -360,7 +360,6 @@ def _observables(posterior=False):
 
     fig, axes = plt.subplots(
         nrows=len(plots), ncols=len(systems),
-        # figsize=(.8*fullwidth, fullwidth),
         figsize=figsize(1.1, aspect=1.25),
         gridspec_kw=dict(
             height_ratios=[p.get('height_ratio', 1) for p in plots]

### 2018-03-24 paper/trento-partons.tex
merge in paper and talk

@@
 \subsection{Bayesian calibration}
+\fig{observables_design}{Caption}
 
@@
 \subsection{Initial condition properties}
+
+\begin{figure}
+  \begin{tikzpicture}
+    \foreach \x in {0pt, 10pt, 20pt}{\draw (\x, 0) ellipse (5pt and 20pt);}
+    \foreach \x in {0pt, 10pt, 20pt, 30pt}{\draw (\x+100pt, 0) ellipse (5pt and 20pt);}
+  \end{tikzpicture}
+  \caption{\label{fig:collision} This is the figure caption}
+\end{figure}
+Consider a relativistic collision between a columnated stack of $n_1$ projectile nucleons and $n_2$ target nucleons as depicted in Fig.~\ref{fig:collision}.
+Moreover, assume that each nucleon has mass $m$ and moves with velocity $\pm u$ along the beam axis $\hat{z}$.
+The four-momentum of each colliding nucleon stack is given by
+\begin{align}
+  p_1 &= n_1 \gamma m\, (1, 0, 0, u),\\
+  p_2 &= n_2 \gamma m\, (1, 0, 0, -u),
+\end{align}
+where $\gamma = 1/\sqrt{1 - u^2}$ is the usual Lorentz factor.
+
+At relativistic energies $\gamma \gg 1$, each stack of nucleons is significantly Lorentz contracted in the lab frame and appears compressed along its direction of motion.
+The nucleons are thus observed to penetrate and pass through each other in the lab frame in time
+\begin{equation}
+  \Delta t = L / \sqrt{\gamma^2 - 1}, 
+\end{equation}
+where $L\sim1$--10~fm is the length of the columnated stack of nucleons in its local rest frame.
+
+Collisions at top RHIC and LHC energies have interpenetration times \mbox{$\Delta t < .1$~fm/$c$} and hence transpire before meaningful dynamics develop in the transverse plane.
+For this reason, it is convenient to say that initial energy deposition occurs instantaneously at time $\Delta t = 0^+$, the moment the two nuclei collide.
+The initial energy (or entropy) deposited into the transverse plane is then modeled in terms of two beam-integrated nuclear thickness functions
+\begin{equation}
+  T_{A,B}(x, y) = \int dz \rho(x, y, z),
+\end{equation}
+where $\rho$ is the density profile of the nucleus in its local rest frame.
 

### 2018-03-24 src/plots.py

@@ -20,6 +20,7 @@ decorator.
 from collections import OrderedDict
 import itertools
 import logging
+import os
 from pathlib import Path
 import subprocess
 import tempfile
@@ -32,12 +33,12 @@ import matplotlib.pyplot as plt
 from matplotlib import lines
 from matplotlib import patches
 from matplotlib import ticker
-from matplotlib import colors
 from matplotlib import cm
+from matplotlib.colors import ListedColormap
 
 from scipy import special
-from scipy.interpolate import PchipInterpolator
-from scipy.ndimage.measurements import center_of_mass
+from scipy.interpolate import interp1d, PchipInterpolator
+from scipy.optimize import brentq
 from sklearn.decomposition import PCA
 from sklearn.gaussian_process import GaussianProcessRegressor as GPR
 from sklearn.gaussian_process import kernels
@@ -47,16 +48,13 @@ from . import workdir, systems, parse_system, expt, model, mcmc
 from .design import Design
 from .emulator import emulators
 
+import freestream
 
-# fontsmall, fontnormal, fontlarge = 5, 6, 7
-# XXX deprecated variables
+
+# golden ratio
 aspect = 1/1.618
-resolution = 72.27
-textwidth = 307.28987/resolution
-textheight = 261.39864/resolution
-fullwidth = 350/resolution
-fullheight = 270/resolution
 
+# font sizes
 fontsize = dict(
     large=11,
     normal=10,
@@ -80,6 +78,7 @@ colors = OrderedDict([
 ])
 
 offblack = '.15'
+theme = '#b01419'
 
 plt.rcdefaults()
 plt.rcParams.update({
@@ -147,9 +146,6 @@ def plot(f):
 
         fig = plt.gcf()
 
-        if not fig.get_tight_layout():
-            set_tight(fig)
-
         plotfile = plotdir / '{}.pdf'.format(f.__name__)
         fig.savefig(str(plotfile))
         logging.info('wrote %s', plotfile)
@@ -158,6 +154,34 @@ def plot(f):
     plot_functions[f.__name__] = wrapper
 
     return wrapper
+
+
+def cmap_to_alpha(cmap=plt.cm.inferno, fraction=.2):
+    """
+    Fade the bottom of a colormap to white.
+
[…]

### 2018-03-24 talk/qm2018.tex
add talk

\documentclass[aspectratio=169]{beamer}

\definecolor{theme}{HTML}{b01419}
\definecolor{accent}{HTML}{0059b3}
\definecolor{offblack}{HTML}{231f20}
\definecolor{link}{HTML}{176ECC}
\definecolor{green}{HTML}{33CC33}
\definecolor{msu}{HTML}{18453B}
\definecolor{duke}{HTML}{001A57}
\setbeamercolor{normal text}{fg=offblack}

\usecolortheme[named=theme]{structure}
\usecolortheme{rose}
\usecolortheme{dolphin}

\makeatletter
\setbeamertemplate{frametitle}{
  \ifbeamercolorempty[bg]{frametitle}{}{\nointerlineskip}%
  \@tempdima=\textwidth%
  \advance\@tempdima by\beamer@leftmargin%
  \advance\@tempdima by\beamer@rightmargin%
  \begin{beamercolorbox}[sep=0.3cm, left, wd=\the\@tempdima]{frametitle}
    \usebeamerfont{frametitle}
    \vbox{}\vskip-2ex%
    \if@tempswa\else\csname beamer@fteleft\endcsname\fi%
    \strut\insertframetitle\strut\par%
    {%
      \ifx\insertframesubtitle\@empty%
      \else%
      {\usebeamerfont{framesubtitle}\usebeamercolor[fg]{
        framesubtitle}\insertframesubtitle\strut\par}%
      \fi
    }%
    \vskip.45ex%
    \hrule %height .6pt%
    \vskip-1.45ex%
    \if@tempswa\else\vskip-.3cm\fi%
  \end{beamercolorbox}%
}
\makeatother

% clean up footer
\beamertemplatenavigationsymbolsempty
\setbeamertemplate{footline}[text line]{
  \parbox{\linewidth}{\vspace*{-10pt}
  \raggedleft \color{offblack} 
  \bf J.\ Scott Moreland (Duke)
  \quad \insertframenumber\,/\,\inserttotalframenumber}
}

% inner theme
\useinnertheme{rectangles}
\setbeamertemplate{itemize item}{
  \raise.30ex\hbox{\vrule width .80ex height .80ex}}
\setbeamertemplate{itemize subitem}{
  \raise.35ex\hbox{\vrule width .70ex height .70ex}}

% for backup slides
\usepackage{appendixnumberbeamer}
\renewcommand\appendixname{Backup}

\usepackage{graphicx}
\usepackage[absolute, overlay]{textpos}
\graphicspath{{../plots/}{fig/}}

\usepackage{amsmath}
\usepackage{amssymb}

\usepackage[T1]{fontenc}
\usepackage[default]{lato}

\usepackage{tikz}
\usetikzlibrary{calc}
\usetikzlibrary{positioning}
\usetikzlibrary{overlay-beamer-styles}

\usepackage[absolute]{textpos}

\usepackage{fontawesome}
\usepackage{ulem}
\newcommand{\soutthick}[1]{%
  \renewcommand{\ULthickness}{.8pt}%
  \sout{#1}%
  }

\newcommand{\trento}{T\raisebox{-0.3ex}{R}ENTo}
\newcommand{\T}{\tilde{T}}
\newcommand{\sqrts}{\sqrt{s_{NN}}}
\newcommand{\sigmaf}{\sigma_\text{fluct}}
\newcommand{\X}{\chi_\text{struct}}
\newcommand{\npartons}{n_\text{partons}}
\newcommand{\dmin}{d_\text{min}^3}
\newcommand{\conference}{\scshape Quark Matter \enskip\textbar\enskip Venice, Italy}
\newcommand{\paddedhline}{\noalign{\smallskip}\hline\noalign{\smallskip}}
\newcommand{\boundellipse}[3]% center, x rad, y rad
{(#1) ellipse [x radius=#2, y radius=#3]
}

\title{Estimating nucleon substructure properties in a\\[.2ex]
unified hydrodynamic model of p-Pb and Pb-Pb collisions}
\author[J.\ S.\ Moreland]{\scshape J.\ S.\ Moreland, J.\ E.\ Bernhard \& S.\ A.\ Bass}
\institute[Duke]{\scshape Duke University}
\date{\scshape 14 May 2018}

\begin{document}

\section{Title}

\begin{frame}[b, plain, noframenumbering]
  \centering
  \textcolor{theme}{\Large \inserttitle} \\[3ex]
  \insertauthor \\
  \insertinstitute \\[2ex]
  \conference \\
  \insertdate \\[6ex] 
  \includegraphics[height=1.2cm]{qcdlogo}\quad
  \includegraphics[height=1.2cm]{nersclogo}\\[3ex]
  \footnotetext[1]{\scriptsize This research used resources of the National Energy Research Scientific Computing Center, a DOE Office of Science User Facility supported by the Office of Science of the U.S.\ Department of Energy under Contract No. DE-AC02-05CH11231.}
\end{frame}

\tikzset{
  box/.style={
    align=center,
    inner sep=1ex,
    background default fill=black!10,
    background default text=black!30,
    background fill=theme!10,
    background text=theme,
  }
}

\newcommand{\boxtitle}[1]{\textbf{#1}\\[.5ex]}

\begin{frame}{\scshape Flow in small systems}
  \smallskip
  \begin{columns}[T]
    \begin{column}{.62\textwidth}
      \bigskip
      Important questions:\\
      \begin{itemize}
        \item Is hydro applicable to small collision systems?
        \item If so, can we fit small system collectivity using current hydro models?
      \end{itemize}
      \smallskip
      First question precludes the second, but also more difficult to answer.\\[1ex]
      Let's tackle the second question \textit{assuming} hydrodynamic models make
      sense down to length scales $\Delta x=.2$~fm.\\[1ex]
      Under this assumption, does a unified hydrodynamic description of p+A and A+A data even exist?
    \end{column}
    \begin{column}{.3\textwidth}
      \centering
      \includegraphics[width=\columnwidth]{pbpb_corr2d}\\
      \includegraphics[width=\columnwidth]{ppb_corr2d}
    \end{column}
  \end{columns}

\end{frame}

\begin{frame}[plain, noframenumbering]
  \begin{center}
    \Large I. Bayesian parameter estimation of heavy-ion collisions
  \end{center}
\end{frame}

\begin{frame}[t]{\scshape Heavy-ion collision model}
  \centering
  \medskip
  \begin{tikzpicture}
    \node (a) {\includegraphics{hannah_collision}};
    \node[align=left, anchor=north west] at (a.north west) {%
      \scriptsize Figure: H. Petersen, MADAI
      };
  \end{tikzpicture}
  \begin{columns}[T]
    \begin{column}{0.3\textwidth}
      \texttt{\scshape\color{theme} TrENTo} {\small initial conditions}\\
      \smallskip\hrule\medskip
      {\scriptsize parametric entropy deposition}\\
      {\tiny \color{theme}PRC.92.011901}\\[2ex]
      \texttt{\scshape\color{theme} freestream} {\small pre-flow}\\
      \smallskip\hrule\medskip
      {\scriptsize infinitely weak coupling limit}\\
      {\tiny \color{theme} PRC.91.064906, ~PRC.80.034902}
    \end{column}
    \hfill
    \begin{column}{0.3\textwidth}
      \texttt{\scshape\color{theme} VISH2+1} {\small viscous hydro}\\
      \smallskip\hrule\medskip
      {\scriptsize 14-mom.\ approx w/ shear \& bulk}\\
      {\tiny \color{theme} PRC.77.064901, ~J.CPC.2015.08.039}\\[2ex]
      \texttt{\scshape\color{theme}frzout} {\small sampler}\\
      \smallskip\hrule\medskip
      {\scriptsize non-RTA shear \& bulk corrections}\\
      {\tiny\color{theme} J.E.\ Bernhard thesis}
    \end{column}
    \hfill
    \begin{column}{0.3\textwidth}
      \texttt{\scshape\color{theme} UrQMD} {\small Boltzmann cascade}
      \smallskip\hrule\medskip
      {\scriptsize hadronic afterburner, simulate scatterings and decays}\\
      {\tiny\color{theme} PPNP.98.00058, ~JPG.25.9.308}\\[2ex]
      {\scriptsize Final observables calculated as similar as possible to experiment}
    \end{column}
  \end{columns}
  \begin{tikzpicture}[overlay]
    \draw (-3, -.2) -- (-2.7, -.2);
    \draw (-2.7, -.2) -- (-2.7, 3.6);
    \draw (-2.7, 3.6) -- (-2.4, 3.6);
    \draw (2.4, -.2) -- (2.7, -.2);
    \draw (2.7, -.2) -- (2.7, 3.6);
    \draw (2.7, 3.6) -- (3.0, 3.6);
  \end{tikzpicture}
\end{frame}

\begin{frame}[t]{\scshape Bayesian parameter estimation}
  \makebox[\textwidth]{
    \begin{tikzpicture}[overlay]
      \node[box=1, font=\small] (input) at (-5, -1) {
      \boxtitle{Input parameters}
      IC and QGP properties
    };
    \node[box, below=of input.south west, anchor=north west, yshift=.6cm, font=\small] (physics) {
      \boxtitle{Physics theory}
      initial stages, hydro, and\\
      Boltzmann transport
    };
    \node[box, right=of input.east, anchor=west, xshift=-.2cm, font=\small] (model) {
      \boxtitle{Computer model}
      minimum bias event-by-\\
      event simulations
    };
    \node[box, right=of model.east, anchor=west, xshift=-.2cm, font=\small] (gp) {
      \boxtitle{Gaussian process emulator}
      surrogate model
    };
    \node[box, below=of gp.south, anchor=north, font=\small] (mcmc) {
      \boxtitle{MCMC}
      calibrate model to data
    };
    \node[box, below=of mcmc.south, anchor=north, font=\small] (posterior) {
      \boxtitle{Posterior distribution}
      quantitative estimates \\
      of each parameter
    };
    \node[box, left=of mcmc.west, anchor=east, font=\small] (exp) {
      \boxtitle{Experimental data}
      yields, mean $p_T$, flows
    };
    \begin{scope}[color=black!70, ->, thick]
      \draw (input) -- (model);
      \draw ([yshift=-.2cm]physics.north east) -| (model.south);
      \draw (model) -- (gp);
      \draw (gp) -- (mcmc);
      \draw (exp) |- (mcmc);
      \draw (mcmc) -- (posterior);
    \end{scope}
  \end{tikzpicture}
  }
  \begin{textblock}{10}(1, 10)
    {\scshape Objective:}
    \begin{itemize}
        \item Explore parameter space of a given model
        \item Quantify posterior probability of every parameter region
          \textit{given the model, data and \textbf{known uncertainties}}
      \end{itemize}
  \end{textblock}
\end{frame}

\begin{frame}[b, plain]{\scshape Parameter estimates from lead-lead collisions}
  \begin{columns}[b]
    \begin{column}{.4\textwidth}
      \begin{tikzpicture}
        \node (A) {\includegraphics<1>[height=.82\textheight]{observables_design}};
        \node (B) {\includegraphics<2>[height=.82\textheight]{observables_posterior}};
        \node [xshift=2.8cm, yshift=3.2cm, rotate=-45] {%
          \textcolor{theme}{\scshape New!}%
          };
      \end{tikzpicture}
    \end{column}
    \begin{column}{.02\textwidth}
    \end{column}
    \begin{column}{.55\textwidth}
      \begin{itemize}
        \item Combined analysis of 2.76 and 5.02 TeV beam energies
        \item Parametric initial conditions and medium parameters
        \item See thesis by Jonah Bernhard {\small [\url{99999}]}
      \end{itemize}
      \begin{tikzpicture}
        \node (0, 0) {\includegraphics<1>{region_shear_blank}};
        \node (0, 0) {\includegraphics<2>{region_shear}};
        \node [xshift=.3cm, yshift=.6cm] {\only<1>{\fontsize{80}{0}\selectfont \color{theme!30} ?}};
        \node [xshift=3cm, yshift=1.8cm, rotate=-45] {%
          \textcolor{theme}{\scshape New!}%
          };
      \end{tikzpicture}
    \end{column}
  \end{columns}
  \medskip
\end{frame}

\begin{frame}[plain, noframenumbering]
  \begin{center}
    \Large II. Adding nucleon substructure to the model
  \end{center}
\end{frame}

\begin{frame}[t]{\scshape Nuclear structure}
  \begin{columns}[T]
    \begin{column}{.4\textwidth}
      \begin{figure}
        \includegraphics[width=.9\columnwidth]{lead_208}
        \caption{Pb$^{208}$ nucleus}
      \end{figure}
    \end{column}
    \begin{column}{.6\textwidth}
      \bigskip
      \textcolor{theme}{Original \trento\ model:}
      \begin{itemize}
        \item Sample nucleon radii from spherical or deformed Woods-Saxon distributions
        \item Euler angles resampled to preserve minimum nucleon distance
          $d_\mathrm{min}$ [fm]
        \item Gaussian nucleons of width $w$ [fm]
      \end{itemize}
      \medskip
      \textcolor{theme}{This work:}
      \begin{itemize}
        \item Trade Gaussian nucleon for lumpy nucleon
      \end{itemize}
      \begin{tikzpicture}[absolute, overlay]
        \node (a) at (2, -.75) {\includegraphics[scale=.5]{proton}};
        \node[right=of a.east] (b) {\includegraphics[scale=.5]{proton_partons}};
        \draw[->, thick] (a) -- (b);
      \end{tikzpicture}
    \end{column}
  \end{columns}
\end{frame}

\begin{frame}[t, plain]{\scshape Nucleon substructure}
  \vspace{2em}
  \begin{columns}[t]
    \begin{column}{.2\textwidth}
      \centering \small Sampling radius
      \begin{figure}
        \begin{tikzpicture}[scale=.9]
          \draw[thick] (0,0) circle [radius= 1cm] node (A) {\normalsize 1 fm};
          \draw[thick] (0,-2) circle [radius=.8cm] node (B) {\small .8 fm};
          \draw[thick] (0,-3.6) circle [radius=.6cm] node (C) {\scriptsize .6 fm};
          \draw [thick] (0,-4.8) circle [radius=.4cm] node (D) {\tiny .4 fm};
        \end{tikzpicture}
      \end{figure}
    \end{column}
    \begin{column}{.2\textwidth}
      \centering \small Parton width
      \begin{figure}
        \begin{tikzpicture}[scale=.9]
          \draw [thick, dashed] (0, 0) circle (1cm);
          \foreach \theta in {0, 120, 240}{
            \draw [thick] ({\theta}:.3) circle (.6cm);
          }
          \draw [thick, dashed] (0, -2.2) circle (1cm);
          \foreach \theta in {0, 120, 240}{
            \draw [thick] ({\theta}:.39) [yshift=-2.2cm] circle (.4cm);
          }
          \draw [thick, dashed] (0, -4.4) circle (1cm);
          \foreach \theta in {0, 120, 240}{
            \draw [thick] ({\theta}:.39) [yshift=-4.4cm] circle (.2cm);
          }
        \end{tikzpicture}
      \end{figure}
    \end{column}
    \begin{column}{.2\textwidth}
      \centering \small Parton number
      \begin{figure}
        \begin{tikzpicture}[scale=.9]
          \draw [thick, dashed] (0,0) circle (1cm);
          \foreach \theta/\radius in {
            36/0.48, 72/0.6, 108/0.6, 140/0.36, 172/0.12,
            190/0.6, 240/0.36, 272/0.6, 308/0.48, 360/0.6
          }{
            \draw [thick] ({\theta}:\radius) circle (.3cm);
          }
          \draw [thick, dashed] (0, -2.2) circle (1cm);
          \foreach \theta in {0, 120, 240}{
            \draw [thick] ({\theta}:.39) [yshift=-2.2cm] circle (.3cm);
          }
          \draw [thick, dashed] (0, -4.4) circle (1cm);
          \foreach \theta in {0, 180}{
            \draw [thick] ({\theta}:.39) [yshift=-4.4cm] circle (.3cm);
          }
        \end{tikzpicture}
      \end{figure}
    \end{column}
    \begin{column}{.4\textwidth}
      \centering
      \textcolor{theme}{Free parameters:}\\[1ex]
      \begin{itemize}
        \small
        \item Width of the Gaussian distribution used to sample
          parton radial coordinates
        \item Number of Gaussian partons inside each nucleon
        \item Width of the Gaussian partons
      \end{itemize}
      \bigskip
      \textcolor{theme}{Absent from this work:}\\[1ex]
      \begin{itemize}
        \small
        \item Parton spatial correlations,\\
          see talk by Alba Soto Ontoso 
      \end{itemize}
    \end{column}
  \end{columns}
\end{frame}

\begin{frame}[t]{\scshape Cross sections}
  \begin{columns}[T]
    \begin{column}{.5\textwidth}
      \begin{figure}
        \includegraphics{trento_model_part}\\
        \small MC-Glauber cross sections\\
        (proton d.o.f. shown)
      \end{figure}
    \end{column}
    \begin{column}{.5\textwidth}
      \centering \bigskip
      \textcolor{theme}{Original \trento\ model:}
      \begin{align*}
        P_\text{coll} &= 1 - \exp[-\sigma_{gg} T_{pp}(b)] \\
        \sigma_\text{nn}^\text{inel} &= \int 2\pi b\, db\, P_\text{coll}(b)
      \end{align*}
      \only<1>{\includegraphics<1>[scale=.8]{collision_profile}}
      \only<2>{
        \textcolor{theme}{This work:}
        \begin{equation*}
          P_\text{coll} \rightarrow 1 - \prod_{i,j=1}^{n_\text{part}}
          [1 - P_\text{coll}(b_{ij})]
        \end{equation*}
        Solve for opacity parameter $\sigma_{gg}$ numerically
      }
    \end{column}
  \end{columns}
\end{frame}

\begin{frame}[t]{\scshape Participant thickness functions}
  \begin{columns}[T]
    \begin{column}{.5\textwidth}
      \begin{figure}
        \includegraphics{trento_model_thickness}
      \end{figure}
    \end{column}
    \begin{column}{.5\textwidth}
      \centering \bigskip
      \textcolor{theme}{Original \trento\ model:}
      \begin{equation*}
        T_\text{nucleus}(x, y) = \sum\limits_{i=1}^{N_\text{part}}
        \gamma_i\, T_\text{proton}(x - x_i, y - y_i)
      \end{equation*}
      {\small Random weight $\gamma_i$ sampled from Gamma distribution with unit mean
      and variance $1/k$.}\\[2ex]
      \textcolor{theme}{This work:}
      \begin{equation*}
        \gamma_i\, T_\text{proton} \rightarrow \sum\limits_{j=1}^{N_\text{parton}}
        \gamma_j\,T_\text{parton} 
      \end{equation*}
    \end{column}
  \end{columns}
\end{frame}

\begin{frame}[t]{\scshape Entropy deposition}
  \begin{columns}[T]
    \begin{column}{.5\textwidth}
      \begin{figure}
        \includegraphics{trento_model_entropy}
      \end{figure}
    \end{column}
    \begin{column}{.5\textwidth}
      \bigskip \centering
      \textcolor{theme}{Generalized Mean Ansatz:}
      \begin{equation*}
        \frac{dS}{d\eta} \propto T_R(T_A, T_B) \equiv
        \left(\frac{T_A^p + T_B^p}{2}\right)^{1/p}
      \end{equation*}
      \begin{equation*}
        \small
        T_R =
        \begin{cases}
          \max(T_A, T_B) & p \rightarrow +\infty, \\[.5ex]
          (T_A + T_B)/2 & p = +1, \\[.5ex]
          \sqrt{T_A T_B} & p = 0, \\[.5ex]
          2\, T_A T_B/(T_A + T_B) & p = -1, \\[.5ex]
          \min(T_A, T_B) & p \rightarrow -\infty.
        \end{cases}
      \end{equation*}
      \smallskip
      \begin{center}
        \small $T$ denotes \emph{participant} thickness function
      \end{center}
    \end{column}
  \end{columns}
\end{frame}

\begin{frame}[t]{\scshape Free streaming}
  \begin{columns}[T]
    \begin{column}{.5\textwidth}
      \begin{figure}
        \includegraphics{trento_model_freestream}\\
        \small arrows: fluid velocity\\weighted by energy density
      \end{figure}
    \end{column}
    \begin{column}{.5\textwidth}
      \bigskip \centering
      \textcolor{theme}{Pre-equillibrium phase:}
      \begin{itemize}
        \small
        \item Massless non-interacting parton gas matched to viscous hydrodynamics
          \textcolor{theme}{\scriptsize PRC.91.064906, PRC.80.034902}.
        \item Must reinterpret initial entropy density as initial gluon density,\,
          $dN_g/y \sim dS/dy$
        \item Initialize hydro with non-zero $u^\mu$ and $\pi^{\mu\nu}$
      \end{itemize}
      \medskip
      \includegraphics{coupling}
    \end{column}
  \end{columns}
\end{frame}

\begin{frame}[plain, noframenumbering]
  \begin{center}
    \Large III. Simultaneous calibration to\\
    p-Pb and Pb-Pb collisions at $\sqrts=5.02$ TeV
  \end{center}
\end{frame}

\begin{frame}[t]{\scshape Computer experiment design}
  \bigskip
  \begin{columns}[T]
    \begin{column}{.55\textwidth}
      \scriptsize
      \begin{tabular}{lll}
        Parameter         & Description                        & Range           \\
        \paddedhline
        Norm              & Overall normalization              & 12--28          \\
        $p$               & Entropy deposition parameter       & $-1$ to $+1$    \\
        $\sigmaf$         & Relative parton fluct. std.\       & 0--2            \\
        $w$               & Parton sampling radius             & 0.4--1.2 fm     \\
        $\X$              & Parton structure parameter         & 0--1            \\
        $\npartons$       & Number of partons                  & 1--10           \\
        $\dmin$           & Nucleon exclusion volume           & 0--4.9 fm$^3$   \\
        $\tau_{fs}$       & Free streaming time                & 0.1--1.5 fm/$c$ \\
        $\eta/s$ min      & Shear viscosity at $T_c$           & 0--0.2          \\
        $\eta/s$ slope    & Slope above $T_c$                  & 0--8 GeV$^{-1}$ \\
        $\eta/s$ crv      & Curvature above $T_c$              & $-1$ to $1$     \\
        $\zeta/s$ norm    & Bulk viscosity peak height         & 0--0.1          \\
        $\zeta/s$ width   & Bulk viscosity peak width          & 0--0.1 GeV      \\
        $\zeta/s$ temp    & Bulk viscosity peak location       & 150--200 MeV    \\
        $T_\text{switch}$ & Particlization temperature         & 135--165 MeV    \\
      \end{tabular}
    \end{column}
    \begin{column}{.45\textwidth}
      \textcolor{theme}{\scshape Important details:}
      \smallskip
      \begin{itemize}
        \small
        \item Parton width is reparametrized:\\[1ex]
          $v = v_\text{min} + \X(v_\text{max} - v_\text{min})$\\[1ex]
          $v_\text{min} = 0.2$~fm,~ $v_\text{max} = w$\\[1ex]
        \item Parton sampling radius is \emph{not} equivalent to the proton radius
          in the proton c.o.m.\ frame, e.g.\ see\\
          \textcolor{theme}{\scriptsize PRC.94.024919}\\[1ex]
        \item Other parameters same as Pb+Pb analysis at 2.76 and 5.02 TeV
      \end{itemize}
    \end{column}
  \end{columns}
  \begin{center}
    \color{theme}
    500 design points per collision system, $\mathcal{O}(10^4)$ events per design point!
  \end{center}
\end{frame}

\begin{frame}[plain]
  \begin{tikzpicture}[remember picture, overlay]
    \node[anchor=west] at (current page.west) {
      \includegraphics[height=.95\paperheight]{observables_design}
      };
    \node[anchor=east] at (current page.east) {
      \includegraphics[height=.95\paperheight]{observables_design}
      };
    \draw[<-] (current page.center) ++(-.75, 1.7) --
      node[above] {\footnotesize Prior}
      node[below] {\tiny (training data)}
      ++(1.5, 0);
    \draw[->] (current page.center) ++(-.75, -1.2) --
      node[above] {\footnotesize Posterior}
      node[below] {\tiny (samples)}
      ++(1.5, 0);
  \end{tikzpicture}
\end{frame}

\begin{frame}[plain]
  \begin{tikzpicture}
    \node {\includegraphics[height=\textheight]{posterior}};
    \node (title) at (6.5, 3.5) {\color{theme}{\scshape \Large Bayesian posterior}};
    \node<1-2>[draw=none, text width=7cm, align=left, below=.2cm of title] (b1) {%
      One row (column) for each parameter
      };
    \node<1-2>[draw=none, text width=7cm, align=left, below=.2cm of b1] (b2) {%
      Diagonal panels: marginal distribution of a single model parameter
      };
    \node<1-2>[draw=none, text width=7cm, align=left, below=.2cm of b2] (b3) {%
      Off-diagonal panels: joint distribution between a pair of model parameters
      };
    \node<2>[draw=none, text width=7cm, align=left, below=.4cm of b3] (b4) {%
      Let's zoom in on some specific regions...
      };
    \draw<1-2>[->, thick, theme] (b2.west) to [out=180, in=90] (-.3, 1.2);
    \draw<1-2>[->, thick, theme] (b3.west) to [out=180, in=90] (.2, -.5);
    \node<3>[xshift=.5cm, yshift=-.5cm, scale=.3] {\includegraphics{magnifying-glass}};
    \node<4>[xshift=-.5cm, yshift=-1cm, scale=.3] {\includegraphics{magnifying-glass}};
  \end{tikzpicture}
\end{frame}

\begin{frame}{\scshape Maximum a posteriori parameters}

\end{frame}

\begin{frame}{\scshape Numerical grid error}
  \begin{figure}
    \includegraphics[width=\textwidth]{grid_error}
    \caption{Numerical grid error tame for $dxy=.2 \times \text{parton width}$}
  \end{figure}
\end{frame}

\usebackgroundtemplate{}
\end{document}

### 2018-03-30 paper/trento-partons.tex
Add some content to paper on trento

--- 2018-03-21 paper/trento-partons.tex
@@
 
-\usepackage{hyperref}
+\usepackage[pdfencoding=auto, psdextra]{hyperref}
+
 \hypersetup{
@@
 
+% hyperref throws warning when a line break is used in the title
+% this removes the warning by redefining the line break in a pdf string
+\pdfstringdefDisableCommands{%
+  \def\\#1{ #1}
+}
+
 % Convenient figure macro.  Usage:
@@
   Estimating initial state and quark-gluon plasma medium properties\\
-  using a hydrodynamic model with nucleon substructure\\calibrated to p-Pb and Pb-Pb collisions at $\sqrts=5.02$~TeV
+  using a hydrodynamic model with nucleon substructure\\
+  calibrated to p-Pb and Pb-Pb collisions \texorpdfstring{$\mathbf{\sqrts=5.02}$}{}~TeV
 }
@@
 \author{Steffen A.\ Bass}
+
 \affiliation{Department of Physics, Duke University, Durham, NC 27708-0305}
@@
 \begin{abstract}
-  We posit a unified hydrodynamic description of ultrarelativistic $p$-Pb and Pb-Pb nuclear collisions at $\sqrts=5.02$~TeV and evaluate the likelihood of our assertion using Bayesian inference.
+  We posit a unified hydrodynamic description of ultrarelativistic p-Pb and Pb-Pb nuclear collisions at $\sqrts=5.02$~TeV and evaluate our assertion using Bayesian inference.
 Specifically, we model the dynamics of both collision systems using QGP initial conditions with parametric nucleon substructure, a pre-equilibrium free-streaming stage, event-by-event viscous hydrodynamics, and a microscopic hadronic afterburner.
 Free parameters of the model which describe the initial state and QGP medium are then simultaneously calibrated to fit charged particle yields, mean $p_T$ and flow cumulants of both systems.
-  We argue that the global agreement of the calibrated model with available experimental data strongly supports the existence of hydrodynamic flow in small collision systems at ultrarelativistic energies, and that this flow necessarily develops at length scales smaller than a proton.
+We argue that the global agreement of the calibrated model with the experimental data strongly supports the existence of hydrodynamic flow in small collision systems at ultrarelativistic energies, and that this flow necessarily develops at length scales smaller than a proton.
 Finally, we present posterior distributions on our model's initial state and QGP medium parameters, including new constraints on the shape and structure of the proton.
@@
 \section{Introduction}
-  Ultrarelativistic collisions of light and light-heavy \mbox{(e.g.\ $p$-$p$ and $p$-Pb)} ion pairs produce long-range multiparticle correlations which are striking similar to those observed in heavy-ion collisions \note, where the observed collectivity is commonly attributed to hydrodynamic flow \note.
-  This observation suggests that hydrodynamic behavior could be manifest in small droplets of quark-gluon-plasma (QGP), and that flow might develop at length scales smaller than a proton.
 
- Lorem ipsum dolor sit amet, consectetur adipiscing elit.
-In tincidunt elit a libero luctus, ut pharetra sem accumsan.
-Nullam eros dui, egestas ut arcu et, commodo suscipit purus.
-Quisque dapibus tincidunt nulla at vehicula.
-Nulla ut libero sapien.
-Suspendisse ac consectetur metus.
-Interdum et malesuada fames ac ante ipsum primis in faucibus.
-Nunc facilisis volutpat suscipit.
-Pellentesque nec est orci.
-Donec id pharetra enim.
+  Ultrarelativistic nuclear collisions between one light-ion \mbox{(e.g.\ p, d, or ${}^3$He)} and one heavy-ion \mbox{(e.g.\ Au or Pb)} generate dense, compact sources of nuclear matter which produce long-range multiparticle correlations that are strikingly similar to the correlations observed in heavy-ion collisions, signals which are most commonly cited as evidence for hydrodynamic flow \note.
+This observation suggests that hydrodynamic behavior could be manifest in small droplets of quark-gluon-plasma (QGP), and that flow might develop at length scales smaller than a single proton \note.
 
-Aliquam porttitor massa risus, hendrerit dignissim purus placerat vel.
-Suspendisse dui ante, porttitor pretium lorem et, placerat posuere mi.
-Donec lacinia nisl leo, ac luctus libero tincidunt non.
-Donec pretium ultricies felis vitae tempus.
-Duis pulvinar nisi justo, sit amet finibus orci volutpat non.
-Proin fermentum a magna id rhoncus.
-Vestibulum posuere, ipsum vel dignissim laoreet, arcu nibh venenatis neque, et aliquet dui dui a ipsum.
-Praesent sagittis condimentum neque, sed aliquam odio volutpat nec.
-Aenean consequat ante eget mauris lacinia, sed ultrices est dapibus.
-Morbi fermentum malesuada turpis, ac placerat est bibendum tincidunt.
-Nullam leo ipsum, sodales sit amet viverra non, iaculis vel nisl.
-Vestibulum purus ipsum, egestas ut bibendum sed, congue quis purus.
-Curabitur iaculis tortor nisl, nec venenatis tortor placerat a.
-Fusce feugiat nisl vitae porta porta.
-Praesent non arcu leo.
-Sed gravida eget neque in auctor.
+  Hydrodynamic models of ultrarelativistic nuclear collisions are complicated by a number of theoretical unknowns, including the detailed geometry of the QGP initial conditions, the strength and duration of pre-equilibrium dynamics, the temperature dependence of QGP transport coefficients, and the boundaries of hydrodynamic applicability \note.
+In general, these theoretical uncertainties tend to grow with decreasing system size, where emergent physics at partonic length scales becomes important to describing bulk properties of the produced system.
 
-Sed ac ullamcorper ipsum.
-Sed quis eleifend mauris.
-Phasellus quis massa neque.
-Proin nec lectus quis ante aliquam tincidunt ut sit amet eros.
-Nullam et odio eu elit aliquam molestie quis ut arcu.
-Aliquam gravida enim non tortor bibendum, at semper tellus egestas.
-In libero lectus, convallis in suscipit nec, bibendum eu felis.
-Vestibulum ante ipsum primis in faucibus orci luctus et ultrices posuere cubilia Curae; Vivamus quis lectus neque.
-Sed ac placerat justo.
-Nullam ultricies vulputate ligula, vel porta mauris scelerisque et.
-Sed hendrerit blandit semper.
-Sed vel bibendum elit, eu posuere elit.
-Phasellus nunc massa, tincidunt ac tempus quis, blandit quis erat.
+  One method for reducing theoretical uncertainties is to test model calculations by varying the species of colliding nuclei at a single beam energy \note.
+Since initial condition and hydrodynamic models generally factorize the structure of the colliding nuclei from the subsequent time dynamics of the collision, a single theory framework can be simultaneously tested and compared to measurements from multiple collision systems using a self consistent set of model parameters where only the nuclear structure in the model is permitted to vary.
+
+  Typically, heavy-ion nuclear structure, characterized e.g.\ by an atomic mass and set of Woods-Saxon parameters, is regarded as a known input to hydrodynamic models which contributes negligible uncertainty to overall model predictions.
+Collisions involving protons and light-ions, meanwhile, are sensitive to detailed geometric properties of the proton \note, such as spatial inhomogeneities inside the proton which are not well constrained by nuclear structure measurements.
+
+One approach to investigate the effect of nucleon substructure on hydrodynamic models is to fix or hand tune initial condition and QGP medium parameters, while varying the shape and size of hot spots inside the proton to measure the effect on bulk observables.
+
+These uncertainties systematically bias the predictions of hydrodynamic models which assert a single theoretical model
 
@@
 
- Lorem ipsum dolor sit amet, consectetur adipiscing elit.
-In tincidunt elit a libero luctus, ut pharetra sem accumsan.
-Nullam eros dui, egestas ut arcu et, commodo suscipit purus.
-Quisque dapibus tincidunt nulla at vehicula.
-Nulla ut libero sapien.
-Suspendisse ac consectetur metus.
-Interdum et malesuada fames ac ante ipsum primis in faucibus.
-Nunc facilisis volutpat suscipit.
-Pellentesque nec est orci.
-Donec id pharetra enim.
-
 \subsection{Initial conditions}
 
-Aliquam porttitor massa risus, hendrerit dignissim purus placerat vel.
-Suspendisse dui ante, porttitor pretium lorem et, placerat posuere mi.
-Donec lacinia nisl leo, ac luctus libero tincidunt non.
-Donec pretium ultricies felis vitae tempus.
-Duis pulvinar nisi justo, sit amet finibus orci volutpat non.
-Proin fermentum a magna id rhoncus.
-Vestibulum posuere, ipsum vel dignissim laoreet, arcu nibh venenatis neque, et aliquet dui dui a ipsum.
-Praesent sagittis condimentum neque, sed aliquam odio volutpat nec.
-Aenean consequat ante eget mauris lacinia, sed ultrices est dapibus.
-Morbi fermentum malesuada turpis, ac placerat est bibendum tincidunt.
-Nullam leo ipsum, sodales sit amet viverra non, iaculis vel nisl.
-Vestibulum purus ipsum, egestas ut bibendum sed, congue quis purus.
-Curabitur iaculis tortor nisl, nec venenatis tortor placerat a.
-Fusce feugiat nisl vitae porta porta.
-Praesent non arcu leo.
-Sed gravida eget neque in auctor.
+  We model the QGP initial conditions in this work using a simple parametric form for initial entropy deposition employed by the \trento\ model \note.
+The model operates in the ultrarelativistic limit with a Lorentz factor $\gamma \gg 1$ such that each nucleus appears as a thin sheet of nuclear density in the laboratory frame.
+The colliding nuclei are then described by their nuclear thickness functions
+\begin{equation}
+  T_{A,B}(\mathbf{x}) = \int dz\, \rho_{A,B}(\mathbf{x}, z),
+\end{equation}
+where $\mathbf{x}$ is a vector in the transverse plane, $z$ lies along the beam axis, and $\rho$ is the density of each nucleus in its local rest frame.
+Individual collisions between the nuclei occur at a random impact parameter offset $b$, and carve out a subset of participant matter $\rho_{A,B}^\text{part}$ which contributes to inelastic processes in the collision.
+The resulting \emph{participant} thickness functions are defined as
+\begin{equation}
+  \T_{A,B}(\mathbf{x}) = \int dz\, \rho^\text{part}_{A,B}(\mathbf{x} \pm \mathbf{b}/2, z).
+\end{equation}
+where $\rho^\text{part}_{A,B}$ is determined using a Monte Carlo Glauber model described later in this section.
+
+The sheets of colliding nuclear density penetrate and pass through each other in time $\Delta t = D_\text{nucl} / \sqrt{\gamma^2 - 1}$ in the laboratory frame, where $D_\text{nucl}$ is the diameter of the nucleus in its rest frame and $\gamma$ is the usual Lorentz factor of the accelerated ions.
+The resulting nuclear overlap time $\Delta t =$ \order{10}~fm/$c$ at top RHIC and LHC energies and hence is sufficiently short to treat all interactions local in the transverse plane.
+
+$t\ll 1$~fm/$c$ in the laboratory frame, and leave little time for spatially separated matter
+The collision deposits entropy at midrapidity, 
+\begin{equation}
+  dS/d\eta \propto f(\T_A, \T_B)
+\end{equation}
 
@@
 
-Sed ac ullamcorper ipsum.
-Sed quis eleifend mauris.
-Phasellus quis massa neque.
-Proin nec lectus quis ante aliquam tincidunt ut sit amet eros.
-Nullam et odio eu elit aliquam molestie quis ut arcu.
-Aliquam gravida enim non tortor bibendum, at semper tellus egestas.
-In libero lectus, convallis in suscipit nec, bibendum eu felis.
-Vestibulum ante ipsum primis in faucibus orci luctus et ultrices posuere cubilia Curae; Vivamus quis lectus neque.
-Sed ac placerat justo.
-Nullam ultricies vulputate ligula, vel porta mauris scelerisque et.
-Sed hendrerit blandit semper.
-Sed vel bibendum elit, eu posuere elit.
-Phasellus nunc massa, tincidunt ac tempus quis, blandit quis erat.
-
 \subsection{Hydrodynamics and Boltzmann transport}
 
- Lorem ipsum dolor sit amet, consectetur adipiscing elit.
-In tincidunt elit a libero luctus, ut pharetra sem accumsan.
-Nullam eros dui, egestas ut arcu et, commodo suscipit purus.
-Quisque dapibus tincidunt nulla at vehicula.
-Nulla ut libero sapien.
-Suspendisse ac consectetur metus.
-Interdum et malesuada fames ac ante ipsum primis in faucibus.
-Nunc facilisis volutpat suscipit.
-Pellentesque nec est orci.
-Donec id pharetra enim.
-
-To first approximation---as with nearly every object in physics---it is convenient to approximate the nucleons in heavy-ion initial condition models as a spherically symmetric particles of finite spatial extent.
-Commonly, a Gaussian is used 
-\begin{figure}
-  \begin{tikzpicture}
-    % spherical proton
-    \draw[dashed, xshift=-2cm] (0,0) circle (.8cm);
-    % three partons
-    \draw[dashed] (0,0) circle (.8cm);
-    \foreach \theta in {0, 120, 240}{
-      \draw ({\theta}:.39) circle (.4cm);
-    }
-    % ten partons
-    \draw[dashed, xshift=2cm] (0,0) circle (.8cm);
-    \foreach \theta/\radius in {
-      36/0.4, 72/0.5, 108/0.5, 140/0.3, 172/0.1,
-      190/0.5, 240/0.3, 272/0.5, 308/0.4, 360/0.5
-    }{
-      \draw[xshift=2cm] ({\theta}:\radius) circle (.2cm);
-    }
-  \end{tikzpicture}
-  \caption{\label{proton_shapes} Schematic of plausible proton shapes.
-  The sketch on the left shows a spherically symmetric proton (dashed line), while the middle and right illustrations depict a fluctuating proton with three and ten partons respectively (solid lines).
-  }
-\end{figure}
+%To first approximation---as with nearly every object in physics---it is convenient to approximate the nucleons in heavy-ion initial condition models as a spherically symmetric particles of finite spatial extent.
+%Commonly, a Gaussian is used
+%\begin{figure}
+%  \begin{tikzpicture}
+%    % spherical proton
+%    \draw[dashed, xshift=-2cm] (0,0) circle (.8cm);
+%    % three partons
+%    \draw[dashed] (0,0) circle (.8cm);
+%    \foreach \theta in {0, 120, 240}{
+%      \draw ({\theta}:.39) circle (.4cm);
+%    }
+%    % ten partons
+%    \draw[dashed, xshift=2cm] (0,0) circle (.8cm);
+%    \foreach \theta/\radius in {
+%      36/0.4, 72/0.5, 108/0.5, 140/0.3, 172/0.1,
+%      190/0.5, 240/0.3, 272/0.5, 308/0.4, 360/0.5
+%    }{
+%      \draw[xshift=2cm] ({\theta}:\radius) circle (.2cm);
+%    }
+%  \end{tikzpicture}
+%  \caption{\label{proton_shapes} Schematic of plausible proton shapes.
+%  The sketch on the left shows a spherically symmetric proton (dashed line), while the middle and right illustrations depict a fluctuating proton with three and ten partons respectively (solid lines).
+%  }
+%\end{figure}
 
@@
 
-Aliquam porttitor massa risus, hendrerit dignissim purus placerat vel.
-Suspendisse dui ante, porttitor pretium lorem et, placerat posuere mi.
-Donec lacinia nisl leo, ac luctus libero tincidunt non.
-Donec pretium ultricies felis vitae tempus.
-Duis pulvinar nisi justo, sit amet finibus orci volutpat non.
-Proin fermentum a magna id rhoncus.
-Vestibulum posuere, ipsum vel dignissim laoreet, arcu nibh venenatis neque, et aliquet dui dui a ipsum.
-Praesent sagittis condimentum neque, sed aliquam odio volutpat nec.
-Aenean consequat ante eget mauris lacinia, sed ultrices est dapibus.
-Morbi fermentum malesuada turpis, ac placerat est bibendum tincidunt.
-Nullam leo ipsum, sodales sit amet viverra non, iaculis vel nisl.
-Vestibulum purus ipsum, egestas ut bibendum sed, congue quis purus.
-Curabitur iaculis tortor nisl, nec venenatis tortor placerat a.
-Fusce feugiat nisl vitae porta porta.
-Praesent non arcu leo.
-Sed gravida eget neque in auctor.
-
-Sed ac ullamcorper ipsum.
-Sed quis eleifend mauris.
-Phasellus quis massa neque.
-Proin nec lectus quis ante aliquam tincidunt ut sit amet eros.
-Nullam et odio eu elit aliquam molestie quis ut arcu.
-Aliquam gravida enim non tortor bibendum, at semper tellus egestas.
-In libero lectus, convallis in suscipit nec, bibendum eu felis.
-Vestibulum ante ipsum primis in faucibus orci luctus et ultrices posuere cubilia Curae; Vivamus quis lectus neque.
-Sed ac placerat justo.
-Nullam ultricies vulputate ligula, vel porta mauris scelerisque et.
-Sed hendrerit blandit semper.
-Sed vel bibendum elit, eu posuere elit.
-Phasellus nunc massa, tincidunt ac tempus quis, blandit quis erat.
+%\fig{observables_design}{Caption}
 
@@
 
+%\begin{figure}
+%  \begin{tikzpicture}
+%    \foreach \x in {0pt, 10pt, 20pt}{\draw (\x, 0) ellipse (5pt and 20pt);}
+%    \foreach \x in {0pt, 10pt, 20pt, 30pt}{\draw (\x+100pt, 0) ellipse (5pt and 20pt);}
+%  \end{tikzpicture}
+%  \caption{\label{fig:collision} This is the figure caption}
+%\end{figure}
+%Consider a relativistic collision between a columnated stack of $n_1$ projectile nucleons and $n_2$ target nucleons as depicted in Fig.~\ref{fig:collision}.
+%Moreover, assume that each nucleon has mass $m$ and moves with velocity $\pm u$ along the beam axis $\hat{z}$.
+%The four-momentum of each colliding nucleon stack is given by
+%\begin{align}
+%  p_1 &= n_1 \gamma m\, (1, 0, 0, u),\\
+%  p_2 &= n_2 \gamma m\, (1, 0, 0, -u),
+%\end{align}
+%where $\gamma = 1/\sqrt{1 - u^2}$ is the usual Lorentz factor.
+%
+%At relativistic energies $\gamma \gg 1$, each stack of nucleons is significantly Lorentz contracted in the lab frame and appears compressed along its direction of motion.
+%The nucleons are thus observed to penetrate and pass through each other in the lab frame in time
+%\begin{equation}
+%  \Delta t = L / \sqrt{\gamma^2 - 1},
+%\end{equation}
+%where $L\sim1$--10~fm is the length of the columnated stack of nucleons in its local rest frame.
+%
+%Collisions at top RHIC and LHC energies have interpenetration times \mbox{$\Delta t < .1$~fm/$c$} and hence transpire before meaningful dynamics develop in the transverse plane.
+%For this reason, it is convenient to say that initial energy deposition occurs instantaneously at time $\Delta t = 0^+$, the moment the two nuclei collide.
+%The initial energy (or entropy) deposited into the transverse plane is then modeled in terms of two beam-integrated nuclear thickness functions
+%\begin{equation}
+%  T_{A,B}(x, y) = \int dz \rho(x, y, z),
+%\end{equation}
+%where $\rho$ is the density profile of the nucleus in its local rest frame.
+
 \subsection{QGP medium properties}
 
-Aliquam porttitor massa risus, hendrerit dignissim purus placerat vel.
-Suspendisse dui ante, porttitor pretium lorem et, placerat posuere mi.
-Donec lacinia nisl leo, ac luctus libero tincidunt non.
-Donec pretium ultricies felis vitae tempus.
-Duis pulvinar nisi justo, sit amet finibus orci volutpat non.
-Proin fermentum a magna id rhoncus.
-Vestibulum posuere, ipsum vel dignissim laoreet, arcu nibh venenatis neque, et aliquet dui dui a ipsum.
-Praesent sagittis condimentum neque, sed aliquam odio volutpat nec.
-Aenean consequat ante eget mauris lacinia, sed ultrices est dapibus.
-Morbi fermentum malesuada turpis, ac placerat est bibendum tincidunt.
-Nullam leo ipsum, sodales sit amet viverra non, iaculis vel nisl.
-Vestibulum purus ipsum, egestas ut bibendum sed, congue quis purus.
-Curabitur iaculis tortor nisl, nec venenatis tortor placerat a.
-Fusce feugiat nisl vitae porta porta.
-Praesent non arcu leo.
-Sed gravida eget neque in auctor.
-
-Sed ac ullamcorper ipsum.
-Sed quis eleifend mauris.
-Phasellus quis massa neque.
-Proin nec lectus quis ante aliquam tincidunt ut sit amet eros.
-Nullam et odio eu elit aliquam molestie quis ut arcu.
-Aliquam gravida enim non tortor bibendum, at semper tellus egestas.
-In libero lectus, convallis in suscipit nec, bibendum eu felis.
-Vestibulum ante ipsum primis in faucibus orci luctus et ultrices posuere cubilia Curae; Vivamus quis lectus neque.
-Sed ac placerat justo.
-Nullam ultricies vulputate ligula, vel porta mauris scelerisque et.
-Sed hendrerit blandit semper.
-Sed vel bibendum elit, eu posuere elit.
-Phasellus nunc massa, tincidunt ac tempus quis, blandit quis erat.
-
 \section{Summary and conclusions}
-
-Aliquam porttitor massa risus, hendrerit dignissim purus placerat vel.
-Suspendisse dui ante, porttitor pretium lorem et, placerat posuere mi.
-Donec lacinia nisl leo, ac luctus libero tincidunt non.
-Donec pretium ultricies felis vitae tempus.
-Duis pulvinar nisi justo, sit amet finibus orci volutpat non.
-Proin fermentum a magna id rhoncus.
-Vestibulum posuere, ipsum vel dignissim laoreet, arcu nibh venenatis neque, et aliquet dui dui a ipsum.
-Praesent sagittis condimentum neque, sed aliquam odio volutpat nec.
-Aenean consequat ante eget mauris lacinia, sed ultrices est dapibus.
-Morbi fermentum malesuada turpis, ac placerat est bibendum tincidunt.
-Nullam leo ipsum, sodales sit amet viverra non, iaculis vel nisl.
-Vestibulum purus ipsum, egestas ut bibendum sed, congue quis purus.
-Curabitur iaculis tortor nisl, nec venenatis tortor placerat a.
-Fusce feugiat nisl vitae porta porta.
-Praesent non arcu leo.
-Sed gravida eget neque in auctor.
-
-Sed ac ullamcorper ipsum.
-Sed quis eleifend mauris.
-Phasellus quis massa neque.
-Proin nec lectus quis ante aliquam tincidunt ut sit amet eros.
-Nullam et odio eu elit aliquam molestie quis ut arcu.
-Aliquam gravida enim non tortor bibendum, at semper tellus egestas.
-In libero lectus, convallis in suscipit nec, bibendum eu felis.
-Vestibulum ante ipsum primis in faucibus orci luctus et ultrices posuere cubilia Curae; Vivamus quis lectus neque.
-Sed ac placerat justo.
-Nullam ultricies vulputate ligula, vel porta mauris scelerisque et.
-Sed hendrerit blandit semper.
-Sed vel bibendum elit, eu posuere elit.
-Phasellus nunc massa, tincidunt ac tempus quis, blandit quis erat.
 
@@
 
+Citations I don't want to forget \cite{Shen:2016mmv}
 \bibliography{trento-partons}

### 2018-04-13 talk/qm2018.tex
Add some content to paper on trento
Add some more content to the talk

@@
   \begin{tikzpicture}
-    \node (a) {\includegraphics{hannah_collision}};
+    \node (a) {\includegraphics[width=.95\textwidth]{hannah_collision}};
     \node[align=left, anchor=north west] at (a.north west) {%
@@
   \begin{columns}[b]
-    \begin{column}{.4\textwidth}
-      \begin{tikzpicture}
-        \node (A) {\includegraphics<1>[height=.82\textheight]{observables_design}};
-        \node (B) {\includegraphics<2>[height=.82\textheight]{observables_posterior}};
-        \node [xshift=2.8cm, yshift=3.2cm, rotate=-45] {%
-          \textcolor{theme}{\scshape New!}%
-          };
-      \end{tikzpicture}
+    \begin{column}{.43\textwidth}
+      \includegraphics<1>[width=\textwidth]{lead_observables_design}
+      \includegraphics<2>[width=\textwidth]{lead_observables_posterior}
     \end{column}
-    \begin{column}{.02\textwidth}
-    \end{column}
-    \begin{column}{.55\textwidth}
+    \begin{column}{.57\textwidth}
       \begin{itemize}
-        \item Combined analysis of 2.76 and 5.02 TeV beam energies
-        \item Parametric initial conditions and medium parameters
+        \item Combined analysis of 2.76 and 5.02
+          TeV beam energies \textcolor{theme}{\scshape (New!)} 
+        \item Parametric initial conditions and medium parameters\
+        \item Quantitative estimates on $\eta/s$ and $\zeta/s$ with meaningful uncertainties
         \item See thesis by Jonah Bernhard {\small [\url{99999}]}
       \end{itemize}
-      \begin{tikzpicture}
-        \node (0, 0) {\includegraphics<1>{region_shear_blank}};
-        \node (0, 0) {\includegraphics<2>{region_shear}};
-        \node [xshift=.3cm, yshift=.6cm] {\only<1>{\fontsize{80}{0}\selectfont \color{theme!30} ?}};
-        \node [xshift=3cm, yshift=1.8cm, rotate=-45] {%
-          \textcolor{theme}{\scshape New!}%
-          };
-      \end{tikzpicture}
+      \bigskip
+      \includegraphics<1>[width=\textwidth]{region_shear_bulk}
+      \includegraphics<2>[width=\textwidth]{region_shear_bulk}
     \end{column}
@@
       \begin{figure}
-        \includegraphics[width=.9\columnwidth]{lead_208}
-        \caption{Pb$^{208}$ nucleus}
+        \includegraphics[width=.9\columnwidth]{pb208}\\
+        Pb$^{208}$ nucleus
       \end{figure}
@@
       \begin{figure}
-        \includegraphics{trento_model_part}\\
+        \includegraphics[width=\textwidth]{trento_model_part}\\
         \small MC-Glauber cross sections\\
@@
       \begin{figure}
-        \includegraphics{trento_model_thickness}
+        \includegraphics[width=\textwidth]{trento_model_thickness}
       \end{figure}
@@
       \begin{figure}
-        \includegraphics{trento_model_entropy}
+        \includegraphics[width=\textwidth]{trento_model_entropy}
       \end{figure}
@@
       \begin{equation*}
-        \small
         T_R =
@@
       \begin{figure}
-        \includegraphics{trento_model_freestream}\\
+        \includegraphics[width=\textwidth]{trento_model_freestream}\\
         \small arrows: fluid velocity\\weighted by energy density
@@
   \begin{columns}[T]
-    \begin{column}{.55\textwidth}
+    \begin{column}{.6\textwidth}
       \scriptsize
@@
     \end{column}
-    \begin{column}{.45\textwidth}
+    \begin{column}{.4\textwidth}
       \textcolor{theme}{\scshape Important details:}
@@
         \small
-        \item Parton width is reparametrized:\\[1ex]
+        \item Parton width reparametrized:\\[1ex]
           $v = v_\text{min} + \X(v_\text{max} - v_\text{min})$\\[1ex]
@@
         \item Parton sampling radius is \emph{not} equivalent to the proton radius
-          in the proton c.o.m.\ frame, e.g.\ see\\
+          in the proton c.o.m.\ frame, see\\
           \textcolor{theme}{\scriptsize PRC.94.024919}\\[1ex]
-        \item Other parameters same as Pb+Pb analysis at 2.76 and 5.02 TeV
+        \item Other parameters same as combined Pb+Pb analysis at 2.76 and 5.02 TeV
       \end{itemize}
@@
     \node[anchor=east] at (current page.east) {
-      \includegraphics[height=.95\paperheight]{observables_design}
+      \includegraphics[height=.95\paperheight]{observables_posterior}
       };
@@
   \begin{tikzpicture}
-    \node {\includegraphics[height=\textheight]{posterior}};
-    \node (title) at (6.5, 3.5) {\color{theme}{\scshape \Large Bayesian posterior}};
-    \node<1-2>[draw=none, text width=7cm, align=left, below=.2cm of title] (b1) {%
-      One row (column) for each parameter
+    \node {\includegraphics[height=.98\paperheight]{posterior}};
+    \node (title) at (6.0, 3.5) {\color{theme}{\scshape \Large Bayesian posterior}};
+    \node<1-2>[draw=none, text width=6.5cm, align=left, below=.2cm of title] (b1) {%
+      \small One row (column) for each parameter
       };
-    \node<1-2>[draw=none, text width=7cm, align=left, below=.2cm of b1] (b2) {%
-      Diagonal panels: marginal distribution of a single model parameter
+    \node<1-2>[draw=none, text width=6.5cm, align=left, below=.1cm of b1] (b2) {%
+      \small Diagonal panels: marginal distribution of a single model parameter
       };
-    \node<1-2>[draw=none, text width=7cm, align=left, below=.2cm of b2] (b3) {%
-      Off-diagonal panels: joint distribution between a pair of model parameters
+    \node<1-2>[draw=none, text width=6.5cm, align=left, below=.1cm of b2] (b3) {%
+      \small Off-diagonal panels: joint distribution between a pair of model parameters
       };
-    \node<2>[draw=none, text width=7cm, align=left, below=.4cm of b3] (b4) {%
-      Let's zoom in on some specific regions...
+    \node<2>[draw=none, text width=6.5cm, align=left, below=.2cm of b3] (b4) {%
+      \textcolor{theme}{Let's look at some specific regions...}
       };
-    \draw<1-2>[->, thick, theme] (b2.west) to [out=180, in=90] (-.3, 1.2);
-    \draw<1-2>[->, thick, theme] (b3.west) to [out=180, in=90] (.2, -.5);
-    \node<3>[xshift=.5cm, yshift=-.5cm, scale=.3] {\includegraphics{magnifying-glass}};
-    \node<4>[xshift=-.5cm, yshift=-1cm, scale=.3] {\includegraphics{magnifying-glass}};
+    \draw<1-2>[->, thick, theme] (b2.west) to [out=180, in=90] (-.58, 1.3);
+    \draw<1-2>[->, thick, theme] (b3.west) to [out=180, in=90] (-.05, 0);
+    \node<3>[xshift=-2.8cm, yshift=3.2cm, scale=.3] {\includegraphics[scale=.15]{magnifying-glass}};
+    \node<4>[xshift=-1.2cm, yshift=1.6cm, scale=.3] {\includegraphics[scale=.15]{magnifying-glass}};
+    \node<5>[xshift=-.7cm, yshift=1.1cm, scale=.3] {\includegraphics[scale=.15]{magnifying-glass}};
   \end{tikzpicture}
@@
 \begin{frame}{\scshape Maximum a posteriori parameters}
-
+  \includegraphics[height=.8\textheight]{find_map}
 \end{frame}

### 2018-04-13 src/plots.py

@@ -17,7 +17,7 @@ In the code, each plot is generated by a function tagged with the ``@plot``
 decorator.
 """
 
-from collections import OrderedDict
+from collections import OrderedDict, Counter
 import itertools
 import logging
 import os
@@ -648,7 +648,13 @@ def find_map():
     fixed_params = {
         'trento_p': 0.,
         'etas_min': .08,
+        'etas_slope': 1.1285,
+        'etas_crv': -0.5365,
         'etas_hrg': .3,
+        'zetas_max': 0.0536,
+        'zetas_width': 0.0194,
+        'zetas_t0': 0.1824,
+        'Tswitch': 0.1515,
         'model_sys_err': .1,
     }
 
@@ -2035,7 +2041,7 @@ def proton_radius():
     set_tight()
 
 
-#@plot
+@plot
 def grid_error():
     """
     Scatter plot observables calculated on a grid with grid scale = 0.2 against
@@ -2064,7 +2070,7 @@ def grid_error():
     )
 
     fine_events, coarse_events = [
-        model.ModelData(system, *design_points).events
+        [ev for p, ev in model.ModelData(system, *design_points).design_events]
         for design_points in (fine_design_points, coarse_design_points)
     ]
 
@@ -2326,7 +2332,7 @@ def trento_model_part():
 @plot
 def trento_model_thickness():
     """
-    Participant thickness functions 
+    Participant thickness functions
 
     """
     fig = plt.figure(figsize=figsize(.5, .75))
@@ -2432,7 +2438,7 @@ def trento_model_freestream():
     TA, TB = thickness_functions(xx, yy)
     TR = np.sqrt(TA*TB)
 
-    fs = freestream.FreeStreamer(TR, xmax, 1) 
+    fs = freestream.FreeStreamer(TR, xmax, 1)
     e = fs.energy_density()
     u0, ux, uy = [fs.flow_velocity(i) for i in range(3)]
     vx, vy = [u/u0 for u in (ux, uy)]
@@ -2495,7 +2501,7 @@ def collision_profile():
 
     # Gaussian protons
     for w in [.4, .7, .9]:
-        norm = 4*np.pi*w**2 
+        norm = 4*np.pi*w**2
 
         def fit_xsection(x, A=6):
             c = A**2/4
@@ -2515,7 +2521,7 @@ def collision_profile():
         bbox_to_anchor=(.5, .15)
     )
 
-    set_tight() 
+    set_tight()
 
 
 @plot
[…]

### 2018-04-30 src/correct.py
Converge on hyperparameters

@@ -18,6 +18,9 @@ from . import cachedir
 from .design import Design
 
 
+import matplotlib.pyplot as plt
+
+
 def run_cmd(*args):
     """
     Run and log a subprocess.

### 2018-04-30 src/design.py

@@ -36,12 +36,13 @@ design and produce far too few particles to be physically reasonable.
 
 """
 bad_points = [
-    384, 129, 392, 273, 276, 277, 412, 156, 285, 159, 162, 418, 421, 422,
-    296, 298, 428, 429, 176, 178, 183, 439, 184, 185, 312, 317, 446, 319,
-    320, 193, 451, 326, 199, 200, 327, 330, 459, 460, 461, 462, 335, 334,
-    464, 466, 83, 211, 341, 338, 471, 472, 346, 219, 476, 349, 223, 480,
-    225, 482, 248, 355, 483, 358, 360, 104, 234, 363, 492, 365, 107, 236,
-    367, 498, 115, 242, 247, 120, 249, 378, 124, 381
+    54, 112, 129, 162, 170, 176, 183, 199, 200, 224, 234, 245, 247, 249, 257,
+    273, 288, 315, 319, 321, 330, 335, 344, 345, 346, 360, 363, 365, 379, 386,
+    418, 422, 445, 451, 460, 461, 466, 480, 482, 492, 498, 120, 176, 200, 273,
+    277, 327, 335, 346, 412, 439, 462, 466, 476, 482, 492, 178, 184, 225, 341,
+    392, 472, 83, 104, 107, 115, 156, 159, 185, 193, 211, 219, 223, 236, 242,
+    248, 276, 285, 296, 298, 312, 317, 326, 334, 338, 355, 358, 367, 378, 392,
+    421, 428, 429, 446, 459, 464, 471, 472, 483
 ]
 
 def generate_lhs(npoints, ndim, seed):

### 2018-04-30 src/emulator.py

@@ -66,19 +66,17 @@ class Emulator:
     """
     # observables to emulate
     # list of 2-tuples: (obs, [list of subobs])
-
     PbPb5020 = [
         ('dNch_deta', [None]),
         ('vnk', [(2, 2), (3, 2), (4, 2)]),
     ]
-
     pPb5020 = [
         ('dNch_deta', [None]),
         ('mean_pT', [None]),
         ('vnk', [(2, 2), (3, 2)]),
     ]
 
-    def __init__(self, system, npc=10, nrestarts=0, exclude_points=[]):
+    def __init__(self, system, npc=7, nrestarts=0, exclude_points=[]):
         logging.info(
             'training emulator for system %s (%d PC, %d restarts)',
             system, npc, nrestarts
@@ -88,7 +86,7 @@ class Emulator:
         self._slices = {}
 
         # System specific observables
-        observables = {
+        self.observables = {
             'pPb5020': self.pPb5020,
             'PbPb5020': self.PbPb5020,
         }[system]
@@ -101,7 +99,7 @@ class Emulator:
 
         # Build an array of all observables to emulate.
         nobs = 0
-        for obs, subobslist in observables:
+        for obs, subobslist in self.observables:
             self._slices[obs] = {}
             for subobs in subobslist:
                 Y.append(model.data[system][obs][subobs]['Y'])
@@ -229,7 +227,7 @@ class Emulator:
             } for obs, slices in self._slices.items()
         }
 
-    def predict(self, X, return_cov=False, extra_std=0):
+    def predict(self, X, return_cov=False):
         """
         Predict model output at `X`.
         X must be a 2D array-like with shape ``(nsamples, ndim)``.  It is passed
@@ -251,9 +249,6 @@ class Emulator:
         ``(nsamples, n_cent_bins_1, n_cent_bins_2)``.
         NB: the covariance is only computed between observables and centrality
         bins, not between sample points.
-        `extra_std` is additional uncertainty which is added to each GP's
-        predictive uncertainty, e.g. to account for model systematic error.  It
-        may either be a scalar or an array-like of length nsamples.
         """
         gp_mean = [gp.predict(X, return_cov=return_cov) for gp in self.gps]
 
@@ -270,10 +265,6 @@ class Emulator:
             gp_var = np.concatenate([
                 c.diagonal()[:, np.newaxis] for c in gp_cov
             ], axis=1)
-
-            # Add extra uncertainty to predictive variance.
[…]

### 2018-04-30 src/expt.py

@@ -323,7 +323,6 @@ def log_transform(data, system):
         for subobs, subobs_data in obs_data.items():
             y, yerr = [subobs_data[k] for k in ('y', 'yerr')]
             ystat, ysys = [yerr[k] for k in ('stat', 'sys')]
-
             np.divide(ystat, y, out=ystat)
             np.divide(ysys, y, out=ysys)
             np.log(y, out=y)
@@ -452,7 +451,10 @@ def cov(
         else:
             return np.zeros((x1.size, x2.size))
 
-    # compute the sys error covariance
+    # Use special sys error covariance matrix for p-Pb system
+    if system == 'pPb5020':
+        sys_corr_length = (5 if 'dNch_deta' not in [obs1, obs2] else 30)
+
     C = (
         np.exp(-.5*(np.subtract.outer(x1, x2)/sys_corr_length)**2) *
         np.outer(sys1, sys2)

### 2018-04-30 src/mcmc.py

@@ -125,7 +125,6 @@ class Chain:
         ('dNch_deta', [None]),
         ('vnk', [(2, 2), (3, 2), (4, 2)]),
     ]
-
     pPb5020 = [
         ('dNch_deta', [None]),
         ('mean_pT', [None]),
@@ -141,12 +140,7 @@ class Chain:
         attr = [(d.keys, d.labels, d.range) for d in designs]
         assert all(klr == next(iter(attr)) for klr in attr)
 
-        # append model systematic error
-        d = Design(next(iter(systems)))
-        d.keys.append('model_sys_err')
-        d.labels.append(r'$\sigma\ \mathrm{model\ sys}$')
-        d.range.append((0., .4))
-
+        d = designs[0]
         self.keys = d.keys
         self.labels = d.labels
         self.range = d.range
@@ -205,18 +199,13 @@ class Chain:
 
         """
         return {
-            sys: emulators[sys].predict(X[:, :-1], **kwargs)
+            sys: emulators[sys].predict(X, **kwargs)
             for n, sys in enumerate(systems)
         }
 
-    def log_posterior(self, X, extra_std_prior_scale=.05):
+    def log_posterior(self, X):
         """
         Evaluate the posterior at `X`.
-
-        `extra_std_prior_scale` is the scale parameter for the prior
-        distribution on the model sys error parameter:
-
-            prior ~ sigma^2 * exp(-sigma/scale)
 
         """
         X = np.array(X, copy=False, ndmin=2)
@@ -229,10 +218,7 @@ class Chain:
         nsamples = np.count_nonzero(inside)
 
         if nsamples > 0:
-            extra_std = X[inside, -1]
-            pred = self._predict(
-                X[inside], return_cov=True, extra_std=extra_std
-            )
+            pred = self._predict(X[inside], return_cov=True)
 
             for sys in systems:
                 nobs = self._expt_y[sys].size
@@ -257,9 +243,6 @@ class Chain:
 
                 # compute log likelihood at each point
                 lp[inside] += list(map(mvn_loglike, dY, cov))
-
-            # add prior for extra_std (model sys error)
-            lp[inside] += 2*np.log(extra_std) - extra_std/extra_std_prior_scale
 
         return lp
 

### 2018-04-30 src/model.py

@@ -258,9 +258,14 @@ class ModelData:
 
                 """
                 minbias_events = events[(trigger == minbias).all(axis=1)]
-                binned_events = correct_centrality(
-                    self.system, design_point, minbias_events, bins
-                )
+                #binned_events = correct_centrality(
+                #    self.system, design_point, minbias_events, bins
+                #)
+                n = minbias_events.size
+                binned_events = [
+                    minbias_events[int((1 - b/100)*n):int((1 - a/100)*n)]
+                    for a, b in bins
+                ]
             elif bin_type == 'mult':
                 """
                 Many p-Pb observables are plotted as a function of <Nch>. These
@@ -345,26 +350,6 @@ def _data(system, dataset='main'):
         data = dict(mean_pT=dict(charged=mean_pT), **data)
 
     data = ModelData(system, *files).observables_like(data)
-
-    #bad_points = set()
-    #for n, y in enumerate(data['dNch_deta'][None]['Y']):
-    #    if np.min(y) < 1:
-    #        bad_points.add(n)
-
-    #for n, y in enumerate(data['mean_pT'][None]['Y']):
-    #    if np.max(y) > 1.5:
-    #        bad_points.add(n)
-
-    #for n, y in enumerate(data['vnk'][(2, 2)]['Y']):
-    #    if np.min(y) < 0.0024787:
-    #        bad_points.add(n)
-
-    #for n, y in enumerate(data['vnk'][(3, 2)]['Y']):
-    #    if np.min(y) < 0.0024787:
-    #        bad_points.add(n)
-
-    #print(list(bad_points))
-    #quit()
 
     logging.info('writing cache file %s', cachefile)
     cachefile.parent.mkdir(parents=True, exist_ok=True)

### 2018-04-30 src/plots.py

@@ -43,6 +43,7 @@ from scipy.optimize import brentq, curve_fit
 from sklearn.decomposition import PCA
 from sklearn.gaussian_process import GaussianProcessRegressor as GPR
 from sklearn.gaussian_process import kernels
+from sklearn.mixture import GaussianMixture
 from sklearn.model_selection import KFold
 
 from . import cachedir, workdir, systems, parse_system, expt, model, mcmc
@@ -310,7 +311,7 @@ def obs_label(obs, subobs, differentials=False, full_cumulants=False):
     }
     if obs in id_parts_labels:
         return id_parts_labels[obs].format(
-            {'pion': '\pi', 'kaon': 'K', 'proton': 'p'}[subobs]
+                {'pion': '\pi', 'kaon': 'K', 'proton': 'p', None: '{}'}[subobs]
         )
 
     if obs == 'pT_fluct':
@@ -347,7 +348,7 @@ def _observables_plots():
         dict(
             title='Mean $p_T$',
             ylabel=r'$\langle p_T \rangle$ [GeV]',
-            ylim=(-2, 1),
+            ylim=(-1.25, 0.7),
             subplots=[
                 ('mean_pT', None, dict(label=r'$\mathrm{mean}\ p_T$', scale=1)),
             ]
@@ -437,6 +438,16 @@ def _observables(posterior=False):
             else:
                 ax.set_xlim(0, 55)
                 ax.set_xlabel('Centrality %')
+
+            if obs == 'dNch_deta':
+                ax.plot([0, 50], [6, 5])
+                ax.plot([0, 50], [2, 1])
+            if obs == 'mean_pT':
+                ax.axhline(0.5)
+                ax.axhline(-1.0)
+            if obs == 'vnk':
+                ax.axhline(-6)
+                ax.axhline(0)
 
         if plot.get('yscale') == 'log':
             ax.set_yscale('log')
@@ -635,123 +646,6 @@ def observables_map():
 
 
 @plot
-def observables_expt_only():
-    """
-    Observables plots of experimental data only.
-
-    """
-    plots = _observables_plots()
-
-    ylim = {
-        'Yields': (2, 1e5),
-        'Flow cumulants': (0, .15),
-        'Mean $p_T$': (0, 1.7),
-        'Mean $p_T$ fluctuations': (0, .045),
-    }
-
-    for n, p in enumerate(plots):
-        p['ylim'] = ylim[p['title']]
-        if p['title'] == 'Flow cumulants':
[…]

### 2018-05-01 src/plots.py
move some of the talk figures to a separate script

@@ -17,13 +17,11 @@ In the code, each plot is generated by a function tagged with the ``@plot``
 decorator.
 """
 
-from collections import OrderedDict, Counter, Mapping
+from collections import OrderedDict, Counter
 import itertools
 import logging
-import os
 from pathlib import Path
 import pickle
-import subprocess
 import tempfile
 import warnings
 
@@ -36,21 +34,17 @@ from matplotlib import patches
 from matplotlib import ticker
 from matplotlib import cm
 from matplotlib.colors import ListedColormap
-
 from scipy import special
-from scipy.interpolate import interp1d, PchipInterpolator
-from scipy.optimize import brentq, curve_fit
+from scipy.interpolate import PchipInterpolator
+from scipy.optimize import curve_fit
 from sklearn.decomposition import PCA
 from sklearn.gaussian_process import GaussianProcessRegressor as GPR
 from sklearn.gaussian_process import kernels
-from sklearn.mixture import GaussianMixture
 from sklearn.model_selection import KFold
 
 from . import cachedir, workdir, systems, parse_system, expt, model, mcmc
 from .design import Design
 from .emulator import Emulator, emulators
-
-import freestream
 
 
 # golden ratio
@@ -80,7 +74,6 @@ colors = OrderedDict([
 ])
 
 offblack = '.15'
-theme = '#b01419'
 
 plt.rcdefaults()
 plt.rcParams.update({
@@ -158,34 +151,6 @@ def plot(f):
     return wrapper
 
 
-def cmap_to_alpha(cmap=plt.cm.inferno, fraction=.2):
-    """
-    Fade the bottom of a colormap to white.
-
-    Currently only works with ListedColormap objects, of which the new cmaps
-    (inferno, magma, plasma, viridis) are examples.
-
-    """
-    try:
-        colors = cmap.colors.copy()
-        n = int(fraction*len(colors))
-
-        for i, rgb in enumerate(colors[:n]):
-            colors[i] = rgb + [i/n]
-
-        return type(cmap)(colors, cmap.name + '_mod')
-    except AttributeError:
-        cmin = 25 if cmap == plt.cm.Oranges_r else 0
-        colors = [list(cmap(n/256)) for n in range(cmin, 200)]
-        n = int(fraction*len(colors))
-
-        for i, rgb in enumerate(colors[:n]):
[…]

### 2018-05-02 talk/qm2018.tex
move some of the talk figures to a separate script
Adapt plots for p-Pb and Pb-Pb

@@
       {\scriptsize non-RTA shear \& bulk corrections}\\
-      {\tiny\color{theme} J.E.\ Bernhard thesis}
+      {\tiny \color{theme} (Bernhard thesis) arXiv:1804.06469}
     \end{column}
@@
         \item Quantitative estimates on $\eta/s$ and $\zeta/s$ with meaningful uncertainties
-        \item See thesis by Jonah Bernhard {\small [\url{99999}]}
+        \item See thesis by Jonah Bernhard {\small [\textcolor{theme}{\url{1804.06469}}]}
       \end{itemize}
@@
       \begin{tikzpicture}[absolute, overlay]
-        \node (a) at (2, -.75) {\includegraphics[scale=.5]{proton}};
-        \node[right=of a.east] (b) {\includegraphics[scale=.5]{proton_partons}};
+        \node (a) at (2, -.75) {\includegraphics[scale=.5]{gaussian_proton}};
+        \node[right=of a.east] (b) {\includegraphics[scale=.5]{deformed_proton}};
         \draw[->, thick] (a) -- (b);
@@
       \begin{figure}
-        \includegraphics[width=\textwidth]{trento_model_part}\\
+        \includegraphics<1>[width=\textwidth]{trento_nucleons}
+        \includegraphics<2>[width=\textwidth]{trento_participants}\\
         \small MC-Glauber cross sections\\
@@
       \begin{figure}
-        \includegraphics[width=\textwidth]{trento_model_thickness}
+        \includegraphics[width=\textwidth]{trento_thickness}
       \end{figure}
@@
       \begin{figure}
-        \includegraphics[width=\textwidth]{trento_model_entropy}
+        \includegraphics[width=\textwidth]{trento_entropy}
       \end{figure}
@@
       \end{equation*}
-      \smallskip
       \begin{center}
-        \small $T$ denotes \emph{participant} thickness function
+        \smallskip \small $T$ denotes \emph{participant} thickness function
       \end{center}
@@
       \begin{figure}
-        \includegraphics[width=\textwidth]{trento_model_freestream}\\
+        \includegraphics[width=\textwidth]{trento_freestream}\\
         \small arrows: fluid velocity\\weighted by energy density
@@
 
-\begin{frame}{\scshape Numerical grid error}
+\begin{frame}{\scshape Posterior nucleon shape}
+  \begin{columns}
+    \begin{column}{.45\textwidth}
       \begin{figure}
-    \includegraphics[width=\textwidth]{grid_error}
-    \caption{Numerical grid error tame for $dxy=.2 \times \text{parton width}$}
+        \includegraphics[width=\textwidth]{proton_posterior_shape}
     \end{figure}
+    \end{column}
+    \begin{column}{.55\textwidth}
+      \begin{itemize}
+        \item Parton sampling radius (approx nucleon width) is not well constrained\\[2ex]
+        \item Analysis prefers small partons,\\
+          $w_\text{parton} = 0.38_{-0.18}^{+0.21}$~fm\\[2ex]
+        \item[\small\faicon{exclamation-triangle}] Results are likely sensitive to assumed
+          details of the model, e.g.\ equation of state, pre-equilibirum dynamics and viscous
+          corrections to freezeout
+      \end{itemize}
+    \end{column}
+  \end{columns}
 \end{frame}

### 2018-05-02 src/__init__.py

@@ -24,7 +24,7 @@ cachedir.mkdir(parents=True, exist_ok=True)
 #: such as ``'PbPb2760'``, ``'AuAu200'``, ``'pPb5020'``.
 #: Even if the project uses only a single system,
 #: this should still be a list of one system string.
-systems = ['pPb5020',]
+systems = ['pPb5020', 'PbPb5020']
 
 
 def parse_system(system):

### 2018-05-02 src/design.py

@@ -35,15 +35,15 @@ Remove outlier design points. These points are near the edge of the
 design and produce far too few particles to be physically reasonable.
 
 """
-bad_points = [
-    54, 112, 129, 162, 170, 176, 183, 199, 200, 224, 234, 245, 247, 249, 257,
-    273, 288, 315, 319, 321, 330, 335, 344, 345, 346, 360, 363, 365, 379, 386,
-    418, 422, 445, 451, 460, 461, 466, 480, 482, 492, 498, 120, 176, 200, 273,
-    277, 327, 335, 346, 412, 439, 462, 466, 476, 482, 492, 178, 184, 225, 341,
-    392, 472, 83, 104, 107, 115, 156, 159, 185, 193, 211, 219, 223, 236, 242,
-    248, 276, 285, 296, 298, 312, 317, 326, 334, 338, 355, 358, 367, 378, 392,
-    421, 428, 429, 446, 459, 464, 471, 472, 483
-]
+bad_points = [61, 62, 79, 83, 104, 107, 112, 115, 123, 126, 129, 132, 141, 143,
+              153, 156, 159, 162, 170, 176, 178, 183, 184, 185, 193, 199, 203,
+              211, 215, 219, 223, 225, 226, 227, 234, 235, 236, 242, 245, 247,
+              248, 249, 266, 267, 273, 276, 285, 291, 294, 296, 297, 298, 301,
+              308, 312, 315, 317, 319, 321, 323, 326, 330, 334, 335, 338, 339,
+              341, 344, 345, 346, 355, 357, 358, 360, 363, 365, 367, 372, 376,
+              378, 379, 382, 386, 392, 395, 405, 414, 418, 421, 422, 428, 429,
+              442, 445, 446, 451, 452, 453, 456, 457, 459, 460, 461, 464, 466,
+              467, 471, 472, 473, 480, 482, 483, 492, 498]
 
 def generate_lhs(npoints, ndim, seed):
     """

### 2018-05-02 src/emulator.py

@@ -68,6 +68,7 @@ class Emulator:
     # list of 2-tuples: (obs, [list of subobs])
     PbPb5020 = [
         ('dNch_deta', [None]),
+        ('mean_pT', [None]),
         ('vnk', [(2, 2), (3, 2), (4, 2)]),
     ]
     pPb5020 = [

### 2018-05-02 src/expt.py

@@ -383,8 +383,9 @@ def _data():
                     maxcent=100
                 )
 
-    """Log transform p-Pb observables"""
+    # Log transform observables
     log_transform(data, 'pPb5020')
+    log_transform(data, 'PbPb5020')
 
     return data
 

### 2018-05-02 src/model.py

@@ -258,14 +258,14 @@ class ModelData:
 
                 """
                 minbias_events = events[(trigger == minbias).all(axis=1)]
-                #binned_events = correct_centrality(
-                #    self.system, design_point, minbias_events, bins
-                #)
                 n = minbias_events.size
                 binned_events = [
                     minbias_events[int((1 - b/100)*n):int((1 - a/100)*n)]
                     for a, b in bins
                 ]
+                #binned_events = correct_centrality(
+                #    self.system, design_point, minbias_events, bins
+                #)
             elif bin_type == 'mult':
                 """
                 Many p-Pb observables are plotted as a function of <Nch>. These
@@ -287,8 +287,9 @@ class ModelData:
             for design_point, events in self.design_events
         ]).squeeze()
 
-        """Log-transform p-Pb events"""
-        return {'x': x, bin_type: bins, 'Y': np.log(Y) if self.pPb_event else Y}
+        # Log-transform events
+        assert np.all(Y > 0)
+        return {'x': x, bin_type: bins, 'Y': np.log(Y)}
 
 
 def _data(system, dataset='main'):
@@ -347,7 +348,7 @@ def _data(system, dataset='main'):
             cent=cent,
             x=np.array([(a + b)/2 for a, b in cent]),
         )
-        data = dict(mean_pT=dict(charged=mean_pT), **data)
+        data = dict({'mean_pT': {None: mean_pT}}, **data)
 
     data = ModelData(system, *files).observables_like(data)
 

### 2018-05-02 src/plots.py

@@ -33,7 +33,7 @@ from matplotlib import lines
 from matplotlib import patches
 from matplotlib import ticker
 from matplotlib import cm
-from matplotlib.colors import ListedColormap
+from matplotlib.colors import ListedColormap, LinearSegmentedColormap
 from scipy import special
 from scipy.interpolate import PchipInterpolator
 from scipy.optimize import curve_fit
@@ -332,10 +332,9 @@ def _observables_plots():
         dict(
             title='Yields',
             ylabel=(r'$dN_\mathrm{ch}/d\eta$'),
-            ylim=(-1, 6),
-            height_ratio=1,
+            ylim=(0, 9),
             subplots=[
-                ('dNch_deta', None, dict(label=r'$N_\mathrm{ch}$', scale=1)),
+                ('dNch_deta', None, dict(label=r'$N_\mathrm{ch}$')),
             ]
         ),
         dict(
@@ -343,7 +342,7 @@ def _observables_plots():
             ylabel=r'$\langle p_T \rangle$ [GeV]',
             ylim=(-1.25, 0.7),
             subplots=[
-                ('mean_pT', None, dict(label=r'$\mathrm{mean}\ p_T$', scale=1)),
+                ('mean_pT', None, dict(label=r'$\mathrm{mean}\ p_T$')),
             ]
         ),
         dict(
@@ -352,7 +351,7 @@ def _observables_plots():
             ylim=(-7, 0),
             subplots=[
                 ('vnk', (n, 2), dict(label='$v_{}$'.format(n)))
-                for n in [2, 3]
+                for n in [2, 3, 4]
             ]
         )
     ]
@@ -419,28 +418,33 @@ def _observables(posterior=False):
                     capsize=0, color='.25', zorder=1000
                 )
 
-            if system == 'pPb5020' and obs == 'mean_pT':
-                ax.set_xlim(1, 6)
-                ax.set_xlabel(r'$N_\mathrm{ch}\,/\langle N_\mathrm{ch} \rangle$')
-            elif system == 'pPb5020' and obs == 'vnk':
-                ax.set_xlim(1, 6)
-                ax.set_xlabel('/'.join([
-                    r'$N^\mathrm{offline}_\mathrm{trk}$',
-                    r'$\langle N^\mathrm{offline}_\mathrm{trk} \rangle$',
-                ]))
-            else:
-                ax.set_xlim(0, 55)
-                ax.set_xlabel('Centrality %')
+            if system == 'pPb5020':
[…]

### 2018-05-04 src/__init__.py
move log transform to init

@@ -1,10 +1,13 @@
 """ Project initialization and common objects. """
 
+import copy
 import logging
 import os
 from pathlib import Path
 import re
 import sys
+
+import numpy as np
 
 
 logging.basicConfig(
@@ -49,3 +52,27 @@ class lazydict(dict):
     def __missing__(self, key):
         self[key] = value = self.function(key, *self.args, **self.kwargs)
         return value
+
+
+def transform(data):
+    """
+    Log transform model or experimental data,
+
+    y = log y,
+    d(log y) = dy/y.
+
+    This transformation improves PCA decomposition and emulator performance.
+
+    """
+    if 'x' in data:
+        d = {}
+        for k, v in data.items():
+            if k in ['y', 'Y']:
+                d[k] = np.log(v)
+            elif k == 'yerr':
+                d[k] = {s: data['yerr'][s]/data['y'] for s in ['stat', 'sys']}
+            else:
+                d[k] = v
+        return d
+    else:
+        return {k: transform(v) for k, v in data.items()}

### 2018-05-04 src/emulator.py

@@ -23,7 +23,7 @@ from sklearn.gaussian_process import GaussianProcessRegressor as GPR
 from sklearn.gaussian_process import kernels
 from sklearn.preprocessing import StandardScaler
 
-from . import cachedir, lazydict, model
+from . import cachedir, lazydict, model, transform
 from .design import Design
 
 
@@ -103,7 +103,7 @@ class Emulator:
         for obs, subobslist in self.observables:
             self._slices[obs] = {}
             for subobs in subobslist:
-                Y.append(model.data[system][obs][subobs]['Y'])
+                Y.append(transform(model.data[system][obs][subobs])['Y'])
                 n = Y[-1].shape[1]
                 self._slices[obs][subobs] = slice(nobs, nobs + n)
                 nobs += n
@@ -211,7 +211,7 @@ class Emulator:
 
         return emu
 
-    def _inverse_transform(self, Z):
+    def _inverse_transform(self, Z, log_space=True):
         """
         Inverse transform principal components to observables.
         Returns a nested dict of arrays.
@@ -221,14 +221,16 @@ class Emulator:
         Y = np.dot(Z, self._trans_matrix[:Z.shape[-1]])
         Y += self.scaler.mean_
 
-        return {
+        mean = {
             obs: {
-                subobs: Y[..., s]
+                subobs: Y[..., s] if log_space else np.exp(Y[..., s])
                 for subobs, s in slices.items()
             } for obs, slices in self._slices.items()
         }
 
-    def predict(self, X, return_cov=False):
+        return mean
+
+    def predict(self, X, return_cov=False, log_space=True):
         """
         Predict model output at `X`.
         X must be a 2D array-like with shape ``(nsamples, ndim)``.  It is passed
@@ -257,7 +259,8 @@ class Emulator:
             gp_mean, gp_cov = zip(*gp_mean)
 
         mean = self._inverse_transform(
-            np.concatenate([m[:, np.newaxis] for m in gp_mean], axis=1)
+            np.concatenate([m[:, np.newaxis] for m in gp_mean], axis=1),
+            log_space=log_space
         )
 
         if return_cov:
@@ -277,27 +280,6 @@ class Emulator:
             return mean, _Covariance(cov, self._slices)
         else:
             return mean
-
-    def sample_y(self, X, n_samples=1, random_state=None):
-        """
-        Sample model output at `X`.
-        Returns a nested dict of observable arrays, each with shape
[…]

### 2018-05-04 src/expt.py

@@ -4,6 +4,7 @@ Prints all data when run as a script.
 """
 
 from collections import defaultdict
+import copy
 import logging
 import pickle
 import re
@@ -13,7 +14,7 @@ from urllib.request import urlopen
 import numpy as np
 import yaml
 
-from . import cachedir, systems
+from . import cachedir, systems, transform
 
 
 class HEPData:
@@ -308,26 +309,6 @@ def pPb5020_flows(mode):
     )
 
 
-def log_transform(data, system):
-    """
-    Transform experimental data and errors to log space,
-
-    y = log y,
-    d(log y) = dy/y.
-
-    This transformation, which is also performed on the model observables,
-    improves PCA decomposition and emulator performance.
-
-    """
-    for obs, obs_data in data[system].items():
-        for subobs, subobs_data in obs_data.items():
-            y, yerr = [subobs_data[k] for k in ('y', 'yerr')]
-            ystat, ysys = [yerr[k] for k in ('stat', 'sys')]
-            np.divide(ystat, y, out=ystat)
-            np.divide(ysys, y, out=ysys)
-            np.log(y, out=y)
-
-
 def _data():
     """
     Curate the experimental data using the `HEPData` class and return a nested
@@ -383,10 +364,6 @@ def _data():
                     maxcent=100
                 )
 
-    # Log transform observables
-    log_transform(data, 'pPb5020')
-    log_transform(data, 'PbPb5020')
-
     return data
 
 
@@ -428,7 +405,7 @@ def cov(
 
     """
     def unpack(obs, subobs):
-        dset = data[system][obs][subobs]
+        dset = transform(data[system][obs][subobs])
         yerr = dset['yerr']
 
         try:

### 2018-05-04 src/mcmc.py

@@ -33,7 +33,7 @@ import h5py
 import numpy as np
 from scipy.linalg import lapack
 
-from . import workdir, systems, expt
+from . import workdir, systems, expt, transform
 from .design import Design
 from .emulator import emulators
 
@@ -152,8 +152,11 @@ class Chain:
         self._expt_y = {}
         self._expt_cov = {}
 
+        # log transform experimental data
+        expt_data = transform(expt.data)
+
         # pre-compute the experimental data vectors and covariance matrices
-        for sys, sysdata in expt.data.items():
+        for sys, sysdata in expt_data.items():
 
             # system specific observables
             observables = {
@@ -187,7 +190,7 @@ class Chain:
             self._expt_cov[sys] = np.empty((nobs, nobs))
 
             for obs1, subobs1, slc1 in self._slices[sys]:
-                self._expt_y[sys][slc1] = expt.data[sys][obs1][subobs1]['y']
+                self._expt_y[sys][slc1] = expt_data[sys][obs1][subobs1]['y']
                 for obs2, subobs2, slc2 in self._slices[sys]:
                     self._expt_cov[sys][slc1, slc2] = expt.cov(
                         sys, obs1, subobs1, obs2, subobs2
@@ -369,7 +372,7 @@ class Chain:
         with self.dataset() as d:
             return np.array(d[:, ::thin, indices]).reshape(-1, ndim)
 
-    def samples(self, n=1):
+    def samples(self, n=1, log_space=False):
         """
         Predict model output at `n` parameter points randomly drawn from the
         chain.
@@ -382,7 +385,7 @@ class Chain:
                 ])
             ])
 
-        return self._predict(X)
+        return self._predict(X, log_space=log_space)
 
 
 def credible_interval(samples, ci=.9):

### 2018-05-04 src/model.py

@@ -36,6 +36,7 @@ Of course, if you have a different data organization scheme and/or format,
 that's fine.  Modify the code for your needs.
 """
 
+import copy
 import logging
 from pathlib import Path
 import pickle
@@ -287,9 +288,7 @@ class ModelData:
             for design_point, events in self.design_events
         ]).squeeze()
 
-        # Log-transform events
-        assert np.all(Y > 0)
-        return {'x': x, bin_type: bins, 'Y': np.log(Y)}
+        return {'x': x, bin_type: bins, 'Y': Y}
 
 
 def _data(system, dataset='main'):
@@ -360,7 +359,6 @@ def _data(system, dataset='main'):
 
 
 data = lazydict(_data, 'main')
-validation_data = lazydict(_data, 'validation')
 map_data = lazydict(_data, 'map')
 
 

### 2018-05-04 src/plots.py

@@ -332,7 +332,8 @@ def _observables_plots():
         dict(
             title='Yields',
             ylabel=(r'$dN_\mathrm{ch}/d\eta$'),
-            ylim=(0, 9),
+            ylim=(1, 5000),
+            yscale='log',
             subplots=[
                 ('dNch_deta', None, dict(label=r'$N_\mathrm{ch}$')),
             ]
@@ -340,7 +341,7 @@ def _observables_plots():
         dict(
             title='Mean $p_T$',
             ylabel=r'$\langle p_T \rangle$ [GeV]',
-            ylim=(-1.25, 0.7),
+            ylim=(0, 1.5),
             subplots=[
                 ('mean_pT', None, dict(label=r'$\mathrm{mean}\ p_T$')),
             ]
@@ -348,7 +349,7 @@ def _observables_plots():
         dict(
             title='Flow cumulants',
             ylabel=r'$v_n\{2\}$',
-            ylim=(-7, 0),
+            ylim=(0, .2),
             subplots=[
                 ('vnk', (n, 2), dict(label='$v_{}$'.format(n)))
                 for n in [2, 3, 4]
@@ -693,7 +694,7 @@ def find_map():
         )
     )
 
-    pred = chain._predict(np.atleast_2d(full_x(res.x)), return_cov=True)
+    pred = chain._predict(np.atleast_2d(full_x(res.x)), log_space=False)
 
     plots = _observables_plots()
 
@@ -715,24 +716,15 @@ def find_map():
             scale = opts.get('scale')
 
             try:
-                model_data = model.data[system][obs][subobs]
+                x = model.data[system][obs][subobs]['x']
             except KeyError:
                 continue
-
-            x = model_data['x']
-            mean, cov = pred[system]
-            y = mean[obs][subobs][0]
-            yerr = np.sqrt(cov[(obs, subobs), (obs, subobs)][0].T.diagonal())
+            y = pred[system][obs][subobs][0]
 
             if scale is not None:
                 y = y*scale
 
-            print(y[0])
-            ax.plot(x, y, color=color, zorder=1)
-            ax.fill_between(
-                x, y - yerr, y + yerr,
-                alpha=.2, color=color, lw=0, zorder=0
-            )
+            ax.plot(x, y, color=color)
 
             if 'label' in opts:
                 ax.text(
@@ -769,8 +761,6 @@ def find_map():
                 color='.9', zorder=-10
             )
 
-            print(yexp[0])
[…]

### 2018-05-05 talk/qm2018.tex
move log transform to init
Add some slides

@@
       \bigskip
-      \includegraphics<1>[width=\textwidth]{region_shear_bulk}
-      \includegraphics<2>[width=\textwidth]{region_shear_bulk}
+      \includegraphics<1>[width=\textwidth]{fig/region_shear_bulk_blank}
+      \includegraphics<2>[width=\textwidth]{fig/region_shear_bulk}
     \end{column}
@@
       \begin{itemize}
-        \item Trade Gaussian nucleon for lumpy nucleon
+        \item Trade Gaussian nucleons for lumpy nucleons
       \end{itemize}
@@
     \node (title) at (6.0, 3.5) {\color{theme}{\scshape \Large Bayesian posterior}};
-    \node<1-2>[draw=none, text width=6.5cm, align=left, below=.2cm of title] (b1) {%
+    \node<1>[draw=none, text width=6.5cm, align=left, below=.4cm of title] (b1) {%
       \small One row (column) for each parameter
       };
-    \node<1-2>[draw=none, text width=6.5cm, align=left, below=.1cm of b1] (b2) {%
+    \node<1>[draw=none, text width=6.5cm, align=left, below=.1cm of b1] (b2) {%
       \small Diagonal panels: marginal distribution of a single model parameter
       };
-    \node<1-2>[draw=none, text width=6.5cm, align=left, below=.1cm of b2] (b3) {%
+    \node<1>[draw=none, text width=6.5cm, align=left, below=.1cm of b2] (b3) {%
       \small Off-diagonal panels: joint distribution between a pair of model parameters
       };
-    \node<2>[draw=none, text width=6.5cm, align=left, below=.2cm of b3] (b4) {%
+    \node<1>[draw=none, text width=6.5cm, align=left, below=.2cm of b3] (b4) {%
       \textcolor{theme}{Let's look at some specific regions...}
       };
-    \draw<1-2>[->, thick, theme] (b2.west) to [out=180, in=90] (-.58, 1.3);
-    \draw<1-2>[->, thick, theme] (b3.west) to [out=180, in=90] (-.05, 0);
-    \node<3>[xshift=-2.8cm, yshift=3.2cm, scale=.3] {\includegraphics[scale=.15]{magnifying-glass}};
-    \node<4>[xshift=-1.2cm, yshift=1.6cm, scale=.3] {\includegraphics[scale=.15]{magnifying-glass}};
-    \node<5>[xshift=-.7cm, yshift=1.1cm, scale=.3] {\includegraphics[scale=.15]{magnifying-glass}};
+    \draw<1>[->, thick, theme] (b2.west) to [out=180, in=90] (-.35, .8);
+    \draw<1>[->, thick, theme] (b3.west) to [out=180, in=90] (.1, -.95);
+    \node<2>[xshift=-2.7cm, yshift=3.2cm, scale=.3] {\includegraphics[scale=.15]{magnifying-glass}};
+    \node<2>[below=.4cm of title, align=center] {
+        \includegraphics[scale=.8]{posterior_p}\\[2ex]
+        \small Initial entropy deposition\\
+        \small consistent with geometric mean\\[.5ex]
+        $dS/dy \sim \sqrt{T_A T_B}$
+      };
+    \node<3>[xshift=-1.05cm, yshift=1.5cm, scale=.3] {\includegraphics[scale=.15]{magnifying-glass}};
+    \node<3>[below=.4cm of title, align=center] {
+        \includegraphics[scale=.8]{posterior_parton_number}\\[2ex]
+        \small Evidence for nucleon substructure $N_\text{partons} > 1$.\\
+        \small Does not prefer a specific parton number.
+      };
+    \node<4>[xshift=.62cm, yshift=-.15cm, scale=.3] {\includegraphics[scale=.15]{magnifying-glass}};
+    \node<4>[below=.4cm of title, align=center] {
+        \includegraphics[scale=.8]{posterior_freestreaming}\\[2ex]
+        \small Analysis cannot tolerate free-streaming\\
+        \small times larger than ${\sim} 1$~fm/$c$
+      };
+    \node<5>[xshift=-.5cm, yshift=.9cm, scale=.3] {\includegraphics[scale=.15]{magnifying-glass}};
+    \node<5>[below=.4cm of title, align=center] {
+        \includegraphics[scale=.8]{posterior_structure}\\[2ex]
+        \small Parton structure parameter $\chi_\mathrm{struct} = 0.28_{-0.16}^{+0.17}$,\\
+        \small Was used to reparametrize parton width as\\[.5ex]
+        \small $v = v_\text{min} + \chi_\text{struct}(w - v_\text{min})$.\\[1ex]
+        \small \hfill Let's revert $\chi_\text{struct}$ back to a width...
+      };
   \end{tikzpicture}
@@
 
-\begin{frame}{\scshape Maximum a posteriori parameters}
-  \includegraphics[height=.8\textheight]{find_map}
-\end{frame}
-
-\begin{frame}{\scshape Posterior nucleon shape}
+\begin{frame}{\scshape Nucleon size and shape}
   \begin{columns}
@@
       \begin{itemize}
-        \item Parton sampling radius (approx nucleon width) is not well constrained\\[2ex]
-        \item Analysis prefers small partons,\\
-          $w_\text{parton} = 0.38_{-0.18}^{+0.21}$~fm\\[2ex]
-        \item[\small\faicon{exclamation-triangle}] Results are likely sensitive to assumed
-          details of the model, e.g.\ equation of state, pre-equilibirum dynamics and viscous
-          corrections to freezeout
+        \item Nucleon width\, $w = 0.9_{-0.14}^{+0.16}$~fm\\[.5ex]
+          (consistent with rms charge radius)\\[2ex]
+        \item Parton width\, $v = 0.4_{-0.14}^{+0.16}$~fm\\[2ex]
+        \item Results likely sensitive to model details,
+          e.g.\ equation of state, pre-equilibirum dynamics and viscous
+          corrections $\delta f$\\[2ex]
+        \item Nevertheless tantalizing question, can hydro be used
+          to measure the size and shape of the proton?
       \end{itemize}
+    \end{column}
+  \end{columns}
+\end{frame}
+
+\begin{frame}{\scshape QGP viscosities}
+  \bigskip
+  \begin{columns}
+    \begin{column}{.6\textwidth}
+      \includegraphics[width=\textwidth]{fig/region_shear_bulk_blank}\\[2ex]
+      \includegraphics[width=\textwidth]{fig/region_shear_bulk}
+    \end{column}
+    \begin{column}{.4\textwidth}
+      p-Pb \& Pb-Pb at 5.02 TeV
     \end{column}

### 2018-05-05 src/plots.py

@@ -953,7 +953,7 @@ def posterior_p():
     Distribution of trento p parameter with annotations for other models.
 
     """
-    plt.figure(figsize=figsize(.8, .35))
+    plt.figure(figsize=figsize(.6, .35))
     ax = plt.axes()
 
     data = mcmc.Chain().load('trento_p').ravel()
@@ -1007,6 +1007,49 @@ def posterior_p():
     ax.set_ylim(0, 1.01*y.max())
 
     set_tight(pad=0)
+
+
+def posterior_parameter(parameter, label, xticks):
+    """
+    Marginal distribution of a single parameter.
+
+    """
+    plt.figure(figsize=figsize(.6, .35))
+    ax = plt.axes()
+
+    data = mcmc.Chain().load(parameter).ravel()
+
+    counts, edges = np.histogram(data, bins=50)
+    x = (edges[1:] + edges[:-1]) / 2
+    y = counts / counts.max()
+    interp = PchipInterpolator(x, y)
+    x = np.linspace(x[0], x[-1], 10*x.size)
+    y = interp(x)
+    ax.plot(x, y, color=plt.cm.Blues(0.8))
+    ax.fill_between(x, y, color=plt.cm.Blues(0.15), zorder=-10)
+    ax.spines['left'].set_visible(False)
+
+    ax.set_xlabel(label)
+    ax.set_xticks(xticks)
+    ax.set_yticks([])
+    ax.set_ylim(0, 1.01*y.max())
+
+    set_tight(pad=0)
+
+
+@plot
+def posterior_parton_number():
+    posterior_parameter('parton_number', 'Parton number', [1, 3, 5, 7, 9])
+
+
+@plot
+def posterior_freestreaming():
+    posterior_parameter('tau_fs', 'Free streaming time [fm/$c$]', [.1, .8, 1.5])
+
+
+@plot
+def posterior_structure():
+    posterior_parameter('parton_struct', '$\chi_\mathrm{struct}$', [0, .5, 1])
 
 
 @plot
@@ -2101,12 +2144,14 @@ def proton_posterior_shape():
         nucleon_width, parton_width, bins=100, cmap=cmap
     )
 
-    plt.plot((.4, 1.2), (.4, 1.2), color='.5', clip_on=False)
-    plt.annotate('$w_\mathrm{parton} = w_\mathrm{nucleon}$',
-                 xy=(.77, .83), xycoords='data', fontsize=fontsize['large'],
[…]

### 2018-05-06 src/design.py
tweak NPC

@@ -35,15 +35,15 @@ Remove outlier design points. These points are near the edge of the
 design and produce far too few particles to be physically reasonable.
 
 """
-bad_points = [61, 62, 79, 83, 104, 107, 112, 115, 123, 126, 129, 132, 141, 143,
-              153, 156, 159, 162, 170, 176, 178, 183, 184, 185, 193, 199, 203,
-              211, 215, 219, 223, 225, 226, 227, 234, 235, 236, 242, 245, 247,
-              248, 249, 266, 267, 273, 276, 285, 291, 294, 296, 297, 298, 301,
-              308, 312, 315, 317, 319, 321, 323, 326, 330, 334, 335, 338, 339,
-              341, 344, 345, 346, 355, 357, 358, 360, 363, 365, 367, 372, 376,
-              378, 379, 382, 386, 392, 395, 405, 414, 418, 421, 422, 428, 429,
-              442, 445, 446, 451, 452, 453, 456, 457, 459, 460, 461, 464, 466,
-              467, 471, 472, 473, 480, 482, 483, 492, 498]
+bad_points = [
+    129, 392, 266, 141, 143, 273, 401, 276, 405, 153, 156, 285, 414,
+    159, 162, 418, 163, 421, 422, 294, 296, 298, 429, 301, 308, 183,
+    184, 312, 442, 446, 319, 451, 453, 326, 199, 457, 330, 459, 460,
+    461, 334, 203, 464, 466, 83, 211, 338, 471, 472, 473, 346, 223,
+    355, 357, 360, 234, 363, 492, 107, 236, 367, 235, 498, 115, 242,
+    376, 378
+]
+
 
 def generate_lhs(npoints, ndim, seed):
     """

### 2018-05-06 src/emulator.py

@@ -322,7 +322,7 @@ if __name__ == '__main__':
     kwargs = vars(args)
 
     for s in kwargs.pop('systems'):
-        npc = {'pPb5020': 7, 'PbPb5020': 12}[s]
+        npc = {'pPb5020': 7, 'PbPb5020': 8}[s]
         emu = Emulator.from_cache(s, npc=npc, **kwargs)
 
         print(s)

### 2018-05-06 src/expt.py

@@ -141,7 +141,7 @@ class HEPData:
             .format(name, quals)
         )
 
-    def dataset(self, name=None, maxcent=100, ignore_bins=[], **quals):
+    def dataset(self, name=None, maxcent=70, ignore_bins=[], **quals):
         """
         Return a dict containing:
 
@@ -361,7 +361,7 @@ def _data():
                     'V{}{{{}{}}}'.format(
                         n, k, ', |DELTAETA|>1' if k == 2 else ''
                     ),
-                    maxcent=100
+                    maxcent=(70 if n == 2 else 50)
                 )
 
     return data

### 2018-05-06 src/plots.py

@@ -42,7 +42,7 @@ from sklearn.gaussian_process import GaussianProcessRegressor as GPR
 from sklearn.gaussian_process import kernels
 from sklearn.model_selection import KFold
 
-from . import cachedir, workdir, systems, parse_system, expt, model, mcmc
+from . import cachedir, workdir, systems, parse_system, expt, model, mcmc, transform
 from .design import Design
 from .emulator import Emulator, emulators
 
@@ -423,29 +423,17 @@ def _observables(posterior=False):
                 if obs == 'dNch_deta':
                     ax.set_xlim(0, 55)
                     ax.set_xlabel('Centrality %')
-
-                    # TODO guidlines
-                    #ax.plot([0, 50], [6, 5])
-                    #ax.plot([0, 50], [2, 1])
                 elif obs == 'mean_pT':
                     ax.set_xlim(1, 6)
                     ax.set_xlabel(r'$n_\mathrm{ch}\,/\langle n_\mathrm{ch} \rangle$')
-
-                    # TODO guidlines
-                    #ax.axhline(0.5)
-                    #ax.axhline(-1.0)
                 elif obs == 'vnk':
                     ax.set_xlim(1, 6)
                     ax.set_xlabel('/'.join([
                         r'$n^\mathrm{offline}_\mathrm{trk}$',
                         r'$\langle n^\mathrm{offline}_\mathrm{trk} \rangle$',
                     ]))
-
-                    # TODO guidlines
-                    #ax.axhline(-6)
-                    #ax.axhline(0)
             else:
-                ax.set_xlim(0, 76)
+                ax.set_xlim(0, 80)
 
         if plot.get('yscale') == 'log':
             ax.set_yscale('log')
@@ -1242,7 +1230,7 @@ def pca():
     ax_y = fig.add_subplot(gs[1:, -1], sharey=ax_j)
 
     x, y = (
-        model.data['pPb5020'][obs][subobs]['Y'][:, 3]
+        model.data['PbPb5020'][obs][subobs]['Y'][:, 3]
         for obs, subobs in [('dNch_deta', None), ('vnk', (2, 2))]
     )
     xlabel = r'$dN_\mathrm{ch}/d\eta$'
@@ -1314,8 +1302,11 @@ def pca():
     set_tight(pad=.1, h_pad=.3, w_pad=.3)
 
 
+default_system = 'PbPb5020'
+
+
 @plot
-def pca_vectors_variance(system='pPb5020'):
[…]

### 2018-05-07 talk/qm2018.tex
Add some validation material to talk slides

@@
     \node {\includegraphics[height=.98\paperheight]{posterior}};
-    \node (title) at (6.0, 3.5) {\color{theme}{\scshape \Large Bayesian posterior}};
+    \node (title) at (5.75, 3.5) {\color{theme}{\scshape \Large Bayesian posterior}};
     \node<1>[draw=none, text width=6.5cm, align=left, below=.4cm of title] (b1) {%
@@
     \node<2>[below=.4cm of title, align=center] {
-        \includegraphics[scale=.8]{posterior_p}\\[2ex]
+        \includegraphics[scale=.7]{posterior_p}\\[2ex]
         \small Initial entropy deposition\\
@@
     \node<3>[below=.4cm of title, align=center] {
-        \includegraphics[scale=.8]{posterior_parton_number}\\[2ex]
+        \includegraphics[scale=.7]{posterior_parton_number}\\[2ex]
         \small Evidence for nucleon substructure $N_\text{partons} > 1$.\\
@@
     \node<4>[below=.4cm of title, align=center] {
-        \includegraphics[scale=.8]{posterior_freestreaming}\\[2ex]
+        \includegraphics[scale=.7]{posterior_freestreaming}\\[2ex]
         \small Analysis cannot tolerate free-streaming\\
@@
     \node<5>[below=.4cm of title, align=center] {
-        \includegraphics[scale=.8]{posterior_structure}\\[2ex]
-        \small Parton structure parameter $\chi_\mathrm{struct} = 0.28_{-0.16}^{+0.17}$,\\
+        \includegraphics[scale=.7]{posterior_structure}\\[2ex]
+        \small Parton structure param.\ $\chi_\mathrm{struct} = 0.28_{-0.16}^{+0.17}$,\\
         \small Was used to reparametrize parton width as\\[.5ex]
@@
       \begin{itemize}
-        \item Nucleon width\, $w = 0.9_{-0.14}^{+0.16}$~fm\\[.5ex]
-          (consistent with rms charge radius)\\[2ex]
-        \item Parton width\, $v = 0.4_{-0.14}^{+0.16}$~fm\\[2ex]
+        \item Parton sampling radius\, $w = 1.04_{-0.15}^{+0.16}$~fm\\[.5ex]
+          $^*$Note: the actual rms radius in the center of mass frame is somewhat smaller.\\[2ex]
+        \item Parton width\, $v = 0.51_{-0.17}^{+0.19}$~fm\\[.5ex]
+          (width of Gaussian constituents)\\[2ex]
         \item Results likely sensitive to model details,
@@
           corrections $\delta f$\\[2ex]
-        \item Nevertheless tantalizing question, can hydro be used
-          to measure the size and shape of the proton?
       \end{itemize}
@@
 
-\begin{frame}{\scshape QGP viscosities}
-  \bigskip
+\begin{frame}{\scshape QGP medium properties}
+\end{frame}
+
+\begin{frame}{\scshape Maximum a posteriori parameters}
+  \begin{center}
+    \Large To do if time allows...
+  \end{center}
+\end{frame}
+
+\appendix
+
+\begin{frame}[plain, noframenumbering]
+  \begin{center}
+    \Large \scshape Backup slides
+  \end{center}
+\end{frame}
+
+\begin{frame}{\scshape Emulator validation}
   \begin{columns}
-    \begin{column}{.6\textwidth}
-      \includegraphics[width=\textwidth]{fig/region_shear_bulk_blank}\\[2ex]
-      \includegraphics[width=\textwidth]{fig/region_shear_bulk}
+    \begin{column}{.35\textwidth}
+      {\scshape Preprocessing}
+      \begin{enumerate}
+        \item log transform data
+        \item subtract mean and divide out variance
+        \item project onto principle component basis
+      \end{enumerate}
+      \medskip
+      {\scshape k-fold cross validation}
+      \begin{itemize}
+        \item Subdivide design into $k$ partitions. Choose one for
+          validation and use the rest for training. Repeat for all partitions.
+      \end{itemize}
     \end{column}
-    \begin{column}{.4\textwidth}
-      p-Pb \& Pb-Pb at 5.02 TeV
+    \begin{column}{.65\textwidth}
+      \includegraphics[width=\textwidth]{validation_example}
     \end{column}
@@
 
+\begin{frame}{\scshape Emulator validation}
+    \includegraphics<1>[width=\textwidth]{validation_pPb5020}
+    \includegraphics<2>[width=\textwidth]{validation_PbPb5020}
+    \begin{itemize}
+      \item Model-data residuals should have zero mean and unit variance.
+      \item Root-mean square error expressed as percentage of overall design variance.
+    \end{itemize}
+\end{frame}
+
 \usebackgroundtemplate{}

### 2018-05-07 src/__init__.py

@@ -56,7 +56,7 @@ class lazydict(dict):
 
 def transform(data):
     """
-    Log transform model or experimental data,
+    Used to log transform model and experimental data,
 
     y = log y,
     d(log y) = dy/y.

### 2018-05-07 src/emulator.py

@@ -322,7 +322,7 @@ if __name__ == '__main__':
     kwargs = vars(args)
 
     for s in kwargs.pop('systems'):
-        npc = {'pPb5020': 7, 'PbPb5020': 8}[s]
+        npc = {'pPb5020': 7, 'PbPb5020': 9}[s]
         emu = Emulator.from_cache(s, npc=npc, **kwargs)
 
         print(s)

### 2018-05-07 src/model.py

@@ -234,28 +234,22 @@ class ModelData:
         compute_bin = _compute_bin()
 
         def compute_all_bins(events, design_point):
+            """
+            Cooper-Frye sampling neglects all energy in the collision which
+            starts below the particlization temperature Tsw. This function
+            corrects dNch/deta for missing particles below Tsw, assuming
+            that particle production follows a simple power law.
+
+            """
             trigger = events['trigger']
             minbias = (0, float('inf'))
 
-            if self.pPb_event:
-                """
-                Cooper-Frye sampling neglects all energy in the collision which
-                starts below the particlization temperature Tsw. This function
-                corrects dNch/deta for missing particles below Tsw, assuming
-                that particle production follows a simple power law.
-
-                """
-                events = correct_yield(events)
+            events = correct_yield(events)
 
             if bin_type == 'cent':
                 """
-                Sorting events into centrality bins is inaccurate for small minimum
-                bias event samples. We correct this bias by using initial entropy
-                (from a far greater number of events) to define centrality classes.
-
-                The `correct_centrality` function runs initial condition events
-                with parameters selected from each design point and uses them
-                to partition the provided minimum bias sample into centrality bins.
+                Sort minimum bias events into bins by their charged particle
+                yield.
 
                 """
                 minbias_events = events[(trigger == minbias).all(axis=1)]
@@ -264,9 +258,6 @@ class ModelData:
                     minbias_events[int((1 - b/100)*n):int((1 - a/100)*n)]
                     for a, b in bins
                 ]
-                #binned_events = correct_centrality(
-                #    self.system, design_point, minbias_events, bins
-                #)
             elif bin_type == 'mult':
                 """
                 Many p-Pb observables are plotted as a function of <Nch>. These

### 2018-05-07 src/plots.py

@@ -396,7 +396,7 @@ def _observables(posterior=False):
                 Y = Y*scale
 
             for y in Y:
-                ax.plot(x, y, color=color, alpha=.08, lw=.3)
+                ax.plot(x, y, color=color, alpha=.15, lw=.3)
 
             try:
                 dset = expt.data[system][obs][subobs]
@@ -1424,7 +1424,7 @@ def boxplot(
         )
 
 
-def validation_data(system, n_splits=20, npc=10):
+def validation_data(system, n_splits=20):
     """
     Partition the design into training and test data using K-fold
     cross validation. Train the emulator on each fold (subset of the design)
@@ -1433,6 +1433,7 @@ def validation_data(system, n_splits=20, npc=10):
     """
     design = Design(system)
     kf = KFold(n_splits=n_splits)
+    npc = {'pPb5020': 7, 'PbPb5020': 9}[system]
 
     mean_folds = []
     cov_folds = []
@@ -1461,8 +1462,7 @@ def validation_data(system, n_splits=20, npc=10):
     return mean_folds, cov_folds
 
 
-@plot
-def validation_all(system='PbPb5020'):
+def validation_all(system):
     """
     Emulator validation: normalized residuals and RMS error for each
     observable.
@@ -1508,7 +1508,7 @@ def validation_all(system='PbPb5020'):
             ):
                 boxplot(ax_box, percentiles, x=i, box_width=.8, color=color)
 
-            Ymin, Ymax = np.percentile(Y, (1, 99))
+            Ymin, Ymax = np.percentile(Y, (.5, 99.5))
             Yerr = (Y_ - Y)/(Ymax - Ymin)
             rms = 100*np.sqrt(np.square(Yerr).mean(axis=0))
             ax_rms.plot(
@@ -1521,11 +1521,14 @@ def validation_all(system='PbPb5020'):
             index = i + 2
 
     ax_box.set_xticks(ticks)
+    ax_box.tick_params(axis='x', pad=-6)
     ax_box.set_xticklabels(ticklabels)
     ax_box.tick_params('x', bottom=False, labelsize=plt.rcParams['font.size'])
 
     ax_box.set_ylim(-2.25, 2.25)
     ax_box.set_ylabel(r'Normalized residuals')
+    nuclei, roots = parse_system(system)
+    ax_box.set_title('{}-{} {} TeV'.format(*nuclei, roots/1000))
 
[…]

### 2018-05-07 src/__init__.py
Remove all traces of the log transform

@@ -7,8 +7,6 @@ from pathlib import Path
 import re
 import sys
 
-import numpy as np
-
 
 logging.basicConfig(
     stream=sys.stdout,
@@ -52,27 +50,3 @@ class lazydict(dict):
     def __missing__(self, key):
         self[key] = value = self.function(key, *self.args, **self.kwargs)
         return value
-
-
-def transform(data):
-    """
-    Used to log transform model and experimental data,
-
-    y = log y,
-    d(log y) = dy/y.
-
-    This transformation improves PCA decomposition and emulator performance.
-
-    """
-    if 'x' in data:
-        d = {}
-        for k, v in data.items():
-            if k in ['y', 'Y']:
-                d[k] = np.log(v)
-            elif k == 'yerr':
-                d[k] = {s: data['yerr'][s]/data['y'] for s in ['stat', 'sys']}
-            else:
-                d[k] = v
-        return d
-    else:
-        return {k: transform(v) for k, v in data.items()}

### 2018-05-07 src/emulator.py

@@ -23,7 +23,7 @@ from sklearn.gaussian_process import GaussianProcessRegressor as GPR
 from sklearn.gaussian_process import kernels
 from sklearn.preprocessing import StandardScaler
 
-from . import cachedir, lazydict, model, transform
+from . import cachedir, lazydict, model
 from .design import Design
 
 
@@ -103,7 +103,7 @@ class Emulator:
         for obs, subobslist in self.observables:
             self._slices[obs] = {}
             for subobs in subobslist:
-                Y.append(transform(model.data[system][obs][subobs])['Y'])
+                Y.append(model.data[system][obs][subobs]['Y'])
                 n = Y[-1].shape[1]
                 self._slices[obs][subobs] = slice(nobs, nobs + n)
                 nobs += n
@@ -211,7 +211,7 @@ class Emulator:
 
         return emu
 
-    def _inverse_transform(self, Z, log_space=True):
+    def _inverse_transform(self, Z):
         """
         Inverse transform principal components to observables.
         Returns a nested dict of arrays.
@@ -223,14 +223,14 @@ class Emulator:
 
         mean = {
             obs: {
-                subobs: Y[..., s] if log_space else np.exp(Y[..., s])
+                subobs: Y[..., s]
                 for subobs, s in slices.items()
             } for obs, slices in self._slices.items()
         }
 
         return mean
 
-    def predict(self, X, return_cov=False, log_space=True):
+    def predict(self, X, return_cov=False):
         """
         Predict model output at `X`.
         X must be a 2D array-like with shape ``(nsamples, ndim)``.  It is passed
@@ -259,8 +259,7 @@ class Emulator:
             gp_mean, gp_cov = zip(*gp_mean)
 
         mean = self._inverse_transform(
-            np.concatenate([m[:, np.newaxis] for m in gp_mean], axis=1),
-            log_space=log_space
+            np.concatenate([m[:, np.newaxis] for m in gp_mean], axis=1)
         )
 
         if return_cov:

### 2018-05-07 src/expt.py

@@ -14,7 +14,7 @@ from urllib.request import urlopen
 import numpy as np
 import yaml
 
-from . import cachedir, systems, transform
+from . import cachedir, systems
 
 
 class HEPData:
@@ -405,7 +405,7 @@ def cov(
 
     """
     def unpack(obs, subobs):
-        dset = transform(data[system][obs][subobs])
+        dset = data[system][obs][subobs]
         yerr = dset['yerr']
 
         try:

### 2018-05-07 src/mcmc.py

@@ -33,7 +33,7 @@ import h5py
 import numpy as np
 from scipy.linalg import lapack
 
-from . import workdir, systems, expt, transform
+from . import workdir, systems, expt
 from .design import Design
 from .emulator import emulators
 
@@ -152,11 +152,8 @@ class Chain:
         self._expt_y = {}
         self._expt_cov = {}
 
-        # log transform experimental data
-        expt_data = transform(expt.data)
-
         # pre-compute the experimental data vectors and covariance matrices
-        for sys, sysdata in expt_data.items():
+        for sys, sysdata in expt.data.items():
 
             # system specific observables
             observables = {
@@ -190,7 +187,7 @@ class Chain:
             self._expt_cov[sys] = np.empty((nobs, nobs))
 
             for obs1, subobs1, slc1 in self._slices[sys]:
-                self._expt_y[sys][slc1] = expt_data[sys][obs1][subobs1]['y']
+                self._expt_y[sys][slc1] = expt.data[sys][obs1][subobs1]['y']
                 for obs2, subobs2, slc2 in self._slices[sys]:
                     self._expt_cov[sys][slc1, slc2] = expt.cov(
                         sys, obs1, subobs1, obs2, subobs2
@@ -372,7 +369,7 @@ class Chain:
         with self.dataset() as d:
             return np.array(d[:, ::thin, indices]).reshape(-1, ndim)
 
-    def samples(self, n=1, log_space=False):
+    def samples(self, n=1):
         """
         Predict model output at `n` parameter points randomly drawn from the
         chain.
@@ -385,7 +382,7 @@ class Chain:
                 ])
             ])
 
-        return self._predict(X, log_space=log_space)
+        return self._predict(X)
 
 
 def credible_interval(samples, ci=.9):

### 2018-05-07 src/plots.py

@@ -42,7 +42,7 @@ from sklearn.gaussian_process import GaussianProcessRegressor as GPR
 from sklearn.gaussian_process import kernels
 from sklearn.model_selection import KFold
 
-from . import cachedir, workdir, systems, parse_system, expt, model, mcmc, transform
+from . import cachedir, workdir, systems, parse_system, expt, model, mcmc
 from .design import Design
 from .emulator import Emulator, emulators
 
@@ -682,7 +682,7 @@ def find_map():
         )
     )
 
-    pred = chain._predict(np.atleast_2d(full_x(res.x)), log_space=False)
+    pred = chain._predict(np.atleast_2d(full_x(res.x)))
 
     plots = _observables_plots()
 
@@ -1487,7 +1487,7 @@ def validation_all(system):
 
             # model data
             try:
-                Y = transform(model_data[obs][subobs])['Y']
+                Y = model_data[obs][subobs]['Y']
             except KeyError:
                 continue
 
@@ -1590,7 +1590,7 @@ def validation_example(
     ax_scatter, ax_hist = axes
 
     # model data
-    model_data  = transform(model.data[system])
+    model_data  = model.data[system]
     vdata = model_data[obs][subobs]
     cent_slc = (slice(None), vdata['cent'].index(cent))
     y = vdata['Y'][cent_slc]

### 2018-05-08 talk/qm2018.tex
Remove all traces of the log transform
work on nucleon width correction

@@
 \title{Estimating nucleon substructure properties in a\\[.2ex]
-unified hydrodynamic model of p-Pb and Pb-Pb collisions}
+unified model of p-Pb and Pb-Pb collisions}
 \author[J.\ S.\ Moreland]{\scshape J.\ S.\ Moreland, J.\ E.\ Bernhard \& S.\ A.\ Bass}
@@
 
-\begin{frame}{\scshape Flow in small systems}
+\begin{frame}{\scshape Collectivity in small systems, is it flow?}
   \smallskip
@@
         \item Is hydro applicable to small collision systems?
-        \item If so, can we fit small system collectivity using current hydro models?
+        \item If so, can we describe small system collectivity using current hydro models?
       \end{itemize}
@@
       sense down to length scales $\Delta x=.2$~fm.\\[1ex]
-      Under this assumption, does a unified hydrodynamic description of p+A and A+A data even exist?
+      Under this assumption, does a unified hydrodynamic description of p+A and A+A data exist?
     \end{column}
@@
   \end{center}
-\end{frame}
-
-\begin{frame}[t]{\scshape Heavy-ion collision model}
-  \centering
-  \medskip
-  \begin{tikzpicture}
-    \node (a) {\includegraphics[width=.95\textwidth]{hannah_collision}};
-    \node[align=left, anchor=north west] at (a.north west) {%
-      \scriptsize Figure: H. Petersen, MADAI
-      };
-  \end{tikzpicture}
-  \begin{columns}[T]
-    \begin{column}{0.3\textwidth}
-      \texttt{\scshape\color{theme} TrENTo} {\small initial conditions}\\
-      \smallskip\hrule\medskip
-      {\scriptsize parametric entropy deposition}\\
-      {\tiny \color{theme}PRC.92.011901}\\[2ex]
-      \texttt{\scshape\color{theme} freestream} {\small pre-flow}\\
-      \smallskip\hrule\medskip
-      {\scriptsize infinitely weak coupling limit}\\
-      {\tiny \color{theme} PRC.91.064906, ~PRC.80.034902}
-    \end{column}
-    \hfill
-    \begin{column}{0.3\textwidth}
-      \texttt{\scshape\color{theme} VISH2+1} {\small viscous hydro}\\
-      \smallskip\hrule\medskip
-      {\scriptsize 14-mom.\ approx w/ shear \& bulk}\\
-      {\tiny \color{theme} PRC.77.064901, ~J.CPC.2015.08.039}\\[2ex]
-      \texttt{\scshape\color{theme}frzout} {\small sampler}\\
-      \smallskip\hrule\medskip
-      {\scriptsize non-RTA shear \& bulk corrections}\\
-      {\tiny \color{theme} (Bernhard thesis) arXiv:1804.06469}
-    \end{column}
-    \hfill
-    \begin{column}{0.3\textwidth}
-      \texttt{\scshape\color{theme} UrQMD} {\small Boltzmann cascade}
-      \smallskip\hrule\medskip
-      {\scriptsize hadronic afterburner, simulate scatterings and decays}\\
-      {\tiny\color{theme} PPNP.98.00058, ~JPG.25.9.308}\\[2ex]
-      {\scriptsize Final observables calculated as similar as possible to experiment}
-    \end{column}
-  \end{columns}
-  \begin{tikzpicture}[overlay]
-    \draw (-3, -.2) -- (-2.7, -.2);
-    \draw (-2.7, -.2) -- (-2.7, 3.6);
-    \draw (-2.7, 3.6) -- (-2.4, 3.6);
-    \draw (2.4, -.2) -- (2.7, -.2);
-    \draw (2.7, -.2) -- (2.7, 3.6);
-    \draw (2.7, 3.6) -- (3.0, 3.6);
-  \end{tikzpicture}
 \end{frame}
@@
 
+\begin{frame}[t]{\scshape Heavy-ion collision model}
+  \centering
+  \medskip
+  \begin{tikzpicture}
+    \node (a) {\includegraphics[width=.95\textwidth]{hannah_collision}};
+    \node[align=left, anchor=north west] at (a.north west) {%
+      \scriptsize Figure: H. Petersen, MADAI
+      };
+  \end{tikzpicture}
+  \begin{columns}[T]
+    \begin{column}{0.3\textwidth}
+      \texttt{\scshape\color{theme} TrENTo} {\small initial conditions}\\
+      \smallskip\hrule\medskip
+      {\scriptsize parametric entropy deposition}\\
+      {\tiny \color{theme}PRC.92.011901}\\[2ex]
+      \texttt{\scshape\color{theme} freestream} {\small pre-flow}\\
+      \smallskip\hrule\medskip
+      {\scriptsize infinitely weak coupling limit}\\
+      {\tiny \color{theme} PRC.91.064906, ~PRC.80.034902}
+    \end{column}
+    \hfill
+    \begin{column}{0.3\textwidth}
+      \texttt{\scshape\color{theme} VISH2+1} {\small viscous hydro}\\
+      \smallskip\hrule\medskip
+      {\scriptsize 14-mom.\ approx w/ shear \& bulk}\\
+      {\tiny \color{theme} PRC.77.064901, ~J.CPC.2015.08.039}\\[2ex]
+      \texttt{\scshape\color{theme}frzout} {\small sampler}\\
+      \smallskip\hrule\medskip
+      {\scriptsize non-RTA shear \& bulk corrections}\\
+      {\tiny \color{theme} (Bernhard thesis) arXiv:1804.06469}
+    \end{column}
+    \hfill
+    \begin{column}{0.3\textwidth}
+      \texttt{\scshape\color{theme} UrQMD} {\small Boltzmann cascade}
+      \smallskip\hrule\medskip
+      {\scriptsize hadronic afterburner, simulate scatterings and decays}\\
+      {\tiny\color{theme} PPNP.98.00058, ~JPG.25.9.308}\\[2ex]
+      {\scriptsize Final observables calculated as similar as possible to experiment}
+    \end{column}
+  \end{columns}
+  \begin{tikzpicture}[overlay]
+    \draw (-3, -.2) -- (-2.7, -.2);
+    \draw (-2.7, -.2) -- (-2.7, 3.6);
+    \draw (-2.7, 3.6) -- (-2.4, 3.6);
+    \draw (2.4, -.2) -- (2.7, -.2);
+    \draw (2.7, -.2) -- (2.7, 3.6);
+    \draw (2.7, 3.6) -- (3.0, 3.6);
+  \end{tikzpicture}
+\end{frame}
+
 \begin{frame}[b, plain]{\scshape Parameter estimates from lead-lead collisions}
@@
         \includegraphics[width=.9\columnwidth]{pb208}\\
-        Pb$^{208}$ nucleus
+        $^{208}$Pb nucleus
       \end{figure}
@@
       \begin{itemize}
-        \item Sample nucleon radii from spherical or deformed Woods-Saxon distributions
-        \item Euler angles resampled to preserve minimum nucleon distance
+        \item Sample nucleon positions from spherical or deformed Woods-Saxon distributions
+        \item Solid angles resampled to preserve minimum nucleon distance
           $d_\mathrm{min}$ [fm]
@@
     \begin{column}{.2\textwidth}
-      \centering \small Parton width
+      \centering \small Constituent width
       \begin{figure}
@@
     \end{column}
-    \begin{column}{.2\textwidth}
-      \centering \small Parton number
+    \begin{column}{.23\textwidth}
+      \centering \small Constituent number
       \begin{figure}
@@
           \draw [thick, dashed] (0, -2.2) circle (1cm);
-          \foreach \theta in {0, 120, 240}{
-            \draw [thick] ({\theta}:.39) [yshift=-2.2cm] circle (.3cm);
+          \foreach \theta/\radius in {
+            0/.4, 65/.5, 115/.45, 190/.53, 240/.44, 310/.62
+          }{
+            \draw [thick] ({\theta}:\radius) [yshift=-2.2cm] circle (.3cm);
           }
           \draw [thick, dashed] (0, -4.4) circle (1cm);
-          \foreach \theta in {0, 180}{
+          \foreach \theta in {0, 120, 240}{
             \draw [thick] ({\theta}:.39) [yshift=-4.4cm] circle (.3cm);
@@
     \end{column}
-    \begin{column}{.4\textwidth}
+    \begin{column}{.37\textwidth}
       \centering
@@
         \small
-        \item Width of the Gaussian distribution used to sample
-          parton radial coordinates
-        \item Number of Gaussian partons inside each nucleon
-        \item Width of the Gaussian partons
+        \item Radius of Gaussian distribution used to sample
+          constituent positions
+        \item Gaussian constituent width
+        \item Number of constituents in each nucleon
       \end{itemize}
@@
         \small
-        \item Parton spatial correlations,\\
+        \item Spatial correlations,
           see talk by Alba Soto Ontoso
@@
       \begin{figure}
-        \includegraphics<1>[width=\textwidth]{trento_nucleons}
-        \includegraphics<2>[width=\textwidth]{trento_participants}\\
+        \includegraphics[width=\textwidth]{trento_participants}\\
         \small MC-Glauber cross sections\\
-        (proton d.o.f. shown)
+        (one constituent)
       \end{figure}
@@
       \begin{align*}
-        P_\text{coll} &= 1 - \exp[-\sigma_{gg} T_{pp}(b)] \\
+        P_\text{coll} &= 1 - \exp[-\sigma_\text{eff} T_\text{pp}(b)] \\
         \sigma_\text{nn}^\text{inel} &= \int 2\pi b\, db\, P_\text{coll}(b)
@@
         \end{equation*}
-        Solve for opacity parameter $\sigma_{gg}$ numerically
+        Solve for effective cross section parameter $\sigma_\text{eff}$ numerically
       }
@@
       \begin{figure}
-        \includegraphics[width=\textwidth]{trento_thickness}
+        \includegraphics[width=\textwidth]{trento_thickness}\\
+        \small Participant matter
       \end{figure}
@@
       \begin{figure}
-        \includegraphics[width=\textwidth]{trento_entropy}
+        \includegraphics[width=\textwidth]{trento_entropy}\\
+        \small Entropy density
       \end{figure}
@@
         \includegraphics[width=\textwidth]{trento_freestream}\\
-        \small arrows: fluid velocity\\weighted by energy density
+        \small Arrows: fluid velocity\\weighted by energy density
       \end{figure}
@@
       \bigskip \centering
-      \textcolor{theme}{Pre-equillibrium phase:}
+      \textcolor{theme}{Pre-equilibrium phase:}
       \begin{itemize}
@@
         \small
-        \item Parton width reparametrized:\\[1ex]
-          $v = v_\text{min} + \X(v_\text{max} - v_\text{min})$\\[1ex]
+        \item Recast constituent width $w$ as:\\[1ex]
+          $w = w_\text{min} + \X(r - w_\text{min})$\\[1ex]
           $v_\text{min} = 0.2$~fm,~ $v_\text{max} = w$\\[1ex]
@@
     \draw[<-] (current page.center) ++(-.75, 1.7) --
-      node[above] {\footnotesize Prior}
-      node[below] {\tiny (training data)}
+      node[above] {\small Prior}
+      node[below] {\scriptsize (training data)}
       ++(1.5, 0);
     \draw[->] (current page.center) ++(-.75, -1.2) --
-      node[above] {\footnotesize Posterior}
-      node[below] {\tiny (samples)}
+      node[above] {\small Posterior}
+      node[below] {\scriptsize (samples)}
       ++(1.5, 0);
@@
 \begin{frame}[plain]
-  \begin{tikzpicture}
-    \node {\includegraphics[height=.98\paperheight]{posterior}};
-    \node (title) at (5.75, 3.5) {\color{theme}{\scshape \Large Bayesian posterior}};
+  \begin{tikzpicture}[remember picture, overlay]
+    \node[anchor=west] at (current page.west) {
+      \includegraphics[height=.98\paperheight]{posterior}
+    };
+    \node[anchor=north, xshift=3cm, yshift=-.5cm] (title) at (current page.north) {
+      \color{theme}{\scshape \Large Bayesian posterior}
+    };
     \node<1>[draw=none, text width=6.5cm, align=left, below=.4cm of title] (b1) {%
@@
       };
-    \draw<1>[->, thick, theme] (b2.west) to [out=180, in=90] (-.35, .8);
-    \draw<1>[->, thick, theme] (b3.west) to [out=180, in=90] (.1, -.95);
-    \node<2>[xshift=-2.7cm, yshift=3.2cm, scale=.3] {\includegraphics[scale=.15]{magnifying-glass}};
+    \draw<1>[->, thick, theme] (b2.west) to [out=180, in=90] (2.6, .8);
+
+    \draw<1>[->, thick, theme] (b3.west) to [out=180, in=90] (2.6, -.6);
+
+    \node<2>[xshift=0cm, yshift=0cm, scale=.3] {
+      \includegraphics[scale=.15]{magnifying-glass}
+    };
     \node<2>[below=.4cm of title, align=center] {
-        \includegraphics[scale=.7]{posterior_p}\\[2ex]
+        \includegraphics[width=.5\textwidth]{posterior_p}\\[1ex]
         \small Initial entropy deposition\\
@@
       };
-    \node<3>[xshift=-1.05cm, yshift=1.5cm, scale=.3] {\includegraphics[scale=.15]{magnifying-glass}};
-    \node<3>[below=.4cm of title, align=center] {
-        \includegraphics[scale=.7]{posterior_parton_number}\\[2ex]
-        \small Evidence for nucleon substructure $N_\text{partons} > 1$.\\
-        \small Does not prefer a specific parton number.
-      };
-    \node<4>[xshift=.62cm, yshift=-.15cm, scale=.3] {\includegraphics[scale=.15]{magnifying-glass}};
-    \node<4>[below=.4cm of title, align=center] {
-        \includegraphics[scale=.7]{posterior_freestreaming}\\[2ex]
-        \small Analysis cannot tolerate free-streaming\\
-        \small times larger than ${\sim} 1$~fm/$c$
-      };
-    \node<5>[xshift=-.5cm, yshift=.9cm, scale=.3] {\includegraphics[scale=.15]{magnifying-glass}};
-    \node<5>[below=.4cm of title, align=center] {
-        \includegraphics[scale=.7]{posterior_structure}\\[2ex]
-        \small Parton structure param.\ $\chi_\mathrm{struct} = 0.28_{-0.16}^{+0.17}$,\\
-        \small Was used to reparametrize parton width as\\[.5ex]
-        \small $v = v_\text{min} + \chi_\text{struct}(w - v_\text{min})$.\\[1ex]
-        \small \hfill Let's revert $\chi_\text{struct}$ back to a width...
-      };
+    %\node<3>[xshift=-1.05cm, yshift=1.5cm, scale=.3] {\includegraphics[scale=.15]{magnifying-glass}};
+    %\node<3>[below=.4cm of title, align=center] {
+    %    \includegraphics[width=.5\textwidth]{posterior_parton_number}\\[2ex]
+    %    \small Evidence for nucleon substructure, $N_\text{partons} > 1$.\\
+    %    \small Does not prefer a specific constituent number.
+    %  };
+    %\node<4>[xshift=-1.6cm, yshift=1.0cm, scale=.3] {\includegraphics[scale=.15]{magnifying-glass}};
+    %\node<4>[below=.4cm of title, align=center] {
+    %    \includegraphics[scale=.7]{proton_posterior_shape}\\[2ex]
+    %  };
   \end{tikzpicture}
@@
       \begin{itemize}
-        \item Parton sampling radius\, $w = 1.04_{-0.15}^{+0.16}$~fm\\[.5ex]
-          $^*$Note: the actual rms radius in the center of mass frame is somewhat smaller.\\[2ex]
-        \item Parton width\, $v = 0.51_{-0.17}^{+0.19}$~fm\\[.5ex]
-          (width of Gaussian constituents)\\[2ex]
+        \item Parton sampling radius,\\[.5ex]
+          $\sigma_\mathrm{const} = 0.99_{-0.15}^{+0.17}$~fm\\[2ex]
+        \item Parton (Gaussian constituent) width,\\[.5ex]
+          $w_\mathrm{const} = 0.47_{-0.15}^{+0.20}$~fm\\[2ex]
         \item Results likely sensitive to model details,
-          e.g.\ equation of state, pre-equilibirum dynamics and viscous
+          e.g.\ equation of state, pre-equilibrium dynamics and viscous
           corrections $\delta f$\\[2ex]
@@
 \begin{frame}{\scshape QGP medium properties}
+  \begin{figure}
+    \includegraphics[width=.9\textwidth]{region_shear_bulk}
+  \end{figure}
+  \begin{center}
+    Pb-Pb at 2.76 and 5.02 TeV, better constraint on $\eta/s$ than p-Pb and Pb-Pb at 5.02 TeV
+  \end{center}
 \end{frame}
@@
       \begin{enumerate}
-        \item log transform data
         \item subtract mean and divide out variance

### 2018-05-08 src/plots.py

@@ -20,6 +20,7 @@ decorator.
 from collections import OrderedDict, Counter
 import itertools
 import logging
+import multiprocessing
 from pathlib import Path
 import pickle
 import tempfile
@@ -35,7 +36,7 @@ from matplotlib import ticker
 from matplotlib import cm
 from matplotlib.colors import ListedColormap, LinearSegmentedColormap
 from scipy import special
-from scipy.interpolate import PchipInterpolator
+from scipy.interpolate import PchipInterpolator, interp2d
 from scipy.optimize import curve_fit
 from sklearn.decomposition import PCA
 from sklearn.gaussian_process import GaussianProcessRegressor as GPR
@@ -281,12 +282,34 @@ def obs_color_hsluv(obs, subobs):
     raise ValueError('unknown observable: {} {}'.format(obs, subobs))
 
 
+def obs_color_tableau(obs, subobs):
+    """
+    Observable tableau colors
+
+    """
+    if obs in {'dNch_deta', 'pT_fluct'}:
+        return colors['blue']
+
+    if obs == 'mean_pT':
+        return colors['purple']
+
+    if obs == 'vnk':
+        return {
+            (2, 2): colors['orange'],
+            (3, 2): colors['red'],
+            (4, 2): colors['purple'],
+        }[subobs]
+
+    raise ValueError('unknown observable: {} {}'.format(obs, subobs))
+
+
 def obs_color(obs, subobs):
     """
     Return a nice color for the given observable.
 
     """
-    return hsluv.hsluv_to_rgb(obs_color_hsluv(obs, subobs))
+    return obs_color_tableau(obs, subobs)
+    #return hsluv.hsluv_to_rgb(obs_color_hsluv(obs, subobs))
 
 
 def obs_label(obs, subobs, differentials=False, full_cumulants=False):
@@ -368,7 +391,7 @@ def _observables(posterior=False):
 
     fig, axes = plt.subplots(
         nrows=len(plots), ncols=len(systems),
-        figsize=figsize(1.1, aspect=1.25),
+        figsize=figsize(.8, aspect=1.25),
         gridspec_kw=dict(
             height_ratios=[p.get('height_ratio', 1) for p in plots]
         )
@@ -434,6 +457,7 @@ def _observables(posterior=False):
                     ]))
             else:
                 ax.set_xlim(0, 80)
+                ax.set_xlabel('Centrality %')
 
[…]

### 2018-05-09 talk/qm2018.tex
Work on MAP figure

@@
         \item Is hydro applicable to small collision systems?
-        \item If so, can we describe small system collectivity using current hydro models?
+        \item If so, can small system collectivity be described using current hydro models?
       \end{itemize}
@@
     };
-    \node[anchor=north, xshift=3cm, yshift=-.5cm] (title) at (current page.north) {
+    \node<1>[anchor=north, xshift=3cm, yshift=-.5cm] (title) at (current page.north) {
       \color{theme}{\scshape \Large Bayesian posterior}
@@
     };
+
     \draw<1>[->, thick, theme] (b2.west) to [out=180, in=90] (2.6, .8);
@@
 
-    \node<2>[xshift=0cm, yshift=0cm, scale=.3] {
-      \includegraphics[scale=.15]{magnifying-glass}
+    \node<2>[xshift=1cm, yshift=2.1cm] {
+      \includegraphics[scale=.065]{magnifying-glass}
     };
-    \node<2>[below=.4cm of title, align=center] {
-        \includegraphics[width=.5\textwidth]{posterior_p}\\[1ex]
+    \node<2>[anchor=north east, align=center, inner sep=.6cm] (post_p) at (current page.north east){
+        \includegraphics[width=.6\textwidth]{posterior_p}\\[1ex]
         \small Initial entropy deposition\\
-        \small consistent with geometric mean\\[.5ex]
+        \small consistent with geometric mean\\[1ex]
         $dS/dy \sim \sqrt{T_A T_B}$
     };
-    %\node<3>[xshift=-1.05cm, yshift=1.5cm, scale=.3] {\includegraphics[scale=.15]{magnifying-glass}};
-    %\node<3>[below=.4cm of title, align=center] {
-    %    \includegraphics[width=.5\textwidth]{posterior_parton_number}\\[2ex]
-    %    \small Evidence for nucleon substructure, $N_\text{partons} > 1$.\\
-    %    \small Does not prefer a specific constituent number.
-    %  };
-    %\node<4>[xshift=-1.6cm, yshift=1.0cm, scale=.3] {\includegraphics[scale=.15]{magnifying-glass}};
-    %\node<4>[below=.4cm of title, align=center] {
-    %    \includegraphics[scale=.7]{proton_posterior_shape}\\[2ex]
-    %  };
+    \node<3>[xshift=2.7cm, yshift=0.5cm] {
+      \includegraphics[scale=.065]{magnifying-glass}
+    };
+    \node<3>[anchor=north east, inner sep=.8cm, align=center] at (current page.north east) {
+        \includegraphics[width=.5\textwidth]{posterior_parton_number}\\[2ex]
+        No strong preference for a specific\\
+        constituent number
+    };
+    \node<3>[align=center] (disfavor) at (6, 2) {
+        One constituent\\
+        (no substructure) disfavored
+    };
+    \draw<3>[->, thick, theme] (disfavor.south) to [out=-90, in=90] (7.55, -1.2);
+    \node<4>[xshift=2.15cm, yshift=0cm] {
+      \includegraphics[scale=.065]{magnifying-glass}
+    };
+    \node<4>[anchor=east] (proton) at (current page.east) {
+      \includegraphics[width=.5\textwidth]{proton_posterior_shape}
+    };
+    \node<4>[xshift=-3.2cm, yshift=-1.6cm, align=right] (proton_label) at (proton.north west) {
+      Informative posterior on constituent\\
+      sampling radius $r$ and width $w$\\[2ex]
+      $r=0.99_{-0.15}^{+0.17}$~fm\\[1ex]
+      $w=0.47_{-0.15}^{+0.20}$~fm
+    };
+    \node<5>[anchor=north east, inner sep=.8cm, align=center] at (current page.north east) {
+      \includegraphics[width=.6\textwidth]{region_shear}
+    };
+    \node<6>[anchor=north east, inner sep=.8cm, align=center] at (current page.north east) {
+      \includegraphics[width=.6\textwidth]{region_bulk}
+    };
   \end{tikzpicture}
@@
 
-\begin{frame}{\scshape Nucleon size and shape}
+%\begin{frame}{\scshape QGP medium properties}
+%  \begin{figure}
+%    \includegraphics[width=.9\textwidth]{region_shear_bulk}
+%  \end{figure}
+%  \begin{center}
+%    Pb-Pb at 2.76 and 5.02 TeV, better constraint on $\eta/s$ than p-Pb and Pb-Pb at 5.02 TeV
+%  \end{center}
+%\end{frame}
+
+\begin{frame}[plain]
   \begin{columns}
-    \begin{column}{.45\textwidth}
-      \begin{figure}
-        \includegraphics[width=\textwidth]{proton_posterior_shape}
-      \end{figure}
+    \begin{column}{.5\textwidth}
+      Right: maximum a posteriori (best fit) predictions from the calibrated emulator; \emph{not a model calculation}.\\[2ex]
+      \begin{tabular}{ll}
+        norm           & 20.7 \\
+        $p$            & 0.0314 \\
+        $\sigma_\text{fluct}$  & 0.835 \\
+        $r$      & 0.903 \\
+        $n$ partons  & 6.48 \\
+        $\chi$  & 0.288 \\
+        $d_\text{min}^3$          & 1.11 \\
+        $\tau$ fs         & 0.595 \\
+        $\eta/s$ min       & 0.122 \\
+        $\eta/s$ slope     & 0.742 \\
+        $\eta/s$ crv       & -0.203 \\
+        $\zeta/s$ max      & 0.0510 \\
+        $\zeta/s$ width    & 0.0275 \\
+        $\zeta/s$ $T_0$       & 0.185 \\
+        $T_\text{sw}$        & 0.143
+      \end{tabular}
     \end{column}
-    \begin{column}{.55\textwidth}
-      \begin{itemize}
-        \item Parton sampling radius,\\[.5ex]
-          $\sigma_\mathrm{const} = 0.99_{-0.15}^{+0.17}$~fm\\[2ex]
-        \item Parton (Gaussian constituent) width,\\[.5ex]
-          $w_\mathrm{const} = 0.47_{-0.15}^{+0.20}$~fm\\[2ex]
-        \item Results likely sensitive to model details,
-          e.g.\ equation of state, pre-equilibrium dynamics and viscous
-          corrections $\delta f$\\[2ex]
-      \end{itemize}
+    \begin{column}{.4\textwidth}
+      \includegraphics[height=.95\textheight]{find_map}
     \end{column}
   \end{columns}
-\end{frame}
-
-\begin{frame}{\scshape QGP medium properties}
-  \begin{figure}
-    \includegraphics[width=.9\textwidth]{region_shear_bulk}
-  \end{figure}
-  \begin{center}
-    Pb-Pb at 2.76 and 5.02 TeV, better constraint on $\eta/s$ than p-Pb and Pb-Pb at 5.02 TeV
-  \end{center}
-\end{frame}
-
-\begin{frame}{\scshape Maximum a posteriori parameters}
-  \begin{center}
-    \Large To do if time allows...
-  \end{center}
 \end{frame}

### 2018-05-09 src/plots.py

@@ -308,8 +308,8 @@ def obs_color(obs, subobs):
     Return a nice color for the given observable.
 
     """
-    return obs_color_tableau(obs, subobs)
-    #return hsluv.hsluv_to_rgb(obs_color_hsluv(obs, subobs))
+    #return obs_color_tableau(obs, subobs)
+    return hsluv.hsluv_to_rgb(obs_color_hsluv(obs, subobs))
 
 
 def obs_label(obs, subobs, differentials=False, full_cumulants=False):
@@ -356,6 +356,7 @@ def _observables_plots():
             title='Yields',
             ylabel=(r'$dN_\mathrm{ch}/d\eta$'),
             ylim=(1, 5000),
+            ylim_tight=(10, 4000),
             yscale='log',
             subplots=[
                 ('dNch_deta', None, dict(label=r'$N_\mathrm{ch}$')),
@@ -365,6 +366,7 @@ def _observables_plots():
             title='Mean $p_T$',
             ylabel=r'$\langle p_T \rangle$ [GeV]',
             ylim=(0, 1.5),
+            ylim_tight=(0, 1.1),
             subplots=[
                 ('mean_pT', None, dict(label=r'$\mathrm{mean}\ p_T$')),
             ]
@@ -373,6 +375,7 @@ def _observables_plots():
             title='Flow cumulants',
             ylabel=r'$v_n\{2\}$',
             ylim=(0, .2),
+            ylim_tight=(0, .12),
             subplots=[
                 ('vnk', (n, 2), dict(label='$v_{}$'.format(n)))
                 for n in [2, 3, 4]
@@ -654,19 +657,17 @@ def observables_map():
 
     set_tight(fig)
 
-
 @plot
 def find_map():
     """
     Find the maximum a posteriori (MAP) point and compare emulator predictions
     to experimental data.
-
     """
     from scipy.optimize import minimize
 
     chain = mcmc.Chain()
 
-    #fixed_params = {
+    fixed_params = {
     #    'trento_p': 0.,
     #    'etas_min': .08,
     #    'etas_slope': 1.1285,
@@ -676,9 +677,7 @@ def find_map():
     #    'zetas_width': 0.0194,
     #    'zetas_t0': 0.1824,
     #    'Tswitch': 0.1515,
-    #}
-
-    fixed_params = {}
+    }
 
     opt_params = [k for k in chain.keys if k not in fixed_params]
 
[…]

### 2018-05-10 talk/qm2018.tex
add intro slide to bayesian parameter estimation

@@
 
+\begin{frame}{Challenge of rigorous model-to-data comparison}
+  \begin{figure}
+    \begin{tikzpicture}
+        \node [anchor=east, left=1cm  of current page.center, yshift=1.5cm]
+          (t1) {\large \color{theme} Parameters};
+        \node [anchor=east, left=1cm  of current page.center, yshift=1cm]
+            (a1) {\small shear viscosity};
+        \node [anchor=east, left=1cm  of current page.center, yshift=.5cm]
+            (a2) {\small bulk viscosity};
+        \node [anchor=east, left=1cm  of current page.center, yshift=0cm]
+            (a3) {\small pre-equilibrium flow};
+        \node [anchor=east, left=1cm  of current page.center, yshift=-.5cm]
+            (a4) {\small nucleon size};
+        \node [anchor=east, left=1cm  of current page.center, yshift=-1cm]
+            (a5) {\small nucleon shape};
+
+        % proton-lead observables
+        \node [anchor=west, right=1cm of current page.center, yshift=2.5cm]
+          (t2) {\large \color{theme} p-Pb observables};
+        \node [anchor=west, right=1cm of current page.center, yshift=2cm]
+            (b1) {\small charged particle yields};
+        \node [anchor=west, right=1cm of current page.center, yshift=1.5cm]
+            (b2) {\small mean $p_T$};
+        \node [anchor=west, right=1cm of current page.center, yshift=1cm]
+            (b3) {\small elliptic flow};
+        \node [anchor=west, right=1cm of current page.center, yshift=.5cm]
+            (b4) {\small triangular flow};
+
+        % lead-lead observables
+        \node [anchor=west, right=1cm of current page.center, yshift=-.5cm]
+          (t3) {\large \color{theme} Pb-Pb observables};
+        \node [anchor=west, right=1cm of current page.center, yshift=-1cm]
+            (c1) {\small charged particle yields};
+        \node [anchor=west, right=1cm of current page.center, yshift=-1.5cm]
+            (c2) {\small mean $p_T$};
+        \node [anchor=west, right=1cm of current page.center, yshift=-2cm]
+            (c3) {\small elliptic flow};
+        \node [anchor=west, right=1cm of current page.center, yshift=-2.5cm]
+            (c4) {\small triangular flow};
+
+        \draw [offblack, line width=0.2pt] (a1.east) -- (b1.west);
+        \draw [offblack, line width=0.2pt] (a1.east) -- (b2.west);
+        \draw [offblack, line width=0.2pt] (a1.east) -- (b3.west);
+        \draw [offblack, line width=0.2pt] (a1.east) -- (b4.west);
+
+        \draw [offblack, line width=0.2pt] (a2.east) -- (b1.west);
+        \draw [offblack, line width=0.2pt] (a2.east) -- (b2.west);
+        \draw [offblack, line width=0.2pt] (a2.east) -- (b3.west);
+        \draw [offblack, line width=0.2pt] (a2.east) -- (b4.west);
+
+        \draw [offblack, line width=0.2pt] (a3.east) -- (b1.west);
+        \draw [offblack, line width=0.2pt] (a3.east) -- (b2.west);
+        \draw [offblack, line width=0.2pt] (a3.east) -- (b3.west);
+        \draw [offblack, line width=0.2pt] (a3.east) -- (b4.west);
+
+        \draw [offblack, line width=0.2pt] (a4.east) -- (b1.west);
+        \draw [offblack, line width=0.2pt] (a4.east) -- (b2.west);
+        \draw [offblack, line width=0.2pt] (a4.east) -- (b3.west);
+        \draw [offblack, line width=0.2pt] (a4.east) -- (b4.west);
+
+        \draw [offblack, line width=0.2pt] (a5.east) -- (b1.west);
+        \draw [offblack, line width=0.2pt] (a5.east) -- (b2.west);
+        \draw [offblack, line width=0.2pt] (a5.east) -- (b3.west);
+        \draw [offblack, line width=0.2pt] (a5.east) -- (b4.west);
+
+        \draw [offblack, line width=0.2pt] (a1.east) -- (c1.west);
+        \draw [offblack, line width=0.2pt] (a1.east) -- (c2.west);
+        \draw [offblack, line width=0.2pt] (a1.east) -- (c3.west);
+        \draw [offblack, line width=0.2pt] (a1.east) -- (c4.west);
+
+        \draw [offblack, line width=0.2pt] (a2.east) -- (c1.west);
+        \draw [offblack, line width=0.2pt] (a2.east) -- (c2.west);
+        \draw [offblack, line width=0.2pt] (a2.east) -- (c3.west);
+        \draw [offblack, line width=0.2pt] (a2.east) -- (c4.west);
+
+        \draw [offblack, line width=0.2pt] (a3.east) -- (c1.west);
+        \draw [offblack, line width=0.2pt] (a3.east) -- (c2.west);
+        \draw [offblack, line width=0.2pt] (a3.east) -- (c3.west);
+        \draw [offblack, line width=0.2pt] (a3.east) -- (c4.west);
+
+        \draw [offblack, line width=0.2pt] (a4.east) -- (c1.west);
+        \draw [offblack, line width=0.2pt] (a4.east) -- (c2.west);
+        \draw [offblack, line width=0.2pt] (a4.east) -- (c3.west);
+        \draw [offblack, line width=0.2pt] (a4.east) -- (c4.west);
+
+        \draw [offblack, line width=0.2pt] (a5.east) -- (c1.west);
+        \draw [offblack, line width=0.2pt] (a5.east) -- (c2.west);
+        \draw [offblack, line width=0.2pt] (a5.east) -- (c3.west);
+        \draw [offblack, line width=0.2pt] (a5.east) -- (c4.west);
+    \end{tikzpicture}
+  \end{figure}
+  \begin{center}
+    Many correlated input parameters---cannot calibrate by hand!
+  \end{center}
+\end{frame}
+
 \begin{frame}[plain, noframenumbering]
   \begin{center}
-    \Large I. Bayesian parameter estimation of heavy-ion collisions
+    \Large I. Bayesian parameter estimation
   \end{center}
@@
     };
+    \node<5>[xshift=5.7cm, yshift=-2.6cm] {
+      \includegraphics[scale=.065]{magnifying-glass}
+    };
     \node<6>[anchor=north east, inner sep=.8cm, align=center] at (current page.north east) {
       \includegraphics[width=.6\textwidth]{region_bulk}
+    };
+    \node<6>[xshift=7.3cm, yshift=-4.2cm] {
+      \includegraphics[scale=.065]{magnifying-glass}
     };
@@
 
-%\begin{frame}{\scshape QGP medium properties}
-%  \begin{figure}
-%    \includegraphics[width=.9\textwidth]{region_shear_bulk}
-%  \end{figure}
-%  \begin{center}
-%    Pb-Pb at 2.76 and 5.02 TeV, better constraint on $\eta/s$ than p-Pb and Pb-Pb at 5.02 TeV
-%  \end{center}
-%\end{frame}
-
-\begin{frame}[plain]
-  \begin{columns}
-    \begin{column}{.5\textwidth}
-      Right: maximum a posteriori (best fit) predictions from the calibrated emulator; \emph{not a model calculation}.\\[2ex]
-      \begin{tabular}{ll}
-        norm           & 20.7 \\
-        $p$            & 0.0314 \\
-        $\sigma_\text{fluct}$  & 0.835 \\
-        $r$      & 0.903 \\
-        $n$ partons  & 6.48 \\
-        $\chi$  & 0.288 \\
-        $d_\text{min}^3$          & 1.11 \\
-        $\tau$ fs         & 0.595 \\
-        $\eta/s$ min       & 0.122 \\
-        $\eta/s$ slope     & 0.742 \\
-        $\eta/s$ crv       & -0.203 \\
-        $\zeta/s$ max      & 0.0510 \\
-        $\zeta/s$ width    & 0.0275 \\
-        $\zeta/s$ $T_0$       & 0.185 \\
-        $T_\text{sw}$        & 0.143
-      \end{tabular}
-    \end{column}
-    \begin{column}{.4\textwidth}
-      \includegraphics[height=.95\textheight]{find_map}
-    \end{column}
-  \end{columns}
+\begin{frame}{\scshape Concluding remarks}
+  \vfill
+  \begin{enumerate}
+    \item Calibrated model simultaneously describes p-Pb and Pb-Pb bulk observables\\[.5ex]
+    \item Nucleon substructure necessary to describe p-Pb and Pb-Pb flows (using hydro)\\[.5ex]
+    \item p-Pb observables constrain the constituent width and sampling radius, but not the constituent number\\[.5ex]
+    \item Varying the beam energy of A-A collisions provides better constraint on temperature dependence of the viscosity than varying system size
+  \end{enumerate}
+  \begin{flushleft}
+    To do: run and analyze events using \textit{maximum a posteriori} parameters\\[2ex]
+    \textcolor{theme}{\trento\ with nucleon substructure:}\\
+    \url{github.com/morelandjs/trento-partons}\\[1ex]
+    \textcolor{theme}{Reproduce my results:}\\
+    \url{github.com/morelandjs/hic-param-est-qm18}
+  \end{flushleft}
 \end{frame}
@@
     \begin{column}{.35\textwidth}
-      {\scshape Preprocessing}
-      \begin{enumerate}
-        \item subtract mean and divide out variance
-        \item project onto principle component basis
-      \end{enumerate}
-      \medskip
-      {\scshape k-fold cross validation}
-      \begin{itemize}
-        \item Subdivide design into $k$ partitions. Choose one for
-          validation and use the rest for training. Repeat for all partitions.
-      \end{itemize}
+      \textcolor{theme}{\scshape k-fold cross validation}\\[.5ex]
+      Subdivide design into $k$ partitions. Choose one for validation and use the
+      rest for training. Repeat for all partitions.\\[2ex]
+      Residuals should sample a unit normal distribution\\
+      (far right).
     \end{column}
@@
 
+\begin{frame}[plain]
+  \begin{columns}
+    \begin{column}{.5\textwidth}
+      Right: maximum a posteriori (best fit) predictions from the calibrated emulator; \emph{not a model calculation}.\\[2ex]
+    \end{column}
+    \begin{column}{.4\textwidth}
+      \includegraphics[height=.95\textheight]{find_map}
+    \end{column}
+  \end{columns}
+\end{frame}
+
 \usebackgroundtemplate{}

### 2018-05-10 src/emulator.py

@@ -321,7 +321,7 @@ if __name__ == '__main__':
     kwargs = vars(args)
 
     for s in kwargs.pop('systems'):
-        npc = {'pPb5020': 7, 'PbPb5020': 9}[s]
+        npc = {'pPb5020': 7, 'PbPb5020': 8}[s]
         emu = Emulator.from_cache(s, npc=npc, **kwargs)
 
         print(s)

### 2018-05-10 talk/qm2018.tex
update figures

@@
     background default text=black!30,
-    background fill=theme!10,
-    background text=theme,
+    background fill=theme!20,
+    background text=offblack,
   }
@@
       \node[box=1, font=\small] (input) at (-5, -1) {
-      \boxtitle{Input parameters}
+          \boxtitle{\color{theme} Input parameters}
       IC and QGP properties
@@
     \node[box, below=of input.south west, anchor=north west, yshift=.6cm, font=\small] (physics) {
-      \boxtitle{Physics theory}
+        \boxtitle{\color{theme} Physics theory}
       initial stages, hydro, and\\
@@
     \node[box, right=of input.east, anchor=west, xshift=-.2cm, font=\small] (model) {
-      \boxtitle{Computer model}
+        \boxtitle{\color{theme} Computer model}
       minimum bias event-by-\\
@@
     \node[box, right=of model.east, anchor=west, xshift=-.2cm, font=\small] (gp) {
-      \boxtitle{Gaussian process emulator}
+        \boxtitle{\color{theme} Gaussian process emulator}
       surrogate model
@@
     \node[box, below=of gp.south, anchor=north, font=\small] (mcmc) {
-      \boxtitle{MCMC}
+        \boxtitle{\color{theme} MCMC}
       calibrate model to data
@@
     \node[box, below=of mcmc.south, anchor=north, font=\small] (posterior) {
-      \boxtitle{Posterior distribution}
+        \boxtitle{\color{theme} Posterior distribution}
       quantitative estimates \\
@@
     \node[box, left=of mcmc.west, anchor=east, font=\small] (exp) {
-      \boxtitle{Experimental data}
+        \boxtitle{\color{theme} Experimental data}
       yields, mean $p_T$, flows
@@
       \smallskip\hrule\medskip
-      {\scriptsize non-RTA shear \& bulk corrections}\\
+      {\scriptsize shear \& bulk corrections}\\
       {\tiny \color{theme} (Bernhard thesis) arXiv:1804.06469}
@@
       \begin{equation*}
-        \gamma_i\, T_\text{proton} \rightarrow \sum\limits_{j=1}^{N_\text{parton}}
-        \gamma_j\,T_\text{parton}
+        \gamma_i\, T_\text{proton} \rightarrow \sum\limits_{j=1}^{M}
+        \gamma_j\,T_\text{constituent}
       \end{equation*}
+      \small where $M$ is the number of constituents.
     \end{column}
@@
         $\sigmaf$         & Relative parton fluct. std.\       & 0--2            \\
-        $w$               & Parton sampling radius             & 0.4--1.2 fm     \\
+        $r$               & Parton sampling radius             & 0.4--1.2 fm     \\
         $\X$              & Parton structure parameter         & 0--1            \\
-        $\npartons$       & Number of partons                  & 1--10           \\
+        $m$               & Number of constituents             & 1--10           \\
         $\dmin$           & Nucleon exclusion volume           & 0--4.9 fm$^3$   \\
@@
         \item Recast constituent width $w$ as:\\[1ex]
-          $w = w_\text{min} + \X(r - w_\text{min})$\\[1ex]
-          $v_\text{min} = 0.2$~fm,~ $v_\text{max} = w$\\[1ex]
+          $w = w_\text{min} + \X(w_\text{max} - w_\text{min})$\\[1ex]
+          $w_\text{min} = 0.2$~fm,~ $w_\text{max} = r$\\[1ex]
         \item Parton sampling radius is \emph{not} equivalent to the proton radius
@@
       ++(1.5, 0);
+    \node[inner sep=1ex, anchor=north, align=center] at (current page.north) {
+      \scriptsize Data: CMS \& ALICE
+    };
   \end{tikzpicture}
@@
   \begin{columns}
-    \begin{column}{.5\textwidth}
-      Right: maximum a posteriori (best fit) predictions from the calibrated emulator; \emph{not a model calculation}.\\[2ex]
+    \begin{column}{.4\textwidth}
+      \begin{tikzpicture}[remember picture, overlay]
+        \node[anchor=west] at (current page.west) {\includegraphics[height=\textheight]{find_map}};
+      \end{tikzpicture}
     \end{column}
-    \begin{column}{.4\textwidth}
-      \includegraphics[height=.95\textheight]{find_map}
+    \begin{column}{.6\textwidth}
+      \textit{Maximum a posteriori} (best fit) predictions from the calibrated \emph{emulator} (not the model).\\[2ex]
+      \begin{table}
+        \scriptsize
+        \begin{tabular}{lll}
+          Parameter         & Description                        & Value           \\
+          \paddedhline
+          Norm              & Overall normalization              & 20.3            \\
+          $p$               & Entropy deposition parameter       & 0.04            \\
+          $\sigmaf$         & Relative constituent fluct. std.\  & 0.92            \\
+          $r$               & Constituent sampling radius        & 0.86 fm         \\
+          $\X$              & Constituent structure parameter    & 0.32            \\
+          $m$               & Number of constituents             & 6               \\
+          $\dmin$           & Nucleon exclusion volume           & 1.1 fm$^3$      \\
+          $\tau_{fs}$       & Free streaming time                & 0.42 fm/$c$     \\
+          $\eta/s$ min      & Shear viscosity at $T_c$           & 0.11            \\
+          $\eta/s$ slope    & Slope above $T_c$                  & 0.98 GeV$^{-1}$ \\
+          $\eta/s$ crv      & Curvature above $T_c$              & -0.14           \\
+          $\zeta/s$ norm    & Bulk viscosity peak height         & 0.06            \\
+          $\zeta/s$ width   & Bulk viscosity peak width          & 0.017 GeV       \\
+          $\zeta/s$ temp    & Bulk viscosity peak location       & 0.18 MeV        \\
+          $T_\text{switch}$ & Particlization temperature         & 142 MeV         \\
+        \end{tabular}
+      \end{table}
     \end{column}

### 2018-05-10 src/design.py

@@ -36,13 +36,13 @@ design and produce far too few particles to be physically reasonable.
 
 """
 bad_points = [
-    129, 392, 266, 141, 143, 273, 401, 276, 405, 153, 156, 285, 414,
-    159, 162, 418, 163, 421, 422, 294, 296, 298, 429, 301, 308, 183,
-    184, 312, 442, 446, 319, 451, 453, 326, 199, 457, 330, 459, 460,
-    461, 334, 203, 464, 466, 83, 211, 338, 471, 472, 473, 346, 223,
-    355, 357, 360, 234, 363, 492, 107, 236, 367, 235, 498, 115, 242,
-    376, 378
+    129, 392, 143, 273, 401, 276, 405, 153, 156, 285, 414, 159, 162, 418, 163,
+    421, 422, 296, 298, 429, 301, 308, 183, 184, 312, 442, 446, 319, 451, 453,
+    326, 199, 457, 330, 459, 460, 461, 334, 203, 464, 466, 83, 211, 338, 471,
+    472, 473, 346, 223, 355, 357, 360, 234, 363, 492, 107, 236, 367, 235, 498,
+    115, 242, 376, 378
 ]
+
 
 
 def generate_lhs(npoints, ndim, seed):

### 2018-05-10 src/plots.py

@@ -30,6 +30,7 @@ import h5py
 import hsluv
 import numpy as np
 import matplotlib.pyplot as plt
+import matplotlib.gridspec as gridspec
 from matplotlib import lines
 from matplotlib import patches
 from matplotlib import ticker
@@ -339,6 +340,19 @@ def obs_label(obs, subobs, differentials=False, full_cumulants=False):
             n,
             (r'\{' + str(k) + r'\}') if full_cumulants else ''
         )
+
+def xlabel(system, obs):
+    """
+    x-axis labels for p-Pb and Pb-Pb systems
+
+    """
+    pPb5020 = dict(
+        dNch_deta='Centrality %',
+        mean_pT=r'$n_\mathrm{ch} / \langle n_\mathrm{ch} \rangle$',
+        vnk=r'$n_\mathrm{trk}^\mathrm{offline} / \langle n_\mathrm{trk}^\mathrm{offline}\rangle$'
+    )
+
+    return 'Centrality %' if system == 'PbPb5020' else pPb5020[obs]
 
 
 def _observables_plots():
@@ -709,19 +723,23 @@ def find_map():
 
     plots = _observables_plots()
 
-    fig, axes = plt.subplots(
-        nrows=2*len(plots), ncols=len(systems),
-        figsize=figsize(1.0, 1.5),
-        gridspec_kw=dict(
-            height_ratios=list(itertools.chain.from_iterable(
-                (p.get('height_ratio', 1), .4) for p in plots
-            )),
-            left=0.09, right=0.94, bottom=0.05, top=0.97, wspace=.2, hspace=0.4
-        )
-    )
+    fig = plt.figure(figsize=figsize(1, 1.5))
+    gs = gridspec.GridSpec(6, 2, height_ratios=[3, 1, 3, 1, 3, 1])
+
+    axes = []
+    ratio_axes = []
+
+    nrow = len(plots)
+    ncol = len(systems)
+
+    for row, col in itertools.product(range(0, 2*nrow, 2), range(ncol)):
+        ax = fig.add_subplot(gs[row, col])
+        ratio_ax = fig.add_subplot(gs[row + 1, col], sharex=ax)
+        axes.append(ax)
+        ratio_axes.append(ratio_ax)
 
     for (plot, system), ax, ratio_ax in zip(
-            itertools.product(plots, systems), axes[::2].flat, axes[1::2].flat
+            itertools.product(plots, systems), axes, ratio_axes
     ):
[…]

### 2018-05-11 talk/qm2018.tex
Working version before departure

@@
 \newcommand{\X}{\chi_\text{struct}}
-\newcommand{\npartons}{n_\text{partons}}
 \newcommand{\dmin}{d_\text{min}^3}
@@
   \begin{columns}[T]
-    \begin{column}{.62\textwidth}
-      \bigskip
-      Important questions:\\
-      \begin{itemize}
-        \item Is hydro applicable to small collision systems?
-        \item If so, can small system collectivity be described using current hydro models?
-      \end{itemize}
-      \smallskip
-      First question precludes the second, but also more difficult to answer.\\[1ex]
-      Let's tackle the second question \textit{assuming} hydrodynamic models make
-      sense down to length scales $\Delta x=.2$~fm.\\[1ex]
-      Under this assumption, does a unified hydrodynamic description of p+A and A+A data exist?
+    \begin{column}{.7\textwidth}
+      \medskip \small
+      \textcolor{theme}{\scshape Narrative:}\\[.5ex]
+      Multiparticle correlations in A-A collisions are explained by flow.\\[1ex]
+      \textcolor{theme}{\scshape Justification:}\\[.5ex]
+      Quantitative agreement of hydrodynamic simulations remains strongest evidence for flow in A-A collisions. Agreement is global, self-consistent and highly non-trivial.\\[1ex]
+      \textcolor{theme}{\scshape Natural questions:}\\[.5ex]
+      Can the same framework simultaneously describe p-A collisions without spoiling its description of A-A collisions? Does the framework continue to make sense?\\[1ex]
+      \textcolor{theme}{\scshape This work:}\\[.5ex]
+      Calibrate a successful hybrid model---based on viscous hydrodynamics and Boltzmann transport---to simultaneously describe p-Pb and Pb-Pb data. Does such a description exist?
     \end{column}
@@
 
-\begin{frame}{Challenge of rigorous model-to-data comparison}
+\begin{frame}{\scshape Challenge of rigorous model-to-data comparison}
   \begin{figure}
@@
             (a5) {\small nucleon shape};
+        \node [anchor=east, left=1cm  of current page.center, yshift=-1.5cm]
+            (a6) {\small and more...};
 
@@
         \small
-        \item Radius of Gaussian distribution used to sample
-          constituent positions
-        \item Gaussian constituent width
+        \item Sampling radius of constituent positions
+        \item Constituent Gaussian width
         \item Number of constituents in each nucleon
@@
         $p$               & Entropy deposition parameter       & $-1$ to $+1$    \\
-        $\sigmaf$         & Relative parton fluct. std.\       & 0--2            \\
-        $r$               & Parton sampling radius             & 0.4--1.2 fm     \\
-        $\X$              & Parton structure parameter         & 0--1            \\
-        $m$               & Number of constituents             & 1--10           \\
+        $\sigmaf$         & Relative constituent fluct. std.\  & 0--2            \\
+        $r$               & Constituent sampling radius        & 0.4--1.2 fm     \\
+        $\X$              & Constituent structure parameter    & 0--1            \\
+        $m$               & Number of constituents             & 1--9            \\
         $\dmin$           & Nucleon exclusion volume           & 0--4.9 fm$^3$   \\
@@
           $w_\text{min} = 0.2$~fm,~ $w_\text{max} = r$\\[1ex]
-        \item Parton sampling radius is \emph{not} equivalent to the proton radius
-          in the proton c.o.m.\ frame, see\\
-          \textcolor{theme}{\scriptsize PRC.94.024919}\\[1ex]
+        \item Constituent sampling radius is \emph{not} equivalent to the proton radius
+          in the proton c.o.m.\ frame, see~ \textcolor{theme}{\scriptsize PRC.94.024919}\\[1ex]
         \item Other parameters same as combined Pb+Pb analysis at 2.76 and 5.02 TeV

### 2018-05-11 src/design.py

@@ -173,6 +173,12 @@ class Design:
 
         keep = [n not in bad_points for n in range(npoints)]
         self.array = self.array[keep]
+
+        # convert constituent number to an integer and correct range
+        index = [i for i, k in enumerate(self.keys) if k == 'parton_number'][0]
+        self.array[:, index] = np.floor(self.array[:, index])
+        self.range[index] = (1, 9)
+
         self.points = list(itertools.compress(self.points, keep))
         logging.debug(
             'removed outlier design points: {}'.format(bad_points)

### 2018-05-11 src/plots.py

@@ -258,8 +258,11 @@ def obs_color_hsluv(obs, subobs):
     Use obs_color() to obtain an RGB color.
 
     """
-    if obs in {'dNch_deta', 'pT_fluct', 'mean_pT'}:
+    if obs in {'dNch_deta', 'pT_fluct'}:
         return 250, 90, 55
+
+    if obs == 'mean_pT':
+        return 230, 90, 65
 
     if obs == 'dET_deta':
         return 10, 65, 55
@@ -274,31 +277,9 @@ def obs_color_hsluv(obs, subobs):
 
     if obs == 'vnk':
         return {
-            (2, 2): (230, 90, 65),
-            (2, 4): (262, 80, 63),
+            (2, 2): (250, 90, 65),
             (3, 2): (150, 90, 67),
-            (4, 2): (310, 70, 50),
-        }[subobs]
-
-    raise ValueError('unknown observable: {} {}'.format(obs, subobs))
-
-
-def obs_color_tableau(obs, subobs):
-    """
-    Observable tableau colors
-
-    """
-    if obs in {'dNch_deta', 'pT_fluct'}:
-        return colors['blue']
-
-    if obs == 'mean_pT':
-        return colors['purple']
-
-    if obs == 'vnk':
-        return {
-            (2, 2): colors['orange'],
-            (3, 2): colors['red'],
-            (4, 2): colors['purple'],
+            (4, 2): (20, 90, 62),
         }[subobs]
 
     raise ValueError('unknown observable: {} {}'.format(obs, subobs))
@@ -309,7 +290,6 @@ def obs_color(obs, subobs):
     Return a nice color for the given observable.
 
     """
-    #return obs_color_tableau(obs, subobs)
     return hsluv.hsluv_to_rgb(obs_color_hsluv(obs, subobs))
 
 
@@ -1580,7 +1560,7 @@ def validation_data(system, n_splits=20):
     """
     design = Design(system)
     kf = KFold(n_splits=n_splits)
-    npc = {'pPb5020': 7, 'PbPb5020': 9}[system]
+    npc = {'pPb5020': 7, 'PbPb5020': 8}[system]
 
     mean_folds = []
     cov_folds = []
@@ -1995,6 +1975,8 @@ def _diag_emu(system=default_system, pcs=None, params=None, label_all=True):
                     lw=0, color=color, alpha=.3, zorder=-20
                 )
 
+            if param == 'parton_number':
[…]

### 2018-06-18 paper/trento-substructure.tex
Resume work on the paper

--- 2018-03-30 paper/trento-partons.tex
@@
   Estimating initial state and quark-gluon plasma medium properties\\
-  using a hydrodynamic model with nucleon substructure\\
-  calibrated to p-Pb and Pb-Pb collisions \texorpdfstring{$\mathbf{\sqrts=5.02}$}{}~TeV
+  using a hybrid model with nucleon substructure\\
+  calibrated to p-Pb and Pb-Pb collisions at \texorpdfstring{$\mathbf{\sqrts=5.02}$}{}~TeV
 }
@@
 \begin{abstract}
-  We posit a unified hydrodynamic description of ultrarelativistic p-Pb and Pb-Pb nuclear collisions at $\sqrts=5.02$~TeV and evaluate our assertion using Bayesian inference.
-Specifically, we model the dynamics of both collision systems using QGP initial conditions with parametric nucleon substructure, a pre-equilibrium free-streaming stage, event-by-event viscous hydrodynamics, and a microscopic hadronic afterburner.
-Free parameters of the model which describe the initial state and QGP medium are then simultaneously calibrated to fit charged particle yields, mean $p_T$ and flow cumulants of both systems.
+  We posit a unified hydrodynamic and microscopic description of the quark-gluon plasma (QGP) produced in ultrarelativistic p-Pb and Pb-Pb nuclear collisions at $\sqrts=5.02$~TeV and evaluate our assertion using Bayesian inference.
+Specifically, we model the dynamics of both collision systems using initial conditions with parametric nucleon substructure, a pre-equilibrium free-streaming stage, event-by-event viscous hydrodynamics, and a microscopic hadronic afterburner.
+Free parameters of the model which describe the initial state and QGP medium are then simultaneously calibrated to fit charged particle yields, mean $p_T$ and flow cumulants.
 We argue that the global agreement of the calibrated model with the experimental data strongly supports the existence of hydrodynamic flow in small collision systems at ultrarelativistic energies, and that this flow necessarily develops at length scales smaller than a proton.
-Finally, we present posterior distributions on our model's initial state and QGP medium parameters, including new constraints on the shape and structure of the proton.
+Posterior estimates for the model's input parameters are obtained, and new insights into the temperature dependence of the QGP transport coefficients and event-by-event structure of the proton discussed.
 \end{abstract}
@@
   Hydrodynamic models of ultrarelativistic nuclear collisions are complicated by a number of theoretical unknowns, including the detailed geometry of the QGP initial conditions, the strength and duration of pre-equilibrium dynamics, the temperature dependence of QGP transport coefficients, and the boundaries of hydrodynamic applicability \note.
-In general, these theoretical uncertainties tend to grow with decreasing system size, where emergent physics at partonic length scales becomes important to describing bulk properties of the produced system.
+In general, these theoretical uncertainties tend to grow with decreasing system size, where emergent physics at partonic length scales becomes important to describe bulk properties of the produced system.
 
@@
 
-  Typically, heavy-ion nuclear structure, characterized e.g.\ by an atomic mass and set of Woods-Saxon parameters, is regarded as a known input to hydrodynamic models which contributes negligible uncertainty to overall model predictions.
-Collisions involving protons and light-ions, meanwhile, are sensitive to detailed geometric properties of the proton \note, such as spatial inhomogeneities inside the proton which are not well constrained by nuclear structure measurements.
+Typically, the macroscopic structure of heavy nuclei, characterized e.g.\ by an atomic mass and set of Woods-Saxon parameters, is regarded as a known input to hydrodynamic models and negligible source of uncertainty when calibrating to data.
+Collisions involving protons and light-ions, meanwhile, are sensitive to geometry of the proton, such as the magnitude of event-by-event shape fluctuations, which is less well understood theoretically and experimentally.
+
+properties of the proton \note, such as spatial inhomogeneities inside the proton which are not well constrained by nuclear structure measurements.
 
@@
 Citations I don't want to forget \cite{Shen:2016mmv}
-\bibliography{trento-partons}
+\bibliography{trento-substructure}
 

### 2018-06-18 src/design.py

@@ -40,7 +40,7 @@ bad_points = [
     421, 422, 296, 298, 429, 301, 308, 183, 184, 312, 442, 446, 319, 451, 453,
     326, 199, 457, 330, 459, 460, 461, 334, 203, 464, 466, 83, 211, 338, 471,
     472, 473, 346, 223, 355, 357, 360, 234, 363, 492, 107, 236, 367, 235, 498,
-    115, 242, 376, 378
+    115, 242, 376, 249, 378
 ]
 
 

### 2018-06-18 src/plots.py

@@ -388,7 +388,7 @@ def _observables(posterior=False):
 
     fig, axes = plt.subplots(
         nrows=len(plots), ncols=len(systems),
-        figsize=figsize(.7, aspect=1.15),
+        figsize=figsize(1.1, aspect=1.25),
         gridspec_kw=dict(
             height_ratios=[p.get('height_ratio', 1) for p in plots]
         )
@@ -662,15 +662,7 @@ def find_map():
     chain = mcmc.Chain()
 
     fixed_params = {
-    #    'trento_p': 0.,
-    #    'etas_min': .08,
-    #    'etas_slope': 1.1285,
-    #    'etas_crv': -0.5365,
-    #    'etas_hrg': .3,
-    #    'zetas_max': 0.0536,
-    #    'zetas_width': 0.0194,
-    #    'zetas_t0': 0.1824,
-    #    'Tswitch': 0.1515,
+        'Tswitch': 0.151,
     }
 
     opt_params = [k for k in chain.keys if k not in fixed_params]
@@ -1152,7 +1144,7 @@ def region_bulk():
     set_tight(fig)
 
 
-#@plot
+@plot
 def region_shear_bulk(cmap=plt.cm.Blues):
     """
     Visual estimates (posterior median and credible region) of the
@@ -2339,14 +2331,14 @@ def trim_design_points():
         Flag 'bad" design points with extreme observable values
 
         """
-        if obs == 'dNch_deta':
+        if any(np.isnan(y)):
+            return True
+        elif obs == 'dNch_deta':
             return any(y < 1)
-        elif obs == 'mean_pT':
-            return any(y < .3)
         elif obs == 'vnk':
             return any(y < 0.002)
         else:
-            return any(np.isnan(y))
+            return False
 
     bad_points = set()
 

### 2018-07-02 paper/trento-substructure.tex
Finishing paper physics model section.

@@
 \usepackage{graphicx}
-\graphicspath{{../plots/}}
+\graphicspath{{../plots/}{/fig/}}
 
@@
 \newcommand{\T}{\tilde{T}}
+\newcommand{\x}{\mathbf{x}}
 \newcommand{\vnk}[2]{v_#1\{#2\}}
@@
 \newcommand{\note}{\textcolor{theblue}{[?]}}
+\newcommand{\smaller}[1]{{\small #1}}
 
@@
 \begin{abstract}
-  We posit a unified hydrodynamic and microscopic description of the quark-gluon plasma (QGP) produced in ultrarelativistic p-Pb and Pb-Pb nuclear collisions at $\sqrts=5.02$~TeV and evaluate our assertion using Bayesian inference.
+  We posit a unified hydrodynamic and microscopic description of the quark-gluon plasma (QGP) produced in ultrarelativistic p-Pb and Pb-Pb collisions at $\sqrts=5.02$~TeV and evaluate our assertion using Bayesian inference.
 Specifically, we model the dynamics of both collision systems using initial conditions with parametric nucleon substructure, a pre-equilibrium free-streaming stage, event-by-event viscous hydrodynamics, and a microscopic hadronic afterburner.
@@
 We argue that the global agreement of the calibrated model with the experimental data strongly supports the existence of hydrodynamic flow in small collision systems at ultrarelativistic energies, and that this flow necessarily develops at length scales smaller than a proton.
-Posterior estimates for the model's input parameters are obtained, and new insights into the temperature dependence of the QGP transport coefficients and event-by-event structure of the proton discussed.
+Posterior estimates for the model's input parameters are obtained, and new insights into the temperature dependence of the QGP transport coefficients and event-by-event structure of the proton are discussed.
 \end{abstract}
@@
 
-  Ultrarelativistic nuclear collisions between one light-ion \mbox{(e.g.\ p, d, or ${}^3$He)} and one heavy-ion \mbox{(e.g.\ Au or Pb)} generate dense, compact sources of nuclear matter which produce long-range multiparticle correlations that are strikingly similar to the correlations observed in heavy-ion collisions, signals which are most commonly cited as evidence for hydrodynamic flow \note.
-This observation suggests that hydrodynamic behavior could be manifest in small droplets of quark-gluon-plasma (QGP), and that flow might develop at length scales smaller than a single proton \note.
+Ultrarelativistic nuclear collisions between one light-ion and one heavy-ion, e.g.\ $^3$He-Au and p-Pb collisions, generate dense, compact sources of nuclear matter which produce long-range multiparticle correlations that are strikingly similar to the correlations observed in heavy-ion collisions where collectivity is commonly explained by the existence of hydrodynamic flow \cite{CMS:2012qk, Abelev:2012ola, Aad:2012gla, Adare:2015ctn}.
+This observation suggests that hydrodynamic behavior could be manifest in small droplets of quark-gluon-plasma (QGP), and that flow might develop at length scales smaller than a single proton.
 
-  Hydrodynamic models of ultrarelativistic nuclear collisions are complicated by a number of theoretical unknowns, including the detailed geometry of the QGP initial conditions, the strength and duration of pre-equilibrium dynamics, the temperature dependence of QGP transport coefficients, and the boundaries of hydrodynamic applicability \note.
-In general, these theoretical uncertainties tend to grow with decreasing system size, where emergent physics at partonic length scales becomes important to describe bulk properties of the produced system.
+Hydrodynamic models of ultrarelativistic nuclear collisions are complicated by a number of theoretical unknowns, including the detailed geometry of the QGP initial conditions, the strength and duration of pre-equilibrium dynamics, the temperature dependence of QGP transport coefficients, and the boundaries of hydrodynamic applicability \cite{Niemi:2014lha, deSouza:2015ena, Ollitrault:2012cm, Song:2012ua}.
+In general, these theoretical uncertainties tend to grow with decreasing system size, where emergent physics at sub-fermi length scales becomes important to describe bulk properties of the produced system.
 
-  One method for reducing theoretical uncertainties is to test model calculations by varying the species of colliding nuclei at a single beam energy \note.
+One method for reducing theoretical uncertainties is to test model calculations by varying the species of colliding nuclei at a single beam energy \cite{Adare:2015bua, Schenke:2014tga, Aidala:2018mcw, Adare:2017wlc, Adamczyk:2015obl, Shen:2016zpp, Aidala:2017ajz, Adare:2006ti}.
 Since initial condition and hydrodynamic models generally factorize the structure of the colliding nuclei from the subsequent time dynamics of the collision, a single theory framework can be simultaneously tested and compared to measurements from multiple collision systems using a self consistent set of model parameters where only the nuclear structure in the model is permitted to vary.
 
-Typically, the macroscopic structure of heavy nuclei, characterized e.g.\ by an atomic mass and set of Woods-Saxon parameters, is regarded as a known input to hydrodynamic models and negligible source of uncertainty when calibrating to data.
-Collisions involving protons and light-ions, meanwhile, are sensitive to geometry of the proton, such as the magnitude of event-by-event shape fluctuations, which is less well understood theoretically and experimentally.
+Typically, the macroscopic structure of heavy nuclei, characterized e.g.\ by an atomic mass and set of Woods-Saxon coefficients \cite{MOLLER1995185, DEVRIES1987495}, is regarded as a known input to hydrodynamic models which contributes negligible uncertainty to simulation predictions, outweighed by large uncertainties in modeling initial energy deposition and off-equilibrium dynamics \cite{Niemi:2014lha, Song:2011hk, Retinskaya:2013gca, Liu:2015nwa, Kurkela:2016vts}.
+The geometry of light ions, meanwhile, is naturally more sensitive to the detailed size and shape of individual protons and neutrons inside the nucleus, which may fluctuate event-by-event and differ signficantly from the round blobs typically used to approximate nucleons in heavy-ion collisions \cite{Schenke:2014zha, Welsh:2016siu, Moreland:2017kdx, Schenke:2014gaa, Schlichting:2014ipa}.
+These nucleon substructure properties are difficult to measure and calculate from first principles and hence contribute significant uncertainty to model predictions of small systems.
 
-properties of the proton \note, such as spatial inhomogeneities inside the proton which are not well constrained by nuclear structure measurements.
+One approach to investigate the effect of nucleon substructure on hydrodynamic simulations is to replace round protons with composite protons, described by a few salient structure parameters such the number of constituents and each constituent's width, in order to study the effect of each structure parameter on simulated observables \cite{Adler:2013aqf, Mitchell:2016jio, Welsh:2016siu}.
+Such sensitivity studies are useful for identifying cause and effect relationships between model inputs and outputs, but lack the ability to constrain nucleon substructure parameters in any kind of global or systematic fashion.
 
-One approach to investigate the effect of nucleon substructure on hydrodynamic models is to fix or hand tune initial condition and QGP medium parameters, while varying the shape and size of hot spots inside the proton to measure the effect on bulk observables.
+An alternative approach to constrain nucleon substructure parameters, is to leverage information from independent experimental measurements which are particularly sensitive to the average size, shape and fluctuations of the proton.
+For instance, measurements by the TOTEM collaboration at $\sqrt{s}=7$~TeV found an unexpected dip in the inelasticity density of $p$-$p$ collisions at nonzero impact parameter \cite{Antchev:2011zz}.
+It was later realized that this depression, or so-called hollowness effect in the $p$-$p$ inelastic collision profile \cite{Arriola2016}, can be explained by the existence of correlated domains inside the proton, and that aspects of these domains, such as their size and correlation strength, may be constrained by comparing model predictions to inelastic $p$-$p$ measurements \cite{Albacete:2016gxu, Albacete:2016pmp}.
 
-These uncertainties systematically bias the predictions of hydrodynamic models which assert a single theoretical model
+Independently, studies of coherent and incoherent $J/\psi$ production based on a color dipole picture of vector meson production were used to simultaneously constrain both the average color charge density of the proton as well as its event-by-event fluctuations in a saturation based framework \cite{Mantysaari:2016ykx, Aaron:2009aa, Abramowicz:2015mha}.
+Meanwhile, initial condition studies using the IP-Glasma model of color-glass condensate effective field theory demonstrated that these color charge fluctuations leave a lasting imprint on the \mbox{small-x} gluon distribution of the proton and the initial geometry of QGP energy deposition \cite{Schlichting:2014ipa}.
+Recent work now shows that IP-Glasma initial conditions with color charge fluctuations tuned to fit coherent and incoherent $J/\psi$ diffraction measured by the H1 and Zeus experiments at HERA \cite{Aaron:2009aa, Abramowicz:2015mha} well reproduces the collective flow observables in small and large collision systems without additional fine tuning of the nucleon substructure parameters \note.
+
+More comprehensive efforts to constrain nucleon substructure properties are complicated by the typically large number of free parameters already used in hydrodynamic transport models.
+Parameters generally affect multiple observables at once, and intercorrelate in complex ways which make it impossible to tune the parameters by hand.
+Evaluating even a single set of parameter values is a computationally intensive process, which generally involves thousands of simulated hydrodynamic events and hundreds of processor hours.
+
+In this work, we take a somewhat different approach to the aforementioned studies and investigate the extent to which nucleon substructure properties can be inferred from a global analysis of p-Pb and Pb-Pb bulk observables at $\sqrts=5.02$~TeV.
+Theoretical ingredients of our model which are uncertain, such as the size and shape of the nucleon, the temperature dependence of the QGP transport coefficients, and the duration of pre-equilibrium dynamics, are described by free parameters with probabilistic values that reflect their inherent uncertainty.
+
+TODO finish introduction.
 
@@
 
-\subsection{Initial conditions}
+We employ a multi-stage hybrid transport model that uses relativistic viscous hydrodynamics to describe the QGP medium and microscopic Boltzmann transport to simulate the dynamics of the system after hadronization \note.
+The hydrodynamic initial conditions are provided by a modified version of the \trento\ model \cite{Moreland:2014oya} with additional parameters to vary the size and shape of the nucleon.
+Each initial condition profile is free streamed to the hydrodynamic starting time and matched onto the hydrodynamic energy-momentum tensor using the Landau matching procedure \note.
+All components of the present model, with the exception of the nucleon substructure implementation, are documented in previous studies \note. We review them here for clarity.
 
-  We model the QGP initial conditions in this work using a simple parametric form for initial entropy deposition employed by the \trento\ model \note.
-The model operates in the ultrarelativistic limit with a Lorentz factor $\gamma \gg 1$ such that each nucleus appears as a thin sheet of nuclear density in the laboratory frame.
-The colliding nuclei are then described by their nuclear thickness functions
+\subsection{Initial state}
+
+We model the QGP initial state in $p$-Pb and Pb-Pb collisions at $\sqrts=5.02$~TeV using a simple parametric form for boost-invariant entropy deposition employed by the \trento\ model \cite{Moreland:2014oya}.
+Generally speaking, the initial three-dimensional distribution of matter produced in relativistic nuclear collisions is \emph{not} boost-invariant; longitudinal entropy deposition fluctuates both locally point-to-point in the transverse plane as well as globally event-by-event due to asymmetries in the sampled density of participant matter \note.
+Nevertheless, boost-invariance has been shown to be a good approximation for both large and small collision systems when hydrodynamic observables are calculated from particles that are detected close to midrapidity \note.
+
+The \trento\ model operates in the ultrarelativistic limit with a Lorentz factor $\gamma \gg 1$ such that each nucleus appears as a thin sheet of nuclear density in the laboratory frame.
+The sheets of colliding nuclear density penetrate and pass through each other in time $\Delta t = D_\text{nucl} / \sqrt{\gamma^2 - 1}$ in the laboratory frame, where $D_\text{nucl}$ is the diameter of the nucleus in its rest frame and $\gamma$ is the usual Lorentz factor of the accelerated ions.
+The resulting nuclear overlap time $\Delta t \lesssim 0.1$~fm/$c$ at top RHIC and LHC energies is thus sufficiently short to neglect initial transverse dynamics which occur while the nuclei pass through each other.
+We therefore assume that the collision produces all secondary particles at uniform proper time $\tau = 0^+$~fm/$c$, and that it deposits entropy (energy) at midrapidity which is a function of the locally varying beam integrated density of each nucleus.
+
+Consider the collision of two protons $A$, $B$ with three-dimensional densities $\rho_{A,B}$ in their local rest frames.
+The proton-proton overlap function
 \begin{equation}
-  T_{A,B}(\mathbf{x}) = \int dz\, \rho_{A,B}(\mathbf{x}, z),
+  \label{Tpp}
+  T_{pp}(b) \equiv \int dx\, dy \int dz\, \rho_A(\x) \int dz\, \rho_B(\x + \mathbf{b}),
 \end{equation}
-where $\mathbf{x}$ is a vector in the transverse plane, $z$ lies along the beam axis, and $\rho$ is the density of each nucleus in its local rest frame.
-Individual collisions between the nuclei occur at a random impact parameter offset $b$, and carve out a subset of participant matter $\rho_{A,B}^\text{part}$ which contributes to inelastic processes in the collision.
-The resulting \emph{participant} thickness functions are defined as
+describes the eikonal overlap of the two proton wave packets at fixed impact parameter $b$ where coordinates $(x, y)$ lie in the transverse plane and $z$ is parallel to the beam axis.
+Here we assume that each proton is comprised of smaller constituents---e.g.\ valence quarks, sea quarks, and small-x gluons---which may collide to produce secondary particles and contribute to the observed inelastic proton-proton cross section.
+
+Within a picture of independent pairwise collisions between the constituents, a Glauber model model may be used to calculate the probability $P_\mathrm{coll}$ that the two protons collide inelastically at impact parameter $b$.
+In the limit when the number of constituents $M \gg 1$, it yields the particularly simple form
 \begin{equation}
-  \T_{A,B}(\mathbf{x}) = \int dz\, \rho^\text{part}_{A,B}(\mathbf{x} \pm \mathbf{b}/2, z).
+  \label{Pcoll}
+  P_\mathrm{coll} = 1 - \exp[-\sigma_\mathrm{eff}\, T_{pp}(b)], \\
 \end{equation}
-where $\rho^\text{part}_{A,B}$ is determined using a Monte Carlo Glauber model described later in this section.
+where $\sigma_\mathrm{eff}$ is an effective cross section for pairwise inelastic collisions between the constituents, and $\sigma_{pp}^\mathrm{inel}$ is the total inelastic proton-proton cross section:
+\begin{equation}
+  \label{sigma_nn}
+  \sigma_{pp}^\mathrm{inel} = \int 2 \pi b\, db\, P_\mathrm{coll}(b).
+\end{equation}
+The proton densities $\rho_{A,B}$ in Eqn.~\eqref{Tpp} are commonly modeled using a spherically symmetric distribution \note.
+For instance, the original implementation of the \trento\ model uses Gaussian protons, largely because it yields a simple analytic solution to Eqn.~\eqref{Pcoll}.
+Needless to say, such approximations are admittedly crude and may have a significant effect on the dynamics of small collision systems where the proton size is comparable to the size of the produced QGP medium.
 
-The sheets of colliding nuclear density penetrate and pass through each other in time $\Delta t = D_\text{nucl} / \sqrt{\gamma^2 - 1}$ in the laboratory frame, where $D_\text{nucl}$ is the diameter of the nucleus in its rest frame and $\gamma$ is the usual Lorentz factor of the accelerated ions.
-The resulting nuclear overlap time $\Delta t =$ \order{10}~fm/$c$ at top RHIC and LHC energies and hence is sufficiently short to treat all interactions local in the transverse plane.
+Several previous studies have investigated the effects of deformed or ``lumpy'' protons \note.
+One common implementation is a superposition of three valence quarks, typically described by Gaussian or exponential form factors.
+The corresponding proton density $\rho(\mathbf{x})$ is then assumed to be that of predominantly small-x gluons, seeded by the distribution of color charge in each of the three valence quarks \note.
 
-$t\ll 1$~fm/$c$ in the laboratory frame, and leave little time for spatially separated matter
-The collision deposits entropy at midrapidity, 
+\begin{figure}
+  \begin{tikzpicture}
+    % spherical proton
+    \draw[dashed, xshift=-2.5cm] (0,0) circle (1cm);
+    % three partons
+    \draw[dashed] (0,0) circle (1cm);
+    \foreach \theta in {0, 120, 240}{
+      \draw ({\theta}:.5) circle (.4cm);
+    }
+    % ten partons
+    \draw[dashed, xshift=2.5cm] (0,0) circle (1cm);
+    \foreach \theta/\radius in {
+      0/0.6, 40/0.5, 90/0.6, 130/0.5, 180/0.6, 220/0.6,
+      260/0.5, 320/0.6, 300/0.2
+    }{
+      \draw[xshift=2.5cm] ({\theta}:\radius) circle (.3cm);
+    }
+  \end{tikzpicture}
+  \caption{\label{fig:substructure} Schematic of plausible proton shapes.
+  The sketch on the left shows a spherically symmetric proton (dashed line), while the middle and right illustrations depict a fluctuating proton with three and nine constituents respectively (solid lines).
+  }
+\end{figure}
+
+In this work, we pursue a less restrictive and more parametric description of the proton where the number of substructure degrees of freedom are uncertain as depicted in Fig.~\ref{fig:substructure}.
+We model each nucleon's density $\rho_{A,B}$ as a sum of $M$ independent constituents
 \begin{equation}
-  dS/d\eta \propto f(\T_A, \T_B)
+  \label{rho}
+  \rho_{A, B}(\x) = \frac{1}{M} \sum\limits_{i=1}^M \rho_c(\mathbf{x} - \mathbf{x_i}),
 \end{equation}
+where each constituent density $\rho_c$ is described by a Gaussian distribution of width $w$
+\begin{equation}
+  \rho_c(\mathbf{x}) = \frac{1}{(2\pi w^2)^{3/2}} \exp\left(-\frac{\x^2}{2 w^2}\right),
+\end{equation}
+and each constituent's position $\mathbf{x_i}$ in Eqn.~\eqref{rho} is sampled from a Gaussian radial distribution of width $r$.
+In the limit $M\to \infty$, the composite proton's root mean square radius is simply the convolution of its sampling radius and constituent width, which add together in quadrature
+\begin{equation}
+  \lim_{M\to\infty} \mathrm{RMS}\{\rho(\mathbf{x})\} = \sqrt{r^2 + w^2}.
+\end{equation}
+We therefore choose to reparametrize the sampling radius $r$ in terms of the variable $\sigma = \sqrt{r^2 - w^2}$ which approximates the proton's root mean square radius when the number of sampled constituents is large.
+
+The two protons $A$, $B$ are assigned a random impact parameter, and Eqn.~\eqref{Pcoll} is used to sample their inelastic collision probability $P_\mathrm{coll}(b)$.
+Note that this proton-proton inelastic collision probability has no direct knowledge of the individual constituent degrees of freedom; it is only \emph{indirectly} sensitive through the geometry of $\rho_{A, B}$ which depends on each of the constituent positions.
+This is an important distinction between the present model and a similar nucleon substructure implementation known as the participant or ``wounded'' quark model which allows for a subset of quarks (constituents) to participate inside a single nucleon \note.
+The proton, unlike the nucleus, cannot produce semi-stable spectator fragments in a high-energy collision.
+Any spectator quarks produced by a wounded quark model would be colored objects that necessarily contribute to secondary particle production as they fragment and recombine to form color-neutral hadrons.
+We correspondingly require that the nucleons in Eqn.~\eqref{rho} participate as singular objects, such that all spectator matter discarded by the simulation is appropriately color-neutral and inert.
+
+Assuming our two protons collide at the sampled impact parameter $\mathbf{b}$, we assign each a \emph{fluctuated} thickness
+\begin{equation}
+  \label{fluctuated_thick}
+  \T_{A, B}(\x) \equiv \int dz\, \frac{1}{M} \sum\limits_{i=1}^M \gamma_i\, \rho_c \,(\mathbf{x} - \mathbf{x_i} \pm \mathbf{b}/2),
+\end{equation}
+equal to the beam-integrated proton density in Eqn.~\eqref{rho}, with each constituent shifted by the appropriate impact parameter offset, and multiplied by a gamma random variable with unit mean and variance $1/k$.
+These \emph{ad hoc} gamma random weights are necessary to describe the large fluctuations observed in minimum bias proton-proton collisions, although their exact physical origin is unknown.
+
+Within the eikonal approximation, the initial entropy deposited at midrapidity at proper time $\tau=0^+$~fm/$c$ can be represented by some mapping ${f(\T_A, \T_B) \mapsto dS/dy}$.
+A natural first guess for the function $f$ is an arithmetic mean $f \propto (\T_A + \T_B)/2$, equal to a wounded nucleon model up to meaningless factor of two in the overall normalization.
+The wounded nucleon model was in fact the first such mapping used as a proxy for initial particle production and entropy deposition in relativistic heavy-ion collisions \note.
+It was quickly realized, however, that the wounded nucleon model predicts the wrong scaling for charged particle production as a function of collision centrality and hence the wrong scaling for initial entropy deposition as a function of participant thickness $\T_{A,B}$ \note.
+
+A simple remedy is to replace the arithmetic mean in the wounded nucleon model with a more flexible parametrization
+\begin{equation}
+  \label{gmean}
+  \frac{dS}{dy}\bigg\vert_{y=0} \propto \left( \frac{\T_A^p + \T_B^p}{2} \right)^{1/p},
+\end{equation}
+based on a family of functions known as the generalized mean(s).
+This parametrization introduces a dimensionless parameter $p$ which varies the scaling behavior of initial entropy deposition at midrapidity.
+For certain discrete values of $p$, it reduces to well known functional forms such as the arithmetic, geometric and harmonic means:
+\begin{equation}
+  \newlength{\extraspace}
+  \setlength{\extraspace}{0.5ex}
+  \frac{dS}{dy}\bigg\vert_{y=0} \propto
+  \begin{cases}
+    \max(T_A, T_B) & p \rightarrow +\infty, \\[\extraspace]
+    (T_A + T_B)/2 & p = +1, \hfill \text{ (arithmetic)} \\[\extraspace]
+    \sqrt{T_A T_B} & p = 0, \hfill \text{ (geometric)} \\[\extraspace]
+    2\, T_A T_B/(T_A + T_B) & p = -1, \hfill \text{ (harmonic)} \\[\extraspace]
+    \min(T_A, T_B) & p \rightarrow -\infty.
+  \end{cases}
+\end{equation}
+Conveniently, it has been shown that the generalized mean ansatz is able to mimic the scaling behavior of certain saturation based initial condition models \note, and hence it should serve as a reasonable parametric form for exploring QGP entropy deposition, assuming imperfect knowledge of saturation effects in nature.
+Of course the model is \emph{not} a substitute for first principle saturation calculations, and may fail to reproduce nuanced features of \emph{ab initio} saturation theory calculations, such as the existence of short-range gluon field fluctuations \note.
+
+\begin{figure}
+  \centering
+  \begin{tikzpicture}
+    \node {\includegraphics[width=\columnwidth]{fig/thickness}};
+    \draw [<->] (0, -1)--(0, 1);
+    \node (0, 1) {${\sim}20$ fm};
+  \end{tikzpicture}
+  \caption{\label{fig:substructure} Effect of nucleon substructure on the nuclear thickness function $T(x, y) \equiv \int dz\, \rho(x, y, z)$ of a $^{208}\mathrm{Pb}$ nucleus.
+    The nucleus on the left has Gaussian nucleons of width $0.88$~fm, while the nucleus on the right has composite nucleons, each containing six Gaussian constituents of width $0.41$~fm.}
+\end{figure}
+
+Equation~\eqref{gmean} is a purely local function of nuclear density in the transverse plane and should (in principle) be equally valid for any pair of colliding nuclei at sufficiently high beam energy.
+The model readily generalizes from individual proton-proton collisions to arbitrary nucleus-nucleus collisions by summing the participant thicknesses $\T_{A,B}$ over all nucleons which participate in one or more inelastic collisions.
+The only modeling difference between $p$-$p$ collisions and Pb-Pb collisions is the number and the position of the nucleons.
+
+When applying the model to heavy-ions, we sample nucleon positions from a Woods-Saxon density distribution subject to a minimum distance criteria $|\mathbf{x_i} - \mathbf{x_j}| > d_\mathrm{min}$ between all pairs of nucleons $i$, $j$.
+The minimum distance algorithm, first described in Ref.~\note, uses a simple trick to resample the nucleon positions without modifying the target Woods-Saxon distribution.
+We first presample the radii of all nucleons in a given nucleus and sort them in ascending order.
+We then sample the solid angles of each nucleon one-by-one, starting with the nucleon closest to the center of the nucleus and working our way outwards.
+If a sampled nucleon position is too close to any of its previously placed neighbors, its solid angle is resampled until the minimum distance criteria is satisfied.
+Similar methods could be used to model correlations between individual constituents inside each nucleon, although the numerical implementation would be somewhat tedious.
 
@@
 
+\begin{figure}[b]
+  \includegraphics{fig/coupling}
+  \caption{\label{fig:freestream} Cartoon of the free-streaming approximation for hydrodynamic initialization.
+The initial state is free-streamed for proper time $\tau_{fs}$ (zero-coupling) before it is matched to hydrodynamics (strong-coupling).
+This piecewise evolution approximates the more realistic scenario expected in nature where the medium's coupling strength smoothly changes as a function of time.}
+\end{figure}
+
+There are of course two limiting cases for the strength of interactions inside the QGP medium immediately at the collision: infinitely weak coupling where the secondary partons free-stream without interacting, and infinitely strong coupling where the fluid's inter-particle mean free path effectively vanishes.
+Realistically, one expects the initial parton interactions to lie somewhere between these two extremes.
+We therefore choose to model the initial off-equilibrium dynamics of the produced QGP medium using a simple step-function approximation, depicted in Fig.~\ref{fig:freestream}, which free-streams the initial state for proper time $\tau_{fs}$ (zero coupling) before instantaneously switching to viscous hydrodynamics (strong coupling) \note.
+The free parameter $\tau_{fs}$ allows us to parametrically vary the \emph{time averaged} coupling strength in the approximate window $0 < \tau_{fs} \lesssim1$~fm/$c$.
+
+The parametric entropy deposition ansatz in Eqn.~\eqref{gmean} does not provide any information about the initial masses or momenta of particles produced in the collision.
+In general, these details will affect the dynamics predicted by the collisionless Boltzmann equation
+\begin{equation}
+  \label{freestream}
+  p^\mu \partial_\mu f(x, p) = 0,
+\end{equation}
+through its dependence on the underlying distribution function $f(x, p)$, and hence are necessary inputs for any free-streaming implementation.
+Equation~\eqref{freestream}, however, simplifies for massless partons with momentum distributions that are locally isotropic.
+Under this assumption, it was shown in Refs.~\cite{Broniowski:2008qk, Liu:2015nwa} that the energy-momentum tensor $T^{\mu\nu}$ of partons at midrapidity and time $\tau$ only depends on its spatial distribution at some earlier time $\tau_0$, but not its $p_\perp$-distribution (which could depend on $\mathbf{x}$).
+
+We thus follow the procedure of the authors and reinterpret our initial transverse entropy density at midrapidity $s(\mathbf{x_\perp}, \tau_0)\vert_{\eta=0}$ as a transverse density of noninteracting massless partons $n(\mathbf{x}, \tau_0)\vert_{\eta=0}$.
+The free-streamed energy-momentum tensor $T^{\mu\nu}(x, y, \eta, \tau)$ at transverse coordinate $(x, y)$, midrapidity $\eta=0$, and proper time $\tau > \tau_0$ is then given by
+\begin{multline}
+  \label{energy-momentum}
+  T^{\mu\nu}(x, y, 0, \tau) = \frac{1}{\tau} \int_{0}^{2\pi} d\phi\, n(x - \Delta \tau \cos \phi, y - \Delta \tau \sin \phi) \\ \times
+  \begin{bmatrix}
+    1 & \cos\phi & \sin\phi  & 0\\
+    \cos\phi & \cos^2\phi & \cos\phi\sin\phi & 0 \\
+    \sin\phi & \sin\phi \cos\phi & \sin^2\phi & 0\\
+    0 & 0 & 0 & 0
+  \end{bmatrix},
+\end{multline}
+where $\Delta\tau$ is the free-streaming time $\Delta\tau = \tau - \tau_0$.
+The solution \eqref{energy-momentum} is decomposed in hydrodynamic form
+\begin{equation}
+  \label{hydro-eqn}
+  T^{\mu\nu} = e u^\mu u^\nu - (P + \Pi) \Delta^{\mu\nu} + \pi^{\mu\nu}
+\end{equation}
+where $e$ and $P$ are the energy density and pressure in the local fluid rest frame, $u^\mu$ is the local fluid velocity, ${\Delta^{\mu\nu} \equiv g^{\mu\nu} - u^\mu u^\nu}$, and $\Pi$ and $\pi^{\mu\nu}$ are the bulk pressure and shear stress tensor respectively.
+We then solve for the energy density $e$ and fluid velocity $u^\mu$ using the Landau matching condition which defines the fluid rest frame velocity as the time-like eigenvector of $T^{\mu\nu}$ with energy density $e$ as its eigenvalue,
+\begin{equation}
+  T^{\mu\nu} u_\nu = e u^\mu.
+\end{equation}
+The initial bulk and shear corrections are finally solved for by subtracting the ideal pressure from the total pressure to find $\Pi$, then solving for $\pi^{\mu\nu}$ using Eqn.~\eqref{hydro-eqn}
+\begin{align}
+  \Pi &= -\frac{1}{3} \mathrm{Tr}(\Delta_{\mu\nu} T^{\mu\nu}) - P,\\
+  \pi^{\mu\nu} &= T^{\mu\nu} - e u^\mu u^\nu + (P + \Pi) \Delta^{\mu\nu}.
+\end{align}
+
+This procedure provides initial values for $T^{\mu\nu}$, $u^\mu$, $\Pi$ and $\pi^{\mu\nu}$ which conserve energy and are consistent with the underlying hydrodynamic equation of state.
+We therefore expect it to provide a more realistic description of the initial stages of the collision as compared to a previous study using the \trento\ initial condition model which set $\Pi$, $\pi^{\mu\nu}$ and $u^\mu$ initially to zero \cite{Bernhard:2016tnd}.
+
 \subsection{Hydrodynamics and Boltzmann transport}
 
-%To first approximation---as with nearly every object in physics---it is convenient to approximate the nucleons in heavy-ion initial condition models as a spherically symmetric particles of finite spatial extent.
-%Commonly, a Gaussian is used
-%\begin{figure}
-%  \begin{tikzpicture}
-%    % spherical proton
-%    \draw[dashed, xshift=-2cm] (0,0) circle (.8cm);
-%    % three partons
-%    \draw[dashed] (0,0) circle (.8cm);
-%    \foreach \theta in {0, 120, 240}{
-%      \draw ({\theta}:.39) circle (.4cm);
-%    }
-%    % ten partons
-%    \draw[dashed, xshift=2cm] (0,0) circle (.8cm);
-%    \foreach \theta/\radius in {
-%      36/0.4, 72/0.5, 108/0.5, 140/0.3, 172/0.1,
-%      190/0.5, 240/0.3, 272/0.5, 308/0.4, 360/0.5
-%    }{
-%      \draw[xshift=2cm] ({\theta}:\radius) circle (.2cm);
-%    }
-%  \end{tikzpicture}
-%  \caption{\label{proton_shapes} Schematic of plausible proton shapes.
-%  The sketch on the left shows a spherically symmetric proton (dashed line), while the middle and right illustrations depict a fluctuating proton with three and ten partons respectively (solid lines).
-%  }
-%\end{figure}
+After free-streaming for proper time $\tau_{fs}$, we transition to viscous hydrodynamics which solves the conservation equations
+\begin{equation}
+  \label{eq:continuity}
+  \partial_\mu T^{\mu\nu} = 0
+\end{equation}
+for the hydrodynamic energy-momentum tensor $T^{\mu\nu}$ expressed in Eqn.~\eqref{hydro-eqn} using a set of second-order Israel-Stewart equations formulated in the 14-moment approximation \note.
+This produces a pair of relaxation-type equations
+\begin{subequations}
+  \label{eq:relaxation}
+  \begin{align}
+    \tau_\Pi \Pi + \dot{\Pi} &=
+      - \zeta \theta - \delta_{\Pi\Pi} \Pi\theta
+      + \lambda_{\Pi\pi} \pi^{\mu\nu} \sigma_{\mu\nu}, \\[1ex]
+    \tau_\pi \dot{\pi}^{\langle \mu\nu \rangle} + \pi^{\mu\nu} &=
+      2\eta\sigma^{\mu\nu} - \delta_{\pi\pi} \pi^{\mu\nu} \theta
+      + \phi_7 \pi_\alpha^{\langle \mu} \pi^{\nu \rangle \alpha} \nonumber \\
+      &\qquad {} - \tau_{\pi\pi} \pi_\alpha^{\langle \mu}\sigma^{\nu \rangle \alpha}
+      + \lambda_{\pi\Pi} \Pi \sigma^{\mu\nu},
+  \end{align}
+\end{subequations}
+for the bulk pressure $\Pi$ and shear-stress $\pi^{\mu\nu}$.
+We model the shear viscosity $\eta$ and bulk viscosity $\zeta$ as free parameters in our study and fix the remaining transport coefficients $\{\tau_\Pi, \delta_{\Pi\Pi}, \lambda_{\Pi\pi}, \tau_\pi, \delta_{\pi\pi}, \phi_7, \tau_{\pi\pi}, \lambda_{\pi\Pi}\}$ using analytic results derived in the limit of small but finite masses \note.
+
+The hydrodynamic equations of motion are closed using an equation of state (EoS) to relate the energy density $e$ and pressure $P$ of the produced medium.
+We use a parametrization for $P(e)$ that matches continuum-extrapolated lattice calculations for (2+1)-flavor QCD by the HotQCD collaboration \note\ to a hadron resonance gas EoS in the temperature interval $165 \le T \le 200$~MeV.
+
+Latest results for the HotQCD trace anomaly calculated using a HISQ/tree fermion action are in good agreement with previous calculations by the Wuppertal-Budapest group using a stout fermion action \note.
+
+Subsequent lattice calculations for (2+1+1)-flavors to investigate the effect of thermalized charm quarks, although studies indicate that the effect on hydrodynamic observables should be relatively small.
+%We implement the hydrodynamic equations using the well-tested \texttt{VISH2+1} viscous hydrodynamics code \note, which solves equations .
+
+\bigskip
 
@@
 
+We parametrize the temperature dependence of the specific shear viscosity $\eta/s$ using a modified linear ansatz
+\begin{equation}
+  (\eta/s)(T) = (\eta/s)_\mathrm{min} + (\eta/s)_\mathrm{slope}\cdot(T - T_c)\cdot(T/T_c)^{(\eta/s)_\mathrm{crv}}
+\end{equation}
+where $\eta/s$ min, slope and curvature are tunable parameters and $T_c=0.154$~GeV is the pseudocritical transition temperature of the HotQCD EoS.
+For the specific bulk viscosity $\zeta/s$, we use an unnormalized Cauchy distribution
+\begin{equation}
+  (\zeta/s)(T) = \frac{(\zeta/s)_\mathrm{max}}{1 + \left(\dfrac{T - (\zeta/s)_{T_0}}{(\zeta/s)_\mathrm{width}}\right)^2}.
+\end{equation}
+described by tunable maximum, width, and location ($T_0$) parameters.
+
 \subsection{Gaussian process emulators}
@@
 
-%\fig{observables_design}{Caption}
-
 \section{Results}
@@
 \subsection{Initial condition properties}
-
-%\begin{figure}
-%  \begin{tikzpicture}
-%    \foreach \x in {0pt, 10pt, 20pt}{\draw (\x, 0) ellipse (5pt and 20pt);}
-%    \foreach \x in {0pt, 10pt, 20pt, 30pt}{\draw (\x+100pt, 0) ellipse (5pt and 20pt);}
-%  \end{tikzpicture}
-%  \caption{\label{fig:collision} This is the figure caption}
-%\end{figure}
-%Consider a relativistic collision between a columnated stack of $n_1$ projectile nucleons and $n_2$ target nucleons as depicted in Fig.~\ref{fig:collision}.
-%Moreover, assume that each nucleon has mass $m$ and moves with velocity $\pm u$ along the beam axis $\hat{z}$.
-%The four-momentum of each colliding nucleon stack is given by
-%\begin{align}
-%  p_1 &= n_1 \gamma m\, (1, 0, 0, u),\\
-%  p_2 &= n_2 \gamma m\, (1, 0, 0, -u),
-%\end{align}
-%where $\gamma = 1/\sqrt{1 - u^2}$ is the usual Lorentz factor.
-%
-%At relativistic energies $\gamma \gg 1$, each stack of nucleons is significantly Lorentz contracted in the lab frame and appears compressed along its direction of motion.
-%The nucleons are thus observed to penetrate and pass through each other in the lab frame in time
-%\begin{equation}
-%  \Delta t = L / \sqrt{\gamma^2 - 1},
-%\end{equation}
-%where $L\sim1$--10~fm is the length of the columnated stack of nucleons in its local rest frame.
-%
-%Collisions at top RHIC and LHC energies have interpenetration times \mbox{$\Delta t < .1$~fm/$c$} and hence transpire before meaningful dynamics develop in the transverse plane.
-%For this reason, it is convenient to say that initial energy deposition occurs instantaneously at time $\Delta t = 0^+$, the moment the two nuclei collide.
-%The initial energy (or entropy) deposited into the transverse plane is then modeled in terms of two beam-integrated nuclear thickness functions
-%\begin{equation}
-%  T_{A,B}(x, y) = \int dz \rho(x, y, z),
-%\end{equation}
-%where $\rho$ is the density profile of the nucleus in its local rest frame.
 
@@
 
-Citations I don't want to forget \cite{Shen:2016mmv}
 \bibliography{trento-substructure}

### 2018-07-02 src/model.py

@@ -102,8 +102,8 @@ def symmetric_cumulant(events, m, n, normalize=False):
     Compute the symmetric cumulant SC(m, n).
 
     """
-    N = np.asarray(events['alice_flow']['N'], dtype=float)
-    Q = dict(enumerate(events['alice_flow']['Qn'].T, start=1))
+    N = np.asarray(events['flow']['alice']['N'], dtype=float)
+    Q = dict(enumerate(events['flow']['alice']['Qn'].T, start=1))
 
     cm2n2 = (
         csq(Q[m]) * csq(Q[n])
@@ -328,17 +328,63 @@ def _data(system, dataset='main'):
     )
 
     data = expt.data[system]
+    species = ['pion', 'kaon', 'proton']
 
-    # Add dummy expt data for PbPb5020 mean pT so that
-    # the model predictions are still calculated.
-    if system.startswith('PbPb'):
-        bins = np.linspace(0, 80, 9)
-        cent = [(a, b) for a, b in zip(bins[:-1], bins[1:])]
-        mean_pT = dict(
-            cent=cent,
-            x=np.array([(a + b)/2 for a, b in cent]),
-        )
-        data = dict({'mean_pT': {None: mean_pT}}, **data)
+    # training data is only missing for Pb-Pb charged particle mean pT
+    if dataset == 'main' and system == 'PbPb5020':
+        edges = np.linspace(0, 70, 15)
+        cent = [(a, b) for a, b in zip(edges[:-1], edges[1:])]
+        x = [(a + b)/2 for a, b in cent]
+        empty = dict(cent=cent, x=x)
+        data = dict({'mean_pT': {None: empty}}, **data)
+
+    # calculate additional observables for the maximum a posteriori parameters
+    if dataset == 'map':
+
+        # p-Pb map observables
+        if system == 'pPb5020':
+
+            # yields
+            cent, x = [data['dNch_deta'][None][k] for k in ('cent', 'x')]
+            empty = dict(cent=cent, x=x)
+            data = dict({'dET_deta': {None: empty}}, **data)
+            data = dict({'iden_dN_dy': {s: empty for s in species}}, **data)
+
+            # mean pT
+            mult, x = [data['mean_pT'][None][k] for k in ('mult', 'x')]
+            empty = dict(mult=mult, x=x)
+            data = dict({'iden_mean_pT': {s: empty for s in species}}, **data)
+            data = dict({'pT_fluct': {None: empty}}, **data)
+
[…]

### 2018-07-02 src/plots.py

@@ -267,7 +267,7 @@ def obs_color_hsluv(obs, subobs):
     if obs == 'dET_deta':
         return 10, 65, 55
 
-    if obs in {'iden_dN_dy'}:
+    if obs in {'iden_dN_dy', 'iden_mean_pT'}:
         return dict(
             charged=(250, 90, 55),
             pion=(210, 85, 70),
@@ -280,6 +280,7 @@ def obs_color_hsluv(obs, subobs):
             (2, 2): (250, 90, 65),
             (3, 2): (150, 90, 67),
             (4, 2): (20, 90, 62),
+            (2, 4): (310, 70, 50),
         }[subobs]
 
     raise ValueError('unknown observable: {} {}'.format(obs, subobs))
@@ -321,19 +322,6 @@ def obs_label(obs, subobs, differentials=False, full_cumulants=False):
             (r'\{' + str(k) + r'\}') if full_cumulants else ''
         )
 
-def xlabel(system, obs):
-    """
-    x-axis labels for p-Pb and Pb-Pb systems
-
-    """
-    pPb5020 = dict(
-        dNch_deta='Centrality %',
-        mean_pT=r'$n_\mathrm{ch} / \langle n_\mathrm{ch} \rangle$',
-        vnk=r'$n_\mathrm{trk}^\mathrm{offline} / \langle n_\mathrm{trk}^\mathrm{offline}\rangle$'
-    )
-
-    return 'Centrality %' if system == 'PbPb5020' else pPb5020[obs]
-
 
 def _observables_plots():
     """
@@ -348,28 +336,55 @@ def _observables_plots():
     return [
         dict(
             title='Yields',
-            ylabel=(r'$dN_\mathrm{ch}/d\eta$'),
-            ylim=(1, 5000),
-            ylim_tight=(10, 4000),
+            xlabel=dict(
+                pPb5020='Centrality %',
+                PbPb5020='Centrality %',
+            ),
+            ylabel=(
+                r'$dN_\mathrm{ch}/d\eta,\ dN/dy,\ dE_T/d\eta\ [\mathrm{GeV}]$'
+            ),
+            xlim=dict(
+                pPb5020=(0, 60),
+                PbPb5020=(0, 75),
+            ),
+            ylim=(2e1, 2e5),
             yscale='log',
+            height_ratio=1.5,
             subplots=[
-                ('dNch_deta', None, dict(label=r'$N_\mathrm{ch}$')),
[…]

### 2018-07-05 paper/trento-substructure.tex
save paper state

@@
 One method for reducing theoretical uncertainties is to test model calculations by varying the species of colliding nuclei at a single beam energy \cite{Adare:2015bua, Schenke:2014tga, Aidala:2018mcw, Adare:2017wlc, Adamczyk:2015obl, Shen:2016zpp, Aidala:2017ajz, Adare:2006ti}.
-Since initial condition and hydrodynamic models generally factorize the structure of the colliding nuclei from the subsequent time dynamics of the collision, a single theory framework can be simultaneously tested and compared to measurements from multiple collision systems using a self consistent set of model parameters where only the nuclear structure in the model is permitted to vary.
+Since initial condition and hydrodynamic models generally factorize the structure of the colliding nuclei from the subsequent time dynamics of the collision, a single theory framework can be simultaneously tested and compared to measurements from multiple collision systems using a self-consistent set of model parameters where only the nuclear structure in the model is permitted to vary.
 
@@
 
-One approach to investigate the effect of nucleon substructure on hydrodynamic simulations is to replace round protons with composite protons, described by a few salient structure parameters such the number of constituents and each constituent's width, in order to study the effect of each structure parameter on simulated observables \cite{Adler:2013aqf, Mitchell:2016jio, Welsh:2016siu}.
-Such sensitivity studies are useful for identifying cause and effect relationships between model inputs and outputs, but lack the ability to constrain nucleon substructure parameters in any kind of global or systematic fashion.
+Early substructure studies replaced round protons with composite protons, described by a few salient model parameters, in order to investigate the effect of each parameter on simulated observables \cite{Adler:2013aqf, Mitchell:2016jio, Welsh:2016siu}.
+These sensitivity studies were able to identify cause and effect relationships between model inputs and outputs, but lacked the ability to constrain nucleon substructure parameters in any kind of global or systematic fashion.
+It quickly became apparent that numerous substructure implementations might be compatible with available data, and that additional work would be required to identify observables which are particularly sensitive to the average size, shape and fluctuations of the proton.
 
-An alternative approach to constrain nucleon substructure parameters, is to leverage information from independent experimental measurements which are particularly sensitive to the average size, shape and fluctuations of the proton.
-For instance, measurements by the TOTEM collaboration at $\sqrt{s}=7$~TeV found an unexpected dip in the inelasticity density of $p$-$p$ collisions at nonzero impact parameter \cite{Antchev:2011zz}.
+Several such observables have been identified in proton-proton and proton-lepton scattering data.
+Measurements by the TOTEM collaboration at $\sqrt{s}=7$~TeV, for instance, found an unexpected dip in the inelasticity density of $p$-$p$ collisions at zero impact parameter \cite{Antchev:2011zz}.
 It was later realized that this depression, or so-called hollowness effect in the $p$-$p$ inelastic collision profile \cite{Arriola2016}, can be explained by the existence of correlated domains inside the proton, and that aspects of these domains, such as their size and correlation strength, may be constrained by comparing model predictions to inelastic $p$-$p$ measurements \cite{Albacete:2016gxu, Albacete:2016pmp}.
@@
 Independently, studies of coherent and incoherent $J/\psi$ production based on a color dipole picture of vector meson production were used to simultaneously constrain both the average color charge density of the proton as well as its event-by-event fluctuations in a saturation based framework \cite{Mantysaari:2016ykx, Aaron:2009aa, Abramowicz:2015mha}.
-Meanwhile, initial condition studies using the IP-Glasma model of color-glass condensate effective field theory demonstrated that these color charge fluctuations leave a lasting imprint on the \mbox{small-x} gluon distribution of the proton and the initial geometry of QGP energy deposition \cite{Schlichting:2014ipa}.
-Recent work now shows that IP-Glasma initial conditions with color charge fluctuations tuned to fit coherent and incoherent $J/\psi$ diffraction measured by the H1 and Zeus experiments at HERA \cite{Aaron:2009aa, Abramowicz:2015mha} well reproduces the collective flow observables in small and large collision systems without additional fine tuning of the nucleon substructure parameters \note.
+Meanwhile, initial condition studies using the IP-Glasma model of color-glass condensate effective field theory demonstrated that these color charge fluctuations leave a lasting imprint on the \mbox{small-x} gluon distribution of the proton and hence the initial geometry of QGP energy deposition \cite{Schlichting:2014ipa}.
+Recent work now shows that hydrodynamic simulations using IP-Glasma initial conditions---with color charge fluctuations calibrated to fit coherent and incoherent $J/\psi$ diffraction measured by the H1 and Zeus experiments at HERA \cite{Aaron:2009aa, Abramowicz:2015mha}---describe collectivity in both small and large collision systems without additional retuning of the nucleon substructure parameters \note.
 
-More comprehensive efforts to constrain nucleon substructure properties are complicated by the typically large number of free parameters already used in hydrodynamic transport models.
-Parameters generally affect multiple observables at once, and intercorrelate in complex ways which make it impossible to tune the parameters by hand.
-Evaluating even a single set of parameter values is a computationally intensive process, which generally involves thousands of simulated hydrodynamic events and hundreds of processor hours.
+Model parameters are of course always in some degree of tension; for instance, fitting one observable may require parameter values which degrade the quantitative description of some other observable.
+Correspondingly, parameters which provide an optimal description of small-system bulk observables may lead to a sub-optimal description of heavy-ion collision bulk observables or \emph{vice versa}.
+It is thus import to look at the experimental data holistically, and to adopt rigorous statistical measures that facilitate systematic model-to-data comparison with meaningful uncertainties.
 
-In this work, we take a somewhat different approach to the aforementioned studies and investigate the extent to which nucleon substructure properties can be inferred from a global analysis of p-Pb and Pb-Pb bulk observables at $\sqrts=5.02$~TeV.
+In this work, we perform a global analysis of soft-sector observables measured in $p$-Pb and Pb-Pb collisions at $\sqrts=5.02$~TeV using established Bayesian methodology to estimate the parameters of a hybrid transport model based on viscous hydrodynamics and Boltzmann transport.
 Theoretical ingredients of our model which are uncertain, such as the size and shape of the nucleon, the temperature dependence of the QGP transport coefficients, and the duration of pre-equilibrium dynamics, are described by free parameters with probabilistic values that reflect their inherent uncertainty.
-
-TODO finish introduction.
 
@@
 \end{equation}
-equal to the beam-integrated proton density in Eqn.~\eqref{rho}, with each constituent shifted by the appropriate impact parameter offset, and multiplied by a gamma random variable with unit mean and variance $1/k$.
-These \emph{ad hoc} gamma random weights are necessary to describe the large fluctuations observed in minimum bias proton-proton collisions, although their exact physical origin is unknown.
+equal to the beam-integrated proton density in Eqn.~\eqref{rho}, with each constituent shifted by the appropriate impact parameter offset, and multiplied by a gamma random variable $\gamma_i$ with unit mean and variance $1/k$.
+These \emph{ad hoc} gamma random weights are necessary to describe the large fluctuations observed in minimum bias proton-proton collisions, although their physical justification is not yet well understood.
 
@@
     \node {\includegraphics[width=\columnwidth]{fig/thickness}};
-    \draw [<->] (0, -1)--(0, 1);
-    \node (0, 1) {${\sim}20$ fm};
+    \node[label=above:{10 fm}] (a) at (0, 1.1) {};
+    \node (b) at (0, -1.1) {};
+    \draw [<->, semithick] (a) to (b);
   \end{tikzpicture}
@@
 When applying the model to heavy-ions, we sample nucleon positions from a Woods-Saxon density distribution subject to a minimum distance criteria $|\mathbf{x_i} - \mathbf{x_j}| > d_\mathrm{min}$ between all pairs of nucleons $i$, $j$.
-The minimum distance algorithm, first described in Ref.~\note, uses a simple trick to resample the nucleon positions without modifying the target Woods-Saxon distribution.
+The minimum distance algorithm, first described in Ref.~\note, uses a simple trick to resample the nucleon positions without modifying the target Woods-Saxon radial distribution.
 We first presample the radii of all nucleons in a given nucleus and sort them in ascending order.
@@
 Realistically, one expects the initial parton interactions to lie somewhere between these two extremes.
-We therefore choose to model the initial off-equilibrium dynamics of the produced QGP medium using a simple step-function approximation, depicted in Fig.~\ref{fig:freestream}, which free-streams the initial state for proper time $\tau_{fs}$ (zero coupling) before instantaneously switching to viscous hydrodynamics (strong coupling) \note.
+We therefore choose to model the QGP's initial off-equilibrium dynamics using a simple step-function approximation, depicted in Fig.~\ref{fig:freestream}, which free-streams the initial state for proper time $\tau_{fs}$ (zero coupling) before instantaneously switching to viscous hydrodynamics (strong coupling) \note.
 The free parameter $\tau_{fs}$ allows us to parametrically vary the \emph{time averaged} coupling strength in the approximate window $0 < \tau_{fs} \lesssim1$~fm/$c$.
@@
 Equation~\eqref{freestream}, however, simplifies for massless partons with momentum distributions that are locally isotropic.
-Under this assumption, it was shown in Refs.~\cite{Broniowski:2008qk, Liu:2015nwa} that the energy-momentum tensor $T^{\mu\nu}$ of partons at midrapidity and time $\tau$ only depends on its spatial distribution at some earlier time $\tau_0$, but not its $p_\perp$-distribution (which could depend on $\mathbf{x}$).
+Under this assumption, it was shown in Refs.~\cite{Broniowski:2008qk, Liu:2015nwa} that the energy-momentum tensor $T^{\mu\nu}$ of partons at midrapidity and time $\tau$ only depends on its spatial distribution at some earlier time $\tau_0$, but not its $p_\perp$-distribution, which could vary as a function of position.
 
@@
 
-\subsection{Hydrodynamics and Boltzmann transport}
+\subsection{Hydrodynamics}
 
@@
 \end{equation}
-for the hydrodynamic energy-momentum tensor $T^{\mu\nu}$ expressed in Eqn.~\eqref{hydro-eqn} using a set of second-order Israel-Stewart equations formulated in the 14-moment approximation \note.
+for the hydrodynamic energy-momentum tensor $T^{\mu\nu}$ expressed in Eqn.~\eqref{hydro-eqn} using a set of second-order Israel-Stewart equations formulated in the 14-moment approximation
+\cite{Israel:1979wp, Israel:1976aa, Denicol:2012cn, Denicol:2010xn}.
 This produces a pair of relaxation-type equations
@@
 for the bulk pressure $\Pi$ and shear-stress $\pi^{\mu\nu}$.
-We model the shear viscosity $\eta$ and bulk viscosity $\zeta$ as free parameters in our study and fix the remaining transport coefficients $\{\tau_\Pi, \delta_{\Pi\Pi}, \lambda_{\Pi\pi}, \tau_\pi, \delta_{\pi\pi}, \phi_7, \tau_{\pi\pi}, \lambda_{\pi\Pi}\}$ using analytic results derived in the limit of small but finite masses \note.
+We model the shear viscosity $\eta$ and bulk viscosity $\zeta$ as unknown temperature dependent quantities and fix the remaining transport coefficients $\{\tau_\Pi, \delta_{\Pi\Pi}, \lambda_{\Pi\pi}, \tau_\pi, \delta_{\pi\pi}, \phi_7, \tau_{\pi\pi}, \lambda_{\pi\Pi}\}$ using analytic results derived in the limit of small but finite masses \cite{Denicol:2014vaa}.
 
-The hydrodynamic equations of motion are closed using an equation of state (EoS) to relate the energy density $e$ and pressure $P$ of the produced medium.
-We use a parametrization for $P(e)$ that matches continuum-extrapolated lattice calculations for (2+1)-flavor QCD by the HotQCD collaboration \note\ to a hadron resonance gas EoS in the temperature interval $165 \le T \le 200$~MeV.
+\begin{figure}
+  \centering
+  \includegraphics{fig/viscosity_dof}
+  \caption{\label{fig:viscosity_dof} Degrees of freedom in the temperature dependent shear and bulk viscosity parametrizations.
+  Lines are chosen for illustrative purposes only and do not represent all possible variability.
+  For instance, $\eta/s$ could have a large slope and negative curvature, or $\zeta/s$ could have a large max and narrow width, neither of which are depicted above.
+}
+\end{figure}
 
-Latest results for the HotQCD trace anomaly calculated using a HISQ/tree fermion action are in good agreement with previous calculations by the Wuppertal-Budapest group using a stout fermion action \note.
+The hydrodynamic equations of motion are necessarily closed using an equation of state (EoS) to relate the energy density $e$ and pressure $P$ of the produced medium.
+We use a parametrization for $P(e)$ that matches a hadron resonance gas EoS at low temperature to a lattice QCD EoS at high temperature by smoothly connecting their trace anomalies in the interval $165 \le T \le 200$~MeV.
+For the lattice EoS, we use a calculation by the HotQCD collaboration for (2+1)-flavor QCD which was extrapolated to the continuum limit \cite{Bazavov:2014pvz}.
+Recent developments in lattice QCD now enable calculations in (2+1+1)-flavors \cite{Borsanyi:2016ksw}, i.e.\ with thermalized charm quarks, and the additional charm flavor has been shown to visibly affect predictions of $p_T$-differential flow observables \cite{Noronha-Hostler:2018zxc}.
+Investigating this sensitivity would thus be a natural target for future improvements to the present work.
 
-Subsequent lattice calculations for (2+1+1)-flavors to investigate the effect of thermalized charm quarks, although studies indicate that the effect on hydrodynamic observables should be relatively small.
-%We implement the hydrodynamic equations using the well-tested \texttt{VISH2+1} viscous hydrodynamics code \note, which solves equations .
+We parametrize the temperature dependence of the QGP viscosities in order to marginalize over their uncertainty when calibrating to data.
+For the specific shear viscosity $\eta/s$, we use a modified linear ansatz
+\begin{equation}
+  \label{eq:shear_viscosity}
+  (\eta/s)(T) = (\eta/s)_\mathrm{min} + (\eta/s)_\mathrm{slope}\cdot(T - T_c)\cdot(T/T_c)^{(\eta/s)_\mathrm{crv}}
+\end{equation}
+where $\eta/s$ min, slope and curvature are tunable parameters and $T_c=0.154$~GeV is the pseudocritical transition temperature of the HotQCD EoS.
+While for the specific bulk viscosity $\zeta/s$, we use an unnormalized Cauchy distribution
+\begin{equation}
+  \label{eq:bulk_viscosity}
+  (\zeta/s)(T) = \frac{(\zeta/s)_\mathrm{max}}{1 + \left(\dfrac{T - (\zeta/s)_{T_0}}{(\zeta/s)_\mathrm{width}}\right)^2},
+\end{equation}
+described by tunable maximum, width, and location ($T_0$) parameters.
+Figure~\ref{fig:viscosity_dof} shows several of the possible curves parametrized by Eqns.~\eqref{eq:shear_viscosity} and \eqref{eq:bulk_viscosity}, although many more are possible.
 
-\bigskip
+The aforementioned hydrodynamic equations are solved numerically using the boost-invariant \texttt{VISH2+1} viscous hydrodynamics code \note\ which is documented extensively in Ref.~\note.
+We simulate each hydrodynamic event on a spacetime grid with transverse extent $x_\mathrm{max}$, spatial grid step $dx$, and time step $d\tau$ which are optimized \emph{event-by-event} to balance trade-offs between numerical accuracy and computation time (see appendix).
+Although these details are somewhat mundane, they are critically important to the present study, since the computation time explodes with the number of grid cells,
+\begin{equation}
+  \mathrm{cpu~time} \sim n_x^2 n_\tau,
+\end{equation}
+and $nx$ and $n\tau$ typically have to be quite large to resolve the small length scales associated with nucleon substructure.
+
+\subsection{Particlization and Boltzmann transport}
+
+We evolve the system hydrodynamically down to a pre-specified switching isotherm $T_{sw}$ at which point the medium is converted into particles using the Cooper-Frye formula \note
+\begin{equation}
+  \label{cooper-frye}
+  E \frac{dN_i}{d^3p} = \frac{g_i}{(2\pi)^3} \int_\Sigma f_i(x, p)\, p^\mu\, d^3\sigma_\mu,
+\end{equation}
+where $i$ is an index over species, $f_i$ is the distribution function of that species, and $d^3\sigma_\mu$ is a volume element of the isothermal hypersurface $\Sigma$ defined by $T_\mathrm{sw}$.
+Thermal particles are then sampled in the rest frame of each fluid cell according to a Bose-Einstein or Fermi-Dirac distribution at zero baryon chemical potential
+\begin{equation}
+  \label{distribution}
+  f(m, p) = \frac{1}{\exp(\sqrt{m^2 + p^2}/T) \mp 1},
+\end{equation}
+where $m$ is the mass of the sampled particle, $p$ is its momentum, and $T$ is the temperature of the fluid cell.
+
+\subsubsection{Sampling with finite resonance widths}
+
+Traditionally, particlization models have sampled resonances using each particle's pole mass in Eqn.~\eqref{distribution}.
+This approximation, however, is somewhat crude and has been known to underpredict pion production, particularly at low $p_T$ \note.
+We thus follow Ref.~\note, and sample particles with a \emph{distribution} of masses
+\begin{equation}
+  f(p) = \int dm\, \mathcal{P}(m)\, f(m, p),
+\end{equation}
+where $\mathcal{P}(m)$ is modeled by a Breit-Wigner distribution
+\begin{equation}
+  \mathcal{P}(m) \propto \frac{\Gamma(m)}{(m - m_0)^2 + \Gamma(m)^2/4}.
+\end{equation}
+Here $m_0$ is the resonance's Breit-Wigner mass and $\Gamma(m)$ is its mass-dependent width, for which we use a simple form:
+\begin{equation}
+  \Gamma(m) = \Gamma_0 \sqrt{\frac{m - m_\mathrm{min}}{m_0 - m_\mathrm{min}}},
+\end{equation}
+where $\Gamma_0$ is the usual Breit-Wigner width and $m_\mathrm{min}$ is a production threshold equal to the total mass of the lightest decay products.
+
+We tabulate the values of $\{\Gamma_0, m_0, m_\mathrm{min}\}$ for all particles and sample the masses of each particle during particlization \note.
+The resonances are then passed to a hadronic transport model, described shortly, which simulates subsequent scatterings and decays.
+
+\subsubsection{Viscous corrections to the distribution function}
+
+When the viscous terms $\pi^{\mu\nu}$ and $\Pi$ are nonzero in Eqn.~\eqref{hydro-eqn}, the distribution function $f$ must be modified to preserve the continuity of $T^{\mu\nu}$ as the system transitions from hydrodynamics to Boltzmann transport.
+We perform the appropriate modification using a general method which transforms the momentum vector \emph{inside} the distribution function
+\begin{align}
+  p_i \rightarrow p'_i &= p_i + \sum\limits_j \lambda_{ij}\, p_j,\\
+  \lambda_{ij} &= (\lambda_\mathrm{shear})_{ij} + \lambda_\mathrm{bulk}\, \delta_{ij},
+\end{align}
+where $\lambda_{ij}$ is a linear transformation matrix consisting of a traceless shear part and a bulk part which is proportional to the identity matrix \note.
+
+We use for the shear viscous correction the form \note
+\begin{equation}
+  (\lambda_\mathrm{shear})_{ij} = \frac{\tau}{2 \eta} \pi_{ij},
+\end{equation}
+with a value for $\eta/\tau$ obtained from the noninteracting hadron resonance gas model
+\begin{equation}
+  \frac{\eta}{\tau} = \frac{1}{15 T} \sum\limits_{sp} g \int \frac{d^3p}{(2\pi)^3}\frac{p^4}{E^2} f_0 (1 \pm f_0).
+\end{equation}
+
+For the bulk viscous correction, we use a novel procedure developed by one of the authors of this work \note.
+
+\subsubsection{Hadronic Boltzmann transport}
+
+\vfill
 
@@
 \subsection{Computer experiment design}
-
-We parametrize the temperature dependence of the specific shear viscosity $\eta/s$ using a modified linear ansatz
-\begin{equation}
-  (\eta/s)(T) = (\eta/s)_\mathrm{min} + (\eta/s)_\mathrm{slope}\cdot(T - T_c)\cdot(T/T_c)^{(\eta/s)_\mathrm{crv}}
-\end{equation}
-where $\eta/s$ min, slope and curvature are tunable parameters and $T_c=0.154$~GeV is the pseudocritical transition temperature of the HotQCD EoS.
-For the specific bulk viscosity $\zeta/s$, we use an unnormalized Cauchy distribution
-\begin{equation}
-  (\zeta/s)(T) = \frac{(\zeta/s)_\mathrm{max}}{1 + \left(\dfrac{T - (\zeta/s)_{T_0}}{(\zeta/s)_\mathrm{width}}\right)^2}.
-\end{equation}
-described by tunable maximum, width, and location ($T_0$) parameters.
 
@@
 
+\appendix*
+
+\section{Adaptive grid sizing}
+\label{appendix}
+This is the appendix
+
 \bibliography{trento-substructure}

### 2018-07-06 paper/trento-substructure.tex
save some writing from home

@@
 \usepackage{multirow}
+
+\usepackage[inline]{enumitem}
 
@@
 Independently, studies of coherent and incoherent $J/\psi$ production based on a color dipole picture of vector meson production were used to simultaneously constrain both the average color charge density of the proton as well as its event-by-event fluctuations in a saturation based framework \cite{Mantysaari:2016ykx, Aaron:2009aa, Abramowicz:2015mha}.
-Meanwhile, initial condition studies using the IP-Glasma model of color-glass condensate effective field theory demonstrated that these color charge fluctuations leave a lasting imprint on the \mbox{small-x} gluon distribution of the proton and hence the initial geometry of QGP energy deposition \cite{Schlichting:2014ipa}.
-Recent work now shows that hydrodynamic simulations using IP-Glasma initial conditions---with color charge fluctuations calibrated to fit coherent and incoherent $J/\psi$ diffraction measured by the H1 and Zeus experiments at HERA \cite{Aaron:2009aa, Abramowicz:2015mha}---describe collectivity in both small and large collision systems without additional retuning of the nucleon substructure parameters \note.
+Initial condition studies using the IP-Glasma model of color-glass condensate effective field theory simultaneously demonstrated that these color charge fluctuations leave a lasting imprint on the \mbox{small-x} gluon distribution of the proton and hence the initial geometry of QGP energy deposition \cite{Schlichting:2014ipa}.
+Just recently, hydrodynamic simulations using IP-Glasma initial conditions with color charge fluctuations calibrated to fit coherent and incoherent $J/\psi$ diffraction measured by the H1 and Zeus experiments at HERA \cite{Aaron:2009aa, Abramowicz:2015mha} were shown to provide a good description of collectivity in small and large collision systems \note.
 
-Model parameters are of course always in some degree of tension; for instance, fitting one observable may require parameter values which degrade the quantitative description of some other observable.
-Correspondingly, parameters which provide an optimal description of small-system bulk observables may lead to a sub-optimal description of heavy-ion collision bulk observables or \emph{vice versa}.
-It is thus import to look at the experimental data holistically, and to adopt rigorous statistical measures that facilitate systematic model-to-data comparison with meaningful uncertainties.
+Model parameters are of course always in some degree of tension; for instance, fitting one observable may require parameter values that degrade the quantitative description of some other observable.
+Similarly, parameters which provide an optimal description of small-system observables may lead to a sub-optimal description of heavy-ion observables or \emph{vice versa}.
+It is thus import to look at the experimental data holistically, and to use model calibration methods which
+\begin{enumerate*}[label=(\arabic*)]
+  \item explore all parameter combinations and
+  \item compare model predictions to all relevant experimental measurements in a statistically rigorous fashion.
+\end{enumerate*}
 
-In this work, we perform a global analysis of soft-sector observables measured in $p$-Pb and Pb-Pb collisions at $\sqrts=5.02$~TeV using established Bayesian methodology to estimate the parameters of a hybrid transport model based on viscous hydrodynamics and Boltzmann transport.
-Theoretical ingredients of our model which are uncertain, such as the size and shape of the nucleon, the temperature dependence of the QGP transport coefficients, and the duration of pre-equilibrium dynamics, are described by free parameters with probabilistic values that reflect their inherent uncertainty.
+We apply, in this work, a model calibration framework known as Bayesian parameter estimation which is ideally suited for this purpose.
+The framework was previously adapted and applied to relativistic heavy-ion collisions, and has since been used in a number of studies to constrain the properties of hot and dense nuclear matter.
+We use it in the present study to search for a unified description of $p$-Pb and Pb-Pb collisions at $\sqrts=5.02$~TeV.
+
+%Section 1 describes our nuclear collision model.
+%Sections 2 and 3 describe the computer experiment design and model calibration process which is used to obtain posterior estimates of the model input parameters.
+%In section 4 we presents the results of our analysis.
+
+%We perform a global analysis of soft-sector observables measured in $p$-Pb and Pb-Pb collisions at $\sqrts=5.02$~TeV and calibrate free parameters of a hybrid transport model to describe the experimental data.
+%
+%to obtain posterior probability estimates for the model parameters given our chosen model, the experimental data and known sources of uncertainty.
 

### 2018-07-06 src/plots.py
save before leaving for wilmington

@@ -813,7 +813,7 @@ def pT_fluct():
     """
     plt.figure(figsize=figsize(.6))
 
-    system = 'PbPb5020'
+    system = 'pPb5020'
     obs = 'pT_fluct'
     subobs = None
 

### 2018-07-10 paper/substructure.tex
Start adding results figures

--- 2018-07-06 paper/trento-substructure.tex
@@
 \usepackage[inline]{enumitem}
+\usepackage{lipsum}
 
@@
 \usepackage[pdfencoding=auto, psdextra]{hyperref}
+
+\usepackage[labelfont={color=theblue}, justification=raggedright, labelsep=quad]{caption}
 
@@
 \usepackage{graphicx}
-\graphicspath{{../plots/}{/fig/}}
+\graphicspath{{../plots/}{./fig/}}
 
 \newcommand{\trento}{T\raisebox{-0.5ex}{R}ENTo}
-\newcommand{\avg}[1]{\langle #1 \rangle}
 \newcommand{\nch}{N_\text{ch}}
-\newcommand{\npart}{N_\text{part}}
 \newcommand{\sqrts}{\sqrt{s_{NN}}}
+\newcommand{\sigmaf}{\sigma_\mathrm{fluct}}
+\newcommand{\taufs}{\tau_\mathrm{fs}}
+\newcommand{\n}{n_\mathrm{c}}
+\newcommand{\R}{R_\mathrm{c}}
+\newcommand{\w}{w_\mathrm{c}}
+\newcommand{\dmin}{d_\mathrm{min}}
 \newcommand{\T}{\tilde{T}}
 \newcommand{\x}{\mathbf{x}}
+\newcommand{\X}{\mathcal{X}_\mathrm{struct}}
 \newcommand{\vnk}[2]{v_#1\{#2\}}
@@
 \newcommand{\note}{\textcolor{theblue}{[?]}}
-\newcommand{\smaller}[1]{{\small #1}}
 
@@
 
-Early substructure studies replaced round protons with composite protons, described by a few salient model parameters, in order to investigate the effect of each parameter on simulated observables \cite{Adler:2013aqf, Mitchell:2016jio, Welsh:2016siu}.
+Early substructure studies replaced round protons with composite protons, described by a few salient model parameters, in order to investigate the effect of each parameter on simulated observables \cite{Adler:2013aqf, Mitchell:2016jio, Welsh:2016siu, Broniowski:2016pvx, Bozek:2017jog}.
 These sensitivity studies were able to identify cause and effect relationships between model inputs and outputs, but lacked the ability to constrain nucleon substructure parameters in any kind of global or systematic fashion.
@@
 Independently, studies of coherent and incoherent $J/\psi$ production based on a color dipole picture of vector meson production were used to simultaneously constrain both the average color charge density of the proton as well as its event-by-event fluctuations in a saturation based framework \cite{Mantysaari:2016ykx, Aaron:2009aa, Abramowicz:2015mha}.
-Initial condition studies using the IP-Glasma model of color-glass condensate effective field theory simultaneously demonstrated that these color charge fluctuations leave a lasting imprint on the \mbox{small-x} gluon distribution of the proton and hence the initial geometry of QGP energy deposition \cite{Schlichting:2014ipa}.
-Just recently, hydrodynamic simulations using IP-Glasma initial conditions with color charge fluctuations calibrated to fit coherent and incoherent $J/\psi$ diffraction measured by the H1 and Zeus experiments at HERA \cite{Aaron:2009aa, Abramowicz:2015mha} were shown to provide a good description of collectivity in small and large collision systems \note.
+Initial condition studies using the IP-Glasma model of color-glass condensate effective field theory \cite{Schenke:2012wb} simultaneously demonstrated that these color charge fluctuations leave a lasting imprint on the \mbox{small-x} gluon distribution of the proton and hence the initial geometry of QGP energy deposition \cite{Schlichting:2014ipa}.
+In addition, it was recently shown that hydrodynamic simulations using IP-Glasma initial conditions with color charge fluctuations calibrated to fit coherent and incoherent $J/\psi$ diffraction measured by the H1 and Zeus experiments at HERA \cite{Aaron:2009aa, Abramowicz:2015mha} provide a good description of collectivity in small and large collision systems \cite{Schenke:2018qmc}.
 
-Model parameters are of course always in some degree of tension; for instance, fitting one observable may require parameter values that degrade the quantitative description of some other observable.
+Model parameters, such as those calibrated by the aforementioned studies, are of course always in some degree of tension.
+For instance, fitting one observable may require parameter values that degrade the quantitative description of some other observable.
 Similarly, parameters which provide an optimal description of small-system observables may lead to a sub-optimal description of heavy-ion observables or \emph{vice versa}.
@@
 
-We apply, in this work, a model calibration framework known as Bayesian parameter estimation which is ideally suited for this purpose.
-The framework was previously adapted and applied to relativistic heavy-ion collisions, and has since been used in a number of studies to constrain the properties of hot and dense nuclear matter.
-We use it in the present study to search for a unified description of $p$-Pb and Pb-Pb collisions at $\sqrts=5.02$~TeV.
+With these considerations in mind, we present progress towards a fully global analysis of $p$-Pb and Pb-Pb bulk observables at $\sqrts=5.02$~TeV using a model calibration framework known as Bayesian parameter estimation.
+We begin, in Sec.~\ref{sec:model}, by constructing a nuclear collision model for $p$-Pb and Pb-Pb collisions using initial conditions with parametric nucleon substructure, and transport dynamics described by a pre-equilibrium free-streaming stage, viscous hydrodynamics and microscopic Boltzmann transport.
+In Sec.~\ref{sec:calibration}, we calibrate free parameters of the model to fit charged particle yields, mean $p_T$ and anisotropic flow coefficients of \emph{both} collision systems at $\sqrts=5.02$~TeV, and finally, in Secs.~\ref{sec:results} and \ref{sec:summary}, we present posterior results for the model input parameters and comment on the implications for hydrodynamics descriptions of small collision systems.
 
-%Section 1 describes our nuclear collision model.
-%Sections 2 and 3 describe the computer experiment design and model calibration process which is used to obtain posterior estimates of the model input parameters.
-%In section 4 we presents the results of our analysis.
+\section{nuclear collision model}
+\label{sec:model}
 
-%We perform a global analysis of soft-sector observables measured in $p$-Pb and Pb-Pb collisions at $\sqrts=5.02$~TeV and calibrate free parameters of a hybrid transport model to describe the experimental data.
-%
-%to obtain posterior probability estimates for the model parameters given our chosen model, the experimental data and known sources of uncertainty.
-
-\section{Nuclear collision model}
-
-We employ a multi-stage hybrid transport model that uses relativistic viscous hydrodynamics to describe the QGP medium and microscopic Boltzmann transport to simulate the dynamics of the system after hadronization \note.
+We employ a multi-stage hybrid transport model that uses relativistic viscous hydrodynamics to describe the QGP medium and microscopic Boltzmann transport to simulate the dynamics of the system after hadronization \cite{Shen:2014vra, Bernhard:2016tnd}.
 The hydrodynamic initial conditions are provided by a modified version of the \trento\ model \cite{Moreland:2014oya} with additional parameters to vary the size and shape of the nucleon.
-Each initial condition profile is free streamed to the hydrodynamic starting time and matched onto the hydrodynamic energy-momentum tensor using the Landau matching procedure \note.
-All components of the present model, with the exception of the nucleon substructure implementation, are documented in previous studies \note. We review them here for clarity.
+Each initial condition profile is free streamed to the hydrodynamic starting time and matched onto the hydrodynamic energy-momentum tensor using the Landau matching procedure \cite{Broniowski:2008qk, Heinz:2015arc}.
+Many of the components of the present model have been documented in previous studies \cite{Moreland:2014oya, Bernhard:2016tnd, Bernhard:2018hnz}; we review each component here for clarity.
 
@@
 We model the QGP initial state in $p$-Pb and Pb-Pb collisions at $\sqrts=5.02$~TeV using a simple parametric form for boost-invariant entropy deposition employed by the \trento\ model \cite{Moreland:2014oya}.
-Generally speaking, the initial three-dimensional distribution of matter produced in relativistic nuclear collisions is \emph{not} boost-invariant; longitudinal entropy deposition fluctuates both locally point-to-point in the transverse plane as well as globally event-by-event due to asymmetries in the sampled density of participant matter \note.
-Nevertheless, boost-invariance has been shown to be a good approximation for both large and small collision systems when hydrodynamic observables are calculated from particles that are detected close to midrapidity \note.
+Generally speaking, the initial three-dimensional distribution of matter produced in relativistic nuclear collisions is \emph{not} boost-invariant; longitudinal entropy deposition fluctuates both locally point-to-point in the transverse plane as well as globally event-by-event due to asymmetries in the sampled density of participant matter \cite{Ke:2016jrd, Bozek:2010vz}.
+Nevertheless, boost-invariance has been shown to be a good approximation for both large and small collision systems when hydrodynamic observables are calculated from particles that are detected close to midrapidity \cite{Shen:2016zpp}.
 
@@
 \begin{equation}
-  \label{Tpp}
+  \label{eq:tpp}
   T_{pp}(b) \equiv \int dx\, dy \int dz\, \rho_A(\x) \int dz\, \rho_B(\x + \mathbf{b}),
@@
 \begin{equation}
-  \label{Pcoll}
+  \label{eq:pcoll}
   P_\mathrm{coll} = 1 - \exp[-\sigma_\mathrm{eff}\, T_{pp}(b)], \\
@@
 \begin{equation}
-  \label{sigma_nn}
+  \label{eq:sigma_nn}
   \sigma_{pp}^\mathrm{inel} = \int 2 \pi b\, db\, P_\mathrm{coll}(b).
 \end{equation}
-The proton densities $\rho_{A,B}$ in Eqn.~\eqref{Tpp} are commonly modeled using a spherically symmetric distribution \note.
-For instance, the original implementation of the \trento\ model uses Gaussian protons, largely because it yields a simple analytic solution to Eqn.~\eqref{Pcoll}.
+The proton densities $\rho_{A,B}$ in Eqn.~\eqref{eq:tpp} are commonly modeled using a spherically symmetric distribution.
+For instance, the original implementation of the \trento\ model uses Gaussian protons, largely because it yields a simple analytic solution to Eqn.~\eqref{eq:pcoll}.
 Needless to say, such approximations are admittedly crude and may have a significant effect on the dynamics of small collision systems where the proton size is comparable to the size of the produced QGP medium.
 
-Several previous studies have investigated the effects of deformed or ``lumpy'' protons \note.
-One common implementation is a superposition of three valence quarks, typically described by Gaussian or exponential form factors.
-The corresponding proton density $\rho(\mathbf{x})$ is then assumed to be that of predominantly small-x gluons, seeded by the distribution of color charge in each of the three valence quarks \note.
+A number of previous studies have investigated the effects of deformed or ``lumpy'' protons.
+One common implementation is a superposition of three valence quarks, typically described by Gaussian or exponential form factors \cite{Welsh:2016siu, Bozek:2017jog, Schenke:2014zha, Schlichting:2014ipa, Adare:2015bua, Broniowski:2016pvx}.
+The corresponding proton density $\rho(\mathbf{x})$ is then assumed to be that of predominantly small-x gluons, seeded by the distribution of color charge in each of the three valence quarks.
 
@@
   \end{tikzpicture}
-  \caption{\label{fig:substructure} Schematic of plausible proton shapes.
+  \caption{
+    \label{fig:substructure}
+    Schematic of plausible proton shapes.
     The sketch on the left shows a spherically symmetric proton (dashed line), while the middle and right illustrations depict a fluctuating proton with three and nine constituents respectively (solid lines).
@@
 \begin{equation}
-  \label{rho}
+  \label{eq:rho}
   \rho_{A, B}(\x) = \frac{1}{M} \sum\limits_{i=1}^M \rho_c(\mathbf{x} - \mathbf{x_i}),
@@
 \end{equation}
-and each constituent's position $\mathbf{x_i}$ in Eqn.~\eqref{rho} is sampled from a Gaussian radial distribution of width $r$.
+and each constituent's position $\mathbf{x_i}$ in Eqn.~\eqref{eq:rho} is sampled from a Gaussian radial distribution of width $r$.
 In the limit $M\to \infty$, the composite proton's root mean square radius is simply the convolution of its sampling radius and constituent width, which add together in quadrature
@@
 \end{equation}
-We therefore choose to reparametrize the sampling radius $r$ in terms of the variable $\sigma = \sqrt{r^2 - w^2}$ which approximates the proton's root mean square radius when the number of sampled constituents is large.
+We therefore choose to reparametrize the sampling radius $r$ in terms of the variable $\w = \sqrt{r^2 - w^2}$ which approximates the rms ``width'' of the proton when the number of sampled constituents is large.
 
-The two protons $A$, $B$ are assigned a random impact parameter, and Eqn.~\eqref{Pcoll} is used to sample their inelastic collision probability $P_\mathrm{coll}(b)$.
+The two protons $A$, $B$ are assigned a random impact parameter, and Eqn.~\eqref{eq:pcoll} is used to sample their inelastic collision probability $P_\mathrm{coll}(b)$.
 Note that this proton-proton inelastic collision probability has no direct knowledge of the individual constituent degrees of freedom; it is only \emph{indirectly} sensitive through the geometry of $\rho_{A, B}$ which depends on each of the constituent positions.
-This is an important distinction between the present model and a similar nucleon substructure implementation known as the participant or ``wounded'' quark model which allows for a subset of quarks (constituents) to participate inside a single nucleon \note.
+This is an important distinction between the present model and a similar nucleon substructure implementation known as the participant or ``wounded'' quark model which allows for a subset of quarks (constituents) to participate inside a single nucleon \cite{ANISOVICH1978477, Broniowski:2016pvx}.
 The proton, unlike the nucleus, cannot produce semi-stable spectator fragments in a high-energy collision.
 Any spectator quarks produced by a wounded quark model would be colored objects that necessarily contribute to secondary particle production as they fragment and recombine to form color-neutral hadrons.
-We correspondingly require that the nucleons in Eqn.~\eqref{rho} participate as singular objects, such that all spectator matter discarded by the simulation is appropriately color-neutral and inert.
+We correspondingly require that the nucleons in Eqn.~\eqref{eq:rho} participate as singular objects, such that all spectator matter discarded by the simulation is appropriately color-neutral and inert.
 
@@
 \begin{equation}
-  \label{fluctuated_thick}
+  \label{eq:fluctuated_thick}
   \T_{A, B}(\x) \equiv \int dz\, \frac{1}{M} \sum\limits_{i=1}^M \gamma_i\, \rho_c \,(\mathbf{x} - \mathbf{x_i} \pm \mathbf{b}/2),
 \end{equation}
-equal to the beam-integrated proton density in Eqn.~\eqref{rho}, with each constituent shifted by the appropriate impact parameter offset, and multiplied by a gamma random variable $\gamma_i$ with unit mean and variance $1/k$.
+equal to the beam-integrated proton density in Eqn.~\eqref{eq:rho}, with each constituent shifted by the appropriate impact parameter offset, and multiplied by a gamma random variable $\gamma_i$ with unit mean and variance $1/k$.
 These \emph{ad hoc} gamma random weights are necessary to describe the large fluctuations observed in minimum bias proton-proton collisions, although their physical justification is not yet well understood.
@@
 A natural first guess for the function $f$ is an arithmetic mean $f \propto (\T_A + \T_B)/2$, equal to a wounded nucleon model up to meaningless factor of two in the overall normalization.
-The wounded nucleon model was in fact the first such mapping used as a proxy for initial particle production and entropy deposition in relativistic heavy-ion collisions \note.
-It was quickly realized, however, that the wounded nucleon model predicts the wrong scaling for charged particle production as a function of collision centrality and hence the wrong scaling for initial entropy deposition as a function of participant thickness $\T_{A,B}$ \note.
+The wounded nucleon model was in fact the first such mapping used as a proxy for initial particle production and entropy deposition in relativistic heavy-ion collisions \cite{Bialas:1976ed}.
+It was quickly realized, however, that the wounded nucleon model predicts the wrong scaling for charged particle production as a function of collision centrality and hence the wrong scaling for initial entropy deposition as a function of participant thickness $\T_{A,B}$ \cite{Kharzeev:2000ph}.
 
@@
 \begin{equation}
-  \label{gmean}
+  \label{eq:gmean}
   \frac{dS}{dy}\bigg\vert_{y=0} \propto \left( \frac{\T_A^p + \T_B^p}{2} \right)^{1/p},
@@
 \end{equation}
-Conveniently, it has been shown that the generalized mean ansatz is able to mimic the scaling behavior of certain saturation based initial condition models \note, and hence it should serve as a reasonable parametric form for exploring QGP entropy deposition, assuming imperfect knowledge of saturation effects in nature.
-Of course the model is \emph{not} a substitute for first principle saturation calculations, and may fail to reproduce nuanced features of \emph{ab initio} saturation theory calculations, such as the existence of short-range gluon field fluctuations \note.
+Conveniently, it has been shown that the generalized mean ansatz is able to mimic the scaling behavior of certain saturation based initial condition models \cite{Bernhard:2016tnd}, and hence it should serve as a reasonable parametric form for exploring QGP entropy deposition, assuming imperfect knowledge of saturation effects in nature.
+Of course the model is \emph{not} a substitute for first principle saturation calculations, and may fail to reproduce nuanced features of \emph{ab initio} saturation theory calculations, such as the existence of short-range gluon field fluctuations \cite{Schenke:2012wb}.
 
@@
   \begin{tikzpicture}
-    \node {\includegraphics[width=\columnwidth]{fig/thickness}};
+    \node {\includegraphics{thickness}};
     \node[label=above:{10 fm}] (a) at (0, 1.1) {};
@@
   \end{tikzpicture}
-  \caption{\label{fig:substructure} Effect of nucleon substructure on the nuclear thickness function $T(x, y) \equiv \int dz\, \rho(x, y, z)$ of a $^{208}\mathrm{Pb}$ nucleus.
+  \caption{
+    \label{fig:thickness} Effect of nucleon substructure on the nuclear thickness function $T(x, y) \equiv \int dz\, \rho(x, y, z)$ of a $^{208}\mathrm{Pb}$ nucleus.
     The nucleus on the left has Gaussian nucleons of width $0.88$~fm, while the nucleus on the right has composite nucleons, each containing six Gaussian constituents of width $0.41$~fm.}
@@
 
-Equation~\eqref{gmean} is a purely local function of nuclear density in the transverse plane and should (in principle) be equally valid for any pair of colliding nuclei at sufficiently high beam energy.
+Equation~\eqref{eq:gmean} is a purely local function of nuclear density in the transverse plane and should (in principle) be equally valid for any pair of colliding nuclei at sufficiently high beam energy.
 The model readily generalizes from individual proton-proton collisions to arbitrary nucleus-nucleus collisions by summing the participant thicknesses $\T_{A,B}$ over all nucleons which participate in one or more inelastic collisions.
@@
 When applying the model to heavy-ions, we sample nucleon positions from a Woods-Saxon density distribution subject to a minimum distance criteria $|\mathbf{x_i} - \mathbf{x_j}| > d_\mathrm{min}$ between all pairs of nucleons $i$, $j$.
-The minimum distance algorithm, first described in Ref.~\note, uses a simple trick to resample the nucleon positions without modifying the target Woods-Saxon radial distribution.
+The minimum distance algorithm, first described in Ref.~\cite{Bernhard:2018hnz}, uses a simple trick to resample the nucleon positions without modifying the target Woods-Saxon radial distribution.
 We first presample the radii of all nucleons in a given nucleus and sort them in ascending order.
@@
 
-\begin{figure}[b]
-  \includegraphics{fig/coupling}
-  \caption{\label{fig:freestream} Cartoon of the free-streaming approximation for hydrodynamic initialization.
-The initial state is free-streamed for proper time $\tau_{fs}$ (zero-coupling) before it is matched to hydrodynamics (strong-coupling).
-This piecewise evolution approximates the more realistic scenario expected in nature where the medium's coupling strength smoothly changes as a function of time.}
-\end{figure}
+\fig[b]{coupling}{
+Cartoon of the free-streaming approximation for hydrodynamic initialization.
+The initial state is free-streamed for proper time $\taufs$ (zero-coupling) before it is matched to hydrodynamics (strong-coupling).
+This piecewise evolution approximates the more realistic scenario expected in nature where the medium's coupling strength smoothly changes as a function of time.
+}
 
@@
 Realistically, one expects the initial parton interactions to lie somewhere between these two extremes.
-We therefore choose to model the QGP's initial off-equilibrium dynamics using a simple step-function approximation, depicted in Fig.~\ref{fig:freestream}, which free-streams the initial state for proper time $\tau_{fs}$ (zero coupling) before instantaneously switching to viscous hydrodynamics (strong coupling) \note.
-The free parameter $\tau_{fs}$ allows us to parametrically vary the \emph{time averaged} coupling strength in the approximate window $0 < \tau_{fs} \lesssim1$~fm/$c$.
+We therefore choose to model the QGP's initial off-equilibrium dynamics using a simple step-function approximation, depicted in Fig.~\ref{fig:coupling}, which free-streams the initial state for proper time $\taufs$ (zero coupling) before instantaneously switching to viscous hydrodynamics (strong coupling) \cite{Liu:2015nwa, Broniowski:2008qk}.
+The free parameter $\taufs$ allows us to parametrically vary the \emph{time averaged} coupling strength in the approximate window $0 < \taufs \lesssim1$~fm/$c$.
 
-The parametric entropy deposition ansatz in Eqn.~\eqref{gmean} does not provide any information about the initial masses or momenta of particles produced in the collision.
+The parametric entropy deposition ansatz in Eqn.~\eqref{eq:gmean} does not provide any information about the initial masses or momenta of particles produced in the collision.
 In general, these details will affect the dynamics predicted by the collisionless Boltzmann equation
@@
 
-After free-streaming for proper time $\tau_{fs}$, we transition to viscous hydrodynamics which solves the conservation equations
+After free-streaming for proper time $\taufs$, we transition to viscous hydrodynamics which solves the conservation equations
 \begin{equation}
@@
 
-\begin{figure}
-  \centering
-  \includegraphics{fig/viscosity_dof}
-  \caption{\label{fig:viscosity_dof} Degrees of freedom in the temperature dependent shear and bulk viscosity parametrizations.
+\fig[t]{viscosity_dof}{
+Degrees of freedom in the temperature dependent shear and bulk viscosity parametrizations.
 Lines are chosen for illustrative purposes only and do not represent all possible variability.
@@
 }
-\end{figure}
 
 The hydrodynamic equations of motion are necessarily closed using an equation of state (EoS) to relate the energy density $e$ and pressure $P$ of the produced medium.
-We use a parametrization for $P(e)$ that matches a hadron resonance gas EoS at low temperature to a lattice QCD EoS at high temperature by smoothly connecting their trace anomalies in the interval $165 \le T \le 200$~MeV.
+We use a parametrization for $P(e)$ that matches a hadron resonance gas EoS at low temperature to a lattice QCD EoS at high temperature by smoothly connecting their trace anomalies in the interval
+$165 \le T \le 200$~MeV.
 For the lattice EoS, we use a calculation by the HotQCD collaboration for (2+1)-flavor QCD which was extrapolated to the continuum limit \cite{Bazavov:2014pvz}.
@@
 
-The aforementioned hydrodynamic equations are solved numerically using the boost-invariant \texttt{VISH2+1} viscous hydrodynamics code \note\ which is documented extensively in Ref.~\note.
+The aforementioned hydrodynamic equations are solved numerically using the boost-invariant \texttt{VISH2+1} viscous hydrodynamics code \cite{Song:2007ux} which is documented extensively in Ref.~\cite{Shen:2014vra}.
 We simulate each hydrodynamic event on a spacetime grid with transverse extent $x_\mathrm{max}$, spatial grid step $dx$, and time step $d\tau$ which are optimized \emph{event-by-event} to balance trade-offs between numerical accuracy and computation time (see appendix).
-Although these details are somewhat mundane, they are critically important to the present study, since the computation time explodes with the number of grid cells,
-\begin{equation}
-  \mathrm{cpu~time} \sim n_x^2 n_\tau,
-\end{equation}
-and $nx$ and $n\tau$ typically have to be quite large to resolve the small length scales associated with nucleon substructure.
+Although these details are somewhat mundane, they are critically important to the present study, since the computation time scales with the number of spacetime cells $n_x^2\, n_\tau$, and $n_x$ and $n_\tau$ typically have to be quite large to resolve the small length scales associated with nucleon substructure.
 
@@
 
-We evolve the system hydrodynamically down to a pre-specified switching isotherm $T_{sw}$ at which point the medium is converted into particles using the Cooper-Frye formula \note
+We evolve the system hydrodynamically down to a pre-specified switching isotherm $T_{sw}$ at which point the medium is converted into particles using the Cooper-Frye formula \cite{PhysRevD.10.186}
 \begin{equation}
@@
 Traditionally, particlization models have sampled resonances using each particle's pole mass in Eqn.~\eqref{distribution}.
-This approximation, however, is somewhat crude and has been known to underpredict pion production, particularly at low $p_T$ \note.
-We thus follow Ref.~\note, and sample particles with a \emph{distribution} of masses
+This approximation, however, is somewhat crude and has been known to underpredict pion production, particularly at low $p_T$ \cite{Sollfrank:1991xm, Huovinen:2016xxq, Vovchenko:2018fmh}.
+We thus follow Ref.~\cite{Bernhard:2018hnz}, and sample particles with a \emph{distribution} of masses
 \begin{equation}
@@
 
-We tabulate the values of $\{\Gamma_0, m_0, m_\mathrm{min}\}$ for all particles and sample the masses of each particle during particlization \note.
+We tabulate the values of $\{\Gamma_0, m_0, m_\mathrm{min}\}$ for all particles and sample the masses of each particle during particlization \cite{PDG:2017}.
 The resonances are then passed to a hadronic transport model, described shortly, which simulates subsequent scatterings and decays.
@@
 When the viscous terms $\pi^{\mu\nu}$ and $\Pi$ are nonzero in Eqn.~\eqref{hydro-eqn}, the distribution function $f$ must be modified to preserve the continuity of $T^{\mu\nu}$ as the system transitions from hydrodynamics to Boltzmann transport.
-We perform the appropriate modification using a general method which transforms the momentum vector \emph{inside} the distribution function
+We perform the appropriate modification using a general method which transforms the momentum vector \emph{inside} the distribution function \cite{Pratt:2010jt}
 \begin{align}
+  \label{viscous_correction}
   p_i \rightarrow p'_i &= p_i + \sum\limits_j \lambda_{ij}\, p_j,\\
@@
 \end{align}
-where $\lambda_{ij}$ is a linear transformation matrix consisting of a traceless shear part and a bulk part which is proportional to the identity matrix \note.
+where $\lambda_{ij}$ is a linear transformation matrix consisting of a traceless shear part and a bulk part which is proportional to the identity matrix.
 
-We use for the shear viscous correction the form \note
+We use for the shear viscous correction the form \cite{Pratt:2010jt}
 \begin{equation}
@@
 \end{equation}
+For the bulk viscous correction, we use a novel procedure developed in Ref.~\cite{Bernhard:2018hnz}.
+The total kinetic pressure of the system is
+\begin{equation}
+  \label{kinetic-pressure}
+  P + \Pi = \sum\limits_\mathrm{sp} g \int \frac{d^3p}{(2\pi)^3} \frac{p^2}{3E} f(p).
+\end{equation}
+For a given bulk pressure, we rescale the momentum $p$ inside the distribution function $f(p) \rightarrow f(p + \lambda_\mathrm{bulk}\, p)$ and adjust the parameter $\lambda_\mathrm{bulk}$ to match the total pressure on the left side of Eqn.~\eqref{kinetic-pressure}.
+This substitution of course also modifies the energy density
+\begin{equation}
+  e = \sum\limits_\mathrm{sp} g \int \frac{d^3p}{(2 \pi)^3} E f(p),
+\end{equation}
+and so a fugacity term $z_\mathrm{bulk}$ is introduced which decreases the yield of all particles by the same overall factor to compensate.
+The full transformation is then given by $f(p) \rightarrow z_\mathrm{bulk}\, f(p + \lambda_\mathrm{bulk}\, p)$, where the parameters $\lambda_\mathrm{bulk}$ and $z_\mathrm{bulk}$ are determined numerically for each value of the bulk pressure.
 
-For the bulk viscous correction, we use a novel procedure developed by one of the authors of this work \note.
+\subsubsection{Boltzmann transport}
 
-\subsubsection{Hadronic Boltzmann transport}
+Once the fluid is converted into particles, we simulate its subsequent microscopic dynamics using the Ultra-relativistic Quantum Molecular Dynamics (\texttt{UrQMD}) transport model \cite{Bass:1998ca, Bleicher:1999xi}.
+It solves the microscopic Boltzmann equation
+\begin{equation}
+  \frac{df_i(x, p)}{dt} = \mathcal{C}_i(x, p)
+\end{equation}
+where $f_i$ is the distribution function for species $i$, and $\mathcal{C}_i$ is its microscopic collision kernel.
+The model propagates all produced hadrons along classical trajectories, and simulates their scatterings, resonance formations and decays until the last interactions cease.
 
-\vfill
+One primary advantage of using a microscopic transport model such as \texttt{UrQMD} as an afterburner, is that it realistically simulates the system break-up when the mean free path becomes large relative to the system size.
+This dilute limit is expected to play a significant role in small collision systems where the produced medium is smaller and shorter lived.
 
-\section{Parameter estimation}
+\section{parameter estimation}
+\label{sec:calibration}
 
-\subsection{Computer experiment design}
+The nuclear collision model constructed in Sec.~\ref{model} includes a number of free parameters $\x$ which describe the initial state, pre-equilibrium dynamics, and hydrodynamic medium.
+We assume that these parameters are embedded in a realistic model such that each parameter is meaningful, but we concede that we know very little about their true values in nature.
 
-\subsection{Gaussian process emulators}
+\begin{table}
+  \caption{
+    \label{tab:design}
+    Input parameter ranges for the initial condition and hydrodynamic models.
+  }
+  \begin{ruledtabular}
+  \begin{tabular}{lll}
+    Parameter         & Description                        & Range             \\
+    \paddedhline
+    Norm              & Normalization factor                 & 9--28           \\
+    $p$               & Entropy deposition parameter         & $-1$ to $+1$    \\
+    $\sigmaf$         & Nucleon fluctuation std.\ dev.\      & 0--2            \\
+    $\w$              & Constituent sampling width           & 0.4--1.2 fm     \\
+    $\n$              & Number of constituents               & 1--9            \\
+    $\X$              & Nucleon structure parameter          & 0--1            \\
+    $\dmin$           & Minimum inter-nucleon distance       & 0--1.7 fm       \\
+    $\taufs$          & Free-streaming time                  & 0.1--1.5 fm/$c$ \\
+    $\eta/s$ min      & Minimum value of $\eta/s$ (at $T_c$) & 0--0.2          \\
+    $\eta/s$ slope    & Slope of $\eta/s$ above $T_c$        & 0--8 GeV$^{-1}$ \\
+    $\eta/s$ crv      & Curvature of $\eta/s$ above $T_c$    & $-1$ to $+1$    \\
+    $\zeta/s$ max     & Maximum value of $\zeta/s$           & 0--0.1          \\
+    $\zeta/s$ width   & Width of $\zeta/s$ peak              & 0--0.1 GeV      \\
+    $\zeta/s$ $T_0$   & Temperature of $\zeta/s$ maximum     & 150--200 MeV    \\
+    $T_\text{switch}$ & Switching/particlization temp.       & 135--165 MeV    \\
+  \end{tabular}
+  \end{ruledtabular}
+\end{table}
 
-\subsection{Bayesian calibration}
+\section{results}
+\label{sec:results}
 
-\section{Results}
+\fig*[t]{validation_example}{
+  Placeholder
+}
 
-\subsection{Initial condition properties}
+\fig*[t]{observables_design}{
+Hybrid model calculations compared to experimental data for $p$-Pb collisions (left column) and Pb-Pb collisions (right column) at $\sqrts=5.02$~TeV.
+Colored lines are model calculations (no emulator) for each of the 500 design points.
+Points with error bars are experimental data from ALICE and CMS \note, with statistical and systematic errors added in quadrature.
+}
 
-\subsection{QGP medium properties}
+\fig*[t]{observables_posterior}{
+Calibrated emulator predictions compared to experimental data for $p$-Pb collisions (left column) and Pb-Pb collisions (right column) at $\sqrts=5.02$~TeV.
+Colored lines are emulator predictions for $n=100$ different parameter sets randomly sampled from the Bayesian posterior.
+Points with error bars are experimental data from ALICE and CMS \note, with statistical and systematic errors added in quadrature.
+}
 
-\section{Summary and conclusions}
+\begin{figure*}[p]
+  \makebox[\textwidth]{
+    \includegraphics[width=.95\paperwidth]{posterior}
+  }
+  \captionsetup{
+    margin={.3\textwidth,0pt},
+    justification=raggedleft,
+    skip=-.75\textheight
+  }
+  \caption{Bayesian posterior distribution of the model input parameters.
+The diagonal panels show the marginalized distributions of individual\\ model parameters, while the off-diagonal panels show the joint\\ distributions for pairs of model parameters, visualizing\\ their correlations.
+The marginalized distribution\\ medians are annotated along the diagonal.}
+  \label{fig:posterior}
+  \vspace{.5\textheight}
+\end{figure*}
+
+\fig*[t]{observables_map}{
+Model calculations using the \emph{maximum a posterior} (MAP) parameters compared to experiment.
+Colored lines are model calculations for $p$-Pb collisions (left) and Pb-Pb collisions (right) at $\sqrts=5.02$~TeV.
+Points with error bars are the experimental data with statistical uncertainties, and gray bands their corresponding systematic uncertainties, from ALICE and CMS \note.
+The ratio axes show the ratio of model over data where where available with gray bands indicating $\pm 10\%$.
+}
+
+\fig*[t]{flow_corr}{
+  Model calculations of the symmetric cumulants for Pb-Pb collisions at $\sqrts=5.02$~TeV using the \emph{maximum a posteriori} (MAP) parameters.
+}
+
+\fig{posterior_proton_shape}{
+  Posterior distribution for the constituent sampling radius $\R$ and constituent width $\w$.
+  The white trapezoidal region (lower-right) covers the initial range of allowed values for $\R$ and $\w$, while the gray triangular region (upper-left) indicates values which are excluded by the prior $\w > \R$.
+  The posterior distribution, shown in blue, indicates the preferred values for $\R$ and $\w$ determined by the analysis.
+}
+
+\fig*[t]{region_shear_bulk}{
+  Left figure: estimated temperature dependence of the QGP specific shear viscosity $(\eta/s)(T)$ determined by the present Bayesian analysis of $p$-Pb and Pb-Pb collisions at $\sqrts=5.02$~TeV (orange line/band) compared to a previous Bayesian analysis of Pb-Pb collisions at $\sqrts=2.76$ and 5.02~TeV (blue line/band) \note.
+  The lines are the medians of each posterior distribution, and the bands are their 90\% credible regions.
+  Right figure: same as before, but for the temperature dependence of the QGP specific bulk viscosity $(\zeta/s)(T)$.
+}
+
+\section{summary}
+\label{sec:summary}
 
@@
 
-\bibliography{trento-substructure}
+\bibliography{substructure}
 

### 2018-07-10 src/model.py

@@ -331,60 +331,60 @@ def _data(system, dataset='main'):
     species = ['pion', 'kaon', 'proton']
 
     # training data is only missing for Pb-Pb charged particle mean pT
-    if dataset == 'main' and system == 'PbPb5020':
+    #if dataset == 'main' and system == 'PbPb5020':
+    #    edges = np.linspace(0, 70, 15)
+    #    cent = [(a, b) for a, b in zip(edges[:-1], edges[1:])]
+    #    x = [(a + b)/2 for a, b in cent]
+    #    empty = dict(cent=cent, x=x)
+    #    data = dict({'mean_pT': {None: empty}}, **data)
+
+    # calculate additional observables for the maximum a posteriori parameters
+    #if dataset == 'map':
+
+    # p-Pb map observables
+    if system == 'pPb5020':
+
+        # yields
+        cent, x = [data['dNch_deta'][None][k] for k in ('cent', 'x')]
+        empty = dict(cent=cent, x=x)
+        data = dict({'dET_deta': {None: empty}}, **data)
+        data = dict({'iden_dN_dy': {s: empty for s in species}}, **data)
+
+        # mean pT
+        mult, x = [data['mean_pT'][None][k] for k in ('mult', 'x')]
+        empty = dict(mult=mult, x=x)
+        data = dict({'iden_mean_pT': {s: empty for s in species}}, **data)
+        data = dict({'pT_fluct': {None: empty}}, **data)
+
+        # symmetric cumulants
+        mult, x = [data['vnk'][(2, 2)][k] for k in ('mult', 'x')]
+        empty = dict(mult=mult, x=x)
+        for obs in ['sc', 'sc_normed', 'sc_central', 'sc_normed_central']:
+            data = dict({obs: {(3, 2): empty, (4, 2): empty}}, **data)
+
+    # Pb-Pb map observables
+    if system == 'PbPb5020':
+
+        # yields
+        cent, x = [data['dNch_deta'][None][k] for k in ('cent', 'x')]
+        empty = dict(cent=cent, x=x)
+        data = dict({'dET_deta': {None: empty}}, **data)
+        data = dict({'iden_dN_dy': {s: empty for s in species}}, **data)
+
+        # mean pT
         edges = np.linspace(0, 70, 15)
         cent = [(a, b) for a, b in zip(edges[:-1], edges[1:])]
         x = [(a + b)/2 for a, b in cent]
         empty = dict(cent=cent, x=x)
         data = dict({'mean_pT': {None: empty}}, **data)
[…]

### 2018-07-10 src/plots.py

@@ -99,6 +99,7 @@ plt.rcParams.update({
     'lines.markersize': 3,
     'lines.markeredgewidth': 0,
     'patch.linewidth': .8,
+    'hatch.linewidth': .6,
     'axes.linewidth': .6,
     'xtick.major.width': .6,
     'ytick.major.width': .6,
@@ -345,14 +346,14 @@ def _observables_plots():
             ),
             xlim=dict(
                 pPb5020=(0, 60),
-                PbPb5020=(0, 75),
+                PbPb5020=(0, 80),
             ),
-            ylim=(2e1, 2e5),
+            ylim=(1e2, 1e5),
             yscale='log',
             height_ratio=1.5,
             subplots=[
-                ('dNch_deta', None, dict(label=r'$N_\mathrm{ch}$ x25', scale=25)),
-                ('dET_deta', None, dict(label=r'$E_T$ x5', scale=5)),
+                ('dNch_deta', None, dict(label=r'$N_\mathrm{ch}^{\times 25}$', scale=25)),
+                ('dET_deta', None, dict(label=r'$E_T^{\times 5}$', scale=5)),
                 *id_parts_plots('iden_dN_dy')
             ]
         ),
@@ -365,7 +366,7 @@ def _observables_plots():
             ylabel=r'$\langle p_T \rangle$ [GeV]',
             xlim=dict(
                 pPb5020=(1, 6),
-                PbPb5020=(0, 75),
+                PbPb5020=(0, 80),
             ),
             ylim=(0, 1.5),
             subplots=[
@@ -374,15 +375,30 @@ def _observables_plots():
             ]
         ),
         dict(
+            title='Mean $p_T$ fluctuations',
+            xlabel=dict(
+                pPb5020=r'$n_\mathrm{ch} / \langle n_\mathrm{ch} \rangle$',
+                PbPb5020='Centrality %'
+            ),
+            ylabel=r'$\delta p_T/\langle p_T \rangle$',
+            xlim=dict(
+                pPb5020=(1, 6),
+                PbPb5020=(0, 75),
+            ),
+            ylim=(0, 0.05),
+            subplots=[('pT_fluct', None, dict())]
+        ),
+        dict(
             title='Flow cumulants',
             xlabel=dict(
-                pPb5020=r'$n_\mathrm{trk}^\mathrm{offline} / \langle n_\mathrm{trk}^\mathrm{offline}\rangle$',
+                pPb5020=(r'$n_\mathrm{trk}^\mathrm{offline}/'
[…]

### 2018-07-11 paper/substructure.tex
Start writing Bayesian section

@@
 
-\usepackage[labelfont={color=theblue}, justification=raggedright, labelsep=quad]{caption}
+\usepackage[font=small, justification=raggedright, labelsep=quad]{caption}
 
@@
 \newcommand{\x}{\mathbf{x}}
+\newcommand{\y}{\mathbf{y}}
+\newcommand{\yexp}{\mathbf{y_\mathrm{exp}}}
 \newcommand{\X}{\mathcal{X}_\mathrm{struct}}
@@
 
-\section{nuclear collision model}
+\section{Nuclear collision model}
 \label{sec:model}
@@
   \end{tikzpicture}
-  \caption{
+  \caption{Schematic of plausible proton shapes.
+The sketch on the left shows a spherically symmetric proton (dashed line), while the middle and right illustrations depict a fluctuating proton with three and nine constituents respectively (solid lines).}
   \label{fig:substructure}
-    Schematic of plausible proton shapes.
-    The sketch on the left shows a spherically symmetric proton (dashed line), while the middle and right illustrations depict a fluctuating proton with three and nine constituents respectively (solid lines).
-  }
 \end{figure}
@@
   \end{tikzpicture}
-  \caption{
-    \label{fig:thickness} Effect of nucleon substructure on the nuclear thickness function $T(x, y) \equiv \int dz\, \rho(x, y, z)$ of a $^{208}\mathrm{Pb}$ nucleus.
+  \caption{Effect of nucleon substructure on the nuclear thickness function $T(x, y) \equiv \int dz\, \rho(x, y, z)$ of a $^{208}\mathrm{Pb}$ nucleus.
 The nucleus on the left has Gaussian nucleons of width $0.88$~fm, while the nucleus on the right has composite nucleons, each containing six Gaussian constituents of width $0.41$~fm.}
+  \label{fig:thickness}
 \end{figure}
@@
 
-\fig[t]{viscosity_dof}{
-Degrees of freedom in the temperature dependent shear and bulk viscosity parametrizations.
+\fig[t]{viscosity_dof}{Degrees of freedom in the temperature dependent shear and bulk viscosity parametrizations.
 Lines are chosen for illustrative purposes only and do not represent all possible variability.
@@
 
-\section{parameter estimation}
+\section{Parameter estimation}
 \label{sec:calibration}
 
-The nuclear collision model constructed in Sec.~\ref{model} includes a number of free parameters $\x$ which describe the initial state, pre-equilibrium dynamics, and hydrodynamic medium.
-We assume that these parameters are embedded in a realistic model such that each parameter is meaningful, but we concede that we know very little about their true values in nature.
+The nuclear collision model constructed in Sec.~\ref{sec:model} includes a number of free parameters $\x$ which describe the initial state, pre-equilibrium dynamics, and hydrodynamic medium.
+Given values for the parameters $\x$, the model may be used to predict a vector of experimental observables $\y$.
+For example, $\y$ might be a vector of fixed length which includes charged particle yields in different centrality bins.
+The physics model thus represents a vector-valued function $f(\x) = \y$
+which maps the parameter values $\x$ to the observable state $\y$.
+
+The goal of this work is to estimate the true model parameters $\x_\star$, given some evidence that our model ${f: X \mapsto Y}$ describes the experimental data $\yexp$ for the simulated observables $\y$.
+According to Bayes' theorem, the probability for the true model parameters $\x_\star$ is
+\begin{equation}
+  \label{eq:bayes}
+  P(\x_\star|f, \yexp) \propto P(f, \yexp|\x_\star) P(\x_\star).
+\end{equation}
+The left-side of this expression is the \emph{posterior}: the probability that $\x_\star$ is the correct value given evidence from the model $f: X \mapsto Y$ and the experimental data $\yexp$.
+
+On the right-side, the \emph{likelihood} function $P(f, \yexp | \x_\star)$ is the compatibility of the experimental evidence $\yexp$ with the hypothetical model parameters $\x_\star$, and the \emph{prior} probability $P(\x_\star)$ is the estimate of the probability of $\x_\star$ in the absence of evidence $\yexp$.
+
+\begin{equation}
+  P(H1 | E, H2) \propto P(E | H1, H2)\, P(H1 | H2)
+\end{equation}
+There are three types of events,
+\begin{itemize}
+  \item $H1$: hypothesis that the model parameters $\x_\star$ are the true model parameters.
+  \item $H2$: hypothesis that nuclear collisions are describable using the hybrid transport model formulated in this work.
+  \item $E$: experimental evidence provided by data $\yexp$.
+\end{itemize}
+
+\subsection{Parameter design}
+
+We create a scaffolding of the parameter space and distribute $n=500$ parameter points in our 15-dimensional parameter space using a Latin hypercube design \note.
+The algorithm maximizes the
+
+We assume a flat prior distribution for each model parameter $x \in \x$ described by the range of values shown in Tbl.~\ref{tab:design}.
+The minimum and maximum values of the parameters are conservative to avoid clipping the calibrated posterior; for example, previous analyses of \trento\ model found $p \sim 0$, but we use the prior $-1 < p < 1$ to account for physics differences in the present model which might modify its posterior.
+
+\subsection{Gaussian process emulator}
 
 \begin{table}
-  \caption{
-    \label{tab:design}
-    Input parameter ranges for the initial condition and hydrodynamic models.
-  }
+  \caption{Input parameter ranges for the initial condition and hydrodynamic models.}
   \begin{ruledtabular}
@@
   \end{ruledtabular}
+  \label{tab:design}
 \end{table}
 
-\section{results}
+\section{Results}
 \label{sec:results}
@@
 
-\fig*[t]{observables_design}{
-Hybrid model calculations compared to experimental data for $p$-Pb collisions (left column) and Pb-Pb collisions (right column) at $\sqrts=5.02$~TeV.
-Colored lines are model calculations (no emulator) for each of the 500 design points.
-Points with error bars are experimental data from ALICE and CMS \note, with statistical and systematic errors added in quadrature.
+\begin{figure}[h]
+  \makebox[\textwidth]{
+    \begin{minipage}[t]{\paperwidth}
+      \includegraphics[width=.95\paperwidth]{posterior}
+      \captionsetup{
+        width={.83\textwidth},
+        justification=raggedright,
       }
-
-\fig*[t]{observables_posterior}{
-Calibrated emulator predictions compared to experimental data for $p$-Pb collisions (left column) and Pb-Pb collisions (right column) at $\sqrts=5.02$~TeV.
-Colored lines are emulator predictions for $n=100$ different parameter sets randomly sampled from the Bayesian posterior.
-Points with error bars are experimental data from ALICE and CMS \note, with statistical and systematic errors added in quadrature.
+      \caption{Bayesian posterior distribution of the model input parameters.
+The diagonal panels show the marginalized distributions of individual model parameters, while off-diagonal panels show the joint distributions for pairs of model parameters, visualizing their correlations.
+The marginalized distribution medians and 90\% credible intervals are annotated along the diagonal.}
+    \end{minipage}
   }
+  \label{fig:posterior}
+\end{figure}
 
 \begin{figure*}[p]
-  \makebox[\textwidth]{
-    \includegraphics[width=.95\paperwidth]{posterior}
-  }
-  \captionsetup{
-    margin={.3\textwidth,0pt},
-    justification=raggedleft,
-    skip=-.75\textheight
-  }
-  \caption{Bayesian posterior distribution of the model input parameters.
-The diagonal panels show the marginalized distributions of individual\\ model parameters, while the off-diagonal panels show the joint\\ distributions for pairs of model parameters, visualizing\\ their correlations.
-The marginalized distribution\\ medians are annotated along the diagonal.}
-  \label{fig:posterior}
-  \vspace{.5\textheight}
+  \includegraphics{observables_pbpb}
+  \caption{Simulated observables compared to experimental data for Pb-Pb collisions at $\sqrts=5.02$~TeV.
+Top row: explicit model calculations (no emulator) for each of the 500 design points; bottom row: emulator predictions for $n=100$ random samples drawn from the posterior.
+Points with error bars are experimental data from ALICE with statistical and systematic errors added in quadrature \note.}
+  \vspace{5ex}
+  \includegraphics{observables_ppb}
+  \caption{Same as above but for $p$-Pb collisions at $\sqrts=5.02$~TeV.
+Note that multiplicity bins are used for mean $p_T$ and flow cumulant observables to match the bins used by experiment.
+Experimental data are from ALICE and CMS \note.}
 \end{figure*}

### 2018-07-11 src/plots.py

@@ -346,7 +346,7 @@ def _observables_plots():
             ),
             xlim=dict(
                 pPb5020=(0, 60),
-                PbPb5020=(0, 80),
+                PbPb5020=(0, 75),
             ),
             ylim=(1e2, 1e5),
             yscale='log',
@@ -366,7 +366,7 @@ def _observables_plots():
             ylabel=r'$\langle p_T \rangle$ [GeV]',
             xlim=dict(
                 pPb5020=(1, 6),
-                PbPb5020=(0, 80),
+                PbPb5020=(0, 75),
             ),
             ylim=(0, 1.5),
             subplots=[
@@ -398,7 +398,7 @@ def _observables_plots():
             ylabel=r'$v_n\{2\}$',
             xlim=dict(
                 pPb5020=(1, 6),
-                PbPb5020=(0, 80),
+                PbPb5020=(0, 75),
             ),
             ylim=(0, .15),
             subplots=[
@@ -409,8 +409,7 @@ def _observables_plots():
     ]
 
 
-def _observables(posterior=False):
-
+def observables(system):
     """
     Model observables at all design points or drawn from the posterior with
     experimental data points.
@@ -421,17 +420,20 @@ def _observables(posterior=False):
 
     plot_fixes = [
         dict(
+            title=r'Yields $dN_\mathrm{ch}/d\eta$',
             ylabel=r'$dN_\mathrm{ch}/d\eta$',
-            ylim=(1, 1e4),
+            ylim=(1e1, 1e4) if system == 'PbPb5020' else (1, 1e3),
             height_ratio=1,
             subplots=[
-                ('dNch_deta', None, dict(label=r'$N_\mathrm{ch}$', scale=1))
+                ('dNch_deta', None, dict(label=''))
             ]
         ),
         dict(
-            subplots=[('mean_pT', None, dict(label='ch'))]
-        ),
-        dict()
+            title=r'Mean $p_T$ [GeV]',
+            subplots=[('mean_pT', None, dict(label=''))]),
+        dict(
+            title=r'Flow cumulants $v_n\{2\}$',
+        )
     ]
 
     for a, b in zip(plots, plot_fixes):
@@ -439,14 +441,16 @@ def _observables(posterior=False):
             a[k] = v
 
     fig, axes = plt.subplots(
-        nrows=len(plots), ncols=len(systems),
-        figsize=figsize(.8, aspect=1.1),
+        nrows=2, ncols=len(plots),
[…]

### 2018-07-13 paper/substructure.tex
Start writing about the experimental observables and gaussian process emulator.

@@
 \newcommand{\sigmaf}{\sigma_\mathrm{fluct}}
+\newcommand{\X}{\chi_\mathrm{struct}}
 \newcommand{\taufs}{\tau_\mathrm{fs}}
-\newcommand{\n}{n_\mathrm{c}}
-\newcommand{\R}{R_\mathrm{c}}
-\newcommand{\w}{w_\mathrm{c}}
 \newcommand{\dmin}{d_\mathrm{min}}
@@
 \newcommand{\y}{\mathbf{y}}
-\newcommand{\yexp}{\mathbf{y_\mathrm{exp}}}
-\newcommand{\X}{\mathcal{X}_\mathrm{struct}}
 \newcommand{\vnk}[2]{v_#1\{#2\}}
@@
 Each initial condition profile is free streamed to the hydrodynamic starting time and matched onto the hydrodynamic energy-momentum tensor using the Landau matching procedure \cite{Broniowski:2008qk, Heinz:2015arc}.
-Many of the components of the present model have been documented in previous studies \cite{Moreland:2014oya, Bernhard:2016tnd, Bernhard:2018hnz}; we review each component here for clarity.
+Many of the components of the present model have been documented in previous studies \cite{Moreland:2014oya, Bernhard:2016tnd, Bernhard:2018hnz}; we review each component here for completeness.
 
 \subsection{Initial state}
+\label{sec:initial_state}
 
@@
 Within a picture of independent pairwise collisions between the constituents, a Glauber model model may be used to calculate the probability $P_\mathrm{coll}$ that the two protons collide inelastically at impact parameter $b$.
-In the limit when the number of constituents $M \gg 1$, it yields the particularly simple form
+In the limit when the number of constituents $n_c \gg 1$, it yields the particularly simple form
 \begin{equation}
@@
 In this work, we pursue a less restrictive and more parametric description of the proton where the number of substructure degrees of freedom are uncertain as depicted in Fig.~\ref{fig:substructure}.
-We model each nucleon's density $\rho_{A,B}$ as a sum of $M$ independent constituents
+We model each nucleon's density $\rho_{A,B}$ as a sum of $n_c$ independent constituents
 \begin{equation}
   \label{eq:rho}
-  \rho_{A, B}(\x) = \frac{1}{M} \sum\limits_{i=1}^M \rho_c(\mathbf{x} - \mathbf{x_i}),
+  \rho_{A, B}(\x) = \frac{1}{n_c} \sum\limits_{i=1}^{n_c} \rho_c(\mathbf{x} - \mathbf{x_i}),
 \end{equation}
-where each constituent density $\rho_c$ is described by a Gaussian distribution of width $w$
+where each constituent density $\rho_c$ is described by a Gaussian distribution of width $v$
 \begin{equation}
-  \rho_c(\mathbf{x}) = \frac{1}{(2\pi w^2)^{3/2}} \exp\left(-\frac{\x^2}{2 w^2}\right),
+  \label{eq:constituent_density}
+  \rho_c(\mathbf{x}) = \frac{1}{(2\pi v^2)^{3/2}} \exp\left(-\frac{\x^2}{2 v^2}\right),
 \end{equation}
 and each constituent's position $\mathbf{x_i}$ in Eqn.~\eqref{eq:rho} is sampled from a Gaussian radial distribution of width $r$.
-In the limit $M\to \infty$, the composite proton's root mean square radius is simply the convolution of its sampling radius and constituent width, which add together in quadrature
-\begin{equation}
-  \lim_{M\to\infty} \mathrm{RMS}\{\rho(\mathbf{x})\} = \sqrt{r^2 + w^2}.
-\end{equation}
-We therefore choose to reparametrize the sampling radius $r$ in terms of the variable $\w = \sqrt{r^2 - w^2}$ which approximates the rms ``width'' of the proton when the number of sampled constituents is large.
 
@@
   \label{eq:fluctuated_thick}
-  \T_{A, B}(\x) \equiv \int dz\, \frac{1}{M} \sum\limits_{i=1}^M \gamma_i\, \rho_c \,(\mathbf{x} - \mathbf{x_i} \pm \mathbf{b}/2),
+  \T_{A, B}(\x) \equiv \int dz\, \frac{1}{n_c} \sum\limits_{i=1}^{n_c} \gamma_i\, \rho_c \,(\mathbf{x} - \mathbf{x_i} \pm \mathbf{b}/2),
 \end{equation}
@@
   \caption{Effect of nucleon substructure on the nuclear thickness function $T(x, y) \equiv \int dz\, \rho(x, y, z)$ of a $^{208}\mathrm{Pb}$ nucleus.
-The nucleus on the left has Gaussian nucleons of width $0.88$~fm, while the nucleus on the right has composite nucleons, each containing six Gaussian constituents of width $0.41$~fm.}
+The nucleus on the left has Gaussian nucleons of width $0.8$~fm, while the nucleus on the right has composite nucleons, each containing six constituents of width $0.4$~fm.}
   \label{fig:thickness}
@@
 The nuclear collision model constructed in Sec.~\ref{sec:model} includes a number of free parameters $\x$ which describe the initial state, pre-equilibrium dynamics, and hydrodynamic medium.
-Given values for the parameters $\x$, the model may be used to predict a vector of experimental observables $\y$.
-For example, $\y$ might be a vector of fixed length which includes charged particle yields in different centrality bins.
-The physics model thus represents a vector-valued function $f(\x) = \y$
-which maps the parameter values $\x$ to the observable state $\y$.
+Given values for the parameters $\x$, the model may be used to predict a vector of simulated experimental observables $\y_m$.
+For example, $\y_m$ might be a vector consisting of charged particle yields in different centrality bins.
+The physics model thus represents a vector-valued function $f(\x) = \y_m$ which maps the parameter values $\x$ to the simulated observables $\y_m$.
 
-The goal of this work is to estimate the true model parameters $\x_\star$, given some evidence that our model ${f: X \mapsto Y}$ describes the experimental data $\yexp$ for the simulated observables $\y$.
-According to Bayes' theorem, the probability for the true model parameters $\x_\star$ is
+The goal of this work is to estimate the true model parameters $\x_\star$ provided some evidence that our model predictions $\y_m$ describe experimental measurements $\y_e$.
+The problem involves three distinct components:
+\begin{enumerate}[itemsep=0pt, leftmargin=2\parindent]
+  \item $H_f$: the hypothesis that the nuclear collision model $f$ formulated in this work provides a realistic description of reality,
+  \item $H_\x$: the hypothesis that the model parameters $\x$ are the true model parameters $\x_\star$ of $f$, and
+  \item $E$: the evidence provided by the experimental data $\y_e$ and its corresponding uncertainties.
+\end{enumerate}
+As a practical matter, we always assume that hypothesis $H_f$ is correct, meaning there are no glaring flaws in our chosen theoretical framework.
+This is a significant assumption; the application of hydrodynamic simulations to small collision systems is speculative, and our conclusions are conditional on the framework making sense.
+
+Subject to this assumption, we can apply Bayes' theorem to evaluate the hypothesis $H_\x$ for the true model parameters
 \begin{equation}
   \label{eq:bayes}
-  P(\x_\star|f, \yexp) \propto P(f, \yexp|\x_\star) P(\x_\star).
+  P(H_\x | E) \propto P(E | H_\x)\, P(H_\x).
 \end{equation}
-The left-side of this expression is the \emph{posterior}: the probability that $\x_\star$ is the correct value given evidence from the model $f: X \mapsto Y$ and the experimental data $\yexp$.
+The left-side of this expression is the \emph{posterior}: the probability for hypothesis $H_\x$ given the experimental evidence $E$.
+On the right-side there are two separate terms.
+The first term $P(E | H_\x)$ is the \emph{likelihood} function: the probability of observing the experimental evidence $E$ given our model and the hypothesis $H_\x$ for the true model parameters $\x_\star$, and the second term $P(H_\x)$ is the \emph{prior}: an estimate of the probability of hypothesis $H_\x$ in the absence of evidence $E$.
 
-On the right-side, the \emph{likelihood} function $P(f, \yexp | \x_\star)$ is the compatibility of the experimental evidence $\yexp$ with the hypothetical model parameters $\x_\star$, and the \emph{prior} probability $P(\x_\star)$ is the estimate of the probability of $\x_\star$ in the absence of evidence $\yexp$.
+We assume that the likelihood function in Eqn.~\eqref{eq:bayes} is described by a multivariate normal distribution
+\begin{equation}
+  \label{eq:likelihood}
+  P(E | H_\x) = \frac{1}{\sqrt{(2\pi)^m \det \Sigma}} \exp \left ( -\frac{1}{2}\Delta\y^\intercal \Sigma^{-1} \Delta\y \right ),
+\end{equation}
+where $\Delta\y = \y_m(\x) - \y_e$ is the discrepancy of the model and experiment, and $\Sigma$ is the total covariance matrix which includes all known sources of uncertainty for the simulated and measured observables.
 
-\begin{equation}
-  P(H1 | E, H2) \propto P(E | H1, H2)\, P(H1 | H2)
-\end{equation}
-There are three types of events,
-\begin{itemize}
-  \item $H1$: hypothesis that the model parameters $\x_\star$ are the true model parameters.
-  \item $H2$: hypothesis that nuclear collisions are describable using the hybrid transport model formulated in this work.
-  \item $E$: experimental evidence provided by data $\yexp$.
-\end{itemize}
-
-\subsection{Parameter design}
-
-We create a scaffolding of the parameter space and distribute $n=500$ parameter points in our 15-dimensional parameter space using a Latin hypercube design \note.
-The algorithm maximizes the
-
-We assume a flat prior distribution for each model parameter $x \in \x$ described by the range of values shown in Tbl.~\ref{tab:design}.
-The minimum and maximum values of the parameters are conservative to avoid clipping the calibrated posterior; for example, previous analyses of \trento\ model found $p \sim 0$, but we use the prior $-1 < p < 1$ to account for physics differences in the present model which might modify its posterior.
-
-\subsection{Gaussian process emulator}
-
-\begin{table}
-  \caption{Input parameter ranges for the initial condition and hydrodynamic models.}
+\begin{table}[t]
+  \caption{Input parameter ranges for the physics model.}
   \begin{ruledtabular}
@@
     $\sigmaf$         & Nucleon fluctuation std.\ dev.\      & 0--2            \\
-    $\w$              & Constituent sampling width           & 0.4--1.2 fm     \\
-    $\n$              & Number of constituents               & 1--9            \\
+    $w$               & Nucleon width parameter              & 0.4--1.2 fm     \\
+    $n_c$             & Number of nucleon constituents       & 1--9            \\
     $\X$              & Nucleon structure parameter          & 0--1            \\
@@
 \end{table}
+
+\subsection{Parameter design}
+
+For the prior $P(H_\x)$, we specify ranges, i.e.\ minimum and maximum values, for each parameter which are listed in Tbl.~\ref{tab:design}.
+We assume the prior distribution is constant and nonzero within each specified range and zero otherwise.
+The selected parameter ranges are intentionally wide to avoid clipping the calibrated posterior; for example, a previous analysis of the \trento\ model \cite{Bernhard:2016tnd} found $p \sim 0$, but we use a prior range $p \in [-1, 1]$ to account for differences in the present model, e.g.\ nucleon substructure, which could modify its posterior.
+Several of the model parameters require special care and are reparametrized accordingly:
+\begin{itemize}[leftmargin=2\parindent]
+  \item The constituent fluctuations, modeled by the gamma random weights $\gamma_i$ in Eqn.~\eqref{eq:fluctuated_thick}, generate overall nucleon fluctuations which are suppressed by the number of constituents $n_c$ inside the nucleon.
+    The observed nucleon fluctuation variance falls like $1 / n_c$ which means the natural range for the constituent fluctuations is larger when the number of constituents is larger and \emph{vice versa}.
+    We therefore reparametrize the constituent fluctuations using the standard deviation of the effective nucleon fluctuations
+    \begin{equation}
+      \sigma_\mathrm{fluct} = 1 / \sqrt{k\, n_c},
+    \end{equation}
+    where $k$ is the shape parameter of the gamma distribution used to fluctuate each constituent.
+  \item In Sec.~\ref{sec:initial_state} we parametrized nucleon substructure using three degrees of freedom:
+    \begin{enumerate*}[label=(\roman*)]
+      \item a parameter $v$ for the Gaussian width of each constituent,
+      \item a parameter $r$ for the Gaussian width of the radial distribution used to sample the constituent centers, and
+      \item a parameter $n_c$ to vary the number of constituents inside the nucleon.
+    \end{enumerate*}
+    In the limit $n_c \to \infty$, the composite nucleon's root mean square (rms) radius is simply the convolution of its sampling radius $r$ and constituent width $v$ which add together in quadrature
+  \begin{equation}
+    \lim_{n_c \to\infty} \mathrm{RMS}\{\rho(\mathbf{x})\} = \sqrt{r^2 + v^2}.
+  \end{equation}
+  We therefore choose to reparametrize the sampling radius $r$ in terms of a new variable
+  \begin{equation}
+    \label{eq:nucleon_width}
+    w = \sqrt{r^2 + v^2},
+  \end{equation}
+  visualized in Fig.~\ref{fig:nucleon_schematic}, which approximates the rms radius of the nucleon when the number of sampled constituents is large.
+  We thus call $w$ a nucleon ``width'', although for smaller numbers of constituents, the \emph{actual} rms radius of our sampled nucleons can be significantly smaller than our width parameter $w$ due to fluctuations in the nucleon's center of mass, and hence once should account for these difference when discussing the nucleon's rms radius.
+  We account for this effect when discussing the nucleon width posterior distribution later in the text.
+
+\begin{figure}[t]
+  \begin{tikzpicture}
+    \draw[dashed] (0,0) circle (1.6cm);
+    \foreach \theta in {60, 180, 300}{
+      \draw ({\theta}:.8) circle (.65cm);
+    }
+    \coordinate (a) at (0, 0);
+    \coordinate (b) at (.8, 1.385);
+    \coordinate (c) at (.4, -.693);
+    \coordinate (d) at (-1.45, 0);
+    \coordinate (e) at (-.8, 0);
+    \draw[<->] (a) -- (b) node[midway, xshift=-1ex, yshift=.8ex]{$w$};
+    \draw[<->] (a) -- (c) node[midway, xshift=1ex, yshift=.8ex]{$r$};
+    \draw[<->] (d) -- (e) node[midway, yshift=1ex]{$v$};
+  \end{tikzpicture}
+  \caption{Schematic illustrating the constituent sampling radius $r$, constituent width $v$ and nucleon width parameter $w = \sqrt{r^2 + v^2}$ for a nucleon with $n_c=3$ constituents.}
+  \label{fig:nucleon_schematic}
+\end{figure}
+
+\begin{table*}
+  \caption{
+    \label{tab:observables}
+    Experimental data to be compared with model calculations.
+  }
+  \begin{ruledtabular}
+  \begin{tabular}{lcccc}
+    Collision system & Observable & Particle selection & Event bins & Ref. \\
+    \paddedhline
+    \multirow{2}{*}{Pb-Pb 5.02 TeV} & Yield $d\nch/d\eta$ & chg \& $|\eta| < 0.5$ & centrality: 0--5, 5--10, 10--20, \ldots, 60--70 & \note \\
+    \noalign{\smallskip}
+    & Flow cumulants $\vnk{n}{2}$ & chg \& $|\eta| < 1$ & centrality: 0--5, 5--10, 10--20, \ldots, 60--70 & \note \\
+    \noalign{\smallskip}
+    %Mean transverse momentum $\langle p_T \rangle$ & $\pi^\pm$, $K^\pm$, $p\bar p$ &
+  \end{tabular}
+  \end{ruledtabular}
+\end{table*}
+
+\item Equation~\eqref{eq:nucleon_width} requires the nucleon width to be larger than the constituent width, i.e.
+$w > v$, lest the sampling radius $r$ turn imaginary, and thus we cannot vary the nucleon width $w$ and the constituent width $v$ independently.
+  We therefore reparametrize the constituent width $v$ using a new variable $\X$ to interpolate between minimum and maximum allowed values:
+  \begin{equation}
+    v = v_\mathrm{min} + \X (v_\mathrm{max} - v_\mathrm{min}).
+  \end{equation}
+  Here we choose a minimum width $v_\mathrm{min} = 0.2$~fm determined by computational limits and a maximum width $v_\mathrm{max} = w$ equal to the nucleon width.
+  Thus for $\X=0$, the nucleons consist of small, distinct hot spots, whereas for $\X=1$ the nucleon is a single Gaussian blob of width $w$.
+  The parameter $\X$ thus varies the granularity of the nucleon while keeping the number of constituents $n_c$ and approximate size of the nucleon $w$ fixed.
+\end{itemize}
+
+\subsection{Calibration observables}
+
+We focus on simple nuclear collision observables which are sensitive to the bulk properties of the produced medium and which can be calculated using a relatively small $\mathcal{O}(10^4)$ number of events.
+
+\subsection{Model emulator}
+
+In principle, one could calculate the likelihood function in Eqn.~\eqref{eq:likelihood} directly, e.g.\ by running the model a large number of times at a given parameter point $\x$ to calculate the model observables $\y_m(\x)$ from the ensemble of simulated events, but in practice such a procedure would be intractable.
+The model is computationally intensive to evaluate, ${\sim}10$~min/event per processor core, and thousands of events are required to calculate the simplest observables for a single parameter point.
+Moreover, we need to evaluate Eqn.~\eqref{eq:likelihood} \emph{numerous} times in order to sample and visualize the multidimensional posterior distribution.
+
+We therefore follow an established framework for computationally intensive models and train a Gaussian process emulator (GPE) to act as a fast surrogate for direct model evaluation.
+The GPE enables essentially instantaneous predictions $\y_m = f(\x)$ which allows us to sample the posterior distribution millions of times.
+With the posterior parameter samples $\{\x_\star\}$ in hand, we simply histogram the samples to marginalize over (integrate out) all but one or two model parameters.
+We can then plot the flattened distribution to examine the constraints on each individual model parameter.
+
+In order to train the GPE, we first generate a scaffolding of the parameter space using a maximin Latin hypercube design to distribute 500 points in our 15-dimensional parameter space according to the ranges in Tbl.~\ref{tab:design}.
+We then run $\mathcal{O}(10^4)$ $p$-Pb and Pb-Pb events at each design point, and calculate the model observables $\y_m$---discussed in the next subsection---from the ensemble of events.
+
+\subsection{Experimental }
 
@@
 \fig{posterior_proton_shape}{
-  Posterior distribution for the constituent sampling radius $\R$ and constituent width $\w$.
+%  Posterior distribution for the constituent sampling radius $\R$ and constituent width $\w$.
 The white trapezoidal region (lower-right) covers the initial range of allowed values for $\R$ and $\w$, while the gray triangular region (upper-left) indicates values which are excluded by the prior $\w > \R$.

### 2018-07-16 paper/substructure.tex
Wrapping up experimental observables section

@@
 \definecolor{theblue}{RGB}{0,50,230}
+
+\usepackage{multirow}
 
@@
 \newcommand{\nch}{N_\text{ch}}
+\newcommand{\ntrk}{N_\text{trk}^\text{offline}}
 \newcommand{\sqrts}{\sqrt{s_{NN}}}
@@
 \newcommand{\note}{\textcolor{theblue}{[?]}}
+
+\renewcommand\labelitemi{$\vcenter{\hbox{\scriptsize$\bullet$}}$}
 
@@
 
-Typically, the macroscopic structure of heavy nuclei, characterized e.g.\ by an atomic mass and set of Woods-Saxon coefficients \cite{MOLLER1995185, DEVRIES1987495}, is regarded as a known input to hydrodynamic models which contributes negligible uncertainty to simulation predictions, outweighed by large uncertainties in modeling initial energy deposition and off-equilibrium dynamics \cite{Niemi:2014lha, Song:2011hk, Retinskaya:2013gca, Liu:2015nwa, Kurkela:2016vts}.
+Typically, the macroscopic structure of heavy nuclei, characterized e.g.\ by an atomic mass number and set of Woods-Saxon coefficients \cite{MOLLER1995185, DEVRIES1987495}, is regarded as a known input to hydrodynamic models which contributes negligible uncertainty to simulation predictions, outweighed by large uncertainties in modeling initial energy deposition and off-equilibrium dynamics \cite{Niemi:2014lha, Song:2011hk, Retinskaya:2013gca, Liu:2015nwa, Kurkela:2016vts}.
 The geometry of light ions, meanwhile, is naturally more sensitive to the detailed size and shape of individual protons and neutrons inside the nucleus, which may fluctuate event-by-event and differ signficantly from the round blobs typically used to approximate nucleons in heavy-ion collisions \cite{Schenke:2014zha, Welsh:2016siu, Moreland:2017kdx, Schenke:2014gaa, Schlichting:2014ipa}.
@@
 Within a picture of independent pairwise collisions between the constituents, a Glauber model model may be used to calculate the probability $P_\mathrm{coll}$ that the two protons collide inelastically at impact parameter $b$.
-In the limit when the number of constituents $n_c \gg 1$, it yields the particularly simple form
+In the limit when the number of constituents is large, it yields the particularly simple form
 \begin{equation}
@@
 \end{equation}
-and each constituent's position $\mathbf{x_i}$ in Eqn.~\eqref{eq:rho} is sampled from a Gaussian radial distribution of width $r$.
+and each constituent's position $\mathbf{x_i}$ in Eqn.~\eqref{eq:rho} is sampled from a Gaussian radial distribution with standard deviation $r$.
 
@@
 Conveniently, it has been shown that the generalized mean ansatz is able to mimic the scaling behavior of certain saturation based initial condition models \cite{Bernhard:2016tnd}, and hence it should serve as a reasonable parametric form for exploring QGP entropy deposition, assuming imperfect knowledge of saturation effects in nature.
-Of course the model is \emph{not} a substitute for first principle saturation calculations, and may fail to reproduce nuanced features of \emph{ab initio} saturation theory calculations, such as the existence of short-range gluon field fluctuations \cite{Schenke:2012wb}.
+Of course the model is \emph{not} a substitute for first principle saturation calculations, and may fail to reproduce nuanced features of \emph{ab initio} models such as the existence of short-range gluon field fluctuations \cite{Schenke:2012wb}.
 
@@
   \label{hydro-eqn}
-  T^{\mu\nu} = e u^\mu u^\nu - (P + \Pi) \Delta^{\mu\nu} + \pi^{\mu\nu}
+  T^{\mu\nu} = e u^\mu u^\nu - (P + \Pi) \Delta^{\mu\nu} + \pi^{\mu\nu},
 \end{equation}
-where $e$ and $P$ are the energy density and pressure in the local fluid rest frame, $u^\mu$ is the local fluid velocity, ${\Delta^{\mu\nu} \equiv g^{\mu\nu} - u^\mu u^\nu}$, and $\Pi$ and $\pi^{\mu\nu}$ are the bulk pressure and shear stress tensor respectively.
+where $e$ and $P$ are the energy density and pressure in the local fluid rest frame, $u^\mu$ is the local fluid velocity, ${\Delta^{\mu\nu} \equiv g^{\mu\nu} - u^\mu u^\nu}$ is the projector onto the space orthogonal to $u^\mu$, and $\Pi$ and $\pi^{\mu\nu}$ are the bulk pressure and shear stress tensor respectively.
 We then solve for the energy density $e$ and fluid velocity $u^\mu$ using the Landau matching condition which defines the fluid rest frame velocity as the time-like eigenvector of $T^{\mu\nu}$ with energy density $e$ as its eigenvalue,
@@
 We use a parametrization for $P(e)$ that matches a hadron resonance gas EoS at low temperature to a lattice QCD EoS at high temperature by smoothly connecting their trace anomalies in the interval
-$165 \le T \le 200$~MeV.
+$165 \le T \le 200$~MeV \cite{Bernhard:2018hnz}.
 For the lattice EoS, we use a calculation by the HotQCD collaboration for (2+1)-flavor QCD which was extrapolated to the continuum limit \cite{Bazavov:2014pvz}.
@@
     The observed nucleon fluctuation variance falls like $1 / n_c$ which means the natural range for the constituent fluctuations is larger when the number of constituents is larger and \emph{vice versa}.
-    We therefore reparametrize the constituent fluctuations using the standard deviation of the effective nucleon fluctuations
+    We therefore reparametrize the constituent fluctuations using the standard deviation of the resulting nucleon fluctuations
     \begin{equation}
@@
   \end{equation}
-  visualized in Fig.~\ref{fig:nucleon_schematic}, which approximates the rms radius of the nucleon when the number of sampled constituents is large.
-  We thus call $w$ a nucleon ``width'', although for smaller numbers of constituents, the \emph{actual} rms radius of our sampled nucleons can be significantly smaller than our width parameter $w$ due to fluctuations in the nucleon's center of mass, and hence once should account for these difference when discussing the nucleon's rms radius.
-  We account for this effect when discussing the nucleon width posterior distribution later in the text.
+  depicted in Fig.~\ref{fig:nucleon_schematic}, which approximates the rms radius of the nucleon when the number of sampled constituents is large.
+  We thus call $w$ a nucleon ``width'', although for smaller numbers of constituents, the \emph{actual} rms radius of our sampled nucleons can be significantly smaller than our width parameter $w$ due to fluctuations in the nucleon's center of mass, and hence one should account for the difference when discussing the nucleon's posterior rms radius.
 
@@
     \label{tab:observables}
-    Experimental data to be compared with model calculations.
+    Experimental data used to calibrate the model.
   }
   \begin{ruledtabular}
-  \begin{tabular}{lcccc}
-    Collision system & Observable & Particle selection & Event bins & Ref. \\
+  \begin{tabular}{cc}
+    Pb-Pb $\sqrts=5.02$~TeV & $p$-Pb $\sqrts=5.02$~TeV \\
     \paddedhline
-    \multirow{2}{*}{Pb-Pb 5.02 TeV} & Yield $d\nch/d\eta$ & chg \& $|\eta| < 0.5$ & centrality: 0--5, 5--10, 10--20, \ldots, 60--70 & \note \\
+    Charged-particle multiplicity $d\nch/d\eta$, $|\eta| < 0.5$ \cite{Adam:2015ptt} & Charged-particle multiplicity $d\nch/d\eta$, $|\eta| < 1.4$ \cite{Adam:2014qja} \\
     \noalign{\smallskip}
-    & Flow cumulants $\vnk{n}{2}$ & chg \& $|\eta| < 1$ & centrality: 0--5, 5--10, 10--20, \ldots, 60--70 & \note \\
+  Two-particle flow cumulants  $\vnk{n}{2}$ for $n=2,3,4$, $|\eta| < 0.8$,  & Two-particle flow cumulants $\vnk{n}{2}$ for $n=2,3$, $|\eta| < 2.4$, \\
+    charged-particles, $|\Delta\eta| > 1$,\, $0.2 < p_T < 5.0$~GeV \cite{Adam:2016izf} & charged-particles, $|\Delta\eta| > 2$,\, $0.3 < p_T < 3.0$~GeV \cite{Chatrchyan:2013nka}\\
     \noalign{\smallskip}
-    %Mean transverse momentum $\langle p_T \rangle$ & $\pi^\pm$, $K^\pm$, $p\bar p$ &
+    & Charged-particle mean $p_T$, $0.15 < p_T < 10$~GeV, $|\eta| < 0.3$ \cite{Abelev:2013bla}\\
   \end{tabular}
@@
 
-\subsection{Calibration observables}
+\subsection{Simulated and measured observables}
 
-We focus on simple nuclear collision observables which are sensitive to the bulk properties of the produced medium and which can be calculated using a relatively small $\mathcal{O}(10^4)$ number of events.
+The likelihood function \eqref{eq:likelihood} provides evidence for \mbox{(or against)} the model parameters $\x$ by comparing the model predictions $\y_m(\x)$ to experimental data $\y_e$.
+We focus on simple experimental observables in the present study which are sensitive to the bulk properties of the produced medium.
+We calculate for each set of model parameters the following observables at midrapidity:
+\begin{itemize}[leftmargin=2\parindent, itemsep=0pt]
+  \item Charged-particle multiplicity $d\nch/d\eta$ \cite{Adam:2015ptt, Adam:2014qja}.
+  \item Identified particle yields $dN/dy$ of pions, kaons, and protons.
+  \item Transverse energy production $dE_T/d\eta$.
+  \item Charged particle mean transverse momentum $\langle p_T \rangle$ ($0.15 < p_T < 10$~GeV) \cite{Abelev:2013bla}.
+  \item Identified particle mean transverse momentum $\langle p_T \rangle$ of pions, kaons and protons.
+  \item Mean transverse momentum fluctuations $\delta p_T / \langle p_T \rangle$ (charged particles, $0.15 < p_T < 2.0$~GeV).
+  \item Two-particle flow cumulants $\vnk{n}{2}$ for $n=2,3,4$\\ (charged particles, $0.2 < p_T < 5.0$~GeV \cite{Adam:2016izf}, and $0.3 < p_T < 3.0$~GeV \cite{Chatrchyan:2013nka}).
+\item Four-particle flow cumulant $\vnk{2}{4}$\\ (charged particles, $0.2 < p_T < 5.0$) \cite{Adam:2016izf}.
+  \item Symmetric cumulants $\mathrm{SC}(4, 2)$ and $\mathrm{SC}(3,2)$.
+\end{itemize}
+Each observable is calculated from the list of final state particles produced by \texttt{UrQMD} using the same methods applied by experiment.
+We generally match the kinematic cuts of all measurements with two exceptions: we use use a larger rapidity interval $|\eta| < 0.8$ than experiment for some boost-invariant observables to improve our finite particle statistics, and we do not apply a finite rapidity gap, e.g.\ $|\Delta \eta| > 1$, between pairs of particles when calculating the two-particle cumulant $\vnk{n}{2}$ since we already oversample particles from each hydrodynamic event, and this oversampling suppresses non-flow correlations.
+
+At the time of this writing, many of the aforementioned experimental observables are not yet published for $p$-Pb and Pb-Pb collisions at $\sqrts=5.02$~TeV.
+We therefore restrict our calibration to the subset of measured and published observables listed in Tbl.~\ref{tab:observables}.
+The Pb-Pb four-particle cumulants $\vnk{n}{4}$ at $\sqrts=5.02$~TeV which are noticeably absent \emph{are} currently available, but we exclude from the present calibration for statistical reasons.
+
+Most of the calibration observables listed in Tbl.~\ref{tab:observables} are calculated as a function of collision centrality, where centrality is defined as some measure of the underlying event activity, e.g.\ the charged particle yield in a given rapidity window.
+When calculating these observables, we generate $\mathcal{O}(10^4)$ minimum-bias events at each design point and divide the events into centrality bins using the charged particle yield at midrapidity, similar to the procedure used by experiment.
+
+However, for some observables such as $p$-Pb mean $p_T$ and flow cumulants, the experiments use a special high-multiplicity trigger to select rare, ultra-central events according to the number of charged particles produced or detector tracks offline \note.
+These high-multiplicity bins are too selective for our modest minimum-bias event sample, and so a different procedure is required.
+We exploit, for this purpose, the approximate correspondence
+\begin{equation}
+  \ntrk \sim \nch \propto \int d^2x\, n(\x)
+\end{equation}
+between each event's initial transverse parton density $n(\x)$ and its final-state event activity, characterized by $\nch$ or $\ntrk$.
+
+and the event's initial parton density $n(\x)$ in Eqn.~\eqref{energy-momentum}, and run a high-multiplicity trigger on the relative event activity
+\begin{equation}
+  \mathrm{bin~low} \le \nch / \langle \nch \rangle \le \mathrm{bin~high}.
+\end{equation}
+
+\fig*{correlation_matrices}{
+  Correlation matrices for Pb-Pb collisions at $\sqrts=5.02$~TeV for the model/emulator (left) and experimental data (right) used to construct the covariance matrix $\Sigma = \Sigma_\mathrm{model} + \Sigma_\mathrm{expt}$ in Eqn.~\eqref{eq:likelihood}.
+  Experimental statistical and systematic errors are from ALICE and CMS \note.
+  The experimental correlation structure is modeled using Eqn.~?.
+}
 

### 2018-07-16 src/expt.py

@@ -213,7 +213,7 @@ def pPb5020_yield():
 
     """
     eta_beam = -0.465
-    eta_cut = 1.4
+    eta_cut = 0.5
 
     # use the V0M centrality estimator
     dset = HEPData(1335350, 2)
@@ -351,9 +351,15 @@ def _data():
         data['PbPb5020']['dNch_deta'] = {None: HEPData(1410589, 2).dataset(name)}
 
         # PbPb5020 flows
-        system, tables_nk = ('PbPb5020', [(1, [(2, 2)]), (2, [(3, 2), (4, 2)]),])
+        system, tables_nk = (
+            'PbPb5020', [
+                (1, [(2, 2), (2, 4)]),
+                (2, [(3, 2), (4, 2)]),
+            ]
+        )
 
         data[system]['vnk'] = {}
+
         for table, nk in tables_nk:
             d = HEPData(1419244, table)
             for n, k in nk:

### 2018-07-16 src/model.py

@@ -330,61 +330,31 @@ def _data(system, dataset='main'):
     data = expt.data[system]
     species = ['pion', 'kaon', 'proton']
 
-    # training data is only missing for Pb-Pb charged particle mean pT
-    #if dataset == 'main' and system == 'PbPb5020':
-    #    edges = np.linspace(0, 70, 15)
-    #    cent = [(a, b) for a, b in zip(edges[:-1], edges[1:])]
-    #    x = [(a + b)/2 for a, b in cent]
-    #    empty = dict(cent=cent, x=x)
-    #    data = dict({'mean_pT': {None: empty}}, **data)
+    def obs_like(d):
+        obs = d.copy()
+        for k in ('y', 'yerr'):
+            del obs[k]
+        return obs
 
-    # calculate additional observables for the maximum a posteriori parameters
-    #if dataset == 'map':
+    # missing yields
+    empty = obs_like(data['dNch_deta'][None])
+    data = dict({'dET_deta': {None: empty}}, **data)
+    data = dict({'iden_dN_dy': {s: empty for s in species}}, **data)
 
-    # p-Pb map observables
-    if system == 'pPb5020':
+    # missing mean pT
+    if 'mean_pT' in data:
+        empty = obs_like(data['mean_pT'][None])
+    else:
+        empty = obs_like(data['vnk'][(2, 2)])
+        data = dict({'mean_pT': {s: empty for s in species}}, **data)
 
-        # yields
-        cent, x = [data['dNch_deta'][None][k] for k in ('cent', 'x')]
-        empty = dict(cent=cent, x=x)
-        data = dict({'dET_deta': {None: empty}}, **data)
-        data = dict({'iden_dN_dy': {s: empty for s in species}}, **data)
+    data = dict({'iden_mean_pT': {s: empty for s in species}}, **data)
+    data = dict({'pT_fluct': {None: empty}}, **data)
 
-        # mean pT
-        mult, x = [data['mean_pT'][None][k] for k in ('mult', 'x')]
-        empty = dict(mult=mult, x=x)
-        data = dict({'iden_mean_pT': {s: empty for s in species}}, **data)
-        data = dict({'pT_fluct': {None: empty}}, **data)
-
-        # symmetric cumulants
-        mult, x = [data['vnk'][(2, 2)][k] for k in ('mult', 'x')]
-        empty = dict(mult=mult, x=x)
[…]

### 2018-07-16 src/plots.py

@@ -553,9 +553,18 @@ def observables_map():
         'Flow cumulants': (0, .12),
     }
 
-    for p in plots:
+    for n, p in enumerate(plots):
         p['ylim'] = ylim[p['title']]
-
+        if p['title'] == 'Flow cumulants':
+            move_index = n
+            p.update(
+                ylabel=r'$v_n\{k\}$',
+                subplots=[
+                    ('vnk', nk, dict(label='$v_{}\{{{}\}}$'.format(*nk)))
+                    for nk in [(2, 2), (2, 4), (3, 2), (4, 2)]
+                ],
+                legend=True
+            )
     fig = plt.figure(figsize=figsize(1.1, 1.5))
 
     yields, mean_pT, mean_pT_fluct, flows = [
@@ -831,16 +840,17 @@ def find_map():
 def flow_corr():
     """
     Symmetric cumulants SC(m, n) at the MAP point compared to experiment.
+
     """
     fig, axes = plt.subplots(
-        figsize=figsize(1.1, .7),
-        nrows=2, ncols=2, gridspec_kw=dict(width_ratios=[.8, 1])
+        figsize=figsize(0.5, 1.2), sharex=True,
+        nrows=2, ncols=1
     )
 
-    systems = ['PbPb5020', 'PbPb5020']
-
-    cmapx_normal = .7
-    cmapx_pred = .5
+    observables = ['sc', 'sc_normed']
+    ylims = [(-2.5e-6, 2.5e-6), (-0.3, 0.8)]
+    labels = ['(4,2)', '(3,2)']
+    system = 'PbPb5020'
 
     def label(*mn, normed=False):
         fmt = r'\mathrm{{SC}}({0}, {1})'
@@ -848,88 +858,37 @@ def flow_corr():
             fmt += r'/\langle v_{0}^2 \rangle\langle v_{1}^2 \rangle'
         return fmt.format(*mn).join('$$')
 
-    for obs, ax in zip(
-            ['sc_central', 'sc', 'sc_normed_central', 'sc_normed'],
-            axes.flat
-    ):
-        for (mn, cmap), sys in itertools.product(
-                [
-                    ((4, 2), 'Blues'),
-                    ((3, 2), 'Oranges'),
-                ],
-                systems
-        ):
-            x = model.map_data[sys][obs][mn]['x']
-            y = model.map_data[sys][obs][mn]['Y']
-
-            pred = obs not in expt.data[sys]
[…]

### 2018-07-17 paper/substructure.tex
Start writing about the emulator

@@
 
-\subsection{Simulated and measured observables}
+\subsection{Observables}
 
@@
 \begin{itemize}[leftmargin=2\parindent, itemsep=0pt]
-  \item Charged-particle multiplicity $d\nch/d\eta$ \cite{Adam:2015ptt, Adam:2014qja}.
+  \item Charged-particle multiplicity $d\nch/d\eta$.
   \item Identified particle yields $dN/dy$ of pions, kaons, and protons.
   \item Transverse energy production $dE_T/d\eta$.
-  \item Charged particle mean transverse momentum $\langle p_T \rangle$ ($0.15 < p_T < 10$~GeV) \cite{Abelev:2013bla}.
+  \item Charged particle mean transverse momentum $\langle p_T \rangle$ ($0.15 < p_T < 10$~GeV).
   \item Identified particle mean transverse momentum $\langle p_T \rangle$ of pions, kaons and protons.
   \item Mean transverse momentum fluctuations $\delta p_T / \langle p_T \rangle$ (charged particles, $0.15 < p_T < 2.0$~GeV).
-  \item Two-particle flow cumulants $\vnk{n}{2}$ for $n=2,3,4$\\ (charged particles, $0.2 < p_T < 5.0$~GeV \cite{Adam:2016izf}, and $0.3 < p_T < 3.0$~GeV \cite{Chatrchyan:2013nka}).
-\item Four-particle flow cumulant $\vnk{2}{4}$\\ (charged particles, $0.2 < p_T < 5.0$) \cite{Adam:2016izf}.
+  \item Two-particle flow cumulants $\vnk{n}{2}$ for $n=2,3,4$\\ (charged particles, $0.2 < p_T < 5.0$~GeV, and \\$0.3 < p_T < 3.0$~GeV).
+  \item Four-particle flow cumulant $\vnk{2}{4}$ \\(charged particles, $0.2 < p_T < 5.0$).
   \item Symmetric cumulants $\mathrm{SC}(4, 2)$ and $\mathrm{SC}(3,2)$.
@@
 We therefore restrict our calibration to the subset of measured and published observables listed in Tbl.~\ref{tab:observables}.
-The Pb-Pb four-particle cumulants $\vnk{n}{4}$ at $\sqrts=5.02$~TeV which are noticeably absent \emph{are} currently available, but we exclude from the present calibration for statistical reasons.
+Notably absent from this list are the four-particle cumulants $\vnk{n}{4}$ at $\sqrts=5.02$~TeV despite being measured and published.
+Unfortunately, the four-particle cumulants require minimum-bias event statistics an order of magnitude larger than those used in this work.
+We therefore refrain from \emph{calibrating} on the four-particle cumulants, although we do show calculations of the four-particle cumulant $\vnk{2}{4}$ later in the text, using a single set of calibration parameters.
 
-Most of the calibration observables listed in Tbl.~\ref{tab:observables} are calculated as a function of collision centrality, where centrality is defined as some measure of the underlying event activity, e.g.\ the charged particle yield in a given rapidity window.
+Most of the calibration observables listed in Tbl.~\ref{tab:observables} are calculated as a function of collision centrality, where centrality is defined using some measure of the underlying event activity, e.g.\ the charged particle yield in a given rapidity window.
 When calculating these observables, we generate $\mathcal{O}(10^4)$ minimum-bias events at each design point and divide the events into centrality bins using the charged particle yield at midrapidity, similar to the procedure used by experiment.
 
-However, for some observables such as $p$-Pb mean $p_T$ and flow cumulants, the experiments use a special high-multiplicity trigger to select rare, ultra-central events according to the number of charged particles produced or detector tracks offline \note.
+However, for some observables such as $p$-Pb mean $p_T$ \cite{Abelev:2013bla} and flow cumulants $\vnk{n}{k}$ \cite{Chatrchyan:2013nka}, the experiments use a special high-multiplicity trigger to select rare, ultra-central events according to the number of charged particles produced or detector tracks offline.
 These high-multiplicity bins are too selective for our modest minimum-bias event sample, and so a different procedure is required.
@@
 \begin{equation}
-  \ntrk \sim \nch \propto \int d^2x\, n(\x)
+  \label{eq:bin_equiv}
+  \ntrk \sim \nch \propto N_\rho
 \end{equation}
-between each event's initial transverse parton density $n(\x)$ and its final-state event activity, characterized by $\nch$ or $\ntrk$.
+between each event's initial secondary-parton multiplicity ${N_\rho = \int d^2x\,n(\x)}$ in Eqn.~\eqref{energy-momentum} and its final-state event activity, characterized by $\nch$ or $\ntrk$ \cite{Abelev:2013bla, Chatrchyan:2013nka}.
 
-and the event's initial parton density $n(\x)$ in Eqn.~\eqref{energy-momentum}, and run a high-multiplicity trigger on the relative event activity
+Consider, for example, an experimental multiplicity bin $(\nch^\text{low}, \nch^\text{high})$ with some kinematic cuts on $p_T$ and $\eta$.
+We rescale this bin by the average multiplicity $\langle \nch \rangle$ of the corresponding minimum-bias event sample, i.e.\
 \begin{equation}
-  \mathrm{bin~low} \le \nch / \langle \nch \rangle \le \mathrm{bin~high}.
+  (\nch^\text{low}, \nch^\text{high}) \rightarrow (\nch^\text{low} / \langle \nch \rangle, \nch^\text{high} / \langle \nch \rangle)
+\end{equation}
+in order to reexpress each bin edge as a unitless variable.
+These rescaled bins are then used to select initial condition events using the equivalence \eqref{eq:bin_equiv}:
+\begin{equation}
+  \label{eq:mult_trigger}
+  \left (\frac{N^\text{low}_\rho}{\langle N_\rho \rangle}, \frac{N^\text{high}_\rho}{\langle N_\rho \rangle} \right) \leftrightarrow \left (\frac{\nch^\text{low}}{\langle \nch \rangle}, \frac{\nch^\text{high}}{\langle \nch \rangle} \right ).
 \end{equation}
 
+Finally, we mimic the method used by experiment and apply \eqref{eq:mult_trigger} to select rare high-multiplicity events from a continuous stream of minimum-bias \trento\ events that satisfy the correct relative multiplicity bin edges.
+This of course means that, in addition to running a large sample of minimum-bias events for centrality binned observables, we must also generate (much like experiment) a separate sample of multiplicity triggered events.
+In practice, we use a few hundred to a few thousand events per multiplicity bin, depending on the type of observable.
+
+We also take stock of the statistical and systematic errors reported by each experiment and incorporate their uncertainty into the likelihood covariance matrix
+\begin{equation}
+  \Sigma = \Sigma_m + \Sigma_e,
+\end{equation}
+which includes uncertainty contributions from both the model $\Sigma_m$ and experimental data $\Sigma_e$.
+The experimental contribution to the covariance $\Sigma_e$ is further broken down into statistical and systematic components,
+\begin{equation}
+  \Sigma_e = \Sigma_e^\text{stat} + \Sigma_e^\text{sys}.
+\end{equation}
+
+The statistical errors are uncorrelated and thus its covariance matrix is diagonal:
+\begin{equation}
+  \Sigma_e^\text{stat} = \text{diag}[(\sigma^\text{stat}_1)^2, (\sigma^\text{stat}_2)^2, \dots (\sigma^\text{stat}_m)^2 ],
+\end{equation}
+where $\sigma^\text{stat}_i$ is the statistical uncertainty of observable $y_i$ in the experimental observable vector $\y_e = [y_1, \dots, y_m]$.
+The systematic errors, meanwhile, are typically correlated, but the correlation structure is not reported by the experiments so we assert a reasonable form.
+We can expand the systematic covariance matrix as
+\begin{equation}
+  (\Sigma_e^\text{sys})_{ij} =  \rho_{ij} \sigma_i \sigma_j,
+\end{equation}
+where $\sigma_i$ and $\sigma_j$ are the systematic errors of observables $y_i$ and $y_j$ respectively, and $\rho_{ij}$ is the Pearson correlation coefficient between observable $y_i$ and $y_j$:
+\begin{equation}
+  \rho_{ij} = \frac{\text{cov}(y_i, y_j)}{\sigma_i \sigma_j},
+\end{equation}
+which satisfies $\rho_{ij}=1$ for $i=j$ and $|\rho_{ij}| \le 1$ for $i \ne j$.
+We assume that each observable is correlated across different centrality/multiplicity bins, and uncorrelated with observables of a different type, e.g.\ correlations between yields and flows.
+This is a crude simplifying assumption but it is better than neglecting the correlation structure of the experimental data entirely.
+
+For the correlation structure between different observable bins, we assert a simple Gaussian form
+\begin{equation}
+  \rho_{ij}^\text{sys} = \exp \left[ -\frac{1}{2} \left(\frac{b_i - b_j}{l} \right)^2 \right],
+\end{equation}
+where $b_i$ and $b_j$ are the midpoints of two observable bins of a single type (centrality or relative multiplicity) and $l$ is a correlation length which describes how quickly the observable bins decorrelate as the distance between the bins increases.
+We use centrality correlation lengths ${l=100\%}$ for all Pb-Pb observables and ${l=30\%}$ for the $p$-Pb charged particle yield $d\nch/d\eta$.
+The $p$-Pb mean $p_T$ and flow observables, meanwhile, use relative multiplicity bins $\nch / \langle \nch \rangle$ and $\ntrk / \langle \ntrk \rangle$ so we use a much smaller correlation length $l=5$ to reflect the smaller range of bin values.
+
 \fig*{correlation_matrices}{
-  Correlation matrices for Pb-Pb collisions at $\sqrts=5.02$~TeV for the model/emulator (left) and experimental data (right) used to construct the covariance matrix $\Sigma = \Sigma_\mathrm{model} + \Sigma_\mathrm{expt}$ in Eqn.~\eqref{eq:likelihood}.
+  Correlation matrices for Pb-Pb collisions at $\sqrts=5.02$~TeV for the model/emulator (left) and experimental data (right) used to construct the covariance matrix ${\Sigma = \Sigma_\mathrm{model} + \Sigma_\mathrm{expt}}$ in Eqn.~\eqref{eq:likelihood}.
   Experimental statistical and systematic errors are from ALICE and CMS \note.
@@
 In principle, one could calculate the likelihood function in Eqn.~\eqref{eq:likelihood} directly, e.g.\ by running the model a large number of times at a given parameter point $\x$ to calculate the model observables $\y_m(\x)$ from the ensemble of simulated events, but in practice such a procedure would be intractable.
-The model is computationally intensive to evaluate, ${\sim}10$~min/event per processor core, and thousands of events are required to calculate the simplest observables for a single parameter point.
-Moreover, we need to evaluate Eqn.~\eqref{eq:likelihood} \emph{numerous} times in order to sample and visualize the multidimensional posterior distribution.
+The model is computationally intensive to evaluate, ${\sim}10$~min/event per processor core, and thousands of events are required to calculate the simplest observables at a single parameter point.
+Moreover, we need to evaluate Eqn.~\eqref{eq:likelihood} \emph{numerous} times in order to sample the multidimensional posterior distribution so that the samples may be histogrammed and visualized.
 
-We therefore follow an established framework for computationally intensive models and train a Gaussian process emulator (GPE) to act as a fast surrogate for direct model evaluation.
-The GPE enables essentially instantaneous predictions $\y_m = f(\x)$ which allows us to sample the posterior distribution millions of times.
-With the posterior parameter samples $\{\x_\star\}$ in hand, we simply histogram the samples to marginalize over (integrate out) all but one or two model parameters.
-We can then plot the flattened distribution to examine the constraints on each individual model parameter.
+We therefore follow an established framework for computationally intensive models and train a Gaussian process emulator (GPE) to act as a fast surrogate for the full physics simulation \note.
+The GPE enables essentially instantaneous predictions for $\y_m = f(\x)$ which allows us to sample the posterior distribution millions of times.
+In order to train the GPE, we first generate a scaffolding of the parameter space using a maximin Latin hypercube design to distribute 500 points uniformly in our 15-dimensional parameter space according to the ranges in Tbl.~\ref{tab:design}.
+We then run minimum-bias and multiplicity triggered $p$-Pb and Pb-Pb events at each design point, and calculate the model observables from the ensemble of events.
 
-In order to train the GPE, we first generate a scaffolding of the parameter space using a maximin Latin hypercube design to distribute 500 points in our 15-dimensional parameter space according to the ranges in Tbl.~\ref{tab:design}.
-We then run $\mathcal{O}(10^4)$ $p$-Pb and Pb-Pb events at each design point, and calculate the model observables $\y_m$---discussed in the next subsection---from the ensemble of events.
+Let $X$ denote the $d \times m$ dimensional matrix of design points where $d=500$ is the number of design points and $m$ is the number of emulated observables $y \in \y_m$.
 

### 2018-07-20 paper/substructure.tex
Get figure placement correct

@@
 
-\usepackage{multirow}
-
 \usepackage[pdfencoding=auto, psdextra]{hyperref}
-
 \usepackage[font=small, justification=raggedright, labelsep=quad]{caption}
@@
 \newcommand{\y}{\mathbf{y}}
+\newcommand{\z}{\mathbf{z}}
 \newcommand{\vnk}[2]{v_#1\{#2\}}
@@
 \begin{figure}
-  \centering
   \begin{tikzpicture}
@@
 
-\fig[b]{coupling}{
-Cartoon of the free-streaming approximation for hydrodynamic initialization.
+\begin{figure}[b]
+  \includegraphics{coupling}
+  \caption{Cartoon of the free-streaming approximation for hydrodynamic initialization.
 The initial state is free-streamed for proper time $\taufs$ (zero-coupling) before it is matched to hydrodynamics (strong-coupling).
-This piecewise evolution approximates the more realistic scenario expected in nature where the medium's coupling strength smoothly changes as a function of time.
-}
+This piecewise evolution approximates the more realistic scenario expected in nature where the medium's coupling strength smoothly changes as a function of time.}
+  \label{fig:coupling}
+\end{figure}
 
@@
 
-\fig[t]{viscosity_dof}{Degrees of freedom in the temperature dependent shear and bulk viscosity parametrizations.
+\begin{figure}[t]
+  \includegraphics{viscosity_dof}
+  \caption{Degrees of freedom in the temperature dependent shear and bulk viscosity parametrizations.
 Lines are chosen for illustrative purposes only and do not represent all possible variability.
-For instance, $\eta/s$ could have a large slope and negative curvature, or $\zeta/s$ could have a large max and narrow width, neither of which are depicted above.
-}
+For instance, $\eta/s$ could have a large slope and negative curvature, or $\zeta/s$ could have a large max and narrow width, neither of which are depicted above.}
+  \label{fig:viscosity_dof}
+\end{figure}
 
@@
 \end{equation}
-where $\Delta\y = \y_m(\x) - \y_e$ is the discrepancy of the model and experiment, and $\Sigma$ is the total covariance matrix which includes all known sources of uncertainty for the simulated and measured observables.
+where $\Delta\y = \y_m(\x) - \y_e$ is the discrepancy of the model and experiment, and $\Sigma = \Sigma_m(\x) + \Sigma_e$ is the \emph{total} covariance matrix, equal to the sum of a modeling component $\Sigma_m(\x)$ and an experimental component $\Sigma_e$ which account for all known sources of uncertainty in the simulated and measured observables.
 
@@
 
+\begin{figure*}[t]
+  \includegraphics{correlation_matrices}
+  \caption{Correlation matrices for Pb-Pb collisions at $\sqrts=5.02$~TeV for the model/emulator (left) at a random point in parameter space and the experimental data (right) used to construct the covariance matrix ${\Sigma = \Sigma_m + \Sigma_e}$ in Eqn.~\eqref{eq:likelihood}.
+Experimental statistical and systematic errors are from ALICE and CMS \note.
+The experimental correlation structure is modeled using Eqn.~\eqref{eq:corr}.}
+  \label{fig:correlation}
+\end{figure*}
+
 The likelihood function \eqref{eq:likelihood} provides evidence for \mbox{(or against)} the model parameters $\x$ by comparing the model predictions $\y_m(\x)$ to experimental data $\y_e$.
@@
 We calculate for each set of model parameters the following observables at midrapidity:
-\begin{itemize}[leftmargin=2\parindent, itemsep=0pt]
+\begin{itemize}[leftmargin=1\parindent, itemsep=0pt]
   \item Charged-particle multiplicity $d\nch/d\eta$.
@@
 \begin{equation}
+  \label{eq:corr}
   \rho_{ij}^\text{sys} = \exp \left[ -\frac{1}{2} \left(\frac{b_i - b_j}{l} \right)^2 \right],
 \end{equation}
+
 where $b_i$ and $b_j$ are the midpoints of two observable bins of a single type (centrality or relative multiplicity) and $l$ is a correlation length which describes how quickly the observable bins decorrelate as the distance between the bins increases.
@@
 The $p$-Pb mean $p_T$ and flow observables, meanwhile, use relative multiplicity bins $\nch / \langle \nch \rangle$ and $\ntrk / \langle \ntrk \rangle$ so we use a much smaller correlation length $l=5$ to reflect the smaller range of bin values.
-
-\fig*{correlation_matrices}{
-  Correlation matrices for Pb-Pb collisions at $\sqrts=5.02$~TeV for the model/emulator (left) and experimental data (right) used to construct the covariance matrix ${\Sigma = \Sigma_\mathrm{model} + \Sigma_\mathrm{expt}}$ in Eqn.~\eqref{eq:likelihood}.
-  Experimental statistical and systematic errors are from ALICE and CMS \note.
-  The experimental correlation structure is modeled using Eqn.~?.
-}
+We show, as an example, the resulting experimental correlation matrix $\text{corr}(\Sigma_e)$ for the Pb-Pb system on the right side of Fig.~\ref{fig:correlation}.
+The correlation matrix is block diagonal, with each block representing the correlations within a single class of observable; in this case, the experimental yield $d\nch/d\eta$ and flows $\vnk{n}{2}$.
 
@@
 
-We therefore follow an established framework for computationally intensive models and train a Gaussian process emulator (GPE) to act as a fast surrogate for the full physics simulation \note.
-The GPE enables essentially instantaneous predictions for $\y_m = f(\x)$ which allows us to sample the posterior distribution millions of times.
-In order to train the GPE, we first generate a scaffolding of the parameter space using a maximin Latin hypercube design to distribute 500 points uniformly in our 15-dimensional parameter space according to the ranges in Tbl.~\ref{tab:design}.
+We therefore follow an established framework for computationally intensive models and train an emulator to act as a fast surrogate for the full physics simulation \note.
+The emulator enables essentially instantaneous predictions for $\y_m = f(\x)$ and allows us to sample the posterior distribution millions of times.
+In order to train the emulator, we first generate a scaffolding of the parameter space using a maximin Latin hypercube design \note\ to distribute 500 points uniformly throughout our 15-dimensional parameter space according to the parameter ranges in Tbl.~\ref{tab:design}.
 We then run minimum-bias and multiplicity triggered $p$-Pb and Pb-Pb events at each design point, and calculate the model observables from the ensemble of events.
 
-Let $X$ denote the $d \times m$ dimensional matrix of design points where $d=500$ is the number of design points and $m$ is the number of emulated observables $y \in \y_m$.
+Specifically, let $X$ denote the ${d \times n}$ design matrix of $d=500$ training points, where each training point is a vector $\x = (x_1, x_2, \dots, x_n)$ of the $n=15$ model parameters in Tbl.~\ref{tab:design}.
+Similarly, let $Y$ denote the corresponding ${d \times m}$ observables matrix, where each row of $Y$ is a vector $\y_m = (y_1, y_2, \dots, y_m)$ of $m$ different simulated observables (here we overload our notation so the subscript $m$ means both \emph{model} and \emph{observable number}).
+Our goal is to train an emulator for the physics model $f$ using the discrete observations $f: X \mapsto Y$.
 
-\subsection{Experimental }
+We use for this purpose a specific type of emulator known as a Gaussian process emulator (GPE).
+The advantage of using GPE's is that they provide an estimate of their own uncertainty which allows us to account for this uncertainty when constructing the covariance matrix $\Sigma$ in Eqn.~\eqref{eq:likelihood}.
+One quirk of GPE's is that they are restricted to scalar-valued functions, i.e.\ functions of one output, whereas we require an emulator for vector-valued functions with multiple outputs.
+This restriction is commonly circumvented using principal component analysis (PCA): a general procedure which transforms a set of correlated variables $\y = (y_1, y_2, \dots, y_m)$ into a new basis representation $\z = (z_1, z_2, \dots, z_m)$ where the linear correlations between $z_i$, $z_j$ vanish for all $i\ne j \in m$ \note.
+
+We first preprocess our model observables by centering and scaling each column of $Y$ (single observable) to zero mean and unit variance to generate a standardized observable matrix $\tilde{Y}$.
+PCA is then used to reexpress each row-vector $\tilde{\y}$ of $\tilde{Y}$ (all standardized observables at a single design point) in the new principal component basis:
+\begin{equation}
+  \tilde{\y}_i = \sum\limits_{j=1}^m z_{ij} \mathbf{v_j},
+\end{equation}
+where $\tilde{\y}_i$ are the standardized observables of the $i$-th row-vector (design point) of matrix $\tilde{Y}$, and $z_{ij}$ and $\mathbf{v}_j$ are the coefficients and corresponding vectors of its $j$-th principal component.
+
+The principal components are reported in order of explained variance, with the first principal component vector $\mathbf{v}_1$ accounting for the most variance in $\tilde{Y}$, and the last principal component vector $\mathbf{v}_m$ accounting for the least.
+We then train a set of independent GPE's,
+\begin{align}
+  \label{eq:gpe}
+  z_1 &= \text{gp}_1(\x), \nonumber \\
+  z_2 &= \text{gp}_2(\x), \nonumber \\
+  \vdots & \nonumber\\
+  z_k &= \text{gp}_k(\x),
+\end{align}
+to predict the principal components (coefficients) $(z_1, \dots, z_k)$ of the first $k$ principal component vectors as a function of the model parameters $\x$ which vary across the design $X$.
+We use $k=7$ principal components when emulating the $p$-Pb system and $k=8$ principal components when emulating Pb-Pb, chosen to describe 99.5\% of the total observed variance of each system.
+
+The GPE's are essentially fancy interpolators applied to the model's training points and PCA transformed observables.
+Each GPE reports a mean value $z(\x)$ as well as an estimated error $\delta z(\x)$ which accounts for statistical noise in the training data and interpolation error between the design points.
+Once the GPE's \eqref{eq:gpe} are trained, we can predict the model observables $\y_m$ at parameter point $\x$ by transforming the vector of principal components
+\begin{equation}
+  \mathbf{z}(\x) = (z_1(\x), z_2(\x), \dots, z_k(\x))
+\end{equation}
+\emph{back} to physical space.
+Similarly, we can construct the covariance matrix of the observables in PCA space,
+\begin{equation}
+\text{Cov}(z_i, z_j) = \text{diag}[(\delta z_{1})^2, (\delta z_2)^2, \dots, (\delta z_k)^2 ],
+\end{equation}
+
+\onecolumngrid
+{\centering\includegraphics{observables_pbpb}\par}
+\captionof{figure}{Simulated observables compared to experimental data for Pb-Pb collisions at $\sqrts=5.02$~TeV.
+Top row: explicit model calculations (no emulator) for each of the 500 design points; bottom row: emulator predictions for $n=100$ random samples drawn from the posterior.
+Points with error bars are experimental data from ALICE with statistical and systematic errors added in quadrature \note.}
+\vspace{.5cm}
+{\centering\includegraphics{observables_ppb}\par}
+\captionof{figure}{Same as Fig.~\ref{fig:obs_pbpb} but for $p$-Pb collisions at $\sqrts=5.02$~TeV.
+Note that multiplicity bins are used for mean $p_T$ and flow cumulant observables to match the bins used by experiment.
+Experimental data are from ALICE and CMS \note.}
+\twocolumngrid
+
+and transform it back to physical space as well to obtain the covariance matrix of the model observables $\y_m$ at a given parameter point $\x$.
+
+The resulting emulator therefore predicts both a mean prediction $\y_m(\x)$ and an uncertainty covariance matrix $\Sigma_m(\x)$ which accounts for multiple sources of model and emulator uncertainty, including the truncation error expected from using a finite number of principal components $k < m$.
+The covariance matrix $\Sigma_\text{emu}$ thus includes the \emph{known-unknowns}, but not the \emph{unknown-unknowns}, such as the overall credibility of viscous hydrodynamics, which are difficult and often impossible to quantify.
+We show in Fig.~\ref{fig:correlation} the correlation structure $\text{Corr}(\Sigma)$ of the model (emulator) covariance $\Sigma_m(\x)$ for a random parameter point $\x$ in the design space (left-side), along side the same correlation matrix for the experimental covariance (right-side) for the Pb-Pb collision system.
+For additional information on the model emulator, we direct the reader to the appendix which includes several validation tests of the emulator prediction accuracy \note.
+
+\subsection{Bayesian calibration}
+
+In order to calibrate the model on two different collision systems, we expand the likelihood function \eqref{eq:likelihood} into a joint likelihood
+\begin{equation}
+  \label{eq:joint_likelihood}
+  P(E | H_\x) = P(E_\text{Pb-Pb} | H_\x) \cdot P(E_\text{$p$-Pb} | H_\x),
+\end{equation}
+where $E$ subsumes all evidence from the $p$-Pb and Pb-Pb collision systems and $H_\x$ is our hypothesis that $\x$ equals the true parameters $\x_\star$.
+
+We then perform Markov-chain Monte Carlo (MCMC) importance sampling on the posterior distribution in Eqn.~\eqref{eq:bayes} to draw random samples for $P(H_\x | E)$, the estimate of the true model parameters given the model and the experimental data.
+We use an affine-invariant ensemble sampler which uses a large ensemble of interdependent walkers \note\ and allow the MCMC chain to ``burn-in'' before generating $\mathcal{O}(10^7)$ posterior samples.
 
@@
 
-\fig*[t]{validation_example}{
-  Placeholder
+We show simulated and emulated model observables (thin colored lines) for Pb-Pb collisions (Fig.~\ref{fig:obs_pbpb}) and \mbox{$p$-Pb} collisions (Fig.~\ref{fig:obs_ppb}) at $\sqrts=5.02$~TeV compared to experimental data from CMS and ALICE \note.
+The top row of each figure shows explicit model calculations at each of the $d=500$ design points (training data), while the bottom row shows emulator predictions for $n=100$ random parameter samples drawn from the Bayesian posterior.
+Each column shows a different class of observable.
+The charged-particle yield $d\nch/d\eta$ is shown on the left, mean $p_T$ is in the middle, and two-particle flow cumulants $\vnk{n}{2}$ for $n=2,3,4$ are on the right.
+The Pb-Pb mean $p_T$ and $p$-Pb $\vnk{4}{2}$ datasets are missing and hence are omitted from the present calibration.
+
+Notice the large spread of the observables calculated at the training points (top row of each figure).
+The design is constructed to vary each parameter across a wide range of values, specified in Tbl.~\ref{tab:design}, and hence the corresponding model calculations are equally uncertain.
+We also point out that there is considerably more variance in the $p$-Pb training data than the Pb-Pb training data.
+The $p$-Pb yields, mean $p_T$ and flow cumulants all vary wildly within the chosen parameter ranges.
+For instance, we can turn the $p$-Pb flows completely \emph{off} with suitably chosen parameters which is not possible in the Pb-Pb system.
+Evidently the $p$-Pb model predictions are far more sensitive to modeling uncertainties.
+
+Conversely, the calibrated emulator predictions (bottom row of each figure) are far better constrained and nicely track the experimental data points.
+We emphasize here that the posterior parameter values are obtained from a \emph{simultaneous} calibration to $p$-Pb and Pb-Pb data, and thus they are self-consistent between the two systems.
+The spread in the posterior samples reflects different sources of model and experimental uncertainty as well as tension in the optimal fit parameters which describe each observable.
+We demonstrate later in the text that a single set of model parameters well describes all of the calibration data, and thus we believe that much of the spread in the posterior samples is uncertainty contributed by our emulator.
+We also note that although the $p$-Pb posterior samples have a somewhat larger spread than the Pb-Pb samples, the percentage uncertainty of the $p$-Pb emulator is similar to that of the Pb-Pb emulator, and thus the difference is likely due to the larger variance of the $p$-Pb training data.
+The uncertainty in the posterior distribution would thus be improved by running the calibration with more design points or with a narrower range of parameter values to increase the density of the training points and reduce interpolation error.
+
+We now direct our attention to Fig.~\ref{fig:posterior} which shows the posterior distribution of the model input parameters.
+Recall that the posterior $P(H_\x | E)$ is the probability that our hypothesis $H_\x$ for the true model parameters $\x_\star$ is correct, given the evidence $E$ provided by experiment.
+The present distribution has 15 dimensions, one for each parameter listed in Tbl.~\ref{tab:design}, and so we sample the distribution and histogram the samples to project the distribution onto a lower number of dimensions for visualization purposes.
+Each diagonal panel in the figure is the distribution of a \emph{single} model parameter (marginalized over all other parameters), and each lower-diagonal panel is the joint distribution of a \emph{pair} of model parameters (marginalized over all other parameters).
+The full posterior distribution includes a wealth of information.
+Here we focus on the most important features...
+
+{
+\onecolumngrid
+\begin{figure}
+  \makebox[\textwidth]{
+      \includegraphics[width=.95\paperwidth]{posterior}
+  }
+\end{figure}
+\captionof{figure}{Bayesian posterior distribution of the model input parameters.
+The diagonal panels show the marginalized distributions of individual model parameters, while off-diagonal panels show the joint distributions for pairs of model parameters, visualizing their correlations.
+The marginalized distribution medians and 90\% credible intervals are annotated along the diagonal.}
+\clearpage
+\twocolumngrid
 }
 
-\begin{figure}[h]
-  \makebox[\textwidth]{
-    \begin{minipage}[t]{\paperwidth}
-      \includegraphics[width=.95\paperwidth]{posterior}
-      \captionsetup{
-        width={.83\textwidth},
-        justification=raggedright,
-      }
-      \caption{Bayesian posterior distribution of the model input parameters.
-The diagonal panels show the marginalized distributions of individual model parameters, while off-diagonal panels show the joint distributions for pairs of model parameters, visualizing their correlations.
-The marginalized distribution medians and 90\% credible intervals are annotated along the diagonal.}
-    \end{minipage}
-  }
-  \label{fig:posterior}
+\begin{itemize}[leftmargin=1\parindent, itemsep=0pt, topsep=5pt]
+  \item The posterior initial condition normalization factor $\text{Norm} = 20.3^{+2.2}_{-2.2}$ (upper-left diagonal) is well constrained and describes both the $p$-Pb and Pb-Pb yields simultaneously (e.g.\ see Figs.~\ref{fig:obs_ppb}).
+    We emphasize that this self-consistency is highly non-trivial since the ratio of $p$-Pb to Pb-Pb yields model can only fit both yields simultaneously if the \trento\ entropy deposition parameter $p$ is appropriately tuned.
+\end{itemize}
+
+\onecolumngrid
+{\centering\includegraphics{observables_map}\par}
+\captionof{figure}{Model calculations using the \emph{maximum a posterior} (MAP) parameters compared to experiment.
+Colored lines are model calculations for $p$-Pb collisions (left) and Pb-Pb collisions (right) at $\sqrts=5.02$~TeV.
+Points with error bars are the experimental data with statistical uncertainties, and gray bands their corresponding systematic uncertainties, from ALICE and CMS \note.
+The ratio axes show the ratio of model over data where where available with gray bands indicating $\pm 10\%$.}
+\twocolumngrid
+
+\begin{figure}
+  \includegraphics{flow_corr}
+  \caption{Model calculations of the symmetric cumulants for Pb-Pb collisions at $\sqrts=5.02$~TeV using the \emph{maximum a posteriori} (MAP) parameters.}
+  \label{fig:flow_corr}
 \end{figure}
 
-\begin{figure*}[p]
-  \includegraphics{observables_pbpb}
-  \caption{Simulated observables compared to experimental data for Pb-Pb collisions at $\sqrts=5.02$~TeV.
-Top row: explicit model calculations (no emulator) for each of the 500 design points; bottom row: emulator predictions for $n=100$ random samples drawn from the posterior.
-Points with error bars are experimental data from ALICE with statistical and systematic errors added in quadrature \note.}
-  \vspace{5ex}
-  \includegraphics{observables_ppb}
-  \caption{Same as above but for $p$-Pb collisions at $\sqrts=5.02$~TeV.
-Note that multiplicity bins are used for mean $p_T$ and flow cumulant observables to match the bins used by experiment.
-Experimental data are from ALICE and CMS \note.}
+\begin{figure}
+  \includegraphics[width=\columnwidth]{posterior_proton_shape}
+  \caption{Posterior distribution for the nucleon width $w$ and constituent width $v$.
+The white trapezoidal region (lower-right) covers the initial range of allowed values for $w$ and $v$, while the gray triangular region (upper-left) indicates values which are excluded by the prior $v > w$.
+The posterior distribution, shown in blue, indicates the preferred values for $w$ and $v$ determined by the analysis.}
+  \label{fig:posterior_proton_shape}
+\end{figure}
+
+\begin{figure*}
+  \includegraphics[width=\textwidth]{region_shear_bulk}
+  \caption{Left figure: estimated temperature dependence of the QGP specific shear viscosity $(\eta/s)(T)$ determined by the present Bayesian analysis of $p$-Pb and Pb-Pb collisions at $\sqrts=5.02$~TeV (orange line/band) compared to a previous Bayesian analysis of Pb-Pb collisions at $\sqrts=2.76$ and 5.02~TeV (blue line/band) \note.
+The lines are the medians of each posterior distribution, and the bands are their 90\% credible regions.
+Right figure: same as before, but for the temperature dependence of the QGP specific bulk viscosity $(\zeta/s)(T)$.}
+  \label{fig:region_shear_bulk}
 \end{figure*}
-
-\fig*[t]{observables_map}{
-  Model calculations using the \emph{maximum a posterior} (MAP) parameters compared to experiment.
-Colored lines are model calculations for $p$-Pb collisions (left) and Pb-Pb collisions (right) at $\sqrts=5.02$~TeV.
-Points with error bars are the experimental data with statistical uncertainties, and gray bands their corresponding systematic uncertainties, from ALICE and CMS \note.
-The ratio axes show the ratio of model over data where where available with gray bands indicating $\pm 10\%$.
-}
-
-\fig*[t]{flow_corr}{
-  Model calculations of the symmetric cumulants for Pb-Pb collisions at $\sqrts=5.02$~TeV using the \emph{maximum a posteriori} (MAP) parameters.
-}
-
-\fig{posterior_proton_shape}{
-%  Posterior distribution for the constituent sampling radius $\R$ and constituent width $\w$.
-The white trapezoidal region (lower-right) covers the initial range of allowed values for $\R$ and $\w$, while the gray triangular region (upper-left) indicates values which are excluded by the prior $\w > \R$.
-The posterior distribution, shown in blue, indicates the preferred values for $\R$ and $\w$ determined by the analysis.
-}
-
-\fig*[t]{region_shear_bulk}{
-  Left figure: estimated temperature dependence of the QGP specific shear viscosity $(\eta/s)(T)$ determined by the present Bayesian analysis of $p$-Pb and Pb-Pb collisions at $\sqrts=5.02$~TeV (orange line/band) compared to a previous Bayesian analysis of Pb-Pb collisions at $\sqrts=2.76$ and 5.02~TeV (blue line/band) \note.
-The lines are the medians of each posterior distribution, and the bands are their 90\% credible regions.
-Right figure: same as before, but for the temperature dependence of the QGP specific bulk viscosity $(\zeta/s)(T)$.
-}
 
@@
 
-\appendix*
+\appendix
 
@@
 
+\section{Emulator validation}
+\label{app:validation}
+\fig*[t]{validation_example}{
+  Placeholder
+}
+
 \bibliography{substructure}

### 2018-07-20 src/plots.py

@@ -346,7 +346,7 @@ def _observables_plots():
             ),
             xlim=dict(
                 pPb5020=(0, 60),
-                PbPb5020=(0, 75),
+                PbPb5020=(0, 80),
             ),
             ylim=(1e2, 1e5),
             yscale='log',
@@ -366,7 +366,7 @@ def _observables_plots():
             ylabel=r'$\langle p_T \rangle$ [GeV]',
             xlim=dict(
                 pPb5020=(1, 6),
-                PbPb5020=(0, 75),
+                PbPb5020=(0, 80),
             ),
             ylim=(0, 1.5),
             subplots=[
@@ -383,7 +383,7 @@ def _observables_plots():
             ylabel=r'$\delta p_T/\langle p_T \rangle$',
             xlim=dict(
                 pPb5020=(1, 6),
-                PbPb5020=(0, 75),
+                PbPb5020=(0, 80),
             ),
             ylim=(0, 0.05),
             subplots=[('pT_fluct', None, dict())]
@@ -398,7 +398,7 @@ def _observables_plots():
             ylabel=r'$v_n\{2\}$',
             xlim=dict(
                 pPb5020=(1, 6),
-                PbPb5020=(0, 75),
+                PbPb5020=(0, 80),
             ),
             ylim=(0, .15),
             subplots=[
@@ -442,7 +442,7 @@ def observables(system):
 
     fig, axes = plt.subplots(
         nrows=2, ncols=len(plots),
-        figsize=figsize(1.1, aspect=.5),
+        figsize=figsize(1.1, aspect=.6),
     )
 
     for (posterior, plot), ax in zip(
@@ -565,12 +565,12 @@ def observables_map():
                 ],
                 legend=True
             )
-    fig = plt.figure(figsize=figsize(1.1, 1.5))
+    fig = plt.figure(figsize=figsize(0.955, 1.5))
 
     yields, mean_pT, mean_pT_fluct, flows = [
         gridspec.GridSpecFromSubplotSpec(
             2, 2, gs, height_ratios=[5, 1] if n == 0 else [3, 1],
-            hspace=0.1, wspace=.1
+            hspace=0.1, wspace=.14
         ) for n, gs in enumerate(
             gridspec.GridSpec(4, 1, height_ratios=[6, 4, 4, 4])
         )
[…]

### 2018-07-21 paper/substructure.tex
Fix some latex formatting issues

@@
 \usepackage{amssymb}
-\usepackage{multirow}
 
 \usepackage[inline]{enumitem}
-\usepackage{lipsum}
 
@@
 \newcommand{\trento}{T\raisebox{-0.5ex}{R}ENTo}
+\newcommand{\sqrts}{\sqrt{s_{NN}}}
 \newcommand{\nch}{N_\text{ch}}
 \newcommand{\ntrk}{N_\text{trk}^\text{offline}}
-\newcommand{\sqrts}{\sqrt{s_{NN}}}
 \newcommand{\sigmaf}{\sigma_\mathrm{fluct}}
@@
 
+% Reduce the size of the itemize bullets
 \renewcommand\labelitemi{$\vcenter{\hbox{\scriptsize$\bullet$}}$}
+
+% Create a temporary full-page single column environment
+\newenvironment{fullpage}{\onecolumngrid}{\clearpage\twocolumngrid}
 
@@
 
-% Convenient figure macro.  Usage:
-%
-%   \fig[placement specifier = t]{filename}{caption}
-%
-% This creates a figure environment, includes the given filename as graphics,
-% puts the given caption below the graphics, and labels it 'fig:filename'.
-% The optional placement specifier defaults to 't' and is passed directly to
-% the figure environment.
-% Use \fig* to make a figure* environment, i.e. a wide figure.
-\usepackage{xparse}
-\NewDocumentCommand\fig{sO{t}mm}{
-  \begin{figure\IfBooleanT{#1}{*}}[#2]
-    \includegraphics{#3}
-    \caption{\label{fig:#3}#4}
-  \end{figure\IfBooleanT{#1}{*}}
-}
-
 \begin{document}
@@
 \title{
-  Estimating initial state and quark-gluon plasma medium properties\\
-  using a hybrid model with nucleon substructure\\
+  Estimating initial state and quark-gluon plasma medium properties \\
+  using a hybrid model with nucleon substructure \\
   calibrated to p-Pb and Pb-Pb collisions at \texorpdfstring{$\mathbf{\sqrts=5.02}$}{}~TeV
@@
 Most of the calibration observables listed in Tbl.~\ref{tab:observables} are calculated as a function of collision centrality, where centrality is defined using some measure of the underlying event activity, e.g.\ the charged particle yield in a given rapidity window.
-When calculating these observables, we generate $\mathcal{O}(10^4)$ minimum-bias events at each design point and divide the events into centrality bins using the charged particle yield at midrapidity, similar to the procedure used by experiment.
+When calculating these observables, we generate \order{10^4} minimum-bias events at each design point and divide the events into centrality bins using the charged particle yield at midrapidity, similar to the procedure used by experiment.
 
@@
 \end{equation}
-
 where $b_i$ and $b_j$ are the midpoints of two observable bins of a single type (centrality or relative multiplicity) and $l$ is a correlation length which describes how quickly the observable bins decorrelate as the distance between the bins increases.
@@
   z_2 &= \text{gp}_2(\x), \nonumber \\
-  \vdots & \nonumber\\
+  \vdots & \nonumber \\
   z_k &= \text{gp}_k(\x),
@@
 
-\onecolumngrid
+\begin{fullpage}
 {\centering\includegraphics{observables_pbpb}\par}
@@
 Experimental data are from ALICE and CMS \note.}
-\twocolumngrid
+\end{fullpage}
 
@@
 We then perform Markov-chain Monte Carlo (MCMC) importance sampling on the posterior distribution in Eqn.~\eqref{eq:bayes} to draw random samples for $P(H_\x | E)$, the estimate of the true model parameters given the model and the experimental data.
-We use an affine-invariant ensemble sampler which uses a large ensemble of interdependent walkers \note\ and allow the MCMC chain to ``burn-in'' before generating $\mathcal{O}(10^7)$ posterior samples.
+We use an affine-invariant ensemble sampler which uses a large ensemble of interdependent walkers \note\ and allow the MCMC chain to ``burn-in'' before generating \order{10^7} posterior samples.
 
@@
 
-{
-\onecolumngrid
+\begin{fullpage}
 \begin{figure}
@@
 The marginalized distribution medians and 90\% credible intervals are annotated along the diagonal.}
-\clearpage
-\twocolumngrid
-}
+\end{fullpage}
 
@@
 
-\onecolumngrid
+\begin{fullpage}
 {\centering\includegraphics{observables_map}\par}
@@
 The ratio axes show the ratio of model over data where where available with gray bands indicating $\pm 10\%$.}
-\twocolumngrid
+\end{fullpage}
 
@@
 \label{app:validation}
-\fig*[t]{validation_example}{
-  Placeholder
-}
+\begin{figure*}
+  \includegraphics{validation_example}
+\end{figure*}
 

### 2018-07-21 paper/substructure.tex
Write appendices

@@
 
-\section{Adaptive grid sizing}
+\section{Event-by-event grid resizing}
 \label{appendix}
-This is the appendix
+
+\begin{figure}
+  \begin{tikzpicture}
+    \draw [black, thin, step=0.3cm] (0.9, 0.9) grid +(4.5, 4.5);
+    \draw [gray, thin, step=0.9cm] (0, 0) grid +(6.3, 6.3);
+    \draw [theblue] (3.15, 3.15) circle (2.25cm);
+  \end{tikzpicture}
+  \caption{Diagram of the adaptive grid resizing algorithm (not drawn to scale).
+Each initial condition event is first run on a very large coarse-grained mesh (large gray grid) of one-third the spatial resolution otherwise required to measure hydrodynamic observables.
+We then measure the maximum transverse radius $R_\text{max}$ (blue circle) of the hypersurface defined by the temperature isotherm $T = T(e_\text{min})$, where $e_\text{min}$ is the largest energy density which can be truncated without modifying the hydrodynamic observables calculated from the event.
+Finally, the initial condition event is rerun on a smaller and finer mesh (smaller black grid) with three-times the cell density of the pre-run event and a smaller transverse extent $-R_\text{max} < x < R_\text{max}$.}
+  \label{fig:adaptive_grid}
+\end{figure}
+
+The boost-invariant \texttt{VISH2+1} hydrodynamics code used in this work \cite{Song:2007ux, Shen:2014vra} runs on a Cartesian transverse grid specified by a maximum grid size $x_\text{max}$ and grid step width $dx$ which fix the transverse grid extent ${-x_\text{max} < x < x_\text{max}}$ and number of grid cells along each dimension ${n_x = 2\, x_\text{max} / dx}$.
+In general, the maximum grid size $x_\text{max}$ should be set large enough to contain the full spacetime evolution of the event.
+This means that the truncation of $T^{\mu\nu}$ at the boundaries of the grid should never modify the final-state observables.
+We enforce this requirement by finding an energy density cutoff $e_\text{min}$ for which the matter $e < e_\text{min}$ can be effectively discarded without significantly modifying the simulation observables.
+We then fix the maximum grid size $x_\text{max}$ such that it fully encloses the isotherm $T=T(e_\text{min})$ for the full lifetime of the fireball.
+
+We find that we can quickly estimate the maximum radius $R_\text{max} = |\x_\text{max}|^2$ of the spacetime hypersurface ${T = T(e_\text{min})}$ by running the event on a coarse-grained spatial grid with one-third the spatial resolution we would otherwise require to resolve typical hydrodynamic observables such as mean $p_T$ and flows.
+The simulation time of a single \texttt{VISH2+1} event scales like ${\sim}n_x^3$ since $dx \propto d\tau$, and thus our ``pre-run'' event requires only ${\sim}1/27$th the time of a production event.
+We therefore run a coarse-grained pre-event on an excessively large grid for \emph{every} minimum-bias event to estimate $R_\text{max}$, then rerun the same event on a thrice finer grid with a trimmed spatial extent $x_\text{max} \equiv R_\text{max}$.
+See Fig.~\ref{fig:adaptive_grid} for a simple diagram of the procedure.
+
+In practice, we find that event-by-event grid resizing leads to a massive speed increase for minimum bias events compared to using a single fixed grid for the entire minimum bias sample.
+This is because the maximum transverse size of each event varies dramatically, from a few fm in peripheral Pb-Pb collisions to 50 fm or more in central Pb-Pb collisions.
+The procedure should generalize to other hydrodynamic codes, including those with $3+1$ spacetime dimensions, where the time savings could be even more pronounced.
 
@@
 \label{app:validation}
+
 \begin{figure*}
   \includegraphics{validation_example}
+  \caption{Example emulator validation for one observable, the Pb-Pb charged particle yield $d\nch/d\eta$ in the 20-30\% centrality class.
+We use the k-fold cross validation method (explained in the text) to partition the model inputs $X$ and outputs $Y$ into training and validation data.
+The scatterplot on the left shows the emulator predictions and one sigma error bars (x-axis) against explicit model calculations (y-axis).
+Perfect emulator/model agreement is indicated by the black like $y_\text{pred}=y_\text{obs}$.
+The histogram on the right shows that the errors are properly accounted for, i.e.\ the normalized residuals follow a normal distribution with unit variance and zero mean.
+  }
+  \label{fig:validation_example}
+\end{figure*}
+
+The emulator is a surrogate for the full physics simulation which generates probabilistic predictions for the model observables $\y_m$ at a given point $\x$.
+Here we validate these probabilistic predictions using a method known as k-fold cross validation.
+We first randomly partition our $d=500$ training points into $k=20$ equal sized subsamples or ``folds''.
+One of the subsamples is used to validate the emulator and the remaining $k-1$ subsamples are used to train it.
+The process is then repeated for each of the subsamples so that we end up validating on all of the training data.
+
+Figure~\ref{fig:validation_example} shows a scatterplot of the emulator predictions with one-sigma error bars (x-axis) against explicit model calculations (y-axis).
+Perfect emulator and model agreement is indicated by the black line $y_\text{pred} = y_\text{obs}$.
+If the emulator errors are properly accounted for, then the normalized residuals ${z=(y_\text{pred} - y_\text{obs})/\sigma_\text{pred}}$ sample a unit normal distribution:
+\begin{equation}
+  P(z) \sim \mathcal{N}(\mu=0,\sigma=1).
+\end{equation}
+
+This comparison is shown by the histogram and box plot on the right side of Fig.~\ref{fig:validation_example}.
+The emulator error is clearly significant, but it is also properly modeled, as indicated by the agreement between the normalized residuals and the unit normal distribution on the right (black curve).
+Moreover, since we include this uncertainty in the likelihood covariance matrix \eqref{eq:likelihood}, we expect our results to be robust to the emulator limitations.
+This is an important point that bears repeating.
+The emulator uncertainty does not erode the veracity of the posterior distribution if it is correctly modeled and accounted for.
+
+More generally, we can perform the validation test in Fig.~\ref{fig:validation_example} for \emph{every} observable $y \in \y_m$ and check that each observable's normalized residuals ${z=(y_\text{pred} - y_\text{obs})/\sigma_\text{pred}}$ follow a unit normal distribution.
+This test is applied to the Pb-Pb system in Fig.~\ref{fig:validation_PbPb5020} and the $p$-Pb system in Fig.~\ref{fig:validation_pPb5020}.
+The top row of each figure shows a box-plot for the normalized residuals of each observable compared to the quantiles of a unit normal distribution.
+The thin horizontal black lines correspond to the 10th and 90th percentiles of a unit normal distribution, and the gray band its interquartile range.
+These visual references should be compared to the whiskers and interquartile range respectively of each box plot, analogous to the comparison test of Fig.~\ref{fig:validation_example}.
+The emulators generally behave as expected, although the validation is somewhat better for the Pb-Pb system than the $p$-Pb system.
+For instance, the $p$-Pb charged particle yield $d\nch/d\eta$ uncertainties are over predicted.
+It is not immediately clear why this would be the case, but the MAP observables in Fig.~\note\ are in good agreement with their emulator predictions which suggests it should not be a grave concern.
+
+We also show in Fig.~\ref{fig:validation_PbPb5020} and Fig.~\ref{fig:validation_pPb5020} an estimate of the emulator error magnitude.
+This error is expressed in terms of the unitless variable
+\begin{equation}
+  \hat{z} = \frac{y_\text{pred} - y_\text{obs}}{(\Delta y)_{99\%}},
+\end{equation}
+where $(\Delta y)_{99\%}$ is 99\% of the full variability of $y$ across the design.
+Thus $\hat{z}$ can be thought of as a fractional emulator error relative to the full design variability.
+The bottom row of each figure shows the root-mean-square value of $\hat{z}$.
+We see that $\text{RMS}\{\hat{z}\}$ ranges from a few percent for most observables to a maximum value of 15\% for the $p$-Pb triangular flow $v_3\{2\}$ in the lowest multiplicity bin.
+This suggests that the present analysis would benefit the most from more $p$-Pb events, in particular, from more multiplicity triggered events which are used to calculate the flows.
+
+\begin{figure*}
+  \includegraphics{validation_PbPb5020}
+  \caption{}
+  \label{fig:validation_PbPb5020}
+\end{figure*}
+
+\begin{figure*}
+  \includegraphics{validation_pPb5020}
+  \caption{}
+  \label{fig:validation_pPb5020}
 \end{figure*}

### 2018-07-21 src/plots.py

@@ -1765,7 +1765,7 @@ def validation_all(system):
     ax_rms.set_xticks([])
     ax_rms.set_yticks(np.arange(0, 21, 5))
     ax_rms.set_ylim(0, 20)
-    ax_rms.set_ylabel('RMS % error')
+    ax_rms.set_ylabel('Fractional error')
 
     for y in ax_rms.get_yticks():
         ax_rms.axhline(y, color='.5', zorder=-10)

### 2018-07-22 paper/substructure.tex
More work on the results section

@@
 
-\newcommand{\trento}{T\raisebox{-0.5ex}{R}ENTo}
+\newcommand{\trento}{\textsc{T\raisebox{-0.5ex}{R}ENTo}}
 \newcommand{\sqrts}{\sqrt{s_{NN}}}
@@
 Free parameters of the model which describe the initial state and QGP medium are then simultaneously calibrated to fit charged particle yields, mean $p_T$ and flow cumulants.
-We argue that the global agreement of the calibrated model with the experimental data strongly supports the existence of hydrodynamic flow in small collision systems at ultrarelativistic energies, and that this flow necessarily develops at length scales smaller than a proton.
+We argue that the global agreement of the calibrated model with the experimental data strongly supports the existence of hydrodynamic flow in small collision systems at ultrarelativistic energies, and that the flow produced develops at length scales smaller than a proton.
 Posterior estimates for the model's input parameters are obtained, and new insights into the temperature dependence of the QGP transport coefficients and event-by-event structure of the proton are discussed.
@@
 Ultrarelativistic nuclear collisions between one light-ion and one heavy-ion, e.g.\ $^3$He-Au and p-Pb collisions, generate dense, compact sources of nuclear matter which produce long-range multiparticle correlations that are strikingly similar to the correlations observed in heavy-ion collisions where collectivity is commonly explained by the existence of hydrodynamic flow \cite{CMS:2012qk, Abelev:2012ola, Aad:2012gla, Adare:2015ctn}.
-This observation suggests that hydrodynamic behavior could be manifest in small droplets of quark-gluon-plasma (QGP), and that flow might develop at length scales smaller than a single proton.
+This observation suggests that hydrodynamic behavior could be manifest in small droplets of quark-gluon-plasma (QGP) \cite{Bozek:2011if, Bozek:2013uha}, and that flow might develop at length scales smaller than a single proton \note.
 
@@
 We begin, in Sec.~\ref{sec:model}, by constructing a nuclear collision model for $p$-Pb and Pb-Pb collisions using initial conditions with parametric nucleon substructure, and transport dynamics described by a pre-equilibrium free-streaming stage, viscous hydrodynamics and microscopic Boltzmann transport.
-In Sec.~\ref{sec:calibration}, we calibrate free parameters of the model to fit charged particle yields, mean $p_T$ and anisotropic flow coefficients of \emph{both} collision systems at $\sqrts=5.02$~TeV, and finally, in Secs.~\ref{sec:results} and \ref{sec:summary}, we present posterior results for the model input parameters and comment on the implications for hydrodynamics descriptions of small collision systems.
+In Sec.~\ref{sec:calibration}, we calibrate free parameters of the model to fit charged particle yields, mean $p_T$ and anisotropic flow coefficients of \emph{both} collision systems at $\sqrts=5.02$~TeV, and finally, in Secs.~\ref{sec:results} and \ref{sec:summary}, we present posterior results for the model input parameters and comment on the implications for hydrodynamic descriptions of small collision systems.
 
@@
 \end{equation}
-where $\sigma_\mathrm{eff}$ is an effective cross section for pairwise inelastic collisions between the constituents, and $\sigma_{pp}^\mathrm{inel}$ is the total inelastic proton-proton cross section:
+where $\sigma_\mathrm{eff}$ is an effective cross section for pairwise inelastic collisions between the constituents, and $\sigma_{pp}^\mathrm{inel}$ is the total inelastic proton-proton cross section
 \begin{equation}
@@
 equal to the beam-integrated proton density in Eqn.~\eqref{eq:rho}, with each constituent shifted by the appropriate impact parameter offset, and multiplied by a gamma random variable $\gamma_i$ with unit mean and variance $1/k$.
-These \emph{ad hoc} gamma random weights are necessary to describe the large fluctuations observed in minimum bias proton-proton collisions, although their physical justification is not yet well understood.
+These \emph{ad hoc} gamma random weights are necessary to describe the large fluctuations observed in minimum bias proton-proton collisions, although their physical origin is not yet well understood.
 
 Within the eikonal approximation, the initial entropy deposited at midrapidity at proper time $\tau=0^+$~fm/$c$ can be represented by some mapping ${f(\T_A, \T_B) \mapsto dS/dy}$.
-A natural first guess for the function $f$ is an arithmetic mean $f \propto (\T_A + \T_B)/2$, equal to a wounded nucleon model up to meaningless factor of two in the overall normalization.
-The wounded nucleon model was in fact the first such mapping used as a proxy for initial particle production and entropy deposition in relativistic heavy-ion collisions \cite{Bialas:1976ed}.
+A natural first guess for the function $f$ is an arithmetic mean $f \propto (\T_A + \T_B)/2$ of participant nuclear matter $\T_{A,B}$, which is simply a wounded nucleon model up to meaningless factor of two in the overall normalization.
+The wounded nucleon model was in fact one of the first such mappings used as a proxy for initial particle production and entropy deposition in relativistic heavy-ion collisions \cite{Bialas:1976ed}.
 It was quickly realized, however, that the wounded nucleon model predicts the wrong scaling for charged particle production as a function of collision centrality and hence the wrong scaling for initial entropy deposition as a function of participant thickness $\T_{A,B}$ \cite{Kharzeev:2000ph}.
@@
 Conveniently, it has been shown that the generalized mean ansatz is able to mimic the scaling behavior of certain saturation based initial condition models \cite{Bernhard:2016tnd}, and hence it should serve as a reasonable parametric form for exploring QGP entropy deposition, assuming imperfect knowledge of saturation effects in nature.
-Of course the model is \emph{not} a substitute for first principle saturation calculations, and may fail to reproduce nuanced features of \emph{ab initio} models such as the existence of short-range gluon field fluctuations \cite{Schenke:2012wb}.
+Of course the model is \emph{not} a substitute for first principle theory calculations, and it may fail to reproduce nuanced features of \emph{ab initio} models such as the existence of short-range gluon field fluctuations \cite{Schenke:2012wb}.
 
@@
   \label{eq:shear_viscosity}
-  (\eta/s)(T) = (\eta/s)_\mathrm{min} + (\eta/s)_\mathrm{slope}\cdot(T - T_c)\cdot(T/T_c)^{(\eta/s)_\mathrm{crv}}
+  (\eta/s)(T) = (\eta/s)_\mathrm{min} + (\eta/s)_\mathrm{slope}\cdot(T - T_c)\cdot(T/T_c)^{(\eta/s)_\mathrm{crv}},
 \end{equation}
@@
 
-The aforementioned hydrodynamic equations are solved numerically using the boost-invariant \texttt{VISH2+1} viscous hydrodynamics code \cite{Song:2007ux} which is documented extensively in Ref.~\cite{Shen:2014vra}.
-We simulate each hydrodynamic event on a spacetime grid with transverse extent $x_\mathrm{max}$, spatial grid step $dx$, and time step $d\tau$ which are optimized \emph{event-by-event} to balance trade-offs between numerical accuracy and computation time (see appendix).
+The aforementioned hydrodynamic equations are solved numerically using the boost-invariant \textsc{VISH2+1} viscous hydrodynamics code \cite{Song:2007ux, Shen:2014vra}.
+We simulate each hydrodynamic event on a spacetime grid with transverse extent $x_\mathrm{max}$, spatial grid step $dx$, and time step $d\tau$ which are optimized \emph{event-by-event} to balance trade-offs between numerical accuracy and computation time (see Appendix~\ref{app:adaptive_grid}).
 Although these details are somewhat mundane, they are critically important to the present study, since the computation time scales with the number of spacetime cells $n_x^2\, n_\tau$, and $n_x$ and $n_\tau$ typically have to be quite large to resolve the small length scales associated with nucleon substructure.
@@
 
-\subsubsection{Sampling with finite resonance widths}
-
 Traditionally, particlization models have sampled resonances using each particle's pole mass in Eqn.~\eqref{distribution}.
@@
 where $\Gamma_0$ is the usual Breit-Wigner width and $m_\mathrm{min}$ is a production threshold equal to the total mass of the lightest decay products.
-
 We tabulate the values of $\{\Gamma_0, m_0, m_\mathrm{min}\}$ for all particles and sample the masses of each particle during particlization \cite{PDG:2017}.
 The resonances are then passed to a hadronic transport model, described shortly, which simulates subsequent scatterings and decays.
-
-\subsubsection{Viscous corrections to the distribution function}
 
@@
 
-\subsubsection{Boltzmann transport}
-
-Once the fluid is converted into particles, we simulate its subsequent microscopic dynamics using the Ultra-relativistic Quantum Molecular Dynamics (\texttt{UrQMD}) transport model \cite{Bass:1998ca, Bleicher:1999xi}.
+Once the fluid is converted into particles, we simulate its subsequent microscopic dynamics using the Ultra-relativistic Quantum Molecular Dynamics (\textsc{UrQMD}) transport model \cite{Bass:1998ca, Bleicher:1999xi}.
 It solves the microscopic Boltzmann equation
@@
 
-One primary advantage of using a microscopic transport model such as \texttt{UrQMD} as an afterburner, is that it realistically simulates the system break-up when the mean free path becomes large relative to the system size.
+One primary advantage of using a microscopic transport model such as \textsc{UrQMD} as an afterburner, is that it realistically simulates the system break-up when the mean free path becomes large relative to the system size.
 This dilute limit is expected to play a significant role in small collision systems where the produced medium is smaller and shorter lived.
@@
 The nuclear collision model constructed in Sec.~\ref{sec:model} includes a number of free parameters $\x$ which describe the initial state, pre-equilibrium dynamics, and hydrodynamic medium.
-Given values for the parameters $\x$, the model may be used to predict a vector of simulated experimental observables $\y_m$.
+Given values for the parameters $\x$, the model may be used to predict a vector of simulated observables $\y_m$.
 For example, $\y_m$ might be a vector consisting of charged particle yields in different centrality bins.
@@
 Several of the model parameters require special care and are reparametrized accordingly:
-\begin{itemize}[leftmargin=2\parindent]
+\begin{itemize}[leftmargin=1\parindent]
   \item The constituent fluctuations, modeled by the gamma random weights $\gamma_i$ in Eqn.~\eqref{eq:fluctuated_thick}, generate overall nucleon fluctuations which are suppressed by the number of constituents $n_c$ inside the nucleon.
@@
     \end{enumerate*}
-    In the limit $n_c \to \infty$, the composite nucleon's root mean square (rms) radius is simply the convolution of its sampling radius $r$ and constituent width $v$ which add together in quadrature
+    In the limit $n_c \to \infty$, the composite nucleon's root-mean-square (RMS) radius is simply the convolution of its sampling radius $r$ and constituent width $v$ which add together in quadrature
   \begin{equation}
-    \lim_{n_c \to\infty} \mathrm{RMS}\{\rho(\mathbf{x})\} = \sqrt{r^2 + v^2}.
+    \lim_{n_c \to\infty} r_\mathrm{RMS}\{\rho\} = \sqrt{r^2 + v^2}.
   \end{equation}
@@
   \end{equation}
-  depicted in Fig.~\ref{fig:nucleon_schematic}, which approximates the rms radius of the nucleon when the number of sampled constituents is large.
-  We thus call $w$ a nucleon ``width'', although for smaller numbers of constituents, the \emph{actual} rms radius of our sampled nucleons can be significantly smaller than our width parameter $w$ due to fluctuations in the nucleon's center of mass, and hence one should account for the difference when discussing the nucleon's posterior rms radius.
+  which approximates the RMS radius of the nucleon when the number of sampled constituents is large (see Fig.~\ref{fig:nucleon_schematic} for an example).
+  We thus call $w$ a nucleon ``width'', although for smaller numbers of constituents, the \emph{actual} RMS radius of our sampled nucleons can be significantly smaller than our width parameter $w$ due to fluctuations in the nucleon's center of mass, and hence one should account for the difference when discussing the nucleon's posterior RMS radius.
 
@@
 
-\begin{figure*}[t]
-  \includegraphics{correlation_matrices}
-  \caption{Correlation matrices for Pb-Pb collisions at $\sqrts=5.02$~TeV for the model/emulator (left) at a random point in parameter space and the experimental data (right) used to construct the covariance matrix ${\Sigma = \Sigma_m + \Sigma_e}$ in Eqn.~\eqref{eq:likelihood}.
-Experimental statistical and systematic errors are from ALICE and CMS \note.
-The experimental correlation structure is modeled using Eqn.~\eqref{eq:corr}.}
-  \label{fig:correlation}
-\end{figure*}
-
 The likelihood function \eqref{eq:likelihood} provides evidence for \mbox{(or against)} the model parameters $\x$ by comparing the model predictions $\y_m(\x)$ to experimental data $\y_e$.
@@
 \end{itemize}
-Each observable is calculated from the list of final state particles produced by \texttt{UrQMD} using the same methods applied by experiment.
-We generally match the kinematic cuts of all measurements with two exceptions: we use use a larger rapidity interval $|\eta| < 0.8$ than experiment for some boost-invariant observables to improve our finite particle statistics, and we do not apply a finite rapidity gap, e.g.\ $|\Delta \eta| > 1$, between pairs of particles when calculating the two-particle cumulant $\vnk{n}{2}$ since we already oversample particles from each hydrodynamic event, and this oversampling suppresses non-flow correlations.
+Each observable is calculated from the list of final state particles produced by \textsc{UrQMD} using the same methods applied by experiment.
+We generally match the kinematic cuts of all measurements with two exceptions: we use use a larger rapidity interval $|\eta| < 0.8$ than experiment for some boost-invariant observables to improve our finite particle statistics, and we do not apply a rapidity gap, e.g.\ $|\Delta \eta| > 1$, between pairs of particles when calculating the two-particle cumulant $\vnk{n}{2}$ since we already oversample particles from each hydrodynamic event, and this oversampling suppresses non-flow correlations.
 
@@
 Most of the calibration observables listed in Tbl.~\ref{tab:observables} are calculated as a function of collision centrality, where centrality is defined using some measure of the underlying event activity, e.g.\ the charged particle yield in a given rapidity window.
-When calculating these observables, we generate \order{10^4} minimum-bias events at each design point and divide the events into centrality bins using the charged particle yield at midrapidity, similar to the procedure used by experiment.
+When calculating these observables, we generate \order{4} minimum-bias events at each design point and divide the events into centrality bins using the charged particle yield at midrapidity, similar to the procedure used by experiment.
 
@@
 
-Finally, we mimic the method used by experiment and apply \eqref{eq:mult_trigger} to select rare high-multiplicity events from a continuous stream of minimum-bias \trento\ events that satisfy the correct relative multiplicity bin edges.
+\begin{figure*}[t]
+  \includegraphics{correlation_matrices}
+  \caption{Visualization of the Pb-Pb correlation matrix $\mathrm{corr}(y_i, y_j) = \mathrm{cov}(y_i,y_j)/(\sigma_i \sigma_j)$ for the model (emulator) at a random point in parameter space (left-side) and for the experimental data (right-side).
+Each cell represents an observable in a single centrality bin.
+Experimental statistical and systematic errors are from ALICE \cite{Adam:2015ptt, Adam:2016izf}.
+The experimental correlation structure is modeled using Eqn.~\eqref{eq:corr}.}
+  \label{fig:correlation}
+\end{figure*}
+
+Finally, we mimic the method used by experiment and apply \eqref{eq:mult_trigger} to select rare high-multiplicity events from a continuous stream of minimum-bias \trento\ events satisfying the correct relative multiplicity bin edges.
 This of course means that, in addition to running a large sample of minimum-bias events for centrality binned observables, we must also generate (much like experiment) a separate sample of multiplicity triggered events.
@@
 \begin{equation}
-  \Sigma = \Sigma_m + \Sigma_e,
+  \Sigma = \Sigma_m + \Sigma_e
 \end{equation}
-which includes uncertainty contributions from both the model $\Sigma_m$ and experimental data $\Sigma_e$.
+in Eqn.~\eqref{eq:likelihood}, which includes uncertainty contributions from both the model $\Sigma_m$ and experimental data $\Sigma_e$.
 The experimental contribution to the covariance $\Sigma_e$ is further broken down into statistical and systematic components,
@@
 \end{equation}
-
-The statistical errors are uncorrelated and thus its covariance matrix is diagonal:
+The statistical errors in $\Sigma_e^\text{stat}$ are uncorrelated and thus its covariance matrix is diagonal:
 \begin{equation}
@@
 where $b_i$ and $b_j$ are the midpoints of two observable bins of a single type (centrality or relative multiplicity) and $l$ is a correlation length which describes how quickly the observable bins decorrelate as the distance between the bins increases.
-We use centrality correlation lengths ${l=100\%}$ for all Pb-Pb observables and ${l=30\%}$ for the $p$-Pb charged particle yield $d\nch/d\eta$.
-The $p$-Pb mean $p_T$ and flow observables, meanwhile, use relative multiplicity bins $\nch / \langle \nch \rangle$ and $\ntrk / \langle \ntrk \rangle$ so we use a much smaller correlation length $l=5$ to reflect the smaller range of bin values.
-We show, as an example, the resulting experimental correlation matrix $\text{corr}(\Sigma_e)$ for the Pb-Pb system on the right side of Fig.~\ref{fig:correlation}.
-The correlation matrix is block diagonal, with each block representing the correlations within a single class of observable; in this case, the experimental yield $d\nch/d\eta$ and flows $\vnk{n}{2}$.
+We use centrality correlation lengths $l=100$ for all of the centrality-binned Pb-Pb observables and $l=30$ for the centrality-binned $p$-Pb charged particle yield $d\nch/d\eta$.
+The \mbox{$p$-Pb} mean $p_T$ and flow observables, meanwhile, use relative multiplicity bins $\nch / \langle \nch \rangle$ and $\ntrk / \langle \ntrk \rangle$ which necessitate a smaller correlation length $l=5$.
+
+We show, as an example, the resulting correlation matrix
+\begin{equation}
+  \mathrm{corr}(y_i, y_j) = \mathrm{cov}(y_i, y_j)/(\sigma_i \sigma_j)
+\end{equation}
+for the Pb-Pb experimental data constructed using Eqn.~\eqref{eq:corr} on the right-side of Fig.~\ref{fig:correlation}.
+Here $y_i$ denotes an element of the experimental data $\y_e$ and $\sigma_i$ its corresponding uncertainty.
+The correlation matrix is block diagonal, with each block representing the correlations within a single class of observable.
 
@@
 In principle, one could calculate the likelihood function in Eqn.~\eqref{eq:likelihood} directly, e.g.\ by running the model a large number of times at a given parameter point $\x$ to calculate the model observables $\y_m(\x)$ from the ensemble of simulated events, but in practice such a procedure would be intractable.
-The model is computationally intensive to evaluate, ${\sim}10$~min/event per processor core, and thousands of events are required to calculate the simplest observables at a single parameter point.
+The model is computationally intensive to evaluate, and thousands of events are required to calculate the simplest observables at a single parameter point.
 Moreover, we need to evaluate Eqn.~\eqref{eq:likelihood} \emph{numerous} times in order to sample the multidimensional posterior distribution so that the samples may be histogrammed and visualized.
 
-We therefore follow an established framework for computationally intensive models and train an emulator to act as a fast surrogate for the full physics simulation \note.
+We therefore follow an established framework for computationally intensive models and train an emulator to act as a fast surrogate for the full physics simulation \cite{OHagan:2006ba, Higdon:2008cmc, Higdon:2014tva}.
 The emulator enables essentially instantaneous predictions for $\y_m = f(\x)$ and allows us to sample the posterior distribution millions of times.
-In order to train the emulator, we first generate a scaffolding of the parameter space using a maximin Latin hypercube design \note\ to distribute 500 points uniformly throughout our 15-dimensional parameter space according to the parameter ranges in Tbl.~\ref{tab:design}.
-We then run minimum-bias and multiplicity triggered $p$-Pb and Pb-Pb events at each design point, and calculate the model observables from the ensemble of events.
+In order to train the emulator, we first generate a scaffolding of the parameter space using a maximin Latin hypercube design \cite{Morris:1995lh} to distribute 500 points uniformly throughout our 15-dimensional parameter space according to the parameter ranges in Tbl.~\ref{tab:design}.
+We then run minimum-bias and multiplicity triggered $p$-Pb and Pb-Pb events at each design point and calculate the model observables from the ensemble of events.
 
@@
 
-We use for this purpose a specific type of emulator known as a Gaussian process emulator (GPE).
+We use for this purpose a specific type of emulator known as a Gaussian process emulator (GPE) \cite{Rasmussen:2006gp}.
 The advantage of using GPE's is that they provide an estimate of their own uncertainty which allows us to account for this uncertainty when constructing the covariance matrix $\Sigma$ in Eqn.~\eqref{eq:likelihood}.
 One quirk of GPE's is that they are restricted to scalar-valued functions, i.e.\ functions of one output, whereas we require an emulator for vector-valued functions with multiple outputs.
-This restriction is commonly circumvented using principal component analysis (PCA): a general procedure which transforms a set of correlated variables $\y = (y_1, y_2, \dots, y_m)$ into a new basis representation $\z = (z_1, z_2, \dots, z_m)$ where the linear correlations between $z_i$, $z_j$ vanish for all $i\ne j \in m$ \note.
+This restriction is commonly circumvented using principal component analysis (PCA): a general procedure which transforms a set of correlated variables $\y = (y_1, y_2, \dots, y_m)$ into a new basis representation $\z = (z_1, z_2, \dots, z_m)$ where the linear correlations between $z_i$, $z_j$ vanish for all $i\ne j \in m$ \cite{Tipping:1999}.
 
@@
 The principal components are reported in order of explained variance, with the first principal component vector $\mathbf{v}_1$ accounting for the most variance in $\tilde{Y}$, and the last principal component vector $\mathbf{v}_m$ accounting for the least.
-We then train a set of independent GPE's,
-\begin{align}
-  \label{eq:gpe}
-  z_1 &= \text{gp}_1(\x), \nonumber \\
-  z_2 &= \text{gp}_2(\x), \nonumber \\
-  \vdots & \nonumber \\
-  z_k &= \text{gp}_k(\x),
-\end{align}
-to predict the principal components (coefficients) $(z_1, \dots, z_k)$ of the first $k$ principal component vectors as a function of the model parameters $\x$ which vary across the design $X$.
-We use $k=7$ principal components when emulating the $p$-Pb system and $k=8$ principal components when emulating Pb-Pb, chosen to describe 99.5\% of the total observed variance of each system.
+We then train a set of independent GPE's $z_i = \mathrm{gp}_i(\x)$ for $i=1$ to $k$ to predict the principal components $(z_1, \dots, z_k)$ of the first $k$ principal component vectors as a function of the model parameters $\x$ which vary across the design $X$.
+For the present study, we use $k=7$ principal components when emulating the $p$-Pb system and $k=8$ principal components when emulating Pb-Pb, chosen to describe 99.5\% of the total observed variance of each system.
 
@@
 Each GPE reports a mean value $z(\x)$ as well as an estimated error $\delta z(\x)$ which accounts for statistical noise in the training data and interpolation error between the design points.
-Once the GPE's \eqref{eq:gpe} are trained, we can predict the model observables $\y_m$ at parameter point $\x$ by transforming the vector of principal components
+Once the GPE's are trained, we can predict the model observables $\y_m$ at parameter point $\x$ by transforming the vector of emulated principal components
 \begin{equation}
@@
 \begin{equation}
-\text{Cov}(z_i, z_j) = \text{diag}[(\delta z_{1})^2, (\delta z_2)^2, \dots, (\delta z_k)^2 ],
+\mathrm{cov}(z_i, z_j) = \mathrm{diag}[(\delta z_{1})^2, (\delta z_2)^2, \dots, (\delta z_k)^2 ],
 \end{equation}
-
-\begin{fullpage}
-{\centering\includegraphics{observables_pbpb}\par}
-\captionof{figure}{Simulated observables compared to experimental data for Pb-Pb collisions at $\sqrts=5.02$~TeV.
-Top row: explicit model calculations (no emulator) for each of the 500 design points; bottom row: emulator predictions for $n=100$ random samples drawn from the posterior.
-Points with error bars are experimental data from ALICE with statistical and systematic errors added in quadrature \note.}
-\vspace{.5cm}
-{\centering\includegraphics{observables_ppb}\par}
-\captionof{figure}{Same as Fig.~\ref{fig:obs_pbpb} but for $p$-Pb collisions at $\sqrts=5.02$~TeV.
-Note that multiplicity bins are used for mean $p_T$ and flow cumulant observables to match the bins used by experiment.
-Experimental data are from ALICE and CMS \note.}
-\end{fullpage}
-
 and transform it back to physical space as well to obtain the covariance matrix of the model observables $\y_m$ at a given parameter point $\x$.
@@
 The resulting emulator therefore predicts both a mean prediction $\y_m(\x)$ and an uncertainty covariance matrix $\Sigma_m(\x)$ which accounts for multiple sources of model and emulator uncertainty, including the truncation error expected from using a finite number of principal components $k < m$.
-The covariance matrix $\Sigma_\text{emu}$ thus includes the \emph{known-unknowns}, but not the \emph{unknown-unknowns}, such as the overall credibility of viscous hydrodynamics, which are difficult and often impossible to quantify.
-We show in Fig.~\ref{fig:correlation} the correlation structure $\text{Corr}(\Sigma)$ of the model (emulator) covariance $\Sigma_m(\x)$ for a random parameter point $\x$ in the design space (left-side), along side the same correlation matrix for the experimental covariance (right-side) for the Pb-Pb collision system.
-For additional information on the model emulator, we direct the reader to the appendix which includes several validation tests of the emulator prediction accuracy \note.
+\begin{fullpage}
+{\centering\includegraphics{observables_pbpb}\par}
+\captionof{figure}{\label{fig:obs_pbpb} Simulated observables compared to experimental data for Pb-Pb collisions at $\sqrts=5.02$~TeV.
+Top row: explicit model calculations (no emulator) for each of the $d=500$ design points; bottom row: emulator predictions for $n=100$ random samples drawn from the posterior.
+Points with error bars are experimental data from ALICE with statistical and systematic errors added in quadrature \cite{Adam:2015ptt, Adam:2016izf}.}
+\vspace{.5cm}
+{\centering\includegraphics{observables_ppb}\par}
+\captionof{figure}{\label{fig:obs_ppb} Same as Fig.~\ref{fig:obs_pbpb} but for $p$-Pb collisions at $\sqrts=5.02$~TeV.
+Note that multiplicity bins are used for mean $p_T$ and flow cumulant observables to match the bins used by experiment.
+Experimental data are from ALICE \cite{Adam:2014qja, Abelev:2013bla} and CMS \cite{Chatrchyan:2013nka}.}
+\end{fullpage}
+The model covariance matrix $\Sigma_m$ includes so-called \emph{known}-unknowns such as statistical error and emulator interpolation error, but not \emph{unknown}-unknowns such as the overall validity of small-system hydrodynamics, i.e.\ things which lack a unified consensus or are difficult to quantify.
+We show in Fig.~\ref{fig:correlation} the resulting Pb-Pb correlation matrix $\mathrm{corr}(y_i,y_j)$ for the model (emulator) at a random parameter point $\x$ in the design space (left-side), along side the same correlation matrix for the experimental data (right-side) discussed previously.
+For additional information on the model emulator, we direct the reader to Appendix~\ref{app:validation} which includes several validation tests of the emulator prediction accuracy.
 
@@
 
-We then perform Markov-chain Monte Carlo (MCMC) importance sampling on the posterior distribution in Eqn.~\eqref{eq:bayes} to draw random samples for $P(H_\x | E)$, the estimate of the true model parameters given the model and the experimental data.
-We use an affine-invariant ensemble sampler which uses a large ensemble of interdependent walkers \note\ and allow the MCMC chain to ``burn-in'' before generating \order{10^7} posterior samples.
+We then perform Markov-chain Monte Carlo (MCMC) importance sampling on the posterior distribution in Eqn.~\eqref{eq:bayes} to draw random samples for $P(H_\x | E)$, the estimate of the true model parameters given the model and the experimental data \cite{Goodman:2010en, FM:2013mc}.
+For this we use an affine-invariant ensemble sampler which uses a large ensemble of interdependent walkers \cite{Goodman:2010en, FM:2013mc} and allow the MCMC chain to ``burn-in'' before generating \order{7} posterior samples.
 
@@
 
-We show simulated and emulated model observables (thin colored lines) for Pb-Pb collisions (Fig.~\ref{fig:obs_pbpb}) and \mbox{$p$-Pb} collisions (Fig.~\ref{fig:obs_ppb}) at $\sqrts=5.02$~TeV compared to experimental data from CMS and ALICE \note.
+We show simulated and emulated model observables (thin colored lines) for Pb-Pb collisions (Fig.~\ref{fig:obs_pbpb}) and \mbox{$p$-Pb} collisions (Fig.~\ref{fig:obs_ppb}) at $\sqrts=5.02$~TeV compared to experimental data from CMS \cite{Chatrchyan:2013nka} and ALICE \cite{Adam:2015ptt, Adam:2016izf, Adam:2014qja, Abelev:2013bla}.
 The top row of each figure shows explicit model calculations at each of the $d=500$ design points (training data), while the bottom row shows emulator predictions for $n=100$ random parameter samples drawn from the Bayesian posterior.
@@
 We also note that although the $p$-Pb posterior samples have a somewhat larger spread than the Pb-Pb samples, the percentage uncertainty of the $p$-Pb emulator is similar to that of the Pb-Pb emulator, and thus the difference is likely due to the larger variance of the $p$-Pb training data.
-The uncertainty in the posterior distribution would thus be improved by running the calibration with more design points or with a narrower range of parameter values to increase the density of the training points and reduce interpolation error.
+The uncertainty in the posterior distribution could thus be improved by running the calibration with more design points or with a narrower range of parameter values to increase the density of the training points and reduce interpolation uncertainty.
 
-We now direct our attention to Fig.~\ref{fig:posterior} which shows the posterior distribution of the model input parameters.
+We now direct our attention to Fig.~\ref{fig:posterior} which shows the main result of this work, the posterior distribution of the model input parameters.
 Recall that the posterior $P(H_\x | E)$ is the probability that our hypothesis $H_\x$ for the true model parameters $\x_\star$ is correct, given the evidence $E$ provided by experiment.
-The present distribution has 15 dimensions, one for each parameter listed in Tbl.~\ref{tab:design}, and so we sample the distribution and histogram the samples to project the distribution onto a lower number of dimensions for visualization purposes.
-Each diagonal panel in the figure is the distribution of a \emph{single} model parameter (marginalized over all other parameters), and each lower-diagonal panel is the joint distribution of a \emph{pair} of model parameters (marginalized over all other parameters).
-The full posterior distribution includes a wealth of information.
-Here we focus on the most important features...
+The present posterior has 15 dimensions, one for each parameter listed in Tbl.~\ref{tab:design}, and thus its joint distribution cannot be summarized by one figure alone.
+We thus sample the distribution and histogram the samples to project the distribution onto one or two dimensions.
+Each diagonal panel in Fig.~\ref{fig:posterior} is the distribution of a \emph{single} model parameter (marginalized over all others), and each lower-diagonal panel is the joint distribution of a \emph{pair} of model parameters (marginalized over all others).
+The posterior includes a wealth of information.
+Here we focus on its most important features, starting with the top-left corner and working our way to the bottom-right.
+
+\subsection{Initial condition parameters}
+
+The \trento\ normalization factor is well constrained ${\text{Norm} = 20.3^{+2.2}_{-2.2}}$, as is the generalized mean parameter $p=0.018^{+0.146}_{-0.159}$ which fixes the scaling of initial QGP entropy deposition.
+
+are well constrained, and we see that the posterior emulator predictions in Fig.~\ref{fig:obs_ppb} and \ref{fig:obs_pbpb} nicely describe the charged particle yields $d\nch/d\eta$ of both collision systems.
+This is perhaps one of the strongest tests of the initial condition model.
 
@@
 \begin{figure}
-  \makebox[\textwidth]{
-      \includegraphics[width=.95\paperwidth]{posterior}
-  }
+  \makebox[\textwidth]{\includegraphics[width=.95\paperwidth]{posterior}}
 \end{figure}
-\captionof{figure}{Bayesian posterior distribution of the model input parameters.
+\captionof{figure}{\label{fig:posterior} Bayesian posterior distribution of the model input parameters.
 The diagonal panels show the marginalized distributions of individual model parameters, while off-diagonal panels show the joint distributions for pairs of model parameters, visualizing their correlations.
@@
 
-\begin{itemize}[leftmargin=1\parindent, itemsep=0pt, topsep=5pt]
-  \item The posterior initial condition normalization factor $\text{Norm} = 20.3^{+2.2}_{-2.2}$ (upper-left diagonal) is well constrained and describes both the $p$-Pb and Pb-Pb yields simultaneously (e.g.\ see Figs.~\ref{fig:obs_ppb}).
+which would not be the case if there was significant tension in fitting the $p$-Pb and Pb-Pb yields, or.
+Moreover, our
+It the emulator predictions using the posterior value describes both the $p$-Pb and Pb-Pb yields simultaneously (e.g.\ see Figs.~\ref{fig:obs_ppb}).
     We emphasize that this self-consistency is highly non-trivial since the ratio of $p$-Pb to Pb-Pb yields model can only fit both yields simultaneously if the \trento\ entropy deposition parameter $p$ is appropriately tuned.
-\end{itemize}
+
+\subsection{Verification of high-probability parameters}
+
+\begin{table}[t]
+  \caption{Maximum a posteriori (MAP) parameters.}
+  \begin{ruledtabular}
+  \begin{tabular}{lll}
+    Parameter         & Description                        & Value\\
+    \paddedhline
+    Norm              & Normalization factor                 & 20.2           \\
+    $p$               & Entropy deposition parameter         & 0.00    \\
+    $\sigmaf$         & Nucleon fluctuation std.\ dev.\      &             \\
+    $w$               & Nucleon width parameter              & 0.4--1.2 fm     \\
+    $n_c$             & Number of nucleon constituents       & 6            \\
+    $\X$              & Nucleon structure parameter          & 0--1            \\
+    $\dmin$           & Minimum inter-nucleon distance       & 0--1.7 fm       \\
+    $\taufs$          & Free-streaming time                  & 0.1--1.5 fm/$c$ \\
+    $\eta/s$ min      & Minimum value of $\eta/s$ (at $T_c$) & 0--0.2          \\
+    $\eta/s$ slope    & Slope of $\eta/s$ above $T_c$        & 0--8 GeV$^{-1}$ \\
+    $\eta/s$ crv      & Curvature of $\eta/s$ above $T_c$    & $-1$ to $+1$    \\
+    $\zeta/s$ max     & Maximum value of $\zeta/s$           & 0--0.1          \\
+    $\zeta/s$ width   & Width of $\zeta/s$ peak              & 0--0.1 GeV      \\
+    $\zeta/s$ $T_0$   & Temperature of $\zeta/s$ maximum     & 150--200 MeV    \\
+    $T_\text{switch}$ & Switching/particlization temp.       & 135--165 MeV    \\
+  \end{tabular}
+  \end{ruledtabular}
+  \label{tab:design}
+\end{table}
+
+The posterior in Fig.~\ref{fig:posterior} is the distribution of true parameter values, given the assumptions of our model and the evidence provided by the experimental data.
+Correspondingly, the probability of nature ``choosing'' any specific set of parameter values is \emph{zero}; we can only assign a probability to a finite region of parameter space.
+Nevertheless, we can define the \emph{maximum a posteriori probability} (MAP) estimate as the single most probable neighborhood (i.e.\ point) in parameter space, equal to the mode of the distribution.
+
+We estimate the MAP parameters for the posterior in Fig.~\ref{fig:posterior} and run \order{6} events for $p$-Pb and Pb-Pb (minimum bias and multiplicity triggered)
 
@@
 {\centering\includegraphics{observables_map}\par}
-\captionof{figure}{Model calculations using the \emph{maximum a posterior} (MAP) parameters compared to experiment.
+\captionof{figure}{\label{fig:obs_map} Model calculations using the \emph{maximum a posterior} (MAP) parameters compared to experiment.
 Colored lines are model calculations for $p$-Pb collisions (left) and Pb-Pb collisions (right) at $\sqrts=5.02$~TeV.
-Points with error bars are the experimental data with statistical uncertainties, and gray bands their corresponding systematic uncertainties, from ALICE and CMS \note.
+Points with error bars are the experimental data with statistical uncertainties, and gray bands their corresponding systematic uncertainties, from CMS \cite{Chatrchyan:2013nka} and ALICE \cite{Adam:2015ptt, Adam:2016izf, Adam:2014qja, Abelev:2013bla}.
 The ratio axes show the ratio of model over data where where available with gray bands indicating $\pm 10\%$.}
@@
   \includegraphics[width=\textwidth]{region_shear_bulk}
-  \caption{Left figure: estimated temperature dependence of the QGP specific shear viscosity $(\eta/s)(T)$ determined by the present Bayesian analysis of $p$-Pb and Pb-Pb collisions at $\sqrts=5.02$~TeV (orange line/band) compared to a previous Bayesian analysis of Pb-Pb collisions at $\sqrts=2.76$ and 5.02~TeV (blue line/band) \note.
+  \caption{Left figure: estimated temperature dependence of the QGP specific shear viscosity $(\eta/s)(T)$ determined by the present Bayesian analysis of $p$-Pb and Pb-Pb collisions at $\sqrts=5.02$~TeV (orange line/band) compared to a previous Bayesian analysis of Pb-Pb collisions at $\sqrts=2.76$ and 5.02~TeV (blue line/band) \cite{Bernhard:2018hnz}.
 The lines are the medians of each posterior distribution, and the bands are their 90\% credible regions.
@@
 \section{Event-by-event grid resizing}
-\label{appendix}
+\label{app:adaptive_grid}
 
@@
 
-The boost-invariant \texttt{VISH2+1} hydrodynamics code used in this work \cite{Song:2007ux, Shen:2014vra} runs on a Cartesian transverse grid specified by a maximum grid size $x_\text{max}$ and grid step width $dx$ which fix the transverse grid extent ${-x_\text{max} < x < x_\text{max}}$ and number of grid cells along each dimension ${n_x = 2\, x_\text{max} / dx}$.
+The boost-invariant \textsc{VISH2+1} hydrodynamics code used in this work \cite{Song:2007ux, Shen:2014vra} runs on a Cartesian transverse grid specified by a maximum grid size $x_\text{max}$ and grid step width $dx$ which fix the transverse grid extent ${-x_\text{max} < x < x_\text{max}}$ and number of grid cells along each dimension ${n_x = 2\, x_\text{max} / dx}$.
 In general, the maximum grid size $x_\text{max}$ should be set large enough to contain the full spacetime evolution of the event.
@@
 We find that we can quickly estimate the maximum radius $R_\text{max} = |\x_\text{max}|^2$ of the spacetime hypersurface ${T = T(e_\text{min})}$ by running the event on a coarse-grained spatial grid with one-third the spatial resolution we would otherwise require to resolve typical hydrodynamic observables such as mean $p_T$ and flows.
-The simulation time of a single \texttt{VISH2+1} event scales like ${\sim}n_x^3$ since $dx \propto d\tau$, and thus our ``pre-run'' event requires only ${\sim}1/27$th the time of a production event.
+The simulation time of a single \textsc{VISH2+1} event scales like ${\sim}n_x^3$ since $dx \propto d\tau$, and thus our ``pre-run'' event requires only ${\sim}1/27$th the time of a production event.
 We therefore run a coarse-grained pre-event on an excessively large grid for \emph{every} minimum-bias event to estimate $R_\text{max}$, then rerun the same event on a thrice finer grid with a trimmed spatial extent $x_\text{max} \equiv R_\text{max}$.
@@
 For instance, the $p$-Pb charged particle yield $d\nch/d\eta$ uncertainties are over predicted.
-It is not immediately clear why this would be the case, but the MAP observables in Fig.~\note\ are in good agreement with their emulator predictions which suggests it should not be a grave concern.
+It is not immediately clear why this would be the case, but the MAP observables in Fig.~\ref{fig:obs_map} are in good agreement with their emulator predictions which suggests it should not be a grave concern.
 

### 2018-07-23 paper/substructure.tex
More work on the results section

@@
 
-\newcommand{\trento}{\textsc{T\raisebox{-0.5ex}{R}ENTo}}
+\newcommand{\trento}{T\raisebox{-0.5ex}{R}ENTo}
 \newcommand{\sqrts}{\sqrt{s_{NN}}}
@@
 \begin{abstract}
-We posit a unified hydrodynamic and microscopic description of the quark-gluon plasma (QGP) produced in ultrarelativistic p-Pb and Pb-Pb collisions at $\sqrts=5.02$~TeV and evaluate our assertion using Bayesian inference.
+We posit a unified hydrodynamic and microscopic description of the quark-gluon plasma (QGP) produced in ultrarelativistic $p$-Pb and Pb-Pb collisions at $\sqrts=5.02$~TeV and evaluate our assertion using Bayesian inference.
 Specifically, we model the dynamics of both collision systems using initial conditions with parametric nucleon substructure, a pre-equilibrium free-streaming stage, event-by-event viscous hydrodynamics, and a microscopic hadronic afterburner.
 Free parameters of the model which describe the initial state and QGP medium are then simultaneously calibrated to fit charged particle yields, mean $p_T$ and flow cumulants.
-We argue that the global agreement of the calibrated model with the experimental data strongly supports the existence of hydrodynamic flow in small collision systems at ultrarelativistic energies, and that the flow produced develops at length scales smaller than a proton.
+We argue that the global agreement of the calibrated model with the experimental data strongly supports the existence of hydrodynamic flow in small collision systems at ultrarelativistic energies, and that the flow produced develops at length scales smaller than a single proton.
 Posterior estimates for the model's input parameters are obtained, and new insights into the temperature dependence of the QGP transport coefficients and event-by-event structure of the proton are discussed.
@@
 
-Ultrarelativistic nuclear collisions between one light-ion and one heavy-ion, e.g.\ $^3$He-Au and p-Pb collisions, generate dense, compact sources of nuclear matter which produce long-range multiparticle correlations that are strikingly similar to the correlations observed in heavy-ion collisions where collectivity is commonly explained by the existence of hydrodynamic flow \cite{CMS:2012qk, Abelev:2012ola, Aad:2012gla, Adare:2015ctn}.
+Ultrarelativistic nuclear collisions between one light-ion and one heavy-ion, e.g.\ $^3$He-Au and $p$-Pb collisions, generate dense, compact sources of nuclear matter which produce long-range multiparticle correlations that are strikingly similar to the correlations observed in heavy-ion collisions where collectivity is commonly explained by the existence of hydrodynamic flow \cite{CMS:2012qk, Abelev:2012ola, Aad:2012gla, Adare:2015ctn}.
 This observation suggests that hydrodynamic behavior could be manifest in small droplets of quark-gluon-plasma (QGP) \cite{Bozek:2011if, Bozek:2013uha}, and that flow might develop at length scales smaller than a single proton \note.
@@
 
-Independently, studies of coherent and incoherent $J/\psi$ production based on a color dipole picture of vector meson production were used to simultaneously constrain both the average color charge density of the proton as well as its event-by-event fluctuations in a saturation based framework \cite{Mantysaari:2016ykx, Aaron:2009aa, Abramowicz:2015mha}.
+Independently, studies of coherent and incoherent $J/\psi$ production based on a color dipole picture of vector meson production were used to simultaneously constrain both the average color charge density of the proton as well as its event-by-event fluctuations in a saturation based framework \cite{Mantysaari:2016ykx, Mantysaari:2016jaz, Aaron:2009aa, Abramowicz:2015mha}.
 Initial condition studies using the IP-Glasma model of color-glass condensate effective field theory \cite{Schenke:2012wb} simultaneously demonstrated that these color charge fluctuations leave a lasting imprint on the \mbox{small-x} gluon distribution of the proton and hence the initial geometry of QGP energy deposition \cite{Schlichting:2014ipa}.
@@
 We begin, in Sec.~\ref{sec:model}, by constructing a nuclear collision model for $p$-Pb and Pb-Pb collisions using initial conditions with parametric nucleon substructure, and transport dynamics described by a pre-equilibrium free-streaming stage, viscous hydrodynamics and microscopic Boltzmann transport.
-In Sec.~\ref{sec:calibration}, we calibrate free parameters of the model to fit charged particle yields, mean $p_T$ and anisotropic flow coefficients of \emph{both} collision systems at $\sqrts=5.02$~TeV, and finally, in Secs.~\ref{sec:results} and \ref{sec:summary}, we present posterior results for the model input parameters and comment on the implications for hydrodynamic descriptions of small collision systems.
+In Sec.~\ref{sec:calibration}, we calibrate free parameters of the model to fit charged particle yields, mean $p_T$ and anisotropic flow cumulants of \emph{both} collision systems at $\sqrts=5.02$~TeV, and finally, in Secs.~\ref{sec:results} and \ref{sec:summary}, we present posterior results for the model input parameters and comment on the implications for hydrodynamic descriptions of small collision systems.
 
@@
 equal to the beam-integrated proton density in Eqn.~\eqref{eq:rho}, with each constituent shifted by the appropriate impact parameter offset, and multiplied by a gamma random variable $\gamma_i$ with unit mean and variance $1/k$.
-These \emph{ad hoc} gamma random weights are necessary to describe the large fluctuations observed in minimum bias proton-proton collisions, although their physical origin is not yet well understood.
+These \emph{ad hoc} gamma random weights are necessary to describe the large fluctuations observed in minimum bias proton-proton collisions, although their exact physical origin is not well understood.
 
 Within the eikonal approximation, the initial entropy deposited at midrapidity at proper time $\tau=0^+$~fm/$c$ can be represented by some mapping ${f(\T_A, \T_B) \mapsto dS/dy}$.
-A natural first guess for the function $f$ is an arithmetic mean $f \propto (\T_A + \T_B)/2$ of participant nuclear matter $\T_{A,B}$, which is simply a wounded nucleon model up to meaningless factor of two in the overall normalization.
+A natural first guess for the mapping is the arithmetic mean
+\begin{equation}
+  \frac{dS}{dy}\bigg\vert_{y=0} \propto \frac{\T_A + \T_B}{2}
+\end{equation}
+of the participant nuclear matter $\T_A$ and $T_B$, which is simply a wounded nucleon model up to meaningless factor of two in the normalization.
 The wounded nucleon model was in fact one of the first such mappings used as a proxy for initial particle production and entropy deposition in relativistic heavy-ion collisions \cite{Bialas:1976ed}.
-It was quickly realized, however, that the wounded nucleon model predicts the wrong scaling for charged particle production as a function of collision centrality and hence the wrong scaling for initial entropy deposition as a function of participant thickness $\T_{A,B}$ \cite{Kharzeev:2000ph}.
-
-A simple remedy is to replace the arithmetic mean in the wounded nucleon model with a more flexible parametrization
+It was quickly realized, however, that the wounded nucleon model predicts the wrong scaling for charged particle production as a function of collision centrality and hence the wrong scaling for initial entropy deposition as a function of participant thicknesses $\T_A$ and $\T_B$ \cite{Kharzeev:2000ph}.
+
+A simple remedy is to replace the arithmetic mean of the wounded nucleon model with a more flexible parametrization
 \begin{equation}
@@
   \end{cases}
+  \label{eq:gmean}
 \end{equation}
@@
   \caption{Cartoon of the free-streaming approximation for hydrodynamic initialization.
-The initial state is free-streamed for proper time $\taufs$ (zero-coupling) before it is matched to hydrodynamics (strong-coupling).
+The initial state is free-streamed for proper time $\taufs$ (zero coupling) before it is matched to hydrodynamics (strong coupling).
 This piecewise evolution approximates the more realistic scenario expected in nature where the medium's coupling strength smoothly changes as a function of time.}
@@
     \end{equation}
-    where $k$ is the shape parameter of the gamma distribution used to fluctuate each constituent.
+    where $k$ is the shape parameter of the gamma distribution used to fluctuate each individual constituent.
   \item In Sec.~\ref{sec:initial_state} we parametrized nucleon substructure using three degrees of freedom:
@@
   \end{equation}
-  Here we choose a minimum width $v_\mathrm{min} = 0.2$~fm determined by computational limits and a maximum width $v_\mathrm{max} = w$ equal to the nucleon width.
+  Here we choose a minimum constituent width $v_\mathrm{min} = 0.2$~fm determined by computational limits and a maximum width $v_\mathrm{max} = w$ equal to the nucleon width.
   Thus for $\X=0$, the nucleons consist of small, distinct hot spots, whereas for $\X=1$ the nucleon is a single Gaussian blob of width $w$.
@@
 \end{equation}
-between each event's initial secondary-parton multiplicity ${N_\rho = \int d^2x\,n(\x)}$ in Eqn.~\eqref{energy-momentum} and its final-state event activity, characterized by $\nch$ or $\ntrk$ \cite{Abelev:2013bla, Chatrchyan:2013nka}.
+between each event's initial secondary-parton density ${N_\rho = \int d^2x\,n(\x)}$ in Eqn.~\eqref{energy-momentum} and its final-state event activity, characterized by $\nch$ or $\ntrk$ \cite{Abelev:2013bla, Chatrchyan:2013nka}.
 
@@
 We show simulated and emulated model observables (thin colored lines) for Pb-Pb collisions (Fig.~\ref{fig:obs_pbpb}) and \mbox{$p$-Pb} collisions (Fig.~\ref{fig:obs_ppb}) at $\sqrts=5.02$~TeV compared to experimental data from CMS \cite{Chatrchyan:2013nka} and ALICE \cite{Adam:2015ptt, Adam:2016izf, Adam:2014qja, Abelev:2013bla}.
-The top row of each figure shows explicit model calculations at each of the $d=500$ design points (training data), while the bottom row shows emulator predictions for $n=100$ random parameter samples drawn from the Bayesian posterior.
+The top row of each figure shows explicit model calculations at each of the $d=500$ design points (training data), while the bottom row shows emulator predictions for $n=100$ random parameter samples drawn from the Bayesian posterior (sampled from the MCMC chain).
 Each column shows a different class of observable.
@@
 
-Conversely, the calibrated emulator predictions (bottom row of each figure) are far better constrained and nicely track the experimental data points.
+Conversely, the calibrated emulator predictions (bottom row of Figs.~\ref{fig:obs_pbpb} and \ref{fig:obs_ppb}) are far better constrained and nicely track the experimental data points.
 We emphasize here that the posterior parameter values are obtained from a \emph{simultaneous} calibration to $p$-Pb and Pb-Pb data, and thus they are self-consistent between the two systems.
@@
 The present posterior has 15 dimensions, one for each parameter listed in Tbl.~\ref{tab:design}, and thus its joint distribution cannot be summarized by one figure alone.
-We thus sample the distribution and histogram the samples to project the distribution onto one or two dimensions.
-Each diagonal panel in Fig.~\ref{fig:posterior} is the distribution of a \emph{single} model parameter (marginalized over all others), and each lower-diagonal panel is the joint distribution of a \emph{pair} of model parameters (marginalized over all others).
-The posterior includes a wealth of information.
-Here we focus on its most important features, starting with the top-left corner and working our way to the bottom-right.
-
-\subsection{Initial condition parameters}
-
-The \trento\ normalization factor is well constrained ${\text{Norm} = 20.3^{+2.2}_{-2.2}}$, as is the generalized mean parameter $p=0.018^{+0.146}_{-0.159}$ which fixes the scaling of initial QGP entropy deposition.
-
-are well constrained, and we see that the posterior emulator predictions in Fig.~\ref{fig:obs_ppb} and \ref{fig:obs_pbpb} nicely describe the charged particle yields $d\nch/d\eta$ of both collision systems.
-This is perhaps one of the strongest tests of the initial condition model.
+We thus sample the distribution and histogram the samples to project the distribution onto one or two dimensions at a time.
+Each diagonal panel is the distribution of a \emph{single} model parameter (marginalized over all others), and each lower-diagonal panel is the joint distribution of a \emph{pair} of model parameters (marginalized over all others).
+
+The posterior in Fig.~\ref{fig:posterior} includes a wealth of information.
+Here we focus on its most important features, starting with the top-left corner of the figure and working our way to the bottom-right.
+We report posterior estimates for individual parameters using their (marginalized) distribution medians and 90\% credible intervals.
+For example, $x=2.45_{-0.15}^{+0.20}$ corresponds to a posterior median $\tilde{x}=2.45$ and 90\% credible interval $x \in (2.30, 2.65)$.
 
@@
 
-which would not be the case if there was significant tension in fitting the $p$-Pb and Pb-Pb yields, or.
-Moreover, our
-It the emulator predictions using the posterior value describes both the $p$-Pb and Pb-Pb yields simultaneously (e.g.\ see Figs.~\ref{fig:obs_ppb}).
-    We emphasize that this self-consistency is highly non-trivial since the ratio of $p$-Pb to Pb-Pb yields model can only fit both yields simultaneously if the \trento\ entropy deposition parameter $p$ is appropriately tuned.
-
-\subsection{Verification of high-probability parameters}
-
-\begin{table}[t]
+\subsection{Initial condition parameters}
+
+The \trento\ normalization factor ${\text{Norm} = 20.3^{+2.2}_{-2.2}}$ and generalized mean parameter $p=0.018^{+0.146}_{-0.159}$ are well constrained by the present analysis.
+Moreover, their posterior values nicely describe the calibration observables of both collision systems as evidenced by Figs.~\ref{fig:obs_pbpb} and \ref{fig:obs_ppb}.
+While it would not be surprising, for example, to fit one or two of these observables using such a narrow range of values, the quality of the combined fit and the number of observables described is highly non-trivial.
+For example, as the entropy deposition parameter $p \rightarrow \pm\infty$, particle production scales like the local maximum (minimum) of the nuclear overlap density.
+This strongly affects the ratio of the predicted $p$-Pb and Pb-Pb yields and consequently---absent any additional information---excludes \emph{most} values of the parameter $p$.
+It just so happens that the small range of $p$-values needed to describe the aforementioned yield ratio are the same values needed to describe all of the calibration observables in the present study, \emph{and} numerous Pb-Pb observables at $\sqrts=2.76$~TeV \note.
+This work thus reaffirms an empirical scaling law reported in several previous studies \note, that entropy deposition at midrapidity in ultrarelativistic nuclear collisions scales approximately as
+\begin{equation}
+  \label{eq:geometric_mean}
+  \frac{dS}{dy}\bigg\vert_{y=0} \propto \sqrt{\T_A \T_B},
+\end{equation}
+where $\T_A$ and $\T_B$ are the participant thickness functions \eqref{eq:fluctuated_thick} of each nucleus.
+We emphasize that this specific analytic form should not be interpreted too literally.
+For instance, a generalized mean described by $p=0.05$ is well within our 90\% credible region, but it is not equal to the geometric mean in Eqn.~\ref{eq:geometric_mean}.
+This approximate form, nevertheless, has been shown to mimic the scaling behavior of several saturation-based theory calculations in high-energy QCD \cite{Bernhard:2016tnd}, and thus it appears to be consistent with first-principle theoretical expectations.
+
+Continuing down the diagonal in Fig.~\ref{fig:posterior}, we see that the nucleon width parameter $w=0.97_{-0.17}^{+0.19}$~fm.
+We caution that this parameter $w$ is \emph{not} the RMS radius of our nucleons due to idiosyncrasies of the constituent sampling procedure.
+We can, however, easily calculate the RMS radius for a specific nucleon width $w$, constituent width $v$ and constituent number $n_c$.
+For example, the single highest posterior probability region of parameter space prefers a nucleon width $w=0.92$~fm, constituent width $v=0.43$~fm and constituent number $n_c=6$.
+The corresponding RMS nucleon radius for these parameters is $R_n = 0.86$~fm, conspicuously close the proton's RMS electric charge radius $R_p = 0.879(8)$~fm \cite{Bernauer:2010wm}.
+
+This is perhaps the single largest difference between our work and the conclusions of recent saturation-based calculations which constrained the event-by-event fluctuations of the proton using a color-dipole picture of vector meson production \cite{Mantysaari:2016ykx, Mantysaari:2016jaz}.
+The studies find that the measured coherent and incoherent $J/\Psi$ spectra at HERA prefer a compact gluon distribution inside each nucleon with an RMS radius $R_g = 0.4$~fm, which is approximately \emph{half} the analogous nucleon length scale in our model and clearly a significant discrepancy.
+
+The posterior constituent number $n_c = 6.2_{-3.1}^{+2.8}$ clearly disfavors $n_c=1$ constituent, although there is no strong peak for $n_c > 1$ constituents.
+This suggests that initial condition models which lack nucleon substructure will perform provide a worse description $p$-Pb and Pb-Pb bulk observables simultaneously which is perhaps at odds with
+
+\begin{table}[b]
   \caption{Maximum a posteriori (MAP) parameters.}
@@
     \paddedhline
-    Norm              & Normalization factor                 & 20.2           \\
-    $p$               & Entropy deposition parameter         & 0.00    \\
-    $\sigmaf$         & Nucleon fluctuation std.\ dev.\      &             \\
-    $w$               & Nucleon width parameter              & 0.4--1.2 fm     \\
+    Norm              & Normalization factor                 & 20.            \\
+    $p$               & Entropy deposition parameter         & 0.0            \\
+    $k$               & Constituent gamma fluct.\ shape      & 0.19           \\
+    $w$               & Nucleon width parameter              & 0.92 fm        \\
     $n_c$             & Number of nucleon constituents       & 6              \\
-    $\X$              & Nucleon structure parameter          & 0--1            \\
-    $\dmin$           & Minimum inter-nucleon distance       & 0--1.7 fm       \\
-    $\taufs$          & Free-streaming time                  & 0.1--1.5 fm/$c$ \\
-    $\eta/s$ min      & Minimum value of $\eta/s$ (at $T_c$) & 0--0.2          \\
-    $\eta/s$ slope    & Slope of $\eta/s$ above $T_c$        & 0--8 GeV$^{-1}$ \\
-    $\eta/s$ crv      & Curvature of $\eta/s$ above $T_c$    & $-1$ to $+1$    \\
-    $\zeta/s$ max     & Maximum value of $\zeta/s$           & 0--0.1          \\
-    $\zeta/s$ width   & Width of $\zeta/s$ peak              & 0--0.1 GeV      \\
-    $\zeta/s$ $T_0$   & Temperature of $\zeta/s$ maximum     & 150--200 MeV    \\
-    $T_\text{switch}$ & Switching/particlization temp.       & 135--165 MeV    \\
+    $v$               & Constituent width                    & 0.43 fm        \\
+    $\dmin$           & Minimum inter-nucleon distance       & 0.81 fm        \\
+    $\taufs$          & Free-streaming time                  & 0.37 fm/$c$    \\
+    $\eta/s$ min      & Minimum value of $\eta/s$ (at $T_c$) & 0.11           \\
+    $\eta/s$ slope    & Slope of $\eta/s$ above $T_c$        & 1.6 GeV$^{-1}$ \\
+    $\eta/s$ crv      & Curvature of $\eta/s$ above $T_c$    & -0.29          \\
+    $\zeta/s$ max     & Maximum value of $\zeta/s$           & 0.032          \\
+    $\zeta/s$ width   & Width of $\zeta/s$ peak              & 0.024 GeV      \\
+    $\zeta/s$ $T_0$   & Temperature of $\zeta/s$ maximum     & 175 MeV        \\
+    $T_\text{switch}$ & Switching/particlization temp.       & 151 MeV        \\
   \end{tabular}
@@
 \end{table}
+
+\subsection{Verification of high-probability parameters}
 

### 2018-07-24 paper/substructure.tex
Write about MAP observables

@@
   \begin{equation}
+    \label{eq:struct_param}
     v = v_\mathrm{min} + \X (v_\mathrm{max} - v_\mathrm{min}).
@@
 \subsection{Observables}
+\label{sec:observables}
 
@@
 
+\begin{figure}
+  \includegraphics[width=\columnwidth]{posterior_proton_shape}
+  \caption{Posterior distribution for the nucleon width $w$ and constituent width $v$.
+The white trapezoidal region (lower-right) covers the initial range of allowed values for $w$ and $v$, while the gray triangular region (upper-left) indicates values which are excluded by the prior $v > w$.
+The posterior distribution, shown in blue, indicates the preferred values for $w$ and $v$ determined by the analysis.}
+  \label{fig:posterior_proton_shape}
+\end{figure}
+
 The \trento\ normalization factor ${\text{Norm} = 20.3^{+2.2}_{-2.2}}$ and generalized mean parameter $p=0.018^{+0.146}_{-0.159}$ are well constrained by the present analysis.
-Moreover, their posterior values nicely describe the calibration observables of both collision systems as evidenced by Figs.~\ref{fig:obs_pbpb} and \ref{fig:obs_ppb}.
+Moreover, their posterior values nicely describe the $p$-Pb and Pb-Pb calibration observables in Figs.~\ref{fig:obs_pbpb} and \ref{fig:obs_ppb}.
 While it would not be surprising, for example, to fit one or two of these observables using such a narrow range of values, the quality of the combined fit and the number of observables described is highly non-trivial.
-For example, as the entropy deposition parameter $p \rightarrow \pm\infty$, particle production scales like the local maximum (minimum) of the nuclear overlap density.
-This strongly affects the ratio of the predicted $p$-Pb and Pb-Pb yields and consequently---absent any additional information---excludes \emph{most} values of the parameter $p$.
-It just so happens that the small range of $p$-values needed to describe the aforementioned yield ratio are the same values needed to describe all of the calibration observables in the present study, \emph{and} numerous Pb-Pb observables at $\sqrts=2.76$~TeV \note.
+For example, consider the ratio of the $p$-Pb yield over the Pb-Pb yield, which imposes a strong constraint on physically reasonable initial condition models.
+As the entropy deposition parameter $p$ trends toward positive (negative) infinity, particle production scales like the local maximum (minimum) of the nuclear overlap density.
+The parameter $p$ thus strongly affects the $p$-Pb and Pb-Pb yield ratio.
+It just so happens that the small range of $p$-values needed to describe this yield ratio are the same values needed to describe all of the calibration observables in the present study, \emph{and} numerous Pb-Pb observables at $\sqrts=2.76$~TeV \note.
 This work thus reaffirms an empirical scaling law reported in several previous studies \note, that entropy deposition at midrapidity in ultrarelativistic nuclear collisions scales approximately as
@@
 We emphasize that this specific analytic form should not be interpreted too literally.
-For instance, a generalized mean described by $p=0.05$ is well within our 90\% credible region, but it is not equal to the geometric mean in Eqn.~\ref{eq:geometric_mean}.
-This approximate form, nevertheless, has been shown to mimic the scaling behavior of several saturation-based theory calculations in high-energy QCD \cite{Bernhard:2016tnd}, and thus it appears to be consistent with first-principle theoretical expectations.
+For instance, a generalized mean described by $p=0.05$ is well within our 90\% credible interval, but it is not equal to the geometric mean in Eqn.~\ref{eq:geometric_mean}.
+This approximate form nevertheless has been shown to mimic the scaling behavior of several saturation-based theory calculations in high-energy QCD \cite{Bernhard:2016tnd}, and thus appears consistent with theoretical expectations.
+
+\begin{figure*}
+  \includegraphics[width=\textwidth]{region_shear_bulk}
+  \caption{Left figure: estimated temperature dependence of the QGP specific shear viscosity $(\eta/s)(T)$ determined by the present Bayesian analysis of $p$-Pb and Pb-Pb collisions at $\sqrts=5.02$~TeV (orange line/band) compared to a previous Bayesian analysis of Pb-Pb collisions at $\sqrts=2.76$ and 5.02~TeV (blue line/band) \cite{Bernhard:2018hnz}.
+The lines are the medians of each posterior distribution, and the bands are their 90\% credible regions.
+Right figure: same as before, but for the temperature dependence of the QGP specific bulk viscosity $(\zeta/s)(T)$.}
+  \label{fig:region_shear_bulk}
+\end{figure*}
 
@@
 We can, however, easily calculate the RMS radius for a specific nucleon width $w$, constituent width $v$ and constituent number $n_c$.
-For example, the single highest posterior probability region of parameter space prefers a nucleon width $w=0.92$~fm, constituent width $v=0.43$~fm and constituent number $n_c=6$.
+For example, the single highest posterior probability region of the parameter space prefers a nucleon width $w=0.92$~fm, constituent width $v=0.43$~fm and constituent number $n_c=6$.
 The corresponding RMS nucleon radius for these parameters is $R_n = 0.86$~fm, conspicuously close the proton's RMS electric charge radius $R_p = 0.879(8)$~fm \cite{Bernauer:2010wm}.
-
 This is perhaps the single largest difference between our work and the conclusions of recent saturation-based calculations which constrained the event-by-event fluctuations of the proton using a color-dipole picture of vector meson production \cite{Mantysaari:2016ykx, Mantysaari:2016jaz}.
-The studies find that the measured coherent and incoherent $J/\Psi$ spectra at HERA prefer a compact gluon distribution inside each nucleon with an RMS radius $R_g = 0.4$~fm, which is approximately \emph{half} the analogous nucleon length scale in our model and clearly a significant discrepancy.
-
-The posterior constituent number $n_c = 6.2_{-3.1}^{+2.8}$ clearly disfavors $n_c=1$ constituent, although there is no strong peak for $n_c > 1$ constituents.
-This suggests that initial condition models which lack nucleon substructure will perform provide a worse description $p$-Pb and Pb-Pb bulk observables simultaneously which is perhaps at odds with
-
-\begin{table}[b]
+Those studies find that the measured coherent and incoherent $J/\Psi$ spectra at HERA prefer a compact gluon distribution inside each nucleon, with an RMS radius $R_g \approx 0.4$~fm, which is roughly \emph{half} the analogous RMS nucleon width preferred by our analysis.
+
+Meanwhile, the posterior on the constituent number $n_c$ is not sharply peaked.
+We therefore refrain from quoting a distribution median and 90\% credible interval, although we do note that the distribution clearly favors $n_c > 1$ constituents.
+This is not surprising.
+The \trento\ model mimics saturation-based initial condition models, and saturation models tend to produce ``proton-sized'' fireballs in $p$-Pb collisions \note.
+When the proton is spherically symmetric, the resulting proton-sized QGP is also largely symmetric and thus produces very little flow.
+Saturation-based models are therefore unable to describe the significant flow measured in high-multiplicity $p$-Pb collisions without nucleon substructure, or alternatively, some other source of additional correlations \note.
+
+The posterior on the nucleon substructure parameter $\X=0.34_{-0.17}^{+0.22}$, on the other hand, is particularly sharply peaked.
+Recall that this parameter, defined in Eqn.~\eqref{eq:struct_param}, interpolates between the minimum and maximum widths of each constituent and hence
+two different limits for the granularity of the nucleon.
+When $\X=0$, the nucleon is populated by $n_c$ small compact hot-spots, and when $\X=1$ the hot-spots are large and fully overlapping, restoring spherical symmetry.
+We use the parameter $\X$ purely for calibration purposes and then transform the variable back to a constituent width $v$ for discussion purposes.
+
+We plot the joint posterior distribution for the nucleon width parameter $w$ and constituent width $v$ in Fig.~\ref{fig:posterior_proton_shape}.
+The white trapezoidal region in the lower-right is the joint parameter prior, and the gray triangular region in the upper-left where $v > w$ is a region that we exclude from the prior.
+The posterior distribution, shown in blue, is remarkably well constrained.
+We report a posterior estimate for the constituent width $v=0.46_{-0.15}^{+0.21}$~fm which is roughly the same magnitude as the \emph{nucleon} hot-spots used in recent saturation-based substructure studies that employed IP-Glasma initial conditions \note.
+Evidently, it may be necessary to place an informative prior on our nucleon substructure parameters in order to resolve the apparent tension between our parameter values and those needed to describe DIS measurements at HERA.
+Alternatively, it is also possible that the fluctuations probed by coherent and incoherent $J/\Psi$ production are different than those probed by minimum-bias particle production.
+
+\subsection{Transport properties}
+
+In this section, we compare several of our posterior estimates to those obtained from a similar Bayesian analysis \cite{Bernhard:2018hnz} which used an (almost) identical version of the present physics model.
+The only modeling difference is the inclusion of nucleon substructure in this work which was absent in Ref.~\cite{Bernhard:2018hnz}.
+Several calibration details, however, are different between the two studies.
+The present work uses a much smaller number of $p$-Pb and Pb-Pb observables at $\sqrts=5.02$~TeV (limited by availability), whereas Ref.~\cite{Bernhard:2018hnz} used a larger cross section of Pb-Pb experimental data at $\sqrts=2.76$ and 5.02~TeV.
+
+The posterior free-streaming time $\taufs=0.37_{-0.27}^{+0.34}$~fm/$c$ obtained in this work is significantly smaller than our previous estimate $\taufs=1.16_{-0.25}^{+0.29}$~fm/$c$ in Ref.~\cite{Bernhard:2018hnz}.
+We point out that the present study is missing several important observables which could affect the estimated free-streaming time, e.g.\ the Pb-Pb mean $p_T$ and mean $p_T$ fluctuations at $\sqrts=5.02$~TeV.
+Nevertheless, it appears that the inclusion of nucleon substructure significantly reduces the maximum allowed free-streaming time, although more work is needed to establish if this is indeed the case.
+
+We also compare in Fig.~\ref{fig:region_shear_bulk} our estimates for the temperature dependence of the QGP specific shear viscosity $(\eta/s)(T)$ and bulk viscosity $(\zeta/s)(T)$ with those of Ref.~\cite{Bernhard:2018hnz}.
+The lines are the distribution medians, and the bands are their 90\% credible regions.
+The results of this work are shown in orange, and the results of Ref.~\cite{Bernhard:2018hnz} are shown in blue.
+In general, our estimates are broader and less certain but otherwise self-consistent.
+Evidently, the combined analysis of Pb-Pb data at $\sqrts=2.76$ and 5.02~TeV in Ref.~\cite{Bernhard:2018hnz} provides a better constraint on the QGP viscosities which is not surprising given the additional observables and multiple beam energies studied.
+The $p$-Pb data, meanwhile, does not appear to provide any unique viscous constraints.
+
+\subsection{Verification of high-probability parameters}
+
+The full posterior in Fig.~\ref{fig:posterior} is the distribution of true parameter values, given the assumptions of our model and the evidence provided by the experimental data.
+Correspondingly, the probability of nature ``choosing'' any specific set of parameter values is \emph{zero}; we can only assign a probability to a finite region of parameter space.
+Nevertheless, we can define the \emph{maximum a posteriori probability} (MAP) estimate as the single most probable neighborhood (i.e.\ point) in parameter space, equal to the mode of the distribution.
+
+We list the posterior MAP parameters in Tbl.~\ref{tab:map_param}.
+The values are estimated using a m
+
+and run \order{6} events minimum-bias and multiplicity triggered $p$-Pb and Pb-Pb events at $\sqrts=5.02$~TeV.
+We then calculate all of the model observables listed in Sec.~\ref{sec:observables}.
+These ``best fit'' model predictions are plotted in Fig.~\ref{fig:obs_map} alongside the experimental data from CMS \cite{Chatrchyan:2013nka} and ALICE \cite{Adam:2015ptt, Adam:2016izf, Adam:2014qja, Abelev:2013bla}.
+They are explicit model calculations; \emph{there is no emulator}.
+
+\begin{table}[t]
   \caption{Maximum a posteriori (MAP) parameters.}
@@
   \end{ruledtabular}
-  \label{tab:design}
+  \label{tab:map_param}
 \end{table}
-
-\subsection{Verification of high-probability parameters}
-
-The posterior in Fig.~\ref{fig:posterior} is the distribution of true parameter values, given the assumptions of our model and the evidence provided by the experimental data.
-Correspondingly, the probability of nature ``choosing'' any specific set of parameter values is \emph{zero}; we can only assign a probability to a finite region of parameter space.
-Nevertheless, we can define the \emph{maximum a posteriori probability} (MAP) estimate as the single most probable neighborhood (i.e.\ point) in parameter space, equal to the mode of the distribution.
-
-We estimate the MAP parameters for the posterior in Fig.~\ref{fig:posterior} and run \order{6} events for $p$-Pb and Pb-Pb (minimum bias and multiplicity triggered)
 
@@
 Points with error bars are the experimental data with statistical uncertainties, and gray bands their corresponding systematic uncertainties, from CMS \cite{Chatrchyan:2013nka} and ALICE \cite{Adam:2015ptt, Adam:2016izf, Adam:2014qja, Abelev:2013bla}.
-The ratio axes show the ratio of model over data where where available with gray bands indicating $\pm 10\%$.}
+The sub-axes show the ratio of model over data where available with gray bands indicating $\pm 10\%$.}
 \end{fullpage}
@@
 
-\begin{figure}
-  \includegraphics[width=\columnwidth]{posterior_proton_shape}
-  \caption{Posterior distribution for the nucleon width $w$ and constituent width $v$.
-The white trapezoidal region (lower-right) covers the initial range of allowed values for $w$ and $v$, while the gray triangular region (upper-left) indicates values which are excluded by the prior $v > w$.
-The posterior distribution, shown in blue, indicates the preferred values for $w$ and $v$ determined by the analysis.}
-  \label{fig:posterior_proton_shape}
-\end{figure}
-
-\begin{figure*}
-  \includegraphics[width=\textwidth]{region_shear_bulk}
-  \caption{Left figure: estimated temperature dependence of the QGP specific shear viscosity $(\eta/s)(T)$ determined by the present Bayesian analysis of $p$-Pb and Pb-Pb collisions at $\sqrts=5.02$~TeV (orange line/band) compared to a previous Bayesian analysis of Pb-Pb collisions at $\sqrts=2.76$ and 5.02~TeV (blue line/band) \cite{Bernhard:2018hnz}.
-The lines are the medians of each posterior distribution, and the bands are their 90\% credible regions.
-Right figure: same as before, but for the temperature dependence of the QGP specific bulk viscosity $(\zeta/s)(T)$.}
-  \label{fig:region_shear_bulk}
-\end{figure*}
-
 \section{summary}
@@
 
+\bibliography{substructure}
 \appendix
@@
 \label{app:adaptive_grid}
+
+The boost-invariant \textsc{VISH2+1} hydrodynamics code used in this work \cite{Song:2007ux, Shen:2014vra} runs on a Cartesian transverse grid specified by a maximum grid size $x_\text{max}$ and grid step width $dx$ which fix the transverse grid extent ${-x_\text{max} < x < x_\text{max}}$ and number of grid cells along each dimension ${n_x = 2\, x_\text{max} / dx}$.
+In general, the maximum grid size $x_\text{max}$ should be set large enough to contain the full spacetime evolution of the event.
+This means that the truncation of $T^{\mu\nu}$ at the boundaries of the grid should never modify the final-state observables.
+We enforce this requirement by finding an energy density cutoff $e_\text{min}$ for which the matter $e < e_\text{min}$ can be effectively discarded without significantly modifying the simulation observables.
+We then fix the maximum grid size $x_\text{max}$ such that it fully encloses the isotherm $T=T(e_\text{min})$ for the full lifetime of the fireball.
+
+We find that we can quickly estimate the maximum radius $R_\text{max} = |\x_\text{max}|^2$ of the spacetime hypersurface ${T = T(e_\text{min})}$ by running the event on a coarse-grained spatial grid with one-third the spatial resolution we would otherwise require to resolve typical hydrodynamic observables such as mean $p_T$ and flows.
+The simulation time of a single \textsc{VISH2+1} event scales like ${\sim}n_x^3$ since $dx \propto d\tau$, and thus our ``pre-run'' event requires only ${\sim}1/27$th the time of a production event.
+We therefore run a coarse-grained pre-event on an excessively large grid for \emph{every} minimum-bias event to estimate $R_\text{max}$, then rerun the same event on a thrice finer grid with a trimmed spatial extent $x_\text{max} \equiv R_\text{max}$.
+See Fig.~\ref{fig:adaptive_grid} for a simple diagram of the procedure.
+
+In practice, we find that event-by-event grid resizing leads to a massive speed increase for minimum bias events compared to using a single fixed grid for the entire minimum bias sample.
+This is because the maximum transverse size of each event varies dramatically, from a few fm in peripheral Pb-Pb collisions to 50 fm or more in central Pb-Pb collisions.
+The procedure should generalize to other hydrodynamic codes, including those with $3+1$ spacetime dimensions, where the time savings could be even more pronounced.
 
@@
 
-The boost-invariant \textsc{VISH2+1} hydrodynamics code used in this work \cite{Song:2007ux, Shen:2014vra} runs on a Cartesian transverse grid specified by a maximum grid size $x_\text{max}$ and grid step width $dx$ which fix the transverse grid extent ${-x_\text{max} < x < x_\text{max}}$ and number of grid cells along each dimension ${n_x = 2\, x_\text{max} / dx}$.
-In general, the maximum grid size $x_\text{max}$ should be set large enough to contain the full spacetime evolution of the event.
-This means that the truncation of $T^{\mu\nu}$ at the boundaries of the grid should never modify the final-state observables.
-We enforce this requirement by finding an energy density cutoff $e_\text{min}$ for which the matter $e < e_\text{min}$ can be effectively discarded without significantly modifying the simulation observables.
-We then fix the maximum grid size $x_\text{max}$ such that it fully encloses the isotherm $T=T(e_\text{min})$ for the full lifetime of the fireball.
-
-We find that we can quickly estimate the maximum radius $R_\text{max} = |\x_\text{max}|^2$ of the spacetime hypersurface ${T = T(e_\text{min})}$ by running the event on a coarse-grained spatial grid with one-third the spatial resolution we would otherwise require to resolve typical hydrodynamic observables such as mean $p_T$ and flows.
-The simulation time of a single \textsc{VISH2+1} event scales like ${\sim}n_x^3$ since $dx \propto d\tau$, and thus our ``pre-run'' event requires only ${\sim}1/27$th the time of a production event.
-We therefore run a coarse-grained pre-event on an excessively large grid for \emph{every} minimum-bias event to estimate $R_\text{max}$, then rerun the same event on a thrice finer grid with a trimmed spatial extent $x_\text{max} \equiv R_\text{max}$.
-See Fig.~\ref{fig:adaptive_grid} for a simple diagram of the procedure.
-
-In practice, we find that event-by-event grid resizing leads to a massive speed increase for minimum bias events compared to using a single fixed grid for the entire minimum bias sample.
-This is because the maximum transverse size of each event varies dramatically, from a few fm in peripheral Pb-Pb collisions to 50 fm or more in central Pb-Pb collisions.
-The procedure should generalize to other hydrodynamic codes, including those with $3+1$ spacetime dimensions, where the time savings could be even more pronounced.
-
 \section{Emulator validation}
@@
 
+\begin{fullpage}
 \begin{figure*}
   \includegraphics{validation_PbPb5020}
-  \caption{}
-  \label{fig:validation_PbPb5020}
+  \includegraphics{validation_pPb5020}
+  \caption{\label{fig:val_pbpb} Emulator validation for the Pb-Pb collision system (top) and $p$-Pb collision system (bottom) at $\sqrts=5.02$~TeV.
+The ``piano keys'' in the top row of each figure are horizontally stacked box plots for the normalized residuals of each model observable.
+The boxes are 50\% interquartile ranges and whiskers are the 90\% interquantiles.}
 \end{figure*}
-
-\begin{figure*}
-  \includegraphics{validation_pPb5020}
-  \caption{}
-  \label{fig:validation_pPb5020}
-\end{figure*}
-
-\bibliography{substructure}
+\end{fullpage}
 

### 2018-07-25 paper/substructure.tex
More plucking away at MAP results and conclusion section

@@
   \end{cases}
-  \label{eq:gmean}
+  \label{eq:trento_p}
 \end{equation}
@@
 
-\subsection{Initial condition parameters}
+\subsection{Initial condition properties}
 
@@
 The parameter $p$ thus strongly affects the $p$-Pb and Pb-Pb yield ratio.
-It just so happens that the small range of $p$-values needed to describe this yield ratio are the same values needed to describe all of the calibration observables in the present study, \emph{and} numerous Pb-Pb observables at $\sqrts=2.76$~TeV \note.
-This work thus reaffirms an empirical scaling law reported in several previous studies \note, that entropy deposition at midrapidity in ultrarelativistic nuclear collisions scales approximately as
+It just so happens that the small range of $p$-values needed to describe this yield ratio are the same values needed to describe all of the calibration observables in the present study, \emph{and} numerous Pb-Pb observables at $\sqrts=2.76$~TeV \cite{Bernhard:2018hnz}.
+This work thus reaffirms an empirical scaling law reported in several previous studies \cite{Moreland:2014oya, Bernhard:2016tnd, Ke:2016jrd, Bernhard:2018hnz}, that entropy deposition at midrapidity in ultrarelativistic nuclear collisions scales approximately as
 \begin{equation}
@@
 
-Meanwhile, the posterior on the constituent number $n_c$ is not sharply peaked.
+Meanwhile, our posterior on the constituent number $n_c$ is not sharply peaked.
 We therefore refrain from quoting a distribution median and 90\% credible interval, although we do note that the distribution clearly favors $n_c > 1$ constituents.
 This is not surprising.
-The \trento\ model mimics saturation-based initial condition models, and saturation models tend to produce ``proton-sized'' fireballs in $p$-Pb collisions \note.
+The \trento\ model mimics saturation-based initial condition models \cite{Bernhard:2016tnd}, and saturation models tend to produce ``proton-sized'' fireballs in $p$-Pb collisions \cite{Bzdak:2013zma}.
 When the proton is spherically symmetric, the resulting proton-sized QGP is also largely symmetric and thus produces very little flow.
-Saturation-based models are therefore unable to describe the significant flow measured in high-multiplicity $p$-Pb collisions without nucleon substructure, or alternatively, some other source of additional correlations \note.
+Saturation-based models are therefore unable to describe the significant flow measured in high-multiplicity $p$-Pb collisions without nucleon substructure, or alternatively, some other source of additional correlations \cite{Schenke:2017bog}.
 
@@
 
-The full posterior in Fig.~\ref{fig:posterior} is the distribution of true parameter values, given the assumptions of our model and the evidence provided by the experimental data.
-Correspondingly, the probability of nature ``choosing'' any specific set of parameter values is \emph{zero}; we can only assign a probability to a finite region of parameter space.
-Nevertheless, we can define the \emph{maximum a posteriori probability} (MAP) estimate as the single most probable neighborhood (i.e.\ point) in parameter space, equal to the mode of the distribution.
-
-We list the posterior MAP parameters in Tbl.~\ref{tab:map_param}.
-The values are estimated using a m
-
-and run \order{6} events minimum-bias and multiplicity triggered $p$-Pb and Pb-Pb events at $\sqrts=5.02$~TeV.
+In this section we select a single set of high-probability model parameters
+
+We verify the emulator predictions and model accuracy by running a large number of $p$-Pb and Pb-Pb events using a single set of high-probability parameters selected from the mode of the posterior distribution.
+These parameters approximate the ``best fit'' values
 We then calculate all of the model observables listed in Sec.~\ref{sec:observables}.
-These ``best fit'' model predictions are plotted in Fig.~\ref{fig:obs_map} alongside the experimental data from CMS \cite{Chatrchyan:2013nka} and ALICE \cite{Adam:2015ptt, Adam:2016izf, Adam:2014qja, Abelev:2013bla}.
-They are explicit model calculations; \emph{there is no emulator}.
+%These ``best fit'' model predictions are plotted in Fig.~\ref{fig:obs_map} alongside the experimental data from CMS \cite{Chatrchyan:2013nka} and ALICE \cite{Adam:2015ptt, Adam:2016izf, Adam:2014qja, Abelev:2013bla}.
+%They are explicit model calculations; \emph{there is no emulator}.
+%
 
 \begin{table}[t]
-  \caption{Maximum a posteriori (MAP) parameters.}
+  \caption{
+    \label{tab:mode_params}
+    High-probability parameters selected from the posterior distribution and used to generate Fig.~\ref{fig:obs_map}.
+    The posterior distribution on the particlization temperature $T_\text{switch}$ is flat (agnostic), so we fix it's value using Ref.~\cite{Bernhard:2018hnz}.
+  }
   \begin{ruledtabular}
-  \begin{tabular}{lll}
-    Parameter         & Description                        & Value\\
+    \begin{tabular}{ll@{\hspace{2em}}ll}
+      \multicolumn{2}{c}{Initial condition} & \multicolumn{2}{c}{QGP medium} \\
       \paddedhline
-    Norm              & Normalization factor                 & 20.            \\
-    $p$               & Entropy deposition parameter         & 0.0            \\
-    $k$               & Constituent gamma fluct.\ shape      & 0.19           \\
-    $w$               & Nucleon width parameter              & 0.92 fm        \\
-    $n_c$             & Number of nucleon constituents       & 6              \\
-    $v$               & Constituent width                    & 0.43 fm        \\
-    $\dmin$           & Minimum inter-nucleon distance       & 0.81 fm        \\
-    $\taufs$          & Free-streaming time                  & 0.37 fm/$c$    \\
-    $\eta/s$ min      & Minimum value of $\eta/s$ (at $T_c$) & 0.11           \\
-    $\eta/s$ slope    & Slope of $\eta/s$ above $T_c$        & 1.6 GeV$^{-1}$ \\
-    $\eta/s$ crv      & Curvature of $\eta/s$ above $T_c$    & -0.29          \\
-    $\zeta/s$ max     & Maximum value of $\zeta/s$           & 0.032          \\
-    $\zeta/s$ width   & Width of $\zeta/s$ peak              & 0.024 GeV      \\
-    $\zeta/s$ $T_0$   & Temperature of $\zeta/s$ maximum     & 175 MeV        \\
-    $T_\text{switch}$ & Switching/particlization temp.       & 151 MeV        \\
+      Norm     & 20.            & $\eta/s$ min      & 0.11           \\
+      $p$      & 0.0            & $\eta/s$ slope    & 1.6 GeV$^{-1}$ \\
+      $k$      & 0.19           & $\eta/s$ curv     & -0.29          \\
+      $n_c$    & 6              & $\zeta/s$ max     & 0.032          \\
+      $w$      & 0.92 fm        & $\zeta/s$ width   & 0.024 GeV      \\
+      $v$      & 0.43 fm        & $\zeta/s$ $T_0$   & 175 MeV        \\
+      $\dmin$  & 0.81 fm        & $T_\text{switch}$ & 151 MeV        \\
+      $\taufs$ & 0.37 fm/$c$
     \end{tabular}
   \end{ruledtabular}
-  \label{tab:map_param}
 \end{table}
@@
 \label{sec:summary}
+
+In this work, we constructed a multi-stage hybrid transport model for relativistic nuclear collisions and calibrated free parameters of the model using Bayesian parameter estimation.
+Specifically, we focused on improving the model's description of small collision systems by investigating the effect of nucleon substructure on bulk observables in $p$-Pb and Pb-Pb collisions at $\sqrts=5.02$~TeV.
+
+The analysis was performed using a parametric data-driven approach to account for theoretical uncertainty (variability) in the QGP initial conditions and hydrodynamic transport coefficients using a handful of equally uncertain model parametersi.
+We marginalized over the theoretical uncertainty in the parameter using a handful of equally uncertain model parameters to vary the size and shape of the nucleon, the scaling behavior of initial QGP entropy deposition, the duration of pre-equilibrium dynamics, and the temperature dependence of the hydrodynamic shear and bulk viscosities.
+
+We placed a conservative prior on the ``true'' value of each parameter, and calculated a global likelihood
+
+varied each parameter within its design range to explored the full 15-dimensional parameter space of our model.
+
+constructed an emulator to interpolate the predictions of our model at arbitrary regions of parameter space.
+
+Bayesian inference was then used to compare the predictions of the model to experimental data from CMS and ALICE, in order to quantify the posterior probability that a given set of model parameters describes the experimental data
 

### 2018-07-26 paper/substructure.tex
Add more content to MAP parameters section

@@
 
-In this section we select a single set of high-probability model parameters
-
-We verify the emulator predictions and model accuracy by running a large number of $p$-Pb and Pb-Pb events using a single set of high-probability parameters selected from the mode of the posterior distribution.
-These parameters approximate the ``best fit'' values
-We then calculate all of the model observables listed in Sec.~\ref{sec:observables}.
-%These ``best fit'' model predictions are plotted in Fig.~\ref{fig:obs_map} alongside the experimental data from CMS \cite{Chatrchyan:2013nka} and ALICE \cite{Adam:2015ptt, Adam:2016izf, Adam:2014qja, Abelev:2013bla}.
-%They are explicit model calculations; \emph{there is no emulator}.
-%
+We verified the emulator and tested the accuracy of our physics model framework using a single set of high-probability parameters selected from the Bayesian posterior.
+These parameters, listed in Tbl.~\ref{tab:mode_params}, are the approximate ``best fit'' values of the calibrated model, and are commonly referred to as the distribution's \emph{maximum a posteriori} (MAP) estimate.
+
+We ran \order{6} minimum-bias and multiplicity triggered events using the MAP estimate and computed all of the model observables listed in Sec.~\ref{sec:observables}.
+The MAP model calculations are plotted in Fig.~\ref{fig:obs_map} and compared to experimental data from CMS \cite{Chatrchyan:2013nka} and ALICE \cite{Adam:2015ptt, Adam:2016izf, Adam:2014qja, Abelev:2013bla} where available.
+The left and right columns show the results for the $p$-Pb and Pb-Pb collision systems respectively, and each row shows a different observable type.
+
+The global agreement of the MAP model calculations with the experimental data is very good.
+The largest discrepancy seems to be the gap between the $p$-Pb two-particle cumulants $\vnk{2}{2}$ and $\vnk{3}{2}$, although even that tension is only about 10--15\%.
+Quite remarkably, the model perfectly describes the shape of the $p$-Pb and Pb-Pb two-particle correlations which is strong evidence that these correlations are hydrodynamic in origin.
+
+We also point to the simultaneous description of the $p$-Pb and Pb-Pb charged-particle yields using a single initial condition normalization and a single entropy deposition factor $p=0$.
+This is the \emph{exact same} generalized mean $p$-value supported by multiple previous studies \note.
+Evidently, the geometric mean scaling in Eqn.~\ref{eq:geometric_mean} appears to be general property of relativistic nuclear collisions.
+
+Similarly, we obtain an excellent description of the \mbox{$p$-Pb} mean $p_T$, although this fit is somewhat less meaningful since we are unable to calibrate on the Pb-Pb mean $p_T$ simultaneously.
+It would also be nice to compare to the mean $p_T$ fluctuations at $\sqrts=5.02$~TeV which have been predicted in this study.
+It was suggested that nucleon substructure may
 

### 2018-07-26 paper/substructure.tex
Just need to write about symmetric cumulants and write the summary

--- 2018-07-25 paper/substructure.tex
@@
 \begin{figure}
-  \includegraphics[width=\columnwidth]{posterior_proton_shape}
-  \caption{Posterior distribution for the nucleon width $w$ and constituent width $v$.
+  \includegraphics{posterior_proton_shape}
+  \caption{\label{fig:posterior_proton_shape}
+  Posterior distribution for the nucleon width $w$ and constituent width $v$.
 The white trapezoidal region (lower-right) covers the initial range of allowed values for $w$ and $v$, while the gray triangular region (upper-left) indicates values which are excluded by the prior $v > w$.
 The posterior distribution, shown in blue, indicates the preferred values for $w$ and $v$ determined by the analysis.}
-  \label{fig:posterior_proton_shape}
+\end{figure}
+
+\begin{figure}[b]
+  \includegraphics{posterior_parton_number}
+  \caption{\label{fig:posterior_parton_number}
+    Posterior distribution for the number of nucleon constituents $n_c$ determined by the analysis.
+    The parameter $n_c$ is an integer (discrete) variable at every design point, but the emulator interpolation produces a posterior distribution which is continuous.
+  }
 \end{figure}
@@
 \begin{figure*}
-  \includegraphics[width=\textwidth]{region_shear_bulk}
+  \includegraphics{region_shear_bulk}
   \caption{Left figure: estimated temperature dependence of the QGP specific shear viscosity $(\eta/s)(T)$ determined by the present Bayesian analysis of $p$-Pb and Pb-Pb collisions at $\sqrts=5.02$~TeV (orange line/band) compared to a previous Bayesian analysis of Pb-Pb collisions at $\sqrts=2.76$ and 5.02~TeV (blue line/band) \cite{Bernhard:2018hnz}.
@@
 \subsection{Verification of high-probability parameters}
-
-In this section we select a single set of high-probability model parameters
-
-We verify the emulator predictions and model accuracy by running a large number of $p$-Pb and Pb-Pb events using a single set of high-probability parameters selected from the mode of the posterior distribution.
-These parameters approximate the ``best fit'' values
-We then calculate all of the model observables listed in Sec.~\ref{sec:observables}.
-%These ``best fit'' model predictions are plotted in Fig.~\ref{fig:obs_map} alongside the experimental data from CMS \cite{Chatrchyan:2013nka} and ALICE \cite{Adam:2015ptt, Adam:2016izf, Adam:2014qja, Abelev:2013bla}.
-%They are explicit model calculations; \emph{there is no emulator}.
-%
 
@@
 
+We verified the emulator and tested the accuracy of our physics model framework using a single set of high-probability parameters selected from the Bayesian posterior.
+These parameters, listed in Tbl.~\ref{tab:mode_params}, are the approximate ``best fit'' values of the calibrated model, commonly referred to as the \emph{maximum a posteriori} (MAP) estimate:
+\begin{equation}
+  \x_\mathrm{MAP} \equiv \operatorname*{arg\, max}_{\x} P(H_\x | E).
+\end{equation}
+We then ran \order{6} minimum-bias and multiplicity triggered events using the MAP estimate and computed all of the model observables listed in Sec.~\ref{sec:observables}.
+The resulting model calculations are shown in Fig.~\ref{fig:obs_map} alongside experimental data from CMS \cite{Chatrchyan:2013nka} and ALICE \cite{Adam:2015ptt, Adam:2016izf, Adam:2014qja, Abelev:2013bla}.
+The left and right columns show the results for the $p$-Pb and Pb-Pb collision systems respectively, and each row shows a different group of related observables.
+
+The global agreement of the MAP model calculations with the experimental data is very good.
+The largest tension is observed in the two-particle cumulants $\vnk{2}{2}$ and $\vnk{3}{2}$ of the $p$-Pb system, although even that tension is only about 10--15\%.
+Quite remarkably, the model perfectly describes the shape of the $p$-Pb and Pb-Pb two-particle correlations which is strong evidence that these correlations are hydrodynamic in origin.
+Moreover, we obtain an excellent description of the \mbox{$p$-Pb} mean $p_T$, although this fit is somewhat less meaningful since we are unable to calibrate on the Pb-Pb mean $p_T$ simultaneously (data is not yet available).
+Additionally, note that the model provides a simultaneous description of the \mbox{$p$-Pb} and Pb-Pb charged-particle yields using a single entropy deposition parameter $p=0$.
+This is the \emph{exact same} generalized mean $p$-value supported by multiple previous studies \cite{Moreland:2014oya, Bernhard:2016tnd, Ke:2016jrd, Bernhard:2018hnz}.
+Evidently, this scaling continues to hold for initial conditions with sizable nucleon substructure.
+
+We also present calculations for several observables which were omitted from the calibration due to the statistical limitations of our training data.
+Here our MAP event sample is several orders-of-magnitude larger so the statistics are no issue.
+The bottom-right panel of Fig.~\ref{fig:obs_map} shows our model calculation for the four-particle elliptic flow cumultant $\vnk{2}{4}$ along with the measured data points from \mbox{ALICE} \cite{Adam:2016izf}.
+We see that the MAP estimate nicely describes the measured $\vnk{2}{4}$ data which is encouraging since this particular observable was never used to calibrate the posterior.
+
+The relative mean $p_T$ fluctuation $\delta p_T / \langle p_T \rangle$ is another important bulk observable to test the predictions of the calibrated model.
+It measures the \emph{dynamical} component of event-by-event mean $p_T$ fluctuations, quantified by the two-particle correlator
+\begin{equation}
+  (\delta p_T)^2 = \langle \langle (p_{T,i} - \langle p_T \rangle) (p_{T,j} - \langle p_T \rangle) \rangle \rangle.
+  \label{eq:mean_pT_corr}
+\end{equation}
+The inner-average in Eqn.~\ref{eq:mean_pT_corr} runs over all pairs of particles $i,j$ in the same event, the outer average runs over all events in a given bin (centrality or multiplicity), and the symbol $\langle p_T \rangle$ denotes the usual mean transverse momentum of particles in the bin.
+The observable is typically presented in terms of the dimensionless ratio $\delta p_T / \langle p_T \rangle$, which quantifies the strength of the dynamical fluctuations in units of the average transverse momentum $\langle p_T \rangle$.
+
+We show the MAP estimate predictions for the $p$-Pb and Pb-Pb relative mean $p_T$ fluctuations $\delta p_T / \langle p_T \rangle$ at $\sqrts=5.02$~TeV in the third row of Fig.~\ref{fig:obs_map}.
+For the Pb-Pb system, we use centrality bins and for the $p$-Pb system we use the same relative multiplicity bins used for two-particle flows.
+The relative mean $p_T$ fluctuations were shown to be particularly sensitive to the existence of nucleon substructure \cite{Bozek:2017elk}, and thus it would be interesting to ultimately include the observable in the calibration when the data becomes available.
+
 \begin{fullpage}
@@
 
+Finally, we compute model predictions for the symmetric cumulants $\mathrm{SC}(m,n)$, which quantify correlations between event-by-event fluctuations of the flow harmonics $v_m$ and $v_n$ \cite{Bilandzic:2013kga, ALICE:2016kpq},
+\begin{align}
+  \mathrm{SC}(m, n) &= \langle\langle \cos[m(\phi_1 - \phi_3) + n(\phi_2-\phi_4)]\rangle\rangle \nonumber \\
+\nonumber &- \langle\langle\cos[m(\phi_1-\phi_2)]\rangle\rangle\langle\langle\cos[n(\phi_1-\phi_2)]\rangle\rangle \label{eq:scmn}\\
+&\approx \langle v_m^2 v_n^2 \rangle - \langle v_m^2\rangle\langle v_n^2\rangle.
+\end{align}
+We show these model prediction in Fig.~\ref{fig:flow_corr} along with the \emph{normalized} symmetric cumulants
+
+\begin{equation}
+  \mathrm{NSC}(m,n) = SC(m,n)/\langle v_m^2\rangle\langle v_n^2\rangle,
+\end{equation}
+which are expected to be less sensitive to the medium response and more sensitive to the properties of the initial state.
+
 \begin{figure}
@@
 
+\medskip
+
+All software used in this work is open source:
+\begin{itemize}[leftmargin=2\parindent, itemsep=0pt, topsep=5pt]
+  \item \trento\ initial conditions \note
+  \item Pre-equilibrium free-streaming \note
+  \item VISH2+1 hydrodynamics \note
+  \item FRZOUT particle sample \note
+  \item UrQMD microscopic transport model \note
+  \item Collision event generator (wrapper) \note
+\end{itemize}
+
 \begin{acknowledgments}
-  The authors thank ...
+  JSM thanks Weiyao Ke, Bj\"orn Schenke and Heikki M\"antysaari for helpful discussions and clarifying comments.
+  This research was completed using 6 million CPU hours provided by the National Energy Research Scientific Computing Center (NERSC), a U.S.
+Department of Energy Office of Science User Facility operated under Contract No.
+DE-AC02-05CH11231.
+  JSM, JEB and SAB acknowledge support by ...
 \end{acknowledgments}

### 2018-07-26 src/emulator.py

@@ -103,6 +103,7 @@ class Emulator:
         for obs, subobslist in self.observables:
             self._slices[obs] = {}
             for subobs in subobslist:
+                print(system, obs, subobs)
                 Y.append(model.data[system][obs][subobs]['Y'])
                 n = Y[-1].shape[1]
                 self._slices[obs][subobs] = slice(nobs, nobs + n)

### 2018-07-26 src/model.py

@@ -346,7 +346,7 @@ def _data(system, dataset='main'):
         empty = obs_like(data['mean_pT'][None])
     else:
         empty = obs_like(data['vnk'][(2, 2)])
-        data = dict({'mean_pT': {s: empty for s in species}}, **data)
+        data = dict({'mean_pT': {None: empty}}, **data)
 
     data = dict({'iden_mean_pT': {s: empty for s in species}}, **data)
     data = dict({'pT_fluct': {None: empty}}, **data)

### 2018-07-26 src/plots.py

@@ -154,13 +154,15 @@ def plot(f):
     return wrapper
 
 
-def figsize(relwidth=1, aspect=.618, refwidth=6):
+
+def figsize(relwidth=1, aspect=.618, refwidth=6.5):
     """
     Return figure dimensions from a relative width (to a reference width) and
     aspect ratio (default: 1/golden ratio).
 
     """
     width = relwidth * refwidth
+
     return width, width*aspect
 
 
@@ -442,7 +444,7 @@ def observables(system):
 
     fig, axes = plt.subplots(
         nrows=2, ncols=len(plots),
-        figsize=figsize(1.1, aspect=.6),
+        figsize=figsize(1, aspect=.6),
     )
 
     for (posterior, plot), ax in zip(
@@ -565,7 +567,7 @@ def observables_map():
                 ],
                 legend=True
             )
-    fig = plt.figure(figsize=figsize(0.955, 1.5))
+    fig = plt.figure(figsize=figsize(0.88, 1.5))
 
     yields, mean_pT, mean_pT_fluct, flows = [
         gridspec.GridSpecFromSubplotSpec(
@@ -680,7 +682,7 @@ def observables_map():
     set_tight(fig, h_pad=.5, rect=[0, 0, .97, 1])
 
 
-@plot
+#@plot
 def find_map():
     """
     Find the maximum a posteriori (MAP) point and compare emulator predictions
@@ -1047,7 +1049,7 @@ def posterior_p():
     Distribution of trento p parameter with annotations for other models.
 
     """
-    plt.figure(figsize=figsize(.6, .5))
+    plt.figure(figsize=figsize(.5, .5))
     ax = plt.axes()
 
     data = mcmc.Chain().load('trento_p').ravel()
@@ -1222,7 +1224,7 @@ def region_shear():
     Region plot for eta/s.
     """
     chain = mcmc.Chain()
-    fig, ax = plt.subplots(figsize=figsize(.6, .65))
+    fig, ax = plt.subplots(figsize=figsize(.5, .65))
     _region(ax, 'shear', chain, legend='upper left')
     set_tight(fig)
 
@@ -1233,7 +1235,7 @@ def region_bulk():
     Region plot for zeta/s.
     """
     chain = mcmc.Chain()
-    fig, ax = plt.subplots(figsize=figsize(.6, .65))
[…]

### 2018-07-27 paper/substructure.tex
Little bit of work from home

@@
 \usepackage{tikz}
+\usepackage[leftcaption]{sidecap}
 
@@
 \end{equation}
-We then ran \order{6} minimum-bias and multiplicity triggered events using the MAP estimate and computed all of the model observables listed in Sec.~\ref{sec:observables}.
+We then ran \order{6} minimum-bias and multiplicity triggered events using the MAP estimate $\x_\mathrm{MAP}$ and computed all of the model observables listed in Sec.~\ref{sec:observables}.
 The resulting model calculations are shown in Fig.~\ref{fig:obs_map} alongside experimental data from CMS \cite{Chatrchyan:2013nka} and ALICE \cite{Adam:2015ptt, Adam:2016izf, Adam:2014qja, Abelev:2013bla}.
@@
 Moreover, we obtain an excellent description of the \mbox{$p$-Pb} mean $p_T$, although this fit is somewhat less meaningful since we are unable to calibrate on the Pb-Pb mean $p_T$ simultaneously (data is not yet available).
-Additionally, note that the model provides a simultaneous description of the \mbox{$p$-Pb} and Pb-Pb charged-particle yields using a single entropy deposition parameter $p=0$.
+Additionally, the model provides a simultaneous description of the \mbox{$p$-Pb} and Pb-Pb charged-particle yields using a single entropy deposition parameter $p=0$.
 This is the \emph{exact same} generalized mean $p$-value supported by multiple previous studies \cite{Moreland:2014oya, Bernhard:2016tnd, Ke:2016jrd, Bernhard:2018hnz}.
@@
 For the Pb-Pb system, we use centrality bins and for the $p$-Pb system we use the same relative multiplicity bins used for two-particle flows.
-The relative mean $p_T$ fluctuations were shown to be particularly sensitive to the existence of nucleon substructure \cite{Bozek:2017elk}, and thus it would be interesting to ultimately include the observable in the calibration when the data becomes available.
+The relative mean $p_T$ fluctuations were shown to be particularly sensitive to the existence of nucleon substructure \cite{Bozek:2017elk}, and thus it would be interesting to ultimately include this observable in the calibration when the data becomes available.
 
@@
 
-Finally, we compute model predictions for the symmetric cumulants $\mathrm{SC}(m,n)$, which quantify correlations between event-by-event fluctuations of the flow harmonics $v_m$ and $v_n$ \cite{Bilandzic:2013kga, ALICE:2016kpq},
+As a final check, we compute the symmetric cumulants $\mathrm{SC}(m,n)$ for the Pb-Pb collision system at $\sqrts=5.02$~TeV.
+The symmetric cumulants quantify correlations between event-by-event fluctuations of the flow harmonics of order $v_m$ and $v_n$ \cite{Bilandzic:2013kga, ALICE:2016kpq},
 \begin{align}
@@
 We show these model prediction in Fig.~\ref{fig:flow_corr} along with the \emph{normalized} symmetric cumulants
-
 \begin{equation}
@@
 which are expected to be less sensitive to the medium response and more sensitive to the properties of the initial state.
+The solid lines are model calculations using the MAP parameters from the present analysis and the dashed lines are model calculations using the MAP parameters of Ref.~\note\ which did not include nucleon substructure.
+The gap between $\mathrm{SC}(3,2)$ and $\mathrm{SC}(4,2)$ is generally wider in the present analysis, as is the gap between the normalized symmetric cumulants $\mathrm{NSC}(3,2)$ and $\mathrm{NSC}(4,2)$.
+This may alleviate a common problem
 
@@
   \includegraphics{flow_corr}
-  \caption{Model calculations of the symmetric cumulants for Pb-Pb collisions at $\sqrts=5.02$~TeV using the \emph{maximum a posteriori} (MAP) parameters.}
+  \caption{Model calculations of the symmetric cumulants (top) and normalized symmetric cumulants (bottom) for Pb-Pb collisions at $\sqrts=5.02$~TeV using the \emph{maximum a posteriori} (MAP) parameters.
+The solid lines are the MAP estimate of the present analysis (with nucleon substructure), and the dashed lines are the MAP estimate of Ref.~\note\ (without nucleon substructure).
+In general, most model parameters are somewhat different between the two studies.}
   \label{fig:flow_corr}
@@
   \item UrQMD microscopic transport model \note
-  \item Collision event generator (wrapper) \note
+  \item DukeQCD event generator (wrapper) \note
 \end{itemize}
@@
 \label{app:adaptive_grid}
-
-The boost-invariant \textsc{VISH2+1} hydrodynamics code used in this work \cite{Song:2007ux, Shen:2014vra} runs on a Cartesian transverse grid specified by a maximum grid size $x_\text{max}$ and grid step width $dx$ which fix the transverse grid extent ${-x_\text{max} < x < x_\text{max}}$ and number of grid cells along each dimension ${n_x = 2\, x_\text{max} / dx}$.
-In general, the maximum grid size $x_\text{max}$ should be set large enough to contain the full spacetime evolution of the event.
-This means that the truncation of $T^{\mu\nu}$ at the boundaries of the grid should never modify the final-state observables.
-We enforce this requirement by finding an energy density cutoff $e_\text{min}$ for which the matter $e < e_\text{min}$ can be effectively discarded without significantly modifying the simulation observables.
-We then fix the maximum grid size $x_\text{max}$ such that it fully encloses the isotherm $T=T(e_\text{min})$ for the full lifetime of the fireball.
-
-We find that we can quickly estimate the maximum radius $R_\text{max} = |\x_\text{max}|^2$ of the spacetime hypersurface ${T = T(e_\text{min})}$ by running the event on a coarse-grained spatial grid with one-third the spatial resolution we would otherwise require to resolve typical hydrodynamic observables such as mean $p_T$ and flows.
-The simulation time of a single \textsc{VISH2+1} event scales like ${\sim}n_x^3$ since $dx \propto d\tau$, and thus our ``pre-run'' event requires only ${\sim}1/27$th the time of a production event.
-We therefore run a coarse-grained pre-event on an excessively large grid for \emph{every} minimum-bias event to estimate $R_\text{max}$, then rerun the same event on a thrice finer grid with a trimmed spatial extent $x_\text{max} \equiv R_\text{max}$.
-See Fig.~\ref{fig:adaptive_grid} for a simple diagram of the procedure.
-
-In practice, we find that event-by-event grid resizing leads to a massive speed increase for minimum bias events compared to using a single fixed grid for the entire minimum bias sample.
-This is because the maximum transverse size of each event varies dramatically, from a few fm in peripheral Pb-Pb collisions to 50 fm or more in central Pb-Pb collisions.
-The procedure should generalize to other hydrodynamic codes, including those with $3+1$ spacetime dimensions, where the time savings could be even more pronounced.
 
@@
 
-\section{Emulator validation}
-\label{app:validation}
-
 \begin{figure*}
-  \includegraphics{validation_example}
+  \includegraphics[width=.6\textwidth]{validation_example}
   \caption{Example emulator validation for one observable, the Pb-Pb charged particle yield $d\nch/d\eta$ in the 20-30\% centrality class.
@@
 \end{figure*}
+
+The boost-invariant \textsc{VISH2+1} hydrodynamics code used in this work \cite{Song:2007ux, Shen:2014vra} runs on a Cartesian transverse grid specified by a maximum grid size $x_\text{max}$ and grid step width $dx$ which fix the transverse grid extent ${-x_\text{max} < x < x_\text{max}}$ and number of grid cells along each dimension ${n_x = 2\, x_\text{max} / dx}$.
+In general, the maximum grid size $x_\text{max}$ should be set large enough to contain the full spacetime evolution of the event.
+This means that the truncation of $T^{\mu\nu}$ at the boundaries of the grid should never modify the final-state observables.
+We enforce this requirement by finding an energy density cutoff $e_\text{min}$ for which the matter $e < e_\text{min}$ can be effectively discarded without significantly modifying the simulation observables.
+We then fix the maximum grid size $x_\text{max}$ such that it fully encloses the isotherm $T=T(e_\text{min})$ for the full lifetime of the fireball.
+
+We find that we can quickly estimate the maximum radius $R_\text{max} = |\x_\text{max}|^2$ of the spacetime hypersurface ${T = T(e_\text{min})}$ by running the event on a coarse-grained spatial grid with one-third the spatial resolution we would otherwise require to resolve typical hydrodynamic observables such as mean $p_T$ and flows.
+The simulation time of a single \textsc{VISH2+1} event scales like ${\sim}n_x^3$ since $dx \propto d\tau$, and thus our ``pre-run'' event requires only ${\sim}1/27$th the time of a production event.
+We therefore run a coarse-grained pre-event on an excessively large grid for \emph{every} minimum-bias event to estimate $R_\text{max}$, then rerun the same event on a thrice finer grid with a trimmed spatial extent $x_\text{max} \equiv R_\text{max}$.
+See Fig.~\ref{fig:adaptive_grid} for a simple diagram of the procedure.
+
+In practice, we find that event-by-event grid resizing leads to a massive speed increase for minimum bias events compared to using a single fixed grid for the entire minimum bias sample.
+This is because the maximum transverse size of each event varies dramatically, from a few fm in peripheral Pb-Pb collisions to 50 fm or more in central Pb-Pb collisions.
+The procedure should generalize to other hydrodynamic codes, including those with $3+1$ spacetime dimensions, where the time savings could be even more pronounced.
+
+\section{Emulator validation}
+\label{app:validation}
 

### 2018-07-27 paper/substructure.tex
Last commit before leaving for Delware

@@
   \end{equation}
-  which approximates the RMS radius of the nucleon when the number of sampled constituents is large (see Fig.~\ref{fig:nucleon_schematic} for an example).
+  which approximates the RMS radius of the nucleon when the number of sampled constituents is large (see Fig.~\ref{fig:nucleon_schematic} for an example proton).
   We thus call $w$ a nucleon ``width'', although for smaller numbers of constituents, the \emph{actual} RMS radius of our sampled nucleons can be significantly smaller than our width parameter $w$ due to fluctuations in the nucleon's center of mass, and hence one should account for the difference when discussing the nucleon's posterior RMS radius.
@@
     \foreach \theta in {60, 180, 300}{
-      \draw ({\theta}:.8) circle (.65cm);
+      \draw ({\theta}:.85) circle (.73cm);
     }
@@
     \coordinate (b) at (.8, 1.385);
-    \coordinate (c) at (.4, -.693);
-    \coordinate (d) at (-1.45, 0);
-    \coordinate (e) at (-.8, 0);
+    \coordinate (c) at (.425, -.736);
+    \coordinate (d) at (-1.58, 0);
+    \coordinate (e) at (-.85, 0);
     \draw[<->] (a) -- (b) node[midway, xshift=-1ex, yshift=.8ex]{$w$};
@@
   \item Mean transverse momentum fluctuations $\delta p_T / \langle p_T \rangle$ (charged particles, $0.15 < p_T < 2.0$~GeV).
-  \item Two-particle flow cumulants $\vnk{n}{2}$ for $n=2,3,4$\\ (charged particles, $0.2 < p_T < 5.0$~GeV, and \\$0.3 < p_T < 3.0$~GeV).
-  \item Four-particle flow cumulant $\vnk{2}{4}$ \\(charged particles, $0.2 < p_T < 5.0$).
+  \item Two-particle flow cumulants $\vnk{n}{2}$ for $n=2,3,4$\\ (charged particles, $0.2 < p_T < 5.0$~GeV for ALICE, and $0.3 < p_T < 3.0$~GeV for CMS).
+  \item Four-particle flow cumulant $\vnk{2}{4}$ \\(charged particles, $0.2 < p_T < 5.0$~GeV).
   \item Symmetric cumulants $\mathrm{SC}(4, 2)$ and $\mathrm{SC}(3,2)$.
@@
 This restriction is commonly circumvented using principal component analysis (PCA): a general procedure which transforms a set of correlated variables $\y = (y_1, y_2, \dots, y_m)$ into a new basis representation $\z = (z_1, z_2, \dots, z_m)$ where the linear correlations between $z_i$, $z_j$ vanish for all $i\ne j \in m$ \cite{Tipping:1999}.
+Independent GPE's can then be used to emulate each $z \in \z$ since the variables $(z_1, \dots, z_m)$ are linearly uncorrelated.
+The emulated vector $\z$ is then easily reexpressed in the basis of $\y$ through its inverse transformation.
 
@@
 The principal components are reported in order of explained variance, with the first principal component vector $\mathbf{v}_1$ accounting for the most variance in $\tilde{Y}$, and the last principal component vector $\mathbf{v}_m$ accounting for the least.
-We then train a set of independent GPE's $z_i = \mathrm{gp}_i(\x)$ for $i=1$ to $k$ to predict the principal components $(z_1, \dots, z_k)$ of the first $k$ principal component vectors as a function of the model parameters $\x$ which vary across the design $X$.
+We then train a set of independent GPE's $\{z_i=\mathrm{gp}_i(\x)\}$ to predict the first $k$ principal components $(z_1, \dots, z_k)$ as a function of the model parameters $\x$ which vary across the design $X$.
 For the present study, we use $k=7$ principal components when emulating the $p$-Pb system and $k=8$ principal components when emulating Pb-Pb, chosen to describe 99.5\% of the total observed variance of each system.
@@
 
-We show simulated and emulated model observables (thin colored lines) for Pb-Pb collisions (Fig.~\ref{fig:obs_pbpb}) and \mbox{$p$-Pb} collisions (Fig.~\ref{fig:obs_ppb}) at $\sqrts=5.02$~TeV compared to experimental data from CMS \cite{Chatrchyan:2013nka} and ALICE \cite{Adam:2015ptt, Adam:2016izf, Adam:2014qja, Abelev:2013bla}.
+We show the simulated and emulated model observables (thin colored lines) for Pb-Pb collisions in Fig.~\ref{fig:obs_pbpb} and for \mbox{$p$-Pb} collisions in Fig.~\ref{fig:obs_ppb} at $\sqrts=5.02$~TeV compared to experimental data from CMS \cite{Chatrchyan:2013nka} and ALICE \cite{Adam:2015ptt, Adam:2016izf, Adam:2014qja, Abelev:2013bla}.
 The top row of each figure shows explicit model calculations at each of the $d=500$ design points (training data), while the bottom row shows emulator predictions for $n=100$ random parameter samples drawn from the Bayesian posterior (sampled from the MCMC chain).
@@
 
-Conversely, the calibrated emulator predictions (bottom row of Figs.~\ref{fig:obs_pbpb} and \ref{fig:obs_ppb}) are far better constrained and nicely track the experimental data points.
+Conversely, the calibrated (posterior sampled) emulator predictions (bottom row of Figs.~\ref{fig:obs_pbpb} and \ref{fig:obs_ppb}) are far better constrained and nicely track the experimental data points.
 We emphasize here that the posterior parameter values are obtained from a \emph{simultaneous} calibration to $p$-Pb and Pb-Pb data, and thus they are self-consistent between the two systems.
@@
 We report posterior estimates for individual parameters using their (marginalized) distribution medians and 90\% credible intervals.
-For example, $x=2.45_{-0.15}^{+0.20}$ corresponds to a posterior median $\tilde{x}=2.45$ and 90\% credible interval $x \in (2.30, 2.65)$.
+For example, the estimate $x=2.45_{-0.15}^{+0.20}$ denotes a median value $\tilde{x}=2.45$ and 90\% credible interval $2.30 < x < 2.65$.
 
@@
 
-\begin{figure}
-  \includegraphics{posterior_proton_shape}
-  \caption{\label{fig:posterior_proton_shape}
-  Posterior distribution for the nucleon width $w$ and constituent width $v$.
-The white trapezoidal region (lower-right) covers the initial range of allowed values for $w$ and $v$, while the gray triangular region (upper-left) indicates values which are excluded by the prior $v > w$.
-The posterior distribution, shown in blue, indicates the preferred values for $w$ and $v$ determined by the analysis.}
-\end{figure}
-
-\begin{figure}[b]
-  \includegraphics{posterior_parton_number}
-  \caption{\label{fig:posterior_parton_number}
-    Posterior distribution for the number of nucleon constituents $n_c$ determined by the analysis.
-    The parameter $n_c$ is an integer (discrete) variable at every design point, but the emulator interpolation produces a posterior distribution which is continuous.
-  }
-\end{figure}
-
 The \trento\ normalization factor ${\text{Norm} = 20.3^{+2.2}_{-2.2}}$ and generalized mean parameter $p=0.018^{+0.146}_{-0.159}$ are well constrained by the present analysis.
@@
 
+Continuing down the diagonal in Fig.~\ref{fig:posterior}, we see that the nucleon width parameter $w=0.97_{-0.17}^{+0.19}$~fm.
+We caution that this parameter $w$ is \emph{not} the RMS radius of our nucleons due to idiosyncrasies of our constituent sampling procedure.
+We can, however, easily calculate the RMS radius for a specific nucleon width $w$, constituent width $v$ and constituent number $n_c$.
+For example, the single highest posterior probability region of the parameter space prefers a nucleon width $w=0.92$~fm, constituent width $v=0.43$~fm and constituent number $n_c=6$.
+The corresponding RMS nucleon radius for these parameters is $R_n = 0.86$~fm, conspicuously close the proton's RMS electric charge radius $R_p = 0.879(8)$~fm \cite{Bernauer:2010wm}.
+This is perhaps the single largest difference between our work and the conclusions of recent saturation-based calculations which constrained the event-by-event fluctuations of the proton using a color-dipole picture of vector meson production \cite{Mantysaari:2016ykx, Mantysaari:2016jaz}.
+Those studies find that the measured coherent and incoherent $J/\Psi$ spectra at HERA prefer a compact gluon distribution inside each nucleon, with an RMS radius $R_g \approx 0.4$~fm, which is roughly \emph{half} the analogous RMS nucleon width preferred by our analysis.
+
+Meanwhile, our posterior on the constituent number $n_c$, shown enlarged in Fig.~\ref{fig:posterior_parton_number}, is not sharply peaked.
+We therefore refrain from quoting a distribution median and 90\% credible interval, although we do note that the distribution clearly favors $n_c > 1$ constituents.
+This is not surprising.
+The \trento\ model mimics saturation-based initial condition models \cite{Bernhard:2016tnd}, and saturation models tend to produce ``proton-sized'' fireballs in $p$-Pb collisions \cite{Bzdak:2013zma}.
+When the proton is spherically symmetric, the resulting proton-sized QGP is also largely symmetric and thus produces very little flow.
+Saturation-based models are therefore unable to describe the significant flow measured in high-multiplicity $p$-Pb collisions without nucleon substructure, or alternatively, some other source of additional correlations \cite{Schenke:2017bog}.
+
+\begin{figure}[t]
+  \includegraphics{posterior_parton_number}
+  \caption{\label{fig:posterior_parton_number}
+    Posterior distribution for the number of nucleon constituents $n_c$ determined by the analysis.
+    The parameter $n_c$ is an integer (discrete) variable at every design point, but the emulator interpolation produces a posterior distribution which is continuous.
+  }
+\end{figure}
+
+The posterior on the nucleon substructure parameter $\X=0.34_{-0.17}^{+0.22}$, on the other hand, is particularly sharply peaked.
+Recall that this parameter, defined in Eqn.~\eqref{eq:struct_param}, interpolates between the minimum and maximum widths of each constituent and hence
+two different limits for the granularity of the nucleon.
+When $\X=0$, the nucleon is populated by $n_c$ small compact hot-spots, and when $\X=1$ the hot-spots are large and fully overlapping, restoring spherical symmetry.
+
+The nucleon structure parameter $\X$ is somewhat awkward to conceptualize, but we can easily transform its value back into a constituent width $v$.
+Figure~\ref{fig:posterior_proton_shape} shows the resulting joint posterior distribution for the nucleon width parameter $w$ and constituent width $v$.
+The white trapezoidal region in the lower-right is the joint parameter prior, and the gray triangular region in the upper-left where $v > w$ is a region that we exclude from the prior.
+The posterior distribution, shown in blue, is remarkably well constrained.
+We report a posterior estimate for the constituent width $v=0.46_{-0.15}^{+0.21}$~fm which is roughly the same magnitude as the \emph{nucleon} hot-spots used in recent saturation-based substructure studies that employed IP-Glasma initial conditions \note.
+Evidently, it may be necessary to place an informative prior on our nucleon substructure parameters in order to resolve the apparent tension between our parameter values and those needed to describe DIS measurements at HERA.
+Alternatively, it is also possible that the fluctuations probed by coherent and incoherent $J/\Psi$ production are different than those probed by minimum-bias particle production.
+
+\begin{figure}
+  \includegraphics{posterior_proton_shape}
+  \caption{\label{fig:posterior_proton_shape}
+  Posterior distribution for the nucleon width $w$ and constituent width $v$.
+The white trapezoidal region (lower-right) covers the initial range of allowed values for $w$ and $v$, while the gray triangular region (upper-left) indicates values which are excluded by the prior $v > w$.
+The posterior distribution, shown in blue, indicates the preferred values for $w$ and $v$ determined by the analysis.}
+\end{figure}
+
+\subsection{Transport properties}
+
 \begin{figure*}
@@
 
-Continuing down the diagonal in Fig.~\ref{fig:posterior}, we see that the nucleon width parameter $w=0.97_{-0.17}^{+0.19}$~fm.
-We caution that this parameter $w$ is \emph{not} the RMS radius of our nucleons due to idiosyncrasies of the constituent sampling procedure.
-We can, however, easily calculate the RMS radius for a specific nucleon width $w$, constituent width $v$ and constituent number $n_c$.
-For example, the single highest posterior probability region of the parameter space prefers a nucleon width $w=0.92$~fm, constituent width $v=0.43$~fm and constituent number $n_c=6$.
-The corresponding RMS nucleon radius for these parameters is $R_n = 0.86$~fm, conspicuously close the proton's RMS electric charge radius $R_p = 0.879(8)$~fm \cite{Bernauer:2010wm}.
-This is perhaps the single largest difference between our work and the conclusions of recent saturation-based calculations which constrained the event-by-event fluctuations of the proton using a color-dipole picture of vector meson production \cite{Mantysaari:2016ykx, Mantysaari:2016jaz}.
-Those studies find that the measured coherent and incoherent $J/\Psi$ spectra at HERA prefer a compact gluon distribution inside each nucleon, with an RMS radius $R_g \approx 0.4$~fm, which is roughly \emph{half} the analogous RMS nucleon width preferred by our analysis.
-
-Meanwhile, our posterior on the constituent number $n_c$ is not sharply peaked.
-We therefore refrain from quoting a distribution median and 90\% credible interval, although we do note that the distribution clearly favors $n_c > 1$ constituents.
-This is not surprising.
-The \trento\ model mimics saturation-based initial condition models \cite{Bernhard:2016tnd}, and saturation models tend to produce ``proton-sized'' fireballs in $p$-Pb collisions \cite{Bzdak:2013zma}.
-When the proton is spherically symmetric, the resulting proton-sized QGP is also largely symmetric and thus produces very little flow.
-Saturation-based models are therefore unable to describe the significant flow measured in high-multiplicity $p$-Pb collisions without nucleon substructure, or alternatively, some other source of additional correlations \cite{Schenke:2017bog}.
-
-The posterior on the nucleon substructure parameter $\X=0.34_{-0.17}^{+0.22}$, on the other hand, is particularly sharply peaked.
-Recall that this parameter, defined in Eqn.~\eqref{eq:struct_param}, interpolates between the minimum and maximum widths of each constituent and hence
-two different limits for the granularity of the nucleon.
-When $\X=0$, the nucleon is populated by $n_c$ small compact hot-spots, and when $\X=1$ the hot-spots are large and fully overlapping, restoring spherical symmetry.
-We use the parameter $\X$ purely for calibration purposes and then transform the variable back to a constituent width $v$ for discussion purposes.
-
-We plot the joint posterior distribution for the nucleon width parameter $w$ and constituent width $v$ in Fig.~\ref{fig:posterior_proton_shape}.
-The white trapezoidal region in the lower-right is the joint parameter prior, and the gray triangular region in the upper-left where $v > w$ is a region that we exclude from the prior.
-The posterior distribution, shown in blue, is remarkably well constrained.
-We report a posterior estimate for the constituent width $v=0.46_{-0.15}^{+0.21}$~fm which is roughly the same magnitude as the \emph{nucleon} hot-spots used in recent saturation-based substructure studies that employed IP-Glasma initial conditions \note.
-Evidently, it may be necessary to place an informative prior on our nucleon substructure parameters in order to resolve the apparent tension between our parameter values and those needed to describe DIS measurements at HERA.
-Alternatively, it is also possible that the fluctuations probed by coherent and incoherent $J/\Psi$ production are different than those probed by minimum-bias particle production.
-
-\subsection{Transport properties}
-
 In this section, we compare several of our posterior estimates to those obtained from a similar Bayesian analysis \cite{Bernhard:2018hnz} which used an (almost) identical version of the present physics model.
@@
 
-\begin{table}[t]
+We verified the emulator and tested the accuracy of our physics model framework using a single set of high-probability parameters selected from the Bayesian posterior.
+These parameters, listed in Tbl.~\ref{tab:mode_params}, are the approximate ``best fit'' values of the calibrated model, commonly referred to as the \emph{maximum a posteriori} (MAP) estimate:
+\begin{equation}
+  \x_\mathrm{MAP} \equiv \operatorname*{arg\, max}_{\x} P(H_\x | E).
+\end{equation}
+We then ran \order{6} minimum-bias and multiplicity triggered events using the MAP estimate $\x_\mathrm{MAP}$ and computed all of the model observables listed in Sec.~\ref{sec:observables}.
+The resulting model calculations are shown in Fig.~\ref{fig:obs_map} alongside experimental data from CMS \cite{Chatrchyan:2013nka} and ALICE \cite{Adam:2015ptt, Adam:2016izf, Adam:2014qja, Abelev:2013bla}.
+The left and right columns show the results for the $p$-Pb and Pb-Pb collision systems respectively, and each row shows a different group of related observables.
+
+The global agreement of the MAP model calculations with the experimental data is very good.
+The largest tension is observed in the two-particle cumulants $\vnk{2}{2}$ and $\vnk{3}{2}$ of the $p$-Pb system, although even that tension is only about 10--15\%.
+Quite remarkably, the model perfectly describes the shape of the $p$-Pb and Pb-Pb two-particle correlations which is strong evidence that these correlations are hydrodynamic in origin.
+Moreover, we obtain an excellent description of the \mbox{$p$-Pb} mean $p_T$, although this fit is somewhat less meaningful since we are unable to calibrate on the Pb-Pb mean $p_T$ simultaneously (data is not yet available).
+Additionally, the model provides a simultaneous description of the \mbox{$p$-Pb} and Pb-Pb charged-particle yields using a single entropy deposition parameter $p=0$.
+This is the \emph{exact same} generalized mean $p$-value supported by multiple previous studies \cite{Moreland:2014oya, Bernhard:2016tnd, Ke:2016jrd, Bernhard:2018hnz}.
+Evidently, this scaling continues to hold for initial conditions with sizable nucleon substructure.
+
+We also present calculations for several observables which were omitted from the calibration due to the statistical limitations of our training data.
+Here our MAP event sample is several orders-of-magnitude larger so the statistics are no issue.
+The bottom-right panel of Fig.~\ref{fig:obs_map} shows our model calculation for the four-particle elliptic flow cumultant $\vnk{2}{4}$ along with the measured data points from \mbox{ALICE} \cite{Adam:2016izf}.
+We see that the MAP estimate nicely describes the measured $\vnk{2}{4}$ data which is encouraging since this particular observable was never used to calibrate the posterior.
+
+The relative mean $p_T$ fluctuation $\delta p_T / \langle p_T \rangle$ is another important bulk observable to test the predictions of the calibrated model.
+It measures the \emph{dynamical} component of event-by-event mean $p_T$ fluctuations, quantified by the two-particle correlator
+\begin{equation}
+  (\delta p_T)^2 = \langle \langle (p_{T,i} - \langle p_T \rangle) (p_{T,j} - \langle p_T \rangle) \rangle \rangle.
+  \label{eq:mean_pT_corr}
+\end{equation}
+The inner-average in Eqn.~\ref{eq:mean_pT_corr} runs over all pairs of particles $i,j$ in the same event, the outer average runs over all events in a given bin (centrality or multiplicity), and the symbol $\langle p_T \rangle$ denotes the usual mean transverse momentum of particles in the bin.
+The observable is typically presented in terms of the dimensionless ratio $\delta p_T / \langle p_T \rangle$, which quantifies the strength of the dynamical fluctuations in units of the average transverse momentum $\langle p_T \rangle$.
+
+We show the MAP estimate predictions for the $p$-Pb and Pb-Pb relative mean $p_T$ fluctuations $\delta p_T / \langle p_T \rangle$ at $\sqrts=5.02$~TeV in the third row of Fig.~\ref{fig:obs_map}.
+For the Pb-Pb system, we use centrality bins and for the $p$-Pb system we use the same relative multiplicity bins used for two-particle flows.
+The relative mean $p_T$ fluctuations were shown to be particularly sensitive to the existence of nucleon substructure \cite{Bozek:2017elk}, and thus it would be interesting to ultimately include this observable in the calibration when the data becomes available.
+
+\begin{table}[b]
   \caption{
@@
 
-We verified the emulator and tested the accuracy of our physics model framework using a single set of high-probability parameters selected from the Bayesian posterior.
-These parameters, listed in Tbl.~\ref{tab:mode_params}, are the approximate ``best fit'' values of the calibrated model, commonly referred to as the \emph{maximum a posteriori} (MAP) estimate:
-\begin{equation}
-  \x_\mathrm{MAP} \equiv \operatorname*{arg\, max}_{\x} P(H_\x | E).
-\end{equation}
-We then ran \order{6} minimum-bias and multiplicity triggered events using the MAP estimate $\x_\mathrm{MAP}$ and computed all of the model observables listed in Sec.~\ref{sec:observables}.
-The resulting model calculations are shown in Fig.~\ref{fig:obs_map} alongside experimental data from CMS \cite{Chatrchyan:2013nka} and ALICE \cite{Adam:2015ptt, Adam:2016izf, Adam:2014qja, Abelev:2013bla}.
-The left and right columns show the results for the $p$-Pb and Pb-Pb collision systems respectively, and each row shows a different group of related observables.
-
-The global agreement of the MAP model calculations with the experimental data is very good.
-The largest tension is observed in the two-particle cumulants $\vnk{2}{2}$ and $\vnk{3}{2}$ of the $p$-Pb system, although even that tension is only about 10--15\%.
-Quite remarkably, the model perfectly describes the shape of the $p$-Pb and Pb-Pb two-particle correlations which is strong evidence that these correlations are hydrodynamic in origin.
-Moreover, we obtain an excellent description of the \mbox{$p$-Pb} mean $p_T$, although this fit is somewhat less meaningful since we are unable to calibrate on the Pb-Pb mean $p_T$ simultaneously (data is not yet available).
-Additionally, the model provides a simultaneous description of the \mbox{$p$-Pb} and Pb-Pb charged-particle yields using a single entropy deposition parameter $p=0$.
-This is the \emph{exact same} generalized mean $p$-value supported by multiple previous studies \cite{Moreland:2014oya, Bernhard:2016tnd, Ke:2016jrd, Bernhard:2018hnz}.
-Evidently, this scaling continues to hold for initial conditions with sizable nucleon substructure.
-
-We also present calculations for several observables which were omitted from the calibration due to the statistical limitations of our training data.
-Here our MAP event sample is several orders-of-magnitude larger so the statistics are no issue.
-The bottom-right panel of Fig.~\ref{fig:obs_map} shows our model calculation for the four-particle elliptic flow cumultant $\vnk{2}{4}$ along with the measured data points from \mbox{ALICE} \cite{Adam:2016izf}.
-We see that the MAP estimate nicely describes the measured $\vnk{2}{4}$ data which is encouraging since this particular observable was never used to calibrate the posterior.
-
-The relative mean $p_T$ fluctuation $\delta p_T / \langle p_T \rangle$ is another important bulk observable to test the predictions of the calibrated model.
-It measures the \emph{dynamical} component of event-by-event mean $p_T$ fluctuations, quantified by the two-particle correlator
-\begin{equation}
-  (\delta p_T)^2 = \langle \langle (p_{T,i} - \langle p_T \rangle) (p_{T,j} - \langle p_T \rangle) \rangle \rangle.
-  \label{eq:mean_pT_corr}
-\end{equation}
-The inner-average in Eqn.~\ref{eq:mean_pT_corr} runs over all pairs of particles $i,j$ in the same event, the outer average runs over all events in a given bin (centrality or multiplicity), and the symbol $\langle p_T \rangle$ denotes the usual mean transverse momentum of particles in the bin.
-The observable is typically presented in terms of the dimensionless ratio $\delta p_T / \langle p_T \rangle$, which quantifies the strength of the dynamical fluctuations in units of the average transverse momentum $\langle p_T \rangle$.
-
-We show the MAP estimate predictions for the $p$-Pb and Pb-Pb relative mean $p_T$ fluctuations $\delta p_T / \langle p_T \rangle$ at $\sqrts=5.02$~TeV in the third row of Fig.~\ref{fig:obs_map}.
-For the Pb-Pb system, we use centrality bins and for the $p$-Pb system we use the same relative multiplicity bins used for two-particle flows.
-The relative mean $p_T$ fluctuations were shown to be particularly sensitive to the existence of nucleon substructure \cite{Bozek:2017elk}, and thus it would be interesting to ultimately include this observable in the calibration when the data becomes available.
-
 \begin{fullpage}
@@
 
-As a final check, we compute the symmetric cumulants $\mathrm{SC}(m,n)$ for the Pb-Pb collision system at $\sqrts=5.02$~TeV.
-The symmetric cumulants quantify correlations between event-by-event fluctuations of the flow harmonics of order $v_m$ and $v_n$ \cite{Bilandzic:2013kga, ALICE:2016kpq},
+Lastly, we compute the symmetric cumulants $\mathrm{SC}(m,n)$ for the Pb-Pb collision system at $\sqrts=5.02$~TeV which quantify correlations between event-by-event fluctuations of the flow harmonics of order $v_m$ and $v_n$ \cite{Bilandzic:2013kga, ALICE:2016kpq},
 \begin{align}
@@
   \caption{Model calculations of the symmetric cumulants (top) and normalized symmetric cumulants (bottom) for Pb-Pb collisions at $\sqrts=5.02$~TeV using the \emph{maximum a posteriori} (MAP) parameters.
-The solid lines are the MAP estimate of the present analysis (with nucleon substructure), and the dashed lines are the MAP estimate of Ref.~\note\ (without nucleon substructure).
-In general, most model parameters are somewhat different between the two studies.}
+The solid lines are the MAP estimate of the present analysis (with nucleon substructure), and the dashed lines are the MAP estimate of Ref.~\cite{Bernhard:2018hnz} (without nucleon substructure).
+In general, most model parameters are different between the two studies.}
   \label{fig:flow_corr}
@@
 \begin{figure*}
-  \includegraphics[width=.6\textwidth]{validation_example}
+  \includegraphics{validation_example}
   \caption{Example emulator validation for one observable, the Pb-Pb charged particle yield $d\nch/d\eta$ in the 20-30\% centrality class.

### 2018-07-27 src/design.py

@@ -129,8 +129,8 @@ class Design:
             ('trento_p',      r'p',                           ( -1.0,    1.0)),
             ('fluct_std',     r'\sigma {fluct}',              (  0.0,    2.0)),
             ('nucleon_width', r'w [{fm}]',                    (  0.4,    1.2)),
-            ('parton_number', r'n {partons}',                 (    1,     10)),
-            ('parton_struct', r'\chi',                        (  0.0,    1.0)),
+            ('parton_number', r'n {constit}',                 (    1,     10)),
+            ('parton_struct', r'\chi {struct}',               (  0.0,    1.0)),
             ('dmin3',         r'd {min} [{fm}]',              (  0.0, 1.7**3)),
             ('tau_fs',        r'\tau {fs} [{fm}/c]',          (  0.1,    1.5)),
             ('etas_min',      r'\eta/s {min}',                (  0.0,    0.2)),

### 2018-07-27 src/model.py

@@ -301,8 +301,7 @@ def _data(system, dataset='main'):
         if dataset == 'map' else
         [
             Path(workdir, 'model_output', dataset, system, '{}.dat'.format(p))
-            for p in
-            Design(system, validation=(dataset == 'validation')).points
+            for p in Design(system, validation=(dataset == 'validation')).points
         ]
     )
 

### 2018-07-27 src/plots.py

@@ -469,7 +469,8 @@ def observables(system):
                 Y = Y*scale
 
             for y in Y:
-                ax.plot(x, y, color=color, alpha=.2, lw=.3)
+                alpha = .2 if system == 'pPb5020' else 0.1
+                ax.plot(x, y, color=color, alpha=alpha, lw=.3)
 
             try:
                 dset = expt.data[system][obs][subobs]
@@ -846,11 +847,11 @@ def flow_corr():
     """
     fig, axes = plt.subplots(
         figsize=figsize(0.5, 1.2), sharex=True,
-        nrows=2, ncols=1
+        nrows=2, gridspec_kw=dict(height_ratios=[4, 5])
     )
 
     observables = ['sc', 'sc_normed']
-    ylims = [(-2.5e-6, 2.5e-6), (-0.3, 0.8)]
+    ylims = [(-2.5e-6, 2.5e-6), (-.8, .8)]
     labels = ['(4,2)', '(3,2)']
     system = 'PbPb5020'
 
@@ -865,11 +866,7 @@ def flow_corr():
             x = model.map_data[system][obs][mn]['x']
             y = model.map_data[system][obs][mn]['Y']
 
-            ax.plot(
-                x, y, lw=.75,
-                color=getattr(plt.cm, cmap)(.7),
-            )
-
+            ax.plot(x, y, color=getattr(plt.cm, cmap)(.7))
             ax.text(1.02*x[-1], y[-1], lbl, va='center', ha='left')
 
         ax.axhline(
@@ -880,7 +877,7 @@ def flow_corr():
         ax.set_xlim(0, 80)
         ax.set_ylim(*ylim)
 
-        auto_ticks(ax, nbins=6, minor=2)
+        auto_ticks(ax, nbins=7, minor=2)
 
         if ax.is_first_col():
             ax.set_ylabel(label('m', 'n', normed='normed' in obs))
@@ -889,6 +886,46 @@ def flow_corr():
             ax.set_title('Pb-Pb 5.02 TeV')
         else:
             ax.set_xlabel('Centrality %')
+
+
+    # MAP estimate for Pb-Pb collisions at 5.02 TeV, calibrated to Pb-Pb
+    # data at 2.76 and 5.02 TeV using a model without nucleon substructure.
+
+    # symmetric cumulants
+    SC = np.array([
+        [2.5e+00, 5.8591e-09,  5.9204e-09],
[…]

### 2018-07-30 paper/substructure.tex
Writing summary

@@
 
-In this work, we constructed a multi-stage hybrid transport model for relativistic nuclear collisions and calibrated free parameters of the model using Bayesian parameter estimation.
-Specifically, we focused on improving the model's description of small collision systems by investigating the effect of nucleon substructure on bulk observables in $p$-Pb and Pb-Pb collisions at $\sqrts=5.02$~TeV.
-
-The analysis was performed using a parametric data-driven approach to account for theoretical uncertainty (variability) in the QGP initial conditions and hydrodynamic transport coefficients using a handful of equally uncertain model parametersi.
-We marginalized over the theoretical uncertainty in the parameter using a handful of equally uncertain model parameters to vary the size and shape of the nucleon, the scaling behavior of initial QGP entropy deposition, the duration of pre-equilibrium dynamics, and the temperature dependence of the hydrodynamic shear and bulk viscosities.
-
-We placed a conservative prior on the ``true'' value of each parameter, and calculated a global likelihood
-
-varied each parameter within its design range to explored the full 15-dimensional parameter space of our model.
-
-constructed an emulator to interpolate the predictions of our model at arbitrary regions of parameter space.
-
-Bayesian inference was then used to compare the predictions of the model to experimental data from CMS and ALICE, in order to quantify the posterior probability that a given set of model parameters describes the experimental data
-
+Relativistic heavy-ion collisions produce long-range multiparticle correlations which are commonly explained by the existence of hydrodynamic flow \note.
+This narrative is evidenced by the global, self-consistent and highly non-trivial quantitative agreement of hydrodynamic models with a large number of heavy-ion bulk observables \note.
+Naturally, such descriptions rely on the validity of hydrodynamic approximations, and these approximations begin to break down in the dilute particle limit where discrete particle degrees-of-freedom dominate and continuous field descriptions of the medium cease to make sense.
+Thus unlike heavy-ion collisions, small system collisions were not expected to exhibit hydrodynamic behavior.
+
+%Unlike heavy-ion collisions, small system collisions , were not expected
+%
+%Generally speaking, hydrodynamic flow signals were not expected to be observed
+%
+%
+%Recent measurements measurements of multiparticle correlations in $p$-Pb collisions are surprisingly similar to those measured in Pb-Pb collisions.
+%
+%The hydrodynamic framework, however, is naturally expected to break down in the dilute limit where discrete particle degrees of freedom dominate and continuous field descriptions of the medium cease to make sense.
+%It thus came somewhat as a surprise when high-multiplicity $p$-Pb collisions were found to produce long-range multiparticle correlations similar in magnitude to those observed in Pb-Pb collisions.
+%The observation challenged preconceived notions of hydrodynamic applicability and suggested that even proton-lead collisions produce small droplets of QGP liquid which evolve hydrodynamically.
+%
+%The potential for a unified hydrodynamic description of small and large collision systems is of course very appealing, although it is by no means given.
+%Multiple alternative explanations for the observed correlations have been proposed in the literature, and these mechanisms may contribute to the observed correlations in part or in full.
+%
+%Relativistic Au-Au and Pb-Pb collisions at RHIC and the LHC produce long-range multiparticle correlations which are commonly explained by the existence of hydrodynamic flow \note.
+%Perhaps the strongest evidence for this claim has been the overwhelming success of hydrodynamic models in describing (and even predicting) a large number of soft-sector observables in relativistic heavy-ion collisions \note.
+%
+%The hydrodynamic framework, however, is naturally expected to break down in the dilute limit where discrete particle degrees of freedom dominate and continuous field descriptions of the medium cease to make sense.
+%It thus came somewhat as a surprise when high-multiplicity $p$-Pb collisions were found to produce long-range multiparticle correlations similar in magnitude to those observed in Pb-Pb collisions.
+%The observation challenged preconceived notions of hydrodynamic applicability and suggested that even proton-lead collisions produce small droplets of QGP liquid which evolve hydrodynamically.
+%
+%The potential for a unified hydrodynamic description of small and large collision systems is of course very appealing, although it is by no means given.
+%Multiple alternative explanations for the observed correlations have been proposed in the literature, and these mechanisms may contribute to the observed correlations in part or in full.
+%
 \medskip

### 2018-07-31 paper/substructure.tex
Partway through summary and conclusion.

--- 2018-07-27 paper/substructure.tex
@@
 
-\section{summary}
+\section{summary and conclusions}
 \label{sec:summary}
 
-In this work, we constructed a multi-stage hybrid transport model for relativistic nuclear collisions and calibrated free parameters of the model using Bayesian parameter estimation.
-Specifically, we focused on improving the model's description of small collision systems by investigating the effect of nucleon substructure on bulk observables in $p$-Pb and Pb-Pb collisions at $\sqrts=5.02$~TeV.
-
-The analysis was performed using a parametric data-driven approach to account for theoretical uncertainty (variability) in the QGP initial conditions and hydrodynamic transport coefficients using a handful of equally uncertain model parametersi.
-We marginalized over the theoretical uncertainty in the parameter using a handful of equally uncertain model parameters to vary the size and shape of the nucleon, the scaling behavior of initial QGP entropy deposition, the duration of pre-equilibrium dynamics, and the temperature dependence of the hydrodynamic shear and bulk viscosities.
-
-We placed a conservative prior on the ``true'' value of each parameter, and calculated a global likelihood
-
-varied each parameter within its design range to explored the full 15-dimensional parameter space of our model.
-
-constructed an emulator to interpolate the predictions of our model at arbitrary regions of parameter space.
-
-Bayesian inference was then used to compare the predictions of the model to experimental data from CMS and ALICE, in order to quantify the posterior probability that a given set of model parameters describes the experimental data
-
-\medskip
+Relativistic heavy-ion collisions produce long-range multiparticle correlations which are commonly explained by the existence of hydrodynamic flow \note.
+This narrative is evidenced by the global, self-consistent and highly non-trivial quantitative agreement of hydrodynamic models with a large number of heavy-ion bulk observables \note.
+Naturally, such descriptions rely on the validity of hydrodynamic approximations, and these approximations begin to break down in the so-called dilute limit where discrete particle degrees-of-freedom dominate and continuous field descriptions of the medium cease to make sense.
+Tell-tale signatures of hydrodynamic collectivity were thus always expected to vanish in smaller nuclear collision systems, e.g.\ $p$-$p$ and $p$-Pb collisions, where the number of produced particles is orders of magnitude smaller than a typical Pb-Pb collision.
+
+These expectations were upended, however, when long-range multiparticle correlations were detected in high-multiplicity $p$-Pb collisions and found to be similar in magnitude to those observed in Pb-Pb collisions \note.
+Nuclear collision systems which were previously thought to be too small for hydrodynamic flow, were subsequently found to generate the same collectivity used to justify hydrodynamic flow in heavy-ion collisions.
+It is thus natural to wonder if a single unified hydrodynamic framework might be able to describe $p$-Pb and Pb-Pb bulk observables simultaneously.
+
+In this work, we performed a semi-exhaustive search for a unified description of $p$-Pb and Pb-Pb collisions at $\sqrts=5.02$~TeV using Bayesian methods to rigorously calibrate and constrain free parameters of a flexible nuclear collision model based on viscous hydrodynamics.
+The goal of our study was two fold.
+First, we aimed to establish whether or not our hydrodynamic framework was able to describe $p$-Pb and Pb-Pb bulk observables at $\sqrts=5.02$~TeV simultaneously, i.e.\ we attempted to falsify the framework.
+And second, given that the former is true, we wished to obtain estimates for the \emph{true} model parameters given the assumptions of our model and the evidence provided by the experimental data.
+
+Perhaps the most important difference between the present study and previous Bayesian analyses of heavy-ion collision models \note\ is the inclusion of variable nucleon substructure.
+We described the QGP initial state using a modified version of the parametric \trento\ model, with new parameters to vary the fluctuating size and shape of each nucleon.
+The nucleon was modeled as a cluster of $n_c$ Gaussian constituents (hot spots), and each constituent was described by a Gaussian density profile of width $v$.
+We also varied a parameter $w$ which controlled the sampling radius of the constituent positions inside each nucleon and hence the resulting RMS width of the ensemble averaged nucleon density.
+
+The transport dynamics were simulated using an event-by-event multi-stage nuclear collision model described by a pre-equilibrium free-streaming stage coupled to boost-invariant viscous hydrodynamics for the early hot and dense regions of the collision, and a microscopic hadronic afterburner to simulate hadronic interactions .
+We parametrized various sources of uncertainty including the duration of the pre-equilibrium free-streaming stage, the temperature dependence of the QGP shear and bulk viscosities, and the switching or particlization temperature where the medium description switches from hydrodynamics to microscopic Boltzmann transport.
+
+%Unfortunately, validating (or invalidating) a hydrodynamic picture of $p$-Pb and Pb-Pb collisions is not as simple as comparing a single model calculation to experimental data.
+%Hydrodynamic simulations typically include a number of free parameters which quantify uncertainties in the initial conditions and QGP medium, and these parameters can dramatically affect model predictions.
+%It's thus imperative to test all possible combinations of parameters in order to compare the full landscape of physically reasonable model calculations to experimental data.
+%If all possible combinations of model parameters fail to describe both collision systems simultaneously, then it's probably safe to say that the underlying assumptions of the chosen framework are either incomplete or invalid.
+%Similarly, if a single set of model parameters describes $p$-Pb and Pb-Pb bulk observables simultaneously, then it's compelling evidence that the hydrodynamic framework generalizes to small collision systems.
+%
+%There is also much more that can be learned about a model by testing all possible parameter combinations as opposed to a single set of values.
+%If the model works exceptionally well, but only for a certain range of parameter values, it's evidence that those parameter values are somehow constrained by the data.
+%Moreover, rigorous statistical methods may then be used to report parameter estimates with meaningful uncertainties which account for the accuracy of the chosen framework and the resolving power of the experimental data.
 

### 2018-08-01 paper/substructure.tex
Some extra work on the summary

@@
 
-Lastly, we compute the symmetric cumulants $\mathrm{SC}(m,n)$ for the Pb-Pb collision system at $\sqrts=5.02$~TeV which quantify correlations between event-by-event fluctuations of the flow harmonics of order $v_m$ and $v_n$ \cite{Bilandzic:2013kga, ALICE:2016kpq},
+Lastly, we compute the symmetric cumulants $\mathrm{SC}(m,n)$ for the Pb-Pb collision system at $\sqrts=5.02$~TeV which quantify correlations between event-by-event fluctuations of the flow harmonics of different order \cite{Bilandzic:2013kga, ALICE:2016kpq},
 \begin{align}
@@
 which are expected to be less sensitive to the medium response and more sensitive to the properties of the initial state.
-The solid lines are model calculations using the MAP parameters from the present analysis and the dashed lines are model calculations using the MAP parameters of Ref.~\note\ which did not include nucleon substructure.
-The gap between $\mathrm{SC}(3,2)$ and $\mathrm{SC}(4,2)$ is generally wider in the present analysis, as is the gap between the normalized symmetric cumulants $\mathrm{NSC}(3,2)$ and $\mathrm{NSC}(4,2)$.
-This may alleviate a common problem
+The solid lines are the MAP estimate of the present study and the dashed lines are the MAP estimate of Ref.~\cite{Bernhard:2018hnz} which did not include nucleon substructure and was calibrated on Pb-Pb collisions at $\sqrts=2.76$ and 5.02~TeV.
+We observe that the gap between $\mathrm{SC}(3,2)$ and $\mathrm{SC}(4,2)$ is generally wider in the present analysis than in Ref.~\cite{Bernhard:2018hnz}, as is the gap between the normalized symmetric cumulants $\mathrm{NSC}(3,2)$ and $\mathrm{NSC}(4,2)$.
+
+We emphasize that multiple aspects of the two analyses are different such as the collision systems and beam energies considered, the observables which were included in each calibration, and the existence of nucleon substructure to name a few.
+Thus we can only speculate what might have lead to the large difference in the symmetric flow cumulants.
+Two reasonable culprits would be the inclusion of nucleon substructure and the large difference in the preferred pre-equilibrium free-streaming time determined by the two studies.
+Identifying the source of the effect would be an interesting study for a future project.
 
@@
   \caption{Model calculations of the symmetric cumulants (top) and normalized symmetric cumulants (bottom) for Pb-Pb collisions at $\sqrts=5.02$~TeV using the \emph{maximum a posteriori} (MAP) parameters.
-The solid lines are the MAP estimate of the present analysis (with nucleon substructure), and the dashed lines are the MAP estimate of Ref.~\cite{Bernhard:2018hnz} (without nucleon substructure).
-In general, most model parameters are different between the two studies.}
+The solid lines are the MAP estimate of the present analysis (with nucleon substructure), and the dashed lines are the MAP estimate of Ref.~\cite{Bernhard:2018hnz} (without nucleon substructure) which was calibrated on Pb-Pb observables at $\sqrts=2.76$ and 5.02~TeV.
+In general, most model parameters are somewhat different between the two studies.}
   \label{fig:flow_corr}
@@
 The goal of our study was two fold.
-First, we aimed to establish whether or not our hydrodynamic framework was able to describe $p$-Pb and Pb-Pb bulk observables at $\sqrts=5.02$~TeV simultaneously, i.e.\ we attempted to falsify the framework.
-And second, given that the former is true, we wished to obtain estimates for the \emph{true} model parameters given the assumptions of our model and the evidence provided by the experimental data.
-
-Perhaps the most important difference between the present study and previous Bayesian analyses of heavy-ion collision models \note\ is the inclusion of variable nucleon substructure.
-We described the QGP initial state using a modified version of the parametric \trento\ model, with new parameters to vary the fluctuating size and shape of each nucleon.
-The nucleon was modeled as a cluster of $n_c$ Gaussian constituents (hot spots), and each constituent was described by a Gaussian density profile of width $v$.
-We also varied a parameter $w$ which controlled the sampling radius of the constituent positions inside each nucleon and hence the resulting RMS width of the ensemble averaged nucleon density.
-
-The transport dynamics were simulated using an event-by-event multi-stage nuclear collision model described by a pre-equilibrium free-streaming stage coupled to boost-invariant viscous hydrodynamics for the early hot and dense regions of the collision, and a microscopic hadronic afterburner to simulate hadronic interactions .
-We parametrized various sources of uncertainty including the duration of the pre-equilibrium free-streaming stage, the temperature dependence of the QGP shear and bulk viscosities, and the switching or particlization temperature where the medium description switches from hydrodynamics to microscopic Boltzmann transport.
-
-%Unfortunately, validating (or invalidating) a hydrodynamic picture of $p$-Pb and Pb-Pb collisions is not as simple as comparing a single model calculation to experimental data.
-%Hydrodynamic simulations typically include a number of free parameters which quantify uncertainties in the initial conditions and QGP medium, and these parameters can dramatically affect model predictions.
-%It's thus imperative to test all possible combinations of parameters in order to compare the full landscape of physically reasonable model calculations to experimental data.
-%If all possible combinations of model parameters fail to describe both collision systems simultaneously, then it's probably safe to say that the underlying assumptions of the chosen framework are either incomplete or invalid.
-%Similarly, if a single set of model parameters describes $p$-Pb and Pb-Pb bulk observables simultaneously, then it's compelling evidence that the hydrodynamic framework generalizes to small collision systems.
-%
-%There is also much more that can be learned about a model by testing all possible parameter combinations as opposed to a single set of values.
-%If the model works exceptionally well, but only for a certain range of parameter values, it's evidence that those parameter values are somehow constrained by the data.
-%Moreover, rigorous statistical methods may then be used to report parameter estimates with meaningful uncertainties which account for the accuracy of the chosen framework and the resolving power of the experimental data.
+First, we aimed to establish whether or not our hydrodynamic framework was able to describe both collision systems simultaneously.
+And second, in the event that the former was true, we wished to obtain estimates for the \emph{true} parameters of our model given the assumptions of our framework and the evidence provided by the experimental data.
+
+We built, for this purpose, a flexible multi-stage nuclear collision model characterized by a number of free parameters which vary theoretically uncertain aspects of the framework such as the QGP initial conditions and hydrodynamic transport properties.
+For the QGP initial conditions, we developed a modified version of the \trento\ model which adds new parameters to vary the fluctuating size and shape of each nucleon.
+Specifically, we modeled each nucleon as a cluster of $n_c$ constituents (hot spots), where each constituent is described by a Gaussian density profile of width $v$.
+We also introduced a parameter $w$ which effectively varies the width of the Gaussian distribution used to sample constituent positions inside each nucleon.
+
+The transport dynamics of the collision were simulated using a pre-equilibrium free-streaming stage of variable duration followed by boost-invariant viscous hydrodynamics for hot and dense regions of the fireball and a microscopic hadronic afterburner for the dilute corona.
+We parametrized various sources of uncertainty in each stage of the collision including the duration of the pre-equilibrium free-streaming stage, the temperature dependence of the QGP shear and bulk viscosities, and the isothermal particlization temperature used to switch from a hydrodynamic description to microscopic Boltzmann transport.
+
+We applied Bayesian parameter estimation which were developed to estimate the parameters of computationally expensive model.
+
+\medskip
+
+...work in progress
+
+\vfill
 

### 2018-08-01 paper/substructure.tex
Start finishing summary and conclusion

@@
 \definecolor{theblue}{RGB}{0,50,230}
+
+\usepackage{booktabs}
 
@@
 \newcommand{\ntrk}{N_\text{trk}^\text{offline}}
+\newcommand{\Tsw}{T_\text{switch}}
 \newcommand{\sigmaf}{\sigma_\mathrm{fluct}}
@@
 \newcommand{\dmin}{d_\mathrm{min}}
+\newcommand{\fmc}{\ensuremath{\text{fm}/c}}
 \newcommand{\T}{\tilde{T}}
@@
 and each constituent's position $\mathbf{x_i}$ in Eqn.~\eqref{eq:rho} is sampled from a Gaussian radial distribution with standard deviation $r$.
+The effect of this addition nucleon substructure on the nuclear thickness functions is visible in Fig.~\ref{fig:thickness}
 
@@
   \end{tikzpicture}
-  \caption{Effect of nucleon substructure on the nuclear thickness function $T(x, y) \equiv \int dz\, \rho(x, y, z)$ of a $^{208}\mathrm{Pb}$ nucleus.
+  \caption{\label{fig:thickness} Effect of nucleon substructure on the nuclear thickness function $T(x, y) \equiv \int dz\, \rho(x, y, z)$ of a $^{208}\mathrm{Pb}$ nucleus.
 The nucleus on the left has Gaussian nucleons of width $0.8$~fm, while the nucleus on the right has composite nucleons, each containing six constituents of width $0.4$~fm.}
-  \label{fig:thickness}
 \end{figure}
@@
 Each diagonal panel is the distribution of a \emph{single} model parameter (marginalized over all others), and each lower-diagonal panel is the joint distribution of a \emph{pair} of model parameters (marginalized over all others).
-
-The posterior in Fig.~\ref{fig:posterior} includes a wealth of information.
-Here we focus on its most important features, starting with the top-left corner of the figure and working our way to the bottom-right.
-We report posterior estimates for individual parameters using their (marginalized) distribution medians and 90\% credible intervals.
+We also report numeric estimates for each model parameter along the distribution diagonal.
+These estimates describe the marginalized distribution median and 90\% credible interval for each parameter.
 For example, the estimate $x=2.45_{-0.15}^{+0.20}$ denotes a median value $\tilde{x}=2.45$ and 90\% credible interval $2.30 < x < 2.65$.
+These posterior parameter estimates are also listed in Tbl.~\ref{tab:post_param} for clarity.
 
@@
 
+\begin{table}[t]
+  \caption{
+    \label{tab:post_param}
+    Posterior parameter estimates corresponding to Fig.~\ref{fig:posterior}.
+    The reported values are for the distribution median and 90\% highest posterior density credible interval.
+  }
+  \begin{ruledtabular}
+    \newlength{\cellwidth}
+    \settowidth{\cellwidth}{$-0.00$}
+    \newcommand{\est}[3]{\parbox{\cellwidth}{\hfill$#1$}$_{-#2}^{+#3}$}
+    \begin{tabular}{llll}
+      \toprule
+      \multicolumn{2}{c}{Initial condition / Pre-eq}     & \multicolumn{2}{c}{QGP medium}              \\
+      \cmidrule(r){1-2}                                    \cmidrule(l){3-4}
+      \addlinespace[.4ex]
+      Norm       & \est{20.3}{2.2}{2.2}          & $\eta/s$ min     & \est{0.10}{0.07}{0.07}            \\[1.1ex]
+      $p$        & \est{0.018}{0.159}{0.146}     & $\eta/s$ slope   & \est{1.16}{1.16}{1.55} GeV$^{-1}$ \\[1.1ex]
+      $\sigmaf$  & \est{0.93}{0.27}{0.27}        & $\eta/s$ crv     & \est{-0.21}{0.79}{0.85}           \\[1.1ex]
+      $w$        & \est{0.97}{0.17}{0.19} fm     & $\zeta/s$ max    & \est{0.027}{0.027}{0.034}         \\[1.1ex]
+      $n_c$      & \est{6.2}{3.1}{2.8}           & $\zeta/s$ width  & \est{0.032}{0.032}{0.047} GeV     \\[1.1ex]
+      $\X$       & \est{0.34}{0.17}{0.22}        & $\zeta/s$ $T_0$  & \est{0.176}{0.020}{0.024} GeV     \\[1.1ex]
+      $\dmin$    & \est{1.12}{0.50}{0.56} fm     & $\Tsw$           & \est{0.150}{0.012}{0.015} GeV     \\[1.1ex]
+      $\taufs$   & \est{0.37}{0.27}{0.34} \fmc   & \\
+      \addlinespace[.4ex]
+      \bottomrule
+    \end{tabular}
+  \end{ruledtabular}
+\end{table}
+
 \subsection{Initial condition properties}
@@
 
+\begin{figure}[b]
+  \includegraphics{posterior_parton_number}
+  \caption{\label{fig:posterior_parton_number}
+    Posterior distribution for the number of nucleon constituents $n_c$ determined by the analysis.
+    The parameter $n_c$ is an integer (discrete) variable at every design point, but the emulator interpolation produces a posterior distribution which is continuous.
+  }
+\end{figure}
+
 Continuing down the diagonal in Fig.~\ref{fig:posterior}, we see that the nucleon width parameter $w=0.97_{-0.17}^{+0.19}$~fm.
@@
 
+\begin{figure}
+  \includegraphics{posterior_proton_shape}
+  \caption{\label{fig:posterior_proton_shape}
+  Posterior distribution for the nucleon width $w$ and constituent width $v$.
+The enclosed trapezoidal region covers the prior range of allowed values for $w$ and $v$.
+The posterior distribution, shown in blue, indicates the preferred values for $w$ and $v$ determined by the analysis.}
+\end{figure}
+
 Meanwhile, our posterior on the constituent number $n_c$, shown enlarged in Fig.~\ref{fig:posterior_parton_number}, is not sharply peaked.
@@
 Saturation-based models are therefore unable to describe the significant flow measured in high-multiplicity $p$-Pb collisions without nucleon substructure, or alternatively, some other source of additional correlations \cite{Schenke:2017bog}.
-
-\begin{figure}[t]
-  \includegraphics{posterior_parton_number}
-  \caption{\label{fig:posterior_parton_number}
-    Posterior distribution for the number of nucleon constituents $n_c$ determined by the analysis.
-    The parameter $n_c$ is an integer (discrete) variable at every design point, but the emulator interpolation produces a posterior distribution which is continuous.
-  }
-\end{figure}
 
@@
 
-\begin{figure}
-  \includegraphics{posterior_proton_shape}
-  \caption{\label{fig:posterior_proton_shape}
-  Posterior distribution for the nucleon width $w$ and constituent width $v$.
-The white trapezoidal region (lower-right) covers the initial range of allowed values for $w$ and $v$, while the gray triangular region (upper-left) indicates values which are excluded by the prior $v > w$.
-The posterior distribution, shown in blue, indicates the preferred values for $w$ and $v$ determined by the analysis.}
-\end{figure}
-
 \subsection{Transport properties}
@@
 
-\begin{table}[b]
+\begin{fullpage}
+{\centering\includegraphics{observables_map}\par}
+\captionof{figure}{\label{fig:obs_map} Model calculations using the \emph{maximum a posterior} (MAP) parameters compared to experiment.
+Colored lines are model calculations for $p$-Pb collisions (left) and Pb-Pb collisions (right) at $\sqrts=5.02$~TeV.
+Points with error bars are the experimental data with statistical uncertainties, and gray bands their corresponding systematic uncertainties, from CMS \cite{Chatrchyan:2013nka} and ALICE \cite{Adam:2015ptt, Adam:2016izf, Adam:2014qja, Abelev:2013bla}.
+The sub-axes show the ratio of model over data where available with gray bands indicating $\pm 10\%$.}
+\end{fullpage}
+
+\begin{table}
   \caption{
@@
     \begin{tabular}{ll@{\hspace{2em}}ll}
-      \multicolumn{2}{c}{Initial condition} & \multicolumn{2}{c}{QGP medium} \\
+      \multicolumn{2}{c}{Initial condition / Pre-eq} & \multicolumn{2}{c}{QGP medium} \\
       \paddedhline
@@
 
-\begin{fullpage}
-{\centering\includegraphics{observables_map}\par}
-\captionof{figure}{\label{fig:obs_map} Model calculations using the \emph{maximum a posterior} (MAP) parameters compared to experiment.
-Colored lines are model calculations for $p$-Pb collisions (left) and Pb-Pb collisions (right) at $\sqrts=5.02$~TeV.
-Points with error bars are the experimental data with statistical uncertainties, and gray bands their corresponding systematic uncertainties, from CMS \cite{Chatrchyan:2013nka} and ALICE \cite{Adam:2015ptt, Adam:2016izf, Adam:2014qja, Abelev:2013bla}.
-The sub-axes show the ratio of model over data where available with gray bands indicating $\pm 10\%$.}
-\end{fullpage}
-
 Lastly, we compute the symmetric cumulants $\mathrm{SC}(m,n)$ for the Pb-Pb collision system at $\sqrts=5.02$~TeV which quantify correlations between event-by-event fluctuations of the flow harmonics of different order \cite{Bilandzic:2013kga, ALICE:2016kpq},
@@
 We built, for this purpose, a flexible multi-stage nuclear collision model characterized by a number of free parameters which vary theoretically uncertain aspects of the framework such as the QGP initial conditions and hydrodynamic transport properties.
-For the QGP initial conditions, we developed a modified version of the \trento\ model which adds new parameters to vary the fluctuating size and shape of each nucleon.
+For the QGP initial conditions, we employed a modified version of the \trento\ model which adds new parameters to vary the fluctuating size and shape of each nucleon.
 Specifically, we modeled each nucleon as a cluster of $n_c$ constituents (hot spots), where each constituent is described by a Gaussian density profile of width $v$.
-We also introduced a parameter $w$ which effectively varies the width of the Gaussian distribution used to sample constituent positions inside each nucleon.
-
-The transport dynamics of the collision were simulated using a pre-equilibrium free-streaming stage of variable duration followed by boost-invariant viscous hydrodynamics for hot and dense regions of the fireball and a microscopic hadronic afterburner for the dilute corona.
-We parametrized various sources of uncertainty in each stage of the collision including the duration of the pre-equilibrium free-streaming stage, the temperature dependence of the QGP shear and bulk viscosities, and the isothermal particlization temperature used to switch from a hydrodynamic description to microscopic Boltzmann transport.
-
-We applied Bayesian parameter estimation which were developed to estimate the parameters of computationally expensive model.
-
-\medskip
-
-...work in progress
-
-\vfill
+Each constituent position inside the nucleon was sampled randomly (without correlations) according to a Gaussian radial distribution of variable width, controlled indirectly by an effective nucleon width parameter $w$.
+
+The transport dynamics of the collision were simulated using a pre-equilibrium free-streaming stage of variable duration followed by boost-invariant viscous hydrodynamics for hot and dense regions of the fireball and a microscopic hadronic afterburner for the relatively dilute corona.
+We parametrized various sources of uncertainty in each stage of the collision including the duration of the pre-equilibrium free-streaming stage, the temperature dependence of the QGP shear and bulk viscosities, and the particlization temperature used to switch from a hydrodynamic description to microscopic Boltzmann transport.
+
+With the full evolution model in hand, we applied Bayesian methods which were developed to estimate the parameters of computationally intensive models \note.
+We first constructed a scaffolding of $n=500$ parameters points distributed throughout our 15-dimensional parameter space and evaluated the nuclear collision model using \order{4} events at each parameter point.
+The ensemble of events was then used to calculate a large number of experimental observables at each design point and train Gaussian process emulators to interpolate the model predictions as a function of the input parameters.
+Finally, we used Markov chain Monte Carlo (MCMC) importance sampling to explore the parameter space and draw samples from the Bayesian posterior distribution for the \emph{true} values of the model parameters, given our emulated model predictions, the experimental data, and their associated uncertainties.
+
+The model calibration process is summarized by Figs.~\ref{fig:obs_pbpb} and \ref{fig:obs_ppb}, and the resulting posterior distribution for the model input parameters is shown in Fig.~\ref{fig:posterior}.
+We also evaluated the model predictions using a single set of high-probability parameters in Fig.~\ref{fig:obs_map}.
+With these results, we are able to address the two primary goals of the study.
+
+First, we demonstrated in Fig.~\ref{fig:obs_map} the existence of a single set of model parameters which can simultaneously describe $p$-Pb and Pb-Pb charged-particle yields, mean $p_T$, and flow cumulants at $\sqrts=5.02$~TeV.
+The excellent quantitative agreement of the model calculation with the experimental data is strong evidence for a unified hydrodynamic description of $p$-A and A-A collisions.
+Moreover, the modifications to the physics model which were required to obtain this agreement are generally modest; one must simply replace Gaussian nucleons with composite nucleons of several or more constituents.
+
+Second, we obtained a posterior distribution for the model input parameters in Fig.~\ref{fig:posterior} and reported quantitative estimates for the model input parameters in Tbl.~\ref{tab:post_param}.
+Here we summarize our key findings about the model:
+\begin{enumerate}[itemsep=0pt, leftmargin=2\parindent]
+  \item
+    Using the \trento\ initial condition model, we find that initial entropy deposition (or particle production) scales like the geometric mean of local participant nuclear densities, in agreement with a number of previous estimates \note.
+  \item
+    Nucleon substructure is necessary to simultaneously describe $p$-Pb and Pb-Pb bulk observables. However, there is no strong preference for a specific number of constituents inside the nucleon.
+  \item
+    Bulk observables in $p$-Pb and Pb-Pb collisions prefer larger nucleons, similar in size to the proton's RMS electric charge radius.
+    This agrees with a similar Bayesian analysis which was calibrated to Pb-Pb bulk observables at $\sqrts=2.76$ and 5.02~TeV using a hybrid model without nucleon substructure \note.
+    We note that our result is in tension an estimate for the effective nucleon width based on the gluon distribution extracted from HERA data \note.
+  \item
+    We obtain an estimate $v=0.46_{-0.15}^{+0.21}$~fm for the Gaussian width of individual constituent hot spots inside the nucleon is which is significantly larger than the length scales typically associated with nucleon substructure.
+    This result may help explain the seemingly unreasonable success of hydrodynamics in small collision systems.
+  \item
+    Our posterior estimate for the pre-equilibrium free-streaming time is $\taufs=0.37_{-0.27}^{+0.34}$~fm/$c$.
+    This is significantly shorter than the estimate obtained from a similar Bayesian study in Ref.~\cite{Bernhard:2018hnz}, which reported $\taufs=1.16_{-0.25}^{+0.29}$~fm/$c$.
+    It is not clear whether the difference is a result of nucleon substructure or the different observables used to calibrate each analysis.
+  \item
+    The posterior estimates for the temperature dependence of the QGP specific shear and bulk viscosities
+\end{enumerate}
+
+The present study would benefit from a number of improvements.
+Perhaps the most obvious improvement would be to add missing experimental data to the calibration once it becomes available.
+For example, the present calibration is missing the Pb-Pb mean $p_T$, as well as the transverse energy, identified particle yields, and mean $p_T$ fluctuations of both collision systems.
+These should certainly influence the quality of the combined fit and the posterior estimates for the model parameters.
+
+Similarly, the analysis would greatly benefit from additional beam energies and collision systems.
+Notably missing are measurements of $p$-Au, $d$-Au and $^3$He-Au collisions at RHIC \note, which would provide an additional constraint on the parameters and more stringent tests of the calibrated model predictions.
+The RHIC data may also help elucidate the beam-energy dependence of the model parameters which would be worth investigating.
+We leave these improvements for future work.
 
@@
 DE-AC02-05CH11231.
-  JSM, JEB and SAB acknowledge support by ...
+  JSM, JEB and SAB are supported by the U.S.\ Department of Energy Grant No.\ DE-FG02-05ER41367 and National Science Foundation Grant No.\ NSF-ACI-1550225.
 \end{acknowledgments}
@@
 
-\begin{figure}
+The boost-invariant \textsc{VISH2+1} hydrodynamics code used in this work \cite{Song:2007ux, Shen:2014vra} runs on a Cartesian transverse grid specified by a maximum grid size $x_\text{max}$ and grid step width $dx$ which fix the transverse grid extent ${-x_\text{max} < x < x_\text{max}}$ and number of grid cells along each dimension ${n_x = 2\, x_\text{max} / dx}$.
+In general, the maximum grid size $x_\text{max}$ should be set large enough to contain the full spacetime evolution of the event.
+This means that the truncation of $T^{\mu\nu}$ at the boundaries of the grid should never modify the final-state observables.
+We enforce this requirement by finding an energy density cutoff $e_\text{min}$ for which the matter $e < e_\text{min}$ can be effectively discarded without significantly modifying the simulation observables.
+We then fix the maximum grid size $x_\text{max}$ such that it fully encloses the isotherm $T=T(e_\text{min})$ for the full lifetime of the fireball.
+
+We find that we can quickly estimate the maximum radius $R_\text{max} = |\x_\text{max}|^2$ of the spacetime hypersurface ${T = T(e_\text{min})}$ by running the event on a coarse-grained spatial grid with one-third the spatial resolution we would otherwise require to resolve typical hydrodynamic observables such as mean $p_T$ and flows.
+The simulation time of a single \textsc{VISH2+1} event scales like ${\sim}n_x^3$ since $dx \propto d\tau$, and thus our ``pre-run'' event requires only ${\sim}1/27$th the time of a production event.
+We therefore run a coarse-grained pre-event on an excessively large grid for \emph{every} minimum-bias event to estimate $R_\text{max}$, then rerun the same event on a thrice finer grid with a trimmed spatial extent $x_\text{max} \equiv R_\text{max}$.
+See Fig.~\ref{fig:adaptive_grid} for a simple diagram of the procedure.
+
+In practice, we find that event-by-event grid resizing leads to a massive speed increase for minimum bias events compared to using a single fixed grid for the entire minimum bias sample.
+This is because the maximum transverse size of each event varies dramatically, from a few fm in peripheral Pb-Pb collisions to 50 fm or more in central Pb-Pb collisions.
+The procedure should generalize to other hydrodynamic codes, including those with $3+1$ spacetime dimensions, where the time savings could be even more pronounced.
+
+\begin{figure}[t]
   \begin{tikzpicture}

### 2018-08-01 src/plots.py

@@ -37,7 +37,7 @@ from matplotlib import ticker
 from matplotlib import cm
 from matplotlib.colors import ListedColormap, LinearSegmentedColormap
 from scipy import special
-from scipy.interpolate import PchipInterpolator, interp2d
+from scipy.interpolate import PchipInterpolator, UnivariateSpline, interp2d
 from scipy.optimize import curve_fit
 from sklearn.decomposition import PCA
 from sklearn.gaussian_process import GaussianProcessRegressor as GPR
@@ -851,7 +851,7 @@ def flow_corr():
     )
 
     observables = ['sc', 'sc_normed']
-    ylims = [(-2.5e-6, 2.5e-6), (-.8, .8)]
+    ylims = [(-2.5e-6, 2.5e-6), (-.9, .8)]
     labels = ['(4,2)', '(3,2)']
     system = 'PbPb5020'
 
@@ -925,7 +925,7 @@ def flow_corr():
     handles = [solid_line, dashed_line]
     labels = ["p-Pb, Pb-Pb 5.02 TeV", "Pb-Pb 2.76, 5.02 TeV"]
 
-    plt.legend(handles, labels, loc=8)
+    plt.legend(handles, labels, loc=8, title='Bayesian calibration on:')
 
     set_tight(fig)
 
@@ -1142,7 +1142,7 @@ def posterior_p():
     set_tight(pad=0)
 
 
-def posterior_parameter(parameter, label, xticks):
+def posterior_parameter(parameter, label, xticks, bins=50):
     """
     Marginal distribution of a single parameter.
 
@@ -1152,11 +1152,11 @@ def posterior_parameter(parameter, label, xticks):
 
     data = mcmc.Chain().load(parameter).ravel()
 
-    counts, edges = np.histogram(data, bins=50)
+    counts, edges = np.histogram(data, bins=bins)
     x = (edges[1:] + edges[:-1]) / 2
     y = counts / counts.max()
     interp = PchipInterpolator(x, y)
-    x = np.linspace(x[0], x[-1], 10*x.size)
+    x = np.linspace(x[0], x[-1], 100*x.size)
     y = interp(x)
     ax.plot(x, y, color=plt.cm.Blues(0.8))
     ax.fill_between(x, y, color=plt.cm.Blues(0.15), zorder=-10)
@@ -1172,17 +1172,57 @@ def posterior_parameter(parameter, label, xticks):
 
 @plot
[…]

### 2018-08-02 paper/substructure.tex
Add missing citations

@@
 \usepackage[utf8]{inputenc}
+
 \usepackage{amsmath}
@@
 
-\usepackage[inline]{enumitem}
+\usepackage{graphicx}
+\graphicspath{{../plots/}{./fig/}}
+
+\usepackage{tikz}
 
@@
 
+\usepackage[font=small, justification=raggedright, labelsep=quad]{caption}
+
+\usepackage[inline]{enumitem}
+
 \usepackage{booktabs}
@@
 \usepackage[pdfencoding=auto, psdextra]{hyperref}
-\usepackage[font=small, justification=raggedright, labelsep=quad]{caption}
-
 \hypersetup{
@@
 
-\usepackage{tikz}
-\usepackage[leftcaption]{sidecap}
-
-\usepackage{graphicx}
-\graphicspath{{../plots/}{./fig/}}
-
 \newcommand{\trento}{T\raisebox{-0.5ex}{R}ENTo}
 \newcommand{\sqrts}{\sqrt{s_{NN}}}
+\newcommand{\fmc}{\ensuremath{\text{fm}/c}}
 \newcommand{\nch}{N_\text{ch}}
 \newcommand{\ntrk}{N_\text{trk}^\text{offline}}
+\newcommand{\vnk}[2]{v_#1\{#2\}}
+\newcommand{\sigmaf}{\sigma_\text{fluct}}
+\newcommand{\X}{\chi_\text{struct}}
+\newcommand{\taufs}{\tau_\text{fs}}
+\newcommand{\dmin}{d_\text{min}}
 \newcommand{\Tsw}{T_\text{switch}}
-\newcommand{\sigmaf}{\sigma_\mathrm{fluct}}
-\newcommand{\X}{\chi_\mathrm{struct}}
-\newcommand{\taufs}{\tau_\mathrm{fs}}
-\newcommand{\dmin}{d_\mathrm{min}}
-\newcommand{\fmc}{\ensuremath{\text{fm}/c}}
 \newcommand{\T}{\tilde{T}}
-\newcommand{\x}{\mathbf{x}}
-\newcommand{\y}{\mathbf{y}}
-\newcommand{\z}{\mathbf{z}}
-\newcommand{\vnk}[2]{v_#1\{#2\}}
+\newcommand{\x}{\mathbf x}
+\newcommand{\y}{\mathbf y}
+\newcommand{\z}{\mathbf z}
 \newcommand{\paddedhline}{\noalign{\smallskip}\hline\noalign{\smallskip}}
@@
 
+\DeclareMathOperator{\diag}{diag}
+\DeclareMathOperator{\cov}{cov}
+\DeclareMathOperator{\corr}{corr}
+\DeclareMathOperator{\SC}{SC}
+\DeclareMathOperator{\NSC}{NSC}
+
 % Reduce the size of the itemize bullets
@@
 % this removes the warning by redefining the line break in a pdf string
-\pdfstringdefDisableCommands{%
-  \def\\#1{ #1}
-}
+\pdfstringdefDisableCommands{\def\\#1{ #1}}
 
@@
 Ultrarelativistic nuclear collisions between one light-ion and one heavy-ion, e.g.\ $^3$He-Au and $p$-Pb collisions, generate dense, compact sources of nuclear matter which produce long-range multiparticle correlations that are strikingly similar to the correlations observed in heavy-ion collisions where collectivity is commonly explained by the existence of hydrodynamic flow \cite{CMS:2012qk, Abelev:2012ola, Aad:2012gla, Adare:2015ctn}.
-This observation suggests that hydrodynamic behavior could be manifest in small droplets of quark-gluon-plasma (QGP) \cite{Bozek:2011if, Bozek:2013uha}, and that flow might develop at length scales smaller than a single proton \note.
+This observation suggests that hydrodynamic behavior could be manifest in small droplets of quark-gluon-plasma (QGP) \cite{Bozek:2011if, Bozek:2013uha}, and that flow even might develop at length scales smaller than a single proton \cite{Schenke:2014zha}.
 
@@
 Initial condition studies using the IP-Glasma model of color-glass condensate effective field theory \cite{Schenke:2012wb} simultaneously demonstrated that these color charge fluctuations leave a lasting imprint on the \mbox{small-x} gluon distribution of the proton and hence the initial geometry of QGP energy deposition \cite{Schlichting:2014ipa}.
-In addition, it was recently shown that hydrodynamic simulations using IP-Glasma initial conditions with color charge fluctuations calibrated to fit coherent and incoherent $J/\psi$ diffraction measured by the H1 and Zeus experiments at HERA \cite{Aaron:2009aa, Abramowicz:2015mha} provide a good description of collectivity in small and large collision systems \cite{Schenke:2018qmc}.
+In addition, it was recently shown that hydrodynamic simulations using IP-Glasma initial conditions with color charge fluctuations calibrated to fit coherent and incoherent $J/\psi$ diffraction measured by the H1 and Zeus experiments at HERA \cite{Aaron:2009aa, Abramowicz:2015mha} provide a good description of collectivity in small and large collision systems \cite{Schenke:2018fci}.
 
@@
 We employ a multi-stage hybrid transport model that uses relativistic viscous hydrodynamics to describe the QGP medium and microscopic Boltzmann transport to simulate the dynamics of the system after hadronization \cite{Shen:2014vra, Bernhard:2016tnd}.
-The hydrodynamic initial conditions are provided by a modified version of the \trento\ model \cite{Moreland:2014oya} with additional parameters to vary the size and shape of the nucleon.
-Each initial condition profile is free streamed to the hydrodynamic starting time and matched onto the hydrodynamic energy-momentum tensor using the Landau matching procedure \cite{Broniowski:2008qk, Heinz:2015arc}.
+The hydrodynamic initial conditions are provided by a modified version of the \trento\ model \cite{Moreland:2014oya} with additional parameters to vary the number and size of hot spots inside the nucleon.
+Each initial condition profile is free-streamed to the hydrodynamic starting time and matched onto the hydrodynamic energy-momentum tensor using the Landau matching procedure \cite{Broniowski:2008qk, Heinz:2015arc}.
 Many of the components of the present model have been documented in previous studies \cite{Moreland:2014oya, Bernhard:2016tnd, Bernhard:2018hnz}; we review each component here for completeness.
@@
 and each constituent's position $\mathbf{x_i}$ in Eqn.~\eqref{eq:rho} is sampled from a Gaussian radial distribution with standard deviation $r$.
-The effect of this addition nucleon substructure on the nuclear thickness functions is visible in Fig.~\ref{fig:thickness}
+The effect of this additional nucleon substructure on the nuclear thickness functions is visible in Fig.~\ref{fig:thickness}
 
@@
 
-The aforementioned hydrodynamic equations are solved numerically using the boost-invariant \textsc{VISH2+1} viscous hydrodynamics code \cite{Song:2007ux, Shen:2014vra}.
-We simulate each hydrodynamic event on a spacetime grid with transverse extent $x_\mathrm{max}$, spatial grid step $dx$, and time step $d\tau$ which are optimized \emph{event-by-event} to balance trade-offs between numerical accuracy and computation time (see Appendix~\ref{app:adaptive_grid}).
+The aforementioned hydrodynamic equations are solved numerically using the boost-invariant VISH2+1 viscous hydrodynamics code \cite{Song:2007ux, Shen:2014vra}.
+We simulate each hydrodynamic event on a spacetime grid with transverse extent $x_\mathrm{max}$, spatial grid step $dx$, and time step $d\tau$ which are optimized event-by-event to balance trade-offs between numerical accuracy and computation time (see Appendix~\ref{app:adaptive_grid}).
 Although these details are somewhat mundane, they are critically important to the present study, since the computation time scales with the number of spacetime cells $n_x^2\, n_\tau$, and $n_x$ and $n_\tau$ typically have to be quite large to resolve the small length scales associated with nucleon substructure.
@@
 
-Once the fluid is converted into particles, we simulate its subsequent microscopic dynamics using the Ultra-relativistic Quantum Molecular Dynamics (\textsc{UrQMD}) transport model \cite{Bass:1998ca, Bleicher:1999xi}.
+Once the fluid is converted into particles, we simulate its subsequent microscopic dynamics using the Ultra-relativistic Quantum Molecular Dynamics (UrQMD) transport model \cite{Bass:1998ca, Bleicher:1999xi}.
 It solves the microscopic Boltzmann equation
@@
 
-One primary advantage of using a microscopic transport model such as \textsc{UrQMD} as an afterburner, is that it realistically simulates the system break-up when the mean free path becomes large relative to the system size.
+One primary advantage of using a microscopic transport model such as UrQMD as an afterburner, is that it realistically simulates the system break-up when the mean free path becomes large relative to the system size.
 This dilute limit is expected to play a significant role in small collision systems where the produced medium is smaller and shorter lived.
@@
 
-Subject to this assumption, we can apply Bayes' theorem to evaluate the hypothesis $H_\x$ for the true model parameters
+Subject to this assumption, we can apply Bayes' theorem to evaluate the hypothesis $H_\x$ for the true model parameters,
 \begin{equation}
@@
 \begin{itemize}[leftmargin=1\parindent]
-  \item The constituent fluctuations, modeled by the gamma random weights $\gamma_i$ in Eqn.~\eqref{eq:fluctuated_thick}, generate overall nucleon fluctuations which are suppressed by the number of constituents $n_c$ inside the nucleon.
+  \item
+    The constituent fluctuations, modeled by the gamma random weights $\gamma_i$ in Eqn.~\eqref{eq:fluctuated_thick}, generate overall nucleon fluctuations which are suppressed by the number of constituents $n_c$ inside the nucleon.
     The observed nucleon fluctuation variance falls like $1 / n_c$ which means the natural range for the constituent fluctuations is larger when the number of constituents is larger and \emph{vice versa}.
@@
     where $k$ is the shape parameter of the gamma distribution used to fluctuate each individual constituent.
-  \item In Sec.~\ref{sec:initial_state} we parametrized nucleon substructure using three degrees of freedom:
+  \item
+    In Sec.~\ref{sec:initial_state} we parametrized nucleon substructure using three degrees of freedom:
     \begin{enumerate*}[label=(\roman*)]
@@
     We thus call $w$ a nucleon ``width'', although for smaller numbers of constituents, the \emph{actual} RMS radius of our sampled nucleons can be significantly smaller than our width parameter $w$ due to fluctuations in the nucleon's center of mass, and hence one should account for the difference when discussing the nucleon's posterior RMS radius.
+  \item
+    Equation~\eqref{eq:nucleon_width} requires the nucleon width to be larger than the constituent width, i.e.
+$w > v$, lest the sampling radius $r$ turn imaginary, and thus we cannot vary the nucleon width $w$ and the constituent width $v$ independently.
+    We therefore reparametrize the constituent width $v$ using a new variable $\X$ to interpolate between minimum and maximum allowed values:
+    \begin{equation}
+      \label{eq:struct_param}
+      v = v_\mathrm{min} + \X (v_\mathrm{max} - v_\mathrm{min}).
+    \end{equation}
+    Here we choose a minimum constituent width $v_\mathrm{min} = 0.2$~fm determined by computational limits and a maximum width $v_\mathrm{max} = w$ equal to the nucleon width.
+    Thus for $\X=0$, the nucleons consist of small, distinct hot spots, whereas for $\X=1$ the nucleon is a single Gaussian blob of width $w$.
+    The parameter $\X$ thus varies the granularity of the nucleon while keeping the number of constituents $n_c$ and approximate size of the nucleon $w$ fixed.
+\end{itemize}
 
@@
   \end{tikzpicture}
-  \caption{Schematic illustrating the constituent sampling radius $r$, constituent width $v$ and nucleon width parameter $w = \sqrt{r^2 + v^2}$ for a nucleon with $n_c=3$ constituents.}
+  \caption{
     \label{fig:nucleon_schematic}
+    Schematic illustrating the constituent sampling radius $r$, constituent width $v$ and nucleon width parameter $w = \sqrt{r^2 + v^2}$ for a nucleon with $n_c=3$ constituents.
+  }
 \end{figure}
@@
 
-\item Equation~\eqref{eq:nucleon_width} requires the nucleon width to be larger than the constituent width, i.e.
-$w > v$, lest the sampling radius $r$ turn imaginary, and thus we cannot vary the nucleon width $w$ and the constituent width $v$ independently.
-  We therefore reparametrize the constituent width $v$ using a new variable $\X$ to interpolate between minimum and maximum allowed values:
-  \begin{equation}
-    \label{eq:struct_param}
-    v = v_\mathrm{min} + \X (v_\mathrm{max} - v_\mathrm{min}).
-  \end{equation}
-  Here we choose a minimum constituent width $v_\mathrm{min} = 0.2$~fm determined by computational limits and a maximum width $v_\mathrm{max} = w$ equal to the nucleon width.
-  Thus for $\X=0$, the nucleons consist of small, distinct hot spots, whereas for $\X=1$ the nucleon is a single Gaussian blob of width $w$.
-  The parameter $\X$ thus varies the granularity of the nucleon while keeping the number of constituents $n_c$ and approximate size of the nucleon $w$ fixed.
-\end{itemize}
-
 \subsection{Observables}
@@
   \item Four-particle flow cumulant $\vnk{2}{4}$ \\(charged particles, $0.2 < p_T < 5.0$~GeV).
-  \item Symmetric cumulants $\mathrm{SC}(4, 2)$ and $\mathrm{SC}(3,2)$.
+  \item Symmetric cumulants $\SC(4, 2)$ and $\SC(3,2)$.
 \end{itemize}
-Each observable is calculated from the list of final state particles produced by \textsc{UrQMD} using the same methods applied by experiment.
-We generally match the kinematic cuts of all measurements with two exceptions: we use use a larger rapidity interval $|\eta| < 0.8$ than experiment for some boost-invariant observables to improve our finite particle statistics, and we do not apply a rapidity gap, e.g.\ $|\Delta \eta| > 1$, between pairs of particles when calculating the two-particle cumulant $\vnk{n}{2}$ since we already oversample particles from each hydrodynamic event, and this oversampling suppresses non-flow correlations.
+Each observable is calculated from the list of final state particles produced by UrQMD using the same methods applied by experiment.
+We generally match the kinematic cuts of all measurements with two exceptions: we use a larger rapidity interval $|\eta| < 0.8$ than experiment for some boost-invariant observables to improve our finite particle statistics, and we do not apply a rapidity gap, e.g.\ $|\Delta \eta| > 1$, between pairs of particles when calculating the two-particle cumulant $\vnk{n}{2}$ since we already oversample particles from each hydrodynamic event, and this oversampling suppresses non-flow correlations.
 
@@
   \includegraphics{correlation_matrices}
-  \caption{Visualization of the Pb-Pb correlation matrix $\mathrm{corr}(y_i, y_j) = \mathrm{cov}(y_i,y_j)/(\sigma_i \sigma_j)$ for the model (emulator) at a random point in parameter space (left-side) and for the experimental data (right-side).
-Each cell represents an observable in a single centrality bin.
-Experimental statistical and systematic errors are from ALICE \cite{Adam:2015ptt, Adam:2016izf}.
-The experimental correlation structure is modeled using Eqn.~\eqref{eq:corr}.}
+  \caption{
     \label{fig:correlation}
+    Visualization of the Pb-Pb correlation matrix $\corr(y_i, y_j) = \cov(y_i,y_j)/(\sigma_i \sigma_j)$ for the model (emulator) at a random point in parameter space (left-side) and for the experimental data (right-side).
+    Each cell represents an observable in a single centrality bin. Experimental statistical and systematic errors are from ALICE \cite{Adam:2015ptt, Adam:2016izf}.
+    The experimental correlation structure is modeled using Eqn.~\eqref{eq:corr}.
+  }
 \end{figure*}
@@
 \begin{equation}
-  \Sigma_e^\text{stat} = \text{diag}[(\sigma^\text{stat}_1)^2, (\sigma^\text{stat}_2)^2, \dots (\sigma^\text{stat}_m)^2 ],
-\end{equation}
-where $\sigma^\text{stat}_i$ is the statistical uncertainty of observable $y_i$ in the experimental observable vector $\y_e = [y_1, \dots, y_m]$.
+  \Sigma_e^\text{stat} = \diag[(\sigma^\text{stat}_1)^2, (\sigma^\text{stat}_2)^2, \dots (\sigma^\text{stat}_m)^2 ],
+\end{equation}
+where $\sigma^\text{stat}_i$ is the statistical uncertainty of observable $y_i$ in the experimental observable vector $\y_e = (y_1, \dots, y_m)$.
 The systematic errors, meanwhile, are typically correlated, but the correlation structure is not reported by the experiments so we assert a reasonable form.
@@
 \begin{equation}
-  \rho_{ij} = \frac{\text{cov}(y_i, y_j)}{\sigma_i \sigma_j},
+  \rho_{ij} = \frac{\cov(y_i, y_j)}{\sigma_i \sigma_j},
 \end{equation}
@@
 \begin{equation}
-  \mathrm{corr}(y_i, y_j) = \mathrm{cov}(y_i, y_j)/(\sigma_i \sigma_j)
+  \mathrm{corr}(y_i, y_j) = \cov(y_i, y_j)/(\sigma_i \sigma_j)
 \end{equation}
@@
 \begin{equation}
-\mathrm{cov}(z_i, z_j) = \mathrm{diag}[(\delta z_{1})^2, (\delta z_2)^2, \dots, (\delta z_k)^2 ],
+  \cov(z_i, z_j) = \diag[(\delta z_{1})^2, (\delta z_2)^2, \dots, (\delta z_k)^2 ],
 \end{equation}
@@
   {\centering\includegraphics{observables_pbpb}\par}
-\captionof{figure}{\label{fig:obs_pbpb} Simulated observables compared to experimental data for Pb-Pb collisions at $\sqrts=5.02$~TeV.
+  \captionof{figure}{
+    \label{fig:obs_pbpb}
+    Simulated observables compared to experimental data for Pb-Pb collisions at $\sqrts=5.02$~TeV.
     Top row: explicit model calculations (no emulator) for each of the $d=500$ design points; bottom row: emulator predictions for $n=100$ random samples drawn from the posterior.
-Points with error bars are experimental data from ALICE with statistical and systematic errors added in quadrature \cite{Adam:2015ptt, Adam:2016izf}.}
+    Points with error bars are experimental data from ALICE with statistical and systematic errors added in quadrature \cite{Adam:2015ptt, Adam:2016izf}.
+  }
   \vspace{.5cm}
   {\centering\includegraphics{observables_ppb}\par}
-\captionof{figure}{\label{fig:obs_ppb} Same as Fig.~\ref{fig:obs_pbpb} but for $p$-Pb collisions at $\sqrts=5.02$~TeV.
+  \captionof{figure}{
+    \label{fig:obs_ppb}
+    Same as Fig.~\ref{fig:obs_pbpb} but for $p$-Pb collisions at $\sqrts=5.02$~TeV.
     Note that multiplicity bins are used for mean $p_T$ and flow cumulant observables to match the bins used by experiment.
-Experimental data are from ALICE \cite{Adam:2014qja, Abelev:2013bla} and CMS \cite{Chatrchyan:2013nka}.}
+    Experimental data are from ALICE \cite{Adam:2014qja, Abelev:2013bla} and CMS \cite{Chatrchyan:2013nka}.
+  }
 \end{fullpage}
@@
   \end{figure}
-\captionof{figure}{\label{fig:posterior} Bayesian posterior distribution of the model input parameters.
+  \captionof{figure}{
+    \label{fig:posterior}
+    Bayesian posterior distribution of the model input parameters.
     The diagonal panels show the marginalized distributions of individual model parameters, while off-diagonal panels show the joint distributions for pairs of model parameters, visualizing their correlations.
-The marginalized distribution medians and 90\% credible intervals are annotated along the diagonal.}
+    The marginalized distribution medians and 90\% credible intervals are annotated along the diagonal.
+  }
 \end{fullpage}
@@
 Moreover, their posterior values nicely describe the $p$-Pb and Pb-Pb calibration observables in Figs.~\ref{fig:obs_pbpb} and \ref{fig:obs_ppb}.
-While it would not be surprising, for example, to fit one or two of these observables using such a narrow range of values, the quality of the combined fit and the number of observables described is highly non-trivial.
+While it would not be surprising, for example, to fit one or two of these observables using such a narrow range of values, the quality of the combined fit (more on this later) and the number of observables described is highly non-trivial.
 For example, consider the ratio of the $p$-Pb yield over the Pb-Pb yield, which imposes a strong constraint on physically reasonable initial condition models.
@@
 It just so happens that the small range of $p$-values needed to describe this yield ratio are the same values needed to describe all of the calibration observables in the present study, \emph{and} numerous Pb-Pb observables at $\sqrts=2.76$~TeV \cite{Bernhard:2018hnz}.
-This work thus reaffirms an empirical scaling law reported in several previous studies \cite{Moreland:2014oya, Bernhard:2016tnd, Ke:2016jrd, Bernhard:2018hnz}, that entropy deposition at midrapidity in ultrarelativistic nuclear collisions scales approximately as
+This work thus reaffirms an empirical scaling law reported in several previous studies \cite{Moreland:2014oya, Bernhard:2016tnd, Ke:2016jrd, Bernhard:2018hnz},
+that ultrarelativistic nuclear collisions deposit entropy at midrapidity approximated by
 \begin{equation}
@@
   \includegraphics{posterior_parton_number}
-  \caption{\label{fig:posterior_parton_number}
+  \caption{
+    \label{fig:posterior_parton_number}
     Posterior distribution for the number of nucleon constituents $n_c$ determined by the analysis.
@@
   \includegraphics{posterior_proton_shape}
-  \caption{\label{fig:posterior_proton_shape}
+  \caption{
+    \label{fig:posterior_proton_shape}
     Posterior distribution for the nucleon width $w$ and constituent width $v$.
     The enclosed trapezoidal region covers the prior range of allowed values for $w$ and $v$.
-The posterior distribution, shown in blue, indicates the preferred values for $w$ and $v$ determined by the analysis.}
+    The posterior distribution, shown in blue, indicates the preferred values for $w$ and $v$ determined by the analysis.
+  }
 \end{figure}
@@
 Figure~\ref{fig:posterior_proton_shape} shows the resulting joint posterior distribution for the nucleon width parameter $w$ and constituent width $v$.
-The white trapezoidal region in the lower-right is the joint parameter prior, and the gray triangular region in the upper-left where $v > w$ is a region that we exclude from the prior.
 The posterior distribution, shown in blue, is remarkably well constrained.
-We report a posterior estimate for the constituent width $v=0.46_{-0.15}^{+0.21}$~fm which is roughly the same magnitude as the \emph{nucleon} hot-spots used in recent saturation-based substructure studies that employed IP-Glasma initial conditions \note.
+We report a posterior estimate for the constituent width $v=0.46_{-0.15}^{+0.21}$~fm which is roughly the same magnitude as the \emph{nucleon} hot-spots used in recent saturation-based substructure studies that employed IP-Glasma initial conditions \cite{Schenke:2018fci}.
 Evidently, it may be necessary to place an informative prior on our nucleon substructure parameters in order to resolve the apparent tension between our parameter values and those needed to describe DIS measurements at HERA.
@@
   \includegraphics{region_shear_bulk}
-  \caption{Left figure: estimated temperature dependence of the QGP specific shear viscosity $(\eta/s)(T)$ determined by the present Bayesian analysis of $p$-Pb and Pb-Pb collisions at $\sqrts=5.02$~TeV (orange line/band) compared to a previous Bayesian analysis of Pb-Pb collisions at $\sqrts=2.76$ and 5.02~TeV (blue line/band) \cite{Bernhard:2018hnz}.
+  \caption{
+    \label{fig:region_shear_bulk}
+    Left figure: estimated temperature dependence of the QGP specific shear viscosity $(\eta/s)(T)$ determined by the present Bayesian analysis of $p$-Pb and Pb-Pb collisions at $\sqrts=5.02$~TeV (orange line/band) compared to a previous Bayesian analysis of Pb-Pb collisions at $\sqrts=2.76$ and 5.02~TeV (blue line/band) \cite{Bernhard:2018hnz}.
     The lines are the medians of each posterior distribution, and the bands are their 90\% credible regions.
-Right figure: same as before, but for the temperature dependence of the QGP specific bulk viscosity $(\zeta/s)(T)$.}
-  \label{fig:region_shear_bulk}
+    Right figure: same as before, but for the temperature dependence of the QGP specific bulk viscosity $(\zeta/s)(T)$.
+  }
 \end{figure*}
@@
 In this section, we compare several of our posterior estimates to those obtained from a similar Bayesian analysis \cite{Bernhard:2018hnz} which used an (almost) identical version of the present physics model.
-The only modeling difference is the inclusion of nucleon substructure in this work which was absent in Ref.~\cite{Bernhard:2018hnz}.
-Several calibration details, however, are different between the two studies.
+The only modeling difference between the two studies is the inclusion of nucleon substructure in this work which was absent in Ref.~\cite{Bernhard:2018hnz}.
+Several calibration details, however, are different as well.
 The present work uses a much smaller number of $p$-Pb and Pb-Pb observables at $\sqrts=5.02$~TeV (limited by availability), whereas Ref.~\cite{Bernhard:2018hnz} used a larger cross section of Pb-Pb experimental data at $\sqrts=2.76$ and 5.02~TeV.
@@
 \begin{equation}
+  \label{eq:mean_pT_corr}
   (\delta p_T)^2 = \langle \langle (p_{T,i} - \langle p_T \rangle) (p_{T,j} - \langle p_T \rangle) \rangle \rangle.
-  \label{eq:mean_pT_corr}
 \end{equation}
@@
   {\centering\includegraphics{observables_map}\par}
-\captionof{figure}{\label{fig:obs_map} Model calculations using the \emph{maximum a posterior} (MAP) parameters compared to experiment.
+  \captionof{figure}{
+    \label{fig:obs_map}
+    Model calculations using the \emph{maximum a posterior} (MAP) parameters compared to experiment.
     Colored lines are model calculations for $p$-Pb collisions (left) and Pb-Pb collisions (right) at $\sqrts=5.02$~TeV.
     Points with error bars are the experimental data with statistical uncertainties, and gray bands their corresponding systematic uncertainties, from CMS \cite{Chatrchyan:2013nka} and ALICE \cite{Adam:2015ptt, Adam:2016izf, Adam:2014qja, Abelev:2013bla}.
-The sub-axes show the ratio of model over data where available with gray bands indicating $\pm 10\%$.}
+    The sub-axes show the ratio of model over data where available with gray bands indicating $\pm 10\%$.
+  }
 \end{fullpage}
@@
 
-Lastly, we compute the symmetric cumulants $\mathrm{SC}(m,n)$ for the Pb-Pb collision system at $\sqrts=5.02$~TeV which quantify correlations between event-by-event fluctuations of the flow harmonics of different order \cite{Bilandzic:2013kga, ALICE:2016kpq},
+Lastly, we compute the symmetric cumulants $\SC(m,n)$ for the Pb-Pb collision system at $\sqrts=5.02$~TeV which quantify correlations between event-by-event fluctuations of the flow harmonics of different order \cite{Bilandzic:2013kga, ALICE:2016kpq},
 \begin{align}
-  \mathrm{SC}(m, n) &= \langle\langle \cos[m(\phi_1 - \phi_3) + n(\phi_2-\phi_4)]\rangle\rangle \nonumber \\
+  \SC(m, n) &= \langle\langle \cos[m(\phi_1 - \phi_3) + n(\phi_2-\phi_4)]\rangle\rangle \nonumber \\
   \nonumber &- \langle\langle\cos[m(\phi_1-\phi_2)]\rangle\rangle\langle\langle\cos[n(\phi_1-\phi_2)]\rangle\rangle \label{eq:scmn}\\
@@
 \begin{equation}
-  \mathrm{NSC}(m,n) = SC(m,n)/\langle v_m^2\rangle\langle v_n^2\rangle,
+  \NSC(m,n) = \SC(m,n)/\langle v_m^2\rangle\langle v_n^2\rangle,
 \end{equation}
@@
 The solid lines are the MAP estimate of the present study and the dashed lines are the MAP estimate of Ref.~\cite{Bernhard:2018hnz} which did not include nucleon substructure and was calibrated on Pb-Pb collisions at $\sqrts=2.76$ and 5.02~TeV.
-We observe that the gap between $\mathrm{SC}(3,2)$ and $\mathrm{SC}(4,2)$ is generally wider in the present analysis than in Ref.~\cite{Bernhard:2018hnz}, as is the gap between the normalized symmetric cumulants $\mathrm{NSC}(3,2)$ and $\mathrm{NSC}(4,2)$.
+We observe that the gap between $\SC(3,2)$ and $\SC(4,2)$ is generally wider in the present analysis than in Ref.~\cite{Bernhard:2018hnz}, as is the gap between the normalized symmetric cumulants $\NSC(3,2)$ and $\NSC(4,2)$.
 
@@
   \includegraphics{flow_corr}
-  \caption{Model calculations of the symmetric cumulants (top) and normalized symmetric cumulants (bottom) for Pb-Pb collisions at $\sqrts=5.02$~TeV using the \emph{maximum a posteriori} (MAP) parameters.
+  \caption{
+    \label{fig:flow_corr}
+    Model calculations of the symmetric cumulants (top) and normalized symmetric cumulants (bottom) for Pb-Pb collisions at $\sqrts=5.02$~TeV using the \emph{maximum a posteriori} (MAP) parameters.
     The solid lines are the MAP estimate of the present analysis (with nucleon substructure), and the dashed lines are the MAP estimate of Ref.~\cite{Bernhard:2018hnz} (without nucleon substructure) which was calibrated on Pb-Pb observables at $\sqrts=2.76$ and 5.02~TeV.
-In general, most model parameters are somewhat different between the two studies.}
-  \label{fig:flow_corr}
+    In general, most model parameters are somewhat different between the two studies.
+  }
 \end{figure}
 
-\section{summary and conclusions}
+\section{Summary and conclusions}
 \label{sec:summary}
 
-Relativistic heavy-ion collisions produce long-range multiparticle correlations which are commonly explained by the existence of hydrodynamic flow \note.
-This narrative is evidenced by the global, self-consistent and highly non-trivial quantitative agreement of hydrodynamic models with a large number of heavy-ion bulk observables \note.
+Relativistic heavy-ion collisions produce long-range multiparticle correlations which are commonly explained by the existence of hydrodynamic flow \cite{deSouza:2015ena}.
+This narrative is evidenced by the global, self-consistent and highly non-trivial quantitative agreement of hydrodynamic models with a large number of heavy-ion bulk observables \cite{Niemi:2015qia, Bernhard:2016tnd, Gale:2012rq}.
 Naturally, such descriptions rely on the validity of hydrodynamic approximations, and these approximations begin to break down in the so-called dilute limit where discrete particle degrees-of-freedom dominate and continuous field descriptions of the medium cease to make sense.
@@
 
-These expectations were upended, however, when long-range multiparticle correlations were detected in high-multiplicity $p$-Pb collisions and found to be similar in magnitude to those observed in Pb-Pb collisions \note.
+These expectations were upended, however, when long-range multiparticle correlations were detected in high-multiplicity $p$-Pb collisions and found to be similar in magnitude to those observed in Pb-Pb collisions \cite{CMS:2012qk, Abelev:2012ola, Aad:2012gla}.
 Nuclear collision systems which were previously thought to be too small for hydrodynamic flow, were subsequently found to generate the same collectivity used to justify hydrodynamic flow in heavy-ion collisions.
@@
 We built, for this purpose, a flexible multi-stage nuclear collision model characterized by a number of free parameters which vary theoretically uncertain aspects of the framework such as the QGP initial conditions and hydrodynamic transport properties.
-For the QGP initial conditions, we employed a modified version of the \trento\ model which adds new parameters to vary the fluctuating size and shape of each nucleon.
+For the QGP initial conditions, we employed a modified version of the \trento\ model \cite{Moreland:2014oya} which adds new parameters to vary the fluctuating size and shape of each nucleon.
 Specifically, we modeled each nucleon as a cluster of $n_c$ constituents (hot spots), where each constituent is described by a Gaussian density profile of width $v$.
@@
 
-With the full evolution model in hand, we applied Bayesian methods which were developed to estimate the parameters of computationally intensive models \note.
+With the full evolution model in hand, we applied Bayesian methods which were developed to estimate the parameters of computationally intensive models \cite{OHagan:2006ba, Higdon:2008cmc, Higdon:2014tva}.
 We first constructed a scaffolding of $n=500$ parameters points distributed throughout our 15-dimensional parameter space and evaluated the nuclear collision model using \order{4} events at each parameter point.
@@
 Finally, we used Markov chain Monte Carlo (MCMC) importance sampling to explore the parameter space and draw samples from the Bayesian posterior distribution for the \emph{true} values of the model parameters, given our emulated model predictions, the experimental data, and their associated uncertainties.
-
 The model calibration process is summarized by Figs.~\ref{fig:obs_pbpb} and \ref{fig:obs_ppb}, and the resulting posterior distribution for the model input parameters is shown in Fig.~\ref{fig:posterior}.
@@
 First, we demonstrated in Fig.~\ref{fig:obs_map} the existence of a single set of model parameters which can simultaneously describe $p$-Pb and Pb-Pb charged-particle yields, mean $p_T$, and flow cumulants at $\sqrts=5.02$~TeV.
-The excellent quantitative agreement of the model calculation with the experimental data is strong evidence for a unified hydrodynamic description of $p$-A and A-A collisions.
+The excellent quantitative agreement of the model calculation with the experimental data is strong evidence for a unified hydrodynamic description of $p$-A and A-A collisions at ultrarelativistic energies.
 Moreover, the modifications to the physics model which were required to obtain this agreement are generally modest; one must simply replace Gaussian nucleons with composite nucleons of several or more constituents.
 
-Second, we obtained a posterior distribution for the model input parameters in Fig.~\ref{fig:posterior} and reported quantitative estimates for the model input parameters in Tbl.~\ref{tab:post_param}.
+Second, we obtained a posterior distribution for the model input parameters in Fig.~\ref{fig:posterior} and reported quantitative estimates for these parameters in Tbl.~\ref{tab:post_param}.
 Here we summarize our key findings about the model:
@@
   \item
-    Using the \trento\ initial condition model, we find that initial entropy deposition (or particle production) scales like the geometric mean of local participant nuclear densities, in agreement with a number of previous estimates \note.
+    Using \trento\ initial conditions, we find that initial entropy deposition (or particle production) scales like the geometric mean of local participant nuclear densities, in agreement with a number of previous estimates \cite{Moreland:2014oya, Bernhard:2016tnd, Bernhard:2018hnz}.
   \item
@@
     Bulk observables in $p$-Pb and Pb-Pb collisions prefer larger nucleons, similar in size to the proton's RMS electric charge radius.
-    This agrees with a similar Bayesian analysis which was calibrated to Pb-Pb bulk observables at $\sqrts=2.76$ and 5.02~TeV using a hybrid model without nucleon substructure \note.
-    We note that our result is in tension an estimate for the effective nucleon width based on the gluon distribution extracted from HERA data \note.
+    This agrees with a similar Bayesian analysis which was calibrated to Pb-Pb bulk observables at $\sqrts=2.76$ and 5.02~TeV using a hybrid model without nucleon substructure \cite{Bernhard:2018hnz}.
+    We also note that our result is in significant tension with an estimate for the effective nucleon width based on the gluon distribution extracted from HERA data \cite{Rezaeian:2012ji}.
   \item
-    We obtain an estimate $v=0.46_{-0.15}^{+0.21}$~fm for the Gaussian width of individual constituent hot spots inside the nucleon is which is significantly larger than the length scales typically associated with nucleon substructure.
-    This result may help explain the seemingly unreasonable success of hydrodynamics in small collision systems.
+    We obtain an estimate $v=0.46_{-0.15}^{+0.21}$~fm for the Gaussian width of the constituent hot spots inside each nucleon.
+    This is significantly larger than the length scales typically associated with nucleon substructure, and it may help explain the seemingly unreasonable success of hydrodynamics in small collision systems.
   \item
@@
   \item
-    The posterior estimates for the temperature dependence of the QGP specific shear and bulk viscosities
+    We compare in Fig.~\ref{fig:region_shear_bulk} our estimate for the temperature dependence of the QGP specific shear and bulk viscosities to those of Ref.~\cite{Bernhard:2018hnz}, which performed a Bayesian calibration to Pb-Pb bulk observables at $\sqrts=2.76$ and 5.02~TeV using a physics model without nucleon substructure.
+    The two studies are in good agreement, although Ref.~\cite{Bernhard:2018hnz} obtains a more precise estimate for $(\eta/s)(T)$, likely due to the additional beam energies and observables included, and the enhanced sensitivity of larger collision systems to the QGP viscosity.
+  \item
+    We make predictions in Fig.~\ref{fig:obs_map} for several quantities which were not included in the model calibration, including the identified yields, transverse energy, symmetric cumulants, and mean $p_T$ fluctuations at $\sqrts=5.02$~TeV.
+    Interestingly, our MAP estimate for the Pb-Pb symmetric cumulants at $\sqrts=5.02$~TeV are significantly different than those estimated in Ref.~\cite{Bernhard:2018hnz}.
+    This could be a direct or indirect result of including nucleon substructure in the model calibration.
 \end{enumerate}
@@
 The present study would benefit from a number of improvements.
-Perhaps the most obvious improvement would be to add missing experimental data to the calibration once it becomes available.
-For example, the present calibration is missing the Pb-Pb mean $p_T$, as well as the transverse energy, identified particle yields, and mean $p_T$ fluctuations of both collision systems.
-These should certainly influence the quality of the combined fit and the posterior estimates for the model parameters.
-
-Similarly, the analysis would greatly benefit from additional beam energies and collision systems.
-Notably missing are measurements of $p$-Au, $d$-Au and $^3$He-Au collisions at RHIC \note, which would provide an additional constraint on the parameters and more stringent tests of the calibrated model predictions.
+Perhaps the most obvious target for improvement is the absence of several important experimental datasets.
+Specifically, we are missing the transverse energy, identified particle yields, and the mean $p_T$ fluctuations of both collision systems, as well the charged-particle mean $p_T$ for the Pb-Pb system.
+These observables would certainly influence the quality of the combined fit and correspondingly our estimates for the posterior model parameters.
+
+Similarly, the results would greatly benefit from additional beam energies and collision systems.
+Notably absent is data from RHIC which includes measurement for $p$-Au, $d$-Au, $^3$He-Au, Cu-Cu, Cu-Au, Au-Au and U-U collision systems at $\sqrts=200$~GeV.
+This data would provide an additional constraint on the model parameters and would enable more stringent tests of the calibrated model predictions.
 The RHIC data may also help elucidate the beam-energy dependence of the model parameters which would be worth investigating.
-We leave these improvements for future work.
+We leave these improvements for future studies.
+
+\medskip
 
@@
 \begin{itemize}[leftmargin=2\parindent, itemsep=0pt, topsep=5pt]
-  \item \trento\ initial conditions \note
-  \item Pre-equilibrium free-streaming \note
-  \item VISH2+1 hydrodynamics \note
-  \item FRZOUT particle sample \note
-  \item UrQMD microscopic transport model \note
-  \item DukeQCD event generator (wrapper) \note
+  \item \trento\ with nucleon substructure (C++) \cite{trento:code}
+  \item Pre-equilibrium free-streaming (Python) \cite{freestream:code}
+  \item VISH2+1 hydrodynamics (Fortran) \cite{osuhydro:code}
+  \item FRZOUT particle sample (Python) \cite{frzout:code}
+  \item UrQMD microscopic transport model (Fortran) \cite{urqmd:code}
+  \item DukeQCD event generator wrapper (Python) \cite{eventgen:code}
+  \item Bayesian parameter estimation (Python) \cite{bayesian:code}
 \end{itemize}
+
+\medskip
 
@@
 DE-AC02-05CH11231.
-  JSM, JEB and SAB are supported by the U.S.\ Department of Energy Grant No.\ DE-FG02-05ER41367 and National Science Foundation Grant No.\ NSF-ACI-1550225.
+  JSM and SAB are supported by the U.S.\ Department of Energy Grant No.\ DE-FG02-05ER41367 and JEB by NSF Grant No.\ NSF-ACI-1550225.
 \end{acknowledgments}
@@
 \bibliography{substructure}
+
 \appendix
@@
 
-The boost-invariant \textsc{VISH2+1} hydrodynamics code used in this work \cite{Song:2007ux, Shen:2014vra} runs on a Cartesian transverse grid specified by a maximum grid size $x_\text{max}$ and grid step width $dx$ which fix the transverse grid extent ${-x_\text{max} < x < x_\text{max}}$ and number of grid cells along each dimension ${n_x = 2\, x_\text{max} / dx}$.
-In general, the maximum grid size $x_\text{max}$ should be set large enough to contain the full spacetime evolution of the event.
-This means that the truncation of $T^{\mu\nu}$ at the boundaries of the grid should never modify the final-state observables.
-We enforce this requirement by finding an energy density cutoff $e_\text{min}$ for which the matter $e < e_\text{min}$ can be effectively discarded without significantly modifying the simulation observables.
-We then fix the maximum grid size $x_\text{max}$ such that it fully encloses the isotherm $T=T(e_\text{min})$ for the full lifetime of the fireball.
-
-We find that we can quickly estimate the maximum radius $R_\text{max} = |\x_\text{max}|^2$ of the spacetime hypersurface ${T = T(e_\text{min})}$ by running the event on a coarse-grained spatial grid with one-third the spatial resolution we would otherwise require to resolve typical hydrodynamic observables such as mean $p_T$ and flows.
-The simulation time of a single \textsc{VISH2+1} event scales like ${\sim}n_x^3$ since $dx \propto d\tau$, and thus our ``pre-run'' event requires only ${\sim}1/27$th the time of a production event.
-We therefore run a coarse-grained pre-event on an excessively large grid for \emph{every} minimum-bias event to estimate $R_\text{max}$, then rerun the same event on a thrice finer grid with a trimmed spatial extent $x_\text{max} \equiv R_\text{max}$.
-See Fig.~\ref{fig:adaptive_grid} for a simple diagram of the procedure.
-
-In practice, we find that event-by-event grid resizing leads to a massive speed increase for minimum bias events compared to using a single fixed grid for the entire minimum bias sample.
-This is because the maximum transverse size of each event varies dramatically, from a few fm in peripheral Pb-Pb collisions to 50 fm or more in central Pb-Pb collisions.
-The procedure should generalize to other hydrodynamic codes, including those with $3+1$ spacetime dimensions, where the time savings could be even more pronounced.
-
-\begin{figure}[t]
+\begin{figure}
   \begin{tikzpicture}
@@
   \end{tikzpicture}
-  \caption{Diagram of the adaptive grid resizing algorithm (not drawn to scale).
+  \caption{
+    \label{fig:adaptive_grid}
+    Diagram of the adaptive grid resizing algorithm (not drawn to scale).
     Each initial condition event is first run on a very large coarse-grained mesh (large gray grid) of one-third the spatial resolution otherwise required to measure hydrodynamic observables.
     We then measure the maximum transverse radius $R_\text{max}$ (blue circle) of the hypersurface defined by the temperature isotherm $T = T(e_\text{min})$, where $e_\text{min}$ is the largest energy density which can be truncated without modifying the hydrodynamic observables calculated from the event.
-Finally, the initial condition event is rerun on a smaller and finer mesh (smaller black grid) with three-times the cell density of the pre-run event and a smaller transverse extent $-R_\text{max} < x < R_\text{max}$.}
-  \label{fig:adaptive_grid}
+    Finally, the initial condition event is rerun on a smaller and finer mesh (smaller black grid) with three-times the cell density of the pre-run event and a smaller transverse extent $-R_\text{max} < x < R_\text{max}$.
+  }
 \end{figure}
@@
   \includegraphics{validation_example}
-  \caption{Example emulator validation for one observable, the Pb-Pb charged particle yield $d\nch/d\eta$ in the 20-30\% centrality class.
+  \caption{
+    \label{fig:validation_example}
+    Example emulator validation for one observable, the Pb-Pb charged particle yield $d\nch/d\eta$ in the 20-30\% centrality class.
     We use the k-fold cross validation method (explained in the text) to partition the model inputs $X$ and outputs $Y$ into training and validation data.
@@
   }
-  \label{fig:validation_example}
 \end{figure*}
+
+The boost-invariant VISH2+1 hydrodynamics code used in this work \cite{Song:2007ux, Shen:2014vra} runs on a Cartesian transverse grid specified by a maximum grid size $x_\text{max}$ and grid step width $dx$ which fix the transverse grid extent ${-x_\text{max} < x < x_\text{max}}$ and number of grid cells along each dimension ${n_x = 2\, x_\text{max} / dx}$.
+In general, the maximum grid size $x_\text{max}$ should be set large enough to contain the full spacetime evolution of the event.
+This means that the truncation of $T^{\mu\nu}$ at the boundaries of the grid should never modify the final-state observables.
+We enforce this requirement by finding an energy density cutoff $e_\text{min}$ for which the matter $e < e_\text{min}$ can be effectively discarded without significantly modifying the simulation observables.
+We then fix the maximum grid size $x_\text{max}$ such that it fully encloses the isotherm $T=T(e_\text{min})$ for the full lifetime of the fireball.
+
+We find that we can quickly estimate the maximum radius $R_\text{max} = |\x_\text{max}|^2$ of the spacetime hypersurface ${T = T(e_\text{min})}$ by running the event on a coarse-grained spatial grid with one-third the spatial resolution we would otherwise require to resolve typical hydrodynamic observables such as mean $p_T$ and flows.
+The simulation time of a single VISH2+1 event scales like ${\sim}n_x^3$ since $dx \propto d\tau$, and thus our ``pre-run'' event requires only ${\sim}1/27$th the time of a production event.
+We therefore run a coarse-grained pre-event on an excessively large grid for \emph{every} minimum-bias event to estimate $R_\text{max}$, then rerun the same event on a thrice finer grid with a trimmed spatial extent $x_\text{max} \equiv R_\text{max}$.
+See Fig.~\ref{fig:adaptive_grid} for a simple diagram of the procedure.
+
+In practice, we find that event-by-event grid resizing leads to a massive speed increase for minimum bias events compared to using a single fixed grid for the entire minimum bias sample.
+This is because the maximum transverse size of each event varies dramatically, from a few fm in peripheral Pb-Pb collisions to 50 fm or more in central Pb-Pb collisions.
+The procedure should generalize to other hydrodynamic codes, including those with $3+1$ spacetime dimensions, where the time savings could be even more pronounced.
 
@@
 More generally, we can perform the validation test in Fig.~\ref{fig:validation_example} for \emph{every} observable $y \in \y_m$ and check that each observable's normalized residuals ${z=(y_\text{pred} - y_\text{obs})/\sigma_\text{pred}}$ follow a unit normal distribution.
-This test is applied to the Pb-Pb system in Fig.~\ref{fig:validation_PbPb5020} and the $p$-Pb system in Fig.~\ref{fig:validation_pPb5020}.
+This test is applied to the $p$-Pb and Pb-Pb collision systems in Fig.~\ref{fig:validation_all}.
 The top row of each figure shows a box-plot for the normalized residuals of each observable compared to the quantiles of a unit normal distribution.
@@
 
-We also show in Fig.~\ref{fig:validation_PbPb5020} and Fig.~\ref{fig:validation_pPb5020} an estimate of the emulator error magnitude.
+We also show in Fig.~\ref{fig:validation_all} an estimate of the emulator error magnitude.
 This error is expressed in terms of the unitless variable
@@
 
-\begin{fullpage}
 \begin{figure*}
-  \includegraphics{validation_PbPb5020}
+  \includegraphics{validation_PbPb5020}\\
   \includegraphics{validation_pPb5020}
-  \caption{\label{fig:val_pbpb} Emulator validation for the Pb-Pb collision system (top) and $p$-Pb collision system (bottom) at $\sqrts=5.02$~TeV.
+  \caption{
+    \label{fig:validation_all}
+    Emulator validation for the Pb-Pb collision system (top) and $p$-Pb collision system (bottom) at $\sqrts=5.02$~TeV.
     The ``piano keys'' in the top row of each figure are horizontally stacked box plots for the normalized residuals of each model observable.
-The boxes are 50\% interquartile ranges and whiskers are the 90\% interquantiles.}
+    The boxes are 50\% interquartile ranges and whiskers are the 90\% interquantiles.
+  }
 \end{figure*}
-\end{fullpage}
 

### 2018-08-02 src/plots.py

@@ -99,7 +99,7 @@ plt.rcParams.update({
     'lines.markersize': 3,
     'lines.markeredgewidth': 0,
     'patch.linewidth': .8,
-    'hatch.linewidth': .6,
+    'hatch.linewidth': .8,
     'axes.linewidth': .6,
     'xtick.major.width': .6,
     'ytick.major.width': .6,
@@ -1847,7 +1847,7 @@ def validation_all(system):
     ax_rms.set_xticks([])
     ax_rms.set_yticks(np.arange(0, 21, 5))
     ax_rms.set_ylim(0, 20)
-    ax_rms.set_ylabel('Fractional error')
+    ax_rms.set_ylabel('Frac. error')
 
     for y in ax_rms.get_yticks():
         ax_rms.axhline(y, color='.5', zorder=-10)
@@ -1881,7 +1881,7 @@ def validation_example(
 
     """
     fig, axes = plt.subplots(
-        ncols=2, figsize=figsize(1, aspect=.6),
+        ncols=2, figsize=figsize(.8, aspect=.6),
         gridspec_kw=dict(width_ratios=[3, 1])
     )
 
@@ -1975,6 +1975,8 @@ def validation_example(
         fontdict=dict(rotation=-90),
         labelpad=4*plt.rcParams['axes.labelpad']
     )
+
+    set_tight(fig, rect=[.02, 0, 1, 1])
 
 
 @plot

### 2018-08-03 paper/substructure.tex
Some small edits to the paper

@@
 \usepackage{tikz}
+\usepackage{subcaption}
 
@@
 The \trento\ model operates in the ultrarelativistic limit with a Lorentz factor $\gamma \gg 1$ such that each nucleus appears as a thin sheet of nuclear density in the laboratory frame.
-The sheets of colliding nuclear density penetrate and pass through each other in time $\Delta t = D_\text{nucl} / \sqrt{\gamma^2 - 1}$ in the laboratory frame, where $D_\text{nucl}$ is the diameter of the nucleus in its rest frame and $\gamma$ is the usual Lorentz factor of the accelerated ions.
-The resulting nuclear overlap time $\Delta t \lesssim 0.1$~fm/$c$ at top RHIC and LHC energies is thus sufficiently short to neglect initial transverse dynamics which occur while the nuclei pass through each other.
-We therefore assume that the collision produces all secondary particles at uniform proper time $\tau = 0^+$~fm/$c$, and that it deposits entropy (energy) at midrapidity which is a function of the locally varying beam integrated density of each nucleus.
+The sheets of colliding nuclear density penetrate and pass through each other in proper time $\Delta \tau \approx D_\text{nucl} / (\gamma\, \beta_z)$ in the laboratory frame, where $D_\text{nucl}$ is the diameter of the nucleus in its rest frame, $\gamma$ is the usual Lorentz factor of the accelerated ions, and $\beta_z$ is their velocity along the beam axis.
+The resulting nuclear overlap time $\Delta \tau \lesssim 0.1\ \fmc$ at top RHIC and LHC energies is thus sufficiently short to neglect initial transverse dynamics which occur while the nuclei pass through each other.
+We therefore assume that the collision produces all secondary particles at uniform proper time $\tau = 0^+\ \fmc$, and that it deposits entropy (energy) at midrapidity which is a function of the locally varying beam integrated density of each nucleus.
 
@@
 
-Within the eikonal approximation, the initial entropy deposited at midrapidity at proper time $\tau=0^+$~fm/$c$ can be represented by some mapping ${f(\T_A, \T_B) \mapsto dS/dy}$.
+Within the eikonal approximation, the initial entropy deposited at midrapidity at proper time $\tau=0^+\ \fmc$ can be represented by some mapping ${f(\T_A, \T_B) \mapsto dS/dy}$.
 A natural first guess for the mapping is the arithmetic mean
@@
 We therefore choose to model the QGP's initial off-equilibrium dynamics using a simple step-function approximation, depicted in Fig.~\ref{fig:coupling}, which free-streams the initial state for proper time $\taufs$ (zero coupling) before instantaneously switching to viscous hydrodynamics (strong coupling) \cite{Liu:2015nwa, Broniowski:2008qk}.
-The free parameter $\taufs$ allows us to parametrically vary the \emph{time averaged} coupling strength in the approximate window $0 < \taufs \lesssim1$~fm/$c$.
+The free parameter $\taufs$ allows us to parametrically vary the \emph{time averaged} coupling strength in the approximate window $0 < \taufs \lesssim1\ \fmc$.
 
@@
 
-We thus follow the procedure of the authors and reinterpret our initial transverse entropy density at midrapidity $s(\mathbf{x_\perp}, \tau_0)\vert_{\eta=0}$ as a transverse density of noninteracting massless partons $n(\mathbf{x}, \tau_0)\vert_{\eta=0}$.
+We thus follow the procedure of the authors and reinterpret our parametrization for the initial transverse entropy density at midrapidity $s(\mathbf{x_\perp}, \tau_0)\vert_{\eta=0}$ as a parametrization for the initial transverse density of massless noninteracting partons $n(\mathbf{x}, \tau_0)\vert_{\eta=0}$.
 The free-streamed energy-momentum tensor $T^{\mu\nu}(x, y, \eta, \tau)$ at transverse coordinate $(x, y)$, midrapidity $\eta=0$, and proper time $\tau > \tau_0$ is then given by
@@
     $\dmin$           & Minimum inter-nucleon distance       & 0--1.7 fm       \\
-    $\taufs$          & Free-streaming time                  & 0.1--1.5 fm/$c$ \\
+    $\taufs$          & Free-streaming time                  & 0.1--1.5 \fmc   \\
     $\eta/s$ min      & Minimum value of $\eta/s$ (at $T_c$) & 0--0.2          \\
@@
 The present posterior has 15 dimensions, one for each parameter listed in Tbl.~\ref{tab:design}, and thus its joint distribution cannot be summarized by one figure alone.
-We thus sample the distribution and histogram the samples to project the distribution onto one or two dimensions at a time.
+We therefore sample the distribution and histogram the samples to project the distribution onto one or two dimensions at a time.
 Each diagonal panel is the distribution of a \emph{single} model parameter (marginalized over all others), and each lower-diagonal panel is the joint distribution of a \emph{pair} of model parameters (marginalized over all others).
@@
 It just so happens that the small range of $p$-values needed to describe this yield ratio are the same values needed to describe all of the calibration observables in the present study, \emph{and} numerous Pb-Pb observables at $\sqrts=2.76$~TeV \cite{Bernhard:2018hnz}.
-This work thus reaffirms an empirical scaling law reported in several previous studies \cite{Moreland:2014oya, Bernhard:2016tnd, Ke:2016jrd, Bernhard:2018hnz},
-that ultrarelativistic nuclear collisions deposit entropy at midrapidity approximated by
+This work thus reaffirms an empirical scaling law reported by several previous studies \cite{Moreland:2014oya, Bernhard:2016tnd, Ke:2016jrd, Bernhard:2018hnz} for the initial transverse entropy density (or massless parton density) of an ultrarelativistic nuclear collision, given by
 \begin{equation}
@@
 
-Meanwhile, our posterior on the constituent number $n_c$, shown enlarged in Fig.~\ref{fig:posterior_parton_number}, is not sharply peaked.
+The posterior on the constituent number $n_c$, shown enlarged in Fig.~\ref{fig:posterior_parton_number}, is not sharply peaked.
 We therefore refrain from quoting a distribution median and 90\% credible interval, although we do note that the distribution clearly favors $n_c > 1$ constituents.
@@
 
-The posterior free-streaming time $\taufs=0.37_{-0.27}^{+0.34}$~fm/$c$ obtained in this work is significantly smaller than our previous estimate $\taufs=1.16_{-0.25}^{+0.29}$~fm/$c$ in Ref.~\cite{Bernhard:2018hnz}.
+The posterior free-streaming time $\taufs=0.37_{-0.27}^{+0.34}\ \fmc$ obtained in this work is significantly smaller than our previous estimate $\taufs=1.16_{-0.25}^{+0.29}\ \fmc$ in Ref.~\cite{Bernhard:2018hnz}.
 We point out that the present study is missing several important observables which could affect the estimated free-streaming time, e.g.\ the Pb-Pb mean $p_T$ and mean $p_T$ fluctuations at $\sqrts=5.02$~TeV.
@@
 We show the MAP estimate predictions for the $p$-Pb and Pb-Pb relative mean $p_T$ fluctuations $\delta p_T / \langle p_T \rangle$ at $\sqrts=5.02$~TeV in the third row of Fig.~\ref{fig:obs_map}.
-For the Pb-Pb system, we use centrality bins and for the $p$-Pb system we use the same relative multiplicity bins used for two-particle flows.
-The relative mean $p_T$ fluctuations were shown to be particularly sensitive to the existence of nucleon substructure \cite{Bozek:2017elk}, and thus it would be interesting to ultimately include this observable in the calibration when the data becomes available.
+For the Pb-Pb system, we use centrality bins and for the $p$-Pb system we use the same relative multiplicity bins used for the $p$-Pb charged-particle mean $p_T$.
+The relative mean $p_T$ fluctuations have been shown to be particularly sensitive to the existence of nucleon substructure \cite{Bozek:2017elk}, and thus it would be interesting to ultimately include this observable in the calibration when the data becomes available.
 
@@
       $\dmin$  & 0.81 fm        & $T_\text{switch}$ & 151 MeV        \\
-      $\taufs$ & 0.37 fm/$c$
+      $\taufs$ & 0.37 \fmc
     \end{tabular}
@@
 \end{align}
-We show these model prediction in Fig.~\ref{fig:flow_corr} along with the \emph{normalized} symmetric cumulants
+We show these model predictions in Fig.~\ref{fig:flow_corr} along with the \emph{normalized} symmetric cumulants
 \begin{equation}
@@
     This agrees with a similar Bayesian analysis which was calibrated to Pb-Pb bulk observables at $\sqrts=2.76$ and 5.02~TeV using a hybrid model without nucleon substructure \cite{Bernhard:2018hnz}.
-    We also note that our result is in significant tension with an estimate for the effective nucleon width based on the gluon distribution extracted from HERA data \cite{Rezaeian:2012ji}.
+    We note that our result is in significant tension with an estimate for the effective nucleon width based on the gluon distribution extracted from HERA data \cite{Rezaeian:2012ji}.
   \item
@@
   \item
-    Our posterior estimate for the pre-equilibrium free-streaming time is $\taufs=0.37_{-0.27}^{+0.34}$~fm/$c$.
-    This is significantly shorter than the estimate obtained from a similar Bayesian study in Ref.~\cite{Bernhard:2018hnz}, which reported $\taufs=1.16_{-0.25}^{+0.29}$~fm/$c$.
+    Our posterior estimate for the pre-equilibrium free-streaming time is $\taufs=0.37_{-0.27}^{+0.34}\ \fmc$.
+    This is significantly shorter than the estimate obtained from a similar Bayesian study in Ref.~\cite{Bernhard:2018hnz}, which reported $\taufs=1.16_{-0.25}^{+0.29}\ \fmc$.
     It is not clear whether the difference is a result of nucleon substructure or the different observables used to calibrate each analysis.
@@
     We compare in Fig.~\ref{fig:region_shear_bulk} our estimate for the temperature dependence of the QGP specific shear and bulk viscosities to those of Ref.~\cite{Bernhard:2018hnz}, which performed a Bayesian calibration to Pb-Pb bulk observables at $\sqrts=2.76$ and 5.02~TeV using a physics model without nucleon substructure.
-    The two studies are in good agreement, although Ref.~\cite{Bernhard:2018hnz} obtains a more precise estimate for $(\eta/s)(T)$, likely due to the additional beam energies and observables included, and the enhanced sensitivity of larger collision systems to the QGP viscosity.
+    The two studies are in good agreement, although Ref.~\cite{Bernhard:2018hnz} obtains a more precise estimate for $(\eta/s)(T)$, likely due to the additional beam energies and observables included, and an enhanced sensitivity of larger collision systems to the QGP viscosity.
   \item
-    We make predictions in Fig.~\ref{fig:obs_map} for several quantities which were not included in the model calibration, including the identified yields, transverse energy, symmetric cumulants, and mean $p_T$ fluctuations at $\sqrts=5.02$~TeV.
+    We make predictions in Figs.~\ref{fig:obs_map} and \ref{fig:flow_corr} for several quantities which were not included in the model calibration, including the identified yields, transverse energy, symmetric cumulants, and mean $p_T$ fluctuations at $\sqrts=5.02$~TeV.
     Interestingly, our MAP estimate for the Pb-Pb symmetric cumulants at $\sqrts=5.02$~TeV are significantly different than those estimated in Ref.~\cite{Bernhard:2018hnz}.
-    This could be a direct or indirect result of including nucleon substructure in the model calibration.
+    This could be a direct (or indirect) result of including nucleon substructure in the model calibration.
 \end{enumerate}
@@
 Specifically, we are missing the transverse energy, identified particle yields, and the mean $p_T$ fluctuations of both collision systems, as well the charged-particle mean $p_T$ for the Pb-Pb system.
-These observables would certainly influence the quality of the combined fit and correspondingly our estimates for the posterior model parameters.
+These observables would certainly influence the quality of the combined fit and correspondingly our estimates for the model parameters.
 
 Similarly, the results would greatly benefit from additional beam energies and collision systems.
-Notably absent is data from RHIC which includes measurement for $p$-Au, $d$-Au, $^3$He-Au, Cu-Cu, Cu-Au, Au-Au and U-U collision systems at $\sqrts=200$~GeV.
-This data would provide an additional constraint on the model parameters and would enable more stringent tests of the calibrated model predictions.
+Notably absent is data from RHIC which includes measurements for $p$-Au, $d$-Au, $^3$He-Au, Cu-Cu, Cu-Au, Au-Au and \mbox{U-U} collision systems at $\sqrts=200$~GeV.
+This data would undoubtedly provide an additional constraint on the model parameters and would enable more stringent tests of the calibrated model predictions.
 The RHIC data may also help elucidate the beam-energy dependence of the model parameters which would be worth investigating.
@@
 \begin{acknowledgments}
-  JSM thanks Weiyao Ke, Bj\"orn Schenke and Heikki M\"antysaari for helpful discussions and clarifying comments.
+  JSM thanks Berndt M\"uller, Weiyao Ke, Bj\"orn Schenke, and Heikki M\"antysaari for helpful discussions and clarifying comments.
   This research was completed using 6 million CPU hours provided by the National Energy Research Scientific Computing Center (NERSC), a U.S.
@@
 \begin{figure}
-  \begin{tikzpicture}
+  \begin{tikzpicture}[scale=.9]
     \draw [black, thin, step=0.3cm] (0.9, 0.9) grid +(4.5, 4.5);
@@
 
-\begin{figure*}
-  \includegraphics{validation_example}
+\begin{figure}[b]
+  \includegraphics{validation_example_vert}
   \caption{
@@
   }
-\end{figure*}
+\end{figure}
 
@@
 
+\begin{figure*}
+  \includegraphics{validation_PbPb5020}\\
+  \includegraphics{validation_pPb5020}
+  \caption{
+    \label{fig:validation_all}
+    Emulator validation for the Pb-Pb collision system (top) and $p$-Pb collision system (bottom) at $\sqrts=5.02$~TeV.
+    The ``piano keys'' in the top row of each figure are horizontally stacked box plots for the normalized residuals of each model observable.
+    The boxes are 50\% interquartile ranges and whiskers are the 90\% interquantiles.
+    The bottom row of each figure is the RMS fractional error defined by Eqn.~\eqref{eq:frac_error}.
+  }
+\end{figure*}
+
 The emulator is a surrogate for the full physics simulation which generates probabilistic predictions for the model observables $\y_m$ at a given point $\x$.
@@
 \begin{equation}
+  \label{eq:frac_error}
   P(z) \sim \mathcal{N}(\mu=0,\sigma=1).
@@
 
-\begin{figure*}
-  \includegraphics{validation_PbPb5020}\\
-  \includegraphics{validation_pPb5020}
-  \caption{
-    \label{fig:validation_all}
-    Emulator validation for the Pb-Pb collision system (top) and $p$-Pb collision system (bottom) at $\sqrts=5.02$~TeV.
-    The ``piano keys'' in the top row of each figure are horizontally stacked box plots for the normalized residuals of each model observable.
-    The boxes are 50\% interquartile ranges and whiskers are the 90\% interquantiles.
-  }
-\end{figure*}
-
 \end{document}

### 2018-08-03 src/plots.py

@@ -1688,7 +1688,7 @@ def pca_vectors_variance(system=default_system):
 def boxplot(
         ax, percentiles, x=0, y=0, box_width=1,
         line_width=plt.rcParams['lines.linewidth'],
-        color=(0, 0, 0), alpha=.6, zorder=10
+        color=(0, 0, 0), alpha=.6, zorder=10, vert=True
 ):
     """
     Draw a minimal boxplot.
@@ -1977,6 +1977,134 @@ def validation_example(
     )
 
     set_tight(fig, rect=[.02, 0, 1, 1])
+
+
+@plot
+def validation_example_vert(
+        system='PbPb5020',
+        obs='dNch_deta', subobs=None,
+        label=r'$dN_\mathrm{ch}/d\eta$',
+        cent=(20, 30)
+):
+    """
+    Example of emulator validation for a single observable.  Scatterplot of
+    model calculations vs emulator predictions with histogram and boxplot of
+    normalized residuals.
+
+    """
+    fig, axes = plt.subplots(
+        nrows=2, figsize=figsize(.5, aspect=1.3),
+        gridspec_kw=dict(height_ratios=[1, 3]),
+    )
+
+    ax_hist, ax_scatter = axes
+
+    # model data
+    model_data  = model.data[system]
+    vdata = model_data[obs][subobs]
+    cent_slc = (slice(None), vdata['cent'].index(cent))
+    y = vdata['Y'][cent_slc]
+
+    # emulator predictions
+    mean_folds, cov_folds = validation_data(system)
+    y_ = np.concatenate(
+        [mean[obs][subobs][cent_slc] for mean in mean_folds], axis=0
+    )
+    std_ = np.concatenate(
+        [np.sqrt(cov[(obs, subobs), (obs, subobs)].T.diagonal()[cent_slc])
+            for cov in cov_folds], axis=0
+    )
+
+    color = obs_color(obs, subobs)
+    alpha = .6
+
+    ax_scatter.set_aspect('equal')
+    ax_scatter.errorbar(
+        y_, y, xerr=std_,
+        fmt='o', mew=.2, mec='white',
+        color=color, alpha=alpha
+    )
+    dy = .03*y.ptp()
+    x = [y.min() - dy, y.max() + dy]
+    ax_scatter.plot(x, x, color='.4')
+    ax_scatter.set_xlabel('Emulator prediction')
[…]

### 2018-08-06 paper/substructure.tex
Update numbers for latest posterior

@@
 
-Within the eikonal approximation, the initial entropy deposited at midrapidity at proper time $\tau=0^+\ \fmc$ can be represented by some mapping ${f(\T_A, \T_B) \mapsto dS/dy}$.
+Within the eikonal approximation, the initial entropy deposited at midrapidity at proper time $\tau=0^+\ \fmc$ can be represented by some mapping ${f: \T_A, \T_B \mapsto dS/dy}$.
 A natural first guess for the mapping is the arithmetic mean
@@
       \addlinespace[.4ex]
-      Norm       & \est{20.3}{2.2}{2.2}          & $\eta/s$ min     & \est{0.10}{0.07}{0.07}            \\[1.1ex]
-      $p$        & \est{0.018}{0.159}{0.146}     & $\eta/s$ slope   & \est{1.16}{1.16}{1.55} GeV$^{-1}$ \\[1.1ex]
-      $\sigmaf$  & \est{0.93}{0.27}{0.27}        & $\eta/s$ crv     & \est{-0.21}{0.79}{0.85}           \\[1.1ex]
-      $w$        & \est{0.97}{0.17}{0.19} fm     & $\zeta/s$ max    & \est{0.027}{0.027}{0.034}         \\[1.1ex]
-      $n_c$      & \est{6.2}{3.1}{2.8}           & $\zeta/s$ width  & \est{0.032}{0.032}{0.047} GeV     \\[1.1ex]
-      $\X$       & \est{0.34}{0.17}{0.22}        & $\zeta/s$ $T_0$  & \est{0.176}{0.020}{0.024} GeV     \\[1.1ex]
-      $\dmin$    & \est{1.12}{0.50}{0.56} fm     & $\Tsw$           & \est{0.150}{0.012}{0.015} GeV     \\[1.1ex]
-      $\taufs$   & \est{0.37}{0.27}{0.34} \fmc   & \\
+      Norm       & \est{20.2}{2.2}{2.2}          & $\eta/s$ min     & \est{0.10}{0.07}{0.07}            \\[1.1ex]
+      $p$        & \est{0.013}{0.155}{0.144}     & $\eta/s$ slope   & \est{1.11}{1.11}{1.42} GeV$^{-1}$ \\[1.1ex]
+      $\sigmaf$  & \est{0.93}{0.27}{0.26}        & $\eta/s$ crv     & \est{-0.14}{0.86}{0.81}           \\[1.1ex]
+      $w$        & \est{0.98}{0.17}{0.18} fm     & $\zeta/s$ max    & \est{0.026}{0.026}{0.033}         \\[1.1ex]
+      $n_c$      & \est{6.3}{3.1}{2.7}           & $\zeta/s$ width  & \est{0.033}{0.033}{0.048} GeV     \\[1.1ex]
+      $\X$       & \est{0.35}{0.17}{0.22}        & $\zeta/s$ $T_0$  & \est{0.176}{0.020}{0.024} GeV     \\[1.1ex]
+      $\dmin$    & \est{1.13}{0.51}{0.55} fm     & $\Tsw$           & \est{0.149}{0.013}{0.014} GeV     \\[1.1ex]
+      $\taufs$   & \est{0.37}{0.27}{0.33} \fmc   & \\
       \addlinespace[.4ex]
@@
 
-The \trento\ normalization factor ${\text{Norm} = 20.3^{+2.2}_{-2.2}}$ and generalized mean parameter $p=0.018^{+0.146}_{-0.159}$ are well constrained by the present analysis.
+The \trento\ normalization factor ${\text{Norm} = 20.2^{+2.2}_{-2.2}}$ and generalized mean parameter $p=0.013^{+0.144}_{-0.155}$ are well constrained by the present analysis.
 Moreover, their posterior values nicely describe the $p$-Pb and Pb-Pb calibration observables in Figs.~\ref{fig:obs_pbpb} and \ref{fig:obs_ppb}.
@@
 
-Continuing down the diagonal in Fig.~\ref{fig:posterior}, we see that the nucleon width parameter $w=0.97_{-0.17}^{+0.19}$~fm.
+Continuing down the diagonal in Fig.~\ref{fig:posterior}, we see that the nucleon width parameter $w=0.98_{-0.17}^{+0.18}$~fm.
 We caution that this parameter $w$ is \emph{not} the RMS radius of our nucleons due to idiosyncrasies of our constituent sampling procedure.
@@
 
-The posterior on the nucleon substructure parameter $\X=0.34_{-0.17}^{+0.22}$, on the other hand, is particularly sharply peaked.
+The posterior on the nucleon substructure parameter $\X=0.35_{-0.17}^{+0.22}$, on the other hand, is particularly sharply peaked.
 Recall that this parameter, defined in Eqn.~\eqref{eq:struct_param}, interpolates between the minimum and maximum widths of each constituent and hence
@@
 The posterior distribution, shown in blue, is remarkably well constrained.
-We report a posterior estimate for the constituent width $v=0.46_{-0.15}^{+0.21}$~fm which is roughly the same magnitude as the \emph{nucleon} hot-spots used in recent saturation-based substructure studies that employed IP-Glasma initial conditions \cite{Schenke:2018fci}.
+We report a posterior estimate for the constituent width $v=0.47_{-0.15}^{+0.20}$~fm which is roughly the same magnitude as the \emph{nucleon} hot-spots used in recent saturation-based substructure studies that employed IP-Glasma initial conditions \cite{Schenke:2018fci}.
 Evidently, it may be necessary to place an informative prior on our nucleon substructure parameters in order to resolve the apparent tension between our parameter values and those needed to describe DIS measurements at HERA.
@@
 
-The posterior free-streaming time $\taufs=0.37_{-0.27}^{+0.34}\ \fmc$ obtained in this work is significantly smaller than our previous estimate $\taufs=1.16_{-0.25}^{+0.29}\ \fmc$ in Ref.~\cite{Bernhard:2018hnz}.
+The posterior free-streaming time $\taufs=0.37_{-0.27}^{+0.33}\ \fmc$ obtained in this work is significantly smaller than our previous estimate $\taufs=1.16_{-0.25}^{+0.29}\ \fmc$ in Ref.~\cite{Bernhard:2018hnz}.
 We point out that the present study is missing several important observables which could affect the estimated free-streaming time, e.g.\ the Pb-Pb mean $p_T$ and mean $p_T$ fluctuations at $\sqrts=5.02$~TeV.
@@
   \item
-    We obtain an estimate $v=0.46_{-0.15}^{+0.21}$~fm for the Gaussian width of the constituent hot spots inside each nucleon.
+    We obtain an estimate $v=0.47_{-0.15}^{+0.20}$~fm for the Gaussian width of the constituent hot spots inside each nucleon.
     This is significantly larger than the length scales typically associated with nucleon substructure, and it may help explain the seemingly unreasonable success of hydrodynamics in small collision systems.
   \item
-    Our posterior estimate for the pre-equilibrium free-streaming time is $\taufs=0.37_{-0.27}^{+0.34}\ \fmc$.
+    Our posterior estimate for the pre-equilibrium free-streaming time is $\taufs=0.37_{-0.27}^{+0.33}\ \fmc$.
     This is significantly shorter than the estimate obtained from a similar Bayesian study in Ref.~\cite{Bernhard:2018hnz}, which reported $\taufs=1.16_{-0.25}^{+0.29}\ \fmc$.

### 2018-08-06 paper/substructure.tex
Final edits before posting to arXiv

@@
 \newcommand{\trento}{T\raisebox{-0.5ex}{R}ENTo}
-\newcommand{\sqrts}{\sqrt{s_{NN}}}
+\newcommand{\sqrts}{\sqrt{s_\mathrm{NN}}}
 \newcommand{\fmc}{\ensuremath{\text{fm}/c}}
@@
 The sheets of colliding nuclear density penetrate and pass through each other in proper time $\Delta \tau \approx D_\text{nucl} / (\gamma\, \beta_z)$ in the laboratory frame, where $D_\text{nucl}$ is the diameter of the nucleus in its rest frame, $\gamma$ is the usual Lorentz factor of the accelerated ions, and $\beta_z$ is their velocity along the beam axis.
-The resulting nuclear overlap time $\Delta \tau \lesssim 0.1\ \fmc$ at top RHIC and LHC energies is thus sufficiently short to neglect initial transverse dynamics which occur while the nuclei pass through each other.
+The resulting nuclear overlap time $\Delta \tau \lesssim 0.1\ \fmc$ at top RHIC and LHC energies, and thus it is safe to neglect the initial transverse dynamics which occur while the nuclei pass through each other.
 We therefore assume that the collision produces all secondary particles at uniform proper time $\tau = 0^+\ \fmc$, and that it deposits entropy (energy) at midrapidity which is a function of the locally varying beam integrated density of each nucleus.
@@
 \end{equation}
-The proton densities $\rho_{A,B}$ in Eqn.~\eqref{eq:tpp} are commonly modeled using a spherically symmetric distribution.
-For instance, the original implementation of the \trento\ model uses Gaussian protons, largely because it yields a simple analytic solution to Eqn.~\eqref{eq:pcoll}.
+The proton densities $\rho_{A,B}$ in Eq.~\eqref{eq:tpp} are commonly modeled using a spherically symmetric distribution.
+For instance, the original implementation of the \trento\ model uses Gaussian protons, largely because it yields a simple analytic solution to Eq.~\eqref{eq:pcoll}.
 Needless to say, such approximations are admittedly crude and may have a significant effect on the dynamics of small collision systems where the proton size is comparable to the size of the produced QGP medium.
@@
 \end{equation}
-and each constituent's position $\mathbf{x_i}$ in Eqn.~\eqref{eq:rho} is sampled from a Gaussian radial distribution with standard deviation $r$.
+and each constituent's position $\mathbf{x_i}$ in Eq.~\eqref{eq:rho} is sampled from a Gaussian radial distribution with standard deviation $r$.
 The effect of this additional nucleon substructure on the nuclear thickness functions is visible in Fig.~\ref{fig:thickness}
-
-The two protons $A$, $B$ are assigned a random impact parameter, and Eqn.~\eqref{eq:pcoll} is used to sample their inelastic collision probability $P_\mathrm{coll}(b)$.
-Note that this proton-proton inelastic collision probability has no direct knowledge of the individual constituent degrees of freedom; it is only \emph{indirectly} sensitive through the geometry of $\rho_{A, B}$ which depends on each of the constituent positions.
-This is an important distinction between the present model and a similar nucleon substructure implementation known as the participant or ``wounded'' quark model which allows for a subset of quarks (constituents) to participate inside a single nucleon \cite{ANISOVICH1978477, Broniowski:2016pvx}.
-The proton, unlike the nucleus, cannot produce semi-stable spectator fragments in a high-energy collision.
-Any spectator quarks produced by a wounded quark model would be colored objects that necessarily contribute to secondary particle production as they fragment and recombine to form color-neutral hadrons.
-We correspondingly require that the nucleons in Eqn.~\eqref{eq:rho} participate as singular objects, such that all spectator matter discarded by the simulation is appropriately color-neutral and inert.
-
-Assuming our two protons collide at the sampled impact parameter $\mathbf{b}$, we assign each a \emph{fluctuated} thickness
-\begin{equation}
-  \label{eq:fluctuated_thick}
-  \T_{A, B}(\x) \equiv \int dz\, \frac{1}{n_c} \sum\limits_{i=1}^{n_c} \gamma_i\, \rho_c \,(\mathbf{x} - \mathbf{x_i} \pm \mathbf{b}/2),
-\end{equation}
-equal to the beam-integrated proton density in Eqn.~\eqref{eq:rho}, with each constituent shifted by the appropriate impact parameter offset, and multiplied by a gamma random variable $\gamma_i$ with unit mean and variance $1/k$.
-These \emph{ad hoc} gamma random weights are necessary to describe the large fluctuations observed in minimum bias proton-proton collisions, although their exact physical origin is not well understood.
-
-Within the eikonal approximation, the initial entropy deposited at midrapidity at proper time $\tau=0^+\ \fmc$ can be represented by some mapping ${f: \T_A, \T_B \mapsto dS/dy}$.
-A natural first guess for the mapping is the arithmetic mean
-\begin{equation}
-  \frac{dS}{dy}\bigg\vert_{y=0} \propto \frac{\T_A + \T_B}{2}
-\end{equation}
-of the participant nuclear matter $\T_A$ and $T_B$, which is simply a wounded nucleon model up to meaningless factor of two in the normalization.
-The wounded nucleon model was in fact one of the first such mappings used as a proxy for initial particle production and entropy deposition in relativistic heavy-ion collisions \cite{Bialas:1976ed}.
-It was quickly realized, however, that the wounded nucleon model predicts the wrong scaling for charged particle production as a function of collision centrality and hence the wrong scaling for initial entropy deposition as a function of participant thicknesses $\T_A$ and $\T_B$ \cite{Kharzeev:2000ph}.
-
-A simple remedy is to replace the arithmetic mean of the wounded nucleon model with a more flexible parametrization
-\begin{equation}
-  \label{eq:gmean}
-  \frac{dS}{dy}\bigg\vert_{y=0} \propto \left( \frac{\T_A^p + \T_B^p}{2} \right)^{1/p},
-\end{equation}
-based on a family of functions known as the generalized mean(s).
-This parametrization introduces a dimensionless parameter $p$ which varies the scaling behavior of initial entropy deposition at midrapidity.
-For certain discrete values of $p$, it reduces to well known functional forms such as the arithmetic, geometric and harmonic means:
-\begin{equation}
-  \newlength{\extraspace}
-  \setlength{\extraspace}{0.5ex}
-  \frac{dS}{dy}\bigg\vert_{y=0} \propto
-  \begin{cases}
-    \max(T_A, T_B) & p \rightarrow +\infty, \\[\extraspace]
-    (T_A + T_B)/2 & p = +1, \hfill \text{ (arithmetic)} \\[\extraspace]
-    \sqrt{T_A T_B} & p = 0, \hfill \text{ (geometric)} \\[\extraspace]
-    2\, T_A T_B/(T_A + T_B) & p = -1, \hfill \text{ (harmonic)} \\[\extraspace]
-    \min(T_A, T_B) & p \rightarrow -\infty.
-  \end{cases}
-  \label{eq:trento_p}
-\end{equation}
-Conveniently, it has been shown that the generalized mean ansatz is able to mimic the scaling behavior of certain saturation based initial condition models \cite{Bernhard:2016tnd}, and hence it should serve as a reasonable parametric form for exploring QGP entropy deposition, assuming imperfect knowledge of saturation effects in nature.
-Of course the model is \emph{not} a substitute for first principle theory calculations, and it may fail to reproduce nuanced features of \emph{ab initio} models such as the existence of short-range gluon field fluctuations \cite{Schenke:2012wb}.
 
@@
 
+The two protons $A$, $B$ are assigned a random impact parameter, and Eq.~\eqref{eq:pcoll} is used to sample their inelastic collision probability $P_\mathrm{coll}(b)$.
+Note that this proton-proton inelastic collision probability has no direct knowledge of the individual constituent degrees of freedom; it is only \emph{indirectly} sensitive through the geometry of $\rho_{A, B}$ which depends on each of the constituent positions.
+This is an important distinction between the present model and a similar nucleon substructure implementation known as the participant or ``wounded'' quark model which allows for a subset of quarks (constituents) to participate inside a single nucleon \cite{ANISOVICH1978477, Broniowski:2016pvx}.
+The proton, unlike the nucleus, cannot produce semi-stable spectator fragments in a high-energy collision.
+Any spectator quarks produced by a wounded quark model would be colored objects that necessarily contribute to secondary particle production as they fragment and recombine to form color-neutral hadrons.
+We correspondingly require that the nucleons in Eq.~\eqref{eq:rho} participate as singular objects, such that all spectator matter discarded by the simulation is appropriately color-neutral and inert.
+
+Assuming our two protons collide at the sampled impact parameter $\mathbf{b}$, we assign each a \emph{fluctuated} thickness
+\begin{equation}
+  \label{eq:fluctuated_thick}
+  \T_{A, B}(\x) \equiv \int dz\, \frac{1}{n_c} \sum\limits_{i=1}^{n_c} \gamma_i\, \rho_c \,(\mathbf{x} - \mathbf{x_i} \pm \mathbf{b}/2),
+\end{equation}
+equal to the beam-integrated proton density in Eq.~\eqref{eq:rho}, with each constituent shifted by the appropriate impact parameter offset, and multiplied by a gamma random variable $\gamma_i$ with unit mean and variance $1/k$.
+These \emph{ad hoc} gamma random weights are necessary to describe the large fluctuations observed in minimum bias proton-proton collisions, although their exact physical origin is not well understood.
+
+Within the eikonal approximation, the initial entropy deposited at midrapidity and proper time $\tau=0^+\ \fmc$ is some function
+\begin{equation}
+  f: \T_A, \T_B \mapsto \frac{dS}{d^2x_\perp d\eta} \bigg\vert_{\eta=0},
+\end{equation}
+of the local density of participant matter $\T_A$, $\T_B$ in each nucleus.
+A natural first guess for this mapping is the arithmetic mean
+\begin{equation}
+  \frac{dS}{d^2x_\perp d\eta}\,\bigg\vert_{\eta=0} \propto \frac{\T_A + \T_B}{2},
+\end{equation}
+which is simply a wounded nucleon model up to meaningless factor of two in the normalization.
+The wounded nucleon model was in fact one of the first such mappings used as a proxy for initial particle production and entropy deposition in relativistic heavy-ion collisions \cite{Bialas:1976ed}.
+It was quickly realized, however, that the wounded nucleon model predicts the wrong scaling for charged particle production as a function of collision centrality and hence the wrong scaling for initial entropy deposition as a function of participant thicknesses $\T_A$ and $\T_B$ \cite{Kharzeev:2000ph}.
+
+A simple remedy is to replace the arithmetic mean of the wounded nucleon model with a more flexible parametrization
+\begin{equation}
+  \label{eq:gmean}
+  \frac{dS}{d^2x_\perp d\eta}\,\bigg\vert_{\eta=0} \propto \bigg( \frac{\T_A^p + \T_B^p}{2} \bigg)^{1/p},
+\end{equation}
+based on a family of functions known as the generalized mean(s).
+This parametrization introduces a dimensionless parameter $p$ which varies the scaling behavior of initial entropy deposition at midrapidity.
+For certain discrete values of $p$, it reduces to well known functional forms such as the arithmetic, geometric and harmonic means:
+\begin{equation}
+  \newlength{\extraspace}
+  \setlength{\extraspace}{0.5ex}
+  \frac{dS}{d\eta}\bigg\vert_{\eta=0} \propto
+  \begin{cases}
+    \max(T_A, T_B) & p \rightarrow +\infty, \\[\extraspace]
+    (T_A + T_B)/2 & p = +1, \hfill \text{ (arithmetic)} \\[\extraspace]
+    \sqrt{T_A T_B} & p = 0, \hfill \text{ (geometric)} \\[\extraspace]
+    2\, T_A T_B/(T_A + T_B) & p = -1, \hfill \text{ (harmonic)} \\[\extraspace]
+    \min(T_A, T_B) & p \rightarrow -\infty.
+  \end{cases}
+  \label{eq:trento_p}
+\end{equation}
+Conveniently, it has been shown that the generalized mean ansatz is able to mimic the scaling behavior of certain saturation based initial condition models \cite{Bernhard:2016tnd}, and hence it should serve as a reasonable parametric form for exploring QGP entropy deposition, assuming imperfect knowledge of saturation effects in nature.
+Of course the model is \emph{not} a substitute for first principle theory calculations, and it may fail to reproduce nuanced features of \emph{ab initio} models such as the existence of short-range gluon field fluctuations \cite{Schenke:2012wb}.
+
 Equation~\eqref{eq:gmean} is a purely local function of nuclear density in the transverse plane and should (in principle) be equally valid for any pair of colliding nuclei at sufficiently high beam energy.
 The model readily generalizes from individual proton-proton collisions to arbitrary nucleus-nucleus collisions by summing the participant thicknesses $\T_{A,B}$ over all nucleons which participate in one or more inelastic collisions.
-The only modeling difference between $p$-$p$ collisions and Pb-Pb collisions is the number and the position of the nucleons.
+The only modeling difference between $p$-$p$, $p$-Pb, and Pb-Pb collisions is the number and the position of the nucleons.
 
@@
 
-The parametric entropy deposition ansatz in Eqn.~\eqref{eq:gmean} does not provide any information about the initial masses or momenta of particles produced in the collision.
+The parametric entropy deposition ansatz in Eq.~\eqref{eq:gmean} does not provide any information about the initial masses or momenta of particles produced in the collision.
 In general, these details will affect the dynamics predicted by the collisionless Boltzmann equation
@@
 
-We thus follow the procedure of the authors and reinterpret our parametrization for the initial transverse entropy density at midrapidity $s(\mathbf{x_\perp}, \tau_0)\vert_{\eta=0}$ as a parametrization for the initial transverse density of massless noninteracting partons $n(\mathbf{x}, \tau_0)\vert_{\eta=0}$.
-The free-streamed energy-momentum tensor $T^{\mu\nu}(x, y, \eta, \tau)$ at transverse coordinate $(x, y)$, midrapidity $\eta=0$, and proper time $\tau > \tau_0$ is then given by
+The entropy density for a gas of massless noninteracting particles is very nearly proportional to its particle density, and so we are able to recast the parametrization in Eq.~\eqref{eq:gmean} in the form
+\begin{equation}
+  \frac{dN}{d^2x_\perp d\eta}\,\bigg\vert_{\eta=0} = \, \text{Norm} \times \bigg(\frac{\T_A^p + \T_B^p}{2}\bigg)^{1/p},
+\end{equation}
+where the left-side is the initial density of free-streaming partons at midrapidity.
+The resulting free-streamed energy-momentum tensor $T^{\mu\nu}(x, y, \eta, \tau)$ at transverse coordinate $(x, y)$, midrapidity $\eta=0$, and proper time $\tau > \tau_0$ is then given by
 \begin{multline}
@@
 \end{multline}
-where $\Delta\tau$ is the free-streaming time $\Delta\tau = \tau - \tau_0$.
-The solution \eqref{energy-momentum} is decomposed in hydrodynamic form
+where $n$ is the local density of massless partons, and $\Delta\tau$ is the free-streaming time $\Delta\tau = \tau - \tau_0$.
+The solution \eqref{energy-momentum} can then be decomposed in hydrodynamic form
 \begin{equation}
@@
 \end{equation}
-The initial bulk and shear corrections are finally solved for by subtracting the ideal pressure from the total pressure to find $\Pi$, then solving for $\pi^{\mu\nu}$ using Eqn.~\eqref{hydro-eqn}
+The initial bulk and shear corrections are finally solved for by subtracting the ideal pressure from the total pressure to find $\Pi$, then solving for $\pi^{\mu\nu}$ using Eq.~\eqref{hydro-eqn}
 \begin{align}
@@
 \end{equation}
-for the hydrodynamic energy-momentum tensor $T^{\mu\nu}$ expressed in Eqn.~\eqref{hydro-eqn} using a set of second-order Israel-Stewart equations formulated in the 14-moment approximation
+for the hydrodynamic energy-momentum tensor $T^{\mu\nu}$ expressed in Eq.~\eqref{hydro-eqn} using a set of second-order Israel-Stewart equations formulated in the 14-moment approximation
 \cite{Israel:1979wp, Israel:1976aa, Denicol:2012cn, Denicol:2010xn}.
@@
 described by tunable maximum, width, and location ($T_0$) parameters.
-Figure~\ref{fig:viscosity_dof} shows several of the possible curves parametrized by Eqns.~\eqref{eq:shear_viscosity} and \eqref{eq:bulk_viscosity}, although many more are possible.
+Figure~\ref{fig:viscosity_dof} shows several of the possible curves parametrized by Eqs.~\eqref{eq:shear_viscosity} and \eqref{eq:bulk_viscosity}, although many more are possible.
 
@@
 \end{equation}
-where $i$ is an index over species, $f_i$ is the distribution function of that species, and $d^3\sigma_\mu$ is a volume element of the isothermal hypersurface $\Sigma$ defined by $T_\mathrm{sw}$.
+where $i$ is an index over species, $f_i$ is the distribution function of that species, and $d^3\sigma_\mu$ is a volume element of the isothermal hypersurface $\Sigma$ defined by $\Tsw$.
 Thermal particles are then sampled in the rest frame of each fluid cell according to a Bose-Einstein or Fermi-Dirac distribution at zero baryon chemical potential
@@
 
-Traditionally, particlization models have sampled resonances using each particle's pole mass in Eqn.~\eqref{distribution}.
+Traditionally, particlization models have sampled resonances using each particle's pole mass in Eq.~\eqref{distribution}.
 This approximation, however, is somewhat crude and has been known to underpredict pion production, particularly at low $p_T$ \cite{Sollfrank:1991xm, Huovinen:2016xxq, Vovchenko:2018fmh}.
@@
 
-When the viscous terms $\pi^{\mu\nu}$ and $\Pi$ are nonzero in Eqn.~\eqref{hydro-eqn}, the distribution function $f$ must be modified to preserve the continuity of $T^{\mu\nu}$ as the system transitions from hydrodynamics to Boltzmann transport.
+When the viscous terms $\pi^{\mu\nu}$ and $\Pi$ are nonzero in Eq.~\eqref{hydro-eqn}, the distribution function $f$ must be modified to preserve the continuity of $T^{\mu\nu}$ as the system transitions from hydrodynamics to Boltzmann transport.
 We perform the appropriate modification using a general method which transforms the momentum vector \emph{inside} the distribution function \cite{Pratt:2010jt}
@@
 \begin{equation}
-  \frac{\eta}{\tau} = \frac{1}{15 T} \sum\limits_{sp} g \int \frac{d^3p}{(2\pi)^3}\frac{p^4}{E^2} f_0 (1 \pm f_0).
-\end{equation}
+  \frac{\eta}{\tau} = \frac{1}{15 T} \sum\limits_{sp} g \int \frac{d^3p}{(2\pi)^3}\frac{p^4}{E^2} f_0 (1 \pm f_0),
+\end{equation}
+where the sum runs over all species in the hadron gas, and $g$ and $f_0$ are the degeneracy factor and equilibrium distribution function of each species respectively.
+
 For the bulk viscous correction, we use a novel procedure developed in Ref.~\cite{Bernhard:2018hnz}.
@@
 \end{equation}
-For a given bulk pressure, we rescale the momentum $p$ inside the distribution function $f(p) \rightarrow f(p + \lambda_\mathrm{bulk}\, p)$ and adjust the parameter $\lambda_\mathrm{bulk}$ to match the total pressure on the left side of Eqn.~\eqref{kinetic-pressure}.
+For a given bulk pressure, we rescale the momentum $p$ inside the distribution function $f(p) \rightarrow f(p + \lambda_\mathrm{bulk}\, p)$ and adjust the parameter $\lambda_\mathrm{bulk}$ to match the total pressure on the left side of Eq.~\eqref{kinetic-pressure}.
 This substitution of course also modifies the energy density
@@
 \end{equation}
-and so a fugacity term $z_\mathrm{bulk}$ is introduced which decreases the yield of all particles by the same overall factor to compensate.
+and so a fugacity term $z_\mathrm{bulk}$ is introduced which modifies the yield of all particles by the same overall factor to compensate.
 The full transformation is then given by $f(p) \rightarrow z_\mathrm{bulk}\, f(p + \lambda_\mathrm{bulk}\, p)$, where the parameters $\lambda_\mathrm{bulk}$ and $z_\mathrm{bulk}$ are determined numerically for each value of the bulk pressure.
 
-Once the fluid is converted into particles, we simulate its subsequent microscopic dynamics using the Ultra-relativistic Quantum Molecular Dynamics (UrQMD) transport model \cite{Bass:1998ca, Bleicher:1999xi}.
+Once the fluid is converted into particles, we simulate its subsequent microscopic dynamics using the Ultra-relativistic Quantum modifies Dynamics (UrQMD) transport model \cite{Bass:1998ca, Bleicher:1999xi}.
 It solves the microscopic Boltzmann equation
@@
 
-We assume that the likelihood function in Eqn.~\eqref{eq:bayes} is described by a multivariate normal distribution
+We assume that the likelihood function in Eq.~\eqref{eq:bayes} is described by a multivariate normal distribution
 \begin{equation}
@@
     $\zeta/s$ $T_0$   & Temperature of $\zeta/s$ maximum     & 150--200 MeV    \\
-    $T_\text{switch}$ & Switching/particlization temp.       & 135--165 MeV    \\
+    $\Tsw$            & Switching/particlization temp.       & 135--165 MeV    \\
   \end{tabular}
@@
 \subsection{Parameter design}
-
-For the prior $P(H_\x)$, we specify ranges, i.e.\ minimum and maximum values, for each parameter which are listed in Tbl.~\ref{tab:design}.
-We assume the prior distribution is constant and nonzero within each specified range and zero otherwise.
-The selected parameter ranges are intentionally wide to avoid clipping the calibrated posterior; for example, a previous analysis of the \trento\ model \cite{Bernhard:2016tnd} found $p \sim 0$, but we use a prior range $p \in [-1, 1]$ to account for differences in the present model, e.g.\ nucleon substructure, which could modify its posterior.
-Several of the model parameters require special care and are reparametrized accordingly:
-\begin{itemize}[leftmargin=1\parindent]
-  \item
-    The constituent fluctuations, modeled by the gamma random weights $\gamma_i$ in Eqn.~\eqref{eq:fluctuated_thick}, generate overall nucleon fluctuations which are suppressed by the number of constituents $n_c$ inside the nucleon.
-    The observed nucleon fluctuation variance falls like $1 / n_c$ which means the natural range for the constituent fluctuations is larger when the number of constituents is larger and \emph{vice versa}.
-    We therefore reparametrize the constituent fluctuations using the standard deviation of the resulting nucleon fluctuations
-    \begin{equation}
-      \sigma_\mathrm{fluct} = 1 / \sqrt{k\, n_c},
-    \end{equation}
-    where $k$ is the shape parameter of the gamma distribution used to fluctuate each individual constituent.
-  \item
-    In Sec.~\ref{sec:initial_state} we parametrized nucleon substructure using three degrees of freedom:
-    \begin{enumerate*}[label=(\roman*)]
-      \item a parameter $v$ for the Gaussian width of each constituent,
-      \item a parameter $r$ for the Gaussian width of the radial distribution used to sample the constituent centers, and
-      \item a parameter $n_c$ to vary the number of constituents inside the nucleon.
-    \end{enumerate*}
-    In the limit $n_c \to \infty$, the composite nucleon's root-mean-square (RMS) radius is simply the convolution of its sampling radius $r$ and constituent width $v$ which add together in quadrature
-    \begin{equation}
-      \lim_{n_c \to\infty} r_\mathrm{RMS}\{\rho\} = \sqrt{r^2 + v^2}.
-    \end{equation}
-    We therefore choose to reparametrize the sampling radius $r$ in terms of a new variable
-    \begin{equation}
-      \label{eq:nucleon_width}
-      w = \sqrt{r^2 + v^2},
-    \end{equation}
-    which approximates the RMS radius of the nucleon when the number of sampled constituents is large (see Fig.~\ref{fig:nucleon_schematic} for an example proton).
-    We thus call $w$ a nucleon ``width'', although for smaller numbers of constituents, the \emph{actual} RMS radius of our sampled nucleons can be significantly smaller than our width parameter $w$ due to fluctuations in the nucleon's center of mass, and hence one should account for the difference when discussing the nucleon's posterior RMS radius.
-  \item
-    Equation~\eqref{eq:nucleon_width} requires the nucleon width to be larger than the constituent width, i.e.
-$w > v$, lest the sampling radius $r$ turn imaginary, and thus we cannot vary the nucleon width $w$ and the constituent width $v$ independently.
-    We therefore reparametrize the constituent width $v$ using a new variable $\X$ to interpolate between minimum and maximum allowed values:
-    \begin{equation}
-      \label{eq:struct_param}
-      v = v_\mathrm{min} + \X (v_\mathrm{max} - v_\mathrm{min}).
-    \end{equation}
-    Here we choose a minimum constituent width $v_\mathrm{min} = 0.2$~fm determined by computational limits and a maximum width $v_\mathrm{max} = w$ equal to the nucleon width.
-    Thus for $\X=0$, the nucleons consist of small, distinct hot spots, whereas for $\X=1$ the nucleon is a single Gaussian blob of width $w$.
-    The parameter $\X$ thus varies the granularity of the nucleon while keeping the number of constituents $n_c$ and approximate size of the nucleon $w$ fixed.
-\end{itemize}
-
-\begin{figure}[t]
-  \begin{tikzpicture}
-    \draw[dashed] (0,0) circle (1.6cm);
-    \foreach \theta in {60, 180, 300}{
-      \draw ({\theta}:.85) circle (.73cm);
-    }
-    \coordinate (a) at (0, 0);
-    \coordinate (b) at (.8, 1.385);
-    \coordinate (c) at (.425, -.736);
-    \coordinate (d) at (-1.58, 0);
-    \coordinate (e) at (-.85, 0);
-    \draw[<->] (a) -- (b) node[midway, xshift=-1ex, yshift=.8ex]{$w$};
-    \draw[<->] (a) -- (c) node[midway, xshift=1ex, yshift=.8ex]{$r$};
-    \draw[<->] (d) -- (e) node[midway, yshift=1ex]{$v$};
-  \end{tikzpicture}
-  \caption{
-    \label{fig:nucleon_schematic}
-    Schematic illustrating the constituent sampling radius $r$, constituent width $v$ and nucleon width parameter $w = \sqrt{r^2 + v^2}$ for a nucleon with $n_c=3$ constituents.
-  }
-\end{figure}
 
@@
 \end{table*}
+
+For the prior $P(H_\x)$, we specify ranges, i.e.\ minimum and maximum values, for each parameter which are listed in Table~\ref{tab:design}.
+We assume the prior distribution is constant and nonzero within each specified range and zero otherwise.
+The selected parameter ranges are intentionally wide to avoid clipping the calibrated posterior; for example, a previous analysis of the \trento\ model \cite{Bernhard:2016tnd} found $p \sim 0$, but we use a prior range $p \in [-1, 1]$ to account for differences in the present model, e.g.\ nucleon substructure, which could modify its posterior.
+Several of the model parameters require special care and are reparametrized accordingly:
+
+\begin{figure}
+  \begin{tikzpicture}
+    \draw[dashed] (0,0) circle (1.6cm);
+    \foreach \theta in {60, 180, 300}{
+      \draw ({\theta}:.85) circle (.73cm);
+    }
+    \coordinate (a) at (0, 0);
+    \coordinate (b) at (.8, 1.385);
+    \coordinate (c) at (.425, -.736);
+    \coordinate (d) at (-1.58, 0);
+    \coordinate (e) at (-.85, 0);
+    \draw[<->] (a) -- (b) node[midway, xshift=-1ex, yshift=.8ex]{$w$};
+    \draw[<->] (a) -- (c) node[midway, xshift=1ex, yshift=.8ex]{$r$};
+    \draw[<->] (d) -- (e) node[midway, yshift=1ex]{$v$};
+  \end{tikzpicture}
+  \caption{
+    \label{fig:nucleon_schematic}
+    Schematic illustrating the constituent sampling radius $r$, constituent width $v$ and nucleon width parameter $w = \sqrt{r^2 + v^2}$ for a nucleon with $n_c=3$ constituents.
+  }
+\end{figure}
+
+\begin{itemize}[leftmargin=1\parindent]
+  \item
+    The constituent fluctuations, modeled by the gamma random weights $\gamma_i$ in Eq.~\eqref{eq:fluctuated_thick}, generate overall nucleon fluctuations which are suppressed by the number of constituents $n_c$ inside the nucleon.
+    The observed nucleon fluctuation variance falls like $1 / n_c$ which means the natural range for the constituent fluctuations is larger when the number of constituents is larger and \emph{vice versa}.
+    We therefore reparametrize the constituent fluctuations using the standard deviation of the resulting nucleon fluctuations
+    \begin{equation}
+      \sigma_\mathrm{fluct} = 1 / \sqrt{k\, n_c},
+    \end{equation}
+    where $k$ is the shape parameter of the gamma distribution used to fluctuate each individual constituent.
+  \item
+    In Sec.~\ref{sec:initial_state} we parametrized nucleon substructure using three degrees of freedom:
+    \begin{enumerate*}[label=(\roman*)]
+      \item a parameter $v$ for the Gaussian width of each constituent,
+      \item a parameter $r$ for the Gaussian width of the radial distribution used to sample the constituent centers, and
+      \item a parameter $n_c$ to vary the number of constituents inside the nucleon.
+    \end{enumerate*}
+    In the limit $n_c \to \infty$, the composite nucleon's root-mean-square (RMS) radius is simply the convolution of its sampling radius $r$ and constituent width $v$ which add together in quadrature
+    \begin{equation}
+      \lim_{n_c \to\infty} r_\mathrm{RMS}\{\rho\} = \sqrt{r^2 + v^2}.
+    \end{equation}
+    We therefore choose to reparametrize the sampling radius $r$ in terms of a new variable
+    \begin{equation}
+      \label{eq:nucleon_width}
+      w = \sqrt{r^2 + v^2},
+    \end{equation}
+    which approximates the RMS radius of the nucleon when the number of sampled constituents is large (see Fig.~\ref{fig:nucleon_schematic} for an example proton).
+    We thus call $w$ a nucleon ``width'', although for smaller numbers of constituents, the \emph{actual} RMS radius of our sampled nucleons can be significantly smaller than our width parameter $w$ due to fluctuations in the nucleon's center of mass, and hence one should account for the difference when discussing the nucleon's posterior RMS radius.
+  \item
+    Equation~\eqref{eq:nucleon_width} requires the nucleon width to be larger than the constituent width, i.e.
+$w > v$, lest the sampling radius $r$ turn imaginary, and thus we cannot vary the nucleon width $w$ and the constituent width $v$ independently.
+    We therefore reparametrize the constituent width $v$ using a new variable $\X$ to interpolate between minimum and maximum allowed values:
+    \begin{equation}
+      \label{eq:struct_param}
+      v = v_\mathrm{min} + \X (v_\mathrm{max} - v_\mathrm{min}).
+    \end{equation}
+    Here we choose a minimum constituent width $v_\mathrm{min} = 0.2$~fm determined by computational limits and a maximum width $v_\mathrm{max} = w$ equal to the nucleon width.
+    Thus for $\X=0$, the nucleons consist of small, distinct hot spots, whereas for $\X=1$ the nucleon is a single Gaussian blob of width $w$.
+    The parameter $\X$ thus varies the granularity of the nucleon while keeping the number of constituents $n_c$ and approximate size of the nucleon $w$ fixed.
+\end{itemize}
 
@@
 At the time of this writing, many of the aforementioned experimental observables are not yet published for $p$-Pb and Pb-Pb collisions at $\sqrts=5.02$~TeV.
-We therefore restrict our calibration to the subset of measured and published observables listed in Tbl.~\ref{tab:observables}.
+We therefore restrict our calibration to the subset of measured and published observables listed in Table~\ref{tab:observables}.
 Notably absent from this list are the four-particle cumulants $\vnk{n}{4}$ at $\sqrts=5.02$~TeV despite being measured and published.
@@
 We therefore refrain from \emph{calibrating} on the four-particle cumulants, although we do show calculations of the four-particle cumulant $\vnk{2}{4}$ later in the text, using a single set of calibration parameters.
-
-Most of the calibration observables listed in Tbl.~\ref{tab:observables} are calculated as a function of collision centrality, where centrality is defined using some measure of the underlying event activity, e.g.\ the charged particle yield in a given rapidity window.
-When calculating these observables, we generate \order{4} minimum-bias events at each design point and divide the events into centrality bins using the charged particle yield at midrapidity, similar to the procedure used by experiment.
-
-However, for some observables such as $p$-Pb mean $p_T$ \cite{Abelev:2013bla} and flow cumulants $\vnk{n}{k}$ \cite{Chatrchyan:2013nka}, the experiments use a special high-multiplicity trigger to select rare, ultra-central events according to the number of charged particles produced or detector tracks offline.
-These high-multiplicity bins are too selective for our modest minimum-bias event sample, and so a different procedure is required.
-We exploit, for this purpose, the approximate correspondence
-\begin{equation}
-  \label{eq:bin_equiv}
-  \ntrk \sim \nch \propto N_\rho
-\end{equation}
-between each event's initial secondary-parton density ${N_\rho = \int d^2x\,n(\x)}$ in Eqn.~\eqref{energy-momentum} and its final-state event activity, characterized by $\nch$ or $\ntrk$ \cite{Abelev:2013bla, Chatrchyan:2013nka}.
-
-Consider, for example, an experimental multiplicity bin $(\nch^\text{low}, \nch^\text{high})$ with some kinematic cuts on $p_T$ and $\eta$.
-We rescale this bin by the average multiplicity $\langle \nch \rangle$ of the corresponding minimum-bias event sample, i.e.\
-\begin{equation}
-  (\nch^\text{low}, \nch^\text{high}) \rightarrow (\nch^\text{low} / \langle \nch \rangle, \nch^\text{high} / \langle \nch \rangle)
-\end{equation}
-in order to reexpress each bin edge as a unitless variable.
-These rescaled bins are then used to select initial condition events using the equivalence \eqref{eq:bin_equiv}:
-\begin{equation}
-  \label{eq:mult_trigger}
-  \left (\frac{N^\text{low}_\rho}{\langle N_\rho \rangle}, \frac{N^\text{high}_\rho}{\langle N_\rho \rangle} \right) \leftrightarrow \left (\frac{\nch^\text{low}}{\langle \nch \rangle}, \frac{\nch^\text{high}}{\langle \nch \rangle} \right ).
-\end{equation}
 
@@
     Each cell represents an observable in a single centrality bin. Experimental statistical and systematic errors are from ALICE \cite{Adam:2015ptt, Adam:2016izf}.
-    The experimental correlation structure is modeled using Eqn.~\eqref{eq:corr}.
+    The experimental correlation structure is modeled using Eq.~\eqref{eq:corr}.
   }
 \end{figure*}
+
+Most of the calibration observables listed in Table~\ref{tab:observables} are calculated as a function of collision centrality, where centrality is defined using some measure of the underlying event activity, e.g.\ the charged particle yield in a given rapidity window.
+When calculating these observables, we generate \order{4} minimum-bias events at each design point and divide the events into centrality bins using the charged particle yield at midrapidity, similar to the procedure used by experiment.
+
+However, for some observables such as $p$-Pb mean $p_T$ \cite{Abelev:2013bla} and flow cumulants $\vnk{n}{k}$ \cite{Chatrchyan:2013nka}, the experiments use a special high-multiplicity trigger to select rare, ultra-central events according to the number of charged particles produced or detector tracks offline.
+These high-multiplicity bins are too selective for our modest minimum-bias event sample, and so a different procedure is required.
+We exploit, for this purpose, the approximate correspondence
+\begin{equation}
+  \label{eq:bin_equiv}
+  \ntrk \sim \nch \propto N_\text{parton}
+\end{equation}
+between each event's initial secondary-parton density ${N_\text{parton} = \int d^2x\,n(\x)}$ in Eq.~\eqref{energy-momentum} and its final-state event activity, characterized by $\nch$ or $\ntrk$ \cite{Abelev:2013bla, Chatrchyan:2013nka}.
+
+Consider, for example, an experimental multiplicity bin $(\nch^\text{low}, \nch^\text{high})$ with some kinematic cuts on $p_T$ and $\eta$.
+We rescale this bin by the average multiplicity $\langle \nch \rangle$ of the corresponding minimum-bias event sample, i.e.\
+\begin{equation}
+  (\nch^\text{low}, \nch^\text{high}) \rightarrow (\nch^\text{low} / \langle \nch \rangle,\, \nch^\text{high} / \langle \nch \rangle)
+\end{equation}
+in order to reexpress each bin edge as a unitless variable.
+These rescaled bins are then used to select initial condition events using the equivalence \eqref{eq:bin_equiv}:
+\begin{equation}
+  \label{eq:mult_trigger}
+  \left (\frac{N^\text{low}_\text{parton}}{\langle N_\text{parton} \rangle}, \frac{N^\text{high}_\text{parton}}{\langle N_\text{parton} \rangle} \right) \leftrightarrow \left (\frac{\nch^\text{low}}{\langle \nch \rangle}, \frac{\nch^\text{high}}{\langle \nch \rangle} \right ).
+\end{equation}
 
@@
 \end{equation}
-in Eqn.~\eqref{eq:likelihood}, which includes uncertainty contributions from both the model $\Sigma_m$ and experimental data $\Sigma_e$.
+in Eq.~\eqref{eq:likelihood}, which includes uncertainty contributions from both the model $\Sigma_m$ and experimental data $\Sigma_e$.
 The experimental contribution to the covariance $\Sigma_e$ is further broken down into statistical and systematic components,
@@
 where $b_i$ and $b_j$ are the midpoints of two observable bins of a single type (centrality or relative multiplicity) and $l$ is a correlation length which describes how quickly the observable bins decorrelate as the distance between the bins increases.
-We use centrality correlation lengths $l=100$ for all of the centrality-binned Pb-Pb observables and $l=30$ for the centrality-binned $p$-Pb charged particle yield $d\nch/d\eta$.
+We use centrality correlation lengths $l=100$ for all of the centrality binned Pb-Pb observables and $l=30$ for the centrality binned $p$-Pb charged particle yield $d\nch/d\eta$.
 The \mbox{$p$-Pb} mean $p_T$ and flow observables, meanwhile, use relative multiplicity bins $\nch / \langle \nch \rangle$ and $\ntrk / \langle \ntrk \rangle$ which necessitate a smaller correlation length $l=5$.
-
-We show, as an example, the resulting correlation matrix
+We show an example correlation matrix
 \begin{equation}
@@
 \end{equation}
-for the Pb-Pb experimental data constructed using Eqn.~\eqref{eq:corr} on the right-side of Fig.~\ref{fig:correlation}.
+for the Pb-Pb experimental data constructed using Eq.~\eqref{eq:corr} on the right-side of Fig.~\ref{fig:correlation}.
 Here $y_i$ denotes an element of the experimental data $\y_e$ and $\sigma_i$ its corresponding uncertainty.
@@
 
-In principle, one could calculate the likelihood function in Eqn.~\eqref{eq:likelihood} directly, e.g.\ by running the model a large number of times at a given parameter point $\x$ to calculate the model observables $\y_m(\x)$ from the ensemble of simulated events, but in practice such a procedure would be intractable.
+In principle, one could calculate the likelihood function in Eq.~\eqref{eq:likelihood} directly, e.g.\ by running the model a large number of times at a given parameter point $\x$ to calculate the model observables $\y_m(\x)$ from the ensemble of simulated events, but in practice such a procedure would be intractable.
 The model is computationally intensive to evaluate, and thousands of events are required to calculate the simplest observables at a single parameter point.
-Moreover, we need to evaluate Eqn.~\eqref{eq:likelihood} \emph{numerous} times in order to sample the multidimensional posterior distribution so that the samples may be histogrammed and visualized.
+Moreover, we need to evaluate Eq.~\eqref{eq:likelihood} \emph{numerous} times in order to sample the multidimensional posterior distribution so that the samples may be histogrammed and visualized.
 
@@
 The emulator enables essentially instantaneous predictions for $\y_m = f(\x)$ and allows us to sample the posterior distribution millions of times.
-In order to train the emulator, we first generate a scaffolding of the parameter space using a maximin Latin hypercube design \cite{Morris:1995lh} to distribute 500 points uniformly throughout our 15-dimensional parameter space according to the parameter ranges in Tbl.~\ref{tab:design}.
+In order to train the emulator, we first generate a scaffolding of the parameter space using a maximin Latin hypercube design \cite{Morris:1995lh} to distribute 500 points uniformly throughout our 15-dimensional parameter space according to the parameter ranges in Table~\ref{tab:design}.
 We then run minimum-bias and multiplicity triggered $p$-Pb and Pb-Pb events at each design point and calculate the model observables from the ensemble of events.
 
-Specifically, let $X$ denote the ${d \times n}$ design matrix of $d=500$ training points, where each training point is a vector $\x = (x_1, x_2, \dots, x_n)$ of the $n=15$ model parameters in Tbl.~\ref{tab:design}.
+Specifically, let $X$ denote the ${d \times n}$ design matrix of $d=500$ training points, where each training point is a vector $\x = (x_1, x_2, \dots, x_n)$ of the $n=15$ model parameters in Table~\ref{tab:design}.
 Similarly, let $Y$ denote the corresponding ${d \times m}$ observables matrix, where each row of $Y$ is a vector $\y_m = (y_1, y_2, \dots, y_m)$ of $m$ different simulated observables (here we overload our notation so the subscript $m$ means both \emph{model} and \emph{observable number}).
@@
 We use for this purpose a specific type of emulator known as a Gaussian process emulator (GPE) \cite{Rasmussen:2006gp}.
-The advantage of using GPE's is that they provide an estimate of their own uncertainty which allows us to account for this uncertainty when constructing the covariance matrix $\Sigma$ in Eqn.~\eqref{eq:likelihood}.
+The advantage of using GPE's is that they provide an estimate of their own uncertainty which allows us to account for this uncertainty when constructing the covariance matrix $\Sigma$ in Eq.~\eqref{eq:likelihood}.
 One quirk of GPE's is that they are restricted to scalar-valued functions, i.e.\ functions of one output, whereas we require an emulator for vector-valued functions with multiple outputs.
@@
 \end{equation}
-where $\tilde{\y}_i$ are the standardized observables of the $i$-th row-vector (design point) of matrix $\tilde{Y}$, and $z_{ij}$ and $\mathbf{v}_j$ are the coefficients and corresponding vectors of its $j$-th principal component.
+where $\tilde{\y}_i$ are the standardized observables of the $i$-th row-vector (design point) of matrix $\tilde{Y}$, and $z_{ij}$ and $\mathbf{v}_j$ are the coefficients and vectors of its $j$-th principal component.
 
@@
 
-The GPE's are essentially fancy interpolators applied to the model's training points and PCA transformed observables.
-Each GPE reports a mean value $z(\x)$ as well as an estimated error $\delta z(\x)$ which accounts for statistical noise in the training data and interpolation error between the design points.
-Once the GPE's are trained, we can predict the model observables $\y_m$ at parameter point $\x$ by transforming the vector of emulated principal components
-\begin{equation}
-  \mathbf{z}(\x) = (z_1(\x), z_2(\x), \dots, z_k(\x))
-\end{equation}
-\emph{back} to physical space.
-Similarly, we can construct the covariance matrix of the observables in PCA space,
-\begin{equation}
-  \cov(z_i, z_j) = \diag[(\delta z_{1})^2, (\delta z_2)^2, \dots, (\delta z_k)^2 ],
-\end{equation}
-and transform it back to physical space as well to obtain the covariance matrix of the model observables $\y_m$ at a given parameter point $\x$.
-
-The resulting emulator therefore predicts both a mean prediction $\y_m(\x)$ and an uncertainty covariance matrix $\Sigma_m(\x)$ which accounts for multiple sources of model and emulator uncertainty, including the truncation error expected from using a finite number of principal components $k < m$.
 \begin{fullpage}
@@
 \end{fullpage}
+
+The GPE's are essentially fancy interpolators applied to the model's training points and PCA transformed observables.
+Each GPE reports a mean value $z(\x)$ as well as an estimated error $\delta z(\x)$ which accounts for statistical noise in the training data and interpolation error between the design points.
+Once the GPE's are trained, we can predict the observables $\y_m$ at parameter point $\x$ by transforming the vector of principal components
+\begin{equation}
+  \mathbf{z}(\x) = (z_1(\x), z_2(\x), \dots, z_k(\x))
+\end{equation}
+\emph{back} to physical space.
+Similarly, we can construct the covariance matrix of the observables in PCA space,
+\begin{equation}
+  \cov(z_i, z_j) = \diag[(\delta z_{1})^2, (\delta z_2)^2, \dots, (\delta z_k)^2 ],
+\end{equation}
+and transform it back to physical space as well to obtain the covariance matrix of the model observables $\y_m$ at a given parameter point $\x$.
+
+The resulting emulator therefore predicts both a mean prediction $\y_m(\x)$ and an uncertainty covariance matrix $\Sigma_m(\x)$ which accounts for multiple sources of model and emulator uncertainty, including the truncation error expected from using a finite number of principal components $k < m$.
 The model covariance matrix $\Sigma_m$ includes so-called \emph{known}-unknowns such as statistical error and emulator interpolation error, but not \emph{unknown}-unknowns such as the overall validity of small-system hydrodynamics, i.e.\ things which lack a unified consensus or are difficult to quantify.
@@
 where $E$ subsumes all evidence from the $p$-Pb and Pb-Pb collision systems and $H_\x$ is our hypothesis that $\x$ equals the true parameters $\x_\star$.
-
-We then perform Markov-chain Monte Carlo (MCMC) importance sampling on the posterior distribution in Eqn.~\eqref{eq:bayes} to draw random samples for $P(H_\x | E)$, the estimate of the true model parameters given the model and the experimental data \cite{Goodman:2010en, FM:2013mc}.
+We then perform Markov-chain Monte Carlo (MCMC) importance sampling on the posterior distribution in Eq.~\eqref{eq:bayes} to draw random samples for $P(H_\x | E)$, the estimate of the true model parameters given the model and the experimental data \cite{Goodman:2010en, FM:2013mc}.
 For this we use an affine-invariant ensemble sampler which uses a large ensemble of interdependent walkers \cite{Goodman:2010en, FM:2013mc} and allow the MCMC chain to ``burn-in'' before generating \order{7} posterior samples.
@@
 Notice the large spread of the observables calculated at the training points (top row of each figure).
-The design is constructed to vary each parameter across a wide range of values, specified in Tbl.~\ref{tab:design}, and hence the corresponding model calculations are equally uncertain.
+The design is constructed to vary each parameter across a wide range of values, specified in Table~\ref{tab:design}, and hence the corresponding model calculations are equally uncertain.
 We also point out that there is considerably more variance in the $p$-Pb training data than the Pb-Pb training data.
@@
 The uncertainty in the posterior distribution could thus be improved by running the calibration with more design points or with a narrower range of parameter values to increase the density of the training points and reduce interpolation uncertainty.
-
-We now direct our attention to Fig.~\ref{fig:posterior} which shows the main result of this work, the posterior distribution of the model input parameters.
-Recall that the posterior $P(H_\x | E)$ is the probability that our hypothesis $H_\x$ for the true model parameters $\x_\star$ is correct, given the evidence $E$ provided by experiment.
-The present posterior has 15 dimensions, one for each parameter listed in Tbl.~\ref{tab:design}, and thus its joint distribution cannot be summarized by one figure alone.
-We therefore sample the distribution and histogram the samples to project the distribution onto one or two dimensions at a time.
-Each diagonal panel is the distribution of a \emph{single} model parameter (marginalized over all others), and each lower-diagonal panel is the joint distribution of a \emph{pair} of model parameters (marginalized over all others).
-We also report numeric estimates for each model parameter along the distribution diagonal.
-These estimates describe the marginalized distribution median and 90\% credible interval for each parameter.
-For example, the estimate $x=2.45_{-0.15}^{+0.20}$ denotes a median value $\tilde{x}=2.45$ and 90\% credible interval $2.30 < x < 2.65$.
-These posterior parameter estimates are also listed in Tbl.~\ref{tab:post_param} for clarity.
 
@@
 
-\begin{table}[t]
+We now direct our attention to Fig.~\ref{fig:posterior} which shows the main result of this work, the posterior distribution of the model input parameters.
+Recall that the posterior $P(H_\x | E)$ is the probability that our hypothesis $H_\x$ for the true model parameters $\x_\star$ is correct, given the evidence $E$ provided by experiment.
+The present posterior has 15 dimensions, one for each parameter listed in Table~\ref{tab:design}, and thus its joint distribution cannot be summarized by one figure alone.
+We therefore sample the distribution and histogram the samples to project the distribution onto one or two dimensions at a time.
+Each diagonal panel is the distribution of a \emph{single} model parameter (marginalized over all others), and each lower-diagonal panel is the joint distribution of a \emph{pair} of model parameters (marginalized over all others).
+We also report numeric estimates for each parameter's median value and 90\% credible interval and annotate their values along the distribution diagonal (see Table~\ref{tab:post_param}).
+For example, the fictitious estimate $x=2.45_{-0.15}^{+0.20}$ reports a median value $\tilde{x}=2.45$ and 90\% credible interval $2.30 < x < 2.65$.
+
+\begin{table}
   \caption{
@@
 It just so happens that the small range of $p$-values needed to describe this yield ratio are the same values needed to describe all of the calibration observables in the present study, \emph{and} numerous Pb-Pb observables at $\sqrts=2.76$~TeV \cite{Bernhard:2018hnz}.
-This work thus reaffirms an empirical scaling law reported by several previous studies \cite{Moreland:2014oya, Bernhard:2016tnd, Ke:2016jrd, Bernhard:2018hnz} for the initial transverse entropy density (or massless parton density) of an ultrarelativistic nuclear collision, given by
+This work thus reaffirms an empirical scaling law reported by several previous studies \cite{Moreland:2014oya, Bernhard:2016tnd, Ke:2016jrd, Bernhard:2018hnz} for the initial transverse entropy density (or massless parton density):
 \begin{equation}
@@
 We emphasize that this specific analytic form should not be interpreted too literally.
-For instance, a generalized mean described by $p=0.05$ is well within our 90\% credible interval, but it is not equal to the geometric mean in Eqn.~\ref{eq:geometric_mean}.
-This approximate form nevertheless has been shown to mimic the scaling behavior of several saturation-based theory calculations in high-energy QCD \cite{Bernhard:2016tnd}, and thus appears consistent with theoretical expectations.
-
-\begin{figure}[b]
+For instance, a generalized mean described by $p=0.05$ is well within our 90\% credible interval, but it is not equal to the geometric mean in Eq.~\eqref{eq:geometric_mean}.
+This approximate form nevertheless has been shown to mimic the scaling behavior of several saturation-based theory calculations in high-energy QCD \cite{Bernhard:2016tnd}, and thus appears consistent with general theoretical expectations.
+
+\begin{figure}
   \includegraphics{posterior_parton_number}
@@
 The posterior on the nucleon substructure parameter $\X=0.35_{-0.17}^{+0.22}$, on the other hand, is particularly sharply peaked.
-Recall that this parameter, defined in Eqn.~\eqref{eq:struct_param}, interpolates between the minimum and maximum widths of each constituent and hence
+Recall that this parameter, defined in Eq.~\eqref{eq:struct_param}, interpolates between the minimum and maximum widths of each constituent and hence
 two different limits for the granularity of the nucleon.
@@
 
-In this section, we compare several of our posterior estimates to those obtained from a similar Bayesian analysis \cite{Bernhard:2018hnz} which used an (almost) identical version of the present physics model.
-The only modeling difference between the two studies is the inclusion of nucleon substructure in this work which was absent in Ref.~\cite{Bernhard:2018hnz}.
-Several calibration details, however, are different as well.
-The present work uses a much smaller number of $p$-Pb and Pb-Pb observables at $\sqrts=5.02$~TeV (limited by availability), whereas Ref.~\cite{Bernhard:2018hnz} used a larger cross section of Pb-Pb experimental data at $\sqrts=2.76$ and 5.02~TeV.
+In this section, we compare several of our posterior estimates to those obtained from a similar Bayesian analysis in Ref.~\cite{Bernhard:2018hnz} which used an (almost) identical version of the present physics model.
+The only modeling difference is the inclusion of nucleon substructure in the present study which was absent in Ref.~\cite{Bernhard:2018hnz}.
+Several calibration details, however, are different between the two analyses.
+This work used a modest number of $p$-Pb and Pb-Pb observables at $\sqrts=5.02$~TeV (limited by availability), whereas Ref.~\cite{Bernhard:2018hnz} calibrated on a much larger number of Pb-Pb observables at $\sqrts=2.76$ and 5.02~TeV.
 
@@
 We verified the emulator and tested the accuracy of our physics model framework using a single set of high-probability parameters selected from the Bayesian posterior.
-These parameters, listed in Tbl.~\ref{tab:mode_params}, are the approximate ``best fit'' values of the calibrated model, commonly referred to as the \emph{maximum a posteriori} (MAP) estimate:
+These parameters, listed in Table~\ref{tab:mode_params}, are the approximate ``best fit'' values of the calibrated model, commonly referred to as the \emph{maximum a posteriori} (MAP) estimate:
 \begin{equation}
@@
 The bottom-right panel of Fig.~\ref{fig:obs_map} shows our model calculation for the four-particle elliptic flow cumultant $\vnk{2}{4}$ along with the measured data points from \mbox{ALICE} \cite{Adam:2016izf}.
-We see that the MAP estimate nicely describes the measured $\vnk{2}{4}$ data which is encouraging since this particular observable was never used to calibrate the posterior.
+We see that the MAP estimate nicely describes the measured $\vnk{2}{4}$ data which is encouraging since this particular observable was never used to calibrate the model.
 
@@
 \end{equation}
-The inner-average in Eqn.~\ref{eq:mean_pT_corr} runs over all pairs of particles $i,j$ in the same event, the outer average runs over all events in a given bin (centrality or multiplicity), and the symbol $\langle p_T \rangle$ denotes the usual mean transverse momentum of particles in the bin.
-The observable is typically presented in terms of the dimensionless ratio $\delta p_T / \langle p_T \rangle$, which quantifies the strength of the dynamical fluctuations in units of the average transverse momentum $\langle p_T \rangle$.
-
-We show the MAP estimate predictions for the $p$-Pb and Pb-Pb relative mean $p_T$ fluctuations $\delta p_T / \langle p_T \rangle$ at $\sqrts=5.02$~TeV in the third row of Fig.~\ref{fig:obs_map}.
-For the Pb-Pb system, we use centrality bins and for the $p$-Pb system we use the same relative multiplicity bins used for the $p$-Pb charged-particle mean $p_T$.
-The relative mean $p_T$ fluctuations have been shown to be particularly sensitive to the existence of nucleon substructure \cite{Bozek:2017elk}, and thus it would be interesting to ultimately include this observable in the calibration when the data becomes available.
 
@@
 
+\noindent The inner-average in Eq.~\eqref{eq:mean_pT_corr} runs over all pairs of particles $i,j$ in the same event, the outer average runs over all events in a given bin (centrality or multiplicity), and the symbol $\langle p_T \rangle$ denotes the usual mean transverse momentum of particles in the bin.
+The observable is typically presented in terms of the dimensionless ratio $\delta p_T / \langle p_T \rangle$, which quantifies the strength of the dynamical fluctuations in units of the average transverse momentum $\langle p_T \rangle$.
+
+We show the MAP estimate predictions for the $p$-Pb and Pb-Pb relative mean $p_T$ fluctuations $\delta p_T / \langle p_T \rangle$ at $\sqrts=5.02$~TeV in the third row of Fig.~\ref{fig:obs_map}.
+For the Pb-Pb system, we use centrality bins and for the $p$-Pb system we use the same relative multiplicity bins used for the $p$-Pb charged-particle mean $p_T$.
+The relative mean $p_T$ fluctuations have been shown to be particularly sensitive to the existence of nucleon substructure \cite{Bozek:2017elk}, and thus it would be interesting to ultimately include this observable in the calibration when the data becomes available.
+
 \begin{table}
@@
     High-probability parameters selected from the posterior distribution and used to generate Fig.~\ref{fig:obs_map}.
-    The posterior distribution on the particlization temperature $T_\text{switch}$ is flat (agnostic), so we fix it's value using Ref.~\cite{Bernhard:2018hnz}.
+    The posterior distribution on the particlization temperature $\Tsw$ is flat (agnostic), so we fix it's value using Ref.~\cite{Bernhard:2018hnz}.
   }
@@
       $v$      & 0.43 fm        & $\zeta/s$ $T_0$   & 175 MeV        \\
-      $\dmin$  & 0.81 fm        & $T_\text{switch}$ & 151 MeV        \\
+      $\dmin$  & 0.81 fm        & $\Tsw$            & 151 MeV        \\
       $\taufs$ & 0.37 \fmc
@@
 
-We emphasize that multiple aspects of the two analyses are different such as the collision systems and beam energies considered, the observables which were included in each calibration, and the existence of nucleon substructure to name a few.
-Thus we can only speculate what might have lead to the large difference in the symmetric flow cumulants.
+We emphasize that multiple aspects of the two analyses are different such as the collision systems and beam energies considered, the observables which were included in each calibration, and the existence of nucleon substructure in the model.
+Thus we can only speculate what might have caused the large difference in the MAP estimate for the symmetric flow cumulants.
 Two reasonable culprits would be the inclusion of nucleon substructure and the large difference in the preferred pre-equilibrium free-streaming time determined by the two studies.
-Identifying the source of the effect would be an interesting study for a future project.
 
@@
 
-Second, we obtained a posterior distribution for the model input parameters in Fig.~\ref{fig:posterior} and reported quantitative estimates for these parameters in Tbl.~\ref{tab:post_param}.
+Second, we obtained a posterior distribution for the model input parameters in Fig.~\ref{fig:posterior} and reported quantitative estimates for these parameters in Table~\ref{tab:post_param}.
 Here we summarize our key findings about the model:
@@
   \item
-    Using \trento\ initial conditions, we find that initial entropy deposition (or particle production) scales like the geometric mean of local participant nuclear densities, in agreement with a number of previous estimates \cite{Moreland:2014oya, Bernhard:2016tnd, Bernhard:2018hnz}.
+    Using \trento\ initial conditions, we find that initial entropy deposition (or massless noninteracting particle production) scales like the geometric mean of local participant nuclear densities, in agreement with a number of previous estimates \cite{Moreland:2014oya, Bernhard:2016tnd, Bernhard:2018hnz}.
   \item
@@
     Bulk observables in $p$-Pb and Pb-Pb collisions prefer larger nucleons, similar in size to the proton's RMS electric charge radius.
-    This agrees with a similar Bayesian analysis which was calibrated to Pb-Pb bulk observables at $\sqrts=2.76$ and 5.02~TeV using a hybrid model without nucleon substructure \cite{Bernhard:2018hnz}.
+    This agrees with a similar Bayesian analysis which was calibrated to Pb-Pb observables at $\sqrts=2.76$ and 5.02~TeV using a hybrid model without nucleon substructure \cite{Bernhard:2018hnz}.
     We note that our result is in significant tension with an estimate for the effective nucleon width based on the gluon distribution extracted from HERA data \cite{Rezaeian:2012ji}.
@@
     We obtain an estimate $v=0.47_{-0.15}^{+0.20}$~fm for the Gaussian width of the constituent hot spots inside each nucleon.
-    This is significantly larger than the length scales typically associated with nucleon substructure, and it may help explain the seemingly unreasonable success of hydrodynamics in small collision systems.
+    This is considerably larger than the length scales typically associated with nucleon substructure, and it may help explain the seemingly unreasonable success of hydrodynamics in small collision systems.
   \item
@@
 
-\begin{figure}[b]
+\begin{figure}[t]
   \includegraphics{validation_example_vert}
@@
     \label{fig:validation_example}
-    Example emulator validation for one observable, the Pb-Pb charged particle yield $d\nch/d\eta$ in the 20-30\% centrality class.
+    Example emulator validation for one observable, the Pb-Pb charged-particle yield $d\nch/d\eta$ in the 20-30\% centrality class.
     We use the k-fold cross validation method (explained in the text) to partition the model inputs $X$ and outputs $Y$ into training and validation data.
@@
     The boxes are 50\% interquartile ranges and whiskers are the 90\% interquantiles.
-    The bottom row of each figure is the RMS fractional error defined by Eqn.~\eqref{eq:frac_error}.
+    The bottom row of each figure is the RMS fractional error defined by Eq.~\eqref{eq:frac_error}.
   }

### 2018-08-06 src/plots.py

@@ -2163,7 +2163,7 @@ def correlation_matrices(system=default_system):
         for s in ax.spines.values():
             s.set_visible(False)
 
-    axes[0].tick_params(labelleft=False)
+    axes[0].tick_params(labelleft=True)
     for t in axes[1].get_yticklabels():
         t.set_horizontalalignment('center')
         t.set_x(-.05)

### 2018-08-06 paper/substructure.tex
Fix Oxford comma inconsistencies

@@
 Specifically, we model the dynamics of both collision systems using initial conditions with parametric nucleon substructure, a pre-equilibrium free-streaming stage, event-by-event viscous hydrodynamics, and a microscopic hadronic afterburner.
-Free parameters of the model which describe the initial state and QGP medium are then simultaneously calibrated to fit charged particle yields, mean $p_T$ and flow cumulants.
+Free parameters of the model which describe the initial state and QGP medium are then simultaneously calibrated to fit charged particle yields, mean $p_T$, and flow cumulants.
 We argue that the global agreement of the calibrated model with the experimental data strongly supports the existence of hydrodynamic flow in small collision systems at ultrarelativistic energies, and that the flow produced develops at length scales smaller than a single proton.
@@
 We begin, in Sec.~\ref{sec:model}, by constructing a nuclear collision model for $p$-Pb and Pb-Pb collisions using initial conditions with parametric nucleon substructure, and transport dynamics described by a pre-equilibrium free-streaming stage, viscous hydrodynamics and microscopic Boltzmann transport.
-In Sec.~\ref{sec:calibration}, we calibrate free parameters of the model to fit charged particle yields, mean $p_T$ and anisotropic flow cumulants of \emph{both} collision systems at $\sqrts=5.02$~TeV, and finally, in Secs.~\ref{sec:results} and \ref{sec:summary}, we present posterior results for the model input parameters and comment on the implications for hydrodynamic descriptions of small collision systems.
+In Sec.~\ref{sec:calibration}, we calibrate free parameters of the model to fit charged particle yields, mean $p_T$, and anisotropic flow cumulants of \emph{both} collision systems at $\sqrts=5.02$~TeV, and finally, in Secs.~\ref{sec:results} and \ref{sec:summary}, we present posterior results for the model input parameters and comment on the implications for hydrodynamic descriptions of small collision systems.
 
@@
 \end{equation}
-describes the eikonal overlap of the two proton wave packets at fixed impact parameter $b$ where coordinates $(x, y)$ lie in the transverse plane and $z$ is parallel to the beam axis.
+describes the eikonal overlap of the two proton wave packets at fixed impact parameter $b$, where coordinates $(x, y)$ lie in the transverse plane, and $z$ is parallel to the beam axis.
 Here we assume that each proton is comprised of smaller constituents---e.g.\ valence quarks, sea quarks, and small-x gluons---which may collide to produce secondary particles and contribute to the observed inelastic proton-proton cross section.
@@
 This parametrization introduces a dimensionless parameter $p$ which varies the scaling behavior of initial entropy deposition at midrapidity.
-For certain discrete values of $p$, it reduces to well known functional forms such as the arithmetic, geometric and harmonic means:
+For certain discrete values of $p$, it reduces to well known functional forms such as the arithmetic, geometric, and harmonic means:
 \begin{equation}
@@
 
-This procedure provides initial values for $T^{\mu\nu}$, $u^\mu$, $\Pi$ and $\pi^{\mu\nu}$ which conserve energy and are consistent with the underlying hydrodynamic equation of state.
+This procedure provides initial values for $T^{\mu\nu}$, $u^\mu$, $\Pi$, and $\pi^{\mu\nu}$ which conserve energy and are consistent with the underlying hydrodynamic equation of state.
 We therefore expect it to provide a more realistic description of the initial stages of the collision as compared to a previous study using the \trento\ initial condition model which set $\Pi$, $\pi^{\mu\nu}$ and $u^\mu$ initially to zero \cite{Bernhard:2016tnd}.
@@
 \end{equation}
-where $\eta/s$ min, slope and curvature are tunable parameters and $T_c=0.154$~GeV is the pseudocritical transition temperature of the HotQCD EoS.
+where $\eta/s$ min, slope, and curvature are tunable parameters, and $T_c=0.154$~GeV is the pseudocritical transition temperature of the HotQCD EoS.
 While for the specific bulk viscosity $\zeta/s$, we use an unnormalized Cauchy distribution
@@
 \end{equation}
-Here $m_0$ is the resonance's Breit-Wigner mass and $\Gamma(m)$ is its mass-dependent width, for which we use a simple form:
+Here $m_0$ is the resonance's Breit-Wigner mass, and $\Gamma(m)$ is its mass-dependent width, for which we use a simple form:
 \begin{equation}
@@
 \end{equation}
-where $\Gamma_0$ is the usual Breit-Wigner width and $m_\mathrm{min}$ is a production threshold equal to the total mass of the lightest decay products.
+where $\Gamma_0$ is the usual Breit-Wigner width, and $m_\mathrm{min}$ is a production threshold equal to the total mass of the lightest decay products.
 We tabulate the values of $\{\Gamma_0, m_0, m_\mathrm{min}\}$ for all particles and sample the masses of each particle during particlization \cite{PDG:2017}.
@@
     \label{fig:nucleon_schematic}
-    Schematic illustrating the constituent sampling radius $r$, constituent width $v$ and nucleon width parameter $w = \sqrt{r^2 + v^2}$ for a nucleon with $n_c=3$ constituents.
+    Schematic illustrating the constituent sampling radius $r$, constituent width $v$, and nucleon width parameter $w = \sqrt{r^2 + v^2}$ for a nucleon with $n_c=3$ constituents.
   }
@@
   \item Charged particle mean transverse momentum $\langle p_T \rangle$ ($0.15 < p_T < 10$~GeV).
-  \item Identified particle mean transverse momentum $\langle p_T \rangle$ of pions, kaons and protons.
+  \item Identified particle mean transverse momentum $\langle p_T \rangle$ of pions, kaons, and protons.
   \item Mean transverse momentum fluctuations $\delta p_T / \langle p_T \rangle$ (charged particles, $0.15 < p_T < 2.0$~GeV).
@@
 \end{equation}
-The statistical errors in $\Sigma_e^\text{stat}$ are uncorrelated and thus its covariance matrix is diagonal:
+The statistical errors in $\Sigma_e^\text{stat}$ are uncorrelated, and thus its covariance matrix is diagonal:
 \begin{equation}
@@
 \end{equation}
-where $b_i$ and $b_j$ are the midpoints of two observable bins of a single type (centrality or relative multiplicity) and $l$ is a correlation length which describes how quickly the observable bins decorrelate as the distance between the bins increases.
+where $b_i$ and $b_j$ are the midpoints of two observable bins of a single type (centrality or relative multiplicity), and $l$ is a correlation length which describes how quickly the observable bins decorrelate as the distance between the bins increases.
 We use centrality correlation lengths $l=100$ for all of the centrality binned Pb-Pb observables and $l=30$ for the centrality binned $p$-Pb charged particle yield $d\nch/d\eta$.
@@
 We also point out that there is considerably more variance in the $p$-Pb training data than the Pb-Pb training data.
-The $p$-Pb yields, mean $p_T$ and flow cumulants all vary wildly within the chosen parameter ranges.
+The $p$-Pb yields, mean $p_T$, and flow cumulants all vary wildly within the chosen parameter ranges.
 For instance, we can turn the $p$-Pb flows completely \emph{off} with suitably chosen parameters which is not possible in the Pb-Pb system.
@@
 We caution that this parameter $w$ is \emph{not} the RMS radius of our nucleons due to idiosyncrasies of our constituent sampling procedure.
-We can, however, easily calculate the RMS radius for a specific nucleon width $w$, constituent width $v$ and constituent number $n_c$.
-For example, the single highest posterior probability region of the parameter space prefers a nucleon width $w=0.92$~fm, constituent width $v=0.43$~fm and constituent number $n_c=6$.
+We can, however, easily calculate the RMS radius for a specific nucleon width $w$, constituent width $v$, and constituent number $n_c$.
+For example, the single highest posterior probability region of the parameter space prefers a nucleon width $w=0.92$~fm, constituent width $v=0.43$~fm, and constituent number $n_c=6$.
 The corresponding RMS nucleon radius for these parameters is $R_n = 0.86$~fm, conspicuously close the proton's RMS electric charge radius $R_p = 0.879(8)$~fm \cite{Bernauer:2010wm}.
@@
 Similarly, the results would greatly benefit from additional beam energies and collision systems.
-Notably absent is data from RHIC which includes measurements for $p$-Au, $d$-Au, $^3$He-Au, Cu-Cu, Cu-Au, Au-Au and \mbox{U-U} collision systems at $\sqrts=200$~GeV.
+Notably absent is data from RHIC which includes measurements for $p$-Au, $d$-Au, $^3$He-Au, Cu-Cu, Cu-Au, Au-Au, and \mbox{U-U} collision systems at $\sqrts=200$~GeV.
 This data would undoubtedly provide an additional constraint on the model parameters and would enable more stringent tests of the calibrated model predictions.

### 2018-08-10 paper/substructure.tex
Fix some of the bibtex entries.

@@
   using a hybrid model with nucleon substructure \\
-  calibrated to p-Pb and Pb-Pb collisions at \texorpdfstring{$\mathbf{\sqrts=5.02}$}{}~TeV
+  calibrated to $\boldsymbol p$-Pb and Pb-Pb collisions at \texorpdfstring{$\mathbf{\sqrts=5.02}$}{}~TeV
 }
@@
 Hydrodynamic models of ultrarelativistic nuclear collisions are complicated by a number of theoretical unknowns, including the detailed geometry of the QGP initial conditions, the strength and duration of pre-equilibrium dynamics, the temperature dependence of QGP transport coefficients, and the boundaries of hydrodynamic applicability \cite{Niemi:2014lha, deSouza:2015ena, Ollitrault:2012cm, Song:2012ua}.
-In general, these theoretical uncertainties tend to grow with decreasing system size, where emergent physics at sub-fermi length scales becomes important to describe bulk properties of the produced system.
+In general, these theoretical uncertainties tend to grow with decreasing system size, where emergent physics at sub-femtometer length scales becomes important to describe bulk properties of the produced system.
 

### 2018-10-10 paper/substructure.tex
save revision work on the manuscript.

@@
 \title{
-  Estimating initial state and quark-gluon plasma medium properties \\
-  using a hybrid model with nucleon substructure \\
-  calibrated to $\boldsymbol p$-Pb and Pb-Pb collisions at \texorpdfstring{$\mathbf{\sqrts=5.02}$}{}~TeV
+  Bayesian calibration of a hybrid nuclear collision model\\using $\boldsymbol p$-Pb and Pb-Pb data from the LHC
 }
@@
 
-We model the QGP initial state in $p$-Pb and Pb-Pb collisions at $\sqrts=5.02$~TeV using a simple parametric form for boost-invariant entropy deposition employed by the \trento\ model \cite{Moreland:2014oya}.
+We model the initial state of $p$-Pb and Pb-Pb collisions at $\sqrts=5.02$~TeV using the boost-invariant \trento\ model \cite{Moreland:2014oya}.
 Generally speaking, the initial three-dimensional distribution of matter produced in relativistic nuclear collisions is \emph{not} boost-invariant; longitudinal entropy deposition fluctuates both locally point-to-point in the transverse plane as well as globally event-by-event due to asymmetries in the sampled density of participant matter \cite{Ke:2016jrd, Bozek:2010vz}.
@@
 The \trento\ model operates in the ultrarelativistic limit with a Lorentz factor $\gamma \gg 1$ such that each nucleus appears as a thin sheet of nuclear density in the laboratory frame.
-The sheets of colliding nuclear density penetrate and pass through each other in proper time $\Delta \tau \approx D_\text{nucl} / (\gamma\, \beta_z)$ in the laboratory frame, where $D_\text{nucl}$ is the diameter of the nucleus in its rest frame, $\gamma$ is the usual Lorentz factor of the accelerated ions, and $\beta_z$ is their velocity along the beam axis.
-The resulting nuclear overlap time $\Delta \tau \lesssim 0.1\ \fmc$ at top RHIC and LHC energies, and thus it is safe to neglect the initial transverse dynamics which occur while the nuclei pass through each other.
-We therefore assume that the collision produces all secondary particles at uniform proper time $\tau = 0^+\ \fmc$, and that it deposits entropy (energy) at midrapidity which is a function of the locally varying beam integrated density of each nucleus.
-
-Consider the collision of two protons $A$, $B$ with three-dimensional densities $\rho_{A,B}$ in their local rest frames.
+The sheets of colliding nuclear density penetrate and pass through each other in time $\Delta t \approx D_\text{nucl} / (\gamma\, \beta_z)$ in the laboratory frame, where $D_\text{nucl}$ is the diameter of the nucleus in its rest frame, $\gamma$ is the usual Lorentz factor of the accelerated ions, and $\beta_z$ is their velocity along the beam axis.
+The resulting nuclear overlap time $\Delta t \lesssim 0.1\ \fmc$ at top RHIC and LHC energies, and thus it is safe to neglect the initial transverse dynamics which occur while the nuclei pass through each other.
+We therefore assume that the collision produces all secondary particles at uniform proper time $\tau_0 \ll 1~\fmc$, and that it deposits energy at midrapidity which is a function of the locally varying beam integrated density of each nucleus.
+
+Consider the collision of two protons, labeled $A$ and $B$, with three-dimensional densities $\rho_A$ and $\rho_B$ in their local rest frames.
 The proton-proton overlap function
@@
   \label{eq:tpp}
-  T_{pp}(b) \equiv \int dx\, dy \int dz\, \rho_A(\x) \int dz\, \rho_B(\x + \mathbf{b}),
-\end{equation}
-describes the eikonal overlap of the two proton wave packets at fixed impact parameter $b$, where coordinates $(x, y)$ lie in the transverse plane, and $z$ is parallel to the beam axis.
+  T_{pp}(b) \equiv \int dx\, dy\, dz\, \rho_A(x, y, z)\, \rho_B(x - b, y, z),
+\end{equation}
+describes the eikonal overlap of the two proton wave packets at fixed impact parameter $b$, where coordinate $(x, y)$ lies in the transverse plane and $z$ is parallel to the beam axis.
 Here we assume that each proton is comprised of smaller constituents---e.g.\ valence quarks, sea quarks, and small-x gluons---which may collide to produce secondary particles and contribute to the observed inelastic proton-proton cross section.
@@
   \label{eq:pcoll}
-  P_\mathrm{coll} = 1 - \exp[-\sigma_\mathrm{eff}\, T_{pp}(b)], \\
+  P_\mathrm{coll}(b) = 1 - \exp[-\sigma_\mathrm{eff}\, T_{pp}(b)], \\
 \end{equation}
@@
 One common implementation is a superposition of three valence quarks, typically described by Gaussian or exponential form factors \cite{Welsh:2016siu, Bozek:2017jog, Schenke:2014zha, Schlichting:2014ipa, Adare:2015bua, Broniowski:2016pvx}.
-The corresponding proton density $\rho(\mathbf{x})$ is then assumed to be that of predominantly small-x gluons, seeded by the distribution of color charge in each of the three valence quarks.
+The corresponding proton density is then assumed to be that of predominantly small-x gluons, seeded by the distribution of color charge in each of the three valence quarks.
 
@@
 In this work, we pursue a less restrictive and more parametric description of the proton where the number of substructure degrees of freedom are uncertain as depicted in Fig.~\ref{fig:substructure}.
-We model each nucleon's density $\rho_{A,B}$ as a sum of $n_c$ independent constituents
+We model each nucleon's three-dimensional density $\rho_{A,B}(\x)$ as a weighted sum of $n_c$ independent constituents,
 \begin{equation}
   \label{eq:rho}
-  \rho_{A, B}(\x) = \frac{1}{n_c} \sum\limits_{i=1}^{n_c} \rho_c(\mathbf{x} - \mathbf{x_i}),
-\end{equation}
-where each constituent density $\rho_c$ is described by a Gaussian distribution of width $v$
-\begin{equation}
-  \label{eq:constituent_density}
-  \rho_c(\mathbf{x}) = \frac{1}{(2\pi v^2)^{3/2}} \exp\left(-\frac{\x^2}{2 v^2}\right),
-\end{equation}
-and each constituent's position $\mathbf{x_i}$ in Eq.~\eqref{eq:rho} is sampled from a Gaussian radial distribution with standard deviation $r$.
-The effect of this additional nucleon substructure on the nuclear thickness functions is visible in Fig.~\ref{fig:thickness}
+  \rho_{A, B}(\x) = \frac{1}{n_c} \sum\limits_{i=1}^{n_c} \frac{\gamma_i}{(2 \pi v^2)^{3/2}} \exp\left(-\frac{|\x - \x_i|^2}{2 v^2}\right),
+\end{equation}
+where each constituent density is assumed to be a Gaussian of width $v$, weighted by a gamma random variable $\gamma_i$ with unit mean and variance $1/k$.
+These \emph{ad hoc} gamma random weights introduce additional multiplicity fluctuations in order to reproduce the large fluctuations observed in experimental proton-proton collisions.
+We sample the constituent positions $\x_i$ independently (ignoring correlations) from a Gaussian radial distribution,
+\begin{equation}
+  \label{eq:radial_dist}
+  P(\x_i) = \frac{1}{(2\pi r^2)^{3/2}} \exp\left(-\frac{|\x_i|^2}{2 r^2}\right),
+\end{equation}
+parametrized by a sampling radius $r$ which varies the average separation of the constituents.
+Figure~\ref{fig:thickness} shows the effect of this additional nucleon substructure on the nuclear thickness functions.
 
@@
 
-Within the eikonal approximation, the initial entropy deposited at midrapidity and proper time $\tau=0^+\ \fmc$ is some function
-\begin{equation}
-  f: \T_A, \T_B \mapsto \frac{dS}{d^2x_\perp d\eta} \bigg\vert_{\eta=0},
+The initial energy density $e$ deposited at midrapidity and proper time $\tau_0 \approx 0~\fmc$ is some function
+\begin{equation}
+  e(\x_\perp, \eta_s=0, \tau_0) = f(\T_A, \T_B)
 \end{equation}
@@
 \begin{equation}
-  \frac{dS}{d^2x_\perp d\eta}\,\bigg\vert_{\eta=0} \propto \frac{\T_A + \T_B}{2},
+  f = \text{const} \times \frac{\T_A + \T_B}{2},
 \end{equation}

### 2018-10-10 src/design.py

@@ -29,6 +29,8 @@ import numpy as np
 
 from . import cachedir, parse_system
 
+import matplotlib.pyplot as plt
+
 
 """
 Remove outlier design points. These points are near the edge of the
@@ -171,6 +173,23 @@ class Design:
             npoints=npoints, ndim=self.ndim, seed=seed
         )
 
+        # alter several design properties (revised manuscript)
+        self.array[:, 5] = .2 + self.array[:, 5]*(self.array[:, 3] - .2)
+        self.array[:, 3] = np.sqrt(self.array[:, 3]**2 - self.array[:, 5]**2)
+
+        design_changes = [
+            (3, 'sampling_radius', r'$r\ [\mathrm{fm}]$', (0.0, 1.2)),
+            (5, 'parton_width', r'$v\ [\mathrm{fm}]$', (0.2, 1.2))
+        ]
+
+        for index, key, label, (range_low, range_high) in design_changes:
+            self.keys[index] = key
+            self.labels[index] = label
+            self.range[index] = (range_low, range_high)
+            self.min[index] = range_low
+            self.max[index] = range_high
+
+        # drop bad points
         keep = [n not in bad_points for n in range(npoints)]
         self.array = self.array[keep]
 
@@ -183,7 +202,6 @@ class Design:
         logging.debug(
             'removed outlier design points: {}'.format(bad_points)
         )
-
 
     def __array__(self):
         return self.array
@@ -274,7 +292,8 @@ def main():
     args = parser.parse_args()
 
     for system in systems:
-        Design(system, validation=False).write_files(args.inputs_dir)
+        #Design(system, validation=False).write_files(args.inputs_dir)
+        Design(system, validation=False)
 
     logging.info('wrote all files to %s', args.inputs_dir)
 

### 2018-10-10 src/model.py

@@ -244,7 +244,7 @@ class ModelData:
             trigger = events['trigger']
             minbias = (0, float('inf'))
 
-            events = correct_yield(events)
+            #events = correct_yield(events)
 
             if bin_type == 'cent':
                 """

### 2018-10-10 src/plots.py

@@ -1005,6 +1005,7 @@ def _posterior(
         interp = PchipInterpolator(x, y)
         x = np.linspace(x[0], x[-1], 10*x.size)
         y = interp(x)
+
         ax.plot(x, y, linewidth=1, color=line_color)
         ax.fill_between(x, lim[0], y, color=fill_color, zorder=-10)
 
@@ -1213,15 +1214,6 @@ def posterior_freestreaming():
         'tau_fs',
         'Free streaming time [fm/$c$]',
         [.1, .8, 1.5]
-    )
-
-
-@plot
-def posterior_structure():
-    posterior_parameter(
-        'parton_struct',
-        '$\chi_\mathrm{struct}$',
-        [0, .5, 1]
     )
 
 
@@ -2411,19 +2403,15 @@ def entropy_scaling():
     set_tight(fig)
 
 
-def rms_nucleon_width(args, samples=10**3):
+def rms_nucleon_width(number, radius, width, samples=10**3):
     """
     Find center-of-mass root-mean-square radius for an ensemble of
     sampled protons with nucleon substructure.
 
     """
-    # unpack arguments
-    number, radius, struct = args
-    width = .2 + struct*(radius - .2)
-
     # sample constituent positions
     positions = np.random.normal(
-            scale=np.sqrt(radius**2 - width**2),
+            scale=radius,
             size=2*number*samples
             ).reshape(number, 2, -1)
 
@@ -2444,6 +2432,11 @@ def rms_nucleon_width(args, samples=10**3):
 
     # return rms radius in com frame
     return np.sqrt(np.average(r**2, weights=rho))
+
+@plot
+def rms_width():
+    R = rms_nucleon_width(3, .84, .3)
+    print(R)
 
 
 def correct_widths(number_values, width_values, struct_values):
@@ -2502,47 +2495,23 @@ def posterior_proton_shape():
 
     """
     chain = mcmc.Chain()
-    nucleon_radius, parton_number, parton_struct = chain.load(
-        'nucleon_width', 'parton_number',  'parton_struct'
+    sampling_radius, parton_number, parton_width = chain.load(
+        'sampling_radius', 'parton_number', 'parton_width'
     ).T
 
-    min_width = .2
-    parton_width = min_width + parton_struct*(nucleon_radius - min_width)
-
[…]

### 2018-10-11 paper/paper.tex
Work on paper revisions.

--- 2018-10-10 paper/substructure.tex
@@
 \title{
-  Bayesian calibration of a hybrid nuclear collision model\\using $\boldsymbol p$-Pb and Pb-Pb data from the LHC
+  Bayesian calibration of a hybrid nuclear collision model\\
+  using $\boldsymbol p$-Pb and Pb-Pb data from the LHC
 }
@@
 The resulting nuclear overlap time $\Delta t \lesssim 0.1\ \fmc$ at top RHIC and LHC energies, and thus it is safe to neglect the initial transverse dynamics which occur while the nuclei pass through each other.
-We therefore assume that the collision produces all secondary particles at uniform proper time $\tau_0 \ll 1~\fmc$, and that it deposits energy at midrapidity which is a function of the locally varying beam integrated density of each nucleus.
+We therefore assume that the collision produces all secondary particles at uniform proper time $\tau_0 \ll 1~\fmc$, and that it deposits energy at midrapidity which is some function of the locally varying beam integrated density of each nucleus.
 
@@
 
-In this work, we pursue a less restrictive and more parametric description of the proton where the number of substructure degrees of freedom are uncertain as depicted in Fig.~\ref{fig:substructure}.
-We model each nucleon's three-dimensional density $\rho_{A,B}(\x)$ as a weighted sum of $n_c$ independent constituents,
+In this work, we pursue a less restrictive and more parametric description of the proton where the number of substructure degrees of freedom is uncertain as depicted in Fig.~\ref{fig:substructure}.
+We model each nucleon's three-dimensional density $\rho_{A,B}$ as a sum of $n_c$ independent constituents,
 \begin{equation}
   \label{eq:rho}
-  \rho_{A, B}(\x) = \frac{1}{n_c} \sum\limits_{i=1}^{n_c} \frac{\gamma_i}{(2 \pi v^2)^{3/2}} \exp\left(-\frac{|\x - \x_i|^2}{2 v^2}\right),
-\end{equation}
-where each constituent density is assumed to be a Gaussian of width $v$, weighted by a gamma random variable $\gamma_i$ with unit mean and variance $1/k$.
-These \emph{ad hoc} gamma random weights introduce additional multiplicity fluctuations in order to reproduce the large fluctuations observed in experimental proton-proton collisions.
-We sample the constituent positions $\x_i$ independently (ignoring correlations) from a Gaussian radial distribution,
+  \rho_{A, B}(\x) = \frac{1}{n_c} \sum\limits_{i=1}^{n_c} \rho_c(\x - \x_i),
+\end{equation}
+where each constituent density $\rho_c$ is described by a Gaussian distribution,
+\begin{equation}
+  \rho_c(\x) = \frac{1}{(2 \pi v^2)^{3/2}} \exp\left(- \frac{|\x|^2}{2 v^2}\right).
+\end{equation}
+of width $v$.
+The constituent positions $\x_i$ are sampled independently (ignoring correlations) from a Gaussian radial distribution,
 \begin{equation}
   \label{eq:radial_dist}
-  P(\x_i) = \frac{1}{(2\pi r^2)^{3/2}} \exp\left(-\frac{|\x_i|^2}{2 r^2}\right),
-\end{equation}
-parametrized by a sampling radius $r$ which varies the average separation of the constituents.
-Figure~\ref{fig:thickness} shows the effect of this additional nucleon substructure on the nuclear thickness functions.
+  P(\x_i; \x_{A,B}) = \frac{1}{(2\pi r^2)^{3/2}} \exp\left(-\frac{|\x_i - \x_{A,B}|^2}{2 r^2}\right),
+\end{equation}
+where $\x_{A,B}$ is the position of each proton, and $r$ is a parameter which varies the sampling radius of the constituent positions about each proton position.
+As a matter of convenience, we sample the proton positions \emph{before} determining the constituent positions.
+This leads to small discrepancies between the designated proton positions $\x_{A,B}$, and the actual position of each proton's center-of-mass, owing to fluctuations in the constituent positions $\x_i$.
+The sampling radius $r$ should thus be interpreted with care; it is \emph{not} the root-mean-square radius of the sampled nucleons.
 
@@
 
-The two protons $A$, $B$ are assigned a random impact parameter, and Eq.~\eqref{eq:pcoll} is used to sample their inelastic collision probability $P_\mathrm{coll}(b)$.
-Note that this proton-proton inelastic collision probability has no direct knowledge of the individual constituent degrees of freedom; it is only \emph{indirectly} sensitive through the geometry of $\rho_{A, B}$ which depends on each of the constituent positions.
+We assign the two protons $A$, $B$ a random impact parameter, and sample their inelastic collision probability using Eq.~\eqref{eq:pcoll}.
+Assuming they collide, we assign each proton a fluctuated thickness function,
+\begin{equation}
+  \label{eq:fluctuated_thick}
+  \T_{A, B}(x, y) \equiv \int dz\, \tilde{\rho}_{A,B}(x \pm b/2, y, z),
+\end{equation}
+which projects each fluctuated proton density $\tilde{\rho}_{A,B}$ onto the transverse plane $(x, y)$ centered at $z=0$.
+
+Note, if any part of the proton collides, the entire proton contributes to inelastic particle production.
+There is no concept of individually wounded constituents; the entire proton either collides (or misses) as a singular connected object.
 This is an important distinction between the present model and a similar nucleon substructure implementation known as the participant or ``wounded'' quark model which allows for a subset of quarks (constituents) to participate inside a single nucleon \cite{ANISOVICH1978477, Broniowski:2016pvx}.
@@
 We correspondingly require that the nucleons in Eq.~\eqref{eq:rho} participate as singular objects, such that all spectator matter discarded by the simulation is appropriately color-neutral and inert.
-
-Assuming our two protons collide at the sampled impact parameter $\mathbf{b}$, we assign each a \emph{fluctuated} thickness
-\begin{equation}
-  \label{eq:fluctuated_thick}
-  \T_{A, B}(\x) \equiv \int dz\, \frac{1}{n_c} \sum\limits_{i=1}^{n_c} \gamma_i\, \rho_c \,(\mathbf{x} - \mathbf{x_i} \pm \mathbf{b}/2),
-\end{equation}
-equal to the beam-integrated proton density in Eq.~\eqref{eq:rho}, with each constituent shifted by the appropriate impact parameter offset, and multiplied by a gamma random variable $\gamma_i$ with unit mean and variance $1/k$.
-These \emph{ad hoc} gamma random weights are necessary to describe the large fluctuations observed in minimum bias proton-proton collisions, although their exact physical origin is not well understood.
 
@@
 
-When applying the model to heavy-ions, we sample nucleon positions from a Woods-Saxon density distribution subject to a minimum distance criteria $|\mathbf{x_i} - \mathbf{x_j}| > d_\mathrm{min}$ between all pairs of nucleons $i$, $j$.
+When applying the model to heavy-ions, we sample nucleon positions from a Woods-Saxon density distribution subject to a minimum distance criteria $|\x_i - \x_j| > d_\mathrm{min}$ between all pairs of nucleons $i$, $j$.
 The minimum distance algorithm, first described in Ref.~\cite{Bernhard:2018hnz}, uses a simple trick to resample the nucleon positions without modifying the target Woods-Saxon radial distribution.
@@
 Similar methods could be used to model correlations between individual constituents inside each nucleon, although the numerical implementation would be somewhat tedious.
+
+Figure~\ref{fig:thickness} shows the qualitative effect of nucleon substructure on a generic lead nuclear thickness function.
 
@@
 
-\bibliography{substructure}
+\bibliography{references}
 

### 2018-10-11 src/design.py

@@ -40,7 +40,7 @@ design and produce far too few particles to be physically reasonable.
 bad_points = [
     129, 392, 143, 273, 401, 276, 405, 153, 156, 285, 414, 159, 162, 418, 163,
     421, 422, 296, 298, 429, 301, 308, 183, 184, 312, 442, 446, 319, 451, 453,
-    326, 199, 457, 330, 459, 460, 461, 334, 203, 464, 466, 83, 211, 338, 471,
+    326, 199, 457, 330, 459, 460, 461, 334, 203, 464, 466,  83, 211, 338, 471,
     472, 473, 346, 223, 355, 357, 360, 234, 363, 492, 107, 236, 367, 235, 498,
     115, 242, 376, 249, 378
 ]
@@ -173,7 +173,10 @@ class Design:
             npoints=npoints, ndim=self.ndim, seed=seed
         )
 
-        # alter several design properties (revised manuscript)
+        # Version 1 of the manuscript parametrized nucleon substructure using
+        # 'nucleon_width' and 'nucleon_structure' parameters. In v2 of the
+        # manuscript, I reverted to a 'sampling_radius' and 'constituent_width'.
+        # These new parameters are more physical and easier to understand.
         self.array[:, 5] = .2 + self.array[:, 5]*(self.array[:, 3] - .2)
         self.array[:, 3] = np.sqrt(self.array[:, 3]**2 - self.array[:, 5]**2)
 

### 2018-10-11 src/model.py

@@ -193,6 +193,7 @@ class ModelData:
             """
             obs_stack = list(keys)
             obs = obs_stack.pop()
+            logging.info('calculating observable: %s', obs)
 
             if obs in ['dNch_deta', 'dET_deta']:
                 return lambda events: events[obs].mean()
@@ -244,7 +245,7 @@ class ModelData:
             trigger = events['trigger']
             minbias = (0, float('inf'))
 
-            #events = correct_yield(events)
+            events = correct_yield(events)
 
             if bin_type == 'cent':
                 """

### 2018-10-12 paper/paper.tex
Save state before weekend

@@
 \newcommand{\T}{\tilde{T}}
-\newcommand{\x}{\mathbf x}
-\newcommand{\y}{\mathbf y}
-\newcommand{\z}{\mathbf z}
+\newcommand{\bv}{\mathbf b}
+\newcommand{\xv}{\mathbf x}
+\newcommand{\yv}{\mathbf y}
+\newcommand{\zv}{\mathbf z}
 \newcommand{\paddedhline}{\noalign{\smallskip}\hline\noalign{\smallskip}}
@@
   \label{eq:tpp}
-  T_{pp}(b) \equiv \int dx\, dy\, dz\, \rho_A(x, y, z)\, \rho_B(x - b, y, z),
+  T_{pp}(b) = \int dx\, dy\, dz\, \rho_A(x, y, z)\, \rho_B(x - b, y, z),
 \end{equation}
@@
 \end{equation}
+
 The proton densities $\rho_{A,B}$ in Eq.~\eqref{eq:tpp} are commonly modeled using a spherically symmetric distribution.
@@
 Needless to say, such approximations are admittedly crude and may have a significant effect on the dynamics of small collision systems where the proton size is comparable to the size of the produced QGP medium.
-
 A number of previous studies have investigated the effects of deformed or ``lumpy'' protons.
@@
   \label{eq:rho}
-  \rho_{A, B}(\x) = \frac{1}{n_c} \sum\limits_{i=1}^{n_c} \rho_c(\x - \x_i),
+  \rho_{A, B}(\xv) = \frac{1}{n_c} \sum\limits_{i=1}^{n_c} \rho_c(\xv - \xv_i),
 \end{equation}
@@
 \begin{equation}
-  \rho_c(\x) = \frac{1}{(2 \pi v^2)^{3/2}} \exp\left(- \frac{|\x|^2}{2 v^2}\right).
+  \rho_c(\xv) = \frac{1}{(2 \pi v^2)^{3/2}} \exp\left(- \frac{|\xv|^2}{2 v^2}\right).
 \end{equation}
 of width $v$.
-The constituent positions $\x_i$ are sampled independently (ignoring correlations) from a Gaussian radial distribution,
+The constituent positions $\xv_i$ are sampled independently (ignoring correlations) from a Gaussian radial distribution,
 \begin{equation}
   \label{eq:radial_dist}
-  P(\x_i; \x_{A,B}) = \frac{1}{(2\pi r^2)^{3/2}} \exp\left(-\frac{|\x_i - \x_{A,B}|^2}{2 r^2}\right),
-\end{equation}
-where $\x_{A,B}$ is the position of each proton, and $r$ is a parameter which varies the sampling radius of the constituent positions about each proton position.
+  P(\xv_i; \xv_{A,B}) = \frac{1}{(2\pi r^2)^{3/2}} \exp\left(-\frac{|\xv_i - \xv_{A,B}|^2}{2 r^2}\right),
+\end{equation}
+where $\xv_{A,B}$ is the position of each proton, and $r$ is a parameter which varies the sampling radius of the constituent positions about each proton position.
 As a matter of convenience, we sample the proton positions \emph{before} determining the constituent positions.
-This leads to small discrepancies between the designated proton positions $\x_{A,B}$, and the actual position of each proton's center-of-mass, owing to fluctuations in the constituent positions $\x_i$.
+This leads to small discrepancies between the designated proton positions $\xv_{A,B}$, and the actual position of each proton's center-of-mass, owing to fluctuations in the constituent positions $\xv_i$.
 The sampling radius $r$ should thus be interpreted with care; it is \emph{not} the root-mean-square radius of the sampled nucleons.
@@
 
-We assign the two protons $A$, $B$ a random impact parameter, and sample their inelastic collision probability using Eq.~\eqref{eq:pcoll}.
-Assuming they collide, we assign each proton a fluctuated thickness function,
-\begin{equation}
-  \label{eq:fluctuated_thick}
-  \T_{A, B}(x, y) \equiv \int dz\, \tilde{\rho}_{A,B}(x \pm b/2, y, z),
-\end{equation}
-which projects each fluctuated proton density $\tilde{\rho}_{A,B}$ onto the transverse plane $(x, y)$ centered at $z=0$.
-
-Note, if any part of the proton collides, the entire proton contributes to inelastic particle production.
-There is no concept of individually wounded constituents; the entire proton either collides (or misses) as a singular connected object.
+The two protons $A$, $B$ are assigned a random impact parameter, and Eq.~\eqref{eq:pcoll} is used to sample their inelastic collision probability $P_\mathrm{coll}(b)$.
+Note, this proton-proton inelastic collision probability has no direct knowledge of the individual constituent positions; it is only indirectly sensitive to their positions through the density $\rho_{A, B}$.
+The protons in this work are described by a \emph{single} density field.
+The constituents are simply used as a mechanism to generate deformity and fluctuations.
 This is an important distinction between the present model and a similar nucleon substructure implementation known as the participant or ``wounded'' quark model which allows for a subset of quarks (constituents) to participate inside a single nucleon \cite{ANISOVICH1978477, Broniowski:2016pvx}.
@@
 
-The initial energy density $e$ deposited at midrapidity and proper time $\tau_0 \approx 0~\fmc$ is some function
-\begin{equation}
-  e(\x_\perp, \eta_s=0, \tau_0) = f(\T_A, \T_B)
-\end{equation}
-of the local density of participant matter $\T_A$, $\T_B$ in each nucleus.
-A natural first guess for this mapping is the arithmetic mean
-\begin{equation}
-  f = \text{const} \times \frac{\T_A + \T_B}{2},
-\end{equation}
-which is simply a wounded nucleon model up to meaningless factor of two in the normalization.
-The wounded nucleon model was in fact one of the first such mappings used as a proxy for initial particle production and entropy deposition in relativistic heavy-ion collisions \cite{Bialas:1976ed}.
-It was quickly realized, however, that the wounded nucleon model predicts the wrong scaling for charged particle production as a function of collision centrality and hence the wrong scaling for initial entropy deposition as a function of participant thicknesses $\T_A$ and $\T_B$ \cite{Kharzeev:2000ph}.
+Hereafter, we switch to Milne coordinates $(\xv_\perp, \eta_s, \tau)$, where $\xv_\perp$ is a Cartesian coordinate $(x, y)$ in the transverse plane, $\eta_s = \tfrac{1}{2} \log[(t+z)/(t-z)]$ is the system's spacetime rapidity, and $\tau = \sqrt{t^2 - z^2}$ its proper time.
+Assuming our protons collide at the sampled impact parameter $\bv$, we assign each proton a participant thickness,
+\begin{equation}
+  \label{eq:fluctuated_thick}
+  \T_{A, B}(\xv_\perp) \equiv \int dz\, \tilde{\rho}_{A,B}(\xv_\perp \pm \bv/2, z),
+\end{equation}
+which projects its \emph{fluctuated} proton density,
+\begin{equation}
+  \tilde{\rho}_{A,B}(\xv) \equiv \frac{1}{n_c} \sum\limits_{i=1}^{n_c} \gamma_i\, \rho_c(\xv - \xv_i),
+\end{equation}
+onto the transverse plane $\xv_\perp$ centered at $z=0$.
+This fluctuated proton density $\tilde{\rho}_{A,B}$ equals the proton density $\rho_{A,B}(\xv)$ in Eq.~\eqref{eq:rho} with additional weights $\gamma_i$ sampled from a gamma distribution with unit mean and variance $1/k$.
+These \emph{ad hoc} gamma random weights are necessary to describe the large fluctuations observed in minimum bias proton-proton collisions, although their exact physical origin is not well understood.
+
+The initial energy density $e$ deposited at midrapidity $\eta_s=0$ and proper time $\tau_0 \approx 0~\fmc$ is then some function,
+\begin{equation}
+  e(\xv_\perp, \eta_s=0, \tau_0) = f(\T_A, \T_B),
+\end{equation}
+of the local participant thicknesses $\T_A$, $\T_B$ in each nucleus.
+A natural first guess for this mapping is an arithmetic mean,
+\begin{equation}
+  e(\xv_\perp, \eta_s=0, \tau_0) = \text{const} \times \frac{\T_A + \T_B}{2},
+\end{equation}
+equal to a participant or ``wounded nucleon'' model for initial energy deposition up to meaningless factor of two in the overall normalization constant.
+This wounded nucleon scaling was in fact one of the first such mappings used as a proxy for initial particle production in relativistic heavy-ion collisions \cite{Bialas:1976ed}.
+It was quickly realized, however, that the wounded nucleon model predicts the wrong scaling for charged particle production as a function of collision centrality and hence the wrong scaling for initial energy deposition at midrapidity as a function of participant thicknesses $\T_A$ and $\T_B$ \cite{Kharzeev:2000ph}.
 
@@
   \label{eq:gmean}
-  \frac{dS}{d^2x_\perp d\eta}\,\bigg\vert_{\eta=0} \propto \bigg( \frac{\T_A^p + \T_B^p}{2} \bigg)^{1/p},
-\end{equation}
-based on a family of functions known as the generalized mean(s).
-This parametrization introduces a dimensionless parameter $p$ which varies the scaling behavior of initial entropy deposition at midrapidity.
+  e(\xv_\perp, \eta_s=0, \tau_0) = \text{const} \times \bigg( \frac{\T_A^p + \T_B^p}{2} \bigg)^{1/p},
+\end{equation}
+based on a family of functions known as the generalized mean(s)\footnote{Previous applications of the \trento\ model used the generalized mean ansatz to parametrize the initial entropy density $s(\xv_\perp, \eta_s=0, \tau_0)$, noting its approximate linear relation with the final charged particle density $d\nch/d\eta$ \cite{Moreland:2014oya, Bernhard:2016tnd}.
+In the present framework, the initial energy density $e(\xv_\perp, \eta_s=0, \tau_0)$ is approximately proportional to $d\nch/d\eta$ as well; hence, the generalized mean works equally well as a parametrization for either quantity.
+Here we use the energy density to make direct contact with the free-streaming equation Eq.~\eqref{energy-momentum}.}.
+This parametrization introduces a dimensionless parameter $p$ which varies the scaling behavior of initial energy deposition at midrapidity.
 For certain discrete values of $p$, it reduces to well known functional forms such as the arithmetic, geometric, and harmonic means:
@@
   \setlength{\extraspace}{0.5ex}
-  \frac{dS}{d\eta}\bigg\vert_{\eta=0} \propto
+  e \propto
   \begin{cases}
-    \max(T_A, T_B) & p \rightarrow +\infty, \\[\extraspace]
-    (T_A + T_B)/2 & p = +1, \hfill \text{ (arithmetic)} \\[\extraspace]
-    \sqrt{T_A T_B} & p = 0, \hfill \text{ (geometric)} \\[\extraspace]
-    2\, T_A T_B/(T_A + T_B) & p = -1, \hfill \text{ (harmonic)} \\[\extraspace]
-    \min(T_A, T_B) & p \rightarrow -\infty.
+    \max(\T_A, \T_B) & p \rightarrow +\infty, \\[\extraspace]
+    (\T_A + \T_B)/2 & p = +1, \hfill \text{ (arithmetic)} \\[\extraspace]
+    \sqrt{\T_A \T_B} & p = 0, \hfill \text{ (geometric)} \\[\extraspace]
+    2\, \T_A \T_B/(\T_A + \T_B) & p = -1, \hfill \text{ (harmonic)} \\[\extraspace]
+    \min(\T_A, \T_B) & p \rightarrow -\infty.
   \end{cases}
@@
 \end{equation}
+
 Conveniently, it has been shown that the generalized mean ansatz is able to mimic the scaling behavior of certain saturation based initial condition models \cite{Bernhard:2016tnd}, and hence it should serve as a reasonable parametric form for exploring QGP entropy deposition, assuming imperfect knowledge of saturation effects in nature.
@@
 
-When applying the model to heavy-ions, we sample nucleon positions from a Woods-Saxon density distribution subject to a minimum distance criteria $|\x_i - \x_j| > d_\mathrm{min}$ between all pairs of nucleons $i$, $j$.
+When applying the model to heavy-ions, we sample nucleon positions from a Woods-Saxon density distribution subject to a minimum distance criteria $|\xv_i - \xv_j| > d_\mathrm{min}$ between all pairs of nucleons $i$, $j$.
 The minimum distance algorithm, first described in Ref.~\cite{Bernhard:2018hnz}, uses a simple trick to resample the nucleon positions without modifying the target Woods-Saxon radial distribution.
@@
 
-The nuclear collision model constructed in Sec.~\ref{sec:model} includes a number of free parameters $\x$ which describe the initial state, pre-equilibrium dynamics, and hydrodynamic medium.
-Given values for the parameters $\x$, the model may be used to predict a vector of simulated observables $\y_m$.
-For example, $\y_m$ might be a vector consisting of charged particle yields in different centrality bins.
-The physics model thus represents a vector-valued function $f(\x) = \y_m$ which maps the parameter values $\x$ to the simulated observables $\y_m$.
-
-The goal of this work is to estimate the true model parameters $\x_\star$ provided some evidence that our model predictions $\y_m$ describe experimental measurements $\y_e$.
+The nuclear collision model constructed in Sec.~\ref{sec:model} includes a number of free parameters $\xv$ which describe the initial state, pre-equilibrium dynamics, and hydrodynamic medium.
+Given values for the parameters $\xv$, the model may be used to predict a vector of simulated observables $\yv_m$.
+For example, $\yv_m$ might be a vector consisting of charged particle yields in different centrality bins.
+The physics model thus represents a vector-valued function $f(\xv) = \yv_m$ which maps the parameter values $\xv$ to the simulated observables $\yv_m$.
+
+The goal of this work is to estimate the true model parameters $\xv_\star$ provided some evidence that our model predictions $\yv_m$ describe experimental measurements $\yv_e$.
 The problem involves three distinct components:
@@
   \item $H_f$: the hypothesis that the nuclear collision model $f$ formulated in this work provides a realistic description of reality,
-  \item $H_\x$: the hypothesis that the model parameters $\x$ are the true model parameters $\x_\star$ of $f$, and
-  \item $E$: the evidence provided by the experimental data $\y_e$ and its corresponding uncertainties.
+  \item $H_\xv$: the hypothesis that the model parameters $\xv$ are the true model parameters $\xv_\star$ of $f$, and
+  \item $E$: the evidence provided by the experimental data $\yv_e$ and its corresponding uncertainties.
 \end{enumerate}
@@
 
-Subject to this assumption, we can apply Bayes' theorem to evaluate the hypothesis $H_\x$ for the true model parameters,
+Subject to this assumption, we can apply Bayes' theorem to evaluate the hypothesis $H_\xv$ for the true model parameters,
 \begin{equation}
   \label{eq:bayes}
-  P(H_\x | E) \propto P(E | H_\x)\, P(H_\x).
-\end{equation}
-The left-side of this expression is the \emph{posterior}: the probability for hypothesis $H_\x$ given the experimental evidence $E$.
+  P(H_\xv | E) \propto P(E | H_\xv)\, P(H_\xv).
+\end{equation}
+The left-side of this expression is the \emph{posterior}: the probability for hypothesis $H_\xv$ given the experimental evidence $E$.
 On the right-side there are two separate terms.
-The first term $P(E | H_\x)$ is the \emph{likelihood} function: the probability of observing the experimental evidence $E$ given our model and the hypothesis $H_\x$ for the true model parameters $\x_\star$, and the second term $P(H_\x)$ is the \emph{prior}: an estimate of the probability of hypothesis $H_\x$ in the absence of evidence $E$.
+The first term $P(E | H_\xv)$ is the \emph{likelihood} function: the probability of observing the experimental evidence $E$ given our model and the hypothesis $H_\xv$ for the true model parameters $\xv_\star$, and the second term $P(H_\xv)$ is the \emph{prior}: an estimate of the probability of hypothesis $H_\xv$ in the absence of evidence $E$.
 
@@
   \label{eq:likelihood}
-  P(E | H_\x) = \frac{1}{\sqrt{(2\pi)^m \det \Sigma}} \exp \left ( -\frac{1}{2}\Delta\y^\intercal \Sigma^{-1} \Delta\y \right ),
-\end{equation}
-where $\Delta\y = \y_m(\x) - \y_e$ is the discrepancy of the model and experiment, and $\Sigma = \Sigma_m(\x) + \Sigma_e$ is the \emph{total} covariance matrix, equal to the sum of a modeling component $\Sigma_m(\x)$ and an experimental component $\Sigma_e$ which account for all known sources of uncertainty in the simulated and measured observables.
+  P(E | H_\xv) = \frac{1}{\sqrt{(2\pi)^m \det \Sigma}} \exp \left ( -\frac{1}{2}\Delta\yv^\intercal \Sigma^{-1} \Delta\yv \right ),
+\end{equation}
+where $\Delta\yv = \yv_m(\xv) - \yv_e$ is the discrepancy of the model and experiment, and $\Sigma = \Sigma_m(\xv) + \Sigma_e$ is the \emph{total} covariance matrix, equal to the sum of a modeling component $\Sigma_m(\xv)$ and an experimental component $\Sigma_e$ which account for all known sources of uncertainty in the simulated and measured observables.
 
@@
 
-For the prior $P(H_\x)$, we specify ranges, i.e.\ minimum and maximum values, for each parameter which are listed in Table~\ref{tab:design}.
+For the prior $P(H_\xv)$, we specify ranges, i.e.\ minimum and maximum values, for each parameter which are listed in Table~\ref{tab:design}.
 We assume the prior distribution is constant and nonzero within each specified range and zero otherwise.
@@
 
-The likelihood function \eqref{eq:likelihood} provides evidence for \mbox{(or against)} the model parameters $\x$ by comparing the model predictions $\y_m(\x)$ to experimental data $\y_e$.
+The likelihood function \eqref{eq:likelihood} provides evidence for \mbox{(or against)} the model parameters $\xv$ by comparing the model predictions $\yv_m(\xv)$ to experimental data $\yv_e$.
 We focus on simple experimental observables in the present study which are sensitive to the bulk properties of the produced medium.
@@
 \end{equation}
-between each event's initial secondary-parton density ${N_\text{parton} = \int d^2x\,n(\x)}$ in Eq.~\eqref{energy-momentum} and its final-state event activity, characterized by $\nch$ or $\ntrk$ \cite{Abelev:2013bla, Chatrchyan:2013nka}.
+between each event's initial secondary-parton density ${N_\text{parton} = \int d^2x\,n(\xv)}$ in Eq.~\eqref{energy-momentum} and its final-state event activity, characterized by $\nch$ or $\ntrk$ \cite{Abelev:2013bla, Chatrchyan:2013nka}.
 
@@
 \end{equation}
-where $\sigma^\text{stat}_i$ is the statistical uncertainty of observable $y_i$ in the experimental observable vector $\y_e = (y_1, \dots, y_m)$.
+where $\sigma^\text{stat}_i$ is the statistical uncertainty of observable $y_i$ in the experimental observable vector $\yv_e = (y_1, \dots, y_m)$.
 The systematic errors, meanwhile, are typically correlated, but the correlation structure is not reported by the experiments so we assert a reasonable form.
@@
 for the Pb-Pb experimental data constructed using Eq.~\eqref{eq:corr} on the right-side of Fig.~\ref{fig:correlation}.
-Here $y_i$ denotes an element of the experimental data $\y_e$ and $\sigma_i$ its corresponding uncertainty.
+Here $y_i$ denotes an element of the experimental data $\yv_e$ and $\sigma_i$ its corresponding uncertainty.
 The correlation matrix is block diagonal, with each block representing the correlations within a single class of observable.
@@
 
-In principle, one could calculate the likelihood function in Eq.~\eqref{eq:likelihood} directly, e.g.\ by running the model a large number of times at a given parameter point $\x$ to calculate the model observables $\y_m(\x)$ from the ensemble of simulated events, but in practice such a procedure would be intractable.
+In principle, one could calculate the likelihood function in Eq.~\eqref{eq:likelihood} directly, e.g.\ by running the model a large number of times at a given parameter point $\xv$ to calculate the model observables $\yv_m(\xv)$ from the ensemble of simulated events, but in practice such a procedure would be intractable.
 The model is computationally intensive to evaluate, and thousands of events are required to calculate the simplest observables at a single parameter point.
@@
 We therefore follow an established framework for computationally intensive models and train an emulator to act as a fast surrogate for the full physics simulation \cite{OHagan:2006ba, Higdon:2008cmc, Higdon:2014tva}.
-The emulator enables essentially instantaneous predictions for $\y_m = f(\x)$ and allows us to sample the posterior distribution millions of times.
+The emulator enables essentially instantaneous predictions for $\yv_m = f(\xv)$ and allows us to sample the posterior distribution millions of times.
 In order to train the emulator, we first generate a scaffolding of the parameter space using a maximin Latin hypercube design \cite{Morris:1995lh} to distribute 500 points uniformly throughout our 15-dimensional parameter space according to the parameter ranges in Table~\ref{tab:design}.
@@
 
-Specifically, let $X$ denote the ${d \times n}$ design matrix of $d=500$ training points, where each training point is a vector $\x = (x_1, x_2, \dots, x_n)$ of the $n=15$ model parameters in Table~\ref{tab:design}.
-Similarly, let $Y$ denote the corresponding ${d \times m}$ observables matrix, where each row of $Y$ is a vector $\y_m = (y_1, y_2, \dots, y_m)$ of $m$ different simulated observables (here we overload our notation so the subscript $m$ means both \emph{model} and \emph{observable number}).
+Specifically, let $X$ denote the ${d \times n}$ design matrix of $d=500$ training points, where each training point is a vector $\xv = (x_1, x_2, \dots, x_n)$ of the $n=15$ model parameters in Table~\ref{tab:design}.
+Similarly, let $Y$ denote the corresponding ${d \times m}$ observables matrix, where each row of $Y$ is a vector $\yv_m = (y_1, y_2, \dots, y_m)$ of $m$ different simulated observables (here we overload our notation so the subscript $m$ means both \emph{model} and \emph{observable number}).
 Our goal is to train an emulator for the physics model $f$ using the discrete observations $f: X \mapsto Y$.
@@
 One quirk of GPE's is that they are restricted to scalar-valued functions, i.e.\ functions of one output, whereas we require an emulator for vector-valued functions with multiple outputs.
-This restriction is commonly circumvented using principal component analysis (PCA): a general procedure which transforms a set of correlated variables $\y = (y_1, y_2, \dots, y_m)$ into a new basis representation $\z = (z_1, z_2, \dots, z_m)$ where the linear correlations between $z_i$, $z_j$ vanish for all $i\ne j \in m$ \cite{Tipping:1999}.
-Independent GPE's can then be used to emulate each $z \in \z$ since the variables $(z_1, \dots, z_m)$ are linearly uncorrelated.
-The emulated vector $\z$ is then easily reexpressed in the basis of $\y$ through its inverse transformation.
+This restriction is commonly circumvented using principal component analysis (PCA): a general procedure which transforms a set of correlated variables $\yv = (y_1, y_2, \dots, y_m)$ into a new basis representation $\zv = (z_1, z_2, \dots, z_m)$ where the linear correlations between $z_i$, $z_j$ vanish for all $i\ne j \in m$ \cite{Tipping:1999}.
+Independent GPE's can then be used to emulate each $z \in \zv$ since the variables $(z_1, \dots, z_m)$ are linearly uncorrelated.
+The emulated vector $\zv$ is then easily reexpressed in the basis of $\yv$ through its inverse transformation.
 
 We first preprocess our model observables by centering and scaling each column of $Y$ (single observable) to zero mean and unit variance to generate a standardized observable matrix $\tilde{Y}$.
-PCA is then used to reexpress each row-vector $\tilde{\y}$ of $\tilde{Y}$ (all standardized observables at a single design point) in the new principal component basis:
-\begin{equation}
-  \tilde{\y}_i = \sum\limits_{j=1}^m z_{ij} \mathbf{v_j},
-\end{equation}
-where $\tilde{\y}_i$ are the standardized observables of the $i$-th row-vector (design point) of matrix $\tilde{Y}$, and $z_{ij}$ and $\mathbf{v}_j$ are the coefficients and vectors of its $j$-th principal component.
+PCA is then used to reexpress each row-vector $\tilde{\yv}$ of $\tilde{Y}$ (all standardized observables at a single design point) in the new principal component basis:
+\begin{equation}
+  \tilde{\yv}_i = \sum\limits_{j=1}^m z_{ij} \mathbf{v_j},
+\end{equation}
+where $\tilde{\yv}_i$ are the standardized observables of the $i$-th row-vector (design point) of matrix $\tilde{Y}$, and $z_{ij}$ and $\mathbf{v}_j$ are the coefficients and vectors of its $j$-th principal component.
 
 The principal components are reported in order of explained variance, with the first principal component vector $\mathbf{v}_1$ accounting for the most variance in $\tilde{Y}$, and the last principal component vector $\mathbf{v}_m$ accounting for the least.
-We then train a set of independent GPE's $\{z_i=\mathrm{gp}_i(\x)\}$ to predict the first $k$ principal components $(z_1, \dots, z_k)$ as a function of the model parameters $\x$ which vary across the design $X$.
+We then train a set of independent GPE's $\{z_i=\mathrm{gp}_i(\xv)\}$ to predict the first $k$ principal components $(z_1, \dots, z_k)$ as a function of the model parameters $\xv$ which vary across the design $X$.
 For the present study, we use $k=7$ principal components when emulating the $p$-Pb system and $k=8$ principal components when emulating Pb-Pb, chosen to describe 99.5\% of the total observed variance of each system.
@@
 The GPE's are essentially fancy interpolators applied to the model's training points and PCA transformed observables.
-Each GPE reports a mean value $z(\x)$ as well as an estimated error $\delta z(\x)$ which accounts for statistical noise in the training data and interpolation error between the design points.
-Once the GPE's are trained, we can predict the observables $\y_m$ at parameter point $\x$ by transforming the vector of principal components
-\begin{equation}
-  \mathbf{z}(\x) = (z_1(\x), z_2(\x), \dots, z_k(\x))
+Each GPE reports a mean value $z(\xv)$ as well as an estimated error $\delta z(\xv)$ which accounts for statistical noise in the training data and interpolation error between the design points.
+Once the GPE's are trained, we can predict the observables $\yv_m$ at parameter point $\xv$ by transforming the vector of principal components
+\begin{equation}
+  \mathbf{z}(\xv) = (z_1(\xv), z_2(\xv), \dots, z_k(\xv))
 \end{equation}
@@
 \end{equation}
-and transform it back to physical space as well to obtain the covariance matrix of the model observables $\y_m$ at a given parameter point $\x$.
-
-The resulting emulator therefore predicts both a mean prediction $\y_m(\x)$ and an uncertainty covariance matrix $\Sigma_m(\x)$ which accounts for multiple sources of model and emulator uncertainty, including the truncation error expected from using a finite number of principal components $k < m$.
+and transform it back to physical space as well to obtain the covariance matrix of the model observables $\yv_m$ at a given parameter point $\xv$.
+
+The resulting emulator therefore predicts both a mean prediction $\yv_m(\xv)$ and an uncertainty covariance matrix $\Sigma_m(\xv)$ which accounts for multiple sources of model and emulator uncertainty, including the truncation error expected from using a finite number of principal components $k < m$.
 The model covariance matrix $\Sigma_m$ includes so-called \emph{known}-unknowns such as statistical error and emulator interpolation error, but not \emph{unknown}-unknowns such as the overall validity of small-system hydrodynamics, i.e.\ things which lack a unified consensus or are difficult to quantify.
-We show in Fig.~\ref{fig:correlation} the resulting Pb-Pb correlation matrix $\mathrm{corr}(y_i,y_j)$ for the model (emulator) at a random parameter point $\x$ in the design space (left-side), along side the same correlation matrix for the experimental data (right-side) discussed previously.
+We show in Fig.~\ref{fig:correlation} the resulting Pb-Pb correlation matrix $\mathrm{corr}(y_i,y_j)$ for the model (emulator) at a random parameter point $\xv$ in the design space (left-side), along side the same correlation matrix for the experimental data (right-side) discussed previously.
 For additional information on the model emulator, we direct the reader to Appendix~\ref{app:validation} which includes several validation tests of the emulator prediction accuracy.
@@
   \label{eq:joint_likelihood}
-  P(E | H_\x) = P(E_\text{Pb-Pb} | H_\x) \cdot P(E_\text{$p$-Pb} | H_\x),
-\end{equation}
-where $E$ subsumes all evidence from the $p$-Pb and Pb-Pb collision systems and $H_\x$ is our hypothesis that $\x$ equals the true parameters $\x_\star$.
-We then perform Markov-chain Monte Carlo (MCMC) importance sampling on the posterior distribution in Eq.~\eqref{eq:bayes} to draw random samples for $P(H_\x | E)$, the estimate of the true model parameters given the model and the experimental data \cite{Goodman:2010en, FM:2013mc}.
+  P(E | H_\xv) = P(E_\text{Pb-Pb} | H_\xv) \cdot P(E_\text{$p$-Pb} | H_\xv),
+\end{equation}
+where $E$ subsumes all evidence from the $p$-Pb and Pb-Pb collision systems and $H_\xv$ is our hypothesis that $\xv$ equals the true parameters $\xv_\star$.
+We then perform Markov-chain Monte Carlo (MCMC) importance sampling on the posterior distribution in Eq.~\eqref{eq:bayes} to draw random samples for $P(H_\xv | E)$, the estimate of the true model parameters given the model and the experimental data \cite{Goodman:2010en, FM:2013mc}.
 For this we use an affine-invariant ensemble sampler which uses a large ensemble of interdependent walkers \cite{Goodman:2010en, FM:2013mc} and allow the MCMC chain to ``burn-in'' before generating \order{7} posterior samples.
@@
 We now direct our attention to Fig.~\ref{fig:posterior} which shows the main result of this work, the posterior distribution of the model input parameters.
-Recall that the posterior $P(H_\x | E)$ is the probability that our hypothesis $H_\x$ for the true model parameters $\x_\star$ is correct, given the evidence $E$ provided by experiment.
+Recall that the posterior $P(H_\xv | E)$ is the probability that our hypothesis $H_\xv$ for the true model parameters $\xv_\star$ is correct, given the evidence $E$ provided by experiment.
 The present posterior has 15 dimensions, one for each parameter listed in Table~\ref{tab:design}, and thus its joint distribution cannot be summarized by one figure alone.
@@
 \begin{equation}
-  \x_\mathrm{MAP} \equiv \operatorname*{arg\, max}_{\x} P(H_\x | E).
-\end{equation}
-We then ran \order{6} minimum-bias and multiplicity triggered events using the MAP estimate $\x_\mathrm{MAP}$ and computed all of the model observables listed in Sec.~\ref{sec:observables}.
+  \xv_\mathrm{MAP} \equiv \operatorname*{arg\, max}_{\xv} P(H_\xv | E).
+\end{equation}
+We then ran \order{6} minimum-bias and multiplicity triggered events using the MAP estimate $\xv_\mathrm{MAP}$ and computed all of the model observables listed in Sec.~\ref{sec:observables}.
 The resulting model calculations are shown in Fig.~\ref{fig:obs_map} alongside experimental data from CMS \cite{Chatrchyan:2013nka} and ALICE \cite{Adam:2015ptt, Adam:2016izf, Adam:2014qja, Abelev:2013bla}.
@@
 
-We find that we can quickly estimate the maximum radius $R_\text{max} = |\x_\text{max}|^2$ of the spacetime hypersurface ${T = T(e_\text{min})}$ by running the event on a coarse-grained spatial grid with one-third the spatial resolution we would otherwise require to resolve typical hydrodynamic observables such as mean $p_T$ and flows.
+We find that we can quickly estimate the maximum radius $R_\text{max} = |\xv_\text{max}|^2$ of the spacetime hypersurface ${T = T(e_\text{min})}$ by running the event on a coarse-grained spatial grid with one-third the spatial resolution we would otherwise require to resolve typical hydrodynamic observables such as mean $p_T$ and flows.
 The simulation time of a single VISH2+1 event scales like ${\sim}n_x^3$ since $dx \propto d\tau$, and thus our ``pre-run'' event requires only ${\sim}1/27$th the time of a production event.
@@
 
-The emulator is a surrogate for the full physics simulation which generates probabilistic predictions for the model observables $\y_m$ at a given point $\x$.
+The emulator is a surrogate for the full physics simulation which generates probabilistic predictions for the model observables $\yv_m$ at a given point $\xv$.
 Here we validate these probabilistic predictions using a method known as k-fold cross validation.
@@
 
-More generally, we can perform the validation test in Fig.~\ref{fig:validation_example} for \emph{every} observable $y \in \y_m$ and check that each observable's normalized residuals ${z=(y_\text{pred} - y_\text{obs})/\sigma_\text{pred}}$ follow a unit normal distribution.
+More generally, we can perform the validation test in Fig.~\ref{fig:validation_example} for \emph{every} observable $y \in \yv_m$ and check that each observable's normalized residuals ${z=(y_\text{pred} - y_\text{obs})/\sigma_\text{pred}}$ follow a unit normal distribution.
 This test is applied to the $p$-Pb and Pb-Pb collision systems in Fig.~\ref{fig:validation_all}.

### 2018-10-12 src/emulator.py

@@ -26,6 +26,7 @@ from sklearn.preprocessing import StandardScaler
 from . import cachedir, lazydict, model
 from .design import Design
 
+import matplotlib.pyplot as plt
 
 class _Covariance:
     """

### 2018-10-12 src/plots.py

@@ -2358,6 +2358,16 @@ def grid_error():
 
     set_tight(fig)
 
+@plot
+def energy_vs_yield():
+    fig = plt.figure(figsize=figsize(.5, aspect=1))
+
+    path = Path(workdir, 'model_output', 'map', 'PbPb5020.dat')
+    model_data = model.ModelData('PbPb5020', path)
+
+    for pt, ev in model_data.design_events:
+        x, y = [ev[k] for k in ('init_entropy', 'dNch_deta')]
+        plt.scatter(x[::1000], y[::1000])
 
 @plot
 def entropy_scaling():

### 2018-10-15 paper/paper.tex
More clearly explain trento parametrization with free streaming.

@@
 We model the initial state of $p$-Pb and Pb-Pb collisions at $\sqrts=5.02$~TeV using the boost-invariant \trento\ model \cite{Moreland:2014oya}.
-Generally speaking, the initial three-dimensional distribution of matter produced in relativistic nuclear collisions is \emph{not} boost-invariant; longitudinal entropy deposition fluctuates both locally point-to-point in the transverse plane as well as globally event-by-event due to asymmetries in the sampled density of participant matter \cite{Ke:2016jrd, Bozek:2010vz}.
+Generally speaking, the initial three-dimensional distribution of matter produced in relativistic nuclear collisions is \emph{not} boost-invariant.
+The longitudinal density of particle production fluctuates both locally point-to-point in the transverse plane as well as globally event-by-event due to asymmetries in the sampled density of participant matter \cite{Ke:2016jrd, Bozek:2010vz}.
 Nevertheless, boost-invariance has been shown to be a good approximation for both large and small collision systems when hydrodynamic observables are calculated from particles that are detected close to midrapidity \cite{Shen:2016zpp}.
@@
 The resulting nuclear overlap time $\Delta t \lesssim 0.1\ \fmc$ at top RHIC and LHC energies, and thus it is safe to neglect the initial transverse dynamics which occur while the nuclei pass through each other.
-We therefore assume that the collision produces all secondary particles at uniform proper time $\tau_0 \ll 1~\fmc$, and that it deposits energy at midrapidity which is some function of the locally varying beam integrated density of each nucleus.
+We therefore assume that the collision produces all secondary particles at a uniform proper time $\tau_0 \ll 1\fmc$, and that it deposits energy at midrapidity which is a function of the locally varying beam-integrated density of each nucleus.
 
@@
   \label{eq:radial_dist}
-  P(\xv_i; \xv_{A,B}) = \frac{1}{(2\pi r^2)^{3/2}} \exp\left(-\frac{|\xv_i - \xv_{A,B}|^2}{2 r^2}\right),
+  \rho_\text{pos}(\xv_i; \xv_{A,B}) = \frac{1}{(2\pi r^2)^{3/2}} \exp\left(-\frac{|\xv_i - \xv_{A,B}|^2}{2 r^2}\right),
 \end{equation}
@@
 This leads to small discrepancies between the designated proton positions $\xv_{A,B}$, and the actual position of each proton's center-of-mass, owing to fluctuations in the constituent positions $\xv_i$.
-The sampling radius $r$ should thus be interpreted with care; it is \emph{not} the root-mean-square radius of the sampled nucleons.
-
-\begin{figure}
-  \begin{tikzpicture}
-    \node {\includegraphics{thickness}};
-    \node[label=above:{10 fm}] (a) at (0, 1.1) {};
-    \node (b) at (0, -1.1) {};
-    \draw [<->, semithick] (a) to (b);
-  \end{tikzpicture}
-  \caption{\label{fig:thickness} Effect of nucleon substructure on the nuclear thickness function $T(x, y) \equiv \int dz\, \rho(x, y, z)$ of a $^{208}\mathrm{Pb}$ nucleus.
-The nucleus on the left has Gaussian nucleons of width $0.8$~fm, while the nucleus on the right has composite nucleons, each containing six constituents of width $0.4$~fm.}
-\end{figure}
-
-The two protons $A$, $B$ are assigned a random impact parameter, and Eq.~\eqref{eq:pcoll} is used to sample their inelastic collision probability $P_\mathrm{coll}(b)$.
+The sampling radius $r$ should thus be interpreted with care.
+It is \emph{not} the root-mean-square radius of the sampled nucleons.
+
+The two protons $A$, $B$ are assigned a random impact parameter $\bv$, and Eq.~\eqref{eq:pcoll} is used to sample their inelastic collision probability $P_\mathrm{coll}(b)$.
 Note, this proton-proton inelastic collision probability has no direct knowledge of the individual constituent positions; it is only indirectly sensitive to their positions through the density $\rho_{A, B}$.
-The protons in this work are described by a \emph{single} density field.
-The constituents are simply used as a mechanism to generate deformity and fluctuations.
+The protons in this work are described by a \emph{single} density field; constituents are simply used as a mechanism to generate deformity and fluctuations.
 This is an important distinction between the present model and a similar nucleon substructure implementation known as the participant or ``wounded'' quark model which allows for a subset of quarks (constituents) to participate inside a single nucleon \cite{ANISOVICH1978477, Broniowski:2016pvx}.
 The proton, unlike the nucleus, cannot produce semi-stable spectator fragments in a high-energy collision.
-Any spectator quarks produced by a wounded quark model would be colored objects that necessarily contribute to secondary particle production as they fragment and recombine to form color-neutral hadrons.
+Any spectator quarks produced by a wounded quark model would be colored objects that necessarily contribute to secondary particle production as they fragment and recombine to form colorless hadrons.
 We correspondingly require that the nucleons in Eq.~\eqref{eq:rho} participate as singular objects, such that all spectator matter discarded by the simulation is appropriately color-neutral and inert.
@@
 
-The initial energy density $e$ deposited at midrapidity $\eta_s=0$ and proper time $\tau_0 \approx 0~\fmc$ is then some function,
+The energy density $e$ deposited at midrapidity $\eta_s=0$ and proper time $\tau_0 \ll 1~\fmc$ is then some function,
 \begin{equation}
@@
 \end{equation}
-of the local participant thicknesses $\T_A$, $\T_B$ in each nucleus.
+of the local participant thicknesses $\T_A$ and $\T_B$.
 A natural first guess for this mapping is an arithmetic mean,
@@
 This wounded nucleon scaling was in fact one of the first such mappings used as a proxy for initial particle production in relativistic heavy-ion collisions \cite{Bialas:1976ed}.
-It was quickly realized, however, that the wounded nucleon model predicts the wrong scaling for charged particle production as a function of collision centrality and hence the wrong scaling for initial energy deposition at midrapidity as a function of participant thicknesses $\T_A$ and $\T_B$ \cite{Kharzeev:2000ph}.
+It was quickly realized, however, that the wounded nucleon model predicts the wrong scaling for charged particle production as a function of collision centrality and hence the wrong scaling for initial energy deposition at midrapidity as a function of participant thicknesses $\T_A$ and $\T_B$ \cite{Kharzeev:2000ph} \footnote{Typically an entropy density $s$ is used, but the shortcomings of the wounded nucleon model are nonetheless the same when depositing energy.}.
 
@@
 \end{equation}
-based on a family of functions known as the generalized mean(s)\footnote{Previous applications of the \trento\ model used the generalized mean ansatz to parametrize the initial entropy density $s(\xv_\perp, \eta_s=0, \tau_0)$, noting its approximate linear relation with the final charged particle density $d\nch/d\eta$ \cite{Moreland:2014oya, Bernhard:2016tnd}.
-In the present framework, the initial energy density $e(\xv_\perp, \eta_s=0, \tau_0)$ is approximately proportional to $d\nch/d\eta$ as well; hence, the generalized mean works equally well as a parametrization for either quantity.
-Here we use the energy density to make direct contact with the free-streaming equation Eq.~\eqref{energy-momentum}.}.
+based on a family of functions known as the generalized mean(s).
 This parametrization introduces a dimensionless parameter $p$ which varies the scaling behavior of initial energy deposition at midrapidity.
@@
 
-Conveniently, it has been shown that the generalized mean ansatz is able to mimic the scaling behavior of certain saturation based initial condition models \cite{Bernhard:2016tnd}, and hence it should serve as a reasonable parametric form for exploring QGP entropy deposition, assuming imperfect knowledge of saturation effects in nature.
-Of course the model is \emph{not} a substitute for first principle theory calculations, and it may fail to reproduce nuanced features of \emph{ab initio} models such as the existence of short-range gluon field fluctuations \cite{Schenke:2012wb}.
-
-Equation~\eqref{eq:gmean} is a purely local function of nuclear density in the transverse plane and should (in principle) be equally valid for any pair of colliding nuclei at sufficiently high beam energy.
-The model readily generalizes from individual proton-proton collisions to arbitrary nucleus-nucleus collisions by summing the participant thicknesses $\T_{A,B}$ over all nucleons which participate in one or more inelastic collisions.
+We note that previous applications of the \trento\ model \cite{Moreland:2014oya, Bernhard:2016tnd} used the generalized mean ansatz to parametrize the \emph{entropy} density $s(\xv_\perp, \eta_s=0, \tau_\text{hydro})$ at midrapidity, motivated (in part) by the approximate linear relation between initial entropy density $(dS/d\eta_s)\vert_{\eta_s=0}$ and final charged-particle density $(d\nch/d\eta_s)\vert_{\eta_s=0}$ \cite{Shen:2015qta}.
+These studies, however, assumed instant thermalization, an approximation which sets $u^\mu$, $\Pi$, and $\pi^{\mu\nu}$ identically to zero at the hydrodynamic starting time $\tau_\text{hydro}$.
+The present study, meanwhile, uses a pre-equilibrium free streaming stage to estimate $u^\mu$, $\Pi$, and $\pi^{\mu\nu}$ as a function of a variable free-streaming time $\taufs$.
+Strictly speaking, the entropy density ${s(\xv_\perp, \eta_s=0, \tau)}$ is ill defined prior to thermalization.
+Hence, it would not make sense to directly parametrize the entropy density at times $\tau_0 < \tau_\text{hydro}$ as required by the present study.
+We therefore reinterpret the quantity produced by the \trento\ model as an energy density $e(\xv_\perp, \eta_s=0, \tau_0)$ and free-stream this quantity to the hydrodynamic starting time $\tau_\text{hydro}$.
+This assumes, of course, that the model's parametrization for entropy density is also suitable as a parametrization for the energy density.
+To this point, we note that the arguments for the generalized mean ansatz are quite general, and that the energy density $dE/d\eta_s$ and entropy density $dS/d\eta_s$ share many similar qualities.
+For example, both are approximately conserved at midrapidity, and both scale approximately linearly with the final charged-particle density $d\nch/d\eta_s$.
+
+\begin{figure}
+  \begin{tikzpicture}
+    \node {\includegraphics{thickness}};
+    \node[label=above:{10 fm}] (a) at (0, 1.1) {};
+    \node (b) at (0, -1.1) {};
+    \draw [<->, semithick] (a) to (b);
+  \end{tikzpicture}
+  \caption{\label{fig:thickness} Effect of nucleon substructure on the nuclear thickness function $T(x, y) \equiv \int dz\, \rho(x, y, z)$ of a $^{208}\mathrm{Pb}$ nucleus.
+The nucleus on the left has Gaussian nucleons of width $0.8$~fm, while the nucleus on the right has composite nucleons, each containing six constituents of width $0.4$~fm.}
+\end{figure}
+
+Up to this point we have restricted our attention to a single proton-proton collision.
+Equation~\eqref{eq:gmean} is a purely local function of nuclear density in the transverse plane and hence it should (in principle) be equally valid for any pair of colliding nuclei at sufficiently high beam energy.
+The \trento\ model readily generalizes from individual proton-proton collisions to arbitrary nucleus-nucleus collisions by summing the participant thicknesses $\T_{A,B}$ in Eq.~\eqref{eq:fluctuated_thick} over all nucleons which participate in one or more inelastic collisions.
 The only modeling difference between $p$-$p$, $p$-Pb, and Pb-Pb collisions is the number and the position of the nucleons.
+Figure~\ref{fig:thickness} shows the qualitative effect of adding nucleon substructure to a lead nucleus.
+Fluctuations emerge over the length scale of a nucleon, but the macroscopic geometry of the ion is largely unchanged.
 
@@
 If a sampled nucleon position is too close to any of its previously placed neighbors, its solid angle is resampled until the minimum distance criteria is satisfied.
-Similar methods could be used to model correlations between individual constituents inside each nucleon, although the numerical implementation would be somewhat tedious.
-
-Figure~\ref{fig:thickness} shows the qualitative effect of nucleon substructure on a generic lead nuclear thickness function.
-
-\subsection{Pre-equilibrium dynamics}
-
-\begin{figure}[b]
+Similar methods could be used to model correlations between individual constituents inside each nucleon, although the implementation would be somewhat tedious.
+
+\begin{figure}
   \includegraphics{coupling}
@@
 
+\subsection{Pre-equilibrium dynamics}
+
 There are of course two limiting cases for the strength of interactions inside the QGP medium immediately at the collision: infinitely weak coupling where the secondary partons free-stream without interacting, and infinitely strong coupling where the fluid's inter-particle mean free path effectively vanishes.
@@
 through its dependence on the underlying distribution function $f(x, p)$, and hence are necessary inputs for any free-streaming implementation.
-Equation~\eqref{freestream}, however, simplifies for massless partons with momentum distributions that are locally isotropic.
-Under this assumption, it was shown in Refs.~\cite{Broniowski:2008qk, Liu:2015nwa} that the energy-momentum tensor $T^{\mu\nu}$ of partons at midrapidity and time $\tau$ only depends on its spatial distribution at some earlier time $\tau_0$, but not its $p_\perp$-distribution, which could vary as a function of position.
-
-The entropy density for a gas of massless noninteracting particles is very nearly proportional to its particle density, and so we are able to recast the parametrization in Eq.~\eqref{eq:gmean} in the form
-\begin{equation}
-  \frac{dN}{d^2x_\perp d\eta}\,\bigg\vert_{\eta=0} = \, \text{Norm} \times \bigg(\frac{\T_A^p + \T_B^p}{2}\bigg)^{1/p},
-\end{equation}
-where the left-side is the initial density of free-streaming partons at midrapidity.
-The resulting free-streamed energy-momentum tensor $T^{\mu\nu}(x, y, \eta, \tau)$ at transverse coordinate $(x, y)$, midrapidity $\eta=0$, and proper time $\tau > \tau_0$ is then given by
+Equation~\eqref{freestream}, however, simplifies for a boost-invariant gas of massless noninteracting partons with locally isotropic transverse momentum distributions.
+Under these conditions, the energy-momentum tensor $T^{\mu\nu}(\xv_\perp, \eta_s=0, \tau)$ at transverse coordinate $\xv_\perp$ and time $\tau > \tau_0$ is given by \cite{Broniowski:2008qk, Liu:2015nwa},
 \begin{multline}
   \label{energy-momentum}
-  T^{\mu\nu}(x, y, 0, \tau) = \frac{1}{\tau} \int_{0}^{2\pi} d\phi\, n(x - \Delta \tau \cos \phi, y - \Delta \tau \sin \phi) \\ \times
-  \begin{bmatrix}
-    1 & \cos\phi & \sin\phi  & 0\\
-    \cos\phi & \cos^2\phi & \cos\phi\sin\phi & 0 \\
-    \sin\phi & \sin\phi \cos\phi & \sin^2\phi & 0\\
-    0 & 0 & 0 & 0
-  \end{bmatrix},
+  T^{\mu\nu}(\xv_\perp, \eta_s=0, \tau) =\\
+  \frac{\tau_0}{\tau} \int_{-\pi}^{\pi} d\phi_q\, \hat{q}^\mu \hat{q}^\nu e(\xv_\perp - (\tau - \tau_0)\hat{\mathbf{q}}_\perp, \eta_s=0, \tau_0),
 \end{multline}
-where $n$ is the local density of massless partons, and $\Delta\tau$ is the free-streaming time $\Delta\tau = \tau - \tau_0$.
-The solution \eqref{energy-momentum} can then be decomposed in hydrodynamic form
+where $\hat{q}^\mu = (1, \cos \phi_q, \sin \phi_q, 0)$ and $\hat{\mathbf{q}}_\perp = (\cos \phi_q, \sin \phi_q)$.
+Here we assume that $\tau_0 \ll 1~\fmc$ is small and define $\taufs \equiv \tau - \tau_0 \approx \tau$.
+We also combine the unknown constant in Eq.~\eqref{eq:gmean} with the pre-factor $\tau_0$ in Eq.~\eqref{energy-momentum} to produce a single normalization factor $\text{Norm} = \text{const} \times \tau_0$ with units of energy.
+The solution \eqref{energy-momentum} is then decomposed in hydrodynamic form,
 \begin{equation}
@@
 
-We evolve the system hydrodynamically down to a pre-specified switching isotherm $T_{sw}$ at which point the medium is converted into particles using the Cooper-Frye formula \cite{PhysRevD.10.186}
+We evolve the system hydrodynamically down to a pre-specified switching isotherm $\Tsw$ at which point the medium is converted into particles using the Cooper-Frye formula \cite{PhysRevD.10.186}
 \begin{equation}
@@
 The selected parameter ranges are intentionally wide to avoid clipping the calibrated posterior; for example, a previous analysis of the \trento\ model \cite{Bernhard:2016tnd} found $p \sim 0$, but we use a prior range $p \in [-1, 1]$ to account for differences in the present model, e.g.\ nucleon substructure, which could modify its posterior.
-Several of the model parameters require special care and are reparametrized accordingly:
-
-\begin{figure}
-  \begin{tikzpicture}
-    \draw[dashed] (0,0) circle (1.6cm);
-    \foreach \theta in {60, 180, 300}{
-      \draw ({\theta}:.85) circle (.73cm);
-    }
-    \coordinate (a) at (0, 0);
-    \coordinate (b) at (.8, 1.385);
-    \coordinate (c) at (.425, -.736);
-    \coordinate (d) at (-1.58, 0);
-    \coordinate (e) at (-.85, 0);
-    \draw[<->] (a) -- (b) node[midway, xshift=-1ex, yshift=.8ex]{$w$};
-    \draw[<->] (a) -- (c) node[midway, xshift=1ex, yshift=.8ex]{$r$};
-    \draw[<->] (d) -- (e) node[midway, yshift=1ex]{$v$};
-  \end{tikzpicture}
-  \caption{
-    \label{fig:nucleon_schematic}
-    Schematic illustrating the constituent sampling radius $r$, constituent width $v$, and nucleon width parameter $w = \sqrt{r^2 + v^2}$ for a nucleon with $n_c=3$ constituents.
-  }
-\end{figure}
-
-\begin{itemize}[leftmargin=1\parindent]
-  \item
+
 The constituent fluctuations, modeled by the gamma random weights $\gamma_i$ in Eq.~\eqref{eq:fluctuated_thick}, generate overall nucleon fluctuations which are suppressed by the number of constituents $n_c$ inside the nucleon.
@@
 where $k$ is the shape parameter of the gamma distribution used to fluctuate each individual constituent.
-  \item
-    In Sec.~\ref{sec:initial_state} we parametrized nucleon substructure using three degrees of freedom:
-    \begin{enumerate*}[label=(\roman*)]
-      \item a parameter $v$ for the Gaussian width of each constituent,
-      \item a parameter $r$ for the Gaussian width of the radial distribution used to sample the constituent centers, and
-      \item a parameter $n_c$ to vary the number of constituents inside the nucleon.
-    \end{enumerate*}
-    In the limit $n_c \to \infty$, the composite nucleon's root-mean-square (RMS) radius is simply the convolution of its sampling radius $r$ and constituent width $v$ which add together in quadrature
-    \begin{equation}
-      \lim_{n_c \to\infty} r_\mathrm{RMS}\{\rho\} = \sqrt{r^2 + v^2}.
-    \end{equation}
-    We therefore choose to reparametrize the sampling radius $r$ in terms of a new variable
-    \begin{equation}
-      \label{eq:nucleon_width}
-      w = \sqrt{r^2 + v^2},
-    \end{equation}
-    which approximates the RMS radius of the nucleon when the number of sampled constituents is large (see Fig.~\ref{fig:nucleon_schematic} for an example proton).
-    We thus call $w$ a nucleon ``width'', although for smaller numbers of constituents, the \emph{actual} RMS radius of our sampled nucleons can be significantly smaller than our width parameter $w$ due to fluctuations in the nucleon's center of mass, and hence one should account for the difference when discussing the nucleon's posterior RMS radius.
-  \item
-    Equation~\eqref{eq:nucleon_width} requires the nucleon width to be larger than the constituent width, i.e.
-$w > v$, lest the sampling radius $r$ turn imaginary, and thus we cannot vary the nucleon width $w$ and the constituent width $v$ independently.
-    We therefore reparametrize the constituent width $v$ using a new variable $\X$ to interpolate between minimum and maximum allowed values:
-    \begin{equation}
-      \label{eq:struct_param}
-      v = v_\mathrm{min} + \X (v_\mathrm{max} - v_\mathrm{min}).
-    \end{equation}
-    Here we choose a minimum constituent width $v_\mathrm{min} = 0.2$~fm determined by computational limits and a maximum width $v_\mathrm{max} = w$ equal to the nucleon width.
-    Thus for $\X=0$, the nucleons consist of small, distinct hot spots, whereas for $\X=1$ the nucleon is a single Gaussian blob of width $w$.
-    The parameter $\X$ thus varies the granularity of the nucleon while keeping the number of constituents $n_c$ and approximate size of the nucleon $w$ fixed.
-\end{itemize}
 
@@
 These high-multiplicity bins are too selective for our modest minimum-bias event sample, and so a different procedure is required.
-We exploit, for this purpose, the approximate correspondence
-\begin{equation}
-  \label{eq:bin_equiv}
-  \ntrk \sim \nch \propto N_\text{parton}
-\end{equation}
-between each event's initial secondary-parton density ${N_\text{parton} = \int d^2x\,n(\xv)}$ in Eq.~\eqref{energy-momentum} and its final-state event activity, characterized by $\nch$ or $\ntrk$ \cite{Abelev:2013bla, Chatrchyan:2013nka}.
+We exploit, for this purpose, the approximate monotonic relationship between each event's initial energy density $dE/d\eta_s$ and its final charged-particle density $d\ntrk/d\eta_s \sim d\nch/d\eta_s$ at midrapidity \cite{Abelev:2013bla, Chatrchyan:2013nka}.
 
 Consider, for example, an experimental multiplicity bin $(\nch^\text{low}, \nch^\text{high})$ with some kinematic cuts on $p_T$ and $\eta$.
-We rescale this bin by the average multiplicity $\langle \nch \rangle$ of the corresponding minimum-bias event sample, i.e.\
-\begin{equation}
-  (\nch^\text{low}, \nch^\text{high}) \rightarrow (\nch^\text{low} / \langle \nch \rangle,\, \nch^\text{high} / \langle \nch \rangle)
+We first rescale this bin by the average charged-particle multiplicity $\langle \nch \rangle$ of the full minimum-bias event sample,
+\begin{equation}
+  (\nch^\text{low}, \nch^\text{high}) \rightarrow (\nch^\text{low} / \langle \nch \rangle,\, \nch^\text{high} / \langle \nch \rangle),
 \end{equation}
 in order to reexpress each bin edge as a unitless variable.
-These rescaled bins are then used to select initial condition events using the equivalence \eqref{eq:bin_equiv}:
+These rescaled bins are then used to select initial condition events according to their midrapidity energy density:
 \begin{equation}
   \label{eq:mult_trigger}
-  \left (\frac{N^\text{low}_\text{parton}}{\langle N_\text{parton} \rangle}, \frac{N^\text{high}_\text{parton}}{\langle N_\text{parton} \rangle} \right) \leftrightarrow \left (\frac{\nch^\text{low}}{\langle \nch \rangle}, \frac{\nch^\text{high}}{\langle \nch \rangle} \right ).
+  \left (\frac{(dE/d\eta_s)_\text{low}}{\langle dE/d\eta_s \rangle}, \frac{N^\text{high}_\text{parton}}{\langle N_\text{parton} \rangle} \right) \leftrightarrow \left (\frac{\nch^\text{low}}{\langle \nch \rangle}, \frac{\nch^\text{high}}{\langle \nch \rangle} \right ).
 \end{equation}

### 2018-10-16 paper/paper.tex
Finish revisions through results dicussion and conclusion.

@@
 Ultrarelativistic nuclear collisions between one light-ion and one heavy-ion, e.g.\ $^3$He-Au and $p$-Pb collisions, generate dense, compact sources of nuclear matter which produce long-range multiparticle correlations that are strikingly similar to the correlations observed in heavy-ion collisions where collectivity is commonly explained by the existence of hydrodynamic flow \cite{CMS:2012qk, Abelev:2012ola, Aad:2012gla, Adare:2015ctn}.
-This observation suggests that hydrodynamic behavior could be manifest in small droplets of quark-gluon-plasma (QGP) \cite{Bozek:2011if, Bozek:2013uha}, and that flow even might develop at length scales smaller than a single proton \cite{Schenke:2014zha}.
+This observation suggests that hydrodynamic behavior could be manifest in small droplets of quark-gluon-plasma (QGP) \cite{Bozek:2011if, Bozek:2013uha}, and that flow might even develop at length scales smaller than a single proton \cite{Schenke:2014zha}.
 
@@
 Generally speaking, the initial three-dimensional distribution of matter produced in relativistic nuclear collisions is \emph{not} boost-invariant.
-The longitudinal density of particle production fluctuates both locally point-to-point in the transverse plane as well as globally event-by-event due to asymmetries in the sampled density of participant matter \cite{Ke:2016jrd, Bozek:2010vz}.
+The longitudinal particle density fluctuates both locally point-to-point in the transverse plane as well as globally event-by-event due to asymmetries in the sampled density of participant matter \cite{Ke:2016jrd, Bozek:2010vz}.
 Nevertheless, boost-invariance has been shown to be a good approximation for both large and small collision systems when hydrodynamic observables are calculated from particles that are detected close to midrapidity \cite{Shen:2016zpp}.
@@
 The resulting nuclear overlap time $\Delta t \lesssim 0.1\ \fmc$ at top RHIC and LHC energies, and thus it is safe to neglect the initial transverse dynamics which occur while the nuclei pass through each other.
-We therefore assume that the collision produces all secondary particles at a uniform proper time $\tau_0 \ll 1\fmc$, and that it deposits energy at midrapidity which is a function of the locally varying beam-integrated density of each nucleus.
+We therefore assume that the collision produces all secondary particles at a uniform proper time $\tau_0 \ll 1~\fmc$, and that it deposits energy at midrapidity which is a function of the locally varying density inside each nucleus.
 
@@
 As a matter of convenience, we sample the proton positions \emph{before} determining the constituent positions.
-This leads to small discrepancies between the designated proton positions $\xv_{A,B}$, and the actual position of each proton's center-of-mass, owing to fluctuations in the constituent positions $\xv_i$.
+This leads to small discrepancies between the designated proton positions $\xv_{A,B}$ and the actual position of each proton's center-of-mass, owing to fluctuations in the constituent positions $\xv_i$.
 The sampling radius $r$ should thus be interpreted with care.
@@
 Hereafter, we switch to Milne coordinates $(\xv_\perp, \eta_s, \tau)$, where $\xv_\perp$ is a Cartesian coordinate $(x, y)$ in the transverse plane, $\eta_s = \tfrac{1}{2} \log[(t+z)/(t-z)]$ is the system's spacetime rapidity, and $\tau = \sqrt{t^2 - z^2}$ its proper time.
-Assuming our protons collide at the sampled impact parameter $\bv$, we assign each proton a participant thickness,
+Assuming the protons collide at the sampled impact parameter $\bv$, we assign each proton a participant thickness
 \begin{equation}
@@
 \begin{equation}
-  \tilde{\rho}_{A,B}(\xv) \equiv \frac{1}{n_c} \sum\limits_{i=1}^{n_c} \gamma_i\, \rho_c(\xv - \xv_i),
+  \tilde{\rho}_{A,B}(\xv) \equiv \frac{1}{n_c} \sum\limits_{i=1}^{n_c} \gamma_i\, \rho_c(\xv - \xv_i)
 \end{equation}
@@
 This fluctuated proton density $\tilde{\rho}_{A,B}$ equals the proton density $\rho_{A,B}(\xv)$ in Eq.~\eqref{eq:rho} with additional weights $\gamma_i$ sampled from a gamma distribution with unit mean and variance $1/k$.
-These \emph{ad hoc} gamma random weights are necessary to describe the large fluctuations observed in minimum bias proton-proton collisions, although their exact physical origin is not well understood.
+Such \emph{ad hoc} random weights are necessary to describe the large fluctuations observed in minimum bias proton-proton collisions, although their exact physical origin is not well understood.
+
+The resulting nucleon fluctuation variance naturally falls like ${\sim}1/ n_c$, where $n_c$ is the number of independent fluctuating sources.
+This means that the natural range for the constituent fluctuations is larger when the number of constituents is larger and \emph{vice versa}.
+We therefore reparametrize the constituent fluctuations using,
+\begin{equation}
+  \sigma_\mathrm{fluct} = 1 / \sqrt{k\, n_c},
+\end{equation}
+where $k$ is the shape parameter (inverse variance) of the gamma distribution weights and $n_c$ is the number of nucleon constituents.
 
@@
 This wounded nucleon scaling was in fact one of the first such mappings used as a proxy for initial particle production in relativistic heavy-ion collisions \cite{Bialas:1976ed}.
-It was quickly realized, however, that the wounded nucleon model predicts the wrong scaling for charged particle production as a function of collision centrality and hence the wrong scaling for initial energy deposition at midrapidity as a function of participant thicknesses $\T_A$ and $\T_B$ \cite{Kharzeev:2000ph} \footnote{Typically an entropy density $s$ is used, but the shortcomings of the wounded nucleon model are nonetheless the same when depositing energy.}.
+It was quickly realized, however, that the wounded nucleon model predicts the wrong scaling for charged particle production as a function of collision centrality \cite{Kharzeev:2000ph} and hence the wrong scaling for initial energy deposition at midrapidity as a function of participant thicknesses $\T_A$ and $\T_B$\footnote{Typically an entropy density $s$ is used, but the shortcomings of the wounded nucleon model are qualitatively the same when depositing energy.}.
 
@@
 
-We note that previous applications of the \trento\ model \cite{Moreland:2014oya, Bernhard:2016tnd} used the generalized mean ansatz to parametrize the \emph{entropy} density $s(\xv_\perp, \eta_s=0, \tau_\text{hydro})$ at midrapidity, motivated (in part) by the approximate linear relation between initial entropy density $(dS/d\eta_s)\vert_{\eta_s=0}$ and final charged-particle density $(d\nch/d\eta_s)\vert_{\eta_s=0}$ \cite{Shen:2015qta}.
+We note that previous applications of the \trento\ model \cite{Moreland:2014oya, Bernhard:2016tnd} used the generalized mean ansatz to parametrize the \emph{entropy} density $s(\xv_\perp, \eta_s=0, \tau_\text{hydro})$, motivated (in part) by the approximate linear relation at midrapidity between initial entropy density $dS/d\eta_s$ and final charged-particle density $d\nch/d\eta_s$ \cite{Shen:2015qta}.
 These studies, however, assumed instant thermalization, an approximation which sets $u^\mu$, $\Pi$, and $\pi^{\mu\nu}$ identically to zero at the hydrodynamic starting time $\tau_\text{hydro}$.
-The present study, meanwhile, uses a pre-equilibrium free streaming stage to estimate $u^\mu$, $\Pi$, and $\pi^{\mu\nu}$ as a function of a variable free-streaming time $\taufs$.
-Strictly speaking, the entropy density ${s(\xv_\perp, \eta_s=0, \tau)}$ is ill defined prior to thermalization.
-Hence, it would not make sense to directly parametrize the entropy density at times $\tau_0 < \tau_\text{hydro}$ as required by the present study.
-We therefore reinterpret the quantity produced by the \trento\ model as an energy density $e(\xv_\perp, \eta_s=0, \tau_0)$ and free-stream this quantity to the hydrodynamic starting time $\tau_\text{hydro}$.
+The present study, meanwhile, uses a pre-equilibrium free streaming stage to estimate $u^\mu$, $\Pi$, and $\pi^{\mu\nu}$ as a function of a variable free-streaming time $\taufs$, at which point the system is matched to hydrodynamics.
+Strictly speaking, the system's entropy is ill defined prior to thermalization while the system free streams.
+We therefore reinterpret the quantity produced by the \trento\ model as an energy density $e(\xv_\perp, \eta_s=0, \tau_0)$ and free-stream this quantity to the hydrodynamic starting time $\tau_\text{hydro} \equiv \taufs$.
 This assumes, of course, that the model's parametrization for entropy density is also suitable as a parametrization for the energy density.
@@
 For example, both are approximately conserved at midrapidity, and both scale approximately linearly with the final charged-particle density $d\nch/d\eta_s$.
+We thus expect the parametrization to be suitable for either quantity.
 
@@
 The only modeling difference between $p$-$p$, $p$-Pb, and Pb-Pb collisions is the number and the position of the nucleons.
-Figure~\ref{fig:thickness} shows the qualitative effect of adding nucleon substructure to a lead nucleus.
-Fluctuations emerge over the length scale of a nucleon, but the macroscopic geometry of the ion is largely unchanged.
+Figure~\ref{fig:thickness} shows the effect of adding nucleon substructure to a generic lead nucleus.
+Additional fluctuations emerge over the length scale of a nucleon, but the macroscopic geometry of the nucleus is largely unchanged.
 
@@
 
-The parametric entropy deposition ansatz in Eq.~\eqref{eq:gmean} does not provide any information about the initial masses or momenta of particles produced in the collision.
+The parametric energy deposition ansatz in Eq.~\eqref{eq:gmean} does not provide any information about the initial masses or momenta of particles produced in the collision.
 In general, these details will affect the dynamics predicted by the collisionless Boltzmann equation
@@
 \end{equation}
-through its dependence on the underlying distribution function $f(x, p)$, and hence are necessary inputs for any free-streaming implementation.
+through its dependence on the underlying distribution function $f(x, p)$.
 Equation~\eqref{freestream}, however, simplifies for a boost-invariant gas of massless noninteracting partons with locally isotropic transverse momentum distributions.
-Under these conditions, the energy-momentum tensor $T^{\mu\nu}(\xv_\perp, \eta_s=0, \tau)$ at transverse coordinate $\xv_\perp$ and time $\tau > \tau_0$ is given by \cite{Broniowski:2008qk, Liu:2015nwa},
+Subject to these assumptions, the energy-momentum tensor $T^{\mu\nu}(\xv_\perp, \eta_s=0, \tau)$ at transverse coordinate $\xv_\perp$ and time $\tau > \tau_0$ equals \cite{Broniowski:2008qk, Liu:2015nwa},
 \begin{multline}
@@
 \end{equation}
-where $e$ and $P$ are the energy density and pressure in the local fluid rest frame, $u^\mu$ is the local fluid velocity, ${\Delta^{\mu\nu} \equiv g^{\mu\nu} - u^\mu u^\nu}$ is the projector onto the space orthogonal to $u^\mu$, and $\Pi$ and $\pi^{\mu\nu}$ are the bulk pressure and shear stress tensor respectively.
+where $e$ and $P$ are the energy density and pressure in the local fluid rest frame, $u^\mu$ is the local fluid velocity, ${\Delta^{\mu\nu} = g^{\mu\nu} - u^\mu u^\nu}$ is the projector onto the space orthogonal to $u^\mu$, and $\Pi$ and $\pi^{\mu\nu}$ are the bulk pressure and shear stress tensor respectively.
 We then solve for the energy density $e$ and fluid velocity $u^\mu$ using the Landau matching condition which defines the fluid rest frame velocity as the time-like eigenvector of $T^{\mu\nu}$ with energy density $e$ as its eigenvalue,
@@
 The aforementioned hydrodynamic equations are solved numerically using the boost-invariant VISH2+1 viscous hydrodynamics code \cite{Song:2007ux, Shen:2014vra}.
-We simulate each hydrodynamic event on a spacetime grid with transverse extent $x_\mathrm{max}$, spatial grid step $dx$, and time step $d\tau$ which are optimized event-by-event to balance trade-offs between numerical accuracy and computation time (see Appendix~\ref{app:adaptive_grid}).
-Although these details are somewhat mundane, they are critically important to the present study, since the computation time scales with the number of spacetime cells $n_x^2\, n_\tau$, and $n_x$ and $n_\tau$ typically have to be quite large to resolve the small length scales associated with nucleon substructure.
+We vary the spatial grid's maximum width $x_\mathrm{max}$ event-by-event to accommodate systems of varying size and determine the optimal spatial step $dx$ and time step $d\tau$ for each design point to balance trade-offs between numerical accuracy and computation time (see Appendix~\ref{app:adaptive_grid}).
+Although these details are somewhat mundane, they are critically important to the present study, since the computation time scales with the number of spacetime cells $n_\text{cell} \sim n_x^2\, n_\tau$, and this quantity grows rapidly when $n_x$ and $n_\tau$ are large, as is typically the case for simulations with nucleon substructure.
 
@@
 For example, $\yv_m$ might be a vector consisting of charged particle yields in different centrality bins.
-The physics model thus represents a vector-valued function $f(\xv) = \yv_m$ which maps the parameter values $\xv$ to the simulated observables $\yv_m$.
+The physics model thus represents a vector-valued function $f(\xv) = \yv_m$ which maps the parameter values $\xv$ to the model observables $\yv_m$.
 
@@
     $\sigmaf$         & Nucleon fluctuation std.\ dev.\      & 0--2            \\
-    $w$               & Nucleon width parameter              & 0.4--1.2 fm     \\
+    $r$               & Constituent sampling radius          & 0--1.2 fm       \\
     $n_c$             & Number of nucleon constituents       & 1--9            \\
-    $\X$              & Nucleon structure parameter          & 0--1            \\
+    $v$               & Constituent width                    & 0.2--1.2 fm     \\
     $\dmin$           & Minimum inter-nucleon distance       & 0--1.7 fm       \\
@@
 
-\subsection{Parameter design}
+\subsection{Parameter design and observables}
+\label{sec:observables}
 
@@
 We assume the prior distribution is constant and nonzero within each specified range and zero otherwise.
-The selected parameter ranges are intentionally wide to avoid clipping the calibrated posterior; for example, a previous analysis of the \trento\ model \cite{Bernhard:2016tnd} found $p \sim 0$, but we use a prior range $p \in [-1, 1]$ to account for differences in the present model, e.g.\ nucleon substructure, which could modify its posterior.
-
-The constituent fluctuations, modeled by the gamma random weights $\gamma_i$ in Eq.~\eqref{eq:fluctuated_thick}, generate overall nucleon fluctuations which are suppressed by the number of constituents $n_c$ inside the nucleon.
-The observed nucleon fluctuation variance falls like $1 / n_c$ which means the natural range for the constituent fluctuations is larger when the number of constituents is larger and \emph{vice versa}.
-We therefore reparametrize the constituent fluctuations using the standard deviation of the resulting nucleon fluctuations
-\begin{equation}
-  \sigma_\mathrm{fluct} = 1 / \sqrt{k\, n_c},
-\end{equation}
-where $k$ is the shape parameter of the gamma distribution used to fluctuate each individual constituent.
-
-\subsection{Observables}
-\label{sec:observables}
+The selected parameter ranges are intentionally wide to avoid clipping the calibrated posterior.
+For example, a previous analysis of the \trento\ model \cite{Bernhard:2016tnd} found $p \sim 0$, but we use a prior range $p \in [-1, 1]$ to account for differences in the present model, e.g.\ nucleon substructure, which could modify its posterior.
+One exception is the constituent number $n_c$ which we limit for practical considerations.
+Recall that each constituent fluctuates independently, weighted by a gamma random variable.
+Hence for constituent numbers $n_c \gg 1$, the fluctuations average out, and the resulting nucleon fluctuations vanish.
+To counteract this effect, the constituent fluctuation variance must increase as $n_c$ increases.
+Eventually, these required fluctuations become unreasonably large.
+We find that for $n_c < 10$, the energy density fluctuations are reasonable, and hence limit $n_c$ to this prior range.
 
@@
 
-However, for some observables such as $p$-Pb mean $p_T$ \cite{Abelev:2013bla} and flow cumulants $\vnk{n}{k}$ \cite{Chatrchyan:2013nka}, the experiments use a special high-multiplicity trigger to select rare, ultra-central events according to the number of charged particles produced or detector tracks offline.
-These high-multiplicity bins are too selective for our modest minimum-bias event sample, and so a different procedure is required.
-We exploit, for this purpose, the approximate monotonic relationship between each event's initial energy density $dE/d\eta_s$ and its final charged-particle density $d\ntrk/d\eta_s \sim d\nch/d\eta_s$ at midrapidity \cite{Abelev:2013bla, Chatrchyan:2013nka}.
-
-Consider, for example, an experimental multiplicity bin $(\nch^\text{low}, \nch^\text{high})$ with some kinematic cuts on $p_T$ and $\eta$.
-We first rescale this bin by the average charged-particle multiplicity $\langle \nch \rangle$ of the full minimum-bias event sample,
-\begin{equation}
-  (\nch^\text{low}, \nch^\text{high}) \rightarrow (\nch^\text{low} / \langle \nch \rangle,\, \nch^\text{high} / \langle \nch \rangle),
+However, for some observables such as $p$-Pb mean $p_T$ \cite{Abelev:2013bla} and flow cumulants $\vnk{n}{k}$ \cite{Chatrchyan:2013nka}, the experiments use a special high-multiplicity trigger to select rare, ultra-central events according to the number of charged-particles produced $\nch$ or detector tracks offline $\ntrk$.
+These high-multiplicity bins are too selective for our modest minimum bias event sample, and so a different procedure is required.
+We exploit for this purpose the approximate monotonic dependence of the midrapidity charged-particle yield on the transverse energy density,
+\begin{equation}
+  E_T = \int d^2x_\perp e(\xv_\perp, \eta_s=0, \tau_0).
+\end{equation}
+
+Consider, for example, an single multiplicity bin $[\nch^\text{low}, \nch^\text{high}]$ subselected from a minimum bias event sample with some kinematic cuts, $p_T^\text{min} < p_T < p_T^\text{max}$ and $|\eta| < \eta^\text{max}$.
+Let $\langle\nch\rangle$ denote the average charged-particle multiplicity of the minimum bias event sample.
+We first rescale the experimental multiplicity bin edges,
+\begin{equation}
+  [\nch^\text{low}, \nch^\text{high}] \rightarrow \left[\frac{\nch^\text{low}}{\langle \nch \rangle},\, \frac{\nch^\text{high}}{\langle \nch \rangle}\right ],
 \end{equation}
 in order to reexpress each bin edge as a unitless variable.
-These rescaled bins are then used to select initial condition events according to their midrapidity energy density:
+These rescaled bins are then associated with the analogous bins edges defined using the \trento\ model's initial transverse energy density,
 \begin{equation}
   \label{eq:mult_trigger}
-  \left (\frac{(dE/d\eta_s)_\text{low}}{\langle dE/d\eta_s \rangle}, \frac{N^\text{high}_\text{parton}}{\langle N_\text{parton} \rangle} \right) \leftrightarrow \left (\frac{\nch^\text{low}}{\langle \nch \rangle}, \frac{\nch^\text{high}}{\langle \nch \rangle} \right ).
-\end{equation}
-
-Finally, we mimic the method used by experiment and apply \eqref{eq:mult_trigger} to select rare high-multiplicity events from a continuous stream of minimum-bias \trento\ events satisfying the correct relative multiplicity bin edges.
+  \left [\frac{E_T^\text{min}}{\langle E_T \rangle}, \frac{E_T^\text{max}}{\langle E_T \rangle} \right] \leftrightarrow \left [\frac{\nch^\text{low}}{\langle \nch \rangle}, \frac{\nch^\text{high}}{\langle \nch \rangle} \right ],
+\end{equation}
+where $\langle E_T \rangle$ denotes the average transverse energy density of the minimum bias event sample.
+
+Finally, we mimic the method used by experiment and apply \eqref{eq:mult_trigger} to select rare high-multiplicity events from a continuous stream of minimum-bias \trento\ events satisfying the correct bin edges.
 This of course means that, in addition to running a large sample of minimum-bias events for centrality binned observables, we must also generate (much like experiment) a separate sample of multiplicity triggered events.
 In practice, we use a few hundred to a few thousand events per multiplicity bin, depending on the type of observable.
+
+\subsection{Experimental uncertainties}
 
@@
 \begin{equation}
-  \mathrm{corr}(y_i, y_j) = \cov(y_i, y_j)/(\sigma_i \sigma_j)
+  \mathrm{corr}(y_i, y_j) = \cov(y_i, y_j)/(\sigma_i \sigma_j),
 \end{equation}
@@
 The emulator enables essentially instantaneous predictions for $\yv_m = f(\xv)$ and allows us to sample the posterior distribution millions of times.
-In order to train the emulator, we first generate a scaffolding of the parameter space using a maximin Latin hypercube design \cite{Morris:1995lh} to distribute 500 points uniformly throughout our 15-dimensional parameter space according to the parameter ranges in Table~\ref{tab:design}.
+In order to train the emulator, we first generate a scaffolding of the parameter space using a maximin Latin hypercube design \cite{Morris:1995lh} to distribute 500 points semi-randomly throughout our 15-dimensional parameter space according to the parameter ranges in Table~\ref{tab:design}.
 We then run minimum-bias and multiplicity triggered $p$-Pb and Pb-Pb events at each design point and calculate the model observables from the ensemble of events.
@@
 For the present study, we use $k=7$ principal components when emulating the $p$-Pb system and $k=8$ principal components when emulating Pb-Pb, chosen to describe 99.5\% of the total observed variance of each system.
+
+The GPE's are essentially fancy interpolators applied to the model's training points and PCA transformed observables.
+Each GPE reports a mean value $z(\xv)$ as well as an estimated error $\delta z(\xv)$ which accounts for statistical noise in the training data and interpolation error between the design points.
+Once the GPE's are trained, we can predict the observables $\yv_m$ at parameter point $\xv$ by transforming the vector of principal components
+\begin{equation}
+  \mathbf{z}(\xv) = (z_1(\xv), z_2(\xv), \dots, z_k(\xv))
+\end{equation}
+\emph{back} to physical space.
+Similarly, we can construct the covariance matrix of the observables in PCA space,
+\begin{equation}
+  \cov(z_i, z_j) = \diag[(\delta z_{1})^2, (\delta z_2)^2, \dots, (\delta z_k)^2 ],
+\end{equation}
+and transform it back to physical space as well to obtain the covariance matrix of the model observables $\yv_m$ at a given parameter point $\xv$.
+
+The resulting emulator therefore predicts both a mean prediction $\yv_m(\xv)$ and an uncertainty covariance matrix $\Sigma_m(\xv)$ which accounts for multiple sources of model and emulator uncertainty, including the truncation error expected from using a finite number of principal components $k < m$.
+The model covariance matrix $\Sigma_m$ includes so-called \emph{known}-unknowns such as statistical error and emulator interpolation error, but not \emph{unknown}-unknowns such as the overall validity of small-system hydrodynamics, i.e.\ things which lack a unified consensus or are difficult to quantify.
+We show in Fig.~\ref{fig:correlation} the resulting Pb-Pb correlation matrix $\mathrm{corr}(y_i,y_j)$ for the model (emulator) at a random parameter point $\xv$ in the design space (left-side), along side the same correlation matrix for the experimental data (right-side) discussed previously.
+For additional information on the model emulator, we direct the reader to Appendix~\ref{app:validation} which includes several validation tests of the emulator prediction accuracy.
+\newpage
 
@@
 
-The GPE's are essentially fancy interpolators applied to the model's training points and PCA transformed observables.
-Each GPE reports a mean value $z(\xv)$ as well as an estimated error $\delta z(\xv)$ which accounts for statistical noise in the training data and interpolation error between the design points.
-Once the GPE's are trained, we can predict the observables $\yv_m$ at parameter point $\xv$ by transforming the vector of principal components
-\begin{equation}
-  \mathbf{z}(\xv) = (z_1(\xv), z_2(\xv), \dots, z_k(\xv))
-\end{equation}
-\emph{back} to physical space.
-Similarly, we can construct the covariance matrix of the observables in PCA space,
-\begin{equation}
-  \cov(z_i, z_j) = \diag[(\delta z_{1})^2, (\delta z_2)^2, \dots, (\delta z_k)^2 ],
-\end{equation}
-and transform it back to physical space as well to obtain the covariance matrix of the model observables $\yv_m$ at a given parameter point $\xv$.
-
-The resulting emulator therefore predicts both a mean prediction $\yv_m(\xv)$ and an uncertainty covariance matrix $\Sigma_m(\xv)$ which accounts for multiple sources of model and emulator uncertainty, including the truncation error expected from using a finite number of principal components $k < m$.
-The model covariance matrix $\Sigma_m$ includes so-called \emph{known}-unknowns such as statistical error and emulator interpolation error, but not \emph{unknown}-unknowns such as the overall validity of small-system hydrodynamics, i.e.\ things which lack a unified consensus or are difficult to quantify.
-We show in Fig.~\ref{fig:correlation} the resulting Pb-Pb correlation matrix $\mathrm{corr}(y_i,y_j)$ for the model (emulator) at a random parameter point $\xv$ in the design space (left-side), along side the same correlation matrix for the experimental data (right-side) discussed previously.
-For additional information on the model emulator, we direct the reader to Appendix~\ref{app:validation} which includes several validation tests of the emulator prediction accuracy.
-
 \subsection{Bayesian calibration}
@@
 
-We show the simulated and emulated model observables (thin colored lines) for Pb-Pb collisions in Fig.~\ref{fig:obs_pbpb} and for \mbox{$p$-Pb} collisions in Fig.~\ref{fig:obs_ppb} at $\sqrts=5.02$~TeV compared to experimental data from CMS \cite{Chatrchyan:2013nka} and ALICE \cite{Adam:2015ptt, Adam:2016izf, Adam:2014qja, Abelev:2013bla}.
+We show the simulated and emulated model observables (thin colored lines) for Pb-Pb collisions in Fig.~\ref{fig:obs_pbpb} and for \mbox{$p$-Pb} collisions in Fig.~\ref{fig:obs_ppb} at $\sqrts=5.02$~TeV compared to experimental data from the CMS \cite{Chatrchyan:2013nka} and ALICE collaborations \cite{Adam:2015ptt, Adam:2016izf, Adam:2014qja, Abelev:2013bla}.
 The top row of each figure shows explicit model calculations at each of the $d=500$ design points (training data), while the bottom row shows emulator predictions for $n=100$ random parameter samples drawn from the Bayesian posterior (sampled from the MCMC chain).
@@
 
-Conversely, the calibrated (posterior sampled) emulator predictions (bottom row of Figs.~\ref{fig:obs_pbpb} and \ref{fig:obs_ppb}) are far better constrained and nicely track the experimental data points.
-We emphasize here that the posterior parameter values are obtained from a \emph{simultaneous} calibration to $p$-Pb and Pb-Pb data, and thus they are self-consistent between the two systems.
-The spread in the posterior samples reflects different sources of model and experimental uncertainty as well as tension in the optimal fit parameters which describe each observable.
-We demonstrate later in the text that a single set of model parameters well describes all of the calibration data, and thus we believe that much of the spread in the posterior samples is uncertainty contributed by our emulator.
-We also note that although the $p$-Pb posterior samples have a somewhat larger spread than the Pb-Pb samples, the percentage uncertainty of the $p$-Pb emulator is similar to that of the Pb-Pb emulator, and thus the difference is likely due to the larger variance of the $p$-Pb training data.
-The uncertainty in the posterior distribution could thus be improved by running the calibration with more design points or with a narrower range of parameter values to increase the density of the training points and reduce interpolation uncertainty.
-
-\begin{fullpage}
-  \begin{figure}
-    \makebox[\textwidth]{\includegraphics[width=.95\paperwidth]{posterior}}
-  \end{figure}
-  \captionof{figure}{
-    \label{fig:posterior}
-    Bayesian posterior distribution of the model input parameters.
-    The diagonal panels show the marginalized distributions of individual model parameters, while off-diagonal panels show the joint distributions for pairs of model parameters, visualizing their correlations.
-    The marginalized distribution medians and 90\% credible intervals are annotated along the diagonal.
-  }
-\end{fullpage}
-
-We now direct our attention to Fig.~\ref{fig:posterior} which shows the main result of this work, the posterior distribution of the model input parameters.
-Recall that the posterior $P(H_\xv | E)$ is the probability that our hypothesis $H_\xv$ for the true model parameters $\xv_\star$ is correct, given the evidence $E$ provided by experiment.
-The present posterior has 15 dimensions, one for each parameter listed in Table~\ref{tab:design}, and thus its joint distribution cannot be summarized by one figure alone.
-We therefore sample the distribution and histogram the samples to project the distribution onto one or two dimensions at a time.
-Each diagonal panel is the distribution of a \emph{single} model parameter (marginalized over all others), and each lower-diagonal panel is the joint distribution of a \emph{pair} of model parameters (marginalized over all others).
-We also report numeric estimates for each parameter's median value and 90\% credible interval and annotate their values along the distribution diagonal (see Table~\ref{tab:post_param}).
-For example, the fictitious estimate $x=2.45_{-0.15}^{+0.20}$ reports a median value $\tilde{x}=2.45$ and 90\% credible interval $2.30 < x < 2.65$.
-
-\begin{table}
+\begin{table}[t]
   \caption{
@@
       \addlinespace[.4ex]
-      Norm       & \est{20.2}{2.2}{2.2}          & $\eta/s$ min     & \est{0.10}{0.07}{0.07}            \\[1.1ex]
+      Norm       & \est{20.0}{2.4}{2.4}          & $\eta/s$ min     & \est{0.10}{0.07}{0.07}            \\[1.1ex]
       $p$        & \est{0.013}{0.155}{0.144}     & $\eta/s$ slope   & \est{1.11}{1.11}{1.42} GeV$^{-1}$ \\[1.1ex]
       $\sigmaf$  & \est{0.93}{0.27}{0.26}        & $\eta/s$ crv     & \est{-0.14}{0.86}{0.81}           \\[1.1ex]
-      $w$        & \est{0.98}{0.17}{0.18} fm     & $\zeta/s$ max    & \est{0.026}{0.026}{0.033}         \\[1.1ex]
+      $r$        & \est{0.88}{0.23}{0.25} fm     & $\zeta/s$ max    & \est{0.026}{0.026}{0.033}         \\[1.1ex]
       $n_c$      & \est{6.3}{3.1}{2.7}           & $\zeta/s$ width  & \est{0.033}{0.033}{0.048} GeV     \\[1.1ex]
-      $\X$       & \est{0.35}{0.17}{0.22}        & $\zeta/s$ $T_0$  & \est{0.176}{0.020}{0.024} GeV     \\[1.1ex]
+      $v$       & \est{0.51}{0.18}{0.26} fm      & $\zeta/s$ $T_0$  & \est{0.176}{0.020}{0.024} GeV     \\[1.1ex]
       $\dmin$    & \est{1.13}{0.51}{0.55} fm     & $\Tsw$           & \est{0.149}{0.013}{0.014} GeV     \\[1.1ex]
@@
 
+Conversely, the calibrated (posterior sampled) emulator predictions (bottom row of Figs.~\ref{fig:obs_pbpb} and \ref{fig:obs_ppb}) are far better constrained and nicely track the experimental data points.
+We emphasize here that the posterior parameter values are obtained from a \emph{simultaneous} calibration to $p$-Pb and Pb-Pb data, and thus they are self-consistent between the two systems.
+The spread in the posterior samples reflects different sources of model and experimental uncertainty as well as tension in the optimal fit parameters which describe each observable.
+We demonstrate later in the text that a single set of model parameters well describes all of the calibration data, and thus we believe that much of the spread in the posterior samples is uncertainty contributed by our emulator.
+We also note that although the $p$-Pb posterior samples have a somewhat larger spread than the Pb-Pb samples, the percentage uncertainty of the $p$-Pb emulator is similar to that of the Pb-Pb emulator, and thus the difference is likely due to the larger variance of the $p$-Pb training data.
+The uncertainty in the posterior distribution could thus be improved by running the calibration with more design points or with a narrower range of parameter values to increase the density of the training points and reduce interpolation uncertainty.
+
+We now direct our attention to Fig.~\ref{fig:posterior} which shows the main result of this work, the posterior distribution of the model input parameters.
+Recall that the posterior $P(H_\xv | E)$ is the probability that our hypothesis $H_\xv$ for the true model parameters $\xv_\star$ is correct, given the evidence $E$ provided by experiment.
+The present posterior has 15 dimensions, one for each parameter listed in Table~\ref{tab:design}, and thus its joint distribution cannot be summarized by one figure alone.
+We therefore sample the distribution and histogram the samples to project the distribution onto one or two dimensions at a time.
+Each diagonal panel is the distribution of a \emph{single} model parameter (marginalized over all others), and each lower-diagonal panel is the joint distribution of a \emph{pair} of model parameters (marginalized over all others).
+We also report numeric estimates for each parameter's median value and 90\% credible interval and annotate their values along the distribution diagonal (see Table~\ref{tab:post_param}).
+For example, the fictitious parameter estimate $q=2.45_{-0.15}^{+0.20}$ reports a median value $\tilde{q}=2.45$ and 90\% credible interval $2.30 < q < 2.65$.
+\newpage
+
+\begin{fullpage}
+  \begin{figure}
+    \makebox[\textwidth]{\includegraphics[width=.95\paperwidth]{posterior}}
+  \end{figure}
+  \captionsetup{skip=-15pt, width=\textwidth}
+  \captionof{figure}{\label{fig:posterior}
+    Bayesian posterior distribution of the model input parameters.
+    The diagonal panels show the marginalized distributions of individual model parameters, while off-diagonal panels show the joint distributions for pairs of model parameters, visualizing their correlations.
+    The marginalized distribution medians and 90\% credible intervals are annotated along the diagonal.
+  }
+\end{fullpage}
+
 \subsection{Initial condition properties}
-
-The \trento\ normalization factor ${\text{Norm} = 20.2^{+2.2}_{-2.2}}$ and generalized mean parameter $p=0.013^{+0.144}_{-0.155}$ are well constrained by the present analysis.
-Moreover, their posterior values nicely describe the $p$-Pb and Pb-Pb calibration observables in Figs.~\ref{fig:obs_pbpb} and \ref{fig:obs_ppb}.
-While it would not be surprising, for example, to fit one or two of these observables using such a narrow range of values, the quality of the combined fit (more on this later) and the number of observables described is highly non-trivial.
-For example, consider the ratio of the $p$-Pb yield over the Pb-Pb yield, which imposes a strong constraint on physically reasonable initial condition models.
-As the entropy deposition parameter $p$ trends toward positive (negative) infinity, particle production scales like the local maximum (minimum) of the nuclear overlap density.
-The parameter $p$ thus strongly affects the $p$-Pb and Pb-Pb yield ratio.
-It just so happens that the small range of $p$-values needed to describe this yield ratio are the same values needed to describe all of the calibration observables in the present study, \emph{and} numerous Pb-Pb observables at $\sqrts=2.76$~TeV \cite{Bernhard:2018hnz}.
-This work thus reaffirms an empirical scaling law reported by several previous studies \cite{Moreland:2014oya, Bernhard:2016tnd, Ke:2016jrd, Bernhard:2018hnz} for the initial transverse entropy density (or massless parton density):
-\begin{equation}
-  \label{eq:geometric_mean}
-  \frac{dS}{dy}\bigg\vert_{y=0} \propto \sqrt{\T_A \T_B},
-\end{equation}
-where $\T_A$ and $\T_B$ are the participant thickness functions \eqref{eq:fluctuated_thick} of each nucleus.
-We emphasize that this specific analytic form should not be interpreted too literally.
-For instance, a generalized mean described by $p=0.05$ is well within our 90\% credible interval, but it is not equal to the geometric mean in Eq.~\eqref{eq:geometric_mean}.
-This approximate form nevertheless has been shown to mimic the scaling behavior of several saturation-based theory calculations in high-energy QCD \cite{Bernhard:2016tnd}, and thus appears consistent with general theoretical expectations.
 
@@
 \end{figure}
+
+The \trento\ normalization factor ${\text{Norm} = 20.2^{+2.2}_{-2.2}}$ and generalized mean parameter $p=0.013^{+0.144}_{-0.155}$ are well constrained by the present analysis.
+Moreover, their posterior values nicely describe the $p$-Pb and Pb-Pb calibration observables in Figs.~\ref{fig:obs_pbpb} and \ref{fig:obs_ppb}.
+While it would not be surprising, for example, to fit one or two of these observables using such a narrow range of values, the quality of the combined fit (more on this later) and the number of observables described is highly non-trivial.
+For example, consider the ratio of the $p$-Pb yield over the Pb-Pb yield, which imposes a strong constraint on physically reasonable initial condition models.
+As the energy deposition parameter $p$ trends toward positive (negative) infinity, particle production scales like the local maximum (minimum) of the nuclear overlap density.
+The parameter $p$ thus strongly affects the $p$-Pb and Pb-Pb yield ratio.
+
+It just so happens that the small range of $p$-values needed to describe this yield ratio are the same values needed to describe all of the calibration observables in the present study,
+\emph{and} numerous Pb-Pb observables at $\sqrts=2.76$~TeV \cite{Bernhard:2018hnz}.
+This work thus reaffirms an empirical scaling law reported by several previous studies \cite{Moreland:2014oya, Bernhard:2016tnd, Ke:2016jrd, Bernhard:2018hnz} for the initial transverse entropy density (or massless parton density):
+\begin{equation}
+  \label{eq:geometric_mean}
+  \frac{dS}{dy}\bigg\vert_{y=0} \propto \sqrt{\T_A \T_B},
+\end{equation}
+where $\T_A$ and $\T_B$ are the participant thickness functions \eqref{eq:fluctuated_thick} of each nucleus.
+We emphasize that this specific analytic form should not be interpreted too literally.
+For instance, a generalized mean described by $p=0.05$ is well within our 90\% credible interval, but it is not equal to the geometric mean in Eq.~\eqref{eq:geometric_mean}.
+This approximate form nevertheless has been shown to mimic the scaling behavior of several saturation-based theory calculations in high-energy QCD \cite{Bernhard:2016tnd}, and thus appears consistent with general theoretical expectations.
 
@@
 
-The posterior on the constituent number $n_c$, shown enlarged in Fig.~\ref{fig:posterior_parton_number}, is not sharply peaked.
-We therefore refrain from quoting a distribution median and 90\% credible interval, although we do note that the distribution clearly favors $n_c > 1$ constituents.
-This is not surprising.
-The \trento\ model mimics saturation-based initial condition models \cite{Bernhard:2016tnd}, and saturation models tend to produce ``proton-sized'' fireballs in $p$-Pb collisions \cite{Bzdak:2013zma}.
-When the proton is spherically symmetric, the resulting proton-sized QGP is also largely symmetric and thus produces very little flow.
-Saturation-based models are therefore unable to describe the significant flow measured in high-multiplicity $p$-Pb collisions without nucleon substructure, or alternatively, some other source of additional correlations \cite{Schenke:2017bog}.
-
-The posterior on the nucleon substructure parameter $\X=0.35_{-0.17}^{+0.22}$, on the other hand, is particularly sharply peaked.
-Recall that this parameter, defined in Eq.~\eqref{eq:struct_param}, interpolates between the minimum and maximum widths of each constituent and hence
-two different limits for the granularity of the nucleon.
-When $\X=0$, the nucleon is populated by $n_c$ small compact hot-spots, and when $\X=1$ the hot-spots are large and fully overlapping, restoring spherical symmetry.
-
-The nucleon structure parameter $\X$ is somewhat awkward to conceptualize, but we can easily transform its value back into a constituent width $v$.
-Figure~\ref{fig:posterior_proton_shape} shows the resulting joint posterior distribution for the nucleon width parameter $w$ and constituent width $v$.
-The posterior distribution, shown in blue, is remarkably well constrained.
-We report a posterior estimate for the constituent width $v=0.47_{-0.15}^{+0.20}$~fm which is roughly the same magnitude as the \emph{nucleon} hot-spots used in recent saturation-based substructure studies that employed IP-Glasma initial conditions \cite{Schenke:2018fci}.
-Evidently, it may be necessary to place an informative prior on our nucleon substructure parameters in order to resolve the apparent tension between our parameter values and those needed to describe DIS measurements at HERA.
-Alternatively, it is also possible that the fluctuations probed by coherent and incoherent $J/\Psi$ production are different than those probed by minimum-bias particle production.
-
-\subsection{Transport properties}
-
 \begin{figure*}
@@
 
+The posterior on the constituent number $n_c$, shown enlarged in Fig.~\ref{fig:posterior_parton_number}, is not sharply peaked.
+We therefore refrain from quoting a distribution median and 90\% credible interval, although we do note that the distribution clearly favors $n_c > 1$ constituents.
+This is not surprising.
+The \trento\ model mimics saturation-based initial condition models \cite{Bernhard:2016tnd}, and saturation models tend to produce ``proton-sized'' fireballs in $p$-Pb collisions \cite{Bzdak:2013zma}.
+When the proton is spherically symmetric, the resulting proton-sized QGP is also largely symmetric and thus produces very little flow.
+Saturation-based models are therefore unable to describe the significant flow measured in high-multiplicity $p$-Pb collisions without nucleon substructure, or alternatively, some other source of additional correlations \cite{Schenke:2017bog}.
+
+The posterior on the nucleon substructure parameter $\X=0.35_{-0.17}^{+0.22}$, on the other hand, is particularly sharply peaked.
+Recall that this parameter, defined in Eq.~\eqref{eq:struct_param}, interpolates between the minimum and maximum widths of each constituent and hence
+two different limits for the granularity of the nucleon.
+When $\X=0$, the nucleon is populated by $n_c$ small compact hot-spots, and when $\X=1$ the hot-spots are large and fully overlapping, restoring spherical symmetry.
+
+The nucleon structure parameter $\X$ is somewhat awkward to conceptualize, but we can easily transform its value back into a constituent width $v$.
+Figure~\ref{fig:posterior_proton_shape} shows the resulting joint posterior distribution for the nucleon width parameter $w$ and constituent width $v$.
+The posterior distribution, shown in blue, is remarkably well constrained.
+We report a posterior estimate for the constituent width $v=0.47_{-0.15}^{+0.20}$~fm which is roughly the same magnitude as the \emph{nucleon} hot-spots used in recent saturation-based substructure studies that employed IP-Glasma initial conditions \cite{Schenke:2018fci}.
+Evidently, it may be necessary to place an informative prior on our nucleon substructure parameters in order to resolve the apparent tension between our parameter values and those needed to describe DIS measurements at HERA.
+Alternatively, it is also possible that the fluctuations probed by coherent and incoherent $J/\Psi$ production are different than those probed by minimum-bias particle production.
+
+\subsection{Transport properties}
+
 In this section, we compare several of our posterior estimates to those obtained from a similar Bayesian analysis in Ref.~\cite{Bernhard:2018hnz} which used an (almost) identical version of the present physics model.
@@
 \subsection{Verification of high-probability parameters}
-
-We verified the emulator and tested the accuracy of our physics model framework using a single set of high-probability parameters selected from the Bayesian posterior.
-These parameters, listed in Table~\ref{tab:mode_params}, are the approximate ``best fit'' values of the calibrated model, commonly referred to as the \emph{maximum a posteriori} (MAP) estimate:
-\begin{equation}
-  \xv_\mathrm{MAP} \equiv \operatorname*{arg\, max}_{\xv} P(H_\xv | E).
-\end{equation}
-We then ran \order{6} minimum-bias and multiplicity triggered events using the MAP estimate $\xv_\mathrm{MAP}$ and computed all of the model observables listed in Sec.~\ref{sec:observables}.
-The resulting model calculations are shown in Fig.~\ref{fig:obs_map} alongside experimental data from CMS \cite{Chatrchyan:2013nka} and ALICE \cite{Adam:2015ptt, Adam:2016izf, Adam:2014qja, Abelev:2013bla}.
-The left and right columns show the results for the $p$-Pb and Pb-Pb collision systems respectively, and each row shows a different group of related observables.
-
-The global agreement of the MAP model calculations with the experimental data is very good.
-The largest tension is observed in the two-particle cumulants $\vnk{2}{2}$ and $\vnk{3}{2}$ of the $p$-Pb system, although even that tension is only about 10--15\%.
-Quite remarkably, the model perfectly describes the shape of the $p$-Pb and Pb-Pb two-particle correlations which is strong evidence that these correlations are hydrodynamic in origin.
-Moreover, we obtain an excellent description of the \mbox{$p$-Pb} mean $p_T$, although this fit is somewhat less meaningful since we are unable to calibrate on the Pb-Pb mean $p_T$ simultaneously (data is not yet available).
-Additionally, the model provides a simultaneous description of the \mbox{$p$-Pb} and Pb-Pb charged-particle yields using a single entropy deposition parameter $p=0$.
-This is the \emph{exact same} generalized mean $p$-value supported by multiple previous studies \cite{Moreland:2014oya, Bernhard:2016tnd, Ke:2016jrd, Bernhard:2018hnz}.
-Evidently, this scaling continues to hold for initial conditions with sizable nucleon substructure.
-
-We also present calculations for several observables which were omitted from the calibration due to the statistical limitations of our training data.
-Here our MAP event sample is several orders-of-magnitude larger so the statistics are no issue.
-The bottom-right panel of Fig.~\ref{fig:obs_map} shows our model calculation for the four-particle elliptic flow cumultant $\vnk{2}{4}$ along with the measured data points from \mbox{ALICE} \cite{Adam:2016izf}.
-We see that the MAP estimate nicely describes the measured $\vnk{2}{4}$ data which is encouraging since this particular observable was never used to calibrate the model.
-
-The relative mean $p_T$ fluctuation $\delta p_T / \langle p_T \rangle$ is another important bulk observable to test the predictions of the calibrated model.
-It measures the \emph{dynamical} component of event-by-event mean $p_T$ fluctuations, quantified by the two-particle correlator
-\begin{equation}
-  \label{eq:mean_pT_corr}
-  (\delta p_T)^2 = \langle \langle (p_{T,i} - \langle p_T \rangle) (p_{T,j} - \langle p_T \rangle) \rangle \rangle.
-\end{equation}
-
-\begin{fullpage}
-  {\centering\includegraphics{observables_map}\par}
-  \captionof{figure}{
-    \label{fig:obs_map}
-    Model calculations using the \emph{maximum a posterior} (MAP) parameters compared to experiment.
-    Colored lines are model calculations for $p$-Pb collisions (left) and Pb-Pb collisions (right) at $\sqrts=5.02$~TeV.
-    Points with error bars are the experimental data with statistical uncertainties, and gray bands their corresponding systematic uncertainties, from CMS \cite{Chatrchyan:2013nka} and ALICE \cite{Adam:2015ptt, Adam:2016izf, Adam:2014qja, Abelev:2013bla}.
-    The sub-axes show the ratio of model over data where available with gray bands indicating $\pm 10\%$.
-  }
-\end{fullpage}
-
-\noindent The inner-average in Eq.~\eqref{eq:mean_pT_corr} runs over all pairs of particles $i,j$ in the same event, the outer average runs over all events in a given bin (centrality or multiplicity), and the symbol $\langle p_T \rangle$ denotes the usual mean transverse momentum of particles in the bin.
-The observable is typically presented in terms of the dimensionless ratio $\delta p_T / \langle p_T \rangle$, which quantifies the strength of the dynamical fluctuations in units of the average transverse momentum $\langle p_T \rangle$.
-
-We show the MAP estimate predictions for the $p$-Pb and Pb-Pb relative mean $p_T$ fluctuations $\delta p_T / \langle p_T \rangle$ at $\sqrts=5.02$~TeV in the third row of Fig.~\ref{fig:obs_map}.
-For the Pb-Pb system, we use centrality bins and for the $p$-Pb system we use the same relative multiplicity bins used for the $p$-Pb charged-particle mean $p_T$.
-The relative mean $p_T$ fluctuations have been shown to be particularly sensitive to the existence of nucleon substructure \cite{Bozek:2017elk}, and thus it would be interesting to ultimately include this observable in the calibration when the data becomes available.
 
@@
 
+We verified the emulator and tested the accuracy of our physics model framework using a single set of high-probability parameters selected from the Bayesian posterior.
+These parameters, listed in Table~\ref{tab:mode_params}, are the approximate ``best fit'' values of the calibrated model, commonly referred to as the \emph{maximum a posteriori} (MAP) estimate:
+\begin{equation}
+  \xv_\mathrm{MAP} \equiv \operatorname*{arg\, max}_{\xv} P(H_\xv | E).
+\end{equation}
+We then ran \order{6} minimum-bias and multiplicity triggered events using the MAP estimate $\xv_\mathrm{MAP}$ and computed all of the model observables listed in Sec.~\ref{sec:observables}.
+The resulting model calculations are shown in Fig.~\ref{fig:obs_map} alongside experimental data from CMS \cite{Chatrchyan:2013nka} and ALICE \cite{Adam:2015ptt, Adam:2016izf, Adam:2014qja, Abelev:2013bla}.
+The left and right columns show the results for the $p$-Pb and Pb-Pb collision systems respectively, and each row shows a different group of related observables.
+
+The global agreement of the MAP model calculations with the experimental data is very good.
+The largest tension is observed in the two-particle cumulants $\vnk{2}{2}$ and $\vnk{3}{2}$ of the $p$-Pb system, although even that tension is only about 10--15\%.
+Quite remarkably, the model perfectly describes the shape of the $p$-Pb and Pb-Pb two-particle correlations which is strong evidence that these correlations are hydrodynamic in origin.
+Moreover, we obtain an excellent description of the \mbox{$p$-Pb} mean $p_T$, although this fit is somewhat less meaningful since we are unable to calibrate on the Pb-Pb mean $p_T$ simultaneously (data is not yet available).
+Additionally, the model provides a simultaneous description of the \mbox{$p$-Pb} and Pb-Pb charged-particle yields using a single entropy deposition parameter $p=0$.
+This is the \emph{exact same} generalized mean $p$-value supported by multiple previous studies \cite{Moreland:2014oya, Bernhard:2016tnd, Ke:2016jrd, Bernhard:2018hnz}.
+Evidently, this scaling continues to hold for initial conditions with sizable nucleon substructure.
+
+We also present calculations for several observables which were omitted from the calibration due to the statistical limitations of our training data.
+Here our MAP event sample is several orders-of-magnitude larger so the statistics are no issue.
+The bottom-right panel of Fig.~\ref{fig:obs_map} shows our model calculation for the four-particle elliptic flow cumultant $\vnk{2}{4}$ along with the measured data points from \mbox{ALICE} \cite{Adam:2016izf}.
+We see that the MAP estimate nicely describes the measured $\vnk{2}{4}$ data which is encouraging since this particular observable was never used to calibrate the model.
+
+The relative mean $p_T$ fluctuation $\delta p_T / \langle p_T \rangle$ is another important bulk observable to test the predictions of the calibrated model.
+It measures the \emph{dynamical} component of event-by-event mean $p_T$ fluctuations, quantified by the two-particle correlator
+\begin{equation}
+  \label{eq:mean_pT_corr}
+  (\delta p_T)^2 = \langle \langle (p_{T,i} - \langle p_T \rangle) (p_{T,j} - \langle p_T \rangle) \rangle \rangle.
+\end{equation}
+
+\noindent The inner-average in Eq.~\eqref{eq:mean_pT_corr} runs over all pairs of particles $i,j$ in the same event, the outer average runs over all events in a given bin (centrality or multiplicity), and the symbol $\langle p_T \rangle$ denotes the usual mean transverse momentum of particles in the bin.
+The observable is typically presented in terms of the dimensionless ratio $\delta p_T / \langle p_T \rangle$, which quantifies the strength of the dynamical fluctuations in units of the average transverse momentum $\langle p_T \rangle$.
+
+We show the MAP estimate predictions for the $p$-Pb and Pb-Pb relative mean $p_T$ fluctuations $\delta p_T / \langle p_T \rangle$ at $\sqrts=5.02$~TeV in the third row of Fig.~\ref{fig:obs_map}.
+For the Pb-Pb system, we use centrality bins and for the $p$-Pb system we use the same relative multiplicity bins used for the $p$-Pb charged-particle mean $p_T$.
+The relative mean $p_T$ fluctuations have been shown to be particularly sensitive to the existence of nucleon substructure \cite{Bozek:2017elk}, and thus it would be interesting to ultimately include this observable in the calibration when the data becomes available.
+
 Lastly, we compute the symmetric cumulants $\SC(m,n)$ for the Pb-Pb collision system at $\sqrts=5.02$~TeV which quantify correlations between event-by-event fluctuations of the flow harmonics of different order \cite{Bilandzic:2013kga, ALICE:2016kpq},
@@
 \end{equation}
+
+\begin{fullpage}
+  {\centering\includegraphics{observables_map}\par}
+  \captionof{figure}{
+    \label{fig:obs_map}
+    Model calculations using the \emph{maximum a posterior} (MAP) parameters compared to experiment.
+    Colored lines are model calculations for $p$-Pb collisions (left) and Pb-Pb collisions (right) at $\sqrts=5.02$~TeV.
+    Points with error bars are the experimental data with statistical uncertainties, and gray bands their corresponding systematic uncertainties, from CMS \cite{Chatrchyan:2013nka} and ALICE \cite{Adam:2015ptt, Adam:2016izf, Adam:2014qja, Abelev:2013bla}.
+    The sub-axes show the ratio of model over data where available with gray bands indicating $\pm 10\%$.
+  }
+\end{fullpage}
 which are expected to be less sensitive to the medium response and more sensitive to the properties of the initial state.

### 2018-10-16 src/plots.py

@@ -2509,7 +2509,7 @@ def posterior_proton_shape():
         'sampling_radius', 'parton_number', 'parton_width'
     ).T
 
-    fig = plt.figure(figsize=figsize(.5, aspect=1))
+    fig = plt.figure(figsize=figsize(.5, aspect=.833))
     ax = plt.gca()
 
     plt.hist2d(sampling_radius, parton_width, bins=100, cmap=plt.cm.Blues)

### 2018-10-19 paper/paper.tex
Finish main text revisions. All that remains, update parameter values.

@@
 \newcommand{\sigmaf}{\sigma_\text{fluct}}
-\newcommand{\X}{\chi_\text{struct}}
 \newcommand{\taufs}{\tau_\text{fs}}
 \newcommand{\dmin}{d_\text{min}}
+\newcommand{\dTb}{\Delta T_\text{bulk}}
+\newcommand{\Tb}{T_\text{bulk}}
 \newcommand{\Tsw}{T_\text{switch}}
@@
 Ultrarelativistic nuclear collisions between one light-ion and one heavy-ion, e.g.\ $^3$He-Au and $p$-Pb collisions, generate dense, compact sources of nuclear matter which produce long-range multiparticle correlations that are strikingly similar to the correlations observed in heavy-ion collisions where collectivity is commonly explained by the existence of hydrodynamic flow \cite{CMS:2012qk, Abelev:2012ola, Aad:2012gla, Adare:2015ctn}.
-This observation suggests that hydrodynamic behavior could be manifest in small droplets of quark-gluon-plasma (QGP) \cite{Bozek:2011if, Bozek:2013uha}, and that flow might even develop at length scales smaller than a single proton \cite{Schenke:2014zha}.
+This observation suggests that hydrodynamic behavior could be manifest in small droplets of quark-gluon-plasma (QGP) \cite{Bozek:2011if, Bozek:2013uha}, and that flow even might develop at length scales smaller than a single proton \cite{Schenke:2014zha}.
 
@@
   \label{eq:tpp}
-  T_{pp}(b) = \int dx\, dy\, dz\, \rho_A(x, y, z)\, \rho_B(x - b, y, z),
+  T_{pp}(b) = \int dx\, dy \int dz_1\, \rho_A(x, y, z_1) \int dz_2\, \rho_B(x - b, y, z_2),
 \end{equation}
@@
 \end{equation}
-of width $v$.
+of variable width $v$.
 The constituent positions $\xv_i$ are sampled independently (ignoring correlations) from a Gaussian radial distribution,
@@
 \end{equation}
-onto the transverse plane $\xv_\perp$ centered at $z=0$.
+onto the transverse plane $\xv_\perp$ centered at midrapidity.
 This fluctuated proton density $\tilde{\rho}_{A,B}$ equals the proton density $\rho_{A,B}(\xv)$ in Eq.~\eqref{eq:rho} with additional weights $\gamma_i$ sampled from a gamma distribution with unit mean and variance $1/k$.
@@
 This wounded nucleon scaling was in fact one of the first such mappings used as a proxy for initial particle production in relativistic heavy-ion collisions \cite{Bialas:1976ed}.
-It was quickly realized, however, that the wounded nucleon model predicts the wrong scaling for charged particle production as a function of collision centrality \cite{Kharzeev:2000ph} and hence the wrong scaling for initial energy deposition at midrapidity as a function of participant thicknesses $\T_A$ and $\T_B$\footnote{Typically an entropy density $s$ is used, but the shortcomings of the wounded nucleon model are qualitatively the same when depositing energy.}.
+It was subsequently realized, however, that the wounded nucleon model predicts the wrong scaling for charged particle production as a function of collision centrality \cite{Kharzeev:2000ph} and hence the wrong scaling for initial energy deposition at midrapidity as a function of participant thicknesses $\T_A$ and $\T_B$\footnote{Typically an entropy density $s$ is used, but the shortcomings of the wounded nucleon model are qualitatively the same when depositing energy.}.
 
@@
 This assumes, of course, that the model's parametrization for entropy density is also suitable as a parametrization for the energy density.
-To this point, we note that the arguments for the generalized mean ansatz are quite general, and that the energy density $dE/d\eta_s$ and entropy density $dS/d\eta_s$ share many similar qualities.
+To this point, we note that the arguments for the generalized mean ansatz are quite general \cite{Moreland:2015oya}, and that the energy density $dE/d\eta_s$ and entropy density $dS/d\eta_s$ share many similar qualities.
 For example, both are approximately conserved at midrapidity, and both scale approximately linearly with the final charged-particle density $d\nch/d\eta_s$.
@@
 
-There are of course two limiting cases for the strength of interactions inside the QGP medium immediately at the collision: infinitely weak coupling where the secondary partons free-stream without interacting, and infinitely strong coupling where the fluid's inter-particle mean free path effectively vanishes.
+There are of course two limiting cases for the strength of interactions inside the QGP medium immediately after the collision: infinitely weak coupling where the secondary partons free-stream without interacting, and infinitely strong coupling where the fluid's inter-particle mean free path effectively vanishes.
 Realistically, one expects the initial parton interactions to lie somewhere between these two extremes.
@@
   \label{eq:bulk_viscosity}
-  (\zeta/s)(T) = \frac{(\zeta/s)_\mathrm{max}}{1 + \left(\dfrac{T - (\zeta/s)_{T_0}}{(\zeta/s)_\mathrm{width}}\right)^2},
-\end{equation}
-described by tunable maximum, width, and location ($T_0$) parameters.
+  (\zeta/s)(T) = \frac{(\zeta/s)_\text{max}}{1 + \left(\dfrac{T - T_\text{bulk}}{\Delta T_\text{bulk}}\right)^2},
+\end{equation}
+described by a tunable maximum $(\zeta/s)_\text{max}$, width $\dTb$, and location parameter $\Tb$.
 Figure~\ref{fig:viscosity_dof} shows several of the possible curves parametrized by Eqs.~\eqref{eq:shear_viscosity} and \eqref{eq:bulk_viscosity}, although many more are possible.
@@
 
-Once the fluid is converted into particles, we simulate its subsequent microscopic dynamics using the Ultra-relativistic Quantum modifies Dynamics (UrQMD) transport model \cite{Bass:1998ca, Bleicher:1999xi}.
+Once the fluid is converted into particles, we simulate its subsequent microscopic dynamics using the Ultra-relativistic Quantum Molecular Dynamics (UrQMD) transport model \cite{Bass:1998ca, Bleicher:1999xi}.
 It solves the microscopic Boltzmann equation
@@
 The nuclear collision model constructed in Sec.~\ref{sec:model} includes a number of free parameters $\xv$ which describe the initial state, pre-equilibrium dynamics, and hydrodynamic medium.
-Given values for the parameters $\xv$, the model may be used to predict a vector of simulated observables $\yv_m$.
-For example, $\yv_m$ might be a vector consisting of charged particle yields in different centrality bins.
-The physics model thus represents a vector-valued function $f(\xv) = \yv_m$ which maps the parameter values $\xv$ to the model observables $\yv_m$.
-
-The goal of this work is to estimate the true model parameters $\xv_\star$ provided some evidence that our model predictions $\yv_m$ describe experimental measurements $\yv_e$.
+Given values for the parameters $\xv$, the model may be used to predict a vector of simulated observables $\yv$.
+For example, $\yv$ might be a vector consisting of charged particle yields in different centrality bins.
+The physics model thus represents a vector-valued function $f(\xv) = \yv$ which maps the parameter values $\xv$ to the model observables $\yv$.
+
+The goal of this work is to estimate the true model parameters $\xv_\star$ provided some evidence that our model predictions $\yv$ describe experimental measurements $\yv_e$.
 The problem involves three distinct components:
@@
 \end{equation}
-where $\Delta\yv = \yv_m(\xv) - \yv_e$ is the discrepancy of the model and experiment, and $\Sigma = \Sigma_m(\xv) + \Sigma_e$ is the \emph{total} covariance matrix, equal to the sum of a modeling component $\Sigma_m(\xv)$ and an experimental component $\Sigma_e$ which account for all known sources of uncertainty in the simulated and measured observables.
+where $\Delta\yv = \yv(\xv) - \yv_e$ is the discrepancy of the model and experiment, and $\Sigma = \Sigma_m(\xv) + \Sigma_e$ is the \emph{total} covariance matrix, equal to the sum of a modeling component $\Sigma_m(\xv)$ and an experimental component $\Sigma_e$ which account for all known sources of uncertainty in the simulated and measured observables.
 
@@
     $\zeta/s$ max     & Maximum value of $\zeta/s$           & 0--0.1          \\
-    $\zeta/s$ width   & Width of $\zeta/s$ peak              & 0--0.1 GeV      \\
-    $\zeta/s$ $T_0$   & Temperature of $\zeta/s$ maximum     & 150--200 MeV    \\
+    $\dTb$            & Width of $\zeta/s$ peak              & 0--0.1 GeV      \\
+    $\Tb$             & Temperature of $\zeta/s$ maximum     & 150--200 MeV    \\
     $\Tsw$            & Switching/particlization temp.       & 135--165 MeV    \\
@@
 
-The likelihood function \eqref{eq:likelihood} provides evidence for \mbox{(or against)} the model parameters $\xv$ by comparing the model predictions $\yv_m(\xv)$ to experimental data $\yv_e$.
+The likelihood function \eqref{eq:likelihood} provides evidence for \mbox{(or against)} the model parameters $\xv$ by comparing the model predictions $\yv(\xv)$ to experimental data $\yv_e$.
 We focus on simple experimental observables in the present study which are sensitive to the bulk properties of the produced medium.
@@
 These high-multiplicity bins are too selective for our modest minimum bias event sample, and so a different procedure is required.
-We exploit for this purpose the approximate monotonic dependence of the midrapidity charged-particle yield on the transverse energy density,
-\begin{equation}
-  E_T = \int d^2x_\perp e(\xv_\perp, \eta_s=0, \tau_0).
-\end{equation}
-
-Consider, for example, an single multiplicity bin $[\nch^\text{low}, \nch^\text{high}]$ subselected from a minimum bias event sample with some kinematic cuts, $p_T^\text{min} < p_T < p_T^\text{max}$ and $|\eta| < \eta^\text{max}$.
-Let $\langle\nch\rangle$ denote the average charged-particle multiplicity of the minimum bias event sample.
+We exploit, for this purpose, the approximate monotonic relation between each event's initial transverse energy density,
+\begin{equation}
+  \frac{dE_T}{d\eta_s} \bigg\vert_{\eta_s=0} = \tau_0 \int d^2x_\perp e(\xv_\perp, \eta_s=0, \tau_0).
+\end{equation}
+and its final charged-particle density $(d\nch/d\eta) \vert_{\eta=0}$ at midrapidity.
+
+Consider, for example, a single multiplicity bin $[\nch^\text{low}, \nch^\text{high}]$ which selects events from a minimum bias event sample with $p_T^\text{min} < p_T < p_T^\text{max}$ and $|\eta| < \Delta \eta$.
+Let $\langle\nch\rangle$ denote the average charged-particle multiplicity of these events.
 We first rescale the experimental multiplicity bin edges,
@@
 in order to reexpress each bin edge as a unitless variable.
-These rescaled bins are then associated with the analogous bins edges defined using the \trento\ model's initial transverse energy density,
+These bin edges are then associated with a pair of transverse energy bin edges,
 \begin{equation}
@@
 \end{equation}
-where $\langle E_T \rangle$ denotes the average transverse energy density of the minimum bias event sample.
-
-Finally, we mimic the method used by experiment and apply \eqref{eq:mult_trigger} to select rare high-multiplicity events from a continuous stream of minimum-bias \trento\ events satisfying the correct bin edges.
+where $E_T \propto (dE_T/d\eta_s) \vert_{\eta_s=0}$ is the midrapidity transverse energy of a single event, and $\langle E_T \rangle$ is the corresponding average over the minimum bias event sample.
+
+Finally, we mimic the method used by experiment and apply Eq.~\eqref{eq:mult_trigger} to select rare high-multiplicity events from a continuous stream of minimum-bias \trento\ events satisfying the correct bin edges.
 This of course means that, in addition to running a large sample of minimum-bias events for centrality binned observables, we must also generate (much like experiment) a separate sample of multiplicity triggered events.
@@
 
-In principle, one could calculate the likelihood function in Eq.~\eqref{eq:likelihood} directly, e.g.\ by running the model a large number of times at a given parameter point $\xv$ to calculate the model observables $\yv_m(\xv)$ from the ensemble of simulated events, but in practice such a procedure would be intractable.
+In principle, one could calculate the likelihood function in Eq.~\eqref{eq:likelihood} directly, e.g.\ by running the model a large number of times at a given parameter point $\xv$ to calculate the model observables $\yv(\xv)$ from the ensemble of simulated events, but in practice such a procedure would be intractable.
 The model is computationally intensive to evaluate, and thousands of events are required to calculate the simplest observables at a single parameter point.
@@
 We therefore follow an established framework for computationally intensive models and train an emulator to act as a fast surrogate for the full physics simulation \cite{OHagan:2006ba, Higdon:2008cmc, Higdon:2014tva}.
-The emulator enables essentially instantaneous predictions for $\yv_m = f(\xv)$ and allows us to sample the posterior distribution millions of times.
-In order to train the emulator, we first generate a scaffolding of the parameter space using a maximin Latin hypercube design \cite{Morris:1995lh} to distribute 500 points semi-randomly throughout our 15-dimensional parameter space according to the parameter ranges in Table~\ref{tab:design}.
+The emulator enables essentially instantaneous predictions for $\yv = f(\xv)$ and allows us to sample the posterior distribution millions of times.
+In order to train the emulator, we first generate a scaffolding of the parameter space using a maximin Latin hypercube design \cite{Morris:1995lh} to distribute 500 points throughout our 15-dimensional parameter space according to the parameter ranges in Table~\ref{tab:design}.
 We then run minimum-bias and multiplicity triggered $p$-Pb and Pb-Pb events at each design point and calculate the model observables from the ensemble of events.
@@
 Specifically, let $X$ denote the ${d \times n}$ design matrix of $d=500$ training points, where each training point is a vector $\xv = (x_1, x_2, \dots, x_n)$ of the $n=15$ model parameters in Table~\ref{tab:design}.
-Similarly, let $Y$ denote the corresponding ${d \times m}$ observables matrix, where each row of $Y$ is a vector $\yv_m = (y_1, y_2, \dots, y_m)$ of $m$ different simulated observables (here we overload our notation so the subscript $m$ means both \emph{model} and \emph{observable number}).
+Similarly, let $Y$ denote the corresponding ${d \times m}$ observables matrix, where each row of $Y$ is a vector $\yv = (y_1, y_2, \dots, y_m)$ of $m$ observables.
 Our goal is to train an emulator for the physics model $f$ using the discrete observations $f: X \mapsto Y$.
@@
 
-We first preprocess our model observables by centering and scaling each column of $Y$ (single observable) to zero mean and unit variance to generate a standardized observable matrix $\tilde{Y}$.
-PCA is then used to reexpress each row-vector $\tilde{\yv}$ of $\tilde{Y}$ (all standardized observables at a single design point) in the new principal component basis:
-\begin{equation}
-  \tilde{\yv}_i = \sum\limits_{j=1}^m z_{ij} \mathbf{v_j},
-\end{equation}
-where $\tilde{\yv}_i$ are the standardized observables of the $i$-th row-vector (design point) of matrix $\tilde{Y}$, and $z_{ij}$ and $\mathbf{v}_j$ are the coefficients and vectors of its $j$-th principal component.
-
-The principal components are reported in order of explained variance, with the first principal component vector $\mathbf{v}_1$ accounting for the most variance in $\tilde{Y}$, and the last principal component vector $\mathbf{v}_m$ accounting for the least.
+We first preprocess our model observables by centering and scaling each column of $Y$ (single observable) to zero mean and unit variance to generate a standardized observable matrix $\hat{Y}$.
+PCA is then used to reexpress each row-vector $\hat{\yv}$ of $\hat{Y}$ (all standardized observables at a single design point) in the new principal component basis:
+\begin{equation}
+  \hat{\yv}_i = \sum\limits_{j=1}^m z_{ij} \mathbf{v_j},
+\end{equation}
+where $\hat{\yv}_i$ are the standardized observables of the $i$-th row-vector (design point) of matrix $\hat{Y}$, and $z_{ij}$ and $\mathbf{v}_j$ are the coefficients and vectors of its $j$-th principal component.
+
+The principal components are reported in order of explained variance, with the first principal component vector $\mathbf{v}_1$ accounting for the most variance in $\hat{Y}$, and the last principal component vector $\mathbf{v}_m$ accounting for the least.
 We then train a set of independent GPE's $\{z_i=\mathrm{gp}_i(\xv)\}$ to predict the first $k$ principal components $(z_1, \dots, z_k)$ as a function of the model parameters $\xv$ which vary across the design $X$.
@@
 Each GPE reports a mean value $z(\xv)$ as well as an estimated error $\delta z(\xv)$ which accounts for statistical noise in the training data and interpolation error between the design points.
-Once the GPE's are trained, we can predict the observables $\yv_m$ at parameter point $\xv$ by transforming the vector of principal components
+Once the GPE's are trained, we can predict the observables $\yv$ at parameter point $\xv$ by transforming the vector of principal components
 \begin{equation}
@@
 \end{equation}
-and transform it back to physical space as well to obtain the covariance matrix of the model observables $\yv_m$ at a given parameter point $\xv$.
-
-The resulting emulator therefore predicts both a mean prediction $\yv_m(\xv)$ and an uncertainty covariance matrix $\Sigma_m(\xv)$ which accounts for multiple sources of model and emulator uncertainty, including the truncation error expected from using a finite number of principal components $k < m$.
+and transform it back to physical space as well to obtain the covariance matrix of the model observables $\yv$ at a given parameter point $\xv$.
+
+The resulting emulator therefore predicts both a mean prediction $\yv(\xv)$ and an uncertainty covariance matrix $\Sigma_m(\xv)$ which accounts for multiple sources of model and emulator uncertainty, including the truncation error expected from using a finite number of principal components $k < m$.
 The model covariance matrix $\Sigma_m$ includes so-called \emph{known}-unknowns such as statistical error and emulator interpolation error, but not \emph{unknown}-unknowns such as the overall validity of small-system hydrodynamics, i.e.\ things which lack a unified consensus or are difficult to quantify.
@@
       $r$        & \est{0.88}{0.23}{0.25} fm     & $\zeta/s$ max    & \est{0.026}{0.026}{0.033}         \\[1.1ex]
-      $n_c$      & \est{6.3}{3.1}{2.7}           & $\zeta/s$ width  & \est{0.033}{0.033}{0.048} GeV     \\[1.1ex]
-      $v$       & \est{0.51}{0.18}{0.26} fm      & $\zeta/s$ $T_0$  & \est{0.176}{0.020}{0.024} GeV     \\[1.1ex]
+      $n_c$      & \est{6.3}{3.1}{2.7}           & $\dTb$           & \est{0.033}{0.033}{0.048} GeV     \\[1.1ex]
+      $v$       & \est{0.51}{0.18}{0.26} fm      & $\Tb$            & \est{0.176}{0.020}{0.024} GeV     \\[1.1ex]
       $\dmin$    & \est{1.13}{0.51}{0.55} fm     & $\Tsw$           & \est{0.149}{0.013}{0.014} GeV     \\[1.1ex]
@@
 \begin{figure}
+  \includegraphics{posterior_proton_shape}
+  \caption{
+    \label{fig:posterior_proton_shape}
+    Posterior distribution for the constituent sampling radius $r$ and constituent width $v$.
+    The prior range for $r$ and $v$ spans the full plot range.
+    The posterior distribution, shown in blue, indicates the preferred values for $r$ and $v$ determined by the analysis.
+  }
+\end{figure}
+
+The \trento\ normalization factor ${\text{Norm} = 20.2^{+2.2}_{-2.2}}$ and energy deposition parameter $p=0.013^{+0.144}_{-0.155}$ are well constrained by the present analysis.
+Moreover, Figs.~\ref{fig:obs_pbpb} and \ref{fig:obs_ppb} show that the model predictions using these values nicely describe both the $p$-Pb and Pb-Pb calibration observables.
+While it would not be surprising to fit one or two of these observables using a narrow range of parameter values, the quality of the combined fit (more on this later) and the number of observables described is highly non-trivial.
+For example, consider the ratio of the $p$-Pb charged-particle yield to the Pb-Pb charged-particle yield.
+As the energy deposition parameter $p$ trends toward positive (negative) infinity, particle production scales like the maximum (minimum) of local nuclear overlap density.
+This has a much stronger effect on the highly asymmetric $p$-Pb system than it does on the Pb-Pb system; hence the parameter $p$ strongly affects the ratio of the two average yields.
+
+It is therefore compelling that $p \sim 0$ correctly describes the charged-particle yield $d\nch/d\eta$ of both systems, while simultaneously describing the centrality dependence of $\vnk{n}{k}$, an observable which is also known to strongly depend on $p$ \cite{Bernhard:2018hnz}.
+Specifically, this value $p \sim 0$ corresponds to an energy deposition mapping proportional to the geometric mean of participant nuclear thickness,
+\begin{equation}
+  \label{eq:geometric_mean}
+  e(\xv_\perp, \eta_s=0, \tau_0) \propto \sqrt{\T_A\, \T_B}.
+\end{equation}
+We caution, however, that this specific analytic form should not be interpreted too literally.
+For instance, a generalized mean described by $p=0.05$ is well within our 90\% credible interval, but it does not equal the geometric mean in Eq.~\eqref{eq:geometric_mean}.
+We also note that this scaling is somewhat different than the scaling obtained from previous analyses of the \trento\ model, which parametrized the \emph{entropy} density using a framework which assumed instant thermalization and zero pre-equilibrium flow.
+Evidently, both prescriptions prefer geometric mean scaling, but each prescription leads to a somewhat different interpretation of the initially produced quantity.
+
+\begin{figure}
   \includegraphics{posterior_parton_number}
@@
 
-The \trento\ normalization factor ${\text{Norm} = 20.2^{+2.2}_{-2.2}}$ and generalized mean parameter $p=0.013^{+0.144}_{-0.155}$ are well constrained by the present analysis.
-Moreover, their posterior values nicely describe the $p$-Pb and Pb-Pb calibration observables in Figs.~\ref{fig:obs_pbpb} and \ref{fig:obs_ppb}.
-While it would not be surprising, for example, to fit one or two of these observables using such a narrow range of values, the quality of the combined fit (more on this later) and the number of observables described is highly non-trivial.
-For example, consider the ratio of the $p$-Pb yield over the Pb-Pb yield, which imposes a strong constraint on physically reasonable initial condition models.
-As the energy deposition parameter $p$ trends toward positive (negative) infinity, particle production scales like the local maximum (minimum) of the nuclear overlap density.
-The parameter $p$ thus strongly affects the $p$-Pb and Pb-Pb yield ratio.
-
-It just so happens that the small range of $p$-values needed to describe this yield ratio are the same values needed to describe all of the calibration observables in the present study,
-\emph{and} numerous Pb-Pb observables at $\sqrts=2.76$~TeV \cite{Bernhard:2018hnz}.
-This work thus reaffirms an empirical scaling law reported by several previous studies \cite{Moreland:2014oya, Bernhard:2016tnd, Ke:2016jrd, Bernhard:2018hnz} for the initial transverse entropy density (or massless parton density):
-\begin{equation}
-  \label{eq:geometric_mean}
-  \frac{dS}{dy}\bigg\vert_{y=0} \propto \sqrt{\T_A \T_B},
-\end{equation}
-where $\T_A$ and $\T_B$ are the participant thickness functions \eqref{eq:fluctuated_thick} of each nucleus.
-We emphasize that this specific analytic form should not be interpreted too literally.
-For instance, a generalized mean described by $p=0.05$ is well within our 90\% credible interval, but it is not equal to the geometric mean in Eq.~\eqref{eq:geometric_mean}.
-This approximate form nevertheless has been shown to mimic the scaling behavior of several saturation-based theory calculations in high-energy QCD \cite{Bernhard:2016tnd}, and thus appears consistent with general theoretical expectations.
-
-Continuing down the diagonal in Fig.~\ref{fig:posterior}, we see that the nucleon width parameter $w=0.98_{-0.17}^{+0.18}$~fm.
-We caution that this parameter $w$ is \emph{not} the RMS radius of our nucleons due to idiosyncrasies of our constituent sampling procedure.
-We can, however, easily calculate the RMS radius for a specific nucleon width $w$, constituent width $v$, and constituent number $n_c$.
-For example, the single highest posterior probability region of the parameter space prefers a nucleon width $w=0.92$~fm, constituent width $v=0.43$~fm, and constituent number $n_c=6$.
-The corresponding RMS nucleon radius for these parameters is $R_n = 0.86$~fm, conspicuously close the proton's RMS electric charge radius $R_p = 0.879(8)$~fm \cite{Bernauer:2010wm}.
-This is perhaps the single largest difference between our work and the conclusions of recent saturation-based calculations which constrained the event-by-event fluctuations of the proton using a color-dipole picture of vector meson production \cite{Mantysaari:2016ykx, Mantysaari:2016jaz}.
-Those studies find that the measured coherent and incoherent $J/\Psi$ spectra at HERA prefer a compact gluon distribution inside each nucleon, with an RMS radius $R_g \approx 0.4$~fm, which is roughly \emph{half} the analogous RMS nucleon width preferred by our analysis.
-
-\begin{figure}
-  \includegraphics{posterior_proton_shape}
-  \caption{
-    \label{fig:posterior_proton_shape}
-    Posterior distribution for the nucleon width $w$ and constituent width $v$.
-    The enclosed trapezoidal region covers the prior range of allowed values for $w$ and $v$.
-    The posterior distribution, shown in blue, indicates the preferred values for $w$ and $v$ determined by the analysis.
-  }
-\end{figure}
+Continuing down the diagonal in Fig.~\ref{fig:posterior}, we see that the constituent sampling radius $r=0.88_{-0.23}^{+0.25}$~fm, and the constituent width $v=0.51_{-0.18}^{+0.26}$~fm.
+Their joint posterior distribution is shown in Fig.~\ref{fig:posterior_proton_shape}, illustrating the constraining power of the analysis.
+Recall that the parameter $r$ controls the dispersion of the constituent positions, while $v$ varies each constituent's Gaussian width.
+We emphasize that $r$ is model parameter, not a physical one.
+It describes the radius of the constituent position sampling distribution, \emph{not} the average proton radius, a quantity that's measured in the proton's center-of-mass frame.
+For example, when $n_c=1$, our model describes each nucleon as a single Gaussian of width $v$, yet the parameter $r$ could nevertheless assume any value within its prior.
+Despite this idiosyncrasy, one can easily calculate a physical nucleon width \emph{ex post facto}, given specific values for $r$, $v$, and $n_c$.
 
@@
 
+For example, using the MAP estimate $r=0.88$~fm, $n_c=6$, and $v=0.51$~fm, we obtain a Gaussian nucleon width $w = 0.88$~fm, corresponding to a three-dimensional root-mean-square radius $R_p = \sqrt{3} w = 1.52$~fm.
+This estimate is consistent with the nucleon width extracted by a similar Bayesian analysis lacking nucleon substructure \note.
+This is perhaps the single largest difference between our work and the conclusions of recent saturation-based calculations which constrained the event-by-event fluctuations of the proton using a color-dipole picture of vector meson production \cite{Mantysaari:2016ykx, Mantysaari:2016jaz}.
+Those studies find that the measured coherent and incoherent $J/\Psi$ spectra at HERA prefer a compact gluon distribution inside each nucleon, with a Gaussian width $w_g \approx 0.4$~fm which is roughly \emph{half} the Gaussian width preferred by our analysis.
+Evidently, it may be necessary to place an informative prior on our nucleon substructure parameters in order to resolve the apparent tension between our parameter values and those needed to describe DIS measurements at HERA.
+Alternatively, it is also possible that the fluctuations probed by coherent and incoherent $J/\Psi$ production are different than those probed by minimum-bias particle production.
+
 The posterior on the constituent number $n_c$, shown enlarged in Fig.~\ref{fig:posterior_parton_number}, is not sharply peaked.
@@
 
-The posterior on the nucleon substructure parameter $\X=0.35_{-0.17}^{+0.22}$, on the other hand, is particularly sharply peaked.
-Recall that this parameter, defined in Eq.~\eqref{eq:struct_param}, interpolates between the minimum and maximum widths of each constituent and hence
-two different limits for the granularity of the nucleon.
-When $\X=0$, the nucleon is populated by $n_c$ small compact hot-spots, and when $\X=1$ the hot-spots are large and fully overlapping, restoring spherical symmetry.
-
-The nucleon structure parameter $\X$ is somewhat awkward to conceptualize, but we can easily transform its value back into a constituent width $v$.
-Figure~\ref{fig:posterior_proton_shape} shows the resulting joint posterior distribution for the nucleon width parameter $w$ and constituent width $v$.
-The posterior distribution, shown in blue, is remarkably well constrained.
-We report a posterior estimate for the constituent width $v=0.47_{-0.15}^{+0.20}$~fm which is roughly the same magnitude as the \emph{nucleon} hot-spots used in recent saturation-based substructure studies that employed IP-Glasma initial conditions \cite{Schenke:2018fci}.
-Evidently, it may be necessary to place an informative prior on our nucleon substructure parameters in order to resolve the apparent tension between our parameter values and those needed to describe DIS measurements at HERA.
-Alternatively, it is also possible that the fluctuations probed by coherent and incoherent $J/\Psi$ production are different than those probed by minimum-bias particle production.
-
 \subsection{Transport properties}
@@
 
-The posterior free-streaming time $\taufs=0.37_{-0.27}^{+0.33}\ \fmc$ obtained in this work is significantly smaller than our previous estimate $\taufs=1.16_{-0.25}^{+0.29}\ \fmc$ in Ref.~\cite{Bernhard:2018hnz}.
+The posterior free-streaming time $\taufs=0.37_{-0.27}^{+0.33}\ \fmc$ obtained in this work is significantly smaller than the previous estimate $\taufs=1.16_{-0.25}^{+0.29}\ \fmc$ quoted in Ref.~\cite{Bernhard:2018hnz}.
 We point out that the present study is missing several important observables which could affect the estimated free-streaming time, e.g.\ the Pb-Pb mean $p_T$ and mean $p_T$ fluctuations at $\sqrts=5.02$~TeV.
@@
       $n_c$    & 6              & $\zeta/s$ max     & 0.032          \\
-      $w$      & 0.92 fm        & $\zeta/s$ width   & 0.024 GeV      \\
-      $v$      & 0.43 fm        & $\zeta/s$ $T_0$   & 175 MeV        \\
+      $r$      & 0.81 fm        & $\dTb$            & 0.024 GeV      \\
+      $v$      & 0.43 fm        & $\Tb$             & 175 MeV        \\
       $\dmin$  & 0.81 fm        & $\Tsw$            & 151 MeV        \\
@@
 Moreover, we obtain an excellent description of the \mbox{$p$-Pb} mean $p_T$, although this fit is somewhat less meaningful since we are unable to calibrate on the Pb-Pb mean $p_T$ simultaneously (data is not yet available).
-Additionally, the model provides a simultaneous description of the \mbox{$p$-Pb} and Pb-Pb charged-particle yields using a single entropy deposition parameter $p=0$.
+Additionally, the model provides a simultaneous description of the \mbox{$p$-Pb} and Pb-Pb charged-particle yields using a single energy deposition parameter $p=0$.
 This is the \emph{exact same} generalized mean $p$-value supported by multiple previous studies \cite{Moreland:2014oya, Bernhard:2016tnd, Ke:2016jrd, Bernhard:2018hnz}.
@@
 We also present calculations for several observables which were omitted from the calibration due to the statistical limitations of our training data.
-Here our MAP event sample is several orders-of-magnitude larger so the statistics are no issue.
+Here our MAP event sample is several orders of magnitude larger so the statistics are no issue.
 The bottom-right panel of Fig.~\ref{fig:obs_map} shows our model calculation for the four-particle elliptic flow cumultant $\vnk{2}{4}$ along with the measured data points from \mbox{ALICE} \cite{Adam:2016izf}.
@@
 For the QGP initial conditions, we employed a modified version of the \trento\ model \cite{Moreland:2014oya} which adds new parameters to vary the fluctuating size and shape of each nucleon.
-Specifically, we modeled each nucleon as a cluster of $n_c$ constituents (hot spots), where each constituent is described by a Gaussian density profile of width $v$.
-Each constituent position inside the nucleon was sampled randomly (without correlations) according to a Gaussian radial distribution of variable width, controlled indirectly by an effective nucleon width parameter $w$.
-
-The transport dynamics of the collision were simulated using a pre-equilibrium free-streaming stage of variable duration followed by boost-invariant viscous hydrodynamics for hot and dense regions of the fireball and a microscopic hadronic afterburner for the relatively dilute corona.
+Specifically, we modeled each nucleon as a cluster of $n_c$ constituents (hot spots), where each constituent is described by a Gaussian density of width $v$.
+The constituent positions were each sampled randomly (without correlations) from a Gaussian radial distribution of width $r$ centered about each pre-defined nucleon position.
+
+The transport dynamics of the collision were simulated using a pre-equilibrium free-streaming stage followed by boost-invariant viscous hydrodynamics for hot and dense regions of the fireball and a microscopic hadronic afterburner for the relatively dilute corona.
 We parametrized various sources of uncertainty in each stage of the collision including the duration of the pre-equilibrium free-streaming stage, the temperature dependence of the QGP shear and bulk viscosities, and the particlization temperature used to switch from a hydrodynamic description to microscopic Boltzmann transport.
@@
   \item
-    Using \trento\ initial conditions, we find that initial entropy deposition (or massless noninteracting particle production) scales like the geometric mean of local participant nuclear densities, in agreement with a number of previous estimates \cite{Moreland:2014oya, Bernhard:2016tnd, Bernhard:2018hnz}.
+    Using \trento\ initial conditions followed by a pre-equilibrium free-streaming stage, we find that initial energy density scales like the geometric mean of local participant nuclear densities.
   \item
@@
   \item
-    Bulk observables in $p$-Pb and Pb-Pb collisions prefer larger nucleons, similar in size to the proton's RMS electric charge radius.
+    Bulk observables in $p$-Pb and Pb-Pb collisions prefer larger nucleons of Gaussian width $w \sim 0.9$~fm.
     This agrees with a similar Bayesian analysis which was calibrated to Pb-Pb observables at $\sqrts=2.76$ and 5.02~TeV using a hybrid model without nucleon substructure \cite{Bernhard:2018hnz}.
@@
   \item VISH2+1 hydrodynamics (Fortran) \cite{osuhydro:code}
-  \item FRZOUT particle sample (Python) \cite{frzout:code}
+  \item FRZOUT particle sampler (Python) \cite{frzout:code}
   \item UrQMD microscopic transport model (Fortran) \cite{urqmd:code}
@@
 
-The emulator is a surrogate for the full physics simulation which generates probabilistic predictions for the model observables $\yv_m$ at a given point $\xv$.
+The emulator is a surrogate for the full physics simulation which generates probabilistic predictions for the model observables $\yv$ at a given point $\xv$.
 Here we validate these probabilistic predictions using a method known as k-fold cross validation.
@@
 
-More generally, we can perform the validation test in Fig.~\ref{fig:validation_example} for \emph{every} observable $y \in \yv_m$ and check that each observable's normalized residuals ${z=(y_\text{pred} - y_\text{obs})/\sigma_\text{pred}}$ follow a unit normal distribution.
+More generally, we can perform the validation test in Fig.~\ref{fig:validation_example} for \emph{every} observable $y \in \yv$ and check that each observable's normalized residuals ${z=(y_\text{pred} - y_\text{obs})/\sigma_\text{pred}}$ follow a unit normal distribution.
 This test is applied to the $p$-Pb and Pb-Pb collision systems in Fig.~\ref{fig:validation_all}.

### 2018-10-19 src/design.py

@@ -29,8 +29,6 @@ import numpy as np
 
 from . import cachedir, parse_system
 
-import matplotlib.pyplot as plt
-
 
 """
 Remove outlier design points. These points are near the edge of the
@@ -38,13 +36,9 @@ design and produce far too few particles to be physically reasonable.
 
 """
 bad_points = [
-    129, 392, 143, 273, 401, 276, 405, 153, 156, 285, 414, 159, 162, 418, 163,
-    421, 422, 296, 298, 429, 301, 308, 183, 184, 312, 442, 446, 319, 451, 453,
-    326, 199, 457, 330, 459, 460, 461, 334, 203, 464, 466,  83, 211, 338, 471,
-    472, 473, 346, 223, 355, 357, 360, 234, 363, 492, 107, 236, 367, 235, 498,
-    115, 242, 376, 249, 378
+    129, 162, 418, 451, 422, 199, 360, 234, 330, 363, 460, 461, 492, 273,
+    466, 498, 183, 249, 346, 319
 ]
-
 
 
 def generate_lhs(npoints, ndim, seed):
@@ -174,32 +168,39 @@ class Design:
         )
 
         # Version 1 of the manuscript parametrized nucleon substructure using
-        # 'nucleon_width' and 'nucleon_structure' parameters. In v2 of the
-        # manuscript, I reverted to a 'sampling_radius' and 'constituent_width'.
+        # 'nucleon_width' and 'nucleon_structure' parameters. In version 2 of the
+        # manuscript, I revert to a 'sampling_radius' and 'constituent_width'.
         # These new parameters are more physical and easier to understand.
-        self.array[:, 5] = .2 + self.array[:, 5]*(self.array[:, 3] - .2)
-        self.array[:, 3] = np.sqrt(self.array[:, 3]**2 - self.array[:, 5]**2)
-
-        design_changes = [
-            (3, 'sampling_radius', r'$r\ [\mathrm{fm}]$', (0.0, 1.2)),
-            (5, 'parton_width', r'$v\ [\mathrm{fm}]$', (0.2, 1.2))
+        nucleon_width, parton_struct = [
+            self.array[:, self.keys.index(k)]
+            for k in ('nucleon_width', 'parton_struct')
         ]
 
-        for index, key, label, (range_low, range_high) in design_changes:
-            self.keys[index] = key
-            self.labels[index] = label
-            self.range[index] = (range_low, range_high)
-            self.min[index] = range_low
[…]

### 2018-10-19 src/emulator.py

@@ -21,12 +21,11 @@ from sklearn.decomposition import PCA
 from sklearn.externals import joblib
 from sklearn.gaussian_process import GaussianProcessRegressor as GPR
 from sklearn.gaussian_process import kernels
-from sklearn.preprocessing import StandardScaler
+from sklearn.preprocessing import StandardScaler, RobustScaler
 
 from . import cachedir, lazydict, model
 from .design import Design
 
-import matplotlib.pyplot as plt
 
 class _Covariance:
     """
@@ -139,7 +138,7 @@ class Emulator:
             ) +
             kernels.WhiteKernel(
                 noise_level=.1**2,
-                noise_level_bounds=(.01**2, 1)
+                noise_level_bounds=(.01**2, 10)
             )
         )
 

### 2018-10-19 src/model.py

@@ -160,7 +160,7 @@ class ModelData:
             logging.debug('loading %s', f)
             d = np.fromfile(str(f), dtype=self.dtype)
             d.sort(order='dNch_deta')
-            return (f.stem, d)
+            return (f.stem, correct_yield(d))
 
         self.design_events = [load_events(f) for f in files]
         self.system = system
@@ -224,7 +224,7 @@ class ModelData:
                 return lambda events: flow.Cumulant(
                     events['flow'][detector]['N'],
                     *events['flow'][detector]['Qn'].T[1:]
-                ).flow(*nk, imaginary='zero')
+                ).flow(*nk, imaginary='negative')
 
             if obs.startswith('sc'):
                 mn = obs_stack.pop()
@@ -244,8 +244,6 @@ class ModelData:
             """
             trigger = events['trigger']
             minbias = (0, float('inf'))
-
-            events = correct_yield(events)
 
             if bin_type == 'cent':
                 """

### 2018-10-19 src/plots.py

@@ -1723,7 +1723,7 @@ def validation_data(system, n_splits=20):
     """
     design = Design(system)
     kf = KFold(n_splits=n_splits)
-    npc = {'pPb5020': 7, 'PbPb5020': 8}[system]
+    npc = {'pPb5020': 4, 'PbPb5020': 8}[system]
 
     mean_folds = []
     cov_folds = []
@@ -2280,6 +2280,8 @@ def _diag_emu(system=default_system, pcs=None, params=None, label_all=True):
             if label_all or ax.is_first_col():
                 ax.set_ylabel('PC {}'.format(pc + 1))
 
+            ax.set_ylim(-4, 4)
+
     set_tight(fig, w_pad=.5, h_pad=.25)
 
 
@@ -2358,16 +2360,6 @@ def grid_error():
 
     set_tight(fig)
 
-@plot
-def energy_vs_yield():
-    fig = plt.figure(figsize=figsize(.5, aspect=1))
-
-    path = Path(workdir, 'model_output', 'map', 'PbPb5020.dat')
-    model_data = model.ModelData('PbPb5020', path)
-
-    for pt, ev in model_data.design_events:
-        x, y = [ev[k] for k in ('init_entropy', 'dNch_deta')]
-        plt.scatter(x[::1000], y[::1000])
 
 @plot
 def entropy_scaling():
@@ -2413,88 +2405,6 @@ def entropy_scaling():
     set_tight(fig)
 
 
-def rms_nucleon_width(number, radius, width, samples=10**3):
-    """
-    Find center-of-mass root-mean-square radius for an ensemble of
-    sampled protons with nucleon substructure.
-
-    """
-    # sample constituent positions
-    positions = np.random.normal(
-            scale=radius,
-            size=2*number*samples
-            ).reshape(number, 2, -1)
-
-    # recenter
-    centers = positions.mean(axis=0)
-    positions -= centers
-
-    # cartesian grid
-    r = np.arange(-4*radius, 4*radius, .2*width)
-    xx, yy = np.meshgrid(r, r)
-
-    # sum over all gaussian constituents
-    rho = np.zeros_like(r)
-    for pos in positions.T:
-        for (xi, yi) in pos.T:
-            rsq = (xx - xi)**2 + (yy - yi)**2
-            rho += np.exp(-rsq/(2*width**2)).sum(axis=0)
-
-    # return rms radius in com frame
[…]

### 2018-10-22 referee_respons.txt
Begin crafting referee response.

Thank you for the constructive criticisms and comments. We have significantly
revised the manuscript to improve its clarity and address several of your concerns.

Response to the specific concerns of the referee:

"A parton model of the proton has direct consequences not only for
heavy-ion collisions, but also for p+p collisions. The authors should
present predictions for p+p scattering using the most probable
parameters (or using the posterior distribution of parameters w, v,
n_c) for the differential elastic p+p scattering cross section. The
prediction can be compared to parametrized data at 5TeV. As a proxy
for the p+p data the parametrization from Phys. Rev. D88 (2013) 094019
can be used."

We agree that it would be interesting to explore the consequences of our
proton structure parameters on p-p observables. However, doing so properly
would involve significant work beyond the scope of the present study. We
therefore believe that such an analysis would be more appropriate for an
independent follow-up study. With this in mind, we've made all software, methods,
and data open source to the broader community.

"The posterior for n_c (the number of constituents) clearly shows
that the prior range is too constrained, a large part of the posterior
extends in the region n_c>9. I do not want to ask the authors to redo
the calculation with different prior, but a comment should be added in
text."

Thank you for pointing this out. We've added the following comment to
explain the truncation of the constituent number prior:

One exception is the constituent number nc which we limit for practical considerations.
Recall that each constituent fluctuates independently, weighted by a gamma random variable.
Hence for constituent numbers $n_c > 1$, the fluctuations average out, and the resulting nucleon fluctuations vanish.
To counteract this effect, the constituent fluctuation variance must increase as nc increases.
Eventually, these required fluctuations become unreasonably large.
We find that for nc < 10, the energy density fluctuations are reasonable, and hence limit nc to this prior range."

"For the model predictions from the posterior experimental data (even preliminary) should be added, whenever available."

We've been careful to include all available data (public or otherwise), but it
is possible that we missed a dataset. In response to your suggestion, I've sent an email to the ALICE
collaboration to request their preliminary data, e.g. Pb-Pb mean pT, which could be
added to the posterior predictions where data is missing. Unfortunately, I have not heard back from them
yet, and I do not believe it is their policy to share preliminary data.
If you can direct me to a specific dataset, I would be happy to include it in
the maximum a posteriori observable plot.

Additional changes to the manuscript:

1. We've replaced the title with a more succinct version.

2. Several technical details of the initialization procedure have been
rewritten based on comments received from the community. The initial condition
parametrization is correspondingly recast as an initial energy density profile
(previously it was a parton density). We've also improved the
discussion of the initial free-streaming evolution and its matching to viscous
hydrodynamics.

3. We've simplified the nucleon substructure parameters, replacing the nucleon
width w [fm] and nucleon structure parameter X with a constituent sampling
radius r and constituent width v [fm]. The new parametrization is simpler and
less prone to misinterpretaiton. This transformation does not change the
conclusions of the study (nor should it).

4. We've corrected several inaccuracies that we made when discussing the
posterior nucleon width. In particular, the RMS radius differs from the Gaussian
width by a factor of sqrt(3).

5.

### 2018-10-22 paper/paper.tex
Begin crafting referee response.

@@
 \newcommand{\dmin}{d_\text{min}}
-\newcommand{\dTb}{\Delta T_\text{bulk}}
-\newcommand{\Tb}{T_\text{bulk}}
 \newcommand{\Tsw}{T_\text{switch}}
 \newcommand{\T}{\tilde{T}}
+
+\newcommand{\smin}{{(\eta/s)_\mathrm{min}}}
+\newcommand{\sslope}{{a_\mathrm{shear}}}
+\newcommand{\scrv}{{c_\mathrm{shear}}}
+\newcommand{\bmax}{{(\zeta/s)_\mathrm{max}}}
+\newcommand{\bloc}{{T_\text{bulk}}}
+\newcommand{\bwidth}{{\Delta T_\text{bulk}}}
+
 \newcommand{\bv}{\mathbf b}
@@
 \end{equation}
-describes the eikonal overlap of the two proton wave packets at fixed impact parameter $b$, where coordinate $(x, y)$ lies in the transverse plane and $z$ is parallel to the beam axis.
+describes the eikonal overlap of the two proton wave packets at fixed impact parameter $b$, where coordinate $(x, y)$ lies in the transverse plane and $z_{1,2}$ is parallel to the beam axis.
 Here we assume that each proton is comprised of smaller constituents---e.g.\ valence quarks, sea quarks, and small-x gluons---which may collide to produce secondary particles and contribute to the observed inelastic proton-proton cross section.
@@
 The sampling radius $r$ should thus be interpreted with care.
-It is \emph{not} the root-mean-square radius of the sampled nucleons.
+It is \emph{not} the Gaussian width of the sampled nucleons in their center-of-mass frame.
 
@@
 This wounded nucleon scaling was in fact one of the first such mappings used as a proxy for initial particle production in relativistic heavy-ion collisions \cite{Bialas:1976ed}.
-It was subsequently realized, however, that the wounded nucleon model predicts the wrong scaling for charged particle production as a function of collision centrality \cite{Kharzeev:2000ph} and hence the wrong scaling for initial energy deposition at midrapidity as a function of participant thicknesses $\T_A$ and $\T_B$\footnote{Typically an entropy density $s$ is used, but the shortcomings of the wounded nucleon model are qualitatively the same when depositing energy.}.
+It was subsequently realized, however, that the wounded nucleon model predicts the wrong scaling for charged particle production as a function of collision centrality \cite{Kharzeev:2000ph} and hence the wrong scaling for initial midrapidity energy\footnote{Typically an entropy density $s$ is used, but the shortcomings of the wounded nucleon model are qualitatively the same the energy density $e$ is used instead.} deposition as a function of participant density.
 
@@
 This assumes, of course, that the model's parametrization for entropy density is also suitable as a parametrization for the energy density.
-To this point, we note that the arguments for the generalized mean ansatz are quite general \cite{Moreland:2015oya}, and that the energy density $dE/d\eta_s$ and entropy density $dS/d\eta_s$ share many similar qualities.
+To this point, we note that the arguments for the generalized mean ansatz are quite general \cite{Moreland:2014oya}, and that the energy density $dE/d\eta_s$ and entropy density $dS/d\eta_s$ share many similar qualities.
 For example, both are approximately conserved at midrapidity, and both scale approximately linearly with the final charged-particle density $d\nch/d\eta_s$.
@@
   \label{eq:shear_viscosity}
-  (\eta/s)(T) = (\eta/s)_\mathrm{min} + (\eta/s)_\mathrm{slope}\cdot(T - T_c)\cdot(T/T_c)^{(\eta/s)_\mathrm{crv}},
-\end{equation}
-where $\eta/s$ min, slope, and curvature are tunable parameters, and $T_c=0.154$~GeV is the pseudocritical transition temperature of the HotQCD EoS.
+  (\eta/s)(T) = \smin + \sslope \cdot(T - T_c)\cdot \left ( \frac{T}{T_c} \right)^\scrv,
+\end{equation}
+where $\smin$, $\sslope$, and $\scrv$ are tunable parameters, and $T_c=0.154$~GeV is the pseudocritical transition temperature of the HotQCD EoS.
 While for the specific bulk viscosity $\zeta/s$, we use an unnormalized Cauchy distribution
@@
   \label{eq:bulk_viscosity}
-  (\zeta/s)(T) = \frac{(\zeta/s)_\text{max}}{1 + \left(\dfrac{T - T_\text{bulk}}{\Delta T_\text{bulk}}\right)^2},
-\end{equation}
-described by a tunable maximum $(\zeta/s)_\text{max}$, width $\dTb$, and location parameter $\Tb$.
+  (\zeta/s)(T) = \frac{\bmax}{1 + \left(\dfrac{T - \bloc}{\bwidth}\right)^2},
+\end{equation}
+described by a tunable maximum $\bmax$, temperature width $\bwidth$, and temperature location parameter $\bloc$.
 Figure~\ref{fig:viscosity_dof} shows several of the possible curves parametrized by Eqs.~\eqref{eq:shear_viscosity} and \eqref{eq:bulk_viscosity}, although many more are possible.
@@
     \paddedhline
-    Norm              & Normalization factor                 & 9--28           \\
-    $p$               & Entropy deposition parameter         & $-1$ to $+1$    \\
+    Norm              & Normalization factor                 & 9--28 GeV       \\
+    $p$               & Energy deposition parameter         & $-1$ to $+1$     \\
     $\sigmaf$         & Nucleon fluctuation std.\ dev.\      & 0--2            \\
@@
     $\taufs$          & Free-streaming time                  & 0.1--1.5 \fmc   \\
-    $\eta/s$ min      & Minimum value of $\eta/s$ (at $T_c$) & 0--0.2          \\
-    $\eta/s$ slope    & Slope of $\eta/s$ above $T_c$        & 0--8 GeV$^{-1}$ \\
-    $\eta/s$ crv      & Curvature of $\eta/s$ above $T_c$    & $-1$ to $+1$    \\
-    $\zeta/s$ max     & Maximum value of $\zeta/s$           & 0--0.1          \\
-    $\dTb$            & Width of $\zeta/s$ peak              & 0--0.1 GeV      \\
-    $\Tb$             & Temperature of $\zeta/s$ maximum     & 150--200 MeV    \\
+    $\smin$           & Minimum value of $\eta/s$ (at $T_c$) & 0--0.2          \\
+    $\sslope$         & Slope of $\eta/s$ above $T_c$        & 0--8 GeV$^{-1}$ \\
+    $\scrv$           & Curvature of $\eta/s$ above $T_c$    & $-1$ to $+1$    \\
+    $\bmax$           & Maximum value of $\zeta/s$           & 0--0.1          \\
+    $\bwidth$         & Width of $\zeta/s$ peak              & 0--0.1 GeV      \\
+    $\bloc$           & Temperature of $\zeta/s$ maximum     & 150--200 MeV    \\
     $\Tsw$            & Switching/particlization temp.       & 135--165 MeV    \\
@@
 \begin{equation}
-  \frac{dE_T}{d\eta_s} \bigg\vert_{\eta_s=0} = \tau_0 \int d^2x_\perp e(\xv_\perp, \eta_s=0, \tau_0).
+  \frac{dE_T}{d\eta_s} \bigg\vert_{\eta_s=0} = \tau_0 \int d^2x_\perp e(\xv_\perp, \eta_s=0, \tau_0),
 \end{equation}
@@
 
-Consider, for example, a single multiplicity bin $[\nch^\text{low}, \nch^\text{high}]$ which selects events from a minimum bias event sample with $p_T^\text{min} < p_T < p_T^\text{max}$ and $|\eta| < \Delta \eta$.
+Consider, for example, a single multiplicity bin $[\nch^\text{low}, \nch^\text{high}]$ which selects events from a minimum bias event sample with $p_T^\text{min} < p_T < p_T^\text{max}$ and $|\eta| < \eta^\text{max}$.
 Let $\langle\nch\rangle$ denote the average charged-particle multiplicity of these events.
@@
 \end{equation}
-where $E_T \propto (dE_T/d\eta_s) \vert_{\eta_s=0}$ is the midrapidity transverse energy of a single event, and $\langle E_T \rangle$ is the corresponding average over the minimum bias event sample.
+where $E_T \propto (dE_T/d\eta_s) \vert_{\eta_s=0}$ is the midrapidity transverse energy of a single event in the desired kinematic range, and $\langle E_T \rangle$ is the corresponding average transverse energy over the full minimum bias event sample.
 
@@
       \addlinespace[.4ex]
-      Norm       & \est{20.0}{2.4}{2.4}          & $\eta/s$ min     & \est{0.10}{0.07}{0.07}            \\[1.1ex]
-      $p$        & \est{0.013}{0.155}{0.144}     & $\eta/s$ slope   & \est{1.11}{1.11}{1.42} GeV$^{-1}$ \\[1.1ex]
-      $\sigmaf$  & \est{0.93}{0.27}{0.26}        & $\eta/s$ crv     & \est{-0.14}{0.86}{0.81}           \\[1.1ex]
-      $r$        & \est{0.88}{0.23}{0.25} fm     & $\zeta/s$ max    & \est{0.026}{0.026}{0.033}         \\[1.1ex]
-      $n_c$      & \est{6.3}{3.1}{2.7}           & $\dTb$           & \est{0.033}{0.033}{0.048} GeV     \\[1.1ex]
-      $v$       & \est{0.51}{0.18}{0.26} fm      & $\Tb$            & \est{0.176}{0.020}{0.024} GeV     \\[1.1ex]
-      $\dmin$    & \est{1.13}{0.51}{0.55} fm     & $\Tsw$           & \est{0.149}{0.013}{0.014} GeV     \\[1.1ex]
-      $\taufs$   & \est{0.37}{0.27}{0.33} \fmc   & \\
+      Norm       & \est{20.0}{2.5}{2.6} GeV      & $\smin$     & \est{0.08}{0.07}{0.07}            \\[1.1ex]
+      $p$        & \est{0.002}{0.180}{0.157}     & $\sslope$   & \est{1.24}{1.24}{1.46} GeV$^{-1}$ \\[1.1ex]
+      $\sigmaf$  & \est{0.91}{0.33}{0.32}        & $\scrv$     & \est{-0.09}{0.91}{0.80}           \\[1.1ex]
+      $r$        & \est{0.88}{0.23}{0.26} fm     & $\bmax$     & \est{0.026}{0.026}{0.032}         \\[1.1ex]
+      $n_c$      & \est{6.0}{3.4}{3.0}           & $\bwidth$   & \est{0.035}{0.035}{0.043} GeV     \\[1.1ex]
+      $v$        & \est{0.52}{0.20}{0.28} fm     & $\bloc$     & \est{0.174}{0.024}{0.020} GeV     \\[1.1ex]
+      $\dmin$    & \est{1.12}{0.49}{0.58} fm     & $\Tsw$      & \est{0.149}{0.014}{0.013} GeV     \\[1.1ex]
+      $\taufs$   & \est{0.47}{0.37}{0.55} \fmc   & \\
       \addlinespace[.4ex]
@@
 
-The \trento\ normalization factor ${\text{Norm} = 20.2^{+2.2}_{-2.2}}$ and energy deposition parameter $p=0.013^{+0.144}_{-0.155}$ are well constrained by the present analysis.
+The \trento\ normalization factor ${\text{Norm} = 20.0^{+2.6}_{-2.5}}$ and energy deposition parameter $p=0.002^{+0.157}_{-0.180}$ are well constrained by the present analysis.
 Moreover, Figs.~\ref{fig:obs_pbpb} and \ref{fig:obs_ppb} show that the model predictions using these values nicely describe both the $p$-Pb and Pb-Pb calibration observables.
@@
 For example, consider the ratio of the $p$-Pb charged-particle yield to the Pb-Pb charged-particle yield.
-As the energy deposition parameter $p$ trends toward positive (negative) infinity, particle production scales like the maximum (minimum) of local nuclear overlap density.
+As the energy deposition parameter $p$ trends toward positive (negative) infinity, particle production scales like the maximum (minimum) of the two nuclear thickness functions.
 This has a much stronger effect on the highly asymmetric $p$-Pb system than it does on the Pb-Pb system; hence the parameter $p$ strongly affects the ratio of the two average yields.
@@
 
-Continuing down the diagonal in Fig.~\ref{fig:posterior}, we see that the constituent sampling radius $r=0.88_{-0.23}^{+0.25}$~fm, and the constituent width $v=0.51_{-0.18}^{+0.26}$~fm.
-Their joint posterior distribution is shown in Fig.~\ref{fig:posterior_proton_shape}, illustrating the constraining power of the analysis.
-Recall that the parameter $r$ controls the dispersion of the constituent positions, while $v$ varies each constituent's Gaussian width.
-We emphasize that $r$ is model parameter, not a physical one.
-It describes the radius of the constituent position sampling distribution, \emph{not} the average proton radius, a quantity that's measured in the proton's center-of-mass frame.
-For example, when $n_c=1$, our model describes each nucleon as a single Gaussian of width $v$, yet the parameter $r$ could nevertheless assume any value within its prior.
-Despite this idiosyncrasy, one can easily calculate a physical nucleon width \emph{ex post facto}, given specific values for $r$, $v$, and $n_c$.
+Continuing down the diagonal in Fig.~\ref{fig:posterior}, we see that the constituent sampling radius $r=0.88_{-0.23}^{+0.26}$~fm, and the constituent width $v=0.52_{-0.20}^{+0.28}$~fm.
+Figure~\ref{fig:posterior_proton_shape} shows the joint posterior distribution of both parameters, illustrating the constraining power of the Bayesian analysis.
+While the sampling radius $r$ varies the size of the nucleons, we caution that its specific meaning should be interpreted with care; it specifies a computational \emph{sampling} radius, not a physical nucleon width.
+Consider, for instance, a single nucleon with $n_c = 2$ constituents.
+If the two constituents positions land on the same side of the nucleon, the effective nucleon size will be smaller than the Gaussian sampling radius $r$.
+Despite this idiosyncrasy, one can easily define a physical nucleon width in the nucleon center-of-mass frame \emph{ex post facto}, given specific values for the sampling radius $r$, constituent width $v$, and constituent number $n_c$.
 
@@
 
-For example, using the MAP estimate $r=0.88$~fm, $n_c=6$, and $v=0.51$~fm, we obtain a Gaussian nucleon width $w = 0.88$~fm, corresponding to a three-dimensional root-mean-square radius $R_p = \sqrt{3} w = 1.52$~fm.
-This estimate is consistent with the nucleon width extracted by a similar Bayesian analysis lacking nucleon substructure \note.
+For example, using the median parameter estimates $r=0.88$~fm, $n_c=6$, and $v=0.52$~fm, we can generate a large ensemble of random nucleon configurations and average their density in each nucleon's center-of-mass frame.
+This produces an ensemble-averaged nucleon density $\langle \rho \rangle$ of Gaussian width $w = 0.96$~fm, i.e.\
+
+obtain an ensemble-averaged Gaussian nucleon width $w = 0.96$~fm, corresponding to a three-dimensional root-mean-square radius $R_p = \sqrt{3} w = 1.66$~fm.
+This estimate is consistent with the larger nucleon width extracted by a similar Bayesian analysis lacking nucleon substructure \cite{Bernhard:2018hnz}.
 This is perhaps the single largest difference between our work and the conclusions of recent saturation-based calculations which constrained the event-by-event fluctuations of the proton using a color-dipole picture of vector meson production \cite{Mantysaari:2016ykx, Mantysaari:2016jaz}.
@@
 
-The posterior free-streaming time $\taufs=0.37_{-0.27}^{+0.33}\ \fmc$ obtained in this work is significantly smaller than the previous estimate $\taufs=1.16_{-0.25}^{+0.29}\ \fmc$ quoted in Ref.~\cite{Bernhard:2018hnz}.
+The posterior free-streaming time $\taufs=0.47_{-0.37}^{+0.55}~\fmc$ obtained in this work is significantly smaller than the previous estimate $\taufs=1.16_{-0.25}^{+0.29}\ \fmc$ quoted in Ref.~\cite{Bernhard:2018hnz}.
 We point out that the present study is missing several important observables which could affect the estimated free-streaming time, e.g.\ the Pb-Pb mean $p_T$ and mean $p_T$ fluctuations at $\sqrts=5.02$~TeV.
@@
       \paddedhline
-      Norm     & 20.            & $\eta/s$ min      & 0.11           \\
-      $p$      & 0.0            & $\eta/s$ slope    & 1.6 GeV$^{-1}$ \\
-      $k$      & 0.19           & $\eta/s$ curv     & -0.29          \\
-      $n_c$    & 6              & $\zeta/s$ max     & 0.032          \\
-      $r$      & 0.81 fm        & $\dTb$            & 0.024 GeV      \\
-      $v$      & 0.43 fm        & $\Tb$             & 175 MeV        \\
+      Norm     & 20. GeV        & $\smin$      & 0.11           \\
+      $p$      & 0.0            & $\sslope$    & 1.6 GeV$^{-1}$ \\
+      $k$      & 0.19           & $\scrv$      & -0.29          \\
+      $n_c$    & 6              & $\bmax$      & 0.032          \\
+      $r$      & 0.81 fm        & $\bwidth$    & 0.024 GeV      \\
+      $v$      & 0.43 fm        & $\bloc$      & 175 MeV        \\
       $\dmin$  & 0.81 fm        & $\Tsw$       & 151 MeV        \\
@@
   \item
-    We obtain an estimate $v=0.47_{-0.15}^{+0.20}$~fm for the Gaussian width of the constituent hot spots inside each nucleon.
+    We obtain an estimate $v=0.52_{-0.20}^{+0.28}$~fm for the Gaussian width of the constituent hot spots inside each nucleon.
     This is considerably larger than the length scales typically associated with nucleon substructure, and it may help explain the seemingly unreasonable success of hydrodynamics in small collision systems.
   \item
-    Our posterior estimate for the pre-equilibrium free-streaming time is $\taufs=0.37_{-0.27}^{+0.33}\ \fmc$.
+    Our posterior estimate for the pre-equilibrium free-streaming time is $\taufs=0.47_{-0.37}^{+0.55}~\fmc$.
     This is significantly shorter than the estimate obtained from a similar Bayesian study in Ref.~\cite{Bernhard:2018hnz}, which reported $\taufs=1.16_{-0.25}^{+0.29}\ \fmc$.
@@
 \begin{acknowledgments}
-  JSM thanks Berndt M\"uller, Weiyao Ke, Bj\"orn Schenke, and Heikki M\"antysaari for helpful discussions and clarifying comments.
+  JSM thanks Ulrich Heinz, Jamie Nagle, Berndt M\"uller, Weiyao Ke, Bj\"orn Schenke, and Heikki M\"antysaari for helpful discussions and clarifying comments.
   This research was completed using 6 million CPU hours provided by the National Energy Research Scientific Computing Center (NERSC), a U.S.

### 2018-10-22 src/design.py

@@ -121,20 +121,20 @@ class Design:
         self.type = 'validation' if validation else 'main'
 
         self.keys, labels, self.range = map(list, zip(*[
-            ('norm',          r'{Norm}',                      (    9,     28)),
+            ('norm',          r'{Norm} [{GeV}]',              (    9,     28)),
             ('trento_p',      r'p',                           ( -1.0,    1.0)),
             ('fluct_std',     r'\sigma {fluct}',              (  0.0,    2.0)),
             ('nucleon_width', r'w [{fm}]',                    (  0.4,    1.2)),
-            ('parton_number', r'n {constit}',                 (    1,     10)),
+            ('parton_number', r'n_c',                         (    1,     10)),
             ('parton_struct', r'\chi {struct}',               (  0.0,    1.0)),
             ('dmin3',         r'd {min} [{fm}]',              (  0.0, 1.7**3)),
             ('tau_fs',        r'\tau {fs} [{fm}/c]',          (  0.1,    1.5)),
             ('etas_min',      r'\eta/s {min}',                (  0.0,    0.2)),
-            ('etas_slope',    r'\eta/s {slope} [{GeV}^{-1}]', (  0.0,    8.0)),
-            ('etas_crv',      r'\eta/s {crv}',                ( -1.0,    1.0)),
+            ('etas_slope',    r'a {shear} [{GeV}^{-1}]',      (  0.0,    8.0)),
+            ('etas_crv',      r'c {shear}',                   ( -1.0,    1.0)),
             ('zetas_max',     r'\zeta/s {max}',               (  0.0,    0.1)),
-            ('zetas_width',   r'\zeta/s {width} [{GeV}]',     (  0.0,    0.1)),
-            ('zetas_t0',      r'\zeta/s T_0 [{GeV}]',         (0.150,  0.200)),
+            ('zetas_width',   r'\Delta \!T {bulk} [{GeV}]',   (  0.0,    0.1)),
+            ('zetas_t0',      r'T {bulk} [{GeV}]',            (0.150,  0.200)),
             ('Tswitch',       r'T {switch} [{GeV}]',          (0.135,  0.165)),
         ]))
 

### 2018-10-22 src/plots.py

@@ -706,7 +706,7 @@ def find_map():
 
     res = minimize(
         lambda x: -chain.log_posterior(full_x(x))[0],
-        x0=np.median(chain.load(*opt_params, thin=1000), axis=0),
+        x0=np.median(chain.load(*opt_params, thin=1), axis=0),
         tol=1e-8,
         bounds=[
             (a + 1e-6*(b - a), b - 1e-6*(b - a))
@@ -1200,7 +1200,7 @@ def posterior_parton_number():
     ax.fill_between(x, y, color=plt.cm.Blues(0.15), zorder=-10)
     ax.spines['left'].set_visible(False)
 
-    ax.set_xlabel('Constituent number')
+    ax.set_xlabel('Constituent number $n_c$')
     ax.set_xticks([1, 3, 5, 7, 9])
     ax.set_yticks([])
     ax.set_ylim(0, 1.01*y.max())
@@ -1241,7 +1241,7 @@ def _region(ax, name, chain, cmap=plt.cm.Blues, legend=False, title=False):
     Tc = .154
 
     samples = chain.load(
-        *['{}s_{}'.format(var, k) for k in keys], thin=100
+        *['{}s_{}'.format(var, k) for k in keys], thin=1
     )
 
     T = np.linspace(Tc if name == 'shear' else Tmin, Tmax, 1000)
@@ -1336,7 +1336,7 @@ def region_shear_bulk(cmap=plt.cm.Blues):
                 lambda T, m, w, T0: m / (1 + ((T - T0)/w)**2),
                 .08)
         ], axes):
-            samples = chain.load(*['{}s_{}'.format(var, k) for k in keys], thin=100)
+            samples = chain.load(*['{}s_{}'.format(var, k) for k in keys], thin=1)
 
             T = np.linspace(Tc if name == 'shear' else Tmin, Tmax, 1000)
 
@@ -2424,8 +2424,8 @@ def posterior_proton_shape():
 
     plt.hist2d(sampling_radius, parton_width, bins=100, cmap=plt.cm.Blues)
 
-    plt.xlabel('Constituent sampling radius [fm]')
-    plt.ylabel('Constituent width [fm]')
+    plt.xlabel('Constituent sampling radius $r$ [fm]')
+    plt.ylabel('Constituent width $v$ [fm]')
 
     ax.set_aspect('equal')
     set_tight(fig, pad=.2)

### 2018-10-23 paper/paper.tex
Update paper revisions

@@
 We posit a unified hydrodynamic and microscopic description of the quark-gluon plasma (QGP) produced in ultrarelativistic $p$-Pb and Pb-Pb collisions at $\sqrts=5.02$~TeV and evaluate our assertion using Bayesian inference.
-Specifically, we model the dynamics of both collision systems using initial conditions with parametric nucleon substructure, a pre-equilibrium free-streaming stage, event-by-event viscous hydrodynamics, and a microscopic hadronic afterburner.
-Free parameters of the model which describe the initial state and QGP medium are then simultaneously calibrated to fit charged particle yields, mean $p_T$, and flow cumulants.
+Specifically, we model the dynamics of both collision systems using initial conditions with parametric nucleon substructure, a pre-equilibrium free streaming stage, event-by-event viscous hydrodynamics, and a microscopic hadronic afterburner.
+Free parameters of the model which describe the initial state and QGP medium are then simultaneously calibrated to fit charged-particle yields, mean $p_T$, and flow cumulants.
 We argue that the global agreement of the calibrated model with the experimental data strongly supports the existence of hydrodynamic flow in small collision systems at ultrarelativistic energies, and that the flow produced develops at length scales smaller than a single proton.
@@
 With these considerations in mind, we present progress towards a fully global analysis of $p$-Pb and Pb-Pb bulk observables at $\sqrts=5.02$~TeV using a model calibration framework known as Bayesian parameter estimation.
-We begin, in Sec.~\ref{sec:model}, by constructing a nuclear collision model for $p$-Pb and Pb-Pb collisions using initial conditions with parametric nucleon substructure, and transport dynamics described by a pre-equilibrium free-streaming stage, viscous hydrodynamics and microscopic Boltzmann transport.
-In Sec.~\ref{sec:calibration}, we calibrate free parameters of the model to fit charged particle yields, mean $p_T$, and anisotropic flow cumulants of \emph{both} collision systems at $\sqrts=5.02$~TeV, and finally, in Secs.~\ref{sec:results} and \ref{sec:summary}, we present posterior results for the model input parameters and comment on the implications for hydrodynamic descriptions of small collision systems.
+We begin, in Sec.~\ref{sec:model}, by constructing a nuclear collision model for $p$-Pb and Pb-Pb collisions using initial conditions with parametric nucleon substructure, and transport dynamics described by a pre-equilibrium free streaming stage, viscous hydrodynamics and microscopic Boltzmann transport.
+In Sec.~\ref{sec:calibration}, we calibrate free parameters of the model to fit charged-particle yields, mean $p_T$, and anisotropic flow cumulants of \emph{both} collision systems at $\sqrts=5.02$~TeV, and finally, in Secs.~\ref{sec:results} and \ref{sec:summary}, we present posterior results for the model input parameters and comment on the implications for hydrodynamic descriptions of small collision systems.
 
@@
 The hydrodynamic initial conditions are provided by a modified version of the \trento\ model \cite{Moreland:2014oya} with additional parameters to vary the number and size of hot spots inside the nucleon.
-Each initial condition profile is free-streamed to the hydrodynamic starting time and matched onto the hydrodynamic energy-momentum tensor using the Landau matching procedure \cite{Broniowski:2008qk, Heinz:2015arc}.
+Each initial condition profile is free streamed to the hydrodynamic starting time and matched onto the hydrodynamic energy-momentum tensor using the Landau matching procedure \cite{Broniowski:2008qk, Heinz:2015arc}.
 Many of the components of the present model have been documented in previous studies \cite{Moreland:2014oya, Bernhard:2016tnd, Bernhard:2018hnz}; we review each component here for completeness.
@@
 This wounded nucleon scaling was in fact one of the first such mappings used as a proxy for initial particle production in relativistic heavy-ion collisions \cite{Bialas:1976ed}.
-It was subsequently realized, however, that the wounded nucleon model predicts the wrong scaling for charged particle production as a function of collision centrality \cite{Kharzeev:2000ph} and hence the wrong scaling for initial midrapidity energy\footnote{Typically an entropy density $s$ is used, but the shortcomings of the wounded nucleon model are qualitatively the same the energy density $e$ is used instead.} deposition as a function of participant density.
+It was subsequently realized, however, that the wounded nucleon model predicts the wrong scaling for charged-particle production as a function of collision centrality \cite{Kharzeev:2000ph} and hence the wrong scaling for midrapidity energy deposition\footnote{Typically, wounded nucleon scaling is used to parametrize the entropy density $s$, but the shortcomings of the model are nevertheless the same when parametrizing the system's energy density $e$.
+Both parametrizations underpredict the steep rise of particle production observed in central collisions.}
+as a function of participant density.
 
@@
 
-We note that previous applications of the \trento\ model \cite{Moreland:2014oya, Bernhard:2016tnd} used the generalized mean ansatz to parametrize the \emph{entropy} density $s(\xv_\perp, \eta_s=0, \tau_\text{hydro})$, motivated (in part) by the approximate linear relation at midrapidity between initial entropy density $dS/d\eta_s$ and final charged-particle density $d\nch/d\eta_s$ \cite{Shen:2015qta}.
+We note that previous applications of the \trento\ model \cite{Moreland:2014oya, Bernhard:2016tnd} used the generalized mean ansatz to parametrize the \emph{entropy} density $s(\xv_\perp, \eta_s=0, \tau_\text{hydro})$, motivated (in part) by the approximate linear relation at midrapidity between initial entropy density $dS/d\eta_s$ and final charged-particle density $d\nch/d\eta$ \cite{Shen:2015qta}.
 These studies, however, assumed instant thermalization, an approximation which sets $u^\mu$, $\Pi$, and $\pi^{\mu\nu}$ identically to zero at the hydrodynamic starting time $\tau_\text{hydro}$.
-The present study, meanwhile, uses a pre-equilibrium free streaming stage to estimate $u^\mu$, $\Pi$, and $\pi^{\mu\nu}$ as a function of a variable free-streaming time $\taufs$, at which point the system is matched to hydrodynamics.
+The present study, meanwhile, uses a pre-equilibrium free streaming stage to estimate $u^\mu$, $\Pi$, and $\pi^{\mu\nu}$ as a function of a variable free streaming time $\taufs$, at which point the system is matched to hydrodynamics.
 Strictly speaking, the system's entropy is ill defined prior to thermalization while the system free streams.
-We therefore reinterpret the quantity produced by the \trento\ model as an energy density $e(\xv_\perp, \eta_s=0, \tau_0)$ and free-stream this quantity to the hydrodynamic starting time $\tau_\text{hydro} \equiv \taufs$.
+We therefore reinterpret the quantity produced by the \trento\ model as an energy density $e(\xv_\perp, \eta_s=0, \tau_0)$ and free stream this quantity to the hydrodynamic starting time $\tau_\text{hydro} \equiv \taufs$.
 This assumes, of course, that the model's parametrization for entropy density is also suitable as a parametrization for the energy density.
 To this point, we note that the arguments for the generalized mean ansatz are quite general \cite{Moreland:2014oya}, and that the energy density $dE/d\eta_s$ and entropy density $dS/d\eta_s$ share many similar qualities.
-For example, both are approximately conserved at midrapidity, and both scale approximately linearly with the final charged-particle density $d\nch/d\eta_s$.
+For example, both are approximately conserved at midrapidity, and both scale approximately linearly with the final charged-particle density $d\nch/d\eta$.
 We thus expect the parametrization to be suitable for either quantity.
@@
   \includegraphics{coupling}
-  \caption{Cartoon of the free-streaming approximation for hydrodynamic initialization.
-The initial state is free-streamed for proper time $\taufs$ (zero coupling) before it is matched to hydrodynamics (strong coupling).
+  \caption{Cartoon of the free streaming approximation for hydrodynamic initialization.
+The initial state is free streamed for proper time $\taufs$ (zero coupling) before it is matched to hydrodynamics (strong coupling).
 This piecewise evolution approximates the more realistic scenario expected in nature where the medium's coupling strength smoothly changes as a function of time.}
@@
 
-There are of course two limiting cases for the strength of interactions inside the QGP medium immediately after the collision: infinitely weak coupling where the secondary partons free-stream without interacting, and infinitely strong coupling where the fluid's inter-particle mean free path effectively vanishes.
+There are of course two limiting cases for the strength of interactions inside the QGP medium immediately after the collision: infinitely weak coupling where the secondary partons free stream without interacting, and infinitely strong coupling where the fluid's inter-particle mean-free-path effectively vanishes.
 Realistically, one expects the initial parton interactions to lie somewhere between these two extremes.
-We therefore choose to model the QGP's initial off-equilibrium dynamics using a simple step-function approximation, depicted in Fig.~\ref{fig:coupling}, which free-streams the initial state for proper time $\taufs$ (zero coupling) before instantaneously switching to viscous hydrodynamics (strong coupling) \cite{Liu:2015nwa, Broniowski:2008qk}.
+We therefore choose to model the QGP's initial off-equilibrium dynamics using a simple step-function approximation, depicted in Fig.~\ref{fig:coupling}, which free streams the initial state for proper time $\taufs$ (zero coupling) before instantaneously switching to viscous hydrodynamics (strong coupling) \cite{Liu:2015nwa, Broniowski:2008qk}.
 The free parameter $\taufs$ allows us to parametrically vary the \emph{time averaged} coupling strength in the approximate window $0 < \taufs \lesssim1\ \fmc$.
@@
 
-After free-streaming for proper time $\taufs$, we transition to viscous hydrodynamics which solves the conservation equations
+After free streaming for proper time $\taufs$, we transition to viscous hydrodynamics which solves the conservation equations
 \begin{equation}
@@
 
-One primary advantage of using a microscopic transport model such as UrQMD as an afterburner, is that it realistically simulates the system break-up when the mean free path becomes large relative to the system size.
+One primary advantage of using a microscopic transport model such as UrQMD as an afterburner, is that it realistically simulates the system break-up when the mean-free-path becomes large relative to the system size.
 This dilute limit is expected to play a significant role in small collision systems where the produced medium is smaller and shorter lived.
@@
 Given values for the parameters $\xv$, the model may be used to predict a vector of simulated observables $\yv$.
-For example, $\yv$ might be a vector consisting of charged particle yields in different centrality bins.
+For example, $\yv$ might be a vector consisting of charged-particle yields in different centrality bins.
 The physics model thus represents a vector-valued function $f(\xv) = \yv$ which maps the parameter values $\xv$ to the model observables $\yv$.
@@
     $\dmin$           & Minimum inter-nucleon distance       & 0--1.7 fm       \\
-    $\taufs$          & Free-streaming time                  & 0.1--1.5 \fmc   \\
+    $\taufs$          & Free streaming time                  & 0.1--1.5 \fmc   \\
     $\smin$           & Minimum value of $\eta/s$ (at $T_c$) & 0--0.2          \\
@@
 Eventually, these required fluctuations become unreasonably large.
-We find that for $n_c < 10$, the energy density fluctuations are reasonable, and hence limit $n_c$ to this prior range.
+We find that for $n_c < 10$, the energy density fluctuations are reasonable, and hence we limit $n_c$ to this prior range.
 
@@
   \item Transverse energy production $dE_T/d\eta$.
-  \item Charged particle mean transverse momentum $\langle p_T \rangle$ ($0.15 < p_T < 10$~GeV).
+  \item Charged-particle mean transverse momentum $\langle p_T \rangle$ ($0.15 < p_T < 10$~GeV).
   \item Identified particle mean transverse momentum $\langle p_T \rangle$ of pions, kaons, and protons.
-  \item Mean transverse momentum fluctuations $\delta p_T / \langle p_T \rangle$ (charged particles, $0.15 < p_T < 2.0$~GeV).
-  \item Two-particle flow cumulants $\vnk{n}{2}$ for $n=2,3,4$\\ (charged particles, $0.2 < p_T < 5.0$~GeV for ALICE, and $0.3 < p_T < 3.0$~GeV for CMS).
-  \item Four-particle flow cumulant $\vnk{2}{4}$ \\(charged particles, $0.2 < p_T < 5.0$~GeV).
+  \item Mean transverse momentum fluctuations $\delta p_T / \langle p_T \rangle$ (charged-particles, $0.15 < p_T < 2.0$~GeV).
+  \item Two-particle flow cumulants $\vnk{n}{2}$ for $n=2,3,4$\\ (charged-particles, $0.2 < p_T < 5.0$~GeV for ALICE, and $0.3 < p_T < 3.0$~GeV for CMS).
+  \item Four-particle flow cumulant $\vnk{2}{4}$ \\(charged-particles, $0.2 < p_T < 5.0$~GeV).
   \item Symmetric cumulants $\SC(4, 2)$ and $\SC(3,2)$.
@@
 
-Most of the calibration observables listed in Table~\ref{tab:observables} are calculated as a function of collision centrality, where centrality is defined using some measure of the underlying event activity, e.g.\ the charged particle yield in a given rapidity window.
-When calculating these observables, we generate \order{4} minimum-bias events at each design point and divide the events into centrality bins using the charged particle yield at midrapidity, similar to the procedure used by experiment.
+Most of the calibration observables listed in Table~\ref{tab:observables} are calculated as a function of collision centrality, where centrality is defined using some measure of the underlying event activity, e.g.\ the charged-particle yield in a given rapidity window.
+When calculating these observables, we generate \order{4} minimum-bias events at each design point and divide the events into centrality bins using the charged-particle yield at midrapidity, similar to the procedure used by experiment.
 
@@
 where $b_i$ and $b_j$ are the midpoints of two observable bins of a single type (centrality or relative multiplicity), and $l$ is a correlation length which describes how quickly the observable bins decorrelate as the distance between the bins increases.
-We use centrality correlation lengths $l=100$ for all of the centrality binned Pb-Pb observables and $l=30$ for the centrality binned $p$-Pb charged particle yield $d\nch/d\eta$.
+We use centrality correlation lengths $l=100$ for all of the centrality binned Pb-Pb observables and $l=30$ for the centrality binned $p$-Pb charged-particle yield $d\nch/d\eta$.
 The \mbox{$p$-Pb} mean $p_T$ and flow observables, meanwhile, use relative multiplicity bins $\nch / \langle \nch \rangle$ and $\ntrk / \langle \ntrk \rangle$ which necessitate a smaller correlation length $l=5$.
@@
 
-For example, using the median parameter estimates $r=0.88$~fm, $n_c=6$, and $v=0.52$~fm, we can generate a large ensemble of random nucleon configurations and average their density in each nucleon's center-of-mass frame.
-This produces an ensemble-averaged nucleon density $\langle \rho \rangle$ of Gaussian width $w = 0.96$~fm, i.e.\
-
-obtain an ensemble-averaged Gaussian nucleon width $w = 0.96$~fm, corresponding to a three-dimensional root-mean-square radius $R_p = \sqrt{3} w = 1.66$~fm.
-This estimate is consistent with the larger nucleon width extracted by a similar Bayesian analysis lacking nucleon substructure \cite{Bernhard:2018hnz}.
+For example, using the posterior distribution's median values, $r=0.88$~fm, $n_c=6$, and $v=0.52$~fm, we can generate a large ensemble of random nucleon configurations and average their density in each nucleon's center-of-mass frame.
+The resulting ensemble-averaged nucleon density,
+\begin{equation}
+  \bar{\rho}_n(\xv) = \frac{1}{(2 \pi w^2)^{3/2}} \exp \left( -\frac{|\xv|^2}{2 w^2} \right),
+\end{equation}
+is described by a single Gaussian of width $w = 0.96$~fm.
+This nucleon width is consistent with a previous estimate, $w = 0.96_{-0.05}^{+0.04}$~fm, obtained by a similar Bayesian analysis of Pb-Pb collisions at $\sqrts=2.76$ and 5.02~TeV using a physics model without nucleon substructure \cite{Bernhard:2018hnz}.
+
 This is perhaps the single largest difference between our work and the conclusions of recent saturation-based calculations which constrained the event-by-event fluctuations of the proton using a color-dipole picture of vector meson production \cite{Mantysaari:2016ykx, Mantysaari:2016jaz}.
@@
 
-The posterior on the constituent number $n_c$, shown enlarged in Fig.~\ref{fig:posterior_parton_number}, is not sharply peaked.
-We therefore refrain from quoting a distribution median and 90\% credible interval, although we do note that the distribution clearly favors $n_c > 1$ constituents.
+Moving on, we direct our attention to the constituent number $n_c$, shown enlarged in Fig.~\ref{fig:posterior_parton_number}.
+The distribution is not sharply peaked, and hence we refrain from quoting a distribution median and 90\% credible interval.
+Note, however, that the posterior clearly favors $n_c > 1$ constituents.
 This is not surprising.
@@
 
-The posterior free-streaming time $\taufs=0.47_{-0.37}^{+0.55}~\fmc$ obtained in this work is significantly smaller than the previous estimate $\taufs=1.16_{-0.25}^{+0.29}\ \fmc$ quoted in Ref.~\cite{Bernhard:2018hnz}.
-We point out that the present study is missing several important observables which could affect the estimated free-streaming time, e.g.\ the Pb-Pb mean $p_T$ and mean $p_T$ fluctuations at $\sqrts=5.02$~TeV.
-Nevertheless, it appears that the inclusion of nucleon substructure significantly reduces the maximum allowed free-streaming time, although more work is needed to establish if this is indeed the case.
+The posterior free streaming time $\taufs=0.47_{-0.37}^{+0.55}~\fmc$ obtained in this work is significantly smaller than the previous estimate $\taufs=1.16_{-0.25}^{+0.29}\ \fmc$ quoted in Ref.~\cite{Bernhard:2018hnz}.
+We point out that the present study is missing several important observables which could affect the estimated free streaming time, e.g.\ the Pb-Pb mean $p_T$ and mean $p_T$ fluctuations at $\sqrts=5.02$~TeV.
+Nevertheless, it appears that the inclusion of nucleon substructure significantly reduces the maximum allowed free streaming time, although more work is needed to establish if this is indeed the case.
 
@@
 \end{equation}
-
 \begin{fullpage}
@@
 \end{fullpage}
+\noindent
 which are expected to be less sensitive to the medium response and more sensitive to the properties of the initial state.
@@
 Thus we can only speculate what might have caused the large difference in the MAP estimate for the symmetric flow cumulants.
-Two reasonable culprits would be the inclusion of nucleon substructure and the large difference in the preferred pre-equilibrium free-streaming time determined by the two studies.
+Two reasonable culprits would be the inclusion of nucleon substructure and the large difference in the preferred pre-equilibrium free streaming time determined by the two studies.
 
@@
 
-The transport dynamics of the collision were simulated using a pre-equilibrium free-streaming stage followed by boost-invariant viscous hydrodynamics for hot and dense regions of the fireball and a microscopic hadronic afterburner for the relatively dilute corona.
-We parametrized various sources of uncertainty in each stage of the collision including the duration of the pre-equilibrium free-streaming stage, the temperature dependence of the QGP shear and bulk viscosities, and the particlization temperature used to switch from a hydrodynamic description to microscopic Boltzmann transport.
+The transport dynamics of the collision were simulated using a pre-equilibrium free streaming stage followed by boost-invariant viscous hydrodynamics for hot and dense regions of the fireball and a microscopic hadronic afterburner for the relatively dilute corona.
+We parametrized various sources of uncertainty in each stage of the collision including the duration of the pre-equilibrium free streaming stage, the temperature dependence of the QGP shear and bulk viscosities, and the particlization temperature used to switch from a hydrodynamic description to microscopic Boltzmann transport.
 
@@
 First, we demonstrated in Fig.~\ref{fig:obs_map} the existence of a single set of model parameters which can simultaneously describe $p$-Pb and Pb-Pb charged-particle yields, mean $p_T$, and flow cumulants at $\sqrts=5.02$~TeV.
-The excellent quantitative agreement of the model calculation with the experimental data is strong evidence for a unified hydrodynamic description of $p$-A and A-A collisions at ultrarelativistic energies.
+The excellent quantitative agreement of the model calculation with the experimental data is strong evidence for a unified hydrodynamic description of $p$-Pb and Pb-Pb collisions at ultrarelativistic energies.
 Moreover, the modifications to the physics model which were required to obtain this agreement are generally modest; one must simply replace Gaussian nucleons with composite nucleons of several or more constituents.
@@
   \item
-    Using \trento\ initial conditions followed by a pre-equilibrium free-streaming stage, we find that initial energy density scales like the geometric mean of local participant nuclear densities.
+    Using \trento\ initial conditions followed by a pre-equilibrium free streaming stage, we find that the initially produced energy density scales like the geometric mean of participant nuclear thickness; see Eq.~\eqref{eq:geometric_mean}.
+    Previous Bayesian studies have reported similar scaling for the initially produced \emph{entropy} density \cite{Bernhard:2016tnd, Ke:2016jrd}, although these studies assumed static initialization, an approximation which neglects the initial values of $u^\mu$, $\pi^{\mu\nu}$ and $\Pi$ at the hydrodynamic starting time.
   \item
-    Nucleon substructure is necessary to simultaneously describe $p$-Pb and Pb-Pb bulk observables. However, there is no strong preference for a specific number of constituents inside the nucleon.
+    We find that nucleon substructure is necessary to simultaneously describe $p$-Pb and Pb-Pb bulk observables.
+    However, we observe no strong preference for a specific number of constituents (hot spots) inside the nucleon.
   \item
-    Bulk observables in $p$-Pb and Pb-Pb collisions prefer larger nucleons of Gaussian width $w \sim 0.9$~fm.
-    This agrees with a similar Bayesian analysis which was calibrated to Pb-Pb observables at $\sqrts=2.76$ and 5.02~TeV using a hybrid model without nucleon substructure \cite{Bernhard:2018hnz}.
+    The present Bayesian analysis prefers larger nucleons, $w \sim 1$~fm, in agreement with a similar Bayesian analysis calibrated to Pb-Pb data at $\sqrts=2.76$ and 5.02~TeV using a hybrid model without nucleon substructure \cite{Bernhard:2018hnz}.
     We note that our result is in significant tension with an estimate for the effective nucleon width based on the gluon distribution extracted from HERA data \cite{Rezaeian:2012ji}.
@@
     We obtain an estimate $v=0.52_{-0.20}^{+0.28}$~fm for the Gaussian width of the constituent hot spots inside each nucleon.
-    This is considerably larger than the length scales typically associated with nucleon substructure, and it may help explain the seemingly unreasonable success of hydrodynamics in small collision systems.
+    This width is considerably larger than the length scales typically associated with nucleon substructure.
+    It may help explain the seemingly unreasonable success of hydrodynamics in small collision systems.
   \item
-    Our posterior estimate for the pre-equilibrium free-streaming time is $\taufs=0.47_{-0.37}^{+0.55}~\fmc$.
-    This is significantly shorter than the estimate obtained from a similar Bayesian study in Ref.~\cite{Bernhard:2018hnz}, which reported $\taufs=1.16_{-0.25}^{+0.29}\ \fmc$.
+    We report an estimate $\taufs=0.47_{-0.37}^{+0.55}~\fmc$ for the model's pre-equilibrium free streaming time.
+    This is significantly shorter than the estimate obtained from a similar Bayesian study in Ref.~\cite{Bernhard:2018hnz}, which found $\taufs=1.16_{-0.25}^{+0.29}\ \fmc$.
     It is not clear whether the difference is a result of nucleon substructure or the different observables used to calibrate each analysis.
@@
     We compare in Fig.~\ref{fig:region_shear_bulk} our estimate for the temperature dependence of the QGP specific shear and bulk viscosities to those of Ref.~\cite{Bernhard:2018hnz}, which performed a Bayesian calibration to Pb-Pb bulk observables at $\sqrts=2.76$ and 5.02~TeV using a physics model without nucleon substructure.
-    The two studies are in good agreement, although Ref.~\cite{Bernhard:2018hnz} obtains a more precise estimate for $(\eta/s)(T)$, likely due to the additional beam energies and observables included, and an enhanced sensitivity of larger collision systems to the QGP viscosity.
+    The two studies are in good agreement, although Ref.~\cite{Bernhard:2018hnz} obtains a more precise estimate for $(\eta/s)(T)$, likely due to the additional beam energies and observables included and an enhanced sensitivity of larger collision systems to the QGP viscosity.
   \item
@@
   \item \trento\ with nucleon substructure (C++) \cite{trento:code}
-  \item Pre-equilibrium free-streaming (Python) \cite{freestream:code}
+  \item Pre-equilibrium free streaming (Python) \cite{freestream:code}
   \item VISH2+1 hydrodynamics (Fortran) \cite{osuhydro:code}
@@
 The emulators generally behave as expected, although the validation is somewhat better for the Pb-Pb system than the $p$-Pb system.
-For instance, the $p$-Pb charged particle yield $d\nch/d\eta$ uncertainties are over predicted.
+For instance, the $p$-Pb charged-particle yield $d\nch/d\eta$ uncertainties are over predicted.
 It is not immediately clear why this would be the case, but the MAP observables in Fig.~\ref{fig:obs_map} are in good agreement with their emulator predictions which suggests it should not be a grave concern.

### 2019-03-20 paper/paper.tex
dissertation

@@
 Typically, the macroscopic structure of heavy nuclei, characterized e.g.\ by an atomic mass number and set of Woods-Saxon coefficients \cite{MOLLER1995185, DEVRIES1987495}, is regarded as a known input to hydrodynamic models which contributes negligible uncertainty to simulation predictions, outweighed by large uncertainties in modeling initial energy deposition and off-equilibrium dynamics \cite{Niemi:2014lha, Song:2011hk, Retinskaya:2013gca, Liu:2015nwa, Kurkela:2016vts}.
-The geometry of light ions, meanwhile, is naturally more sensitive to the detailed size and shape of individual protons and neutrons inside the nucleus, which may fluctuate event-by-event and differ signficantly from the round blobs typically used to approximate nucleons in heavy-ion collisions \cite{Schenke:2014zha, Welsh:2016siu, Moreland:2017kdx, Schenke:2014gaa, Schlichting:2014ipa}.
+The geometry of light ions, meanwhile, is naturally more sensitive to the detailed size and shape of individual protons and neutrons inside the nucleus, which may fluctuate event-by-event and differ significantly from the round blobs typically used to approximate nucleons in heavy-ion collisions \cite{Schenke:2014zha, Welsh:2016siu, Moreland:2017kdx, Schenke:2014gaa, Schlichting:2014ipa}.
 These nucleon substructure properties are difficult to measure and calculate from first principles and hence contribute significant uncertainty to model predictions of small systems.

### 2019-03-20 src/design.py

@@ -127,14 +127,14 @@ class Design:
             ('nucleon_width', r'w [{fm}]',                    (  0.4,    1.2)),
             ('parton_number', r'n_c',                         (    1,     10)),
             ('parton_struct', r'\chi {struct}',               (  0.0,    1.0)),
-            ('dmin3',         r'd {min} [{fm}]',              (  0.0, 1.7**3)),
+            ('dmin3',         r'd_{min}^3 [{fm}]',            (  0.0, 1.7**3)),
             ('tau_fs',        r'\tau {fs} [{fm}/c]',          (  0.1,    1.5)),
             ('etas_min',      r'\eta/s {min}',                (  0.0,    0.2)),
-            ('etas_slope',    r'a {shear} [{GeV}^{-1}]',      (  0.0,    8.0)),
-            ('etas_crv',      r'c {shear}',                   ( -1.0,    1.0)),
+            ('etas_slope',    r'\eta/s {slope} [{GeV}^{-1}]', (  0.0,    8.0)),
+            ('etas_crv',      r'\eta/s {crv}',                ( -1.0,    1.0)),
             ('zetas_max',     r'\zeta/s {max}',               (  0.0,    0.1)),
-            ('zetas_width',   r'\Delta \!T {bulk} [{GeV}]',   (  0.0,    0.1)),
-            ('zetas_t0',      r'T {bulk} [{GeV}]',            (0.150,  0.200)),
+            ('zetas_width',   r'\zeta/s {width} [{GeV}]',     (  0.0,    0.1)),
+            ('zetas_t0',      r'\zeta/s {loc} [{GeV}]',       (0.150,  0.200)),
             ('Tswitch',       r'T {switch} [{GeV}]',          (0.135,  0.165)),
         ]))
 

### 2019-03-20 src/plots.py

@@ -155,7 +155,7 @@ def plot(f):
 
 
 
-def figsize(relwidth=1, aspect=.618, refwidth=6.5):
+def figsize(relwidth=1, aspect=.618, refwidth=4.98):
     """
     Return figure dimensions from a relative width (to a reference width) and
     aspect ratio (default: 1/golden ratio).
@@ -243,7 +243,7 @@ def format_system(system):
     else:
         prefix = 'G'
 
-    return '{} {} {}eV'.format('-'.join(proj), energy, prefix)
+    return '{} {} {}eV'.format('-'.join(proj).replace('p', '$p$'), energy, prefix)
 
 
 def darken(color_hex, amount=.5):
@@ -348,7 +348,7 @@ def _observables_plots():
                 r'$dN_\mathrm{ch}/d\eta,\ dN/dy,\ dE_T/d\eta\ [\mathrm{GeV}]$'
             ),
             xlim=dict(
-                pPb5020=(0, 60),
+                pPb5020=(0, 59),
                 PbPb5020=(0, 80),
             ),
             ylim=(1e2, 1e5),
@@ -363,44 +363,30 @@ def _observables_plots():
         dict(
             title='Mean $p_T$',
             xlabel=dict(
-                pPb5020=r'$n_\mathrm{ch} / \langle n_\mathrm{ch} \rangle$',
+                pPb5020=r'$N_\mathrm{ch} / \langle N_\mathrm{ch} \rangle$',
                 PbPb5020='Centrality %'
             ),
-            ylabel=r'$\langle p_T \rangle$ [GeV]',
+            ylabel=r'$\langle p_T \rangle\ [\mathrm{GeV}]$',
             xlim=dict(
-                pPb5020=(1, 6),
+                pPb5020=(1, 6.5),
                 PbPb5020=(0, 80),
             ),
-            ylim=(0, 1.5),
+            ylim=(0, 1.65),
             subplots=[
                 ('mean_pT', None, dict(label='ch')),
-                *id_parts_plots('iden_mean_pT')
+                *id_parts_plots('iden_mean_pT'),
             ]
-        ),
-        dict(
-            title='Mean $p_T$ fluctuations',
-            xlabel=dict(
-                pPb5020=r'$n_\mathrm{ch} / \langle n_\mathrm{ch} \rangle$',
-                PbPb5020='Centrality %'
-            ),
-            ylabel=r'$\delta p_T/\langle p_T \rangle$',
[…]

### 2019-04-29 paper/response.txt

--- 2018-10-22 referee_respons.txt
@@
-Thank you for the constructive criticisms and comments. We have significantly
-revised the manuscript to improve its clarity and address several of your concerns.
+Thank you for the constructive criticisms and comments. We have significantly revised the manuscript to improve its clarity and address several of your concerns.
 
@@
 
-"A parton model of the proton has direct consequences not only for
-heavy-ion collisions, but also for p+p collisions. The authors should
-present predictions for p+p scattering using the most probable
-parameters (or using the posterior distribution of parameters w, v,
-n_c) for the differential elastic p+p scattering cross section. The
-prediction can be compared to parametrized data at 5TeV. As a proxy
-for the p+p data the parametrization from Phys. Rev. D88 (2013) 094019
-can be used."
+"A parton model of the proton has direct consequences not only for heavy-ion collisions, but also for p+p collisions. The authors should present predictions for p+p scattering using the most probable
+parameters (or using the posterior distribution of parameters w, v, n_c) for the differential elastic p+p scattering cross section.
+The prediction can be compared to parametrized data at 5TeV.
+As a proxy
+for the p+p data the parametrization from Phys. Rev. D88 (2013) 094019 can be used."
 
-We agree that it would be interesting to explore the consequences of our
-proton structure parameters on p-p observables. However, doing so properly
-would involve significant work beyond the scope of the present study. We
-therefore believe that such an analysis would be more appropriate for an
-independent follow-up study. With this in mind, we've made all software, methods,
-and data open source to the broader community.
+We agree that it would be interesting to explore the consequences of our proton structure parameters on p-p observables.
+However, doing so properly would involve significant work beyond the scope of the present study.
+We therefore believe that such an analysis would be more appropriate for an independent follow-up study.
+With this in mind, we've made all software, methods, and data open source to the broader community.
 
-"The posterior for n_c (the number of constituents) clearly shows
-that the prior range is too constrained, a large part of the posterior
-extends in the region n_c>9. I do not want to ask the authors to redo
-the calculation with different prior, but a comment should be added in
-text."
+"The posterior for n_c (the number of constituents) clearly shows that the prior range is too constrained, a large part of the posterior extends in the region n_c>9.
+I do not want to ask the authors to redo the calculation with different prior, but a comment should be added in text."
 
-Thank you for pointing this out. We've added the following comment to
-explain the truncation of the constituent number prior:
+Thank you for pointing this out. We've added the following comment to explain the truncation of the constituent number prior:
 
-One exception is the constituent number nc which we limit for practical considerations.
+One exception is the constituent number $n_c$ which we limit for practical considerations.
 Recall that each constituent fluctuates independently, weighted by a gamma random variable.
 Hence for constituent numbers $n_c > 1$, the fluctuations average out, and the resulting nucleon fluctuations vanish.
-To counteract this effect, the constituent fluctuation variance must increase as nc increases.
+To counteract this effect, the constituent fluctuation variance must increase as $n_c$ increases.
 Eventually, these required fluctuations become unreasonably large.
-We find that for nc < 10, the energy density fluctuations are reasonable, and hence limit nc to this prior range."
+We find that for $n_c < 10$, the energy density fluctuations are reasonable, and hence limit $n_c$ to this prior range.
 
@@
 
-We've been careful to include all available data (public or otherwise), but it
-is possible that we missed a dataset. In response to your suggestion, I've sent an email to the ALICE
-collaboration to request their preliminary data, e.g. Pb-Pb mean pT, which could be
-added to the posterior predictions where data is missing. Unfortunately, I have not heard back from them
-yet, and I do not believe it is their policy to share preliminary data.
-If you can direct me to a specific dataset, I would be happy to include it in
-the maximum a posteriori observable plot.
+We've been careful to include all available data (public or otherwise), but it is possible that we missed a dataset.
+In response to your suggestion, we've sent an email to the ALICE collaboration to request their preliminary data, e.g.
+Pb-Pb mean pT, which could be added to the posterior predictions where data is missing.
+Unfortunately, I have not heard back from them yet, and I do not believe it is their policy to share preliminary data.
+If you can direct us to a specific dataset, we would be happy to include it in the maximum a posteriori observable plot.
 
@@
 
-2. Several technical details of the initialization procedure have been
-rewritten based on comments received from the community. The initial condition
-parametrization is correspondingly recast as an initial energy density profile
-(previously it was a parton density). We've also improved the
-discussion of the initial free-streaming evolution and its matching to viscous
-hydrodynamics.
+2.
+Several technical details of the initialization procedure have been rewritten based on comments received from the community.
+The initial condition parametrization is now presented as an energy density profile.
+Previously it was presented as a parton density with each parton carry some average energy per particle.
+We've also improved the discussion of the initial free-streaming evolution and its matching to viscous hydrodynamics.
 
-3. We've simplified the nucleon substructure parameters, replacing the nucleon
-width w [fm] and nucleon structure parameter X with a constituent sampling
-radius r and constituent width v [fm]. The new parametrization is simpler and
-less prone to misinterpretaiton. This transformation does not change the
-conclusions of the study (nor should it).
+3.
+We've simplified the nucleon substructure parameters, replacing the nucleon width w [fm] and nucleon structure parameter X with a constituent sampling radius r and constituent width v [fm].
+The new parametrization is simpler and less prone to misinterpretaiton.
+This transformation does not change the conclusions of the study (nor should it); we simply applied a transformation to the training data before emulation.
 
-4. We've corrected several inaccuracies that we made when discussing the
-posterior nucleon width. In particular, the RMS radius differs from the Gaussian
-width by a factor of sqrt(3).
-
-5.
+4. We've corrected several inaccuracies that were made when discussing the posterior nucleon width.

### 2019-05-03 src/design.py
Second round of PRC edits

@@ -143,7 +143,7 @@ class Design:
         #   - escape spaces
         #   - surround with $$
         self.labels = [
-            re.sub(r'({[A-Za-z]+})', r'\mathrm\1', i)
+            re.sub(r'({[A-Za-z]+})', r'\\mathrm\1', i)
             .replace(' ', r'\ ')
             .join('$$')
             for i in labels

### 2019-05-03 src/emulator.py

@@ -21,7 +21,7 @@ from sklearn.decomposition import PCA
 from sklearn.externals import joblib
 from sklearn.gaussian_process import GaussianProcessRegressor as GPR
 from sklearn.gaussian_process import kernels
-from sklearn.preprocessing import StandardScaler, RobustScaler
+from sklearn.preprocessing import StandardScaler
 
 from . import cachedir, lazydict, model
 from .design import Design
@@ -234,7 +234,7 @@ class Emulator:
     def predict(self, X, return_cov=False):
         """
         Predict model output at `X`.
-        X must be a 2D array-like with shape ``(nsamples, ndim)``.  It is passed
+        X must be a 2D array-like with shape ``(nsamples, ndim)``. It is passed
         directly to sklearn :meth:`GaussianProcessRegressor.predict`.
         If `return_cov` is true, return a tuple ``(mean, cov)``, otherwise only
         return the mean.

### 2019-05-03 src/expt.py

@@ -4,11 +4,9 @@ Prints all data when run as a script.
 """
 
 from collections import defaultdict
-import copy
 import logging
 import pickle
 import re
-from statistics import mean
 from urllib.request import urlopen
 
 import numpy as np
@@ -228,7 +226,7 @@ def pPb5020_yield():
 
     y, stat, sys = np.array([[
         (y['value'], y['errors'][0]['symerror'], y['errors'][1]['symerror'])
-        for (x, y) in zip(dset.x('$\eta_{lab}$'), dset.y(name))
+        for (x, y) in zip(dset.x(r'$\eta_{lab}$'), dset.y(name))
         if eta_lab_min < x['low'] and x['high'] < eta_lab_max
     ] for name in dset.names[:-drop_bins]]).T
 
@@ -285,10 +283,10 @@ def pPb5020_mean_pT():
     )
 
 
-def pPb5020_flows(mode):
+def pPb5020_flows(n, k):
     """
-    The CMS p+Pb flows are not posted on HEP data so the data files are included
-    in the git repo parent directory under expt.
+    The CMS p+Pb flows are not posted on HEP data so the data files
+    are included in the git repo parent directory under expt.
 
     reference: https://inspirehep.net/record/1231945
 
@@ -296,16 +294,14 @@ def pPb5020_flows(mode):
     # Mean Ntrk offline 0-100% centrality
     Ntrk_avg = 40.
 
-    # Drop top-5 most central bins
-    xlo, xhi, x, y, stat, sys = np.loadtxt(
-        'expt/CMS_pPb5020_v{}2_sub.txt'.format(mode)
-    )[:-5].T
+    Ntrk_lo, Ntrk_hi, Ntrk, vnk, stat, sys = np.loadtxt(
+        f'expt/CMS_pPb5020_v{n}{k}.txt').T
 
     return dict(
-        mult=list(zip(xlo/Ntrk_avg, xhi/Ntrk_avg)),
-        x=x/Ntrk_avg,
-        y=y,
-        yerr=dict(stat=stat, sys=sys),
+        mult=list(zip(Ntrk_lo/Ntrk_avg, Ntrk_hi/Ntrk_avg)),
+        x=Ntrk/Ntrk_avg,
+        y=vnk,
+        yerr=dict(stat=stat, sys=sys)
     )
 
 
@@ -331,44 +327,41 @@ def _data():
     """
     data = {s: {} for s in systems}
 
-    if 'pPb5020' in systems:
[…]

### 2019-05-03 src/mcmc.py

@@ -123,6 +123,7 @@ class Chain:
     # list of 2-tuples: (obs, [list of subobs])
     PbPb5020 = [
         ('dNch_deta', [None]),
+        ('mean_pT', [None]),
         ('vnk', [(2, 2), (3, 2), (4, 2)]),
     ]
     pPb5020 = [

### 2019-05-03 src/model.py

@@ -34,9 +34,9 @@ collision event generator
 
 Of course, if you have a different data organization scheme and/or format,
 that's fine.  Modify the code for your needs.
+
 """
 
-import copy
 import logging
 from pathlib import Path
 import pickle
@@ -47,12 +47,12 @@ from sklearn.externals import joblib
 
 from . import workdir, cachedir, systems, lazydict, expt
 from .design import Design
-from .correct import correct_yield, correct_centrality
+from .correct import correct_yield
 
 
 def pT_fluct(events):
     """
-    Compute the relative mean pT fluctuation \sqrt{C_m}/M(p_T)_m, defined in
+    Compute the relative mean pT fluctuation \\sqrt{C_m}/M(p_T)_m, defined in
     Eqs. (2-3) of the ALICE paper https://inspirehep.net/record/1307102.
 
     """
@@ -149,9 +149,14 @@ class ModelData:
         ('dET_deta', float_t),
         ('mean_pT', [('N', float_t), ('pT', float_t)]),
         ('iden_dN_dy', [(s, float_t) for s in species]),
-        ('iden_mean_pT', [(s, [('N', float_t), ('pT', float_t)]) for s in species]),
-        ('pT_fluct', [('N', int_t), ('sum_pT', float_t), ('sum_pTsq', float_t)]),
-        ('flow', [(d, [('N', int_t), ('Qn', complex_t, 8)]) for d in ('alice', 'cms')]),
+        ('iden_mean_pT', [(s, [
+            ('N', float_t), ('pT', float_t)
+        ]) for s in species]),
+        ('pT_fluct', [
+            ('N', int_t), ('sum_pT', float_t), ('sum_pTsq', float_t)
+        ]),
+        ('flow', [(d, [('N', int_t), ('Qn', complex_t, 8)])
+                  for d in ('alice', 'cms')]),
     ])
 
     def __init__(self, system, *files):
@@ -292,7 +297,7 @@ def _data(system, dataset='main'):
         - 'map' (maximum a posteriori, i.e. "best-fit" point)
 
     """
-    if dataset not in {'main', 'validation', 'map'}:
+    if dataset not in {'main', 'map'}:
         raise ValueError('invalid dataset: {}'.format(dataset))
 
     files = (

### 2019-05-03 src/plots.py

@@ -2484,7 +2484,7 @@ def grid_error():
     set_tight(fig)
 
 
-@plot
+#@plot
 def entropy_scaling():
     """
     Plot initial entropy vs final dNch/deta
@@ -2554,7 +2554,7 @@ def posterior_proton_shape():
     set_tight(fig, pad=.2)
 
 
-@plot
+#@plot
 def statistics(system='pPb5020'):
     """
     Event statistics in each p-Pb trigger bin

### 2019-05-03 paper/paper.tex
Additional PRC edits

--- 2018-10-23 paper/paper.tex
@@
 \newcommand{\zv}{\mathbf z}
+\newcommand{\ym}{{\mathbf y}_\text{model}}
+\newcommand{\ye}{{\mathbf y}_\text{expt}}
+\newcommand{\Sigmam}{\Sigma_\text{model}}
+\newcommand{\Sigmae}{\Sigma_\text{expt}}
 \newcommand{\paddedhline}{\noalign{\smallskip}\hline\noalign{\smallskip}}
 \newcommand{\order}[1]{$\mathcal O(10^{#1})$}
-\newcommand{\note}{\textcolor{theblue}{[?]}}
 
@@
 One method for reducing theoretical uncertainties is to test model calculations by varying the species of colliding nuclei at a single beam energy \cite{Adare:2015bua, Schenke:2014tga, Aidala:2018mcw, Adare:2017wlc, Adamczyk:2015obl, Shen:2016zpp, Aidala:2017ajz, Adare:2006ti}.
-Since initial condition and hydrodynamic models generally factorize the structure of the colliding nuclei from the subsequent time dynamics of the collision, a single theory framework can be simultaneously tested and compared to measurements from multiple collision systems using a self-consistent set of model parameters where only the nuclear structure in the model is permitted to vary.
+Since initial condition and hydrodynamic models generally factorize the structure of the colliding nuclei from the subsequent time dynamics of the collision, a given theory framework can be validated by testing its predictions for multiple collision systems, using a single self-consistent set of model parameters where only the nuclear structure in the model is permitted to vary.
 
 Typically, the macroscopic structure of heavy nuclei, characterized e.g.\ by an atomic mass number and set of Woods-Saxon coefficients \cite{MOLLER1995185, DEVRIES1987495}, is regarded as a known input to hydrodynamic models which contributes negligible uncertainty to simulation predictions, outweighed by large uncertainties in modeling initial energy deposition and off-equilibrium dynamics \cite{Niemi:2014lha, Song:2011hk, Retinskaya:2013gca, Liu:2015nwa, Kurkela:2016vts}.
-The geometry of light ions, meanwhile, is naturally more sensitive to the detailed size and shape of individual protons and neutrons inside the nucleus, which may fluctuate event-by-event and differ signficantly from the round blobs typically used to approximate nucleons in heavy-ion collisions \cite{Schenke:2014zha, Welsh:2016siu, Moreland:2017kdx, Schenke:2014gaa, Schlichting:2014ipa}.
+The geometry of light-ions, meanwhile, is naturally more sensitive to the detailed size and shape of individual protons and neutrons inside the nucleus, which may fluctuate event-by-event and differ significantly from the round blobs typically used to approximate nucleons in heavy-ion collisions \cite{Schenke:2014zha, Welsh:2016siu, Moreland:2017kdx, Schenke:2014gaa, Schlichting:2014ipa}.
 These nucleon substructure properties are difficult to measure and calculate from first principles and hence contribute significant uncertainty to model predictions of small systems.
 
-Early substructure studies replaced round protons with composite protons, described by a few salient model parameters, in order to investigate the effect of each parameter on simulated observables \cite{Adler:2013aqf, Mitchell:2016jio, Welsh:2016siu, Broniowski:2016pvx, Bozek:2017jog}.
+Early substructure studies replaced round nucleons with composite nucleons, described by a few salient model parameters, in order to investigate the effect of each parameter on simulated observables \cite{Adler:2013aqf, Mitchell:2016jio, Welsh:2016siu, Broniowski:2016pvx, Bozek:2017jog}.
 These sensitivity studies were able to identify cause and effect relationships between model inputs and outputs, but lacked the ability to constrain nucleon substructure parameters in any kind of global or systematic fashion.
-It quickly became apparent that numerous substructure implementations might be compatible with available data, and that additional work would be required to identify observables which are particularly sensitive to the average size, shape and fluctuations of the proton.
+It quickly became apparent that numerous substructure implementations might be compatible with available data, and that additional work would be required to identify observables which are particularly sensitive to the average size, shape, and fluctuations of the nucleon.
 
@@
 Independently, studies of coherent and incoherent $J/\psi$ production based on a color dipole picture of vector meson production were used to simultaneously constrain both the average color charge density of the proton as well as its event-by-event fluctuations in a saturation based framework \cite{Mantysaari:2016ykx, Mantysaari:2016jaz, Aaron:2009aa, Abramowicz:2015mha}.
-Initial condition studies using the IP-Glasma model of color-glass condensate effective field theory \cite{Schenke:2012wb} simultaneously demonstrated that these color charge fluctuations leave a lasting imprint on the \mbox{small-x} gluon distribution of the proton and hence the initial geometry of QGP energy deposition \cite{Schlichting:2014ipa}.
+Initial condition studies using the IP-Glasma model of Color Glass Condensate effective field theory \cite{Schenke:2012wb} simultaneously demonstrated that these color charge fluctuations leave a lasting imprint on the \mbox{small-x} gluon distribution of the proton and hence the initial geometry of QGP energy deposition \cite{Schlichting:2014ipa}.
 In addition, it was recently shown that hydrodynamic simulations using IP-Glasma initial conditions with color charge fluctuations calibrated to fit coherent and incoherent $J/\psi$ diffraction measured by the H1 and Zeus experiments at HERA \cite{Aaron:2009aa, Abramowicz:2015mha} provide a good description of collectivity in small and large collision systems \cite{Schenke:2018fci}.
@@
 With these considerations in mind, we present progress towards a fully global analysis of $p$-Pb and Pb-Pb bulk observables at $\sqrts=5.02$~TeV using a model calibration framework known as Bayesian parameter estimation.
-We begin, in Sec.~\ref{sec:model}, by constructing a nuclear collision model for $p$-Pb and Pb-Pb collisions using initial conditions with parametric nucleon substructure, and transport dynamics described by a pre-equilibrium free streaming stage, viscous hydrodynamics and microscopic Boltzmann transport.
+We begin in Sec.~\ref{sec:model} by constructing a nuclear collision model for $p$-Pb and Pb-Pb collisions using initial conditions with parametric nucleon substructure and transport dynamics described by a pre-equilibrium free streaming stage, viscous hydrodynamics, and microscopic Boltzmann transport.
 In Sec.~\ref{sec:calibration}, we calibrate free parameters of the model to fit charged-particle yields, mean $p_T$, and anisotropic flow cumulants of \emph{both} collision systems at $\sqrts=5.02$~TeV, and finally, in Secs.~\ref{sec:results} and \ref{sec:summary}, we present posterior results for the model input parameters and comment on the implications for hydrodynamic descriptions of small collision systems.
@@
 We model the initial state of $p$-Pb and Pb-Pb collisions at $\sqrts=5.02$~TeV using the boost-invariant \trento\ model \cite{Moreland:2014oya}.
-Generally speaking, the initial three-dimensional distribution of matter produced in relativistic nuclear collisions is \emph{not} boost-invariant.
-The longitudinal particle density fluctuates both locally point-to-point in the transverse plane as well as globally event-by-event due to asymmetries in the sampled density of participant matter \cite{Ke:2016jrd, Bozek:2010vz}.
+Generally speaking, the initial three-dimensional energy density deposited by relativistic nuclear collisions is \emph{not} boost-invariant.
+The longitudinal energy density fluctuates both locally point-to-point in the transverse plane as well as globally event-by-event due to asymmetries in the sampled density of participant matter \cite{Ke:2016jrd, Bozek:2010vz}.
 Nevertheless, boost-invariance has been shown to be a good approximation for both large and small collision systems when hydrodynamic observables are calculated from particles that are detected close to midrapidity \cite{Shen:2016zpp}.
@@
 The \trento\ model operates in the ultrarelativistic limit with a Lorentz factor $\gamma \gg 1$ such that each nucleus appears as a thin sheet of nuclear density in the laboratory frame.
-The sheets of colliding nuclear density penetrate and pass through each other in time $\Delta t \approx D_\text{nucl} / (\gamma\, \beta_z)$ in the laboratory frame, where $D_\text{nucl}$ is the diameter of the nucleus in its rest frame, $\gamma$ is the usual Lorentz factor of the accelerated ions, and $\beta_z$ is their velocity along the beam axis.
-The resulting nuclear overlap time $\Delta t \lesssim 0.1\ \fmc$ at top RHIC and LHC energies, and thus it is safe to neglect the initial transverse dynamics which occur while the nuclei pass through each other.
+The sheets of colliding nuclear density penetrate and pass through each other in time $\tau_\text{overlap} \approx D_\text{nucl} / (\gamma\, \beta_z)$ in the laboratory frame, where $D_\text{nucl}$ is the diameter of the nucleus in its rest frame, $\gamma$ is the usual Lorentz factor of the accelerated ions, and $\beta_z$ is their velocity along the beam axis.
+The resulting nuclear overlap time $\tau_\text{overlap} \lesssim 0.1\ \fmc$ at top RHIC and LHC energies, and thus it is reasonable to neglect the initial transverse dynamics which occur while the nuclei pass through each other.
 We therefore assume that the collision produces all secondary particles at a uniform proper time $\tau_0 \ll 1~\fmc$, and that it deposits energy at midrapidity which is a function of the locally varying density inside each nucleus.
 
-Consider the collision of two protons, labeled $A$ and $B$, with three-dimensional densities $\rho_A$ and $\rho_B$ in their local rest frames.
-The proton-proton overlap function
+Consider the collision of two nucleons, labeled $A$ and $B$, with three-dimensional densities $\rho_A$ and $\rho_B$ in their local rest frames.
+The nucleon-nucleon overlap function
 \begin{equation}
   \label{eq:tpp}
-  T_{pp}(b) = \int dx\, dy \int dz_1\, \rho_A(x, y, z_1) \int dz_2\, \rho_B(x - b, y, z_2),
-\end{equation}
-describes the eikonal overlap of the two proton wave packets at fixed impact parameter $b$, where coordinate $(x, y)$ lies in the transverse plane and $z_{1,2}$ is parallel to the beam axis.
-Here we assume that each proton is comprised of smaller constituents---e.g.\ valence quarks, sea quarks, and small-x gluons---which may collide to produce secondary particles and contribute to the observed inelastic proton-proton cross section.
-
-Within a picture of independent pairwise collisions between the constituents, a Glauber model model may be used to calculate the probability $P_\mathrm{coll}$ that the two protons collide inelastically at impact parameter $b$.
+  T_{nn}(\bv) = \int d^2x_\perp \int dz_1\, \rho_A(\xv_\perp, z_1) \int dz_2\, \rho_B(\xv_\perp - \,\bv, z_2)
+\end{equation}
+describes the eikonal overlap of the two nucleons at impact parameter $\bv$ in the transverse plane $\xv_\perp$, orthogonal to the beam axis coordinates $z_{1,2}$.
+Here we assume that each nucleon is comprised of smaller constituents---e.g.\ valence quarks, sea quarks, and small-x gluons---which may collide to produce secondary particles and contribute to the observed inelastic nucleon-nucleon cross section.
+
+Within a picture of independent pairwise collisions between the constituents, a Glauber model model may be used to calculate the probability $P_\mathrm{coll}$ that the two nucleons collide inelastically at impact parameter $\bv$.
 In the limit when the number of constituents is large, it yields the particularly simple form
@@
   \label{eq:pcoll}
-  P_\mathrm{coll}(b) = 1 - \exp[-\sigma_\mathrm{eff}\, T_{pp}(b)], \\
-\end{equation}
-where $\sigma_\mathrm{eff}$ is an effective cross section for pairwise inelastic collisions between the constituents, and $\sigma_{pp}^\mathrm{inel}$ is the total inelastic proton-proton cross section
+  P_\mathrm{coll}(\bv) = 1 - \exp[-\sigma_\mathrm{eff}\, T_{nn}(\bv)],
+\end{equation}
+where $\sigma_\mathrm{eff}$ is the effective cross section for inelastic collisions between each pair of constituents, determined by fitting the inelastic proton-proton cross section
 \begin{equation}
   \label{eq:sigma_nn}
-  \sigma_{pp}^\mathrm{inel} = \int 2 \pi b\, db\, P_\mathrm{coll}(b).
-\end{equation}
-
-The proton densities $\rho_{A,B}$ in Eq.~\eqref{eq:tpp} are commonly modeled using a spherically symmetric distribution.
-For instance, the original implementation of the \trento\ model uses Gaussian protons, largely because it yields a simple analytic solution to Eq.~\eqref{eq:pcoll}.
-Needless to say, such approximations are admittedly crude and may have a significant effect on the dynamics of small collision systems where the proton size is comparable to the size of the produced QGP medium.
-A number of previous studies have investigated the effects of deformed or ``lumpy'' protons.
+  \sigma_{pp}^\mathrm{inel} = \int d^2b\, P_\mathrm{coll}(\bv)
+\end{equation}
+at the specified collision energy $\sqrts$.
+
+The nucleon densities $\rho_{A,B}$ in Eq.~\eqref{eq:tpp} are commonly modeled using a spherically symmetric distribution.
+For instance, the original implementation of the \trento\ model uses Gaussian nucleons, largely because it yields a simple analytic solution to Eq.~\eqref{eq:pcoll}.
+Needless to say, such approximations are admittedly crude and may have a significant effect on the dynamics of small collision systems where the nucleon size is comparable to the size of the produced QGP medium.
+
+A number of previous studies have investigated the effects of deformed or ``lumpy'' nucleons.
 One common implementation is a superposition of three valence quarks, typically described by Gaussian or exponential form factors \cite{Welsh:2016siu, Bozek:2017jog, Schenke:2014zha, Schlichting:2014ipa, Adare:2015bua, Broniowski:2016pvx}.
-The corresponding proton density is then assumed to be that of predominantly small-x gluons, seeded by the distribution of color charge in each of the three valence quarks.
+The corresponding nucleon density is then assumed to be that of predominantly small-x gluons, seeded by the distribution of color charge in each of the three valence quarks.
 
@@
   \begin{tikzpicture}
-    % spherical proton
+    % spherical nucleon
     \draw[dashed, xshift=-2.5cm] (0,0) circle (1cm);
@@
   \end{tikzpicture}
-  \caption{Schematic of plausible proton shapes.
-The sketch on the left shows a spherically symmetric proton (dashed line), while the middle and right illustrations depict a fluctuating proton with three and nine constituents respectively (solid lines).}
+  \caption{
     \label{fig:substructure}
+    Schematic of plausible nucleon shapes.
+    The sketch on the left shows a spherically symmetric nucleon (dashed line), while the middle and right illustrations depict a fluctuating nucleon with three and nine constituents respectively (solid lines).
+  }
 \end{figure}
 
-In this work, we pursue a less restrictive and more parametric description of the proton where the number of substructure degrees of freedom is uncertain as depicted in Fig.~\ref{fig:substructure}.
-We model each nucleon's three-dimensional density $\rho_{A,B}$ as a sum of $n_c$ independent constituents,
+In this work, we pursue a less restrictive and more parametric description of the nucleon where the number of substructure degrees of freedom is uncertain as depicted in Fig.~\ref{fig:substructure}.
+We model each nucleon's three-dimensional density $\rho_{A,B}$ as a sum of $n_c$ independent constituents
 \begin{equation}
@@
 \end{equation}
-where each constituent density $\rho_c$ is described by a Gaussian distribution,
-\begin{equation}
-  \rho_c(\xv) = \frac{1}{(2 \pi v^2)^{3/2}} \exp\left(- \frac{|\xv|^2}{2 v^2}\right).
+where each constituent density $\rho_c$ is described by a Gaussian distribution
+\begin{equation}
+  \rho_c(\xv) = \frac{1}{(2 \pi v^2)^{3/2}} \exp\left(- \frac{|\xv|^2}{2 v^2}\right)
 \end{equation}
 of variable width $v$.
-The constituent positions $\xv_i$ are sampled independently (ignoring correlations) from a Gaussian radial distribution,
+The constituent positions $\xv_i$ are sampled independently (ignoring correlations) from a Gaussian radial distribution
 \begin{equation}
@@
 \end{equation}
-where $\xv_{A,B}$ is the position of each proton, and $r$ is a parameter which varies the sampling radius of the constituent positions about each proton position.
-As a matter of convenience, we sample the proton positions \emph{before} determining the constituent positions.
-This leads to small discrepancies between the designated proton positions $\xv_{A,B}$ and the actual position of each proton's center-of-mass, owing to fluctuations in the constituent positions $\xv_i$.
-The sampling radius $r$ should thus be interpreted with care.
-It is \emph{not} the Gaussian width of the sampled nucleons in their center-of-mass frame.
-
-The two protons $A$, $B$ are assigned a random impact parameter $\bv$, and Eq.~\eqref{eq:pcoll} is used to sample their inelastic collision probability $P_\mathrm{coll}(b)$.
-Note, this proton-proton inelastic collision probability has no direct knowledge of the individual constituent positions; it is only indirectly sensitive to their positions through the density $\rho_{A, B}$.
-The protons in this work are described by a \emph{single} density field; constituents are simply used as a mechanism to generate deformity and fluctuations.
-This is an important distinction between the present model and a similar nucleon substructure implementation known as the participant or ``wounded'' quark model which allows for a subset of quarks (constituents) to participate inside a single nucleon \cite{ANISOVICH1978477, Broniowski:2016pvx}.
-The proton, unlike the nucleus, cannot produce semi-stable spectator fragments in a high-energy collision.
-Any spectator quarks produced by a wounded quark model would be colored objects that necessarily contribute to secondary particle production as they fragment and recombine to form colorless hadrons.
+where $r$ is a parameter that varies the sampling radius of the constituent positions $\xv_i$ about the nucleon position $\xv_{A,B}$.
+As a matter of convenience, we sample the nucleon positions \emph{before} determining the constituent positions.
+This leads to small discrepancies between the designated nucleon positions $\xv_{A,B}$ and the actual position of each nucleon's center-of-mass, owing to fluctuations in the constituent positions $\xv_i$.
+The parameter $r$ should thus be interpreted with care.
+It is a computational sampling radius, \emph{not} the Gaussian width of the sampled nucleons in their center-of-mass frame.
+
+The two nucleons $A$, $B$ are assigned a random impact parameter offset $\bv$, and Eq.~\eqref{eq:pcoll} is used to sample their inelastic collision probability $P_\mathrm{coll}(\bv)$.
+Note, this collision probability has no direct knowledge of the individual constituent positions; it is only \emph{indirectly} sensitive to the constituent positions via their effect on the nucleon densities $\rho_{A, B}$.
+The constituents are merely used as a mechanism to deform and fluctuate each nucleon profile.
+
+This is an important distinction between the present model and a similar nucleon substructure implementation known as the participant quark model, which allows for a subset of quarks (constituents) to participate inside a single nucleon \cite{ANISOVICH1978477, Broniowski:2016pvx}.
+The nucleon, unlike the nucleus, cannot produce semi-stable spectator fragments in a high-energy collision.
+Any spectator quarks produced by a wounded quark model would be colored objects that necessarily contribute to secondary particle production.
 We correspondingly require that the nucleons in Eq.~\eqref{eq:rho} participate as singular objects, such that all spectator matter discarded by the simulation is appropriately color-neutral and inert.
@@
 Hereafter, we switch to Milne coordinates $(\xv_\perp, \eta_s, \tau)$, where $\xv_\perp$ is a Cartesian coordinate $(x, y)$ in the transverse plane, $\eta_s = \tfrac{1}{2} \log[(t+z)/(t-z)]$ is the system's spacetime rapidity, and $\tau = \sqrt{t^2 - z^2}$ its proper time.
-Assuming the protons collide at the sampled impact parameter $\bv$, we assign each proton a participant thickness
+Assuming the nucleons collide at the sampled impact parameter $\bv$, we assign each nucleon a participant thickness
 \begin{equation}
@@
 \end{equation}
-which projects its \emph{fluctuated} proton density,
+which projects its \emph{fluctuated} nucleon density
 \begin{equation}
@@
 onto the transverse plane $\xv_\perp$ centered at midrapidity.
-This fluctuated proton density $\tilde{\rho}_{A,B}$ equals the proton density $\rho_{A,B}(\xv)$ in Eq.~\eqref{eq:rho} with additional weights $\gamma_i$ sampled from a gamma distribution with unit mean and variance $1/k$.
+This fluctuated nucleon density $\tilde{\rho}_{A,B}$ equals the nucleon density $\rho_{A,B}$ in Eq.~\eqref{eq:rho} with additional weights $\gamma_i$ sampled from a gamma distribution with unit mean and variance $1/k$.
 Such \emph{ad hoc} random weights are necessary to describe the large fluctuations observed in minimum bias proton-proton collisions, although their exact physical origin is not well understood.
@@
 This means that the natural range for the constituent fluctuations is larger when the number of constituents is larger and \emph{vice versa}.
-We therefore reparametrize the constituent fluctuations using,
+We therefore reparametrize the constituent fluctuations using
 \begin{equation}
@@
 
-The energy density $e$ deposited at midrapidity $\eta_s=0$ and proper time $\tau_0 \ll 1~\fmc$ is then some function,
-\begin{equation}
-  e(\xv_\perp, \eta_s=0, \tau_0) = f(\T_A, \T_B),
+The energy density $e$ deposited at midrapidity $\eta_s=0$ and proper time $\tau_0 \ll 1~\fmc$ is then some function
+\begin{equation}
+  e(\xv_\perp, \eta_s=0, \tau_0) = f(\T_A, \T_B)
 \end{equation}
 of the local participant thicknesses $\T_A$ and $\T_B$.
-A natural first guess for this mapping is an arithmetic mean,
+A natural first guess for this mapping is an arithmetic mean
 \begin{equation}
@@
 This wounded nucleon scaling was in fact one of the first such mappings used as a proxy for initial particle production in relativistic heavy-ion collisions \cite{Bialas:1976ed}.
-It was subsequently realized, however, that the wounded nucleon model predicts the wrong scaling for charged-particle production as a function of collision centrality \cite{Kharzeev:2000ph} and hence the wrong scaling for midrapidity energy deposition\footnote{Typically, wounded nucleon scaling is used to parametrize the entropy density $s$, but the shortcomings of the model are nevertheless the same when parametrizing the system's energy density $e$.
+It was subsequently realized, however, that the wounded nucleon model predicts the wrong scaling for charged-particle production as a function of collision centrality \cite{Kharzeev:2000ph} and hence the wrong scaling for midrapidity energy deposition
+as a function of participant density.\footnote{Typically, wounded nucleon scaling is used to parametrize the entropy density $s$, but the shortcomings of the model are nevertheless the same when parametrizing the system's energy density $e$.
 Both parametrizations underpredict the steep rise of particle production observed in central collisions.}
-as a function of participant density.
 
@@
   \label{eq:gmean}
-  e(\xv_\perp, \eta_s=0, \tau_0) = \text{const} \times \bigg( \frac{\T_A^p + \T_B^p}{2} \bigg)^{1/p},
-\end{equation}
-based on a family of functions known as the generalized mean(s).
+  e(\xv_\perp, \eta_s=0, \tau_0) = \text{const} \times M_p(\T_A, \T_B),
+\end{equation}
+where $M_p$ is a family of functions known as the generalized means
+\begin{equation}
+  M_p(x, y) = \left( \frac{x^p + y^p}{2} \right)^{1/p}.
+\end{equation}
 This parametrization introduces a dimensionless parameter $p$ which varies the scaling behavior of initial energy deposition at midrapidity.
@@
   \setlength{\extraspace}{0.5ex}
-  e \propto
+  M_p(x, y) =
   \begin{cases}
-    \max(\T_A, \T_B) & p \rightarrow +\infty, \\[\extraspace]
-    (\T_A + \T_B)/2 & p = +1, \hfill \text{ (arithmetic)} \\[\extraspace]
-    \sqrt{\T_A \T_B} & p = 0, \hfill \text{ (geometric)} \\[\extraspace]
-    2\, \T_A \T_B/(\T_A + \T_B) & p = -1, \hfill \text{ (harmonic)} \\[\extraspace]
-    \min(\T_A, \T_B) & p \rightarrow -\infty.
+    \max(x, y) & p \rightarrow +\infty, \\[\extraspace]
+    (x + y)/2 & p = +1, \hfill \text{ (arithmetic)} \\[\extraspace]
+    \sqrt{x y} & p = 0, \hfill \text{ (geometric)} \\[\extraspace]
+    2\, x y/(x + y) & p = -1, \hfill \text{ (harmonic)} \\[\extraspace]
+    \min(x, y) & p \rightarrow -\infty.
   \end{cases}
@@
 
-We note that previous applications of the \trento\ model \cite{Moreland:2014oya, Bernhard:2016tnd} used the generalized mean ansatz to parametrize the \emph{entropy} density $s(\xv_\perp, \eta_s=0, \tau_\text{hydro})$, motivated (in part) by the approximate linear relation at midrapidity between initial entropy density $dS/d\eta_s$ and final charged-particle density $d\nch/d\eta$ \cite{Shen:2015qta}.
-These studies, however, assumed instant thermalization, an approximation which sets $u^\mu$, $\Pi$, and $\pi^{\mu\nu}$ identically to zero at the hydrodynamic starting time $\tau_\text{hydro}$.
-The present study, meanwhile, uses a pre-equilibrium free streaming stage to estimate $u^\mu$, $\Pi$, and $\pi^{\mu\nu}$ as a function of a variable free streaming time $\taufs$, at which point the system is matched to hydrodynamics.
-Strictly speaking, the system's entropy is ill defined prior to thermalization while the system free streams.
-We therefore reinterpret the quantity produced by the \trento\ model as an energy density $e(\xv_\perp, \eta_s=0, \tau_0)$ and free stream this quantity to the hydrodynamic starting time $\tau_\text{hydro} \equiv \taufs$.
-This assumes, of course, that the model's parametrization for entropy density is also suitable as a parametrization for the energy density.
-To this point, we note that the arguments for the generalized mean ansatz are quite general \cite{Moreland:2014oya}, and that the energy density $dE/d\eta_s$ and entropy density $dS/d\eta_s$ share many similar qualities.
-For example, both are approximately conserved at midrapidity, and both scale approximately linearly with the final charged-particle density $d\nch/d\eta$.
-We thus expect the parametrization to be suitable for either quantity.
+Note, the form of Eq.~\eqref{eq:gmean} differs somewhat from our previous studies \cite{Moreland:2014oya, Bernhard:2016tnd}, which parametrized the system's \emph{entropy} density using the generalized mean ansatz:
+\begin{equation}
+  s(\xv_\perp, \eta_s=0, \tau_0) \propto M_p(\T_A, \T_B).
+\end{equation}
+Our motivation for reinterpretting the left-side of this equation as an energy density in the present work is to enable the application of pre-equilibrium free streaming equations (described shortly) which require the system's initial energy density as input.
+This modification is permissible since the \trento\ model is based on particle yield observations coupled with the approximate scaling relation $d\nch/d\eta \propto M_p(\T_A, \T_B)$, which remains valid for both static entropy density and free streamed energy density initialization.
 
@@
 
-Up to this point we have restricted our attention to a single proton-proton collision.
+Up to this point we have restricted our attention to a single nucleon-nucleon collision.
 Equation~\eqref{eq:gmean} is a purely local function of nuclear density in the transverse plane and hence it should (in principle) be equally valid for any pair of colliding nuclei at sufficiently high beam energy.
-The \trento\ model readily generalizes from individual proton-proton collisions to arbitrary nucleus-nucleus collisions by summing the participant thicknesses $\T_{A,B}$ in Eq.~\eqref{eq:fluctuated_thick} over all nucleons which participate in one or more inelastic collisions.
+The \trento\ model readily generalizes from individual nucleon-nucleon collisions to arbitrary nucleus-nucleus collisions by summing the participant thicknesses $\T_{A,B}$ in Eq.~\eqref{eq:fluctuated_thick} over all nucleons which participate in one or more inelastic collisions.
 The only modeling difference between $p$-$p$, $p$-Pb, and Pb-Pb collisions is the number and the position of the nucleons.
@@
 We then sample the solid angles of each nucleon one-by-one, starting with the nucleon closest to the center of the nucleus and working our way outwards.
-If a sampled nucleon position is too close to any of its previously placed neighbors, its solid angle is resampled until the minimum distance criteria is satisfied.
+If a sampled nucleon position is too close to any of its previously placed neighbors, its solid angle is resampled---but not its radial coordinate---until the minimum distance criteria is satisfied.
 Similar methods could be used to model correlations between individual constituents inside each nucleon, although the implementation would be somewhat tedious.
@@
 We therefore choose to model the QGP's initial off-equilibrium dynamics using a simple step-function approximation, depicted in Fig.~\ref{fig:coupling}, which free streams the initial state for proper time $\taufs$ (zero coupling) before instantaneously switching to viscous hydrodynamics (strong coupling) \cite{Liu:2015nwa, Broniowski:2008qk}.
-The free parameter $\taufs$ allows us to parametrically vary the \emph{time averaged} coupling strength in the approximate window $0 < \taufs \lesssim1\ \fmc$.
+The free parameter $\taufs$ allows us to parametrically vary the \emph{time averaged} coupling strength during the initial stage of the collision.
 
@@
 Equation~\eqref{freestream}, however, simplifies for a boost-invariant gas of massless noninteracting partons with locally isotropic transverse momentum distributions.
-Subject to these assumptions, the energy-momentum tensor $T^{\mu\nu}(\xv_\perp, \eta_s=0, \tau)$ at transverse coordinate $\xv_\perp$ and time $\tau > \tau_0$ equals \cite{Broniowski:2008qk, Liu:2015nwa},
+Subject to these assumptions, the energy-momentum tensor $T^{\mu\nu}(\xv_\perp, \eta_s=0, \tau)$ at transverse coordinate $\xv_\perp$ and time $\tau > \tau_0$ equals \cite{Broniowski:2008qk, Liu:2015nwa}
 \begin{multline}
@@
   T^{\mu\nu}(\xv_\perp, \eta_s=0, \tau) =\\
-  \frac{\tau_0}{\tau} \int_{-\pi}^{\pi} d\phi_q\, \hat{q}^\mu \hat{q}^\nu e(\xv_\perp - (\tau - \tau_0)\hat{\mathbf{q}}_\perp, \eta_s=0, \tau_0),
+  \frac{\tau_0}{\tau} \int_{-\pi}^{\pi} d\phi_p\, \hat{p}^\mu \hat{p}^\nu e(\xv_\perp - (\tau - \tau_0)\hat{\mathbf{p}}_\perp, \eta_s=0, \tau_0),
 \end{multline}
-where $\hat{q}^\mu = (1, \cos \phi_q, \sin \phi_q, 0)$ and $\hat{\mathbf{q}}_\perp = (\cos \phi_q, \sin \phi_q)$.
-Here we assume that $\tau_0 \ll 1~\fmc$ is small and define $\taufs \equiv \tau - \tau_0 \approx \tau$.
+where $\hat{p}^\mu = (1, \cos \phi_q, \sin \phi_q, 0)$ and $\hat{\mathbf{p}}_\perp = (\cos \phi_q, \sin \phi_q)$ are momentum unit vectors.
+Here we assume that the time $\tau_0 \ll 1~\fmc$ is small and define $\taufs \equiv \tau - \tau_0 \approx \tau$.
 We also combine the unknown constant in Eq.~\eqref{eq:gmean} with the pre-factor $\tau_0$ in Eq.~\eqref{energy-momentum} to produce a single normalization factor $\text{Norm} = \text{const} \times \tau_0$ with units of energy.
-The solution \eqref{energy-momentum} is then decomposed in hydrodynamic form,
+The solution \eqref{energy-momentum} is then decomposed in hydrodynamic form
 \begin{equation}
@@
 where $e$ and $P$ are the energy density and pressure in the local fluid rest frame, $u^\mu$ is the local fluid velocity, ${\Delta^{\mu\nu} = g^{\mu\nu} - u^\mu u^\nu}$ is the projector onto the space orthogonal to $u^\mu$, and $\Pi$ and $\pi^{\mu\nu}$ are the bulk pressure and shear stress tensor respectively.
-We then solve for the energy density $e$ and fluid velocity $u^\mu$ using the Landau matching condition which defines the fluid rest frame velocity as the time-like eigenvector of $T^{\mu\nu}$ with energy density $e$ as its eigenvalue,
+We then solve for the energy density $e$ and fluid velocity $u^\mu$ using the Landau matching condition which defines the fluid rest frame velocity as the time-like eigenvector of $T^{\mu\nu}$ with energy density $e$ as its eigenvalue
 \begin{equation}
@@
 This procedure provides initial values for $T^{\mu\nu}$, $u^\mu$, $\Pi$, and $\pi^{\mu\nu}$ which conserve energy and are consistent with the underlying hydrodynamic equation of state.
-We therefore expect it to provide a more realistic description of the initial stages of the collision as compared to a previous study using the \trento\ initial condition model which set $\Pi$, $\pi^{\mu\nu}$ and $u^\mu$ initially to zero \cite{Bernhard:2016tnd}.
+We therefore expect it to provide a more realistic description of the initial stages of the collision as compared to a previous study using the \trento\ initial condition model which set $\Pi$, $\pi^{\mu\nu}$, and $u^\mu$ initially to zero \cite{Bernhard:2016tnd}.
 
@@
 where $\smin$, $\sslope$, and $\scrv$ are tunable parameters, and $T_c=0.154$~GeV is the pseudocritical transition temperature of the HotQCD EoS.
-While for the specific bulk viscosity $\zeta/s$, we use an unnormalized Cauchy distribution
+Meanwhile, for the specific bulk viscosity $\zeta/s$, we use an unnormalized Cauchy distribution
 \begin{equation}
@@
 described by a tunable maximum $\bmax$, temperature width $\bwidth$, and temperature location parameter $\bloc$.
-Figure~\ref{fig:viscosity_dof} shows several of the possible curves parametrized by Eqs.~\eqref{eq:shear_viscosity} and \eqref{eq:bulk_viscosity}, although many more are possible.
+Figure~\ref{fig:viscosity_dof} shows several of the possible curves parametrized by Eqs.~\eqref{eq:shear_viscosity} and \eqref{eq:bulk_viscosity}, although infinitely more are possible.
 
@@
 \end{equation}
-Here $m_0$ is the resonance's Breit-Wigner mass, and $\Gamma(m)$ is its mass-dependent width, for which we use a simple form:
+Here $m_0$ is the resonance's Breit-Wigner mass, and $\Gamma(m)$ is its mass-dependent width, for which we use the simple form
 \begin{equation}
@@
 \begin{equation}
-  \frac{df_i(x, p)}{dt} = \mathcal{C}_i(x, p)
+  \frac{df_i(x, p)}{dt} = \mathcal{C}_i(x, p),
 \end{equation}
 where $f_i$ is the distribution function for species $i$, and $\mathcal{C}_i$ is its microscopic collision kernel.
-The model propagates all produced hadrons along classical trajectories, and simulates their scatterings, resonance formations and decays until the last interactions cease.
+The model propagates all produced hadrons along classical trajectories and simulates their scatterings, resonance formations, and decays until the last interactions cease.
 
@@
 The nuclear collision model constructed in Sec.~\ref{sec:model} includes a number of free parameters $\xv$ which describe the initial state, pre-equilibrium dynamics, and hydrodynamic medium.
-Given values for the parameters $\xv$, the model may be used to predict a vector of simulated observables $\yv$.
-For example, $\yv$ might be a vector consisting of charged-particle yields in different centrality bins.
-The physics model thus represents a vector-valued function $f(\xv) = \yv$ which maps the parameter values $\xv$ to the model observables $\yv$.
-
-The goal of this work is to estimate the true model parameters $\xv_\star$ provided some evidence that our model predictions $\yv$ describe experimental measurements $\yv_e$.
+Given values for the parameters $\xv$, the model may be used to predict a vector of simulated observables $\ym$.
+For example, $\ym$ might be a vector consisting of charged-particle yields in different centrality bins.
+The physics model thus represents a vector-valued function $\ym = f(\xv)$ which maps the parameter values $\xv$ to the model observables $\ym$.
+
+The goal of this work is to estimate the true model parameters $\xv_\star$ provided some evidence that our model predictions $\ym$ describe experimental measurements $\ye$.
 The problem involves three distinct components:
@@
   \item $H_\xv$: the hypothesis that the model parameters $\xv$ are the true model parameters $\xv_\star$ of $f$, and
-  \item $E$: the evidence provided by the experimental data $\yv_e$ and its corresponding uncertainties.
+  \item $E$: the evidence provided by the experimental data $\ye$ and its corresponding uncertainties.
 \end{enumerate}
@@
 
-Subject to this assumption, we can apply Bayes' theorem to evaluate the hypothesis $H_\xv$ for the true model parameters,
+Subject to this assumption, we can apply Bayes' theorem to evaluate the hypothesis $H_\xv$, \emph{given} the evidence provided by $E$.
+Simplifying notation and writing $H_\xv$ as just $\xv$, Bayes' theorem yields
 \begin{equation}
   \label{eq:bayes}
-  P(H_\xv | E) \propto P(E | H_\xv)\, P(H_\xv).
-\end{equation}
-The left-side of this expression is the \emph{posterior}: the probability for hypothesis $H_\xv$ given the experimental evidence $E$.
+  P(\xv | E) \propto P(E | \xv)\, P(\xv).
+\end{equation}
+The left-side of this expression is the \emph{posterior}: the probability of $\xv = \xv_\star$ given the experimental evidence $E$.
 On the right-side there are two separate terms.
-The first term $P(E | H_\xv)$ is the \emph{likelihood} function: the probability of observing the experimental evidence $E$ given our model and the hypothesis $H_\xv$ for the true model parameters $\xv_\star$, and the second term $P(H_\xv)$ is the \emph{prior}: an estimate of the probability of hypothesis $H_\xv$ in the absence of evidence $E$.
+The first term $P(E | \xv)$ is the \emph{likelihood} function: the probability of observing the evidence $E$ provided that $\xv = \xv_\star$, and the second term $P(\xv)$ is the \emph{prior}: an estimate of the probability of hypothesis $\xv = \xv_\star$ in the absence of evidence $E$.
 
@@
   \label{eq:likelihood}
-  P(E | H_\xv) = \frac{1}{\sqrt{(2\pi)^m \det \Sigma}} \exp \left ( -\frac{1}{2}\Delta\yv^\intercal \Sigma^{-1} \Delta\yv \right ),
-\end{equation}
-where $\Delta\yv = \yv(\xv) - \yv_e$ is the discrepancy of the model and experiment, and $\Sigma = \Sigma_m(\xv) + \Sigma_e$ is the \emph{total} covariance matrix, equal to the sum of a modeling component $\Sigma_m(\xv)$ and an experimental component $\Sigma_e$ which account for all known sources of uncertainty in the simulated and measured observables.
+  P(E | \xv) = \frac{1}{\sqrt{(2\pi)^m \det \Sigma}} \exp \left ( -\frac{1}{2}\Delta\yv^\intercal \Sigma^{-1} \Delta\yv \right ),
+\end{equation}
+where $\Delta\yv = \ym(\xv) - \ye$ is a vector of size $m$, equal to the discrepancy of the model and experiment, and $\Sigma = \Sigmam(\xv) + \Sigmae$ is the \emph{total} covariance matrix, equal to the sum of a modeling component $\Sigmam(\xv)$ and an experimental component $\Sigmae$ which account for all known sources of uncertainty in the simulated and measured observables.
 
@@
 
-For the prior $P(H_\xv)$, we specify ranges, i.e.\ minimum and maximum values, for each parameter which are listed in Table~\ref{tab:design}.
+For the prior $P(\xv)$, we specify ranges, i.e.\ minimum and maximum values, for each parameter which are listed in Table~\ref{tab:design}.
 We assume the prior distribution is constant and nonzero within each specified range and zero otherwise.
@@
 
-The likelihood function \eqref{eq:likelihood} provides evidence for \mbox{(or against)} the model parameters $\xv$ by comparing the model predictions $\yv(\xv)$ to experimental data $\yv_e$.
+The likelihood function \eqref{eq:likelihood} provides evidence for (or against) the model parameters $\xv$ by comparing the model predictions $\ym$ to experimental data $\ye$.
 We focus on simple experimental observables in the present study which are sensitive to the bulk properties of the produced medium.
@@
 These high-multiplicity bins are too selective for our modest minimum bias event sample, and so a different procedure is required.
-We exploit, for this purpose, the approximate monotonic relation between each event's initial transverse energy density,
-\begin{equation}
-  \frac{dE_T}{d\eta_s} \bigg\vert_{\eta_s=0} = \tau_0 \int d^2x_\perp e(\xv_\perp, \eta_s=0, \tau_0),
+We exploit, for this purpose, the approximate monotonic relation between each event's initial energy density
+\begin{equation}
+  \frac{dE}{d\eta_s} \bigg\vert_{\eta_s=0} = \tau_0 \int d^2x_\perp e(\xv_\perp, \eta_s=0, \tau_0)
 \end{equation}
@@
 Let $\langle\nch\rangle$ denote the average charged-particle multiplicity of these events.
-We first rescale the experimental multiplicity bin edges,
-\begin{equation}
-  [\nch^\text{low}, \nch^\text{high}] \rightarrow \left[\frac{\nch^\text{low}}{\langle \nch \rangle},\, \frac{\nch^\text{high}}{\langle \nch \rangle}\right ],
+We first rescale the experimental multiplicity bin edges
+\begin{equation}
+  [\nch^\text{low}, \nch^\text{high}] \rightarrow \left[\frac{\nch^\text{low}}{\langle \nch \rangle},\, \frac{\nch^\text{high}}{\langle \nch \rangle}\right ]
 \end{equation}
 in order to reexpress each bin edge as a unitless variable.
-These bin edges are then associated with a pair of transverse energy bin edges,
+These bin edges are then associated with a pair of energy bin edges
 \begin{equation}
   \label{eq:mult_trigger}
-  \left [\frac{E_T^\text{min}}{\langle E_T \rangle}, \frac{E_T^\text{max}}{\langle E_T \rangle} \right] \leftrightarrow \left [\frac{\nch^\text{low}}{\langle \nch \rangle}, \frac{\nch^\text{high}}{\langle \nch \rangle} \right ],
-\end{equation}
-where $E_T \propto (dE_T/d\eta_s) \vert_{\eta_s=0}$ is the midrapidity transverse energy of a single event in the desired kinematic range, and $\langle E_T \rangle$ is the corresponding average transverse energy over the full minimum bias event sample.
+  \Bigg [\frac{E^\text{min}}{\langle E \rangle}, \frac{E^\text{max}}{\langle E \rangle} \Bigg ] \leftrightarrow \Bigg [\frac{\nch^\text{low}}{\langle \nch \rangle}, \frac{\nch^\text{high}}{\langle \nch \rangle} \Bigg ],
+\end{equation}
+where $E \propto (dE/d\eta_s) \vert_{\eta_s=0}$ is the midrapidity energy of a single event in the desired kinematic range, and $\langle E \rangle$ is the corresponding average energy over the full minimum bias event sample.
 
@@
 \begin{equation}
-  \Sigma = \Sigma_m + \Sigma_e
-\end{equation}
-in Eq.~\eqref{eq:likelihood}, which includes uncertainty contributions from both the model $\Sigma_m$ and experimental data $\Sigma_e$.
-The experimental contribution to the covariance $\Sigma_e$ is further broken down into statistical and systematic components,
-\begin{equation}
-  \Sigma_e = \Sigma_e^\text{stat} + \Sigma_e^\text{sys}.
-\end{equation}
-The statistical errors in $\Sigma_e^\text{stat}$ are uncorrelated, and thus its covariance matrix is diagonal:
-\begin{equation}
-  \Sigma_e^\text{stat} = \diag[(\sigma^\text{stat}_1)^2, (\sigma^\text{stat}_2)^2, \dots (\sigma^\text{stat}_m)^2 ],
-\end{equation}
-where $\sigma^\text{stat}_i$ is the statistical uncertainty of observable $y_i$ in the experimental observable vector $\yv_e = (y_1, \dots, y_m)$.
+  \Sigma = \Sigmam + \Sigmae
+\end{equation}
+appearing in Eq.~\eqref{eq:likelihood}, which includes uncertainty contributions from both the model $\Sigmam$ and experimental data $\Sigmae$.
+The experimental contribution to the covariance $\Sigmae$ is further broken down into statistical and systematic components:
+\begin{equation}
+  \Sigmae = \Sigmae^\text{stat} + \Sigmae^\text{sys}.
+\end{equation}
+The statistical errors in $\Sigmae^\text{stat}$ are uncorrelated, and thus its covariance matrix is diagonal,
+\begin{equation}
+  \Sigmae^\text{stat} = \diag[(\sigma^\text{stat}_1)^2, (\sigma^\text{stat}_2)^2, \dots (\sigma^\text{stat}_m)^2 ],
+\end{equation}
+where $\sigma^\text{stat}_i$ is the statistical uncertainty of observable $y_i$ in the observable vector $\ye = (y_1, \dots, y_m)$.
 The systematic errors, meanwhile, are typically correlated, but the correlation structure is not reported by the experiments so we assert a reasonable form.
-We can expand the systematic covariance matrix as
-\begin{equation}
-  (\Sigma_e^\text{sys})_{ij} =  \rho_{ij} \sigma_i \sigma_j,
-\end{equation}
-where $\sigma_i$ and $\sigma_j$ are the systematic errors of observables $y_i$ and $y_j$ respectively, and $\rho_{ij}$ is the Pearson correlation coefficient between observable $y_i$ and $y_j$:
+We can expand the systematic covariance as
+\begin{equation}
+  (\Sigmae^\text{sys})_{ij} =  \rho_{ij} \sigma_i \sigma_j,
+\end{equation}
+where $\sigma_i$ and $\sigma_j$ are the systematic errors of observables $y_i$ and $y_j$ respectively, and $\rho_{ij}$ is the Pearson correlation coefficient
 \begin{equation}
@@
 \end{equation}
-which satisfies $\rho_{ij}=1$ for $i=j$ and $|\rho_{ij}| \le 1$ for $i \ne j$.
+between observable $y_i$ and $y_j$, which satisfies $\rho_{ij}=1$ for $i=j$ and $|\rho_{ij}| \le 1$ for $i \ne j$.
 We assume that each observable is correlated across different centrality/multiplicity bins, and uncorrelated with observables of a different type, e.g.\ correlations between yields and flows.
@@
 
-For the correlation structure between different observable bins, we assert a simple Gaussian form
+For the correlation structure between different observable bins, we assert the simple Gaussian form
 \begin{equation}
   \label{eq:corr}
-  \rho_{ij}^\text{sys} = \exp \left[ -\frac{1}{2} \left(\frac{b_i - b_j}{l} \right)^2 \right],
-\end{equation}
-where $b_i$ and $b_j$ are the midpoints of two observable bins of a single type (centrality or relative multiplicity), and $l$ is a correlation length which describes how quickly the observable bins decorrelate as the distance between the bins increases.
-We use centrality correlation lengths $l=100$ for all of the centrality binned Pb-Pb observables and $l=30$ for the centrality binned $p$-Pb charged-particle yield $d\nch/d\eta$.
-The \mbox{$p$-Pb} mean $p_T$ and flow observables, meanwhile, use relative multiplicity bins $\nch / \langle \nch \rangle$ and $\ntrk / \langle \ntrk \rangle$ which necessitate a smaller correlation length $l=5$.
+  \rho_{ij}^\text{sys} = \exp \left[ -\frac{1}{2} \left(\frac{b_i - b_j}{\ell} \right)^2 \right],
+\end{equation}
+where $b_i$ and $b_j$ are the midpoints of two observable bins of a single type (centrality or relative multiplicity), and $\ell$ is a correlation length which describes how quickly the observable bins decorrelate as the distance between the bins increases.
+We use centrality correlation lengths ${\ell=100}$ for all of the centrality binned Pb-Pb observables and $\ell=30$ for the centrality binned $p$-Pb charged-particle yield $d\nch/d\eta$.
+The \mbox{$p$-Pb} mean $p_T$ and flow observables, meanwhile, use relative multiplicity bins $\nch / \langle \nch \rangle$ and $\ntrk / \langle \ntrk \rangle$ which necessitate a smaller correlation length $\ell=5$.
 We show an example correlation matrix
@@
 for the Pb-Pb experimental data constructed using Eq.~\eqref{eq:corr} on the right-side of Fig.~\ref{fig:correlation}.
-Here $y_i$ denotes an element of the experimental data $\yv_e$ and $\sigma_i$ its corresponding uncertainty.
-The correlation matrix is block diagonal, with each block representing the correlations within a single class of observable.
+Here $y_i$ denotes an element of the experimental data $\ye$ and $\sigma_i$ its corresponding uncertainty.
 
@@
 
-In principle, one could calculate the likelihood function in Eq.~\eqref{eq:likelihood} directly, e.g.\ by running the model a large number of times at a given parameter point $\xv$ to calculate the model observables $\yv(\xv)$ from the ensemble of simulated events, but in practice such a procedure would be intractable.
+In principle, one could calculate the likelihood function in Eq.~\eqref{eq:likelihood} directly, e.g.\ by running the model a large number of times at a given parameter point $\xv$ to calculate the model observables $\ym(\xv)$ from the ensemble of simulated events, but in practice such a procedure would be intractable.
 The model is computationally intensive to evaluate, and thousands of events are required to calculate the simplest observables at a single parameter point.
@@
 We therefore follow an established framework for computationally intensive models and train an emulator to act as a fast surrogate for the full physics simulation \cite{OHagan:2006ba, Higdon:2008cmc, Higdon:2014tva}.
-The emulator enables essentially instantaneous predictions for $\yv = f(\xv)$ and allows us to sample the posterior distribution millions of times.
+The emulator enables essentially instantaneous predictions for $\ym = f(\xv)$ and allows us to sample the posterior distribution millions of times.
 In order to train the emulator, we first generate a scaffolding of the parameter space using a maximin Latin hypercube design \cite{Morris:1995lh} to distribute 500 points throughout our 15-dimensional parameter space according to the parameter ranges in Table~\ref{tab:design}.
@@
 Specifically, let $X$ denote the ${d \times n}$ design matrix of $d=500$ training points, where each training point is a vector $\xv = (x_1, x_2, \dots, x_n)$ of the $n=15$ model parameters in Table~\ref{tab:design}.
-Similarly, let $Y$ denote the corresponding ${d \times m}$ observables matrix, where each row of $Y$ is a vector $\yv = (y_1, y_2, \dots, y_m)$ of $m$ observables.
+Similarly, let $Y$ denote the corresponding ${d \times m}$ observables matrix, where each row of $Y$ is a vector $\ym = (y_1, y_2, \dots, y_m)$ of $m$ model observables.
+The emulator operates strictly on model outputs, so let's drop the subscript on $\ym$ hereafter to declutter the notation.
 Our goal is to train an emulator for the physics model $f$ using the discrete observations $f: X \mapsto Y$.
 
-We use for this purpose a specific type of emulator known as a Gaussian process emulator (GPE) \cite{Rasmussen:2006gp}.
-The advantage of using GPE's is that they provide an estimate of their own uncertainty which allows us to account for this uncertainty when constructing the covariance matrix $\Sigma$ in Eq.~\eqref{eq:likelihood}.
-One quirk of GPE's is that they are restricted to scalar-valued functions, i.e.\ functions of one output, whereas we require an emulator for vector-valued functions with multiple outputs.
+We use for this purpose a specific type of emulator known as a Gaussian process (GP) emulator \cite{Rasmussen:2006gp}.
+The advantage of using GPs is that they provide an estimate of their own uncertainty which allows us to account for this uncertainty when constructing the covariance matrix $\Sigma$ in Eq.~\eqref{eq:likelihood}.
+One quirk of GPs is that they are restricted to scalar-valued functions, i.e.\ functions of one output, whereas we require an emulator for vector-valued functions with multiple outputs.
 This restriction is commonly circumvented using principal component analysis (PCA): a general procedure which transforms a set of correlated variables $\yv = (y_1, y_2, \dots, y_m)$ into a new basis representation $\zv = (z_1, z_2, \dots, z_m)$ where the linear correlations between $z_i$, $z_j$ vanish for all $i\ne j \in m$ \cite{Tipping:1999}.
-Independent GPE's can then be used to emulate each $z \in \zv$ since the variables $(z_1, \dots, z_m)$ are linearly uncorrelated.
+Independent GPs can then be used to emulate each $z \in \zv$ since the variables $(z_1, \dots, z_m)$ are linearly uncorrelated.
 The emulated vector $\zv$ is then easily reexpressed in the basis of $\yv$ through its inverse transformation.
@@
 We first preprocess our model observables by centering and scaling each column of $Y$ (single observable) to zero mean and unit variance to generate a standardized observable matrix $\hat{Y}$.
-PCA is then used to reexpress each row-vector $\hat{\yv}$ of $\hat{Y}$ (all standardized observables at a single design point) in the new principal component basis:
+PCA is then used to reexpress each row-vector $\hat{\yv}_i$ of $\hat{Y}$ (all standardized observables at a single design point) in the new principal component basis:
 \begin{equation}
@@
 The principal components are reported in order of explained variance, with the first principal component vector $\mathbf{v}_1$ accounting for the most variance in $\hat{Y}$, and the last principal component vector $\mathbf{v}_m$ accounting for the least.
-We then train a set of independent GPE's $\{z_i=\mathrm{gp}_i(\xv)\}$ to predict the first $k$ principal components $(z_1, \dots, z_k)$ as a function of the model parameters $\xv$ which vary across the design $X$.
+We then train a set of independent GPs $\{z_i=\mathrm{gp}_i(\xv)\}$ to predict the first $k$ principal components $(z_1, \dots, z_k)$ as a function of the model parameters $\xv$ which vary across the design $X$.
 For the present study, we use $k=7$ principal components when emulating the $p$-Pb system and $k=8$ principal components when emulating Pb-Pb, chosen to describe 99.5\% of the total observed variance of each system.
 
-The GPE's are essentially fancy interpolators applied to the model's training points and PCA transformed observables.
-Each GPE reports a mean value $z(\xv)$ as well as an estimated error $\delta z(\xv)$ which accounts for statistical noise in the training data and interpolation error between the design points.
-Once the GPE's are trained, we can predict the observables $\yv$ at parameter point $\xv$ by transforming the vector of principal components
+The GPs are essentially fancy interpolators applied to the model's training points and PCA transformed observables.
+Each GP reports a mean value $z(\xv)$ as well as an estimated error $\delta z(\xv)$ which accounts for statistical noise in the training data and interpolation error between the design points.
+Once the GPs are trained, we can predict the observables $\ym$ at parameter point $\xv$ by transforming the vector of principal components
 \begin{equation}
@@
 \emph{back} to physical space.
-Similarly, we can construct the covariance matrix of the observables in PCA space,
+Similarly, we can construct the covariance matrix of the observables in PCA space
 \begin{equation}
@@
 \end{equation}
-and transform it back to physical space as well to obtain the covariance matrix of the model observables $\yv$ at a given parameter point $\xv$.
-
-The resulting emulator therefore predicts both a mean prediction $\yv(\xv)$ and an uncertainty covariance matrix $\Sigma_m(\xv)$ which accounts for multiple sources of model and emulator uncertainty, including the truncation error expected from using a finite number of principal components $k < m$.
-The model covariance matrix $\Sigma_m$ includes so-called \emph{known}-unknowns such as statistical error and emulator interpolation error, but not \emph{unknown}-unknowns such as the overall validity of small-system hydrodynamics, i.e.\ things which lack a unified consensus or are difficult to quantify.
+and transform it back to physical space as well to obtain the covariance matrix $\Sigmam$ of the model observables at parameter point $\xv$.
+
+The resulting emulator therefore predicts both a mean prediction $\ym(\xv)$ and an uncertainty covariance matrix $\Sigmam(\xv)$ which accounts for multiple sources of model and emulator uncertainty, including the truncation error expected from using a finite number of principal components $k < m$.
+The model covariance matrix $\Sigmam$ includes so-called \emph{known}-unknowns such as statistical error and emulator interpolation error, but not \emph{unknown}-unknowns such as the overall validity of small-system hydrodynamics, i.e.\ things which lack a unified consensus or are difficult to quantify.
 We show in Fig.~\ref{fig:correlation} the resulting Pb-Pb correlation matrix $\mathrm{corr}(y_i,y_j)$ for the model (emulator) at a random parameter point $\xv$ in the design space (left-side), along side the same correlation matrix for the experimental data (right-side) discussed previously.
@@
   \label{eq:joint_likelihood}
-  P(E | H_\xv) = P(E_\text{Pb-Pb} | H_\xv) \cdot P(E_\text{$p$-Pb} | H_\xv),
-\end{equation}
-where $E$ subsumes all evidence from the $p$-Pb and Pb-Pb collision systems and $H_\xv$ is our hypothesis that $\xv$ equals the true parameters $\xv_\star$.
-We then perform Markov-chain Monte Carlo (MCMC) importance sampling on the posterior distribution in Eq.~\eqref{eq:bayes} to draw random samples for $P(H_\xv | E)$, the estimate of the true model parameters given the model and the experimental data \cite{Goodman:2010en, FM:2013mc}.
+  P(E | \xv) = P(E_\text{Pb-Pb} | \xv) \cdot P(E_\text{$p$-Pb} | \xv),
+\end{equation}
+where $E$ subsumes all evidence from the $p$-Pb and \mbox{Pb-Pb} collision systems.
+We then perform Markov-chain Monte Carlo (MCMC) importance sampling on the posterior distribution $P(\xv | E)$ in Eq.~\eqref{eq:bayes} to draw random samples of the true model parameters $\xv_\star$, given the predictions of the model and the evidence provided by the experimental data \cite{Goodman:2010en, FM:2013mc}.
 For this we use an affine-invariant ensemble sampler which uses a large ensemble of interdependent walkers \cite{Goodman:2010en, FM:2013mc} and allow the MCMC chain to ``burn-in'' before generating \order{7} posterior samples.
@@
 We now direct our attention to Fig.~\ref{fig:posterior} which shows the main result of this work, the posterior distribution of the model input parameters.
-Recall that the posterior $P(H_\xv | E)$ is the probability that our hypothesis $H_\xv$ for the true model parameters $\xv_\star$ is correct, given the evidence $E$ provided by experiment.
+Recall that the posterior $P(\xv | E)$ is the probability that our hypothesis $\xv = \xv_\star$ is correct, given the evidence $E$ provided by the predictions of the model and the experimental data.
 The present posterior has 15 dimensions, one for each parameter listed in Table~\ref{tab:design}, and thus its joint distribution cannot be summarized by one figure alone.
@@
 It is therefore compelling that $p \sim 0$ correctly describes the charged-particle yield $d\nch/d\eta$ of both systems, while simultaneously describing the centrality dependence of $\vnk{n}{k}$, an observable which is also known to strongly depend on $p$ \cite{Bernhard:2018hnz}.
-Specifically, this value $p \sim 0$ corresponds to an energy deposition mapping proportional to the geometric mean of participant nuclear thickness,
+Specifically, this value $p \sim 0$ corresponds to an energy deposition mapping proportional to the geometric mean of participant nuclear thickness
 \begin{equation}
@@
 For example, using the posterior distribution's median values, $r=0.88$~fm, $n_c=6$, and $v=0.52$~fm, we can generate a large ensemble of random nucleon configurations and average their density in each nucleon's center-of-mass frame.
-The resulting ensemble-averaged nucleon density,
-\begin{equation}
-  \bar{\rho}_n(\xv) = \frac{1}{(2 \pi w^2)^{3/2}} \exp \left( -\frac{|\xv|^2}{2 w^2} \right),
+The resulting ensemble-averaged nucleon density
+\begin{equation}
+  \bar{\rho}_n(\xv) = \frac{1}{(2 \pi w^2)^{3/2}} \exp \left( -\frac{|\xv|^2}{2 w^2} \right)
 \end{equation}
 is described by a single Gaussian of width $w = 0.96$~fm.
-This nucleon width is consistent with a previous estimate, $w = 0.96_{-0.05}^{+0.04}$~fm, obtained by a similar Bayesian analysis of Pb-Pb collisions at $\sqrts=2.76$ and 5.02~TeV using a physics model without nucleon substructure \cite{Bernhard:2018hnz}.
+This nucleon width is consistent with a previous estimate $w = 0.96_{-0.05}^{+0.04}$~fm obtained by a similar Bayesian analysis of Pb-Pb collisions at $\sqrts=2.76$ and 5.02~TeV using a physics model without nucleon substructure \cite{Bernhard:2018hnz}.
 
@@
     \label{tab:mode_params}
-    High-probability parameters selected from the posterior distribution and used to generate Fig.~\ref{fig:obs_map}.
+    \emph{Maximum a posteriori} (MAP) parameters determined from the posterior distribution and used to generate Fig.~\ref{fig:obs_map}.
     The posterior distribution on the particlization temperature $\Tsw$ is flat (agnostic), so we fix it's value using Ref.~\cite{Bernhard:2018hnz}.
@@
 We verified the emulator and tested the accuracy of our physics model framework using a single set of high-probability parameters selected from the Bayesian posterior.
-These parameters, listed in Table~\ref{tab:mode_params}, are the approximate ``best fit'' values of the calibrated model, commonly referred to as the \emph{maximum a posteriori} (MAP) estimate:
-\begin{equation}
-  \xv_\mathrm{MAP} \equiv \operatorname*{arg\, max}_{\xv} P(H_\xv | E).
+These parameters, listed in Table~\ref{tab:mode_params}, are the approximate ``best fit'' values of the calibrated model, commonly referred to as the \emph{maximum a posteriori} (MAP) estimate
+\begin{equation}
+  \xv_\mathrm{MAP} \equiv \operatorname*{arg\, max}_{\xv} P(\xv | E).
 \end{equation}
@@
 
-We also present calculations for several observables which were omitted from the calibration due to the statistical limitations of our training data.
+We also present calculations for several observables which were omitted from the calibration due to missing experimental data and the statistical limitations of our training data.
 Here our MAP event sample is several orders of magnitude larger so the statistics are no issue.
-The bottom-right panel of Fig.~\ref{fig:obs_map} shows our model calculation for the four-particle elliptic flow cumultant $\vnk{2}{4}$ along with the measured data points from \mbox{ALICE} \cite{Adam:2016izf}.
+The bottom-right panel of Fig.~\ref{fig:obs_map} shows our model calculation for the four-particle elliptic flow cumulant $\vnk{2}{4}$ along with the measured data points from \mbox{ALICE} \cite{Adam:2016izf}.
 We see that the MAP estimate nicely describes the measured $\vnk{2}{4}$ data which is encouraging since this particular observable was never used to calibrate the model.
@@
 
-Lastly, we compute the symmetric cumulants $\SC(m,n)$ for the Pb-Pb collision system at $\sqrts=5.02$~TeV which quantify correlations between event-by-event fluctuations of the flow harmonics of different order \cite{Bilandzic:2013kga, ALICE:2016kpq},
+Lastly, we compute the symmetric cumulants $\SC(m,n)$ for the Pb-Pb collision system at $\sqrts=5.02$~TeV which quantify correlations between event-by-event fluctuations of the flow harmonics of different order \cite{Bilandzic:2013kga, ALICE:2016kpq}
 \begin{align}
@@
 which are expected to be less sensitive to the medium response and more sensitive to the properties of the initial state.
-The solid lines are the MAP estimate of the present study and the dashed lines are the MAP estimate of Ref.~\cite{Bernhard:2018hnz} which did not include nucleon substructure and was calibrated on Pb-Pb collisions at $\sqrts=2.76$ and 5.02~TeV.
+The solid lines are the MAP estimate of the present study, and the dashed lines are the MAP estimate of Ref.~\cite{Bernhard:2018hnz} which did not include nucleon substructure and was calibrated on Pb-Pb collisions at $\sqrts=2.76$ and 5.02~TeV.
 We observe that the gap between $\SC(3,2)$ and $\SC(4,2)$ is generally wider in the present analysis than in Ref.~\cite{Bernhard:2018hnz}, as is the gap between the normalized symmetric cumulants $\NSC(3,2)$ and $\NSC(4,2)$.
@@
 Relativistic heavy-ion collisions produce long-range multiparticle correlations which are commonly explained by the existence of hydrodynamic flow \cite{deSouza:2015ena}.
-This narrative is evidenced by the global, self-consistent and highly non-trivial quantitative agreement of hydrodynamic models with a large number of heavy-ion bulk observables \cite{Niemi:2015qia, Bernhard:2016tnd, Gale:2012rq}.
+This narrative is evidenced by the global, self-consistent, and highly non-trivial quantitative agreement of hydrodynamic models with a large number of heavy-ion bulk observables \cite{Niemi:2015qia, Bernhard:2016tnd, Gale:2012rq}.
 Naturally, such descriptions rely on the validity of hydrodynamic approximations, and these approximations begin to break down in the so-called dilute limit where discrete particle degrees-of-freedom dominate and continuous field descriptions of the medium cease to make sense.
@@
     We find that nucleon substructure is necessary to simultaneously describe $p$-Pb and Pb-Pb bulk observables.
-    However, we observe no strong preference for a specific number of constituents (hot spots) inside the nucleon.
+    However, we observe no strong preference for a specific number of constituents $n_c$ (hot spots) inside the nucleon.
+    In particular, we find no evidence to support the specific number $n_c=3$ which is commonly used in the literature.
   \item
@@
 Notably absent is data from RHIC which includes measurements for $p$-Au, $d$-Au, $^3$He-Au, Cu-Cu, Cu-Au, Au-Au, and \mbox{U-U} collision systems at $\sqrts=200$~GeV.
-This data would undoubtedly provide an additional constraint on the model parameters and would enable more stringent tests of the calibrated model predictions.
+This data would undoubtedly help constrain the model parameters and would enable more stringent tests of the calibrated model predictions.
 The RHIC data may also help elucidate the beam-energy dependence of the model parameters which would be worth investigating.
@@
 
-\begin{figure}
+\begin{figure}[b]
   \begin{tikzpicture}[scale=.9]
@@
     We use the k-fold cross validation method (explained in the text) to partition the model inputs $X$ and outputs $Y$ into training and validation data.
-    The scatterplot on the left shows the emulator predictions and one sigma error bars (x-axis) against explicit model calculations (y-axis).
+    The scatter plot on the left shows the emulator predictions and one sigma error bars (x-axis) against explicit model calculations (y-axis).
     Perfect emulator/model agreement is indicated by the black like $y_\text{pred}=y_\text{obs}$.
@@
 
-We find that we can quickly estimate the maximum radius $R_\text{max} = |\xv_\text{max}|^2$ of the spacetime hypersurface ${T = T(e_\text{min})}$ by running the event on a coarse-grained spatial grid with one-third the spatial resolution we would otherwise require to resolve typical hydrodynamic observables such as mean $p_T$ and flows.
+We find that we can quickly estimate the maximum transverse radius $R_\text{max}$ of the spacetime hypersurface ${T = T(e_\text{min})}$ by running the event on a coarse-grained spatial grid with one-third the spatial resolution we would otherwise require to resolve typical hydrodynamic observables such as mean $p_T$ and flows.
 The simulation time of a single VISH2+1 event scales like ${\sim}n_x^3$ since $dx \propto d\tau$, and thus our ``pre-run'' event requires only ${\sim}1/27$th the time of a production event.
-We therefore run a coarse-grained pre-event on an excessively large grid for \emph{every} minimum-bias event to estimate $R_\text{max}$, then rerun the same event on a thrice finer grid with a trimmed spatial extent $x_\text{max} \equiv R_\text{max}$.
+We therefore start by running a coarse-grained event on an excessively large grid for \emph{every} minimum-bias event to estimate $R_\text{max}$, then rerun the same event on a thrice finer grid with a trimmed spatial extent $x_\text{max} = R_\text{max}$.
 See Fig.~\ref{fig:adaptive_grid} for a simple diagram of the procedure.
@@
 This is because the maximum transverse size of each event varies dramatically, from a few fm in peripheral Pb-Pb collisions to 50 fm or more in central Pb-Pb collisions.
-The procedure should generalize to other hydrodynamic codes, including those with $3+1$ spacetime dimensions, where the time savings could be even more pronounced.
+The procedure should generalize to other hydrodynamic codes.
 
@@
 
-\begin{figure*}
+The emulator is a surrogate for the full physics simulation which generates probabilistic predictions for the model observables $\ym$ at a given point $\xv$.
+Here we validate these probabilistic predictions using a method known as k-fold cross validation.
+We first randomly partition our $d=500$ training points into $k=20$ equal sized subsamples or ``folds''.
+One of the subsamples is used to validate the emulator and the remaining $k-1$ subsamples are used to train it.
+The process is then repeated for each of the subsamples so that we end up validating on all of the training data.
+
+\begin{figure*}[b]
   \includegraphics{validation_PbPb5020}\\
@@
 
-The emulator is a surrogate for the full physics simulation which generates probabilistic predictions for the model observables $\yv$ at a given point $\xv$.
-Here we validate these probabilistic predictions using a method known as k-fold cross validation.
-We first randomly partition our $d=500$ training points into $k=20$ equal sized subsamples or ``folds''.
-One of the subsamples is used to validate the emulator and the remaining $k-1$ subsamples are used to train it.
-The process is then repeated for each of the subsamples so that we end up validating on all of the training data.
-
-Figure~\ref{fig:validation_example} shows a scatterplot of the emulator predictions with one-sigma error bars (x-axis) against explicit model calculations (y-axis).
+Figure~\ref{fig:validation_example} shows a scatter plot of the emulator predictions with one-sigma error bars (x-axis) against explicit model calculations (y-axis).
 Perfect emulator and model agreement is indicated by the black line $y_\text{pred} = y_\text{obs}$.
-If the emulator errors are properly accounted for, then the normalized residuals ${z=(y_\text{pred} - y_\text{obs})/\sigma_\text{pred}}$ sample a unit normal distribution:
+If the emulator errors are properly accounted for, then the normalized residuals ${z=(y_\text{pred} - y_\text{obs})/\sigma_\text{pred}}$ sample a unit normal distribution
 \begin{equation}
@@
 These visual references should be compared to the whiskers and interquartile range respectively of each box plot, analogous to the comparison test of Fig.~\ref{fig:validation_example}.
-The emulators generally behave as expected, although the validation is somewhat better for the Pb-Pb system than the $p$-Pb system.
+The emulators generally behave as expected, although the validation is better for the Pb-Pb system than the $p$-Pb system.
 For instance, the $p$-Pb charged-particle yield $d\nch/d\eta$ uncertainties are over predicted.
@@
 Thus $\hat{z}$ can be thought of as a fractional emulator error relative to the full design variability.
-The bottom row of each figure shows the root-mean-square value of $\hat{z}$.
-We see that $\text{RMS}\{\hat{z}\}$ ranges from a few percent for most observables to a maximum value of 15\% for the $p$-Pb triangular flow $v_3\{2\}$ in the lowest multiplicity bin.
+The bottom row of each figure shows the root-mean-square (RMS) value of $\hat{z}$.
+We see that RMS $\hat{z}$ ranges from a few percent for most observables to a maximum value of 15\% for the $p$-Pb triangular flow $v_3\{2\}$ in the lowest multiplicity bin.
 This suggests that the present analysis would benefit the most from more $p$-Pb events, in particular, from more multiplicity triggered events which are used to calculate the flows.

### 2019-05-07 paper/response.txt
Final revisions

@@
 
-"A parton model of the proton has direct consequences not only for heavy-ion collisions, but also for p+p collisions. The authors should present predictions for p+p scattering using the most probable
-parameters (or using the posterior distribution of parameters w, v, n_c) for the differential elastic p+p scattering cross section.
-The prediction can be compared to parametrized data at 5TeV.
-As a proxy
-for the p+p data the parametrization from Phys. Rev. D88 (2013) 094019 can be used."
+"A parton model of the proton has direct consequences not only for heavy-ion collisions, but also for p+p collisions.
+The authors should present predictions for p+p scattering using the most probable parameters (or using the posterior distribution of parameters w, v, n_c) for the differential elastic p+p scattering cross section.
+The prediction can be compared to parametrized data at 5 TeV.
+As a proxy for the p+p data the parametrization from Phys.
+Rev.
+D88 (2013) 094019 can be used."
 
@@
 We've been careful to include all available data (public or otherwise), but it is possible that we missed a dataset.
-In response to your suggestion, we've sent an email to the ALICE collaboration to request their preliminary data, e.g.
+In response to your suggestion, we sent an email to the ALICE collaboration to request their preliminary data, e.g.
 Pb-Pb mean pT, which could be added to the posterior predictions where data is missing.
-Unfortunately, I have not heard back from them yet, and I do not believe it is their policy to share preliminary data.
+Unfortunately, it is not their policy to share preliminary data, and our request was denied.
 If you can direct us to a specific dataset, we would be happy to include it in the maximum a posteriori observable plot.

### 2019-05-07 paper/paper.tex
Final revisions

@@
 \newcommand{\Tsw}{T_\text{switch}}
+\newcommand{\Tc}{T_c}
 \newcommand{\T}{\tilde{T}}
 
+\newcommand{\rc}{r_{cp}}
+\newcommand{\nc}{n_c}
+\newcommand{\wc}{w_c}
+
 \newcommand{\smin}{{(\eta/s)_\mathrm{min}}}
-\newcommand{\sslope}{{a_\mathrm{shear}}}
-\newcommand{\scrv}{{c_\mathrm{shear}}}
+\newcommand{\sslope}{{(\eta/s)_\mathrm{slope}}}
+\newcommand{\scrv}{{(\eta/s)_\mathrm{crv}}}
 \newcommand{\bmax}{{(\zeta/s)_\mathrm{max}}}
-\newcommand{\bloc}{{T_\text{bulk}}}
-\newcommand{\bwidth}{{\Delta T_\text{bulk}}}
+\newcommand{\bwidth}{{(\zeta/s)_\mathrm{width}}}
+\newcommand{\bloc}{{(\zeta/s)_{T_\mathrm{peak}}}}
 
@@
 
-Hydrodynamic models of ultrarelativistic nuclear collisions are complicated by a number of theoretical unknowns, including the detailed geometry of the QGP initial conditions, the strength and duration of pre-equilibrium dynamics, the temperature dependence of QGP transport coefficients, and the boundaries of hydrodynamic applicability \cite{Niemi:2014lha, deSouza:2015ena, Ollitrault:2012cm, Song:2012ua}.
+Hydrodynamic models of ultrarelativistic nuclear collisions are complicated by a number of theoretical unknowns including the detailed geometry of the QGP initial conditions, the strength and duration of pre-equilibrium dynamics, the temperature dependence of QGP transport coefficients, and the boundaries of hydrodynamic applicability \cite{Niemi:2014lha, deSouza:2015ena, Ollitrault:2012cm, Song:2012ua}.
 In general, these theoretical uncertainties tend to grow with decreasing system size, where emergent physics at sub-femtometer length scales becomes important to describe bulk properties of the produced system.
@@
 One method for reducing theoretical uncertainties is to test model calculations by varying the species of colliding nuclei at a single beam energy \cite{Adare:2015bua, Schenke:2014tga, Aidala:2018mcw, Adare:2017wlc, Adamczyk:2015obl, Shen:2016zpp, Aidala:2017ajz, Adare:2006ti}.
-Since initial condition and hydrodynamic models generally factorize the structure of the colliding nuclei from the subsequent time dynamics of the collision, a given theory framework can be validated by testing its predictions for multiple collision systems, using a single self-consistent set of model parameters where only the nuclear structure in the model is permitted to vary.
+Since initial condition and hydrodynamic models generally factorize the structure of the colliding nuclei from the subsequent time dynamics of the collision, a proposed theory framework can be validated by testing its predictions for multiple collision systems using a single self-consistent set of model parameters, where only the nuclear structure in the model is permitted to vary.
 
@@
 Measurements by the TOTEM collaboration at $\sqrt{s}=7$~TeV, for instance, found an unexpected dip in the inelasticity density of $p$-$p$ collisions at zero impact parameter \cite{Antchev:2011zz}.
-It was later realized that this depression, or so-called hollowness effect in the $p$-$p$ inelastic collision profile \cite{Arriola2016}, can be explained by the existence of correlated domains inside the proton, and that aspects of these domains, such as their size and correlation strength, may be constrained by comparing model predictions to inelastic $p$-$p$ measurements \cite{Albacete:2016gxu, Albacete:2016pmp}.
+It was later realized that this depression, or so-called hollowness effect, in the $p$-$p$ inelastic collision profile \cite{Arriola2016} can be explained by the existence of correlated domains inside the proton, and that aspects of these domains, such as their size and correlation strength, may be constrained by comparing model predictions to inelastic $p$-$p$ measurements \cite{Albacete:2016gxu, Albacete:2016pmp}.
 
@@
 \begin{enumerate*}[label=(\arabic*)]
-  \item explore all parameter combinations and
-  \item compare model predictions to all relevant experimental measurements in a statistically rigorous fashion.
+  \item
+    explore all parameter combinations and
+  \item
+    compare model predictions to all relevant experimental measurements in a statistically rigorous fashion.
 \end{enumerate*}
 
-With these considerations in mind, we present progress towards a fully global analysis of $p$-Pb and Pb-Pb bulk observables at $\sqrts=5.02$~TeV using a model calibration framework known as Bayesian parameter estimation.
+With these considerations in mind, we present progress toward a fully global analysis of $p$-Pb and Pb-Pb bulk observables at $\sqrts=5.02$~TeV using a model calibration framework known as Bayesian parameter estimation.
 We begin in Sec.~\ref{sec:model} by constructing a nuclear collision model for $p$-Pb and Pb-Pb collisions using initial conditions with parametric nucleon substructure and transport dynamics described by a pre-equilibrium free streaming stage, viscous hydrodynamics, and microscopic Boltzmann transport.
@@
 The resulting nuclear overlap time $\tau_\text{overlap} \lesssim 0.1\ \fmc$ at top RHIC and LHC energies, and thus it is reasonable to neglect the initial transverse dynamics which occur while the nuclei pass through each other.
-We therefore assume that the collision produces all secondary particles at a uniform proper time $\tau_0 \ll 1~\fmc$, and that it deposits energy at midrapidity which is a function of the locally varying density inside each nucleus.
+We therefore assume that the collision produces all secondary particles at a uniform proper time $\tau_0 \ll 1~\fmc$, and that it deposits energy at midrapidity which is a function of the locally varying transverse density inside each nucleus.
 
@@
 In this work, we pursue a less restrictive and more parametric description of the nucleon where the number of substructure degrees of freedom is uncertain as depicted in Fig.~\ref{fig:substructure}.
-We model each nucleon's three-dimensional density $\rho_{A,B}$ as a sum of $n_c$ independent constituents
+We model each nucleon's three-dimensional density $\rho_{A,B}$ as a sum of $\nc$ independent constituents
 \begin{equation}
   \label{eq:rho}
-  \rho_{A, B}(\xv) = \frac{1}{n_c} \sum\limits_{i=1}^{n_c} \rho_c(\xv - \xv_i),
+  \rho_{A, B}(\xv) = \frac{1}{\nc} \sum\limits_{i=1}^{\nc} \rho_c(\xv - \xv_i),
 \end{equation}
@@
 \begin{equation}
-  \rho_c(\xv) = \frac{1}{(2 \pi v^2)^{3/2}} \exp\left(- \frac{|\xv|^2}{2 v^2}\right)
-\end{equation}
-of variable width $v$.
-The constituent positions $\xv_i$ are sampled independently (ignoring correlations) from a Gaussian radial distribution
+  \rho_c(\xv) = \frac{1}{(2 \pi \wc^2)^{3/2}} \exp\left(- \frac{|\xv|^2}{2 \wc^2}\right)
+\end{equation}
+of variable width $\wc$.
+The constituent positions $\xv_i$ are sampled independently (ignoring correlations) from a Gaussian probability distribution
 \begin{equation}
   \label{eq:radial_dist}
-  \rho_\text{pos}(\xv_i; \xv_{A,B}) = \frac{1}{(2\pi r^2)^{3/2}} \exp\left(-\frac{|\xv_i - \xv_{A,B}|^2}{2 r^2}\right),
-\end{equation}
-where $r$ is a parameter that varies the sampling radius of the constituent positions $\xv_i$ about the nucleon position $\xv_{A,B}$.
+  P(\xv_i) = \frac{1}{(2\pi \rc^2)^{3/2}} \exp\left(-\frac{|\xv_i - \xv_n|^2}{2\, \rc^2}\right),
+\end{equation}
+where $\rc$ is a free parameter that varies the dispersion of the constituent positions $\xv_i$ about each nucleon position $\xv_n$.
 As a matter of convenience, we sample the nucleon positions \emph{before} determining the constituent positions.
-This leads to small discrepancies between the designated nucleon positions $\xv_{A,B}$ and the actual position of each nucleon's center-of-mass, owing to fluctuations in the constituent positions $\xv_i$.
-The parameter $r$ should thus be interpreted with care.
+This leads to small discrepancies between the designated nucleon positions and the actual position of each nucleon's center-of-mass, owing to fluctuations in the constituent positions.
+The parameter $\rc$ should thus be interpreted with care.
 It is a computational sampling radius, \emph{not} the Gaussian width of the sampled nucleons in their center-of-mass frame.
@@
 \begin{equation}
-  \tilde{\rho}_{A,B}(\xv) \equiv \frac{1}{n_c} \sum\limits_{i=1}^{n_c} \gamma_i\, \rho_c(\xv - \xv_i)
+  \tilde{\rho}_{A,B}(\xv) \equiv \frac{1}{\nc} \sum\limits_{i=1}^{\nc} \gamma_i\, \rho_c(\xv - \xv_i)
 \end{equation}
@@
 
-The resulting nucleon fluctuation variance naturally falls like ${\sim}1/ n_c$, where $n_c$ is the number of independent fluctuating sources.
-This means that the natural range for the constituent fluctuations is larger when the number of constituents is larger and \emph{vice versa}.
+The resulting nucleon fluctuation variance naturally falls like ${\sim}1/ \nc$, where $\nc$ is the number of constituents.
+This means that the natural range for the constituent fluctuations is larger when the number of constituents is larger.
 We therefore reparametrize the constituent fluctuations using
 \begin{equation}
-  \sigma_\mathrm{fluct} = 1 / \sqrt{k\, n_c},
-\end{equation}
-where $k$ is the shape parameter (inverse variance) of the gamma distribution weights and $n_c$ is the number of nucleon constituents.
+  \sigma_\mathrm{fluct} = 1 / \sqrt{k\, \nc},
+\end{equation}
+where $k$ is the shape parameter (inverse variance) of the gamma distribution weights and $\nc$ is the number of nucleon constituents.
 
@@
 This wounded nucleon scaling was in fact one of the first such mappings used as a proxy for initial particle production in relativistic heavy-ion collisions \cite{Bialas:1976ed}.
-It was subsequently realized, however, that the wounded nucleon model predicts the wrong scaling for charged-particle production as a function of collision centrality \cite{Kharzeev:2000ph} and hence the wrong scaling for midrapidity energy deposition
-as a function of participant density.\footnote{Typically, wounded nucleon scaling is used to parametrize the entropy density $s$, but the shortcomings of the model are nevertheless the same when parametrizing the system's energy density $e$.
+It was subsequently realized, however, that the wounded nucleon model predicts the wrong scaling for charged-particle production as a function of collision centrality \cite{Kharzeev:2000ph} and hence the wrong scaling for midrapidity energy deposition as a function of participant density.\footnote{Typically, wounded nucleon scaling is used to parametrize the entropy density $s$, but the shortcomings of the model are nevertheless the same when parametrizing the system's energy density $e$.
 Both parametrizations underpredict the steep rise of particle production observed in central collisions.}
@@
 This parametrization introduces a dimensionless parameter $p$ which varies the scaling behavior of initial energy deposition at midrapidity.
-For certain discrete values of $p$, it reduces to well known functional forms such as the arithmetic, geometric, and harmonic means:
+For certain discrete $p$ values, it reduces to well known functional forms such as the arithmetic, geometric, and harmonic means:
 \begin{equation}
@@
 \end{equation}
-Our motivation for reinterpretting the left-side of this equation as an energy density in the present work is to enable the application of pre-equilibrium free streaming equations (described shortly) which require the system's initial energy density as input.
-This modification is permissible since the \trento\ model is based on particle yield observations coupled with the approximate scaling relation $d\nch/d\eta \propto M_p(\T_A, \T_B)$, which remains valid for both static entropy density and free streamed energy density initialization.
+Our motivation for reinterpreting the left-side of this equation as an energy density in the present work is to enable the application of pre-equilibrium free streaming equations (described shortly) which require the system's initial energy density as input.
+This modification is permissible since the \trento\ model is formulated using particle yield observations and the approximate scaling relation $d\nch/d\eta \propto M_p(\T_A, \T_B)$, which remains valid for both static entropy density and free streamed energy density initialization.
 
@@
   \end{tikzpicture}
-  \caption{\label{fig:thickness} Effect of nucleon substructure on the nuclear thickness function $T(x, y) \equiv \int dz\, \rho(x, y, z)$ of a $^{208}\mathrm{Pb}$ nucleus.
-The nucleus on the left has Gaussian nucleons of width $0.8$~fm, while the nucleus on the right has composite nucleons, each containing six constituents of width $0.4$~fm.}
+  \caption{
+    \label{fig:thickness}
+    Effect of nucleon substructure on the nuclear thickness function $T(x, y) \equiv \int dz\, \rho(x, y, z)$ of a $^{208}\mathrm{Pb}$ nucleus.
+    The nucleus on the left has Gaussian nucleons of width $0.8$~fm, while the nucleus on the right has composite nucleons, each containing six constituents of width $0.4$~fm.
+  }
 \end{figure}
@@
 Up to this point we have restricted our attention to a single nucleon-nucleon collision.
-Equation~\eqref{eq:gmean} is a purely local function of nuclear density in the transverse plane and hence it should (in principle) be equally valid for any pair of colliding nuclei at sufficiently high beam energy.
+Equation~\eqref{eq:gmean} is a purely local function of nuclear density in the transverse plane and hence it should, in principle, be equally valid for any pair of colliding nuclei at sufficiently high beam energy.
 The \trento\ model readily generalizes from individual nucleon-nucleon collisions to arbitrary nucleus-nucleus collisions by summing the participant thicknesses $\T_{A,B}$ in Eq.~\eqref{eq:fluctuated_thick} over all nucleons which participate in one or more inelastic collisions.
@@
   \includegraphics{coupling}
-  \caption{Cartoon of the free streaming approximation for hydrodynamic initialization.
+  \caption{
+    \label{fig:coupling}
+    Cartoon of the free streaming approximation for hydrodynamic initialization.
 The initial state is free streamed for proper time $\taufs$ (zero coupling) before it is matched to hydrodynamics (strong coupling).
-This piecewise evolution approximates the more realistic scenario expected in nature where the medium's coupling strength smoothly changes as a function of time.}
-  \label{fig:coupling}
+    This piecewise evolution approximates the more realistic scenario expected in nature where the medium's coupling strength smoothly changes as a function of time.
+  }
 \end{figure}
@@
   \label{freestream}
-  p^\mu \partial_\mu f(x, p) = 0,
+  p^\mu \partial_\mu f(x, p) = 0
 \end{equation}
@@
 \end{equation}
-The initial bulk and shear corrections are finally solved for by subtracting the ideal pressure from the total pressure to find $\Pi$, then solving for $\pi^{\mu\nu}$ using Eq.~\eqref{hydro-eqn}
+The initial bulk and shear corrections are finally solved for by subtracting the ideal pressure from the total pressure to find $\Pi$, then solving for $\pi^{\mu\nu}$ using Eq.~\eqref{hydro-eqn}:
 \begin{align}
@@
   \includegraphics{viscosity_dof}
-  \caption{Degrees of freedom in the temperature dependent shear and bulk viscosity parametrizations.
-Lines are chosen for illustrative purposes only and do not represent all possible variability.
-For instance, $\eta/s$ could have a large slope and negative curvature, or $\zeta/s$ could have a large max and narrow width, neither of which are depicted above.}
+  \caption{
     \label{fig:viscosity_dof}
+    Degrees of freedom in the temperature dependent shear and bulk viscosity parametrizations. Lines are chosen for illustrative purposes only and do not represent all possible variability.
+    For instance, $\eta/s$ could have a large slope and negative curvature, or $\zeta/s$ could have a large max and narrow width, neither of which are depicted above.
+  }
 \end{figure}
@@
 The hydrodynamic equations of motion are necessarily closed using an equation of state (EoS) to relate the energy density $e$ and pressure $P$ of the produced medium.
-We use a parametrization for $P(e)$ that matches a hadron resonance gas EoS at low temperature to a lattice QCD EoS at high temperature by smoothly connecting their trace anomalies in the interval
-$165 \le T \le 200$~MeV \cite{Bernhard:2018hnz}.
+We use a parametrization for $P(e)$ that matches a hadron resonance gas EoS at low temperature to a lattice QCD EoS at high temperature by smoothly connecting their trace anomalies in the interval $165 \le T \le 200$~MeV \cite{Bernhard:2018hnz}.
 For the lattice EoS, we use a calculation by the HotQCD collaboration for (2+1)-flavor QCD which was extrapolated to the continuum limit \cite{Bazavov:2014pvz}.
@@
   \label{eq:shear_viscosity}
-  (\eta/s)(T) = \smin + \sslope \cdot(T - T_c)\cdot \left ( \frac{T}{T_c} \right)^\scrv,
-\end{equation}
-where $\smin$, $\sslope$, and $\scrv$ are tunable parameters, and $T_c=0.154$~GeV is the pseudocritical transition temperature of the HotQCD EoS.
+(\eta/s)(T) = \smin + \sslope \cdot (T - \Tc) \left ( \frac{T}{\Tc} \right)^\scrv,
+\end{equation}
+where $\smin$, $\sslope$, and $\scrv$ are tunable parameters, and $\Tc=0.154$~GeV is the pseudocritical transition temperature of the HotQCD EoS.
 Meanwhile, for the specific bulk viscosity $\zeta/s$, we use an unnormalized Cauchy distribution
@@
 The aforementioned hydrodynamic equations are solved numerically using the boost-invariant VISH2+1 viscous hydrodynamics code \cite{Song:2007ux, Shen:2014vra}.
-We vary the spatial grid's maximum width $x_\mathrm{max}$ event-by-event to accommodate systems of varying size and determine the optimal spatial step $dx$ and time step $d\tau$ for each design point to balance trade-offs between numerical accuracy and computation time (see Appendix~\ref{app:adaptive_grid}).
+We vary the spatial grid's maximum width $x_\mathrm{max}$ event-by-event to accommodate systems of varying size and determine the optimal spatial step $dx$ and time step $d\tau$ for each set of model parameters to balance trade-offs between numerical accuracy and computation time (see Appendix~\ref{app:adaptive_grid}).
 Although these details are somewhat mundane, they are critically important to the present study, since the computation time scales with the number of spacetime cells $n_\text{cell} \sim n_x^2\, n_\tau$, and this quantity grows rapidly when $n_x$ and $n_\tau$ are large, as is typically the case for simulations with nucleon substructure.
@@
 \begin{enumerate}[itemsep=0pt, leftmargin=2\parindent]
-  \item $H_f$: the hypothesis that the nuclear collision model $f$ formulated in this work provides a realistic description of reality,
-  \item $H_\xv$: the hypothesis that the model parameters $\xv$ are the true model parameters $\xv_\star$ of $f$, and
-  \item $E$: the evidence provided by the experimental data $\ye$ and its corresponding uncertainties.
+  \item
+    $H_f$: the hypothesis that the nuclear collision model $f$ formulated in this work provides a realistic description of reality,
+  \item
+    $H_\xv$: the hypothesis that the model parameters $\xv$ are the true model parameters $\xv_\star$ of $f$, and
+  \item
+    $E$: the evidence provided by the model $\ym$, the experimental data $\ye$, and their corresponding uncertainties.
 \end{enumerate}
@@
 \begin{table}[t]
-  \caption{Input parameter ranges for the physics model.}
+  \caption{
+    \label{tab:design}
+    Input parameter ranges for the physics model.
+  }
   \begin{ruledtabular}
@@
     $\sigmaf$         & Nucleon fluctuation std.\ dev.\      & 0--2             \\
-    $r$               & Constituent sampling radius          & 0--1.2 fm       \\
-    $n_c$             & Number of nucleon constituents       & 1--9            \\
-    $v$               & Constituent width                    & 0.2--1.2 fm     \\
-    $\dmin$           & Minimum inter-nucleon distance       & 0--1.7 fm       \\
+    $\rc$             & Constituent position radius          & 0--1.2 fm        \\
+    $\nc$             & Number of constituents               & 1--9             \\
+    $\wc$             & Width of constituents                & 0.2--1.2 fm      \\
+    $\dmin$           & Minimum inter-nucleon dist.\         & 0--1.7 fm        \\
     $\taufs$          & Free streaming time                  & 0.1--1.5 \fmc    \\
-    $\smin$           & Minimum value of $\eta/s$ (at $T_c$) & 0--0.2          \\
-    $\sslope$         & Slope of $\eta/s$ above $T_c$        & 0--8 GeV$^{-1}$ \\
-    $\scrv$           & Curvature of $\eta/s$ above $T_c$    & $-1$ to $+1$    \\
+    $\smin$           & Minimum value of $\eta/s$ (at $\Tc$) & 0--0.2           \\
+    $\sslope$         & Slope of $\eta/s$ above $\Tc$        & 0--8 GeV$^{-1}$  \\
+    $\scrv$           & Curvature of $\eta/s$ above $\Tc$    & $-1$ to $+1$     \\
     $\bmax$           & Maximum value of $\zeta/s$           & 0--0.1           \\
     $\bwidth$         & Width of $\zeta/s$ peak              & 0--0.1 GeV       \\
-    $\bloc$           & Temperature of $\zeta/s$ maximum     & 150--200 MeV    \\
-    $\Tsw$            & Switching/particlization temp.       & 135--165 MeV    \\
+    $\bloc$           & Temp.\ of $\zeta/s$ maximum          & 0.150--0.200 GeV \\
+    $\Tsw$            & Switching/particlization temp.       & 0.135--0.165 GeV \\
   \end{tabular}
   \end{ruledtabular}
-  \label{tab:design}
 \end{table}
@@
 For example, a previous analysis of the \trento\ model \cite{Bernhard:2016tnd} found $p \sim 0$, but we use a prior range $p \in [-1, 1]$ to account for differences in the present model, e.g.\ nucleon substructure, which could modify its posterior.
-One exception is the constituent number $n_c$ which we limit for practical considerations.
+One exception is the constituent number $\nc$ which we limit for practical considerations.
 Recall that each constituent fluctuates independently, weighted by a gamma random variable.
-Hence for constituent numbers $n_c \gg 1$, the fluctuations average out, and the resulting nucleon fluctuations vanish.
-To counteract this effect, the constituent fluctuation variance must increase as $n_c$ increases.
+Hence for constituent numbers $\nc \gg 1$, the fluctuations average out, and the resulting nucleon fluctuations vanish.
+To counteract this effect, the constituent fluctuation variance must increase as $\nc$ increases.
 Eventually, these required fluctuations become unreasonably large.
-We find that for $n_c < 10$, the energy density fluctuations are reasonable, and hence we limit $n_c$ to this prior range.
+We find that for $\nc < 10$, the energy density fluctuations are reasonable, and hence we limit $\nc$ to this prior range.
 
@@
 \begin{itemize}[leftmargin=1\parindent, itemsep=0pt]
-  \item Charged-particle multiplicity $d\nch/d\eta$.
-  \item Identified particle yields $dN/dy$ of pions, kaons, and protons.
-  \item Transverse energy production $dE_T/d\eta$.
-  \item Charged-particle mean transverse momentum $\langle p_T \rangle$ ($0.15 < p_T < 10$~GeV).
-  \item Identified particle mean transverse momentum $\langle p_T \rangle$ of pions, kaons, and protons.
-  \item Mean transverse momentum fluctuations $\delta p_T / \langle p_T \rangle$ (charged-particles, $0.15 < p_T < 2.0$~GeV).
-  \item Two-particle flow cumulants $\vnk{n}{2}$ for $n=2,3,4$\\ (charged-particles, $0.2 < p_T < 5.0$~GeV for ALICE, and $0.3 < p_T < 3.0$~GeV for CMS).
-  \item Four-particle flow cumulant $\vnk{2}{4}$ \\(charged-particles, $0.2 < p_T < 5.0$~GeV).
-  \item Symmetric cumulants $\SC(4, 2)$ and $\SC(3,2)$.
+  \item
+    Charged-particle multiplicity $d\nch/d\eta$.
+  \item
+    Identified particle yields $dN/dy$ of pions, kaons, and protons.
+  \item
+    Transverse energy production $dE_T/d\eta$.
+  \item
+    Charged-particle mean transverse momentum $\langle p_T \rangle$ ($0.15 < p_T < 10$~GeV).
+  \item
+    Identified particle mean transverse momentum $\langle p_T \rangle$ of pions, kaons, and protons.
+  \item
+    Mean transverse momentum fluctuations $\delta p_T / \langle p_T \rangle$ (charged-particles, $0.15 < p_T < 2.0$~GeV).
+  \item
+    Two-particle flow cumulants $\vnk{n}{2}$ for $n=2,3,4$\\ (charged-particles, $0.2 < p_T < 5.0$~GeV for ALICE, and $0.3 < p_T < 3.0$~GeV for CMS).
+  \item
+    Four-particle flow cumulant $\vnk{2}{4}$ \\(charged-particles, $0.2 < p_T < 5.0$~GeV).
+  \item
+    Symmetric cumulants $\SC(4, 2)$ and $\SC(3,2)$.
 \end{itemize}
+
 Each observable is calculated from the list of final state particles produced by UrQMD using the same methods applied by experiment.
@@
 Unfortunately, the four-particle cumulants require minimum-bias event statistics an order of magnitude larger than those used in this work.
-We therefore refrain from \emph{calibrating} on the four-particle cumulants, although we do show calculations of the four-particle cumulant $\vnk{2}{4}$ later in the text, using a single set of calibration parameters.
+We therefore refrain from \emph{calibrating} on the four-particle cumulants, although we do show calculations of the Pb-Pb four-particle cumulant $\vnk{2}{4}$ later in the text, using a single set of calibration parameters.
 
@@
 appearing in Eq.~\eqref{eq:likelihood}, which includes uncertainty contributions from both the model $\Sigmam$ and experimental data $\Sigmae$.
-The experimental contribution to the covariance $\Sigmae$ is further broken down into statistical and systematic components:
+The experimental contribution to the covariance $\Sigmae$ is further broken down into statistical and systematic components
 \begin{equation}
@@
 \end{equation}
-The statistical errors in $\Sigmae^\text{stat}$ are uncorrelated, and thus its covariance matrix is diagonal,
+The statistical errors in $\Sigmae^\text{stat}$ are uncorrelated, and thus its covariance matrix is diagonal
 \begin{equation}
@@
 \begin{equation}
-  \rho_{ij} = \frac{\cov(y_i, y_j)}{\sigma_i \sigma_j},
+  \rho_{ij} = \frac{\cov(y_i, y_j)}{\sigma_i \sigma_j}
 \end{equation}
@@
 We assume that each observable is correlated across different centrality/multiplicity bins, and uncorrelated with observables of a different type, e.g.\ correlations between yields and flows.
-This is a crude simplifying assumption but it is better than neglecting the correlation structure of the experimental data entirely.
+This is a crude simplifying assumption, but it is better than neglecting the correlation structure of the experimental data entirely.
 
@@
 \begin{equation}
-  \mathrm{corr}(y_i, y_j) = \cov(y_i, y_j)/(\sigma_i \sigma_j),
+  \mathrm{corr}(y_i, y_j) = \cov(y_i, y_j)/(\sigma_i \sigma_j)
 \end{equation}
@@
 The model is computationally intensive to evaluate, and thousands of events are required to calculate the simplest observables at a single parameter point.
-Moreover, we need to evaluate Eq.~\eqref{eq:likelihood} \emph{numerous} times in order to sample the multidimensional posterior distribution so that the samples may be histogrammed and visualized.
+Moreover, we need to evaluate Eq.~\eqref{eq:likelihood} numerous times in order to sample the multidimensional posterior distribution so that the samples may be histogrammed and visualized.
 
@@
 Similarly, let $Y$ denote the corresponding ${d \times m}$ observables matrix, where each row of $Y$ is a vector $\ym = (y_1, y_2, \dots, y_m)$ of $m$ model observables.
-The emulator operates strictly on model outputs, so let's drop the subscript on $\ym$ hereafter to declutter the notation.
+The emulator operates strictly on model outputs, so let's temporarily drop the subscript on $\ym$ to declutter the notation.
 Our goal is to train an emulator for the physics model $f$ using the discrete observations $f: X \mapsto Y$.
@@
 We first preprocess our model observables by centering and scaling each column of $Y$ (single observable) to zero mean and unit variance to generate a standardized observable matrix $\hat{Y}$.
-PCA is then used to reexpress each row-vector $\hat{\yv}_i$ of $\hat{Y}$ (all standardized observables at a single design point) in the new principal component basis:
+PCA is then used to reexpress each row-vector $\hat{\yv}_i$ of $\hat{Y}$ (all standardized observables at a single design point) in the new principal component basis
 \begin{equation}
@@
 \begin{equation}
-  \mathbf{z}(\xv) = (z_1(\xv), z_2(\xv), \dots, z_k(\xv))
+  \mathbf{z}(\xv) = [z_1(\xv), z_2(\xv), \dots, z_k(\xv)]
 \end{equation}
@@
 For additional information on the model emulator, we direct the reader to Appendix~\ref{app:validation} which includes several validation tests of the emulator prediction accuracy.
-\newpage
 
@@
 where $E$ subsumes all evidence from the $p$-Pb and \mbox{Pb-Pb} collision systems.
-We then perform Markov-chain Monte Carlo (MCMC) importance sampling on the posterior distribution $P(\xv | E)$ in Eq.~\eqref{eq:bayes} to draw random samples of the true model parameters $\xv_\star$, given the predictions of the model and the evidence provided by the experimental data \cite{Goodman:2010en, FM:2013mc}.
-For this we use an affine-invariant ensemble sampler which uses a large ensemble of interdependent walkers \cite{Goodman:2010en, FM:2013mc} and allow the MCMC chain to ``burn-in'' before generating \order{7} posterior samples.
+We then perform Markov-chain Monte Carlo (MCMC) importance sampling on the posterior distribution $P(\xv | E)$ in Eq.~\eqref{eq:bayes} to draw random samples of the ``true'' model parameters $\xv_\star$, given the evidence provided by the model predictions and the experimental data \cite{Goodman:2010en, FM:2013mc}.
+For this we use an affine-invariant sampler which uses a large ensemble of interdependent walkers \cite{Goodman:2010en, FM:2013mc} and allow the MCMC chain to burn-in before generating \order{7} posterior samples.
 
@@
       Norm       & \est{20.0}{2.5}{2.6} GeV      & $\smin$     & \est{0.08}{0.07}{0.07}            \\[1.1ex]
-      $p$        & \est{0.002}{0.180}{0.157}     & $\sslope$   & \est{1.24}{1.24}{1.46} GeV$^{-1}$ \\[1.1ex]
-      $\sigmaf$  & \est{0.91}{0.33}{0.32}        & $\scrv$     & \est{-0.09}{0.91}{0.80}           \\[1.1ex]
-      $r$        & \est{0.88}{0.23}{0.26} fm     & $\bmax$     & \est{0.026}{0.026}{0.032}         \\[1.1ex]
-      $n_c$      & \est{6.0}{3.4}{3.0}           & $\bwidth$   & \est{0.035}{0.035}{0.043} GeV     \\[1.1ex]
-      $v$        & \est{0.52}{0.20}{0.28} fm     & $\bloc$     & \est{0.174}{0.024}{0.020} GeV     \\[1.1ex]
+      $p$        & \est{0.002}{0.181}{0.156}     & $\sslope$   & \est{1.23}{1.23}{1.46} GeV$^{-1}$ \\[1.1ex]
+      $\sigmaf$  & \est{0.91}{0.33}{0.32}        & $\scrv$     & \est{-0.08}{0.92}{0.79}           \\[1.1ex]
+      $\rc$      & \est{0.88}{0.23}{0.26} fm     & $\bmax$     & \est{0.026}{0.026}{0.032}         \\[1.1ex]
+      $\nc$      & \est{6.0}{3.4}{3.0}           & $\bwidth$   & \est{0.036}{0.036}{0.043} GeV     \\[1.1ex]
+      $\wc$      & \est{0.52}{0.20}{0.28} fm     & $\bloc$     & \est{0.174}{0.024}{0.020} GeV     \\[1.1ex]
       $\dmin$    & \est{1.12}{0.49}{0.58} fm     & $\Tsw$      & \est{0.149}{0.014}{0.013} GeV     \\[1.1ex]
@@
 For example, the fictitious parameter estimate $q=2.45_{-0.15}^{+0.20}$ reports a median value $\tilde{q}=2.45$ and 90\% credible interval $2.30 < q < 2.65$.
-\newpage
 
 \begin{fullpage}
-  \begin{figure}
-    \makebox[\textwidth]{\includegraphics[width=.95\paperwidth]{posterior}}
-  \end{figure}
-  \captionsetup{skip=-15pt, width=\textwidth}
-  \captionof{figure}{\label{fig:posterior}
+  \centering
+  \vspace*{10pt}
+  \makebox[\textwidth]{
+    \includegraphics[width=.95\paperwidth]{posterior}
+  }
+  \captionof{figure}{
+    \label{fig:posterior}
     Bayesian posterior distribution of the model input parameters.
@@
     \label{fig:posterior_proton_shape}
-    Posterior distribution for the constituent sampling radius $r$ and constituent width $v$.
-    The prior range for $r$ and $v$ spans the full plot range.
-    The posterior distribution, shown in blue, indicates the preferred values for $r$ and $v$ determined by the analysis.
+    Posterior distribution (blue histogram) for the constituent position sampling radius $\rc$ and constituent width $\wc$.
+    The prior for $\rc$ and $\wc$ spans the full plot range.
   }
@@
 
-The \trento\ normalization factor ${\text{Norm} = 20.0^{+2.6}_{-2.5}}$ and energy deposition parameter $p=0.002^{+0.157}_{-0.180}$ are well constrained by the present analysis.
+The \trento\ normalization factor ${\text{Norm} = 20.0^{+2.6}_{-2.5}}$ and energy deposition parameter $p=0.002^{+0.156}_{-0.181}$ are well constrained by the present analysis.
 Moreover, Figs.~\ref{fig:obs_pbpb} and \ref{fig:obs_ppb} show that the model predictions using these values nicely describe both the $p$-Pb and Pb-Pb calibration observables.
@@
 We caution, however, that this specific analytic form should not be interpreted too literally.
-For instance, a generalized mean described by $p=0.05$ is well within our 90\% credible interval, but it does not equal the geometric mean in Eq.~\eqref{eq:geometric_mean}.
+For instance, a generalized mean described by $p=0.05$ is well within our 90\% credible interval, but it does not equal the geometric mean of Eq.~\eqref{eq:geometric_mean}.
 We also note that this scaling is somewhat different than the scaling obtained from previous analyses of the \trento\ model, which parametrized the \emph{entropy} density using a framework which assumed instant thermalization and zero pre-equilibrium flow.
@@
     \label{fig:posterior_parton_number}
-    Posterior distribution for the number of nucleon constituents $n_c$ determined by the analysis.
-    The parameter $n_c$ is an integer (discrete) variable at every design point, but the emulator interpolation produces a posterior distribution which is continuous.
+    Posterior distribution for the number of nucleon constituents $\nc$ determined by the analysis.
+    The parameter $\nc$ is an integer (discrete) variable at every design point, but the emulator interpolation produces a posterior distribution which is continuous.
   }
@@
 
-Continuing down the diagonal in Fig.~\ref{fig:posterior}, we see that the constituent sampling radius $r=0.88_{-0.23}^{+0.26}$~fm, and the constituent width $v=0.52_{-0.20}^{+0.28}$~fm.
+Continuing down the diagonal in Fig.~\ref{fig:posterior}, we see that the constituent position sampling radius $\rc=0.88_{-0.23}^{+0.26}$~fm, and the constituent width $\wc=0.52_{-0.20}^{+0.28}$~fm.
 Figure~\ref{fig:posterior_proton_shape} shows the joint posterior distribution of both parameters, illustrating the constraining power of the Bayesian analysis.
-While the sampling radius $r$ varies the size of the nucleons, we caution that its specific meaning should be interpreted with care; it specifies a computational \emph{sampling} radius, not a physical nucleon width.
-Consider, for instance, a single nucleon with $n_c = 2$ constituents.
-If the two constituents positions land on the same side of the nucleon, the effective nucleon size will be smaller than the Gaussian sampling radius $r$.
-Despite this idiosyncrasy, one can easily define a physical nucleon width in the nucleon center-of-mass frame \emph{ex post facto}, given specific values for the sampling radius $r$, constituent width $v$, and constituent number $n_c$.
+While the sampling radius $\rc$ varies the size of the nucleons, we caution that its specific meaning should be interpreted with care; it specifies a computational \emph{sampling} radius, not a physical nucleon width.
+Consider, for instance, a single nucleon with $\nc = 2$ constituents.
+If the two constituents positions land on the same side of the nucleon, the effective nucleon size will be smaller than the Gaussian sampling radius $\rc$.
+Despite this idiosyncrasy, one can easily define a physical nucleon width in the nucleon center-of-mass frame \emph{ex post facto}, given specific values for the sampling radius $\rc$, constituent width $\wc$, and constituent number $\nc$.
 
@@
 
-For example, using the posterior distribution's median values, $r=0.88$~fm, $n_c=6$, and $v=0.52$~fm, we can generate a large ensemble of random nucleon configurations and average their density in each nucleon's center-of-mass frame.
+For example, using the posterior distribution's median values, $\rc=0.88$~fm, $\nc=6$, and $\wc=0.52$~fm, we can generate a large ensemble of random nucleon configurations and average their density in each nucleon's center-of-mass frame.
 The resulting ensemble-averaged nucleon density
 \begin{equation}
-  \bar{\rho}_n(\xv) = \frac{1}{(2 \pi w^2)^{3/2}} \exp \left( -\frac{|\xv|^2}{2 w^2} \right)
+  \langle \rho_n(\xv)\rangle = \frac{1}{(2 \pi w^2)^{3/2}} \exp \left( -\frac{|\xv|^2}{2 w^2} \right)
 \end{equation}
@@
 
-This is perhaps the single largest difference between our work and the conclusions of recent saturation-based calculations which constrained the event-by-event fluctuations of the proton using a color-dipole picture of vector meson production \cite{Mantysaari:2016ykx, Mantysaari:2016jaz}.
+This is perhaps the single largest difference between our work and the conclusions of recent saturation-based calculations which constrained the event-by-event fluctuations of the proton using a color dipole picture of vector meson production \cite{Mantysaari:2016ykx, Mantysaari:2016jaz}.
 Those studies find that the measured coherent and incoherent $J/\Psi$ spectra at HERA prefer a compact gluon distribution inside each nucleon, with a Gaussian width $w_g \approx 0.4$~fm which is roughly \emph{half} the Gaussian width preferred by our analysis.
@@
 
-Moving on, we direct our attention to the constituent number $n_c$, shown enlarged in Fig.~\ref{fig:posterior_parton_number}.
+Moving on, we direct our attention to the constituent number $\nc$, shown enlarged in Fig.~\ref{fig:posterior_parton_number}.
 The distribution is not sharply peaked, and hence we refrain from quoting a distribution median and 90\% credible interval.
-Note, however, that the posterior clearly favors $n_c > 1$ constituents.
+Note, however, that the posterior clearly favors $\nc > 1$ constituents.
 This is not surprising.
-The \trento\ model mimics saturation-based initial condition models \cite{Bernhard:2016tnd}, and saturation models tend to produce ``proton-sized'' fireballs in $p$-Pb collisions \cite{Bzdak:2013zma}.
-When the proton is spherically symmetric, the resulting proton-sized QGP is also largely symmetric and thus produces very little flow.
-Saturation-based models are therefore unable to describe the significant flow measured in high-multiplicity $p$-Pb collisions without nucleon substructure, or alternatively, some other source of additional correlations \cite{Schenke:2017bog}.
+The \trento\ model mimics saturation-based initial condition models \cite{Bernhard:2016tnd}, and saturation models tend to produce proton-sized fireballs in $p$-Pb collisions \cite{Bzdak:2013zma}.
+When the proton is spherically symmetric, the resulting QGP is also largely symmetric and thus produces very little anisotropic flow.
+Saturation-based models are therefore unable to describe the significant anisotropic flow measured in central $p$-Pb collisions without nucleon substructure, or alternatively, some other source of additional correlations \cite{Schenke:2017bog}.
 
@@
 
-The posterior free streaming time $\taufs=0.47_{-0.37}^{+0.55}~\fmc$ obtained in this work is significantly smaller than the previous estimate $\taufs=1.16_{-0.25}^{+0.29}\ \fmc$ quoted in Ref.~\cite{Bernhard:2018hnz}.
+The posterior free streaming time $\taufs=0.47_{-0.37}^{+0.55}~\fmc$ obtained in this work is significantly smaller than the previous estimate $\taufs=1.16_{-0.25}^{+0.29}~\fmc$ quoted in Ref.~\cite{Bernhard:2018hnz}.
 We point out that the present study is missing several important observables which could affect the estimated free streaming time, e.g.\ the Pb-Pb mean $p_T$ and mean $p_T$ fluctuations at $\sqrts=5.02$~TeV.
@@
 Evidently, the combined analysis of Pb-Pb data at $\sqrts=2.76$ and 5.02~TeV in Ref.~\cite{Bernhard:2018hnz} provides a better constraint on the QGP viscosities which is not surprising given the additional observables and multiple beam energies studied.
-The $p$-Pb data, meanwhile, does not appear to provide any unique viscous constraints.
+The $p$-Pb data, meanwhile, does not appear to provide strong viscous constraints.
 
@@
       $k$      & 0.19           & $\scrv$      & -0.29          \\
-      $n_c$    & 6              & $\bmax$      & 0.032          \\
-      $r$      & 0.81 fm        & $\bwidth$    & 0.024 GeV      \\
-      $v$      & 0.43 fm        & $\bloc$      & 175 MeV        \\
-      $\dmin$  & 0.81 fm        & $\Tsw$       & 151 MeV        \\
+      $\nc$    & 6              & $\bmax$      & 0.032          \\
+      $\rc$    & 0.81 fm        & $\bwidth$    & 0.024 GeV      \\
+      $\wc$    & 0.43 fm        & $\bloc$      & 0.175 GeV      \\
+      $\dmin$  & 0.81 fm        & $\Tsw$       & 0.151 GeV      \\
       $\taufs$ & 0.37 \fmc
@@
 Here our MAP event sample is several orders of magnitude larger so the statistics are no issue.
-The bottom-right panel of Fig.~\ref{fig:obs_map} shows our model calculation for the four-particle elliptic flow cumulant $\vnk{2}{4}$ along with the measured data points from \mbox{ALICE} \cite{Adam:2016izf}.
+The bottom-right panel of Fig.~\ref{fig:obs_map} shows our model calculation for the Pb-Pb four-particle elliptic flow cumulant $\vnk{2}{4}$ along with the measured data points from \mbox{ALICE} \cite{Adam:2016izf}.
 We see that the MAP estimate nicely describes the measured $\vnk{2}{4}$ data which is encouraging since this particular observable was never used to calibrate the model.
@@
 \end{equation}
-
-\noindent The inner-average in Eq.~\eqref{eq:mean_pT_corr} runs over all pairs of particles $i,j$ in the same event, the outer average runs over all events in a given bin (centrality or multiplicity), and the symbol $\langle p_T \rangle$ denotes the usual mean transverse momentum of particles in the bin.
+The inner-average in Eq.~\eqref{eq:mean_pT_corr} runs over all pairs of particles $i,j$ in the same event, the outer average runs over all events in a given bin (centrality or multiplicity), and the symbol $\langle p_T \rangle$ denotes the usual mean transverse momentum of particles in the bin.
 The observable is typically presented in terms of the dimensionless ratio $\delta p_T / \langle p_T \rangle$, which quantifies the strength of the dynamical fluctuations in units of the average transverse momentum $\langle p_T \rangle$.
@@
 \end{equation}
+
 \begin{fullpage}
@@
 \end{fullpage}
-\noindent
-which are expected to be less sensitive to the medium response and more sensitive to the properties of the initial state.
+
+\noindent which are expected to be less sensitive to the medium response and more sensitive to the properties of the initial state.
 The solid lines are the MAP estimate of the present study, and the dashed lines are the MAP estimate of Ref.~\cite{Bernhard:2018hnz} which did not include nucleon substructure and was calibrated on Pb-Pb collisions at $\sqrts=2.76$ and 5.02~TeV.
@@
 Naturally, such descriptions rely on the validity of hydrodynamic approximations, and these approximations begin to break down in the so-called dilute limit where discrete particle degrees-of-freedom dominate and continuous field descriptions of the medium cease to make sense.
-Tell-tale signatures of hydrodynamic collectivity were thus always expected to vanish in smaller nuclear collision systems, e.g.\ $p$-$p$ and $p$-Pb collisions, where the number of produced particles is orders of magnitude smaller than a typical Pb-Pb collision.
+Tell-tale signatures of hydrodynamic collectivity were thus generally expected to vanish in smaller nuclear collision systems, e.g.\ $p$-$p$ and $p$-Pb collisions, where the number of produced particles is orders of magnitude smaller than a typical Pb-Pb collision.
 
@@
 First, we aimed to establish whether or not our hydrodynamic framework was able to describe both collision systems simultaneously.
-And second, in the event that the former was true, we wished to obtain estimates for the \emph{true} parameters of our model given the assumptions of our framework and the evidence provided by the experimental data.
+And second, in the event that the former was true, we wished to obtain estimates for the \emph{true} parameters of our model given the assumptions of our framework and the evidence provided by the model predictions and the experimental data.
 
@@
 For the QGP initial conditions, we employed a modified version of the \trento\ model \cite{Moreland:2014oya} which adds new parameters to vary the fluctuating size and shape of each nucleon.
-Specifically, we modeled each nucleon as a cluster of $n_c$ constituents (hot spots), where each constituent is described by a Gaussian density of width $v$.
-The constituent positions were each sampled randomly (without correlations) from a Gaussian radial distribution of width $r$ centered about each pre-defined nucleon position.
+Specifically, we modeled each nucleon as a cluster of $\nc$ constituents (hot spots), where each constituent is described by a Gaussian density of width $\wc$.
+The constituent positions were each sampled randomly (without correlations) from a Gaussian radial distribution of width $\rc$ centered about each pre-defined nucleon position.
 
@@
     We find that nucleon substructure is necessary to simultaneously describe $p$-Pb and Pb-Pb bulk observables.
-    However, we observe no strong preference for a specific number of constituents $n_c$ (hot spots) inside the nucleon.
-    In particular, we find no evidence to support the specific number $n_c=3$ which is commonly used in the literature.
+    However, we observe no strong preference for a specific number of constituents $\nc$ (hot spots) inside the nucleon.
+    In particular, we find no evidence to support the specific number $\nc=3$ which is commonly used in the literature.
   \item
@@
   \item
-    We obtain an estimate $v=0.52_{-0.20}^{+0.28}$~fm for the Gaussian width of the constituent hot spots inside each nucleon.
+    We obtain an estimate $\wc=0.52_{-0.20}^{+0.28}$~fm for the Gaussian width of the constituent hot spots inside each nucleon.
     This width is considerably larger than the length scales typically associated with nucleon substructure.
@@
     We report an estimate $\taufs=0.47_{-0.37}^{+0.55}~\fmc$ for the model's pre-equilibrium free streaming time.
-    This is significantly shorter than the estimate obtained from a similar Bayesian study in Ref.~\cite{Bernhard:2018hnz}, which found $\taufs=1.16_{-0.25}^{+0.29}\ \fmc$.
+    This is significantly shorter than the estimate obtained from a similar Bayesian study in Ref.~\cite{Bernhard:2018hnz}, which found $\taufs=1.16_{-0.25}^{+0.29}~\fmc$.
     It is not clear whether the difference is a result of nucleon substructure or the different observables used to calibrate each analysis.
@@
 
-\begin{figure}[t]
+\begin{figure}
   \includegraphics{validation_example_vert}
@@
 
-\begin{figure*}[b]
+\begin{figure*}[t]
   \includegraphics{validation_PbPb5020}\\

### 2019-05-07 src/design.py

@@ -130,11 +130,11 @@ class Design:
             ('dmin3',         r'd {min} [{fm}]',              (  0.0, 1.7**3)),
             ('tau_fs',        r'\tau {fs} [{fm}/c]',          (  0.1,    1.5)),
             ('etas_min',      r'\eta/s {min}',                (  0.0,    0.2)),
-            ('etas_slope',    r'a {shear} [{GeV}^{-1}]',      (  0.0,    8.0)),
-            ('etas_crv',      r'c {shear}',                   ( -1.0,    1.0)),
+            ('etas_slope',    r'\eta/s {slope} [{GeV}^{-1}]', (  0.0,    8.0)),
+            ('etas_crv',      r'\eta/s {crv}',                ( -1.0,    1.0)),
             ('zetas_max',     r'\zeta/s {max}',               (  0.0,    0.1)),
-            ('zetas_width',   r'\Delta \!T {bulk} [{GeV}]',   (  0.0,    0.1)),
-            ('zetas_t0',      r'T {bulk} [{GeV}]',            (0.150,  0.200)),
+            ('zetas_width',   r'\zeta/s {width} [{GeV}]',     (  0.0,    0.1)),
+            ('zetas_t0',      r'\zeta/s T_{peak} [{GeV}]',    (0.150,  0.200)),
             ('Tswitch',       r'T {switch} [{GeV}]',          (0.135,  0.165)),
         ]))
 
@@ -180,8 +180,8 @@ class Design:
         sampling_radius = np.sqrt(nucleon_width**2 - parton_width**2)
 
         design_changes = [
-            ('nucleon_width', 'sampling_radius', sampling_radius, r'$r\ [\mathrm{fm}]$', (0.0, 1.2)),
-            ('parton_struct', 'parton_width', parton_width, r'$v\ [\mathrm{fm}]$', (0.2, 1.2))
+            ('nucleon_width', 'sampling_radius', sampling_radius, r'$r_{cp}\ [\mathrm{fm}]$', (0.0, 1.2)),
+            ('parton_struct', 'parton_width', parton_width, r'$w_c\ [\mathrm{fm}]$', (0.2, 1.2))
         ]
 
         for key_old, key_new, points, label, (low, high) in design_changes:

### 2019-05-07 src/expt.py

@@ -336,7 +336,7 @@ def _data():
     # pPb5020 flows
     data['pPb5020']['vnk'] = {
         (n, k): pPb5020_flows(n, k)
-        for (n, k) in [(2, 2), (3, 2), (2, 4)]
+        for (n, k) in [(2, 2), (3, 2)]
     }
 
     # PbPb5020 dNch/deta

### 2019-05-07 src/model.py

@@ -359,12 +359,6 @@ def _data(system, dataset='main'):
     for obs in ['sc', 'sc_normed', 'sc_central', 'sc_normed_central']:
         data = dict({obs: {(3, 2): empty, (4, 2): empty}}, **data)
 
-    # missing flows
-    if system == 'pPb5020':
-        flow_dict = data['vnk'].copy()
-        flow_dict[(2, 4)] = empty
-        data['vnk'] = flow_dict
-
     data = ModelData(system, *files).observables_like(data)
 
     logging.info('writing cache file %s', cachefile)

### 2019-05-07 src/plots.py

@@ -2547,8 +2547,8 @@ def posterior_proton_shape():
 
     plt.hist2d(sampling_radius, parton_width, bins=50, cmap=plt.cm.Blues)
 
-    plt.xlabel('Constituent sampling radius $r$ [fm]')
-    plt.ylabel('Constituent width $v$ [fm]')
+    plt.xlabel('Constituent position radius $r_{cp}$ [fm]')
+    plt.ylabel('Constituent width $w_c$ [fm]')
 
     ax.set_aspect('equal')
     set_tight(fig, pad=.2)

### 2019-10-14 paper/paper.tex
Begin implementing referee edits.

@@
   Bayesian calibration of a hybrid nuclear collision model\\
-  using $\boldsymbol p$-Pb and Pb-Pb data from the LHC
+  using $\boldsymbol p$-Pb and Pb-Pb data at energies available\\
+  at the CERN Large Hadron Collider
 }
@@
 at the specified collision energy $\sqrts$.
+We tune $\sigma_\mathrm{eff}$ in the present work to fit the experimental inelastic nucleon-nucleon cross section $\sigma_\text{NN}^\text{inel}= 7.0$~fm$^2$ at $\sqrts=5.02$~TeV for comparison with our chosen datasets \cite{ALICE:2012xs}.
+The resulting \trento\ inelastic nucleon-nucleon cross section agrees with the experimental value to better than 2\% accuracy, as verified by one of the model's standard unit tests.
 
@@
 \emph{back} to physical space.
-Similarly, we can construct the covariance matrix of the observables in PCA space
-\begin{equation}
-  \cov(z_i, z_j) = \diag[(\delta z_{1})^2, (\delta z_2)^2, \dots, (\delta z_k)^2 ],
-\end{equation}
-and transform it back to physical space as well to obtain the covariance matrix $\Sigmam$ of the model observables at parameter point $\xv$.
-
-The resulting emulator therefore predicts both a mean prediction $\ym(\xv)$ and an uncertainty covariance matrix $\Sigmam(\xv)$ which accounts for multiple sources of model and emulator uncertainty, including the truncation error expected from using a finite number of principal components $k < m$.
-The model covariance matrix $\Sigmam$ includes so-called \emph{known}-unknowns such as statistical error and emulator interpolation error, but not \emph{unknown}-unknowns such as the overall validity of small-system hydrodynamics, i.e.\ things which lack a unified consensus or are difficult to quantify.
-We show in Fig.~\ref{fig:correlation} the resulting Pb-Pb correlation matrix $\mathrm{corr}(y_i,y_j)$ for the model (emulator) at a random parameter point $\xv$ in the design space (left-side), along side the same correlation matrix for the experimental data (right-side) discussed previously.
-For additional information on the model emulator, we direct the reader to Appendix~\ref{app:validation} which includes several validation tests of the emulator prediction accuracy.
 
@@
 \end{fullpage}
+
+Similarly, we can construct the covariance matrix of the observables in PCA space
+\begin{equation}
+  \cov(z_i, z_j) = \diag[(\delta z_{1})^2, (\delta z_2)^2, \dots, (\delta z_k)^2 ],
+\end{equation}
+and transform it back to physical space as well to obtain the covariance matrix $\Sigmam$ of the model observables at parameter point $\xv$.
+
+The resulting emulator therefore predicts both a mean prediction $\ym(\xv)$ and an uncertainty covariance matrix $\Sigmam(\xv)$ which accounts for multiple sources of model and emulator uncertainty, including the truncation error expected from using a finite number of principal components $k < m$.
+The model covariance matrix $\Sigmam$ includes so-called \emph{known}-unknowns such as statistical error and emulator interpolation error, but not \emph{unknown}-unknowns such as the overall validity of small-system hydrodynamics, i.e.\ things which lack a unified consensus or are difficult to quantify.
+We show in Fig.~\ref{fig:correlation} the resulting Pb-Pb correlation matrix $\mathrm{corr}(y_i,y_j)$ for the model (emulator) at a random parameter point $\xv$ in the design space (left-side), along side the same correlation matrix for the experimental data (right-side) discussed previously.
+For additional information on the model emulator, we direct the reader to Appendix~\ref{app:validation} which includes several validation tests of the emulator prediction accuracy.
 
@@
 We demonstrate later in the text that a single set of model parameters well describes all of the calibration data, and thus we believe that much of the spread in the posterior samples is uncertainty contributed by our emulator.
+
+\begin{fullpage}
+  \centering
+  \vspace*{10pt}
+  \makebox[\textwidth]{
+    \includegraphics[width=.95\paperwidth]{posterior}
+  }
+  \captionof{figure}{
+    \label{fig:posterior}
+    Bayesian posterior distribution of the model input parameters.
+    The diagonal panels show the marginalized distributions of individual model parameters, while off-diagonal panels show the joint distributions for pairs of model parameters, visualizing their correlations.
+    The marginalized distribution medians and 90\% credible intervals are annotated along the diagonal.
+  }
+\end{fullpage}
+
+\noindent
 We also note that although the $p$-Pb posterior samples have a somewhat larger spread than the Pb-Pb samples, the percentage uncertainty of the $p$-Pb emulator is similar to that of the Pb-Pb emulator, and thus the difference is likely due to the larger variance of the $p$-Pb training data.
@@
 
-\begin{fullpage}
-  \centering
-  \vspace*{10pt}
-  \makebox[\textwidth]{
-    \includegraphics[width=.95\paperwidth]{posterior}
-  }
-  \captionof{figure}{
-    \label{fig:posterior}
-    Bayesian posterior distribution of the model input parameters.
-    The diagonal panels show the marginalized distributions of individual model parameters, while off-diagonal panels show the joint distributions for pairs of model parameters, visualizing their correlations.
-    The marginalized distribution medians and 90\% credible intervals are annotated along the diagonal.
-  }
-\end{fullpage}
-
 \subsection{Initial condition properties}
-
-\begin{figure}
-  \includegraphics{posterior_proton_shape}
-  \caption{
-    \label{fig:posterior_proton_shape}
-    Posterior distribution (blue histogram) for the constituent position sampling radius $\rc$ and constituent width $\wc$.
-    The prior for $\rc$ and $\wc$ spans the full plot range.
-  }
-\end{figure}
 
@@
 \begin{figure}
-  \includegraphics{posterior_parton_number}
-  \caption{
-    \label{fig:posterior_parton_number}
-    Posterior distribution for the number of nucleon constituents $\nc$ determined by the analysis.
-    The parameter $\nc$ is an integer (discrete) variable at every design point, but the emulator interpolation produces a posterior distribution which is continuous.
+  \includegraphics{posterior_proton_shape}
+  \caption{
+    \label{fig:posterior_proton_shape}
+    Posterior distribution (blue histogram) for the constituent position sampling radius $\rc$ and constituent width $\wc$.
+    The prior for $\rc$ and $\wc$ spans the full plot range.
   }
@@
 
-\begin{figure*}
-  \includegraphics{region_shear_bulk}
-  \caption{
-    \label{fig:region_shear_bulk}
-    Left figure: estimated temperature dependence of the QGP specific shear viscosity $(\eta/s)(T)$ determined by the present Bayesian analysis of $p$-Pb and Pb-Pb collisions at $\sqrts=5.02$~TeV (orange line/band) compared to a previous Bayesian analysis of Pb-Pb collisions at $\sqrts=2.76$ and 5.02~TeV (blue line/band) \cite{Bernhard:2018hnz}.
-    The lines are the medians of each posterior distribution, and the bands are their 90\% credible regions.
-    Right figure: same as before, but for the temperature dependence of the QGP specific bulk viscosity $(\zeta/s)(T)$.
-  }
-\end{figure*}
-
 For example, using the posterior distribution's median values, $\rc=0.88$~fm, $\nc=6$, and $\wc=0.52$~fm, we can generate a large ensemble of random nucleon configurations and average their density in each nucleon's center-of-mass frame.
@@
 Evidently, it may be necessary to place an informative prior on our nucleon substructure parameters in order to resolve the apparent tension between our parameter values and those needed to describe DIS measurements at HERA.
-Alternatively, it is also possible that the fluctuations probed by coherent and incoherent $J/\Psi$ production are different than those probed by minimum-bias particle production.
+
+\begin{figure}[b]
+  \includegraphics{inelasticity_profile}
+  \caption{
+    \label{fig:inelasticity_profile}
+    Glauber model estimates of the normalized proton-proton inelasticity density $G_\mathrm{inel}(b) = d^2\sigma_\mathrm{inel}/d^2b$ at two different LHC beam energies.
+    Blue line: A multi-source Glauber model fit to LHC data at $\sqrts=7$~TeV \cite{Albacete:2016pmp}.
+    Orange line: \trento\ model predictions at $\sqrts=5.02$~TeV using median parameters from the present study's Bayesian posterior.
+  }
+\end{figure}
+
+Additional constraints on the proton size and shape are provided by the proton-proton inelasticity density $G_\mathrm{inel}(b) = d^2\sigma_\mathrm{inel}(b)/d^2b$, which measures the proton-proton inelastic collision probability as a function of its impact parameter.
+We compare, in Fig~\ref{fig:inelasticity_profile}, the normalized proton-proton inelasticity density $G_\mathrm{inel}(b) / G_\mathrm{inel}(0)$ at $\sqrts=5.02$~TeV predicted by \trento\ using the present study's posterior median parameters with a calculation from reference \cite{Albacete:2016pmp}, which extracted the proton-proton inelasticity density from LHC data at $\sqrts=7$~TeV using a multi-source Glauber model \cite{Albacete:2016pmp}.
+Our Bayesian median estimate (orange line) prefers a larger width for the proton-proton normalized inelasticity density compared to that of reference \cite{Albacete:2016pmp}, which is \emph{opposite} the trend one would expect based on the differences in beam energies alone.
+This suggests that our posterior estimate for the proton radius is somewhat oversized.
+Nevertheless, it is fascinating that the present Bayesian estimate is as close as it is, given that the model is fit to quantities which are not typically used to extract the proton-proton inelasticity density.
+
+\begin{figure}
+  \includegraphics{posterior_parton_number}
+  \caption{
+    \label{fig:posterior_parton_number}
+    Posterior distribution for the number of nucleon constituents $\nc$ determined by the analysis.
+    The parameter $\nc$ is an integer (discrete) variable at every design point, but the emulator interpolation produces a posterior distribution which is continuous.
+  }
+\end{figure}
 
@@
 
+\begin{figure*}
+  \includegraphics{region_shear_bulk}
+  \caption{
+    \label{fig:region_shear_bulk}
+    Left figure: estimated temperature dependence of the QGP specific shear viscosity $(\eta/s)(T)$ determined by the present Bayesian analysis of $p$-Pb and Pb-Pb collisions at $\sqrts=5.02$~TeV (orange line/band) compared to a previous Bayesian analysis of Pb-Pb collisions at $\sqrts=2.76$ and 5.02~TeV (blue line/band) \cite{Bernhard:2018hnz}.
+    The lines are the medians of each posterior distribution, and the bands are their 90\% credible regions.
+    Right figure: same as before, but for the temperature dependence of the QGP specific bulk viscosity $(\zeta/s)(T)$.
+  }
+\end{figure*}
+
 In this section, we compare several of our posterior estimates to those obtained from a similar Bayesian analysis in Ref.~\cite{Bernhard:2018hnz} which used an (almost) identical version of the present physics model.
@@
 \subsection{Verification of high-probability parameters}
+
+We verified the emulator and tested the accuracy of our physics model framework using a single set of high-probability parameters selected from the Bayesian posterior.
+These parameters, listed in Table~\ref{tab:mode_params}, are the approximate ``best fit'' values of the calibrated model, commonly referred to as the \emph{maximum a posteriori} (MAP) estimate
+\begin{equation}
+  \xv_\mathrm{MAP} \equiv \operatorname*{arg\, max}_{\xv} P(\xv | E).
+\end{equation}
+We then ran \order{6} minimum-bias and multiplicity triggered events using the MAP estimate $\xv_\mathrm{MAP}$ and computed all of the model observables listed in Sec.~\ref{sec:observables}.
+The resulting model calculations are shown in Fig.~\ref{fig:obs_map} alongside experimental data from CMS \cite{Chatrchyan:2013nka} and ALICE \cite{Adam:2015ptt, Adam:2016izf, Adam:2014qja, Abelev:2013bla}.
+The left and right columns show the results for the $p$-Pb and Pb-Pb collision systems respectively, and each row shows a different group of related observables.
+
+The global agreement of the MAP model calculations with the experimental data is generally quite good.
+The largest tension is observed in the two-particle cumulants $\vnk{2}{2}$ and $\vnk{3}{2}$ of the $p$-Pb system, although even that tension is only about 10--15\%.
+Quite remarkably, the model perfectly describes the shape of the $p$-Pb and Pb-Pb two-particle correlations which is strong evidence that these correlations are hydrodynamic in origin.
 
@@
 
-We verified the emulator and tested the accuracy of our physics model framework using a single set of high-probability parameters selected from the Bayesian posterior.
-These parameters, listed in Table~\ref{tab:mode_params}, are the approximate ``best fit'' values of the calibrated model, commonly referred to as the \emph{maximum a posteriori} (MAP) estimate
-\begin{equation}
-  \xv_\mathrm{MAP} \equiv \operatorname*{arg\, max}_{\xv} P(\xv | E).
-\end{equation}
-We then ran \order{6} minimum-bias and multiplicity triggered events using the MAP estimate $\xv_\mathrm{MAP}$ and computed all of the model observables listed in Sec.~\ref{sec:observables}.
-The resulting model calculations are shown in Fig.~\ref{fig:obs_map} alongside experimental data from CMS \cite{Chatrchyan:2013nka} and ALICE \cite{Adam:2015ptt, Adam:2016izf, Adam:2014qja, Abelev:2013bla}.
-The left and right columns show the results for the $p$-Pb and Pb-Pb collision systems respectively, and each row shows a different group of related observables.
-
-The global agreement of the MAP model calculations with the experimental data is very good.
-The largest tension is observed in the two-particle cumulants $\vnk{2}{2}$ and $\vnk{3}{2}$ of the $p$-Pb system, although even that tension is only about 10--15\%.
-Quite remarkably, the model perfectly describes the shape of the $p$-Pb and Pb-Pb two-particle correlations which is strong evidence that these correlations are hydrodynamic in origin.
 Moreover, we obtain an excellent description of the \mbox{$p$-Pb} mean $p_T$, although this fit is somewhat less meaningful since we are unable to calibrate on the Pb-Pb mean $p_T$ simultaneously (data is not yet available).
@@
 \end{equation}
+which are expected to be less sensitive to the medium response and more sensitive to the properties of the initial state.
+The solid lines are the MAP estimate of the present study, and the dashed lines are the MAP estimate of Ref.~\cite{Bernhard:2018hnz} which did not include nucleon substructure and was calibrated on Pb-Pb collisions at $\sqrts=2.76$ and 5.02~TeV.
+We observe that the gap between $\SC(3,2)$ and $\SC(4,2)$ is generally wider in the present analysis than in Ref.~\cite{Bernhard:2018hnz}, as is the gap between the normalized symmetric cumulants $\NSC(3,2)$ and $\NSC(4,2)$.
+
+We emphasize that multiple aspects of the two analyses are different such as the collision systems and beam energies considered, the observables which were included in each calibration, and the existence of nucleon substructure in the model.
+Thus we can only speculate what might have caused the large difference in the MAP estimate for the symmetric flow cumulants.
+Two reasonable culprits would be the inclusion of nucleon substructure and the large difference in the preferred pre-equilibrium free streaming time determined by the two studies.
 
@@
 \end{fullpage}
-
-\noindent which are expected to be less sensitive to the medium response and more sensitive to the properties of the initial state.
-The solid lines are the MAP estimate of the present study, and the dashed lines are the MAP estimate of Ref.~\cite{Bernhard:2018hnz} which did not include nucleon substructure and was calibrated on Pb-Pb collisions at $\sqrts=2.76$ and 5.02~TeV.
-We observe that the gap between $\SC(3,2)$ and $\SC(4,2)$ is generally wider in the present analysis than in Ref.~\cite{Bernhard:2018hnz}, as is the gap between the normalized symmetric cumulants $\NSC(3,2)$ and $\NSC(4,2)$.
-
-We emphasize that multiple aspects of the two analyses are different such as the collision systems and beam energies considered, the observables which were included in each calibration, and the existence of nucleon substructure in the model.
-Thus we can only speculate what might have caused the large difference in the MAP estimate for the symmetric flow cumulants.
-Two reasonable culprits would be the inclusion of nucleon substructure and the large difference in the preferred pre-equilibrium free streaming time determined by the two studies.
 

### 2019-10-14 src/plots.py

@@ -23,6 +23,7 @@ import logging
 import multiprocessing
 from pathlib import Path
 import pickle
+import subprocess
 import tempfile
 import warnings
 
@@ -2625,6 +2626,90 @@ def trim_design_points():
                         bad_points.add(point)
 
     print('{} bad points\n{}'.format(len(bad_points), bad_points))
+
+
+def run_cmd(*args):
+    """
+    Run and log a subprocess.
+
+    """
+    cmd = ' '.join(args)
+    logging.info('running command: %s', cmd)
+
+    try:
+        proc = subprocess.run(
+            cmd.split(), check=True,
+            stdout=subprocess.PIPE,
+            stderr=subprocess.STDOUT,
+            universal_newlines=True
+        )
+    except subprocess.CalledProcessError as e:
+        logging.error(
+            'command failed with status %d:\n%s',
+            e.returncode, e.output.strip('\n')
+        )
+        raise
+    else:
+        logging.debug(
+            'command completed successfully:\n%s',
+            proc.stdout.strip('\n')
+        )
+        return proc
+
+
+def trento_pcoll(impact_parameter):
+    """
+    Returns collision probability at the specified
+    impact parameter.
+
+    """
+    proc = run_cmd(
+            'trento p p',
+            '--number-events 100000',
+            '--reduced-thickness 0',
+            '--fluctuation 0.19',
+            '--nucleon-width 0.855',
+            '--constit-width 0.43',
+            '--constit-number 6',
+            '--cross-section 7.0',
+            '--b-min {}'.format(impact_parameter),
+            '--b-max {}'.format(impact_parameter),
+            '--ncoll'
+    )
+
+    ncoll = np.loadtxt(proc.stdout.splitlines(), usecols=2)
+    return np.count_nonzero(ncoll) / ncoll.size
+
+
+@plot
+def inelasticity_profile():
+    """
+    proton-proton inelasticity density as a function of collision impact
+    parameter.
+
+    """
+    fig = plt.figure(figsize=figsize(.5, aspect=.7))
+    ncpu = multiprocessing.cpu_count()
+
+    # inelasticity density at 7 TeV by Alba Ontoso et. al.
+    fname = workdir / 'expt/inelasticity_density_lhc.txt'
[…]

### 2019-10-27 paper/paper.tex
Finish referee revisions

@@
 
-\begin{table}[t]
+Conversely, the calibrated (posterior sampled) emulator predictions (bottom row of Figs.~\ref{fig:obs_pbpb} and \ref{fig:obs_ppb}) are far better constrained and nicely track the experimental data points.
+We emphasize here that the posterior parameter values are obtained from a \emph{simultaneous} calibration to $p$-Pb and Pb-Pb data, and thus they are self-consistent between the two systems.
+The spread in the posterior samples reflects different sources of model and experimental uncertainty as well as tension in the optimal fit parameters which describe each observable.
+We demonstrate later in the text that a single set of model parameters well describes all of the calibration data, and thus we believe that much of the spread in the posterior samples is uncertainty contributed by our emulator.
+
+We also note that although the $p$-Pb posterior samples have a somewhat larger spread than the Pb-Pb samples, the percentage uncertainty of the $p$-Pb emulator is similar to that of the Pb-Pb emulator, and thus the difference is likely due to the larger variance of the $p$-Pb training data.
+The uncertainty in the posterior distribution could thus be improved by running the calibration with more design points or with a narrower range of parameter values to increase the density of the training points and reduce interpolation uncertainty.
+
+We now direct our attention to Fig.~\ref{fig:posterior} which shows the main result of this work, the posterior distribution of the model input parameters.
+Recall that the posterior $P(\xv | E)$ is the probability that our hypothesis $\xv = \xv_\star$ is correct, given the evidence $E$ provided by the predictions of the model and the experimental data.
+The present posterior has 15 dimensions, one for each parameter listed in Table~\ref{tab:design}, and thus its joint distribution cannot be summarized by one figure alone.
+We therefore sample the distribution and histogram the samples to project the distribution onto one or two dimensions at a time.
+
+\begin{fullpage}
+  \centering
+  \vspace*{10pt}
+  \makebox[\textwidth]{
+    \includegraphics[width=.95\paperwidth]{posterior}
+  }
+  \captionof{figure}{
+    \label{fig:posterior}
+    Bayesian posterior distribution of the model input parameters.
+    The diagonal panels show the marginalized distributions of individual model parameters, while off-diagonal panels show the joint distributions for pairs of model parameters, visualizing their correlations.
+    The marginalized distribution medians and 90\% credible intervals are annotated along the diagonal.
+  }
+\end{fullpage}
+
+Each diagonal panel is the distribution of a \emph{single} model parameter (marginalized over all others), and each lower-diagonal panel is the joint distribution of a \emph{pair} of model parameters (marginalized over all others).
+We also report numeric estimates for each parameter's median value and 90\% credible interval and annotate their values along the distribution diagonal (see Table~\ref{tab:post_param}).
+For example, the fictitious parameter estimate $q=2.45_{-0.15}^{+0.20}$ reports a median value $\tilde{q}=2.45$ and 90\% credible interval $2.30 < q < 2.65$.
+
+\begin{table}
   \caption{
@@
 
-Conversely, the calibrated (posterior sampled) emulator predictions (bottom row of Figs.~\ref{fig:obs_pbpb} and \ref{fig:obs_ppb}) are far better constrained and nicely track the experimental data points.
-We emphasize here that the posterior parameter values are obtained from a \emph{simultaneous} calibration to $p$-Pb and Pb-Pb data, and thus they are self-consistent between the two systems.
-The spread in the posterior samples reflects different sources of model and experimental uncertainty as well as tension in the optimal fit parameters which describe each observable.
-We demonstrate later in the text that a single set of model parameters well describes all of the calibration data, and thus we believe that much of the spread in the posterior samples is uncertainty contributed by our emulator.
-
-\begin{fullpage}
-  \centering
-  \vspace*{10pt}
-  \makebox[\textwidth]{
-    \includegraphics[width=.95\paperwidth]{posterior}
-  }
-  \captionof{figure}{
-    \label{fig:posterior}
-    Bayesian posterior distribution of the model input parameters.
-    The diagonal panels show the marginalized distributions of individual model parameters, while off-diagonal panels show the joint distributions for pairs of model parameters, visualizing their correlations.
-    The marginalized distribution medians and 90\% credible intervals are annotated along the diagonal.
-  }
-\end{fullpage}
-
-\noindent
-We also note that although the $p$-Pb posterior samples have a somewhat larger spread than the Pb-Pb samples, the percentage uncertainty of the $p$-Pb emulator is similar to that of the Pb-Pb emulator, and thus the difference is likely due to the larger variance of the $p$-Pb training data.
-The uncertainty in the posterior distribution could thus be improved by running the calibration with more design points or with a narrower range of parameter values to increase the density of the training points and reduce interpolation uncertainty.
-
-We now direct our attention to Fig.~\ref{fig:posterior} which shows the main result of this work, the posterior distribution of the model input parameters.
-Recall that the posterior $P(\xv | E)$ is the probability that our hypothesis $\xv = \xv_\star$ is correct, given the evidence $E$ provided by the predictions of the model and the experimental data.
-The present posterior has 15 dimensions, one for each parameter listed in Table~\ref{tab:design}, and thus its joint distribution cannot be summarized by one figure alone.
-We therefore sample the distribution and histogram the samples to project the distribution onto one or two dimensions at a time.
-Each diagonal panel is the distribution of a \emph{single} model parameter (marginalized over all others), and each lower-diagonal panel is the joint distribution of a \emph{pair} of model parameters (marginalized over all others).
-We also report numeric estimates for each parameter's median value and 90\% credible interval and annotate their values along the distribution diagonal (see Table~\ref{tab:post_param}).
-For example, the fictitious parameter estimate $q=2.45_{-0.15}^{+0.20}$ reports a median value $\tilde{q}=2.45$ and 90\% credible interval $2.30 < q < 2.65$.
-
 \subsection{Initial condition properties}
@@
     \label{fig:inelasticity_profile}
-    Glauber model estimates of the normalized proton-proton inelasticity density $G_\mathrm{inel}(b) = d^2\sigma_\mathrm{inel}/d^2b$ at two different LHC beam energies.
-    Blue line: A multi-source Glauber model fit to LHC data at $\sqrts=7$~TeV \cite{Albacete:2016pmp}.
-    Orange line: \trento\ model predictions at $\sqrts=5.02$~TeV using median parameters from the present study's Bayesian posterior.
+    Estimates of the normalized proton-proton inelasticity density $G_\mathrm{inel}(b) = d^2\sigma_\mathrm{inel}/d^2b$ at two different LHC beam energies.
+    Blue line: Inelasticity density obtained by fitting a parametric form to LHC data at $\sqrts=7$~TeV \cite{Albacete:2016pmp}.
+    Orange line: \trento\ predictions at $\sqrts=5.02$~TeV using median parameters from the present study's Bayesian posterior.
   }
@@
 Additional constraints on the proton size and shape are provided by the proton-proton inelasticity density $G_\mathrm{inel}(b) = d^2\sigma_\mathrm{inel}(b)/d^2b$, which measures the proton-proton inelastic collision probability as a function of its impact parameter.
-We compare, in Fig~\ref{fig:inelasticity_profile}, the normalized proton-proton inelasticity density $G_\mathrm{inel}(b) / G_\mathrm{inel}(0)$ at $\sqrts=5.02$~TeV predicted by \trento\ using the present study's posterior median parameters with a calculation from reference \cite{Albacete:2016pmp}, which extracted the proton-proton inelasticity density from LHC data at $\sqrts=7$~TeV using a multi-source Glauber model \cite{Albacete:2016pmp}.
-Our Bayesian median estimate (orange line) prefers a larger width for the proton-proton normalized inelasticity density compared to that of reference \cite{Albacete:2016pmp}, which is \emph{opposite} the trend one would expect based on the differences in beam energies alone.
+We compare in Fig~\ref{fig:inelasticity_profile} the normalized proton-proton inelasticity density $G_\mathrm{inel}(b) / G_\mathrm{inel}(0)$ at $\sqrts=5.02$~TeV predicted by \trento\ using the present study's posterior median parameters with a calculation that extracted the proton-proton inelasticity density using a parametrization fit to proton-proton differential scattering data at $\sqrts=7$~TeV \cite{Albacete:2016pmp}.
+Our Bayesian median estimate (orange line) prefers a larger width for the proton-proton normalized inelasticity density compared to that of reference \cite{Albacete:2016pmp}, and this trend is \emph{opposite} what one would expect based on the difference in beam energy alone.
 This suggests that our posterior estimate for the proton radius is somewhat oversized.
@@
 
+\section{Summary and conclusions}
+\label{sec:summary}
+
+Relativistic heavy-ion collisions produce long-range multiparticle correlations which are commonly explained by the existence of hydrodynamic flow \cite{deSouza:2015ena}.
+This narrative is evidenced by the global, self-consistent, and highly non-trivial quantitative agreement of hydrodynamic models with a large number of heavy-ion bulk observables \cite{Niemi:2015qia, Bernhard:2016tnd, Gale:2012rq}.
+Naturally, such descriptions rely on the validity of hydrodynamic approximations, and these approximations begin to break down in the so-called dilute limit where discrete particle degrees-of-freedom dominate and continuous field descriptions of the medium cease to make sense.
+Tell-tale signatures of hydrodynamic collectivity were thus generally expected to vanish in smaller nuclear collision systems, e.g.\ $p$-$p$ and $p$-Pb collisions, where the number of produced particles is orders of magnitude smaller than a typical Pb-Pb collision.
+
 \begin{fullpage}
@@
 
-\section{Summary and conclusions}
-\label{sec:summary}
-
-Relativistic heavy-ion collisions produce long-range multiparticle correlations which are commonly explained by the existence of hydrodynamic flow \cite{deSouza:2015ena}.
-This narrative is evidenced by the global, self-consistent, and highly non-trivial quantitative agreement of hydrodynamic models with a large number of heavy-ion bulk observables \cite{Niemi:2015qia, Bernhard:2016tnd, Gale:2012rq}.
-Naturally, such descriptions rely on the validity of hydrodynamic approximations, and these approximations begin to break down in the so-called dilute limit where discrete particle degrees-of-freedom dominate and continuous field descriptions of the medium cease to make sense.
-Tell-tale signatures of hydrodynamic collectivity were thus generally expected to vanish in smaller nuclear collision systems, e.g.\ $p$-$p$ and $p$-Pb collisions, where the number of produced particles is orders of magnitude smaller than a typical Pb-Pb collision.
-
 These expectations were upended, however, when long-range multiparticle correlations were detected in high-multiplicity $p$-Pb collisions and found to be similar in magnitude to those observed in Pb-Pb collisions \cite{CMS:2012qk, Abelev:2012ola, Aad:2012gla}.
@@
     We note that our result is in significant tension with an estimate for the effective nucleon width based on the gluon distribution extracted from HERA data \cite{Rezaeian:2012ji}.
+    Our model calculations also predict a broader proton-proton inelasticity density at $\sqrts=5.02$~TeV than supported by the data \cite{Albacete:2016pmp}.
+    This suggests that our reconstructed protons are somewhat oversized.
   \item
@@
     This width is considerably larger than the length scales typically associated with nucleon substructure.
-    It may help explain the seemingly unreasonable success of hydrodynamics in small collision systems.
+    However, it is natural to wonder if this estimate is oversized, given the aforementioned comments regarding our extracted nucleon width.
   \item
@@
 Similarly, the results would greatly benefit from additional beam energies and collision systems.
-Notably absent is data from RHIC which includes measurements for $p$-Au, $d$-Au, $^3$He-Au, Cu-Cu, Cu-Au, Au-Au, and \mbox{U-U} collision systems at $\sqrts=200$~GeV.
-This data would undoubtedly help constrain the model parameters and would enable more stringent tests of the calibrated model predictions.
+For example, future studies should incorporate proton-proton collisions at top LHC energies, as well as the numerous collisions systems studied at top RHIC energy.
+Such data would undoubtedly improve constraints on the model parameters and would enable more stringent tests of the calibrated model predictions.
 The RHIC data may also help elucidate the beam-energy dependence of the model parameters which would be worth investigating.

### 2019-10-27 src/plots.py

@@ -2694,7 +2694,7 @@ def inelasticity_profile():
     # inelasticity density at 7 TeV by Alba Ontoso et. al.
     fname = workdir / 'expt/inelasticity_density_lhc.txt'
     b_values, Pcoll = np.loadtxt(fname).T
-    plt.plot(b_values, Pcoll, label=r'$7$ TeV, Glauber fit')
+    plt.plot(b_values, Pcoll, label=r'$7$ TeV, Fit to $d(\sigma/dt)(t)$')
 
     # trento inelasticity density at 5.02 TeV
     b_values = np.linspace(0, 5, 20)
